\section{Lauds of the Dead}\label{laudsdead}
\fancyhead[RO,LE]{\textit{Lauds}}
\begin{secrubric}
    %The Monastic Diurnal is ambiguous on this rubric, but the Diurnale is clear.
    Before Lauds---if it be said separately from Matins---the Lord's Prayer and Angelic Salutation are said secretly. Otherwise, the Office begins at once with the Antiphon.
\end{secrubric}
\subsec{Psalm 51. \textit{Miserere mei, Deus}}\par\noindent
\textit{Ant.} The bones which thou hast broken {\dag} shall rejoice in the Lord.\par
\lett{H}{ave} mercy upon me, O God, after thy great goodness : according to the multitude of thy mercies do away mine offences.\par
\secondlinepsalm{Wash me throughly from my wickedness : and cleanse me from my sin.}
\thirdlinepsalm{For I acknowledge my faults : and my sin is ever before me.}
\psanum{4}Against thee only have I sinned, and done this evil in thy sight : that thou mightest be justified in thy saying, and clear when thou art judged.\par
\psanum{5}Behold, I was shapen in wickedness : and in sin hath my mother conceived me.\par
\psanum{6}But lo, thou requirest truth in the inward parts: and shalt make me to understand wisdom secretly.\par
\psanum{7}Thou shalt purge me with hyssop, and I shall be clean : thou shalt wash me, and I shall be whiter than snow.\par
\psanum{8}Thou shalt make me hear of joy and gladness : that the bones which thou hast broken may rejoice.\par
\psanum{9}Turn thy face from my sins : and put out all my misdeeds.\par
\psanum{10}Make me a clean heart, O God : and renew a right spirit within me.\par
\psanum{11}Cast me not away from thy presence : and take not thy holy Spirit from me.\par
\psanum{12}O give me the comfort of thy help again : and stablish me with thy free Spirit.\par
\psanum{13}Then shall I teach thy ways unto the wicked : and sinners shall be converted unto thee.\par
\psanum{14}Deliver me from blood-guiltiness, O God, thou that art the God of my health : and my tongue shall sing of thy righteousness.\par
\psanum{15}Thou shalt open my lips, O Lord : and my mouth shall shew thy praise.\par
\psanum{16}For thou desirest no sacrifice, else would I give it thee : but thou delightest not in burnt-offerings.\par
\psanum{17}The sacrifice of God is a troubled spirit : a broken and contrite heart, O God, shalt thou not despise.\par
\psanum{18}O be favourable and gracious unto Sion : build thou the walls of Jerusalem.\par
\psanum{19}Then shalt thou be pleased with the sacrifice of righteousness, with the burnt-offerings and oblations : then shall they offer young bullocks upon thine altar.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\par\noindent
\textit{Ant.} The bones which thou hast broken shall rejoice in the Lord.\par
\subsec{Psalm 65. \textit{Te decet hymnus}}\par\noindent
\textit{Ant.} Thou, O Lord, {\dag} that hearest the prayer, unto thee shall all flesh come.\par
\lett{T}{hou}, O God, art praised in Sion : and unto thee shall the vow be performed in Jerusalem.\par
\secondlinepsalm{Thou that hearest the prayer : unto thee shall all flesh come.}
\thirdlinepsalm{My misdeeds prevail against me : O be thou merciful unto our sins.}
\psanum{4}Blessed is the man whom thou choosest, and receivest unto thee : he shall dwell in thy court, and shall be satisfied with the pleasures of thy house, even of thy holy temple.\par
\psanum{5}Thou shalt shew us wonderful things in thy righteousness, O God of our salvation : thou that art the hope of all the ends of the earth, and of them that remain in the broad sea.\par
\psanum{6}Who in his strength setteth fast the mountains : and is girded about with power.\par
\psanum{7}Who stilleth the raging of the sea : and the noise of his waves, and the madness of the people.\par
\psanum{8}They also that dwell in the uttermost parts of the earth shall be afraid at thy tokens : thou that makest the outgoings of the morning and evening to praise thee.\par
\psanum{9}Thou visitest the earth, and blessest it : thou makest it very plenteous.\par
\psanum{10}The river of God is full of water : thou preparest their corn, for so thou providest for the earth.\par
\psanum{11}Thou waterest her furrows, thou sendest rain into the little valleys thereof : thou makest it soft with the drops of rain, and blessest the increase of it.\par
\psanum{12}Thou crownest the year with thy goodness : and thy clouds drop fatness.\par
\psanum{13}They shall drop upon the dwellings of the wilderness : and the little hills shall rejoice on every side.\par
\psanum{14}The folds shall be full of sheep : the valleys also shall stand so thick with corn, that they shall laugh and sing.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\par\noindent
\textit{Ant.} Thou, O Lord, that hearest the prayer, unto thee shall all flesh come.\par
\subsec{Psalm 63. \textit{Deus, Deus meus}}\par\noindent
\textit{Ant.} Thy right hand {\dag} hath upholden me, O Lord.\par
\lett{O}{God,} thou art my God : early will I seek thee.\par
\secondlinepsalm{My soul thirsteth for thee, my flesh also longeth after thee : in a barren and dry land where no water is.}
\thirdlinepsalm{Thus have I looked for thee in holiness : that I might behold thy power and glory.}
\psanum{4}For thy loving-kindness is better than the life itself : my lips shall praise thee.\par
\psanum{5}As long as I live will I magnify thee on this manner : and lift up my hands in thy Name.\par
\psanum{6}My soul shall be satisfied, even as it were with marrow and fatness : when my mouth praiseth thee with joyful lips.\par
\psanum{7}Have I not remembered thee in my bed : and thought upon thee when I was waking?\par
\psanum{8}Because thou hast been my helper : therefore under the shadow of thy wings will I rejoice.\par
\psanum{9}My soul hangeth upon thee : thy right hand hath upholden me.\par
\psanum{10}These also that seek the hurt of my soul : they shall go under the earth.\par
\psanum{11}Let them fall upon the edge of the sword : that they may be a portion for foxes.\par
\psanum{12}But the King shall rejoice in God; all they also that swear by him shall be commended : for the mouth of them that speak lies shall be stopped.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\par\noindent
\textit{Ant.} Thy right hand hath upholden me, O Lord.\par
\subsec{Ego dixi}\par\noindent
\textit{Ant.} From the gate of hell, {\dag} deliver my soul, O Lord.\par
\lett{I}{said} in the cutting off of my days, {\dag} I shall go to the gates of the grave: * I am deprived of the residue of my years.\par
\secondline{I said, I shall not see the Lord, {\dag} even the Lord, in the land of the living: * I shall behold man no more with the inhabitants of the world.}
\thirdline{Mine age is departed, * and is removed from me as a shepherd's tent;}
I have cut off like a weaver my life: {\dag} he will cut me off with pining sickness: * from day even to night wilt thou make an end of me.\par
I reckoned till morning, {\dag} that, as a lion, so will he break all my bones: * from day even to night wilt thou make an end of me.\par
Like a crane or a swallow, so did I chatter: * I did mourn as a dove.\par
Mine eyes fail with looking upward: * O Lord, I am oppressed; undertake for me.\par
What shall I say? {\dag} he hath both spoken unto me, and himself hath done it: * I shall go softly all my years in the bitterness of my soul.\par
O Lord, by these things men live, {\dag} and in all these things is the life of my spirit: * so wilt thou recover me, and make me to live.\par
Behold, for peace I had great bitterness: {\dag} but thou hast in love to my soul delivered it from the pit of corruption: * for thou hast cast all my sins behind thy back.\par
For the grave cannot praise thee, {\dag} death can not celebrate thee: * they that go down into the pit cannot hope for thy truth.\par
The living, the living, he shall praise thee, as I do this day: * the father to the children shall make known thy truth.\par
The Lord was ready to save me: * therefore we will sing my songs to the stringed instruments all the days of our life in the house of the Lord.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\par\noindent
\textit{Ant.} From the gate of hell, deliver my soul, O Lord.\par
%MANUAL ADJUSTMENT:
\vspace{-1.75ex}
\subsec{Psalm 150. \textit{Laudate Dominum}}\par\noindent
\textit{Ant.} Let everything that hath breath {\dag} praise the Lord.\par
\lett{O}{praise} God in his holiness : praise him in the firmament of his power.\par
\secondlinepsalm{Praise him in his noble acts : praise him according to his excellent greatness.}
\thirdlinepsalm{Praise him in the sound of the trumpet : praise him upon the lute and harp.}
\psanum{4}Praise him in the cymbals and dances : praise him upon the strings and pipe.\par
\psanum{5}Praise him upon the well-tuned cymbals : praise him upon the loud cymbals.\par
\psanum{6}Let every thing that hath breath : praise the Lord.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\par\noindent
\textit{Ant.} Let everything that hath breath praise the Lord.\par
℣. I heard a voice from heaven, saying unto me.\par
℟. Blessed are the dead which die in the Lord.\par
%\elcol{℣. I heard a voice from heaven, saying unto me.}{℣. Audívi vocem de c{\ae}lo dicéntem mihi.}
%\elcol{℟. Blessed are the dead which die in the Lord.}{℟. Beáti mórtui qui in Dómino moriúntur.}
%MANUAL ADJUSTMENT:
\vspace{-1ex}
\subby{Benedictus}\par\noindent
\textit{Ant.} I am the resurrection {\dag} and the life: he that believeth in me, though he were dead, yet shall he live: and whosoever liveth and believeth in me, shall never die.
\fakecol{\lett{B}{lessed} {\ding{64}} be the Lord God of Israel; * for he hath visited and redeemed his people;\par
    \secondline{And hath raised up a mighty salvation for us, * in the house of his servant David;}
    \thirdline{As he spake by the mouth of his holy Prophets, * which have been since the world began;}
    That we should be saved from our enemies, * and from the hand of all that hate us.\par
    To perform the mercy promised to our forefathers, * and to remember his holy covenant;\par
    To perform the oath which he sware to our forefather Abraham, * that he would give us;\par
    That we being delivered out of the hand of our enemies * might serve him without fear;\par
    In holiness and righteousness before him, * all the days of our life.\par
    And thou, child, shalt be called the prophet of the Highest: * for thou shalt go before the face of the Lord to prepare his ways;\par
    To give knowledge of salvation unto his people * for the remission of their sins,\par
    Through the tender mercy of our God; * whereby the day-spring from on high hath visited us;\par
    To give light to them that sit in darkness, and in the shadow of death, * and to guide our feet into the way of peace.\par
    ℣. Rest eternal * grant unto them, O Lord.\par
    ℟. And let light perpetual * shine upon them.}
{\lett{B}{ened\smash{í}ctus} {\ding{64}} Dóminus, Deus Isra{\"e}l: * quia visitávit, et fecit redemptiónem plebi su{\ae}:\par
\secondline{Et eréxit cornu salútis nobis: * in domo David, púeri sui.}
\thirdline{Sicut locútus est per os sanctórum, * qui a sǽculo sunt, prophetárum ejus:}
Salútem ex inimícis nostris, * et de manu ómnium, qui odérunt nos.\par
Ad faciéndam misericórdiam cum pátribus nostris: * et memorári testaménti sui sancti.\par
Jusjurándum, quod jurávit ad Ábraham patrem nostrum, * datúrum se nobis:\par
Ut sine timóre, de manu inimicórum nostrórum liberáti, * serviámus illi.\par
In sanctitáte, et justítia coram ipso, * ómnibus diébus nostris.\par
Et tu, puer, Prophéta Altíssimi vocáberis: * pr{\ae}íbis enim ante fáciem Dómini, paráre vias ejus:\par
Ad dandam sciéntiam salútis plebi ejus: * in remissiónem peccatórum eórum:\par
Per víscera misericórdi{\ae} Dei nostri: * in quibus visitávit nos, óriens ex alto:\par
Illumináre his, qui in ténebris, et in umbra mortis sedent: * ad dirigéndos pedes nostros in viam pacis.\par
℣. Réquiem {\ae}térnam dona eis, Dómine.\par
℟. Et lux perpétua lúceat eis.}\par\noindent
\textit{Ant.} I am the resurrection and the life: he that believeth in me, though he were dead, yet shall he live; and whosoever liveth and believeth in me, shall never die.
%MANUAL ADJUSTMENT:
\vspace{-3ex}
\subsec{Lord's Prayer}
\begin{rubric}
The Lord's Prayer is said, in secret, kneeling, ending with,
\end{rubric}
\fakecol{℣. And lead us not into temptation.}{℣. Et ne nos indúcas in tentatiónem:}
\fakecol{℟. But deliver us from evil.}{℟. Sed líbera nos a malo.}
\clearpage
\subby{Psalm 130. \textit{De profundis}}
\begin{rubric}
{Psalm 130 is not said on All Souls Day, on the day of death or burial, nor at any time when the Office is recited with Double rite.}
\end{rubric}
\lett{O}{ut} of the deep have I called unto thee, O Lord : Lord, hear my voice.\par
\secondlinepsalm{O let thine ears consider well : the voice of my complaint.}
\thirdlinepsalm{If thou, Lord, wilt be extreme to mark what is done amiss : O Lord, who may abide it?}
\psanum{4}For there is mercy with thee : therefore shalt thou be feared.\par
\psanum{5}I look for the Lord; my soul doth wait for him : in his word is my trust.\par
\psanum{6}My soul fleeth unto the Lord : before the morning watch, I say, before the morning watch.\par
\psanum{7}O Israel, trust in the Lord, for with the Lord there is mercy : and with him is plenteous redemption.\par
\psanum{8}And he shall redeem Israel : from all his sins.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.

\subsec{Responsory}
\fakecol{℣. From the gate of hell.}{℣. A porta ínferi.}
\fakecol{℟. Deliver \textit{his soul}, O Lord.}{℟. Erue, Dómine, \textit{ánimam ejus}.}
\fakecol{℣. May \textit{he} rest in peace.}{℣. Requiésca\textit{n}t in pace.}
\fakecol{℟. Amen.}{℟. Amen.}

\subsec{Conclusion}
\fakecol{℣. O Lord, hear my prayer.}{℣. Dómine, exáudi oratiónem meam.}
\fakecol{℟. And let my cry come unto thee.}{℟. Et clamor meus ad te véniat.}
\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\begin{rubric}
{The appropriate collect is here said (p. \pageref{dead}).}
\end{rubric}
\fakecol{℣. Rest eternal * grant unto them, O Lord.}{℣. Réquiem {\ae}térnam dona eis, Dómine.}
\fakecol{℟. And let light perpetual * shine upon them.}{℟. Et lux perpétua lúceat eis.}
\fakecol{℣. May they rest in peace.}{℣. Requiéscant in pace.}
\fakecol{℟. Amen.}{℟. Amen.}\par
\begin{center}
    \textsc{Here endeth the Order of Lauds of the Dead.}
\end{center}