%It is very difficult to find public domain translations of the Rite of Baptism of Adults. We have tried to remain consistent, but in some points this has been impossible. The following resources were followed in order of priority: BCP \& KJV, the Rituale Baptism of Infants, the Orthodox Ritual, the Old Catholic Baptism of Infants, old Roman devotionals, mine own translation of the Latin.
\section{Sacrament of Baptism for Adults}
\fancyhead[RO,LE]{\textit{Baptism of Adults}}
%Rubric from Orthodox Ritual:
\begin{secrubric}
    The clean Baptismal font is filled with clear water. The container for salt and the phials of Holy Chrism and Oil of Catechumens are placed nearby. The Baptismal Candle and Chrisom (white cloth) are prepared. The Priest (vested in surplice and purple stole and cope), with acolytes and assisting clerics, kneels at the steps of the Altar and silently offers devout prayers to God that he may worthily administer this great Sacrament. Rising, he makes the sign of the Cross and says or sings:
\end{secrubric}

\gregorioscore{resources/gabc/OGod.gabc}

\gregorioscore{resources/gabc/OLord.gabc}

\gregorioscore{resources/gabc/GloryBeBaptism.gabc}

\gregorioscore{resources/gabc/AsItWasBaptism.gabc}
\clearpage
%Antiphon is from Ezekiel 36:25.
%The word effundum means pour (though sprinkle here is a fine translation), but how ironic the KJV brings us to sprinkle in a rite where sprinkling is prohibited!
%\elcol{\antiphon{}{I will sprinkle clean water upon you, and ye shall be clean: from all your filthiness, saith the Lord.}}{\antiphon{}{Effúndam super vos aquam mundam, et mundabímini ab ómnibus inquinaméntis vestris, dicit Dóminus.}}
%Unsure where the Orthodox Ritual's translation comes from:
\subby{Psalm 8. \textit{Domine, Dominus noster}}\par\noindent
\antiphon{}{I will pour clean water over you, and ye shall be washed of all your iniquities, saith the Lord.}\par
\lett{O}{Lord} our Governor, how excellent is thy Name in all the world : thou that hast set thy glory above the heavens!\par
\secondlinepsalm{Out of the mouth of very babes and sucklings hast thou ordained strength, because of thine enemies : that thou mightest still the enemy and the avenger.}
\thirdlinepsalm{For I will consider thy heavens, even the works of thy fingers : the moon and the stars, which thou hast ordained.}
\psanum{4}What is man, that thou art mindful of him : and the son of man, that thou visitest him?\par
\psanum{5}Thou madest him lower than the angels : to crown him with glory and worship.\par
\psanum{6}Thou makest him to have dominion of the works of thy hands : and thou hast put all things in subjection under his feet;\par
\psanum{7}All sheep and oxen : yea, and the beasts of the field;\par
\psanum{8}The fowls of the air, and the fishes of the sea : and whatsoever walketh through the paths of the seas.\par
\psanum{9}O Lord our Governor : how excellent is thy Name in all the world!\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.\par
\subby{Psalm 29. \textit{Afferte Domino}}
\lett{B}{ring} unto the Lord, O ye mighty, bring young rams unto the Lord : ascribe unto the Lord worship and strength.\par
\secondlinepsalm{Give the Lord the honour due unto his Name : worship the Lord with holy worship.}
\thirdlinepsalm{It is the Lord that commandeth the waters : it is the glorious God that maketh the thunder.}
\psanum{4}It is the Lord that ruleth the sea; the voice of the Lord is mighty in operation : the voice of the Lord is a glorious voice.\par
\psanum{5}The voice of the Lord breaketh the cedar-trees : yea, the Lord breaketh the cedars of Libanus.\par
\psanum{6}He maketh them also to skip like a calf : Libanus also, and Sirion, like a young unicorn.\par
\psanum{7}The voice of the Lord divideth the flames of fire; the voice of the Lord shaketh the wilderness : yea, the Lord shaketh the wilderness of Cades.\par
\psanum{8}The voice of the Lord maketh the hinds to bring forth young, and discovereth the thick bushes : in his temple doth every man speak of his honour.\par
\psanum{9}The Lord sitteth above the water-flood : and the Lord remaineth a King for ever.\par
\psanum{10}The Lord shall give strength unto his people : the Lord shall give his people the blessing of peace.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.\par
\subby{Psalm 42. \textit{Quemadmodum}}
\lett{L}{ike} as the hart desireth the water-brooks : so longeth my soul after thee, O God.\par
\secondlinepsalm{My soul is athirst for God, yea, even for the living God : when shall I come to appear before the presence of God?}
\thirdlinepsalm{My tears have been my meat day and night : while they daily say unto me, Where is now thy God?}
\psanum{4}Now when I think thereupon, I pour out my heart by myself : for I went with the multitude, and brought them forth into the house of God;\par
\psanum{5}In the voice of praise and thanksgiving : among such as keep holy-day.\par
\psanum{6}Why art thou so full of heaviness, O my soul : and why art thou so disquieted within me?\par
\psanum{7}Put thy trust in God : for I will yet give him thanks for the help of his countenance.\par
\psanum{8}My God, my soul is vexed within me : therefore will I remember thee concerning the land of Jordan, and the little hill of Hermon.\par
\psanum{9}One deep calleth another, because of the noise of the water-pipes : all thy waves and storms are gone over me.\par
\psanum{10}The Lord hath granted his loving-kindness in the day-time : and in the night-season did I sing of him, and made my prayer unto the God of my life.\par
\psanum{11}I will say unto the God of my strength, Why hast thou forgotten me : why go I thus heavily, while the enemy oppresseth me?\par
\psanum{12}My bones are smitten asunder as with a sword : while mine enemies that trouble me cast me in the teeth;\par
\psanum{13}Namely, while they say daily unto me : Where is now thy God?\par
\psanum{14}Why art thou so vexed, O my soul : and why art thou so disquieted within me?\par
\psanum{15}O put thy trust in God : for I will yet thank him, which is the help of my countenance, and my God.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.\par
%\elcol{\antiphon{}{I will sprinkle clean water upon you, and ye shall be clean: from all your filthiness, saith the Lord.}}{\antiphon{}{Effúndam super vos aquam mundam, et mundabímini ab ómnibus inquinaméntis vestris, dicit Dóminus.}}
%\antiphon{}{I will sprinkle clean water upon you, and ye shall be clean: from all your filthiness, saith the Lord.}
\antiphon{}{I will pour clean water over you, and ye shall be washed of all your iniquities, saith the Lord.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us.\par
℣. And lead us not into temptation;\par
℟. But deliver us from evil.\par
℣. O Lord, hear my prayer.\par
℟. And let my cry come unto thee.\par
℣. The Lord be with you.\par
℟. And with thy spirit.\par
\letuspray
%BCP Collect of Trinity Sunday:
\lett{A}{lmighty} and everlasting God, who hast given unto us thy servants grace, by the confession of a true faith, to acknowledge the glory of the eternal Trinity, and in the power of the Divine Majesty to worship the Unity; We beseech thee that thou wouldest keep us stedfast in this faith, and evermore defend us from all adversities.
%Orthodox Ritual:
\needspace{4\baselineskip}
\lett{A}{lmighty} God, we beseech thee to hear our prayers: may the working of thy power accomplish what is to be done through our lowly ministry.
%Orthodox Ritual:
\needspace{4\baselineskip}
\lett{G}{rant,} O Lord, we beseech thee, to thy \textit{servant}, that, being instructed in thy holy mysteries, \textit{he} may be born again in the water of baptism and counted among the members of thy holy Church. Through Christ our Lord.\par
℟. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris:\par
℣. Et ne nos indúcas in tentatiónem:\par
℟. Sed líbera nos a malo.\par
℣. Dómine, exáudi oratiónem meam.\par
℟. Et clamor meus ad te véniat.\par
℣. Dóminus vobíscum.\par
℟. Et cum spíritu tuo.\par
\oremuslatin
\lett{O}{mn\smash{í}potens} sempitérne Deus, qui dedísti fámulis tuis in confessióne ver{\ae} Fídei, {\ae}térn{\ae} Trinitátis glóriam agnóscere, et in poténtia majestátis adoráre unitatem, qu{\ae}sumus, ut ejusdem Fidei firmitáte ab ómnibus semper muniámur advérsis.
\needspace{4\baselineskip}
\lett{A}{d\smash{é}sto} supplicatiónibus nostris, Omnípotens Deus; et quod humilitátis nostr{\ae} geréndum est ministério, tu{\ae} virtútis impleátur efféctu.
\needspace{4\baselineskip}
%Orthodox Ritual:
\lett{D}{a,} qu{\ae}sumus, Dómine,* \textit{Elécto nostro}, ut sanctis * \textit{edóctus} mystériis, et renovétur Fonte Baptísmatis, et inter Ecclési{\ae} tu{\ae} membra numerétur. Per Christum Dóminum nostrum.\par
℟. Amen.}

\begin{rubric}
	The Priest then interrogates the Catechumen.
\end{rubric}
%The Orthodox Ritual goes back and forth between sacral and secular language. Orthodox Ritual adopted, except where noted, adjusted for sacral language.
\elcol{℣. \textit{N.}, what dost thou ask of the Church of God?}{℣. \textit{N.}, Quid petis ab Ecclesia Dei?}

\elcol{℟. Faith.}{℟. Fidem.}

%Orthodox Ritual translation is errant:
\elcol{℣. What doth Faith bring to thee?}{℣. Fides quid tibi pr{\ae}stat?}

\elcol{℟. Life everlasting.}{℟. Vitam {\ae}térnam.}

%Changed to conform to BCP translation.
\elcol{℣. If then thou desirest to enter into life, keep the commandments: Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind and thy neighbour as thyself. On these two commandments hang all the Law and the Prophets.
%A quote of the Athanasian Creed
%Fides autem est, ut unum Deum in Trinitáte, et Trinitátem in unitáte veneréris, neque confundéndo Persónas, neque substantiam separándo. Alia est enim Persóna Patris, ália Fili, ália Spíritus Sancti; sed horum trium una est substantia, et non nisi una Divinitas.
And the Faith is, that we worship one God in Trinity, and Trinity in Unity; neither confusing the Persons: nor dividing the substance. For there is one Person of the Father, another of the Son: and another of the Holy Ghost. 
%My translation based off of the Athanasian Creed, with reflection on the Orthodox Ritual:
But the Godhead of these three is all one.
%Sed horum trium una est substantia, et non nisi una Divinitas.
}{℣. Si vis habére vitam {\ae}térnam, serva mandáta. Díliges Dóminum Deum tuum ex toto corde tuo, et ex tota ánima tua, et ex tota mente tua, et próximum tuum sicut te ipsum. In his duóbus mandátis tota Lex pendet, et Prophét{\ae}. Fides autem est, ut unum Deum in Trinitáte, et Trinitátem in unitáte veneréris, neque confundéndo persónas, neque substántiam separándo. Alia est enim persóna Patris, ália Fílii, ália Spíritus Sancti: sed horum trium una est substántia, et nónnisi una Divínitas.}
\begin{rubric}
	The Priest then questions the Catechumen,
\end{rubric}
\elcol{℣. \textit{N.}, Dost thou renounce Satan?}{℣. Abrenúntias Sátan{\ae}?}

\elcol{℟. I do renounce him.}{℟. Abrenúntio.}

\elcol{℣. And all his works?}{℣. Et ómnibus opéribus ejus?}

\elcol{℟. I do renounce them.}{℟. Abrenúntio.}

\elcol{℣. And all his allurements?}{℣. Et ómnibus pompis ejus?}

\elcol{℟. I do renounce them.}{℟. Abrenúntio.}

\elcol{℣. \textit{N.}, Dost thou believe in God the Father Almighty, Maker of heaven and earth?}{℣. Credis in Deum Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}?}

\elcol{℟. I do believe.}{℟. Credo.}

\elcol{℣. Dost thou believe in Jesus Christ, his only Son, our Lord, who was born into this world, and who suffered?}{℣. Credis in Jesum Christum Fílium ejus únicum, Dóminum nostrum, natum et passum?}

\elcol{℟. I do believe.}{℟. Credo.}

\elcol{℣. Dost thou also believe in the Holy Ghost, the Holy Catholic Church, the Communion of Saints, the Forgiveness of Sins, the Resurrection of the Body, and the Life Everlasting?}{℣. Credis et in Spíritum Sanctum, sanctam Ecclésiam Cathólicam, Sanctórum communiónem, remissiónem peccatórum, carnis resurrectiónem et vitam {\ae}térnam ?}

\elcol{℟. I do believe.}{℟. Credo.}

\begin{rubric}
	The Priest now breathes three times softly upon the face of the Catechumen, and says:
\end{rubric}
\elcoldent{Go out of \textit{him} thou  unclean spirit, and give place unto the Holy Spirit the Paraclete.}{Exi ab \textit{eo}, spíritus immúnde, et da locum Spirítui Sancto Paráclito.}
\begin{rubric}
	The Priest then breathes in the form of the Cross upon his face, saying,
\end{rubric}
%The Golden Manual is consulted (https://www.google.com/books/edition/The_Golden_Manual_Or_Guide_to_Catholic_D/TLJVAAAAcAAJ?hl=en&gbpv=1&dq=Accipe+Spiritum+bonum+per+istam+insufflationem&pg=PT7&printsec=frontcover).
\elcol{\noindent\textit{N.}, through this rite of breathing upon thee, receive the good spirit and the bless {\ding{64}} ing of God. Peace be with thee.}{\noindent\textit{N.} Accipe Spíritum bonum per istam insufflatiónem, et Dei bene {\ding{64}} dictiónem. Pax tibi.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\begin{rubric}
	He then makes the sign of the cross upon the forehead and breast of the Catechumen, while he says:
\end{rubric}
%Golden Manual:
\elcoldent{Receive the sign of the Cross both upon thy forehead {\ding{64}} and also on the heart {\ding{64}}; take unto thee the faith of the heavenly precepts: and be thou such in thy conversation, that thou mayest now be the temple of God. Having entered into the Church of God, be happy in knowing that thou hast escaped the snares of death.}{Accipe signum Crucis tam in fron {\ding{64}} te, quam in cor {\ding{64}} de, sume fidem c{\ae}léstium pr{\ae}ceptórum. Talis esto móribus, ut templum Dei jam esse possis: \textit{ingressúsque} ecclésiam Dei, evasísse te láqueos mortis, \textit{l{\ae}tus} agnósce.}
%Orthodox Ritual omits these parts of the Rituale. Golden Manual translation used:
\begin{rubric}
	If the Catechumen is from paganism or idolatry,
\end{rubric}
\elcoldent{Abhor false gods. Reject idols.}{Horrésce idóla, réspue simulácra.}

%MANUAL ADJUSTMENT:
\needspace{3\baselineskip}
\begin{rubric}
	from Judaism,
\end{rubric}
\elcoldent{Abhor the Jewish perfidy. Reject the Hebrew superstition.}{Horrésce Judáicam perfídiam, réspue Hebráicam superstitiónem.}
\begin{rubric}
	from Islam,
\end{rubric}
\elcoldent{Abhor the perfidy of the Mahotemans. Reject the depraved sect of infidelity.}{Horrésce Mahuméticam perfídiam, réspue pravam sectam infidelitátis.}
\begin{rubric}
	from Heretics, whose Baptism is not accepted,
\end{rubric}
\elcoldent{Abhor heretical depravity. Reject the foul sects of the impious.}{Horrésce h{\ae}réticam pravitátem, réspue nefárias sectas impiórum.}
\begin{rubric}
	For all, it continues,
\end{rubric}
%Orthodox Ritual deviates from Latin. Adjusted properly:
\elcoldent{Worship only God, the Father almighty, and Jesus Christ, his only Son, our Lord, who shall come to judge the quick and the dead, and the world by fire.\par
℟. Amen.}{Cole Deum Patrem omnipoténtem, et Jesum Christum, Fílium ejus únicum, Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\oremus
%Orthodox Ritual:
\elcol{\lett{O}{Lord,} holy Father, almighty and eternal God, I pray thee, show thy \textit{servant}, \textit{N.}, the way of the true faith and the knowledge of thyself. For \textit{he} wandereth in uncertainty and doubt through the darkness of this world. Then, when the eyes of \textit{his} heart have been opened, \textit{he} will profess thee to be the one God, the Father, the Son, and the Holy Spirit. May \textit{he} enjoy the fruit of this profession of faith both here and in the world to come. Through Christ our Lord.\par
℟. Amen.}{\lett{T}{e} déprecor, Dómine sancte, Pater omnípotens, {\ae}térne Deus: ut huic \textit{fámulo tuo \textit{N.}, qui} in hujus s{\ae}culi nocte vagátur \textit{incértus} ac \textit{dúbius}, viam veritátis et agnitiónis tu{\ae} júbeas demonstrári; quátenus, reserátis óculis cordis sui, te unum Deum Patrem in Fílio, et Fílium in Patre cum Spíritu Sancto recognóscat, atque hujus confessiónis fructum, et hic, et in futúro s{\ae}culo percípere mereátur. Per Christum Dóminum nostrum.\par
℟. Amen.}
\begin{rubric}
	The Priest, with his thumb, then signs the seven parts of the Body with the Cross, starting with the forehead,
\end{rubric}
%The Orthodox Ritual's translation here is very free.
\elcoldent{I sign thee on the forehead, {\ding{64}} that thou mayest take up the Lord's Cross.}{Signo tibi fron {\ding{64}} tem, ut suscípias Crucem Dómini.}
\elcoldent{\inrub{On the ears:} I sign thee on the ears, {\ding{64}} that thou mayest listen attentively to God's commands.}{\inrub{In auribus:}  Signo tibi au {\ding{64}} res, ut áudias divína pr{\ae}cépta.}
\elcoldent{\inrub{On the eyes:} I sign thee on the eyes, {\ding{64}} that thou mayest see God's glory.}{\inrub{In oculis:} Signo tibi ócu {\ding{64}} los, ut vídeas claritátem Dei.}
\elcoldent{\inrub{On the nostrils:} I sign thee on the nostrils, {\ding{64}} that thou mayest perceive the sweet fragrance of Christ.}{\inrub{In naribus:} Signo tibi na {\ding{64}} res, ut odórem suavitátis Christi sántias}
\elcoldent{\inrub{On the mouth:} I sign thee on the mouth, {\ding{64}} that thou mayest speak the words of life.}{\inrub{In ore:} Signo tibi os {\ding{64}}, ut loquáris verba vit{\ae}.}
\elcoldent{\inrub{On the breast:} I sign thee on the breast, {\ding{64}} that thou mayest believe in God.}{\inrub{In pectore:} Signo tibi pe {\ding{64}} ctus, ut credas in Deum.}
\elcoldent{\inrub{On the shoulders:} I sign thee on the shoulders, {\ding{64}} that thou mayest take upon thyself the yoke of his service.}{\inrub{In scapulis:} Signo tibi scá {\ding{64}} pulas, ut suscípias jugum servitútis ejus.}
\begin{rubric}
	The Priest then signs the Catechumen's whole body (without physical touch),
\end{rubric}
%orthodox Ritual:
\elcoldent{I sign thy whole being, in the name of the Father, {\ding{64}} and of the Son, {\ding{64}} and of the Holy {\ding{64}} Spirit, that thou mayest have etenal life and live for ever.\par
℟. Amen.}{Signo te totum in nómine Pa {\ding{64}} tris, et Fí {\ding{64}} lii, et Spíritus {\ding{64}} Sancti, ut hábeas vitam {\ae}térnam, et vivas in s{\ae}cula s{\ae}culórum.\par
℟. Amen.}
\oremus
\elcol{\lett{W}{e} beseech thee, O Lord, graciously hear our prayers, and keep with thy perpetual power this thine Elect, \textit{N.}, who hath been signed with the sign of the Cross of the Lord: that \textit{he}, preserving the rudiments of the greatness of thy glory, may by keeping of thy commandments be worthy to attain unto the glory of regeneration. Through Christ our Lord.\par
℟. Amen.}{\lett{P}{reces} nostras, qu{\ae}sumus Dómine, cleménter exáudi: et \textit{hunc Eléctum tuum}, \textit{N.}, Crucis Domínic{\ae} impressióne \textit{signátum} perpétua virtúte custódi: ut magnitúdinis glóri{\ae} tu{\ae} rudiménta servans, per custódiam mandatórum tuórum ad regeneratiónis glóriam perveníre mereátur. Per Christum Dominum nostrum.\par
℟. Amen.}
\oremus
\elcol{\lett{O}{God,} Creator of the human race and also its Redeemer, be merciful to the people thou hast adopted. Enrol these reborn children in the New Covenant. Then, as children of promise, they will joyfully receive by grace what they could not attain by nature. Through Christ our Lord.\par
℟. Amen.}{\lett{D}{eus,} qui humáni géneris ita es cónditor, ut sis étiam reformátor, propitiáre pópulis adoptívis, et novo testaménto sóbolem nov{\ae} prolis adscríbe: ut fílii promissiónis, quod non potuérunt ássequi per naúram, gáudeant suscepísse per grátiam. Per Christum Dóminum nostrum.\par
℟. Amen.}
\begin{rubric}
	The Priest imposes his hands upon the hand of the Elect and then extends them over him, saying,
\end{rubric}
\oremus
%From Baptism of Infants until the last part:
\elcol{\lett{A}{lmighty} and everlasting God, Father of Our Lord Jesus Christ, vouchsafe to look upon this thy \textit{servant} \textit{N.}, whom thou hast been pleased to call to the rudiments of the faith: drive far from \textit{him} all blindness of heart: break all the snares of Satan, wherewith \textit{he} hath been bound: open to \textit{him}, O Lord, the gate of thy mercy, that, being filled with the sign of thy wisdom, \textit{he} may be free from all evil desires, and in the sweet savour of thy commandments may joyfully serve thee in thy Church, and prosper from day to day. Through the same Christ our Lord.\par
℟. Amen.}{\lett{O}{mn\smash{í}potens,} sempitérne Deus, Pater Dómini nostri Jesu Christi, respícere dignáre super \textit{hunc fámulum tuum N., quem} ad rudiménta fídei vocáre dignátus es: omnem c{\ae}citátem cordis ab \textit{eo} expélle: disrúmpe omnes láqueos sátan{\ae}, quibus fúerat \textit{colligátus}: áperi ei, Dómine, jánuam pietátis tu{\ae}, ut signo sapiénti{\ae} tu{\ae} \textit{imbútus}, ómnium cupiditátum f{\oe}tóribus cáreat, et ad suávem odórem pr{\ae}ceptórum tuórum \textit{l{\ae}tus} tibi in Ecclésia tua desérviat, et profíciat de die in diem, ut \textit{idóneus} efficiátur accédere ad grátiam Baptísmi tui, percépta medicína. Per eúmdem Christum Dóminum nostrum.\par
℟. Amen.}
\begin{rubric}
If the salt has not already been blest, it is here blest by the Priest.
\end{rubric}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par\noindent
\subby{Blessing of Salt}
\elcol{\lett{I}{exorcise} thee, O creature of salt, in the Name of God the Father {\ding{64}} almighty, and in the charity of our Lord Jesus {\ding{64}} Christ, and in the power of the Holy {\ding{64}} Ghost.}{\lett{E}{xorc\smash{í}zo} te, creatúra salis, in nómine Dei {\ding{64}} Patris omnipoténtis, et in caritáte Dómini nostri Jesu {\ding{64}} Christi, et in virtúte Spíritus {\ding{64}} Sancti.}

\elcol{\lett{I}{exorcise} thee through God {\ding{64}} the living, through God {\ding{64}} the true, through God {\ding{64}} the holy, through God {\ding{64}} who created thee for the protection of mankind, and commanded thee to be consecrated through his servants for the people that should come to believe; that in the Name fo the holy Trinity thou be made a saving sacrament to put to flight the enemy.}{\lett{E}{xor\smash{í}zo} te per Deum {\ding{64}} vivum, per Deum {\ding{64}} verum, per Deum {\ding{64}} sanctum, per Deum {\ding{64}}, qui te ad tutélam humáni géneris procreávit, et pópulo veniénti ad credulitátem per servos suos consecrári pr{\ae}cépit: ut in nómine sanct{\ae} Trinitátis efficiáris salutáre Sacraméntum ad effugándum inimícum.}

\elcol{\lett{T}{herefore,} we pray thee, O Lord our God, that sanctifying thou wouldest {\ding{64}} sanctify, and blessing thou wouldest {\ding{64}} bless this creature of salt, that it be made to all who receive it perfect healing, abiding in their members, in the Name of the same our Lord Jesus Christ, who shall come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{P}{ro\smash{í}nde} rogámus te, Dómine, Deus noster, ut hanc creatúram salis sanctificándo sanctí {\ding{64}} fices, et benedicéndo bene {\ding{64}} dícas ut fiat ómnibus accipiéntibus perfécta medicína, pérmanens in viscéribus eórum, in nómine ejúsdem Dómini nostri Jesu Christi, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par\noindent
\subby{Imposition of Salt}
\begin{rubric}
	If the Catechumen comes from paganism or idolatry, the following Collect is said by the Priest,
\end{rubric}
\oremus
%Orthodox Ritual deviates from Latin. Golden Manual used:
\elcol{\lett{O}{Lord,} holy Father, almighty and eternal God, yesterday, who art, who wast, and who remainest even unto the end, whose beginning is unknown, nor can the end be found out; we humbly beseech thee for this thy \textit{servant}, \textit{N.}, whom thou hast delivered from the error of the Gentiles, and their abominable conversation; vouchsafe to hear \textit{him} who bows down unto thee \textit{his} neck unto the font of purification, that being regenerate by water and the Holy Spirit, having put off the old man, \textit{he} may put on the new, which is created according to thee, may receive the uncorrupt and spotless garment, and may merit to serve thee, our God. Through Christ our Lord.
Through Christ our Lord.\par
℟. Amen.}{\lett{D}{\smash{ó}mine,} sancte, Pater omnípotens, {\ae}térne Deus, qui es, qui eras, et qui pérmanes usque in finem, cujus orígo nescítur, nec finis comprehéndi potest: te súpplices invocámus super \textit{hunc fámulum tuum N., quem} liberásti de erróre gentílium, et conversatióne turpíssima: dignáre exaudíre eum, \textit{qui} tibi cervíces suas humíliat ad lavácri fontem, ut, \textit{renátus} ex aqua et Spíritu Sancto, \textit{exspoliátus} véterem hóminem, índuat novum, qui secúndum te creátus est; accípiat vestem incorrúptam, et immaculátam, tibíque Deo nostro servíre mereátur. Per Christum Dóminum nostrum.\par
℟. Amen.}
\begin{rubric}
The Priest now puts a small quantity of the blessed salt into the mouth of the Catechumen, saying:
\end{rubric}
%Orthodox Ritual deviates from Latin. Golden Manual used:
\elcoldent{\textit{N.}, receive the salt of wisdom: may it be to thee a propitiation unto life everlasting.}{\textit{N.}, Accipe sal sapiénti{\ae}: propitiátio sit tibi in vitam {\ae}térnam.}

\elcol{℟. Amen.}{℟. Amen.}

\elcol{℣. Peace be with thee.}{℣. Pax tibi.}

\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\oremus
\elcol{\lett{O}{God} of our fathers, O God the author of all truth, we humbly beseech thee, that thou wouldest vouchsafe mercifully to look upon this thy \textit{servant} \textit{N.}, and that thou wouldest suffer \textit{him} no longer to hunger, now tasting this first relish of salt, to the end that, being fulfilled with heavenly food, \textit{he} may ever be fervent in spirit, rejoicing in hope, ever serving thy Name.}{\lett{D}{eus} patrum nostrórum, Deus univers{\ae}	 cónditor veritátis, te súpplices exorámus, ut \textit{hunc fámulum tuum N.} respícere dignéris propítius, et hoc primum pábulum salis gustántem, non diútius esuríre permíttas, quo minus cibo expleátur c{\ae}lésti, quátenus sit semper spíritu fervens, spe gaudens, tuo semper nómini sérviens.}

\needspace{4\baselineskip}
\elcol{\lett{L}{ead} \textit{him}, O Lord, we beseech thee, to the washing of the new birth, that \textit{he} may with thy faithful be worthy to attain unto the eternal rewards of thy promises. Through Christ our Lord.\par
℟. Amen.}{\lett{P}{er} duc \textit{eum}, Dómine, qu{\ae}sumus, ad nov{\ae} regeneratiónis lavácrum, ut cum fidélibus tuis promissiónum tuárum {\ae}térna pr{\ae}mia cónsequi mereátur. Per Christum Dóminum nostrum.\par
℟. Amen.}

\subby{Exorcisms for Men}
\begin{rubric}
	If there be women among the Elect, they stand to the side for while the males receive the following exorcisms.
\end{rubric}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par\noindent
%Orthodox Ritual deviates from Latin here and uses language which deviates from the Missal (`let us bow the knee' \& `Arise'), so the latter is preferred.
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécte, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather:} Make the sign of the Cross on him.}{℣. \inrub{Patrino:} Signa eum.}
\begin{rubric}
	The Priest says,
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches them out over him, saying,
\end{rubric}
\oremus
\elcol{\lett{O}{God} of Abraham, God of Isaac, God of Jacob, thou appearedst on Mount Sinai to thy servant, Moses; thou leddest the children of Israel out of Egypt, graciously appointing an Angel to guard them day and night. We ask thee, Lord, to send thy holy Angel from heaven to guard in the same way thy servant, \textit{N.}, and to lead him to the grace of thy baptism. Through Christ our Lord.\par
℟. Amen.}{\lett{D}{eus} Abraham, Deus Isaac, Deus Jacob, Deus, qui Móysi fámulo tuo in monte Sínai apparuísti, et fílios Isra{\"e}l de terra {\AE}gypti eduxísti, députans eis Angelum pietátis tu{\ae}, qui custodíret eos die ac nocte: te qu{\ae}sumus, Dómine; ut míttere dignéris sanctum Angelum tuum de c{\ae}lis, qui simíliter custódiat et hunc fámulum tuum \textit{N.} et perdúcat eum ad grátiam Baptísmi tui. Per Christum Dóminum nostrum.\par
℟. Amen.}
\needspace{4\baselineskip}
\elcol{\lett{T}{herefore,} accursed devil, acknowledge thy condemnation, and pay homage to the living and true God; pay homage to Jesus Christ, his Son, and to the Holy Spirit, and depart from this servant of God, \textit{N.}, for Jesus Christ, our God and Lord, hath called him to his holy grace and to the Font of baptism. Accursed devil, never dare to desecrate this sign of the holy {\ding{64}} Cross which we are tracing upon his forehead. Through the same Christ our Lord, who is come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{E}{rgo,} maledícte diábole, recognósce senténtiam tuam, et da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, et Spirítui Sancto, et recéde ab hoc fámulo Dei \textit{N.}, quia istum sibi Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam, fontémque Baptísmatis vocáre dignátus est: et hoc signum sanct{\ae} cru {\ding{64}} cis \inrub{signat singulos}, quod nos fronti ejus damus, tu, maledícte diábole, numquam áudeas violáre. Per eúmdem Christum Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
%This ritual is intentionally repeated, as indicated in the Rituale.
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécte, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather:} Make the sign of the Cross on him.}{℣. \inrub{Patrino:} Signa eum.}
\begin{rubric}
	The Priest says,
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches them out over him, saying,
\end{rubric}
\oremus
%Orthodox Ritual, corrected for sacral English
\elcol{\lett{O}{God,} thou art the unfailing support of those who call upon thee, the liberator of those who pray to thee, the peace of those who appeal to thee, the life of those who believe in thee, the resurrection of those who die in thee. I call upon thee in behalf of thy servant, \textit{N.}, that asketh for the gift of baptism and hopeth to obtain, through this spiritual rebirth, everlasting grace. Accept him, O Lord. Thou hast graciously said, `Ask, and ye shall receive' seek, and ye shall find; knock, and it shall be opened unto you.' Then grant thy favour to him that knocketh, so that, having received an everlasting blessing in his baptism, he may possess the kingdom of grace that thou hast promised. Who, with the Father and the Holy Spirit, livest and reignest ever one God, world without end.\par
℟. Amen.}{\lett{D}{eus,} immortále pr{\ae}sídium ómnium postulántium, liberátio súpplicum, pax rogántium, vita credéntium, resurréctio mortuórum: te ínvoco super hunc fámulum tuum \textit{N.}, qui, Baptísmi tui donum petens, {\ae}térnam cónsequi grátiam spirituáli regeneratióne desíderat: áccipe eum, Dómine, et quia dignátus es dícere: Pétite, et accipiétis; qu{\ae}rite, et inveniétis; pulsáte, et aperiétur vobis: peténti pr{\ae}mium pórrige, et jánuam pande pulsánti, ut, {\ae}térnam c{\ae}léstis lavácri benedictiónem consecútus, promíssa tui múneris regna percípiat : Qui cum Patre, et Spíritu Sancto vivis et regnas Deus, in s{\ae}cula s{\ae}culórum.\par
℟. Amen.}
\needspace{4\baselineskip}
%Orthodox Ritual, corrected for sacral English
\elcol{\lett{H}{ear,} accursed Satan, I adjure thee by the Name of the eternal God and of our Saviour Jesus Christ his Son. Depart with thy envy: conquered, trembling and groaning. May have thou no part in the servant of God, \textit{N.}, that already hath thoughts of heaven and is about to renounce thee and thy world and achieve a blessed immortality. Therefore, give honour to the Holy Spirit who, descending from the high throne of heaven, cometh to upset thy wiles and to make perfect the heart cleansed at the divine fountain, a temple and dwelling place dedicated to God. And thus entirely freed from the harmful effects of past sins, may this servant of God give thanks always to the eternal God and bless his Holy Name for ever and ever.\par
℟. Amen.}{\lett{A}{udi,} maledícte sátana, adjurátus per nomen {\ae}térni Dei, et Salvatóris nostri Jesu Christi Fílii ejus, cum tua victus invídia, tremens, geménsque discéde: nihil tibi sit commúne cum servo Dei \textit{N.}, jam c{\ae}léstia cogitánte, renuntiatúro tibi et s{\ae}culo tuo, et beát{\ae} immortalitáti victúro. Da ígitur honórem adveniénti Spirítui Sancto, qui, ex summa c{\ae}li arce descéndens, proturbátis fráudibus tuis, divíno fonte purgátum pectus, sanctificátum Deo templum et habitáculum perfíciat ut, ab ómnibus pénitus nóxiis pr{\ae}teritórum críminum liberátus, servus Dei grátias perénni Deo réferat semper, et benedícat nomen sanctum ejus in s{\ae}cula s{\ae}culórum.\par
℟. Amen.}
%This ritual is intentionally repeated, as indicated in the Rituale.
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécte, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather:} Make the sign of the Cross on him.}{℣. \inrub{Patrino:} Signa eum.}
\begin{rubric}
	The Priest says,
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches them out over him, saying,
\end{rubric}
\needspace{4\baselineskip}
\elcol{\lett{I}{exorcise} thee, unclean spirit, in the Name of the Father {\ding{64}}, and of the Son {\ding{64}}, and of the Holy {\ding{64}} Ghost. Come forth, depart from this servant of God, \textit{N.}, for he commandeth thee, accursed and damned spirit, he who walked upon the sea and extended his right hand to Peter as he was sinking.\par
\lett{T}{herefore,} accursed devil, acknowledge thy condemnation and pay homage to the living and true God; pay homage to Jesus Christ, his Son, and to the Holy Spirit, and depart from this servant of God, \textit{N.}, for Jesus Christ, our Lord and God, hath called him to his holy grace and to the Font of baptism. Accursed devil, never dare to desecrate this sign of the holy {\ding{64}} Cross which we are tracing upon his forehead. Through the same Christ our Lord, who is come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{E}{xorc\smash{í}zo} te, immúnde spíritus, in nómine Pa {\ding{64}} tris, et Fí {\ding{64}} lii, et Spíritus {\ding{64}} Sancti, ut éxeas, et recédas ab hoc fámulo Dei \textit{N.} Ipse enim tibi ímperat, maledícte damnáte, qui pédibus super mare ambulávit, et Petro mergénti déxteram porréxit.\par
\lett{E}{rgo,} maledícte diábole, recognósce senténtiam tuam, et da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, et Spirítui Sancto, et recéde ab hoc fámulo Dei \textit{N.}, quia istum sibi Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam, fontémque Baptísmatis vocáre dignátus est: et hoc signum sanct{\ae} cru {\ding{64}} cis \inrub{signat singulos}, quod nos fronti ejus damus, tu, maledícte diábole, numquam áudeas violáre. Per eúmdem Christum Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\subby{Exorcisms for Women}
\begin{rubric}
	If there be women among the Elect, the men stand to the side for while the females receive the following exorcisms.
\end{rubric}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par\noindent
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécta, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather or Godmother:} Make the sign of the Cross on her.}{℣. \inrub{Patrino:} Signa eam.}
\begin{rubric}
	The Priest says 
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather or Godmother approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches it out over her, saying,
\end{rubric}
\oremus
%Not present in the Orthodox Ritual. Golden Manual used:
\elcol{\lett{O}{God} of heaven, God of earth, God of angels, God of archangels, God of patriarchs, God of prophets, God of apostles, God of martyrs, God of confessors, God of virgins, God of all that live well, God, to whom every tongue confesseth, and every knee boweth, of those that are in heaven, on earth, and under the earth; I call upon thee, O Lord, for this thy handmaid, \textit{N.}, that thou mayest vouchsafe to guard her, and to bring her to the grace of thy baptism. Through Christ our Lord.\par
℟. Amen.}{\lett{D}{eus} c{\ae}li, Deus terr{\ae}, Deus Angelórum, Deus Archangelórum, Deus Patriarchárum, Deus Prophetárum, Deus Apostolórum, Deus Mártyrum, Deus Confessórum, Deus Vírginum, Deus ómnium bene vivéntium, Deus, cui omnis lingua confitétur, et omne genu fléctitur, c{\ae}léstium, terréstrium, et infernórum: te ínvoco, Dómine, super hanc fámulam tuam \textit{N.}, ut eam custodíre, et perdúcere dignéris ad grátiam Baptísmi tui. Per Christum Dóminum nostrum.\par
℟. Amen.}
%Orthodox Ritual:
\elcol{\lett{T}{herefore,} accursed devil, acknowledge thy condemnation, and pay homage to the living and true God; pay homage to Jesus Christ, his Son, and to the Holy Spirit, and depart from this handmaid of God, \textit{N.}, for Jesus Christ, our God and Lord, hath called her to his holy grace and to the Font of baptism. Accursed devil, never dare to desecrate this sign of the holy {\ding{64}} Cross which we are tracing upon her forehead. Through the same Christ our Lord, who is come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{E}{rgo,} maledícte diábole, recognósce senténtiam tuam, et da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, et Spirítui Sancto, et recéde ab hac fámula Dei \textit{N.}, quia istam sibi Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam, fontémque Baptísmatis vocáre dignátus est: et hoc signum sanct{\ae} cru {\ding{64}} cis \inrub{signat singulos}, quod nos fronti ejus damus, tu, maledícte diábole, numquam áudeas violáre. Per eúmdem Christum Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécta, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather or Godmother:} Make the sign of the Cross on her.}{℣. \inrub{Patrino:} Signa eam.}
\begin{rubric}
	The Priest says 
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather or Godmother approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches it out over her, saying,
\end{rubric}
\oremus
\elcol{\lett{O}{God} of Abraham, God of Isaac, God of Jacob, thou appearedst on Mount Sinai to thy servant, Moses; thou leddest the children of Israel out of Egypt, graciously appointing an Angel to guard them day and night. We ask thee, Lord, to send thy holy Angel from heaven to guard in the same way thy handmaid, \textit{N.}, and to lead her to the grace of thy baptism. Through Christ our Lord.\par
℟. Amen.}{\lett{D}{eus} Abraham, Deus Isaac, Deus Jacob, Deus, qui Móysi fámulo tuo in monte Sínai apparuísti, et fílios Isra{\"e}l de terra {\AE}gypti eduxísti, députans eis Angelum pietátis tu{\ae}, qui custodíret eos die ac nocte: te qu{\ae}sumus, Dómine; ut míttere dignéris sanctum Angelum tuum de c{\ae}lis, qui simíliter custódiat et hanc fámulam tuam \textit{N.} et perdúcat eam ad grátiam Baptísmi tui. Per Christum Dóminum nostrum.\par
℟. Amen.}
\subby{Exorcism}
\elcol{\lett{T}{herefore,} accursed devil, acknowledge thy condemnation, and pay homage to the living and true God; pay homage to Jesus Christ, his Son, and to the Holy Spirit, and depart from this handmaid of God, \textit{N.}, for Jesus Christ, our God and Lord, hath called her to his holy grace and to the Font of baptism. Accursed devil, never dare to desecrate this sign of the holy {\ding{64}} Cross which we are tracing upon her forehead. Through the same Christ our Lord, who is come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{E}{rgo,} maledícte diábole, recognósce senténtiam tuam, et da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, et Spirítui Sancto, et recéde ab hac fámula Dei \textit{N.}, quia istam sibi Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam, fontémque Baptísmatis vocáre dignátus est: et hoc signum sanct{\ae} cru {\ding{64}} cis \inrub{signat singulos}, quod nos fronti ejus damus, tu, maledícte diábole, numquam áudeas violáre. Per eúmdem Christum Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\elcol{℣. Pray, O elect, and bow the knee, and say: Our Father.}{℣. Ora, Elécta, flecte génua, et dic: Pater noster.}
\elcol{℟. Our Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil.}{℟. Pater noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo.}
\elcol{℣. Arise, complete thy prayer, and say: Amen.}{℣. Leva, comple oratiónem tuam, et dic: Amen.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. \inrub{To the Godfather or Godmother:} Make the sign of the Cross on her.}{℣. \inrub{Patrino:} Signa eam.}
\begin{rubric}
	The Priest says 
\end{rubric}
\elcol{℣. \inrub{To the Elect:} Come forward.}{℣. \inrub{Electo:} Accéde.}
\begin{rubric}
	The Godfather or Godmother approaches and signs the Elect on the forehead, saying:
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest then also makes the sign of the Cross on the forehead of the Elect (only once), saying,
\end{rubric}
\elcoldent{In the Name of the Father, and of the Son, and of the Holy Spirit.}{In nómine Patris, et Fílii, et Spíritus Sancti.}
\begin{rubric}
	The Priest imposes his hand upon the Elect, then stretches it out over her, saying,
\end{rubric}
\oremus
%Not in Orthodox Ritual:
\elcol{\lett{I}{exorcise} thee, unclean spirit, by the Father {\ding{64}}, and the Son {\ding{64}}, and the Holy {\ding{64}} Ghost, that thou go out and depart from this handmaid of God, \textit{N.}; for he commandeth thee, accursed one, who opened the eyes of him that was born blind, and raised Lazarus, after four days, from the grave.}{\lett{E}{xorc\smash{í}zo} te, immúnde spíritus, per Pa {\ding{64}} trem, et Fí {\ding{64}} lium, et Spíritum {\ding{64}} Sanctum, ut éxeas, et recédas ab hac fámula Dei \textit{N.} Ipse enim tibi ímperat, maledícte damnáte, qui c{\ae}co nato óculos apéruit, et quatriduánum Lázarum de monuménto suscitávit.}
\needspace{4\baselineskip}
\elcol{\lett{T}{herefore,} accursed devil, acknowledge thy condemnation, and pay homage to the living and true God; pay homage to Jesus Christ, his Son, and to the Holy Spirit, and depart from this handmaid of God, \textit{N.}, for Jesus Christ, our God and Lord, hath called her to his holy grace and to the Font of baptism. Accursed devil, never dare to desecrate this sign of the holy {\ding{64}} Cross which we are tracing upon her forehead. Through the same Christ our Lord, who is come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{E}{rgo,} maledícte diábole, recognósce senténtiam tuam, et da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, et Spirítui Sancto, et recéde ab hac fámula Dei \textit{N.}, quia istam sibi Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam, fontémque Baptísmatis vocáre dignátus est: et hoc signum sanct{\ae} cru {\ding{64}} cis \inrub{signat singulos}, quod nos fronti ejus damus, tu, maledícte diábole, numquam áudeas violáre. Per eúmdem Christum Dóminum nostrum, qui ventúrus est judicáre vivos et mórtuos, et s{\ae}culum per ignem.\par
℟. Amen.}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par\noindent
\subby{Blessing of the Catechumens}
\begin{rubric}
	If there be many Catechumens of both sexes, the men return, the men being on the Priest's right and the women on his left, as at the beginning.
\end{rubric}
\begin{rubric}
    The Priest then imposes his hand upon the Catechumen (if there be many, he imposes on each individually and then raises it over all of them) and says,
\end{rubric}
\oremus
%Orthodox Ritual:
\elcol{\lett{I}{entreat} thy eternal and most just mercy, O Lord holy, Father Almighty, everlasting God, author of light and truth, upon \textit{this servant} of God, \textit{N.}, that thou wouldest vouchsafe to enlighten \textit{him} with the light of thine understanding: cleanse \textit{him} and sanctify \textit{him}: give \textit{him} true knowledge, that being made worthy of the grace of thy Baptism, \textit{he} may hold firm hope, right counsel, and holy doctrine. Through Christ our Lord.\par
℟. Amen.}{\lett{AE}{t\smash{é}rnam,} ac justíssimam pietátem tuam déprecor, Dómine sancte, Pater omnípotens, {\ae}térne Deus, auctor lúminis et veritátis, super \textit{hunc fámulum tuum} \textit{N.} ut dignéris \textit{eum} illumináre lúmine intelligénti{\ae} tu{\ae}: munda \textit{eum}, et sanctífica: da ei sciéntiam veram, ut \textit{dignus} efficiátur accédere ad grátiam Baptísmi tui, téneat firmam spem, consílium rectum, doctrínam sanctam, ut \textit{aptus} sit ad percipiéndam grátiam tuam. Per Christum Dóminum nostrum.\par
℟. Amen.}
\subby{Entrance into the Church}
\begin{rubric}
	The Priest holds out his left hand on the Elect's right arm, or the end of his stole on the left side to the Elect, introducing him into the Church (if there be many Catechumens, it is done individually with each), while saying,
\end{rubric}%Orthodox Ritual deviates from the Latin. Golden Manual followed:
\elcoldent{\textit{N.}, Enter into the holy Church of God, that thou mayest receive heavenly blessing from the Lord Jesus Christ, and may have thy portion with him and with his saints.}{\textit{N.}, Ingrédere in sanctam ecclésiam Dei, ut accípias benedictiónem c{\ae}léstem a Dómino Jesu Christo, et hábeas partem cum illo et Sanctis ejus. }
\elcol{℟. Amen.}{℟. Amen.}
\begin{rubric}
    When the Elect enters, he either prostrates at the entrance or prostrates himself on the Pavement (in the centre right before the Altar Rail and Chancel) and adores (if there be many, they all do so together).
\end{rubric}
\begin{rubric}
    Then he rises (if there be many, they all do so together), and the Priest lays his hand on his head (each individually, if there be many), and the Elect recites with him the Nicene Creed and the Lord's Prayer (if there be many, they all do so together).
\end{rubric}
\elcol{\lett{I}{believe} in one God, the Father Almighty, Maker of heaven and earth, And of all things visible and invisible:\par
    And in one Lord Jesus Christ, the only-begotten Son of God; Begotten of his Father before all worlds, God of God, Light of Light, Very God of very God; Begotten, not made; Being of one substance with the Father; By whom all things were made: Who for us men and for our salvation came down from heaven, And was incarnate by the Holy Ghost of the Virgin Mary, And was made man: And was crucified also for us under Pontius Pilate; He suffered and was buried: And the third day he rose again according to the Scriptures: And ascended into heaven, And sitteth on the right hand of the Father: And he shall come again, with glory, to judge both the quick and the dead; Whose kingdom shall have no end.\par
    And I believe in the Holy Ghost, The Lord, and Giver of Life, Who proceedeth from the Father; Who with the Father and the Son together is worshiped and glorified; Who spake by the Prophets: And I believe one, holy, catholic, and apostolic Church: I acknowledge one Baptism for the remission of sins: And I look for the Resurrection of the dead: {\ding{64}} And the Life of the world to come. Amen.}
    {\lett{C}{redo} in unum Deum, Patrem omnipoténtem, factórem c{\ae}li et terr{\ae}, visibílium ómnium et invisibílium.\par
    Et in unum Dóminum Jesum Christum, Fílium Dei unigénitum. Et ex Patre natum ante ómnia sǽcula. Deum de Deo, lumen de lúmine, Deum verum de Deo vero. Génitum, non factum, consubstantiálem Patri: per quem ómnia facta sunt. Qui propter nos hómines et propter nostram salútem descéndit de c{\ae}lis. Et incarnátus est de Spíritu Sancto ex María Vírgine: Et homo factus est. Crucifíxus étiam pro nobis: sub Póntio Piláto passus, et sepúltus est. Et resurréxit tértia die, secúndum Scriptúras. Et ascéndit in c{\ae}lum: sedet ad déxteram Patris. Et íterum ventúrus est cum glória judicáre vivos et mórtuos: cujus regni non erit finis.\par
    Et in Spíritum Sanctum, Dóminum et vivificántem: qui ex Patre procédit. Qui cum Patre et Fílio simul adorátur et conglorificátur: qui locútus est per Prophétas. Et unam sanctam cathólicam et apostólicam Ecclésiam. Confíteor unum baptísma in remissiónem peccatórum. Et exspécto resurrectiónem mortuórum. {\ding{64}} Et vitam ventúri s{\ae}culi. Amen.}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\begin{rubric}
	The Priest then turning his back to the doors of the Baptistry, imposes his hand upon the head of the Elect (each Elect), and afterwards, extending his hands, he says,
\end{rubric}
\elcol{\lett{T}{he} threats of punishments and tortures are not hidden from thee, Satan, the day of judgement, the day of pain, the day which is to come like a fiery furnace, in which unending destruction is prepared for thee and all thine angels. Therefore, accursed and damnable one, pay homage to the living and true God; pay homage to Jesus Christ, his Son, give homage to the Holy Spirit, the Consoler, in whose Name and power I command thee, unclean spirit, whoever thou mayest be, go forth and depart from \textit{this servant} of God, \textit{N.}, whom this day Jesus Christ, our God and Lord, hath called to his holy grace and blessing, and to the Font of baptism. May \textit{he} become the temple of God through the water of rebirth unto remission of all \textit{his} sins. In the Name of the same Jesus Christ our Lord, who shall come to judge the living and the dead, and the world by fire.\par
℟. Amen.}{\lett{N}{ec} te latet, sátana, imminére tibi p{\oe}nas, imminére tibi torménta, imminére tibi diem judícii, diem supplícii sempitérni; diem, qui ventúrus est velut clíbanus ardens, in quo tibi, atque univérsis ángelis tuis pr{\ae}parátus sempitérnus erit intéritus. Proínde, damnáte, atque damnánde, da honórem Deo vivo et vero, da honórem Jesu Christo Fílio ejus, da honórem Spirítui Sancto Paráclito, in cujus nómine atque virtúte pr{\ae}cípio tibi, quicúmque es, spíritus immúnde, ut éxeas, et recédas ab \textit{hoc fámulo Dei \textit{N.}, quem} hódie idem Deus et Dóminus noster Jesus Christus ad suam sanctam grátiam et benedictiónem, fontémque Baptísmatis dono vocáre dignátus est: ut fiat ejus templum per aquam regeneratiónis in remissiónem ómnium peccatórum. In nómine ejúsdem Dómini nostri Jesu Christi, qui ventúrus est judicáre vivos et mórtuos et s{\ae}culum per ignem.\par
℟. Amen.}
\begin{rubric}
	Then the Priest puts spit from his mouth into his thumb, then makes the sign of the Cross on the right, and afterwards on the left, ear of the Catechumen, and says once only:
\end{rubric}
\elcol{Ephthatha, {\ding{64}} that is, {\ding{64}} Be thou opened.}{Ephpheta, {\ding{64}} quod est, {\ding{64}} Adaperíre.}
\begin{rubric}
	And then, signing the nostrils, he adds:
\end{rubric}
\elcol{For a savour {\ding{64}} of sweetness.}{In odórem suavitátis.}
\begin{rubric}
	Lastly, in a louder tone of voice, he says,
\end{rubric}
\elcol{But thou, devil, flee away; for the judgement of God draweth nigh.}{Tu autem effugáre, diábole; appropinquábit enim judícium Dei.}
\begin{rubric}
	He then questions the Catechumen, asking,
\end{rubric}
\elcol{℣. What is thy name?}{℣. Quis vocáris?}

\elcol{℟. N.}{℟. N.}

\elcol{℣. \textit{N.}, Dost thou renounce Satan?}{℣. \textit{N.}, Abrenúntias Sátan{\ae}?}

\elcol{℟. I do renounce him.}{℟. Abrenúntio.}

\elcol{℣. And all his works?}{℣. Et ómnibus opéribus ejus?}

\elcol{℟. I do renounce them.}{℟. Abrenúntio.}

\elcol{℣. And all his allurements?}{℣. Et ómnibus pompis ejus?}

\elcol{℟. I do renounce them.}{℟. Abrenúntio.}
\begin{rubric}
	The Priest then dips his right thumb into the Oil of Catechumens, and with it makes the sign of the Cross on the breast, and between the shoulders of the Catechumen, saying:
\end{rubric}
\elcoldent{I anoint thee {\ding{64}} with the Oil of salvation, in Christ Jesus our Lord, that thou mayest have eternal life.}{Ego te línio {\ding{64}} óleo salútis in Christo Jesu Dómino nostro, ut habeas vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. Peace be to thee.}{℣. Pax tibi.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\begin{rubric}
	Then he wipes his thumb and the places anointed with cotton or something similar and says,
\end{rubric}
%Orthodox Ritual:
\elcol{\lett{G}{o} forth, unclean spirit, and pay homage to the living and true God. Depart, unclean spirit, and give place to Jesus Christ, his Son. Depart, unclean spirit, and give place to the Holy Spirit, the Consoler.}{\lett{E}{xi,} immúnde spíritus, et da honórem Deo vivo et vero. Fuge, immúnde spíritus, et da locum Jesu Christo Fílio ejus. Recéde, immúnde spíritus, et da locum Spirítui Sancto Paráclito.}
\begin{rubric}
	Standing outside the Chancel, the Priest takes off his violet cope and stole and puts on the same of white colour.
\end{rubric}
\begin{rubric}
    The Priest leads the Catechumen to the Baptistry, where if the water has not been blessed, it is blessed according to the appropriate rite (p. \pageref{BaptismalBlessing}).
\end{rubric}
\begin{rubric}
    When at the Font, the Priest asks,
\end{rubric}
\elcol{℣. What is thy name?}{℣. Quis vocáris?}

\elcol{℟. N.}{℟. N.}

\elcol{℣. \textit{N.}, Dost thou believe in God the Father Almighty, Creator of heaven and earth?}{℣. \textit{N.}, Credis in Deum Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}?}

\elcol{℟. I do believe.}{℟. Credo.}

\elcol{℣. Dost thou believe in Jesus Christ, his only Son our Lord, who was born into this world, and who suffered for us?}{℣. Credis in Jesum Christum Fílium ejus únicum, Dóminum nostrum, natum et passum?}

\elcol{℟. I do believe.}{℟. Credo.}

\elcol{℣. Dost thou also believe in the Holy Ghost, the Holy Catholic Church, the Communion of Saints, the Forgiveness of Sins, the Resurrection of the Body, and Life Everlasting?}{℣. Credis et in Spíritum Sanctum, sanctam Ecclésiam Cathólicam, Sanctórum communiónem, remissiónem peccatórum, carnis resurrectiónem et vitam {\ae}térnam ?}

%Orthodox Ritual deviates from Rituale:
\elcol{℟. I do believe.}{℟. Credo.}

\elcol{℣. \textit{N.}, What seekest thou?}{℣. \textit{N.}, Quid petis?}

\elcol{℟. Baptism.}{℟. Baptísmum.}

\elcol{℣. Wilt thou be baptised?}{℣. Vis baptizári?}

\elcol{℟. I will.}{℟. Volo.}

\begin{rubric}
	The Priest takes the Elect by his arms, near the shoulders, and, the upper part of his body being uncovered, the rest decently covered, the Priest, thrice immersing him, and raising him as often, thus invoking the Most Holy Trinity once only.
\end{rubric}
\begin{rubric}
    Meanwhile, the Godfather or Godmother, or both, either holds or touches him.
\end{rubric}
\elcoldent{\textit{N.}, I baptise thee in the name of the Father, \inrub{he immerses} and of the Son, \inrub{he immerses} and of the Holy Ghost \inrub{he immerses}.}{\textit{N.} Ego te baptízo in nómine Patris, \inrub{mergit} et Fílii, \inrub{mergit} et Spíritus Sancti \inrub{mergit}.}
\begin{rubric}
	After cleansing his hands, the Priest immediately moves onto Confirmation (p. \pageref{confirmation}).
\end{rubric}