\fancyhead[RO,LE]{\textit{Burial of a Child}}
\fancyhead[RE,LO]{\textit{}}
\section{At the Burial of a Child}
\begin{rubric}
    When a baptised Child dies before the age of reason, it receives this special funeral rite.\par
    \textsc{Note,} the Ministers, Holy Water, and incense should be present similar to the standard Burial rite.
\end{rubric}
\begin{rubric}
    Unlike funerals for adults, bells should be rung joyfully. The Altar should be decorated festively, and the liturgical colour is white.
\end{rubric}
\begin{rubric}
    The Priest, vested in surplice and white stole, meeting the Body, and going before it, either into the Church or towards the Grave, shall say:
\end{rubric}
\lett{I}{am} the resurrection and the life, saith the Lord: he that believeth in me, though he were dead, yet shall he live: and whosoever liveth and believeth in me, shall never die.

\lett{J}{esus} called them unto him and said, Suffer the little children to come unto me, and forbid them not: for of such is the kingdom of God.\par
\vspace{2ex}
\lett{H}{e} shall feed his flock like a shepherd: he shall gather the lambs with his arms, and carry them in his bosom.

\begin{rubric}
    When they are come into the Church, shall be said the following Psalms.
\end{rubric}

\subbypsalm{Psalm 23. \textit{Dominus regit me.}}
\lett{T}{he} Lord is my shepherd : therefore can I lack nothing.\par
\secondlinepsalm{He shall feed me in a green pasture : and lead me forth beside the waters of comfort.}
\thirdlinepsalm{He shall convert my soul : and bring me forth in the paths of righteousness, for his Name's sake.}
\psanum{4}Yea, though I walk through the valley of the shadow of death, I will fear no evil : for thou art with me; thy rod and thy staff comfort me.\par
\psanum{5}Thou shalt prepare a table before me against them that trouble me : thou hast anointed my head with oil, and my cup shall be full.\par
\psanum{6}But thy loving-kindness and mercy shall follow me all the days of my life : and I will dwell in the house of the Lord for ever.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 121. \textit{Levavi oculus}}
\lett{I}{will} lift up mine eyes unto the hills : from whence cometh my help.\par
\secondlinepsalm{My help cometh even from the Lord : who hath made heaven and earth.}
\thirdlinepsalm{He will not suffer thy foot to be moved : and he that keepeth thee will not sleep.}
\psanum{4}Behold, he that keepeth Israel : shall neither slumber nor sleep.\par
\psanum{5}The Lord himself is thy keeper : the Lord is thy defence upon thy right hand;\par
\psanum{6}So that the sun shall not burn thee by day : neither the moon by night.\par
\psanum{7}The Lord shall preserve thee from all evil : yea, it is even he that shall keep thy soul.\par
\psanum{8}The Lord shall preserve thy going out, and thy coming in : from this time forth for evermore.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\vspace{-1ex}
\readingcitation{Lesson}{Matthew 18:1}
 \lett{A}{t} the same time came the disciples unto Jesus, saying, Who is the greatest in the kingdom of heaven? And Jesus called a little child unto him, and set him in the midst of them, and said, Verily I say unto you, Except ye be converted, and become as little children, ye shall not enter into the kingdom of heaven. Whosoever therefore shall humble himself as this little child, the same is greatest in the kingdom of heaven. And whoso shall receive one such little child in my name receiveth me. Take heed that ye despise not one of these little ones; for I say unto you, That in heaven their angels do always behold the face of my Father which is in heaven.

 \begin{rubric}
    If a Mass is to be said, it may be offered here, either a Votive Mass for the Holy Angels (p. \pageref{MassOfTheAngels})---if rubrics permit---or the Mass of the Day.\par
    Otherwise, Here may be sung a Hymn or an Anthem; then shall the Minister say,
\end{rubric}

\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\fakecol{℣. Let us pray.}{℣. Orémus.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\needspace{4\baselineskip}
\fakecol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\begin{rubric}
    While the Lord's Prayer is said, the Priest thrice sprinkles the Coffin with Holy Water.\par
    \textsc{Note,} He does not go round it, nor is it censed.
\end{rubric}\par
℣. Blessed are the pure in heart;\par
℟. For they shall see God.\par
℣. Blessed be the name of the Lord;\par
℟. Henceforth, world without end.\par
℣. O Lord, hear our prayer;\par
℟. And let our cry come unto thee.\par
\begin{rubric}
    Here shall be said the following Prayers.
\end{rubric}\par
\needspace{4\baselineskip}
\lett{O}{merciful} Father, whose face the angels of thy little ones do always behold in heaven; Grant us stedfastly to believe that this thy child hath been taken into the safe keeping of thine eternal love. Through.\par
\needspace{4\baselineskip}
\lett{A}{lmighty} and merciful Father, who dost grant to children an abundant entrance into thy kingdom; Grant us grace so to conform our lives to their innocency and perfect faith, that at length, united with them, we may stand in thy presence in fulness of joy. Through.\par
\needspace{4\baselineskip}
\lett{T}{he} grace of our Lord Jesus {\ding{64}} Christ, and the love of God, and the fellowship of the Holy Ghost, be with us all evermore. \textit{Amen.}\\

\begin{rubric}
    While the Body is being borne from the church to the Grave, the following is said or sung,
\end{rubric}

\subbypsalm{Psalm 148. \textit{Laudate Dominum}}
\lett{O}{praise} the Lord of heaven : praise him in the height.\par
\secondlinepsalm{Praise him, all ye angels of his : praise him, all his host.}
\thirdlinepsalm{Praise him, sun and moon : praise him, all ye stars and light.}
\psanum{4}Praise him, all ye heavens : and ye waters that are above the heavens.\par
\psanum{5}Let them praise the Name of the Lord : for he spake the word, and they were made; he commanded, and they were created.\par
\psanum{6}He hath made them fast for ever and ever : he hath given them a law which shall not be broken.\par
\psanum{7}Praise the Lord upon earth : ye dragons, and all deeps;\par
\psanum{8}Fire and hail, snow and vapours : wind and storm, fulfilling his word;\par
\psanum{9}Mountains and all hills : fruitful trees and all cedars;\par
\psanum{10}Beasts and all cattle : worms and feathered fowls;\par
\psanum{11}Kings of the earth and all people : princes and all judges of the world;\par
\psanum{12}Young men and maidens, old men and children, praise the Name of the Lord : for his Name only is excellent, and his praise above heaven and earth.\par
\psanum{13}He shall exalt the horn of his people; all his saints shall praise him : even the children of Israel, even the people that serveth him.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%From the English 1928:
\lett{O}{God,} whose ways are hidden and thy works most wonderful, who makest nothing in vain and lovest all that thou hast made: Comfort thou thy servants, whose hearts are sore smitten and oppressed and grant that they may so love and serve thee in this life, that together with this thy child, they may obtain the fulness of thy promises in the world to come. Through.

\subby{At the Grave}
\begin{rubric}
    When they are come to the Grave shall be said or sung,
\end{rubric}
\lett{J}{esus} saith to his disciples, Ye now therefore have sorrow: but I will see you again, and your heart shall rejoice, and your joy no man taketh from you.

\begin{rubric}
    Meanwhile the Priest sprinkles both the coffin and the Grave with Holy Water and censes them.\par
    While the earth is being cast upon the Body, the Minister shall say,
\end{rubric}\par
\needspace{4\baselineskip}
\lett{I}{n} sure and certain hope of the Resurrection to eternal life through our Lord Jesus Christ, we commit the body of this child to the ground. The \divineName{Lord} bless \textit{him} and keep \textit{him}, the \divineName{Lord} make his face to shine upon \textit{him} and be gracious unto \textit{him}, the \divineName{Lord} lift up his countenance upon \textit{him}, and give \textit{him} peace, both now and evermore.

\begin{rubric}
    Then shall be said or sung,
\end{rubric}

\lett{T}{herefore} are they before the throne of God, and serve him day and night in his temple: and he that sitteth on the throne shall dwell among them.\par
    \secondline{They shall hunger no more, neither thirst any more; neither shall the sun light on them, nor any heat.}
    \thirdline{For the Lamb which is in the midst of the throne shall feed them, and shall lead them unto living fountains of waters: and God shall wipe away all tears from their eyes.}

\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{O}{God,} whose most dear Son did take little children into his arms and bless them; Give us grace, we beseech thee, to entrust the soul of this child to thy neverfailing care and love, and bring us all to thy heavenly kingdom. Through the same.\par
\needspace{4\baselineskip}
\lett{A}{lmighty} God, Father of mercies and giver of all comfort; Deal graciously, we pray thee, with all those who mourn, that, casting every care on thee, they may know the consolation of thy love. Through.\par
\needspace{4\baselineskip}
\lett{M}{ay} Almighty God, the Father, the {\ding{64}} Son, and the Holy Ghost, bless you and keep you, now and for evermore. \textit{Amen.}

\begin{rubric}
    Before leaving, the Priest sprinkles the Grave one more time with Holy Water.
\end{rubric}