%The Rite of Communion of the Sick comes from the Orthodox Ritual
\fancyhead[RO,LE]{\textit{Communion of the Sick}}
\fancyhead[RE,LO]{\textit{}}
\section{Communion of the Sick}
\begin{secrubric}
    Upon entering the room, the Priest makes preparation and says,
\end{secrubric}
\elcol{℣. Peace be to this house.}{℣. Pax huic dómui.}
\elcol{℟. And to all that dwell in it.}{℟. Et ómnibus habitántibus in ea.}
\begin{rubric}
    The Priest then takes Holy Water and sprinkles the sick person and the room while saying,
\end{rubric}
Thou shalt purge me, * O Lord, with hyssop, and I shall be clean: thou shalt wash me, and I shall be whiter than snow. \textit{Ps.} Have mercy upon me, O God, after thy great goodness. ℣. Glory be to the Father, and to the Son, and to the Holy Ghost; ℟. As it was in the beginning, is now and ever shall be, world without end. Amen.\par
Thou shalt purge me, * O Lord, with hyssop, and I shall be clean: thou shalt wash me, and I shall be whiter than snow.

\elcol{℣. O Lord, show thy mercy upon us.}{℣. Osténde nobis, Dómine, misericórdiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutáre tuum da nobis.}
\elcol{℣. O Lord, hear my prayer.}{℣. Dómine, exáudi oratiónem meam.}
\elcol{℟. And let my cry come unto thee.}{℟. Et clamor meus ad te véniat.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{G}{raciously} hear us, O Lord holy, Father almighty, everlasting God: and vouchsafe to send thy holy Angel from heaven; to guard and cherish, to protect and visit, and to defend all who dwell in this place. Through Christ our Lord. \textit{Amen.}

\begin{rubric}
    The following general confession is then said, unless a private confession be made.
\end{rubric}
\elcol{\lett{I}{confess} to God Almighty, to Blessed Mary ever Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, to all the Saints, and to thee, Father, that I have sinned exceedingly in thought, word, and deed; \inrub{Strike breast thrice.} by my fault, my own fault, my own most grievous fault. Therefore I beg blessed Mary ever Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, all the Saints, and thee, Father, to pray for me to the Lord our God.}{\lett{C}{onf\smash{í}teor} Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et tibi, pater: quia peccávi nimis cogitatióne, verbo et ópere: \inrub{Strike breast thrice.} mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et te, pater, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℣. God Almighty have mercy upon \textit{you}, forgive \textit{you} \textit{your} sins, and bring \textit{you} to everlasting life.}{℣. Misereátur \textit{vestri} omnípotens Deus, et, dimíssis peccátis \textit{vestris}, perdúcat \textit{vos} ad vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. The almighty and merciful Lord grant unto \textit{you} pardon, absolution, {\ding{64}} and remission of \textit{your} sins.}{℣. Indulgéntiam, {\ding{64}} absolutiónem et remissiónem peccatórum \textit{vestrórum} tríbuat \textit{vobis} omnípotens et miséricors Dóminus.}
\elcol{℟. Amen.}{℟. Amen.}
\begin{rubric}
    The Priest then takes the Sacrament from the Pyx and raises it saying,
\end{rubric}
\elcol{℣. Behold {\ding{64}} the Lamb of God; behold him who taketh away the sins of the world.}{℣. Ecce Agnus Dei, ecce, qui tollit peccáta mundi.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\begin{rubric}
    The Priest administers the Sacrament to the sick person saying,\par
    \textsc{Note,} the Priest may use a form of delivering the Sacrament to the People from the Mass.
\end{rubric}
\begin{center}
    \textsc{The Body and Blood of Our Lord Jesus Christ preserve thy soul unto everlasting life.}
\end{center}
\begin{rubric}
    The Priest cleanses his fingers and then says.
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{O}{Lord} holy, Father almighty, everlasting God, we faithfully beseech thee that the most Sacred Body of our Lord Jesus Christ thy Son, which our \textit{brother} hath received, may avail for the healing both of body and soul: who liveth and reigneth with thee, in the unity of the Holy Ghost, ever one God, world without end. \textit{Amen.}

\lett{T}{he} blessing of God almighty, the Father, {\ding{64}} the Son, and the Holy Ghost, descend upon \textit{you} and remain with \textit{you} always. \textit{Amen.}

%\begin{center}
%    \textsc{Here Endeth Communion of the Sick.}
%\end{center}