\section{Sacrament of Penance}
\fancyhead[RO,LE]{\textit{Penance}}
%Penance comes from the Monastic Diurnal
\begin{multicols}{2}
\begin{rubric}
	The Pentient begins with the following.
\end{rubric}
Bless me, Father, for I have sinned.\par

℣. The Lord be in thy heart and upon thy lips, that so thou mayest worthily confess all thy sins; In the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost. Amen.\par
\begin{rubric}
	{The Penitent then confesses his sins using the following or similar formula:}
\end{rubric}
\label{Confiteor}
I confess to God Almighty, to Blessed Mary Ever-Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, to all the Saints, and to thee, Father, that I have sinned exceedingly in thought, word, and deed, by my fault, by my own fault, by my own most grievous fault,\par
Especially I accuse myself that since my last Confession, which was \inrub{Say how long} ago, I have committed the following sins. \inrub{Confess your sins and conclude with:}\par
For these and for all my other sins which I cannot now remember, I am heartily sorry, firmly purpose amendment, and humbly ask pardon of God; and of thee, my spiritual father, penance, counsel, and Absolution.\par
Wherefore I beg blessed Mary Ever-Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, all the Saints, and thee, father, to pray for me to the Lord our God.
\begin{rubric}
	{Here the Priest gives penance, counsel, and then absolution saying,}
\end{rubric}
%\elcol{
\lett{A}{lmighty} God have mercy upon thee, forgive thee thy sins, and bring thee to everlasting life. \textit{Amen.}%}{\lett{M}{isere\smash{á}tur} tui omnípotens Deus, et dimíssis peccátis tuis, perdúcat te ad vitam {\ae}térnam. Amen.}
%\elcol{
\lett{T}{he} Almighty and merciful Lord grant thee pardon, {\ding{64}} absolution, and remission of thy sins. \textit{Amen.}%}{\lett{I}{ndulg\smash{é}ntiam,} {\ding{64}} absolutiónem, et remissiónem peccatórum tuórum tríbuat tibi omnípotens et miséricors Dóminus. Amen.}
\par\noindent
%	\centerline{\rule{0.5\textwidth}{0.4pt}}
%\par\noindent
%\elcol{Our Lord Jesus Christ absolves thee; and I, by his authority, absolve thee from every bond of excommunication, (suspension,) and interdict, so far as I have power, and thou hast need. I absolve thee from thy sins, in the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost. Amen.}{Dóminus noster Jesus Christus te absólvat; et ego auctoritáte ipsíus te absólvo ab omni vínculo excommunicatiónis, (suspensiónis,) et interdícti, in quantum possum, et tu índiges. Ego te absólvo a peccátis tuis, in nómine Patris, {\ding{64}} et Fílii, et Spíritus Sancti. Amen.}
%\begin{inhead}
%	or
%\end{inhead}
%\elcol{
\lett{O}{ur} Lord Jesus Christ, who hath left power to his Church to absolve all sinners who truly repent and believe in him, of his great mercy forgive thee thine offences: And by his authority committed to me, I absolve thee from all thy sins, In the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost. \textit{Amen.}%\footnote{The following Absolution may instead be said: Our Lord Jesus Christ, who hath left power to his Church to absolve all sinners who truly repent and believe in him, of his great mercy forgive thee thine offences: And by his authority committed to me, I absolve thee from all thy sins, In the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost.}%}{Dóminus noster Jesus Christus, qui dedit potestátem Ecclési{\ae} absolvéndi a peccátis p{\oe}niténtes, et credéntes Evangélio, ipse ex infiníta misericórdia indúlgeat tibi peccáta tua : ego vere autoritáte ipsius mihi commissa absólvo te ab ómnibus peccátis, in nómine Patris {\ding{64}} et Fílii, et Spíritus Sancti. Amen.}
%\par\noindent
%	\centerline{\rule{0.5\textwidth}{0.4pt}}
%\par\noindent
%\begin{rubric}
%	The Priest concludes with,
%\end{rubric}
%\elcol{
\lett{T}{he} merits of the Passion of our Lord Jesus Christ, the prayers of his holy Mother the Blessed Virgin Mary, and of all the Saints, whatsoever good thou hast done, or evil thou hast endured, be unto thee for the remission of sins, the increase of grace, and the reward of eternal life: And the blessing of God Almighty, the Father, {\ding{64}} the Son, and the Holy Ghost, be upon thee and remain with thee for ever. \textit{Amen.}%}{\lett{P}{\smash{á}ssio} Dómini nostri Jesu Christi, mérita Beát{\ae} Marí{\ae} Vírginis et ómnium Sanctórum, quidquid boni féceris, et mali sustinúeris, sint tibi in remissiónem peccatórum, augméntum gráti{\ae}, et pr{\ae}mium vit{\ae} {\ae}térn{\ae}. Et benedícat te omnípotens Deus, Pater, {\ding{64}} et Fílius, et Spíritus Sanctus.}
\par\noindent
\begin{center}
	Go in peace, the Lord hath put away all thy sins. 
	%
	And pray for me, a sinner.
\end{center}
\end{multicols}