\section{Reception of Converts}
\fancyhead[RO,LE]{\textit{Reception of Converts}}
\fancyhead[RE,LO]{}
%Rubric added to contextualise this ritual.
\begin{rubric}
    The Rite of Reception is to be used for catechumens who are to be received by Confirmation immediately before the Rite of Confirmation.
\end{rubric}
\begin{rubric}
    If there be many such catechumens, then each such catechumen must recite the Profession of Faith and individually receive absolution.
\end{rubric}
\begin{rubric}
    The Priest, kneeling before the middle of the altar, says alternately with the attendants:
\end{rubric}
\subby{Veni, Creator Spiritus}
\begin{multicols}{2}
\lett{C}{ome,} Holy Ghost, Creator blest,\par
\secondline{Vouchsafe within our souls to rest;}
\thirdline{Come with thy grace and heav'nly aid}
And fill the hearts which thou hast made.\\

To thee, the Comforter, we cry,

To thee, the Gift of God Most High,

The Fount of life, the Fire of love,

The soul's Anointing from above.\\

The sev'n-fold gifts of grace are thine,

O Finger of the Hand Divine;

True Promise of the Father thou,

Who dost the tongue with speech endow.\\

Thy light to every thought impart

And shed thy love in every heart;

Thine own unfailing might supply

To strengthen our infirmity.\\

Drive far away our ghostly foe,

And thine abiding peace bestow;

If thou be our preventing Guide,

No evil can our steps betide.\\

Make thou to us the Father known;

Teach us the eternal Son to own

Be this our never-changing creed,

%Translation changed to avoid heresy in English translation:
That from the Father dost proceed.\\

To thee who, dead, again dost live,

All glory, Lord, thy people give;

Whom with the Father we adore

And Holy Ghost for evermore. Amen.\\

%Versicle translation from the Alleluia of Whitsunday.
℣. O send forth thy Spirit and they shall be made.\par
℟. And thou shalt renew the face of the earth.
\end{multicols}

\letuspray
\lett{O}{God,} who as at this time didst teach the hearts of thy faithful people, by sending to them the light of thy Holy Spirit; Grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort. Through Christ our Lord. \textit{Amen.}%}{%
%\lett{V}{eni,} Creátor Spíritus,\par
%\secondline{Mentes tuórum vísita,}
%\thirdline{Imple supérna grátia,}
%Qu{\ae} tu creásti péctora.\\

%Qui díceris Paráclitus,

%Altíssimi donum Dei,

%Fons vivus, ignis, cáritas,

%Et spiritális únctio.\\

%Tu septifórmis múnere,

%Dígitus patérnæ déxteræ,

%Tu rite promíssum Patris,

%Sermóne ditans gúttura.\\

%Accénde lumen sénsibus:

%Infúnde amórem córdibus:

%Infírma nostri córporis

%Virtúte firmans pérpeti.\\

%Hostem repéllas lóngius,

%Pacémque dones prótinus:

%Ductóre sic te prǽvio

%Vitémus omne nóxium.\\

%Per te sciámus da Patrem,

%Noscámus atque Fílium,

%The `of both' refers here not to procession per se but relation. Since the Spirit is the Spirit of the Father and the Son, the Latin can stay.
%Teque utriúsque Spíritum

%Credámus omni témpore.\\

%Deo Patri sit glória,

%Et Fílio, qui a mórtuis

%Surréxit, ac Paráclito,

%In s{\ae}culórum sǽcula.

%Amen.\\

%℣. Emítte Spíritum tuum et creabúntur.\par
%℟. Et renovábis fáciem terr{\ae}.

%\oremuslatin
%\lett{D}{eus,} qui corda fidélium Sancti Spíritus illustratióne docuísti: da nobis in eódem Spíritu recta sápere, et de ejus semper consolatióne gaudére. Per Christum Dóminum nostrum. \textit{Amen.}}

\begin{rubric}
    The Priest then sits, and the new convert kneels before him: who, touching the Holy Gospel with his right hand, utters the profession of faith as stated below; or, if he does not know how to read, the Priest recites the same profession to him slowly, so that he can understand it when he turns and pronounce it with the priest in distinct words.
\end{rubric}
\subby{Profession of Faith}
\lett{I}{\textit{N.N.,}} having before my eyes the Holy Gospels, which I touch with my hand, and knowing that no one can be saved without that faith which the Holy, Catholic, and Apostolic 
%Change due to necessity:
Orthodox Church holds, believes, and teaches; against which I grieve that I have greatly erred, inasmuch as, I have held and believed doctrines opposed to her teaching.
\lett{I}{now} by the help of God’s grace, profess that I believe the Holy, Catholic Apostolic 
%Change due to necessity:
Orthodox Church to be the only and true Church established on earth by Jesus Christ, to which I submit myself with my whole heart. I firmly believe all the articles that she proposes to my belief; I reject and condemn all that she rejects and condemns, and I am ready to observe all that she commands me. And especially I profess that I believe:

One only God in three divine Persons, distinct from and equal to each other -- that is to say, the Father, the Son, and the Holy Ghost;

The Catholic doctrine of the Incarnation, Passion, Death, and Resurrection of our Lord Jesus Christ; and the personal union of the two Natures, the divine and the human: the Divine Maternity of the most holy Mary together with her most spotless virginity; and also her 
%Change due to necessity:
Assumption into heaven;

The true, real, and substantial presence of the Body of our Lord Jesus Christ, together with his Soul and Divinity, in the most holy sacrament of the Eucharist;

The seven Sacraments instituted by Jesus Christ, for the salvation of mankind: that is to say, Baptism, Confirmation, Holy Eucharist, Penance, 
%Change due to necessity:
Holy Unction, Holy Orders, Matrimony;

I also believe in 
%Change due to necessity:
the resurrection of the dead and Everlasting Life;

%Change due to both necessity and expediency (taken from inverting the Eastern Reception Repudiation):
That Our Lord Jesus Christ is the head of the Catholic Church; that no man can be the head of Christ's Body, that is to say, of the whole Church.

That the holy Apostles received from our Lord equal spiritual power, and every bishop is equally successor of the Apostles.

The veneration of the Saints, and of their images;

The authority of the apostolic and ecclesiastical Traditions, and of the Holy Scriptures, which we must interpret and understand only in the sense which our holy Mother the Catholic Church has held, and does hold, to whom alone it belongs to judge of their meaning and interpretation.

And everything else that has been defined and declared by the sacred Canons, by the 
%Change due to both necessity and expediency, according to the Service Book:
%The latter portion would be expedient to add, but it seems too controversial and unnecessary.
Seven Holy Ecumenical Councils, and the other traditions of the Holy Orthodox Catholic Apostolic Church.%, the Eighth Ecumenical Council in A.D. 879, the Holy Palamite Councils, and the Great Council of Jerusalem led by blessed Dositheus of Jerusalem.

With a sincere heart, therefore, and with unfeigned faith, I detest and abjure every error, heresy, and sect opposed to the said Catholic, Apostolic, and 
%Change due to necessity:
Orthodox Church. So help me God, and these his holy Gospels, which I touch with my hand.
\begin{rubric}
    The new convert then kneels---the Priest remaining sitting---and says either Psalm 51 as follows (or Psalm 130), ending in the \emph{Gloria Patri},
\end{rubric}
\subby{Psalm 51}
\lett{H}{ave} mercy upon me, O God, after thy great goodness : according to the multitude of thy mercies do away mine offences.\par
\secondlinepsalm{Wash me throughly from my wickedness : and cleanse me from my sin.}
\thirdlinepsalm{For I acknowledge my faults : and my sin is ever before me.}
\psanum{4}Against thee only have I sinned, and done this evil in thy sight : that thou mightest be justified in thy saying, and clear when thou art judged.\par
\psanum{5}Behold, I was shapen in wickedness : and in sin hath my mother conceived me.\par
\psanum{6}But lo, thou requirest truth in the inward parts : and shalt make me to understand wisdom secretly.\par
\psanum{7}Thou shalt purge me with hyssop, and I shall be clean : thou shalt wash me, and I shall be whiter than snow.\par
\psanum{8}Thou shalt make me hear of joy and gladness : that the bones which thou hast broken may rejoice.\par
\psanum{9}Turn thy face from my sins : and put out all my misdeeds.\par
\psanum{10}Make me a clean heart, O God : and renew a right spirit within me.\par
\psanum{11}Cast me not away from thy presence : and take not thy holy Spirit from me.\par
\psanum{12}O give me the comfort of thy help again : and stablish me with thy free Spirit.\par
\psanum{13}Then shall I teach thy ways unto the wicked : and sinners shall be converted unto thee.\par
\psanum{14}Deliver me from blood-guiltiness, O God, thou that art the God of my health : and my tongue shall sing of thy righteousness.\par
\psanum{15}Thou shalt open my lips, O Lord : and my mouth shall shew thy praise.\par
\psanum{16}For thou desirest no sacrifice, else would I give it thee : but thou delightest not in burnt-offerings.\par
\psanum{17}The sacrifice of God is a troubled spirit : a broken and contrite heart, O God, shalt thou not despise.\par
\psanum{18}O be favourable and gracious unto Sion : build thou the walls of Jerusalem.\par
\psanum{19}Then shalt thou be pleased with the sacrifice of righteousness, with the burnt-offerings and oblations : then shall they offer young bullocks upon thine altar.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.\par
\vspace{-3ex}
\subby{Preces}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subby{Our Father}
\begin{rubric}
    Our Father is said secretly until,
\end{rubric}
\fakecol{℣. And lead us not into temptation;}{℣. Et ne nos indúcas in tentatiónem:}
\fakecol{℟. But deliver us from evil.}{℟. Sed líbera nos a malo.}
%Translation from the Lenten Ember Friday Gradual:
\fakecol{℣. Save thy \textit{servant}, O Lord.}{℣. Salv\textit{um} fac \textit{servum tuum}, Dómine.}
\fakecol{℟. My God, that putteth \textit{his} trust in thee.}{℟. Deus meus, sperántem in te.}
\fakecol{℣. O Lord, hear my prayer.}{℣. Dómine, exáudi oratiónem meam.}
\fakecol{℟. And let my cry come unto thee.}{℟. Et clamor meus ad te véniat.}
\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
%\oremus
%\letuspray

\centerline{\textit{Let us pray.}}

%Translation from Litany of Saints and adapted:
\fakecol{\lett{O}{God,} whose nature and property is ever to have mercy and to forgive: receive our humble petitions, so that this thy \textit{servant}---being bound by the sentence of excommunication---may be loosed by the pitifulness of thy great mercy. \textit{Amen.}}{\lett{D}{eus,} cui próprium est miseréri semper et párcere: súscipe deprecatiónem nostram, ut \textit{hunc fámulum tuum quem} excommunicatiónis senténtia constríngit, miserátio tu{\ae} pietátis cleménter absólvat. Per Christum Dominum nostrum. \textit{Amen.}}

\vspace{-0.5ex}
\begin{rubric}
    A conditional baptism here follows, if it is to be conferred.
\end{rubric}
\subby{Absolution}
\begin{rubric}
    The Priest immediately sits down, turns to the convert, and---covering his head with the stole---absolves him of heresy in the following manner.
\end{rubric}
%Translation based on Confession
\fakecol{\lett{W}{ith} Apostolic Authority, I, by such authority, absolve thee from the bond of excommunication, which thou hadst incurred,\margrub{In grave or light doubt as to whether the penitent incurred excommunication by the profession of heresy, the priest here inserts the word `perchance'.} and I restore thee to the Sacred Sacraments of the Church, Communion, and the Unity of the Faithful, in the Name of the Father, and of the Son, {\ding{64}} and of the Holy Ghost. Amen.}{\lett{A}{uctorit\smash{á}te} Apostólica, qua fungor in hac parte, absólvo te a vínculo excommunicatiónis quam incurrísti*, et restítuo te sánctis Ecclési{\ae} sacraméntis, communióni et unitáti fidélium, in nómine Patris, et Fílii, {\ding{64}} et Spíritus Sancti. Amen.}

\begin{rubric}
    The Priest then enjoins some salutary penance upon the convert.
\end{rubric}
\subby{Te Deum laudamus}
\fakecol{\lett{W}{e} praise thee, O God; we acknowledge thee to be the Lord.\par
\secondline{All the earth doth worship thee, the Father everlasting.}
\thirdline{To thee all Angels cry aloud; the Heavens, and all the Powers therein;}
    To thee Cherubim and Seraphim continually do cry,\par
    Holy, Holy, Holy, Lord God of Sabaoth;\par
    Heaven and earth are full of the Majesty of thy glory.\par
    The glorious company of the Apostles praise thee.\par
    The goodly fellowship of the Prophets praise thee.\par
    The noble army of Martyrs praise thee.\par
    The holy Church throughout all the world doth acknowledge thee;\par
    The Father of an infinite Majesty;\par
    %Adorable reverted to honourable in line with Prayer Book Tradition and the Latin text.
    Thine honourable, true and only Son;\par
    Also the Holy Ghost the Comforter.\par
    \needspace{4\baselineskip}
\lett{T}{hou} art the King of Glory, O Christ.\par
\secondline{Thou art the everlasting Son of the Father.}
%`thou didst humble thyself to be born of a Virgin' reverted to `thou didst not abhor the Virgin’s womb' in line with the Prayer Book Tradition and Latin text.
\thirdline{When thou tookest upon thee to deliver man, thou didst not abhor the Virgin's womb.}
    When thou hadst overcome the sharpness of death, thou didst open the Kingdom of Heaven to all believers.\par
    Thou sittest at the right hand of God, in the glory of the Father.\par
    We believe that thou shalt come to be our Judge. \inrub{The following is said kneeling.}
    \par
    We therefore pray thee, help thy servants, whom thou hast redeemed with thy precious blood.\par
    Make them to be numbered with thy Saints, in glory everlasting.\par
    \needspace{4\baselineskip}
\lett{O}{Lord,} save thy people, and bless thine heritage.\par
\secondline{Govern them and lift them up for ever.}
\thirdline{Day by day we magnify thee;}
    And we worship thy Name ever, world without end.\par
    Vouchsafe, O Lord, to keep us this day without sin.\par
    O Lord, have mercy upon us, have mercy upon us.\par
    %Be reverted to lightened in line with Prayer Book Tradition.
    O Lord, let thy mercy lighten upon us, as our trust is in thee.\par
    O Lord, in thee have I trusted; let me never be confounded.}
{\lett{T}{e} Deum laudámus: * te Dóminum confitémur.\par
\secondline{Te {\ae}térnum Patrem * omnis terra venerátur.}
\thirdline{Tibi omnes Ángeli, * tibi C{\ae}li, et univérs{\ae} Potestátes}
Tibi Chérubim et Séraphim * incessábili voce proclámant:\par
Sanctus, Sanctus, Sanctus * Dóminus Deus Sábaoth.\par
Pleni sunt c{\ae}li et terra * majestátis glóri{\ae} tu{\ae}.\par
Te gloriósus * Apostolórum chorus,\par
Te Prophetárum * laudábilis númerus,\par
Te Mártyrum candidátus * laudat exércitus.\par
Te per orbem terrárum * sancta confitétur Ecclésia,\par
Patrem * imméns{\ae} majestátis;\par
Venerándum tuum verum * et únicum Fílium;\par
Sanctum quoque * Paráclitum Spíritum.

\lett{T}{u} Rex glóri{\ae}, * Christe.\par
\secondline{Tu Patris * sempitérnus es Fílius.}
\thirdline{Tu, ad liberándum susceptúrus hóminem: * non horruísti Vírginis úterum.}
Tu, devícto mortis acúleo, * aperuísti credéntibus regna c{\ae}lórum.\par
Tu ad déxteram Dei sedes, * in glória Patris.\par
Judex créderis * esse ventúrus. \inrub{Sequens versus dicitur flexis genibus.}
\par
Te ergo quǽsumus, tuis fámulis súbveni, * quos pretióso sánguine redemísti.\par
{\AE}térna fac cum Sanctis tuis * in glória numerári.

\lett{S}{alvum} fac pópulum tuum, Dómine, * et bénedic hereditáti tu{\ae}.\par
\secondline{Et rege eos, * et extólle illos usque in {\ae}térnum.}
\thirdline{Per síngulos dies * benedícimus te.}
Et laudámus nomen tuum in sǽculum, * et in sǽculum sǽculi.\par
Dignáre, Dómine, die isto * sine peccáto nos custodíre.\par
Miserére nostri, Dómine, * miserére nostri.\par
Fiat misericórdia tua, Dómine, super nos, * quemádmodum sperávimus in te.\par
In te, Dómine, sperávi: * non confúndar in {\ae}térnum.}

\subby{Conclusion}
\begin{rubric}
    The Priest then stands and says,
\end{rubric}
\fakecol{℣. Blessed art thou, O Lord, in the firmament of heaven.}{℣. Benedíctus es, Dómine, in firmaménto c{\ae}li.}
\fakecol{℟. And above all to be praised and glorified for ever.}{℟. Et laudábilis et gloriósus in s{\ae}cula.}
\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
%\oremus
\letuspray
%Adapted Collect of Trinity Sunday:
\fakecol{\lett{A}{lmighty} and everlasting God, who hast given unto us thy servants grace by the confession of a true faith to acknowledge the glory of the eternal Trinity, and in the power of the Divine Majesty to worship the Unity: We beseech thee; that thou wouldest keep us stedfast in this faith, and evermore defend us from all adversities. Who livest and reignest, one God, world without end. \textit{Amen.}}{\lett{O}{mn\smash{í}potens} sempitérne Deus, qui dedísti fámulis tuis in confessióne ver{\ae} fídei, {\ae}térn{\ae} Trinitátis glóriam agnóscere, et in poténtia majestátis adoráre unitátem: qu{\ae}sumus; ut, ejúsdem fídei firmitáte, ab ómnibus semper muniámur advérsis. Per Christum Dóminum nostrum. \textit{Amen.}}
\fakecol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\fakecol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\fakecol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\fakecol{℟. Thanks be to God.}{℟. Deo grátias.}
\par\noindent
\fakecol{\par\noindent
The blessing of Almighty God, the Father, {\ding{64}} the Son, and the Holy Ghost, descend upon you and remain with you always. \textit{Amen.}}{Benedíctio Dei omnipoténtis, Patris {\ding{64}} et Fílii, et Spíritus Sancti, descéndat super vos, et máneat semper. \textit{Amen.}}