\section{Blessing of Baptismal Waters}
\fancyhead[RO,LE]{\textit{Baptismal Blessing}}
\fancyhead[RE,LO]{}
\begin{rubric}
    The baptismal container being washed and cleaned, it is filled with clean water. The Priest and his clerics (or other Priests) proceed---with a Cross, two candles, and a thurible with incense---and with salt and the phials of Sacred Chrism \& the Oil of Catechumens placed nearby, they approach the Font where he says the Litany of Saints (p. \pageref{LitanySaints}) after the Seven Penitential Psalms (Psalms 6, 32, 38, 51, 102, 130, \& 143).\par
    \textsc{Note,} The Priest may instead say the shorter Litany from Holy Saturday in the Missal.
\end{rubric}
\begin{rubric}
%FIND TRANSLATION:
    Before the \emph{That thou wouldest vouchsafe} (\emph{Ut nos exáudire dignéris,}) is said and then repeated: \emph{That thou wouldest vouchsafe to {\ding{64}} bless and {\ding{64}} consecrate this Font for the regeneration unto thee of new offspring. ℟. We beseech thee to hear us.} (\emph{Ut Fontem istum ad regenerándam tibi novam prolem bene {\ding{64}} dícere, et conse {\ding{64}} cráre dignéris. ℟. Te rogámus, audi nos.})
\end{rubric}
\begin{rubric}
    After the final \emph{Kyrie, eléison}, the Priest says the Lord's Prayer (as noted) and the Apostles' Creed, all in a clear voice. He then proceeds with the following.
\end{rubric}
\elcol{℣. With thee, O Lord, is the well of life.}{℣. Apud te, Dómine, est fons vit{\ae}.}

\elcol{℟. And in thy light shall we see light.}{℟. Et in lúmine tuo vidébimus lumen.}

\elcol{℣. O Lord, hear my prayer.}{℣. Dómine, exáudi oratiónem meam.}

\elcol{℟. And let my cry come unto thee.}{℟. Et clamor meus ad te véniat.}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}

\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\oremus

\elcol{\lett{A}{lmighty} and everlasting God, be present at the mysteries, be present at the sacraments of thy great goodness: and send forth the spirit of adoption for the regeneration of the new peoples whom the Font of Baptism doth bring forth unto thee: that what is to be done by our humble ministry, may be effectually fulfilled by thy power. Through.\par
℟. Amen.}{\lett{O}{mn\smash{í}potens} sempitérne Deus, adésto magn{\ae} pietátis tu{\ae} mystériis, adésto sacraméntis: et ad recreándos novos pópulos, quos tibi fons Baptísmatis párturit, spíritum adoptiónis emítte; ut, quod nostr{\ae} humilitátis geréndum est ministério, virtútis tu{\ae} impleátur efféctu. Per Dóminum nostrum.%Jesum Christum, Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti Deus, per ómnia sǽcula saculórum.\par

℟. Amen.}

\subby{Exorcism of Water}
\elcol{\lett{I}{exorcise} thee, O creature of water, by the living {\ding{64}} God, by the true {\ding{64}} God, by the holy {\ding{64}} God, by God, who in the beginning through his Word divided thee form the dry land: whose Spirit moved upon thee, who commanded thee to flow from paradise.}{\lett{E}{xorc\smash{í}zo} te, creatúra aqu{\ae}, per Deum {\ding{64}} vivum, per Deum {\ding{64}} verum, per Deum {\ding{64}} sanctum, per Deum, qui te in princípio verbo separávit ab árida: cujus Spíritus super te ferebátur, qui te de paradíso manáre jussit.}
\begin{rubric}
The Priest divides the waters with his hand and then scatters it beyond the edge of the Font, towards the four quarters of the world, proceeding with the following.
\end{rubric}
\elcoldent{And commanded thee to water the whole earth with thy four rivers, Who in the desert by wood bestowed upon thee sweetness when thou wast bitter, that men might drink. Who brought thee forth from the rock, that he might refresh the people, wearied with thirst, whom he had delivered out of Egypt. I exorcise thee through Jesus Christ, his only Son, our Lord: who in Cana of Galilee by a wondrous miracle did change thee through his power into wine: who walked upon thee with his feet, and was baptized in thee by John in Jordan. Who brought thee forth together with blood from his wide: and commanded his disciples, that they should baptize in thee them that believe, saying: Go ye, teach all nations, baptizing them in the Name of the Father, and of the Son, and of the Holy Ghost, that thou mayest be made holy water, blessed water, water that washeth away stains and cleanseth sins. I command thee therefore, every unclean spirit, every phantom, every lie, be rooted out, and flee away from this creature of water; that to them who shall be baptized therein, it may become a Font of water springing up unto life eternal, regenerating them unto God the Father, and the Son, and the Holy Ghost, in the Name of the same our Lord Jesus Christ, who shall come to judge the living and the dead, and the world by fire. \textit{Amen.}

\letuspray
\lett{O}{Lord} holy, Father almighty, everlasting God, who dost sanctify spiritual waters, we humbly entreat thee: that thou wouldest vouchsafe to look upon this ministry of our humble service, and to send forth upon these waters, made ready for the washing and purifying of men, the Angel of holiness; to the end that, the sins of their former life being washed, and their guilt cleansed, they being regenerate may be worthy to be made a spotless habitation for the Holy Spirit. Through.\par
℟. Amen.}{Et in quátuor flumínibus totam terram rigáre pr{\ae}cépit: qui te in desérto amáram per lignum, dulcem fecit atque potábilem; qui te de petra prodúxit, ut pópulum, quem ex {\AE}gýpto liberáverat, siti fatigátum recreáret. Exorcízo te per Jesum Christum, Fílium ejus únicum, Dóminum nostrum: qui te in Cana Galilǽ{\ae} signo admirábili sua poténtia convértit in vinum: qui super te pédibus ambulávit, et a Joánne in Jordáne in te baptizátus est. Qui te una cum sánguine de látere suo prodúxit: et discípulis suis jussit, ut credéntes baptizárent in te, dicens: Ite, docéte omnes gentes, baptizántes eos in nómine Patris, et Fílii, et Spíritus Sancti; ut efficiáris aqua sancta, aqua benedícta, aqua, qu{\ae} lavat sordes, et mundat peccáta. Tibi ígitur pr{\ae}cípio, omnis spíritus immúnde, omne phantásma, omne mendácium, eradicáre, et effugáre ab hac creatúra aqu{\ae}, ut qui in ipsa baptizándi erunt, fiat eis fons aqu{\ae} saliéntis in vitam {\ae}térnam, regénerans eos Deo Patri, et Fílio, et Spirítui Sancto, in nómine ejúsdem Dómini nostri Jesu Christi, qui ventúrus est judicáre vivos et mórtuos, et sǽculum per ignem. \textit{Amen.}

\oremuslatin

\lett{D}{\smash{ó}mine} sancte, Pater omnípotens, {\ae}térne Deus, aquárum spirituálium sanctificátor, te supplíciter deprecámur: ut ad hoc ministérium humilitátis nostr{\ae} respícere dignéris, et super has aquas, abluéndis et purificándis homínibus pr{\ae}parátas, Angelum sanctitátis emíttas, quo, peccátis vit{\ae} prióris ablútis, reatúque detérso, purum Sancto Spirítui habitáculum regeneráti éffici mereántur. Per Dóminum nostrum.\par
℟. Amen.}


\begin{rubric}
    The Priest breathes thrice into the water in the form of a \textsc{Ψ}. Then he places incense in the thurible, blesses it, and censes the Font with three simple swings. He then pours some of the Oil of Catechumens into the water in the form of Cross saying in a clear voice,
\end{rubric}
\elcol{\lett{M}{ay} this Font be sanctified and made fruitful by the Oil of salvation, for such as shall be born again therefrom unto life eternal, in the Name of the {\ding{64}} Father, and of the {\ding{64}} Son, and of the Holy {\ding{64}} Ghost. \textit{Amen.}}{\lett{S}{anctific\smash{é}tur,} et fecundétur fons iste óleo salútis renascéntibus ex eo in vitam {\ae}térnam, in nómine Pa {\ding{64}} tris, et Fí {\ding{64}} lii, et Spíritus {\ding{64}} Sancti. \textit{Amen.}}

\begin{adjustwidth}{0in}{-1em}
\begin{rubric}
    The Priest then pours in some of the Sacred Chrism in the same mode, saying the following.
\end{rubric}
\end{adjustwidth}

\elcol{\lett{M}{ay} this inpouring of the Chrism of our Lord Jesus Christ, and of the Holy Spirit the Paraclete, be wrought in the Name of the holy Trinity. \textit{Amen.}}{\lett{I}{nf\smash{ú}sio} Chrísmatis Dómini nostri Jesu Christi, et Spíritus Sancti Parácliti, fiat in nómine sanct{\ae} Trinitátis. \textit{Amen.}}

\begin{rubric}
    The Priest then receives both vessels for the Sacred Chrism and the Oil of Catechumens and pours them both in the form of a Cross, saying the following.
\end{rubric}
\elcol{\lett{M}{ay} this commingling of the Chrism of sanctification, and of the Oil of Unction, and of the water of Baptism be likewise wrought in the Name of the {\ding{64}} Father, and of the {\ding{64}} Son, and of the Holy {\ding{64}} Ghost. \textit{Amen.}}{\lett{C}{ommm\smash{í}xtio} Chrísmatis sanctificatiónis, et Olei unctiónis, et aqu{\ae} Baptísmatis páriter fiat in nómine Pa {\ding{64}} tris, et Fí {\ding{64}} lii, et Spíritus {\ding{64}} Sancti. \textit{Amen.}}

\begin{rubric}
Then setting down the phials, with his right hand he mingles the holy Oil and the infused Chrism with the Water, and scatters it over the whole Font. Then he wipes his hand with crumbs of bread. And if there are those to be baptised, he baptises them, changing his vestments from violet to white. But if he baptises no{\"o}ne, then he washes his hands and pours the ablution into the sacrarium.
\end{rubric}