\fancyhead[RO,LE]{\textit{Penitential Office}}\label{penitential}
\fancyhead[RE,LO]{}
\section{A Penitential Office}
%From the 1928 Book of Common Prayer
\begin{rubric}
%Language of rubric from the 1928 American is adjusted for readability
	On the First Day of Lent, the Penitential Office is read immediately after the Great Litany during Morning Prayer. It may also be used during Evening Prayer or as a separate Office.
\end{rubric}
\begin{rubric}
    The Penitential Office may be read at other times, at the discretion of the Minister.
\end{rubric}
\begin{rubric}
    The Minister and the People kneeling, then shall be said by them this Psalm following.
\end{rubric}
\subbypsalm{Psalm 51. \textit{Miserere mei, Deus}}
\lett{H}{ave} mercy upon me, O God, after thy great goodness : according to the multitude of thy mercies do away mine offences.\par
\secondlinepsalm{Wash me throughly from my wickedness : and cleanse me from my sin.}
\thirdlinepsalm{For I acknowledge my faults : and my sin is ever before me.}
\psanum{4}Against thee only have I sinned, and done this evil in thy sight : that thou mightest be justified in thy saying, and clear when thou art judged.\par
\psanum{5}Behold, I was shapen in wickedness : and in sin hath my mother conceived me.\par
\psanum{6}But lo, thou requirest truth in the inward parts : and shalt make me to understand wisdom secretly.\par
\psanum{7}Thou shalt purge me with hyssop, and I shall be clean : thou shalt wash me, and I shall be whiter than snow.\par
\psanum{8}Thou shalt make me hear of joy and gladness : that the bones which thou hast broken may rejoice.\par
\psanum{9}Turn thy face from my sins : and put out all my misdeeds.\par
\psanum{10}Make me a clean heart, O God : and renew a right spirit within me.\par
\psanum{11}Cast me not away from thy presence : and take not thy holy Spirit from me.\par
\psanum{12}O give me the comfort of thy help again : and stablish me with thy free Spirit.\par
\psanum{13}Then shall I teach thy ways unto the wicked : and sinners shall be converted unto thee.\par
\psanum{14}Deliver me from blood-guiltiness, O God, thou that art the God of my health : and my tongue shall sing of thy righteousness.\par
\psanum{15}Thou shalt open my lips, O Lord : and my mouth shall shew thy praise.\par
\psanum{16}For thou desirest no sacrifice, else would I give it thee : but thou delightest not in burnt-offerings.\par
\psanum{17}The sacrifice of God is a troubled spirit : a broken and contrite heart, O God, shalt thou not despise.\par
\psanum{18}O be favourable and gracious unto Sion : build thou the walls of Jerusalem.\par
\psanum{19}Then shalt thou be pleased with the sacrifice of righteousness, with the burnt-offerings and oblations : then shall they offer young bullocks upon thine altar.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
\begin{rubric}
    If the Litany hath been already said, the Minister may pass at once to \emph{O Lord, save thy servants;} etc.
\end{rubric}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par
    ℣. Lord, have mercy upon us.

    ℟. Christ, have mercy upon us.
    
    ℣. Lord, have mercy upon us.
    
\lett{O}{ur} Father who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}
\par
    ℣. O Lord, save thy servants;

    ℟. That put their trust in thee.

    ℣. Send unto them help from above.

    ℟. And evermore mightily defend them.

    ℣. Help us, O God our Saviour.

    ℟. And for the glory of thy Name deliver us; be merciful to us sinners, for thy Name's sake.
    
    ℣. O Lord, hear our prayer.
    
    ℟. And let our cry come unto thee.

\letuspray
\lett{O}{Lord,} we beseech thee, mercifully hear our prayers, and spare all those who confess their sins unto thee; that they, whose consciences by sin are accused, by thy merciful pardon may be absolved; through Christ our Lord. \textit{Amen.}
\needspace{4\baselineskip}
\lett{O}{most} mighty God, and merciful Father, who hast compassion upon all men, and who wouldest not the death of a sinner, but rather that he should turn from his sin, and be saved; Mercifully forgive us our trespasses; receive and comfort us, who are grieved and wearied with the burden of our sins. Thy property is always to have mercy; to thee only it appertaineth to forgive sins. Spare us therefore, good Lord, spare thy people, whom thou hast redeemed; enter not into judgment with thy servants; but so turn thine anger from us, who meekly acknowledge our transgressions, and truly repent us of our faults, and so make haste to help us in this world, that we may ever live with thee in the world to come; through Jesus Christ our Lord. \textit{Amen.}

%Rubric language slightly adjusted for clarity.
\begin{rubric}
    Then shall the People say this that followeth, with the Minister.
\end{rubric}

\lett{T}{urn} thou us, O good Lord, and so shall we be turned. Be favourable, O Lord, Be favourable to thy people, Who turn to thee in weeping, fasting, and praying. For thou art a merciful God, Full of compassion, Long-suffering, and of great pity. Thou sparest when we deserve punishment, And in thy wrath thinkest upon mercy. Spare thy people, good Lord, spare them, And let not thine heritage be brought to confusion. Hear us, O Lord, for thy mercy is great, And after the multitude of thy mercies look upon us; Through the merits and mediation of thy blessed Son, Jesus Christ our Lord. Amen.

\begin{rubric}
    Then the Minister shall say,
\end{rubric}

\lett{O}{God,} whose nature and property is ever to have mercy and to forgive; Receive our humble petitions; and though we be tied and bound with the chain of our sins, yet let the pitifulness of thy great mercy loose us; for the honour of Jesus Christ, our Mediator and Advocate. \textit{Amen.}

\lett{T}{he} Lord bless us, and keep us. The Lord make his face to shine upon us, and be gracious unto us. The Lord lift up his countenance upon us, and give us peace, both now and evermore. \textit{Amen.}
%\begin{center}
%    \textsc{Here endeth the Penitential Office.}
%\end{center}