\fancyhead[C]{{\LARGE Kalendar of the Year}}\label{kalendar}
\phantomsection
\addcontentsline{toc}{section}{Kalendar of the Year}
\begin{longtable}{p{2mm}|p{94mm}}
\multicolumn{2}{l}{\textsc{January}}\\
1&{\color{RubricRed}Circumcision of Our Lord}, \textit{\nth{2} Double}\\
&(Octave Day of the Nativity)\\
2&{\color{RubricRed}Most Holy Name of Jesus}, \textit{\nth{2} Double}\\
&Octave Day of St. Stephen, Protomartyr, \textit{Simple}\\
3&Octave Day of St. John, Apostle \& Evangelist, \textit{Simple}\\
4&Octave Day of the Holy Innocents, \textit{Simple}\\
&St. Titus, Bishop \& Confessor, \textit{Memorial}\\
5&The Vigil of the Epiphany, \textit{\nth{2} Vigil}\\
&St. Telesphorus of Rome, Bishop \& Martyr, \textit{Memorial}\\
6&{\color{RubricRed}The Epiphany of Our Lord}, \textit{\nth{1} Double}, \nth{2} Octave\\
7&Day II within the Octave of the Epiphany, \textit{Semidouble}\\
8&Day III within the Octave of the Epiphany, \textit{Semidouble}\\
9&Day IV within the Octave of the Epiphany, \textit{Semidouble}\\
10&Day V within the Octave of the Epiphany, \textit{Semidouble}\\
&St. Paul the First Hermit, \textit{Memorial}\\
11&Day VI within the Octave of the Epiphany, \textit{Semidouble}\\
&St. Hyginus of Rome, Bishop \& Martyr, \textit{Memorial}\\
12&Day VII within the Octave of the Epiphany, \textit{Semidouble}\\
&St. Benedict, Bishop \& Abbot, \textit{Memorial}\\
%Emphasis is here laid on the Baptism to be consistent with other Octave Days
13&{\color{RubricRed}Baptism of Our Lord}, \textit{Greater Double}\\
&(Octave Day of the Epiphany)\\
14&St. Hilary, Bishop, Confessor \& Doctor, \textit{Double}\\
&St. Felix of Nola, Priest \& Martyr, \textit{Memorial}\\
15&St. Maurus, Abbot, \textit{Greater Double}\\
16&St. Marcellus I of Rome, Bishop \& Martyr, \textit{Memorial}\\
17&St. Anthony, Abbot, \textit{Double}\\
18&Chair of St. Peter in Rome, \textit{Greater Double}\\
&Commemoration of St. Paul, \textit{Commemoration}\\
&St. Prisca of Rome, Virgin Martyr, \textit{Memorial}\\
19&Ss. Marius, Martha, Audifax, \& Abachum, Martyrs, \textit{Memorial}\\
&St. Mark of Ephesus, Bishop \& Confessor, \textit{Memorial}\\
20&Ss. Fabian (Bishop) \& Sebastian, Martyrs, \textit{Double}\\
21&St. Agnes of Rome, Virgin \& Martyr, \textit{Greater Double}\\
22&Ss. Vincent, Martyr, \textit{Memorial}\\
23&St. Emerentiana, Virgin \& Martyr, \textit{Memorial}\\
24&St. Timothy, Bishop \& Martyr, \textit{Double}\\
25&{\color{RubricRed}The Conversion Of St. Paul the Apostle}, \textit{Greater Double}\\
&{Commemoration of St. Peter}, \textit{Commemoration}\\
26&St. Polycarp of Smyrna, Bishop \& Martyr, \textit{Memorial}\\
27&St. John Chrysostom, Bishop, Confessor \& Doctor, \textit{Double}\\
28&	St. Cyril of Alexandria, Bishop, Confessor, \& Doctor, \textit{Double}\\
&The Second Feast of St. Agnes, Virgin \& Martyr, \textit{Memorial}\\
29&\\
30&St. Martina, Virgin \& Martyr, \textit{Memorial}\\
31&(Earliest Date of Septuagesima)\\
\multicolumn{2}{l}{\textsc{February}}\\
1&St. Ignatius of Antioch, Bishop \& Martyr, \textit{Double}\\
&St. Bridget of Ireland, Virgin, \textit{Memorial}\\
2&{\color{RubricRed}Purification of the B.V.M.}, \textit{\nth{2} Double}\\
3&St. Blaise, Bishop \& Martyr, \textit{Memorial}\\
4&St. Joseph of Aleppo, Martyr, \textit{Memorial}\\
&The New Martyrs of Russia, \textit{Memorial}\\
5&St. Agatha, Virgin \& Martyr, \textit{Greater Double}\\
6&St. Dorothea, Virgin \& Martyr, \textit{Memorial}\\
7&St. Romuald, Abbot, \textit{Double}\\
8&\\
9&St. Apollonia of Alexandria, Virgin \& Martyr, \textit{Memorial}\\
%Some make this a second double:
10&St. Scholastica, Virgin, \textit{Greater Double}\\
11&St. Gregory II, Pope \& Confessor, \textit{Memorial}\\
12&\\
13&\\
14&St. Valentine, Priest \& Martyr, \textit{Memorial}\\
15&Ss. Faustinus \& Jovita, Martyrs, \textit{Memorial}\\
16&\\
17&(Earliest Date of Lent)\\
18&St. Simeon of Jerusalem, Bishop \& Martyr, \textit{Memorial}\\
19&\\
20&\\
21&\\
22&{\color{RubricRed}Chair of St. Peter at Antioch}, \textit{\nth{2} Double}\\
&{Commemoration of St. Paul}, \textit{Commemoration}\\
23&Vigil of St. Matthias, \textit{Vigil}\\
24&{\color{RubricRed}St. Matthias, Apostle}, \textit{\nth{2} Double}\\
25&St. Walburga of Heidenheim, Virgin, \textit{Memorial}\\
26&St. Alexander of Alexandria, Bishop \& Confessor, \textit{Memorial}\\
27&St. Raphael of Brooklyn, Bishop \& Confessor, \textit{Greater Double}\\
28&\\
\multicolumn{2}{l}{\textsc{March}}\\
1&St. David of Wales, Bishop \& Confessor, \textit{Memorial}\\
2&St. Chad, Bishop \& Confessor, \textit{Memorial}\\
3&\\
4&St. Lucius I, Pope \& Martyr, \textit{Memorial}\\
5&\\
6&St. Perpetua \& Felicitas, Martyrs, \textit{Double}\\
&(Latest Date of Septuagesima)\\
7&\\
8&\\
9&St. Gregory of Nyssa, Bishop, Confessor, \& Doctor, \textit{Memorial}\\
10&The Forty Holy Martyrs, \textit{Memorial}\\
11&\\
12&{\color{RubricRed}St. Gregory, Pope, Confessor, \& Martyr}, \textit{\nth{2} Double}\\
13&\\
14&\\
15&\\
16&\\
17&St. Patrick, Bishop \& Confessor, \textit{Double}\\
&St. Joseph of Arimathea, Confessor, \textit{Memorial}\\
18&St. Cyril of Jerusalem, Bishop, Confessor, \& Doctor, \textit{Double}\\
&St. Edward, King \& Martyr, \textit{Memorial}\\
19&{\color{RubricRed}St. Joseph, Spouse of the B.V.M., Confessor}, \textit{\nth{1} Double}\\
20&St. Cuthbert, Bishop \& Confessor, \textit{Double}\\
21&{\color{RubricRed}St. Benedict, Abbot}, \textit{\nth{2} Double}\\
22&\\
23&(Latest Date of Lent)\\
24&St. Gabriel the Archangel, \textit{Greater Double}\\
25&{\color{RubricRed}The Annunciation of the B.V.M.}, \textit{\nth{1} Double}\\
26&\\
27&St. John Damascene, Confessor \& Doctor, \textit{Double}\\
28&\\
29&\\
30&St. John Climacus, Abbot, \textit{Memorial}\\
31&St. Innocent of Alaska, Bishop \& Confessor, \textit{Memorial}\\
\multicolumn{2}{l}{\textsc{April}}\\
1&\\
2&\\
3&\\
4&St. Isidore of Seville, Bishop, Confessor, \& Doctor, \textit{Double}\\
&(Earliest Date of Easter)\\
5&\\
6&\\
7&{\color{RubricRed}St. Tikhon, Bishop, \& Confessor}, \textit{\nth{1} Double}, Common Octave\\
8&\\
9&\\
10&\\
11&St. Leo the Great, Pope, Confessor, \& Doctor, \textit{Double}\\
12&\\
13&St. Hermengild, Martyr, \textit{Memorial}\\
14&St. Justin, Martyr, \textit{Memorial}\\
&St. Tiburtius, Valerian, \& Maximus, Martyrs, \textit{Memorial}\\
15&\\
16&\\
17&St. Anicetus, Pope \& Martyr, \textit{Memorial}\\
18&\\
19&\\
20&\\
21&\\
22&Sts. Soter \& Caius, Popes \& Martyrs, \textit{Memorial}\\
23&{\color{RubricRed}St. George, Martyr}, \textit{\nth{1} Double}, Common Octave\\
24&Day II within the Octave of St. George, \textit{Semidouble}\\
25&{\color{RubricRed}St. Mark, Evangelist}, \textit{\nth{2} Double}\\
&Day III within the Octave of St. George, \textit{Semidouble}\\
26&Day IV within the Octave of St. George, \textit{Semidouble}\\
&Sts. Cletus \& Marcellinus, Martyrs, \textit{Memorial}\\
27&Day V within the Octave of St. George, \textit{Semidouble}\\
28&Day VI within the Octave of St. George, \textit{Semidouble}\\
&St. Vitalis, Martyr, \textit{Memorial}\\
29&Day VII within the Octave of St. George, \textit{Semidouble}\\
30&Octave Day of St. George, \textit{Greater Double}\\
\multicolumn{2}{l}{\textsc{May}}\\
1&{\color{RubricRed}Sts. Philip \& James, Apostles}, \textit{\nth{2} Double}\\
2&St. Athanasius, Bishop, Confessor, \& Doctor, \textit{Double}\\
3&{\color{RubricRed}Invention of the Holy Cross}, \textit{\nth{2} Double}\\
&Sts. Alexander (Pope), Eventius, Theodulus (Martyrs), \& Juvenalis (Bishop \& Confessor), \textit{Memorial}\\
4&St. Monica, Widow, \textit{Memorial}\\
5&\\
6&{\color{RubricRed}St. John before the Latin Gate}, \textit{Greater Double}\\
7&St. Alexis Toth, Priest \& Confessor, \textit{Memorial}\\
8&The Apparition of St. Michael the Archangel, \textit{Greater Double}\\
&St. Boniface IV, Bishop \& Confessor, \textit{Memorial}\\
&(Latest Date of Easter)\\
9&St. Gregory Nazianzen, Bishop, Confessor, \& Doctor, \textit{Double}\\
10&Sts. Gordian \& Epimachus, Martyrs, \textit{Memorial}\\
11&\\
12&Sts. Nereus, Achilles, Domitilla, \& Pancras, Martyrs, \textit{Memorial}\\
13&\\
14&St. Boniface of Tarsus, Martyr, \textit{Memorial}\\
15&\\
16&\\
17&\\
18&St. Venantius, Martyr, \textit{Double}\\
19&St. Pudentiana, Virgin, \textit{Memorial}\\
20&\\
21&\\
22&St. Romanus, Abbot, \textit{Memorial}\\
23&\\
24&St. Vincent of Lerins, Confessor, \textit{Memorial}\\
25&Pope St. Urban, Bishop \& Martyr, \textit{Memorial}\\
26&St. Augustine of Canterbury, Bishop \& Confessor, \textit{Double}\\
&St. Eleutherius, Pope \& Martyr, \textit{Memorial}\\
27&St. Bede the Venerable, Confessor \& Doctor, \textit{Double}\\
&St. John I, Pope \& Martyr, \textit{Memorial}\\
28&\\
29&\\
30&St. Felix I, Pope \& Martyr, \textit{Memorial}\\
31&St. Petronilla, Virgin, \textit{Memorial}\\
\multicolumn{2}{l}{\textsc{June}}\\
1&\\
2&St. Marcellinus, Peter, \& Erasmus, Martyrs, \textit{Memorial}\\
3&\\
4&\\
5&St. Boniface, Bishop \& Martyr, \textit{Double}\\
6&\\
7&\\
8&\\
9&St. Columba of Iona, Abbot, \textit{Double}\\
&Sts. Primus \& Felician, Martyrs, \textit{Memorial}\\
10&St. Margaret of Scotland, Queen \& Widow, \textit{Memorial}\\
11&{\color{RubricRed}St. Barnabas, Apostle}, \textit{Greater Double}\\
12&Sts. Basilides, Cyrinus, Nabor, \& Nazarius, Martyrs, \textit{Memorial}\\
13&\\
14&St. Basil the Great, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
15&Sts. Vitus, Modestus, \& Crescentia, Martyrs, \textit{Memorial}\\
16&\\
17&\\
18&St. Ephrem the Syrian, Deacon, Confessor, \& Doctor, \textit{Double}\\
&Sts. Marcus \& Marcellianus, Martyrs, \textit{Memorial}\\
19&Sts. Gervase \& Protase, Martyrs, \textit{Memorial}\\
20&St. Silverius, Pope \& Martyr, \textit{Memorial}\\
21&\\
22&St. Alban, Protomartyr of England, \textit{Double}\\
&St. Paulinus, Bishop \& Confessor, \textit{Memorial}\\
23&St. Etheldreda, Queen \& Virgin, \textit{Memorial}\\
&Vigil of St. John Baptist, \textit{Vigil}\\
24&{\color{RubricRed}Nativity of St. John Baptist}, \textit{\nth{1} Double}, Common Octave\\
25&Day II within the Octave of St. John Baptist, \textit{Semidouble}\\
26&Sts. John \& Paul, Martyrs, \textit{Double}\\
& Day III within the Octave of St. John Baptist, \textit{Semidouble}\\
27& Day IV within the Octave of St. John Baptist, \textit{Semidouble}\\
28& Day V within the Octave of St. John Baptist, \textit{Semidouble}\\
&Vigil of St. Peter, \textit{Vigil}\\
&St. Leo II, Pope \& Confessor, \textit{Memorial}\\
29&{\color{RubricRed}St. Peter, Apostle}, \textit{\nth{1} Double}, Common Octave\\
&{Day VI within the Octave of St. John Baptist}, \textit{Semidouble}\\
30&{\color{RubricRed}Commemoration of St. Paul, Apostle}, \textit{Greater Double}\\
&{Commemoration of St. Peter, Apostle}, \textit{Commemoration}\\
&{Day VII within the Octave of St. John Baptist}, \textit{Semidouble}\\
\multicolumn{2}{l}{\textsc{July}}\\
1&{\color{RubricRed}Most Precious Blood of Our Lord Jesus Christ}, \textit{\nth{2} Double}\\
&{Octave Day of St. John Baptist}, \textit{Greater Double}\\
&Day III within the Octave of St. Peter, \textit{Semidouble}\\
2&{\color{RubricRed}Visitation of the B.V.M.}, \textit{\nth{2} Double}\\
&St. John of San Francisco, Bishop \& Confessor, \textit{Memorial}\\
&Sts. Processus \& Martinian, Martyrs, \textit{Memorial}\\
&Day IV within the Octave of St. Peter, \textit{Semidouble}\\
3&St. Iren{\ae}us of Lyon, Bishop \& Martyr, \textit{Double}\\
&Day V within the Octave of St. Peter, \textit{Semidouble}\\
4&Day VI within the Octave of St. Peter, \textit{Semidouble}\\
5&Day VII within the Octave of St. Peter, \textit{Semidouble}\\
6&The Octave Day of St. Peter, \textit{Greater Double}\\
7&Sts. Cyril \& Methodius, Bishops \& Confessors, \textit{Double}\\
8&\\
9&\\
10&Seven Holy Brothers, with Sts. Rufina \& Secunda, Martyrs, \textit{Memorial}\\
&St. Joseph of Damascus (Priest) \& Companions, Martyrs, \textit{Memorial}\\
11&St. Pius I, Pope \& Martyr, \textit{Memorial}\\
&[Solemnity of St. Benedict (Monastics \& Oblates only)], \textit{Greater Double}\\
12&Sts. Nabor \& Felix, Martyrs, \textit{Memorial}\\
13&St. Anacletus, Pope \& Martyr, \textit{Memorial}\\
14&\\
15&St. Vladimir of Kiev, King \& Confessor, \textit{Double}\\
%Deviation from Ordo:
&Translation of St. Swithun, \textit{Memorial}\\
16&Our Lady of Einsiedeln, \textit{Greater Double}\\
17&St. Alexius, Confessor, \textit{Memorial}\\
18&Sts. Symphorosa \& Her Seven Sons, Martyrs, \textit{Memorial}\\
&St. Sergius of Radonezh, Abbot, \textit{Memorial}\\
19&St. Seraphim of Sarov, Priest \& Confessor, \textit{Memorial}\\
20&St. Elias the Prophet, Confessor, \textit{Double}\\
&St. Margaret of Antioch, Virgin \& Martyr, \textit{Memorial}\\
21&St. Praxedes, Virgin, \textit{Memorial}\\
22&{\color{RubricRed}St. Mary Magdalene, Penitent}, \textit{Greater Double}\\
23&St. John Cassian, Abbot, \textit{Memorial}\\
&St. Apollinaris of Ravenna, Bishop \& Martyr, \textit{Memorial}\\
&St. Liborius, Bishop \& Confessor, \textit{Memorial}\\
24&Vigil of St. James, \textit{Vigil}\\
&St. Christina, Virgin \& Martyr, \textit{Memorial}\\
25&{\color{RubricRed}St. James, Apostle}, \textit{\nth{2} Double}\\
&St. Christopher, Martyr, \textit{Memorial}\\
26&{\color{RubricRed}St. Anne, Mother of the B.V.M.}, \textit{\nth{2} Double}\\
27&St. Pantaleon, Martyr, \textit{Memorial}\\
28&Sts. Nazarius, Celsus, Martyrs; St. Victor, Pope \& Martyr; \& Innocent, Bishop \& Confessor, \textit{Memorial}\\
29&St. Martha of Bethany, Virgin, \textit{Double}\\
&Sts. Felix II (Bishop), Simplicius, Faustinus, \& Beatrice, Martyrs, \textit{Memorial}\\
30&Sts. Abdon \& Sennen, Martyrs, \textit{Memorial}\\
31&\\
\multicolumn{2}{l}{\textsc{August}}\\
1&{\color{RubricRed}Chains of St. Peter, Apostle}, \textit{Greater Double}\\
&{Commemoration of St. Paul, Apostle}, \textit{Commemoration}\\
&Holy Maccabees, Martyrs, \textit{Memorial}\\
2&St. Stephen, Pope \& Martyr, \textit{Memorial}\\
3&The Invention of St. Stephen, Protomartyr, \textit{Double}\\
4&\\
5&Dedication of Our Lady of the Snows, \textit{Greater Double}\\
&St. Oswald, King \& Martyr, \textit{Memorial}\\
6&{\color{RubricRed}Transfiguration of Our Lord Jesus Christ}, \textit{\nth{2} Double}\\
&Sts. Sixtus II (Pope \& Martyr), Felicissimus \& Agapitus (Deacons), Martyrs, \textit{Memorial}\\
7&St. Donatus, Bishop \& Martyr, \textit{Memorial}\\
8&Sts. Cyriacus, Largus, \& Smaragdus, Martyrs, \textit{Memorial}\\
9&Vigil of St. Laurence, \textit{Vigil}\\
&St. Romanus, Martyr, \textit{Memorial}\\
10&{\color{RubricRed}St. Laurence, Deacon, \& Martyr}, \textit{\nth{2} Double}, Simple Octave\\
11&St. Tiburtius \& Susanna, Martyrs, \textit{Memorial}\\
12&\\
13&St. Maximus of Constantinople, Confessor, \textit{Double}\\
&Sts. Hippolytus \& Cassian, Martyrs, \textit{Memorial}\\
14&Vigil of the Assumption, \textit{Vigil}\\
&St. Eusebius, Priest \& Confessor, \textit{Memorial}\\
15&{\color{RubricRed}Assumption of the B.V.M.}, \textit{\nth{1} Double}, Common Octave\\
16&{\color{RubricRed}St. Joachim, Father of the B.V.M.}, \textit{\nth{2} Double}\\
17&Day III within the Octave of the Assumption, \textit{Semidouble}\\
&Octave Day of St. Laurence, \textit{Simple}\\
18&Day IV within the Octave of the Assumption, \textit{Semidouble}\\
&St. Helen, Empress, \textit{Memorial}\\
&St. Agapitus, Martyr, \textit{Memorial}\\
19&Day V within the Octave of the Assumption, \textit{Semidouble}\\
20&Day VI within the Octave of the Assumption, \textit{Semidouble}\\
21&Day VII within the Octave of the Assumption, \textit{Semidouble}\\
22&Octave Day of the Assumption, \textit{Greater Double}\\
&St. Timothy (Bishop), Hippolytus, \& Symphorian, Martyrs, \textit{Memorial}\\
23&Vigil of St. Bartholomew, \textit{Vigil}\\
24&{\color{RubricRed}St. Bartholomew, Apostle}, \textit{\nth{2} Double}\\
25&\\
26&St. Zephyrinus, Pope \& Martyr, \textit{Memorial}\\
27&\\
28&St. Augustine of Hippo, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
&St. Hermes, Martyr, \textit{Memorial}\\
29&{\color{RubricRed}Beheading of St. John Baptist}, \textit{Greater Double}\\
&St. Sabina, Martyr, \textit{Memorial}\\
30&Sts. Felix \& Adauctus, Martyrs, \textit{Memorial}\\
31&St. Aidan of Lindisfarne, Bishop \& Confessor, \textit{Memorial}\\
\multicolumn{2}{l}{\textsc{September}}\\
1&St. Giles, Abbot, \textit{Memorial}\\
&Twelve Holy Brothers, Martyrs, \textit{Memorial}\\
2&St. Stephen of Hungary, King \& Confessor, \textit{Memorial}\\
3&\\
4&St. Gorazde of Prague, Bishop \& Martyr, \textit{Double}\\
5&\\
6&\\
7&\\
8&{\color{RubricRed}Nativity of the B.V.M.}, \textit{\nth{2} Double}, Simple Octave\\
&St. Hadrian, Martyr, \textit{Memorial}\\
9&St. Gorgonius, Martyr, \textit{Memorial}\\
10&\\
11&Sts. Protus \& Hyacinth, Martyrs, \textit{Memorial}\\
12&Most Holy Name of Mary, \textit{Greater Double}\\
13&\\
14&{\color{RubricRed}Exaltation of the Holy Cross}, \textit{Greater Double}\\
15&{\color{RubricRed}Seven Sorrows of the B.V.M.}, \textit{\nth{2} Double}\\
&St. Nicomedes, Martyr, \textit{Memorial}\\
16&Sts. Cornelius (Pope) \& Cyprian (Bishop), Martyrs, \textit{Double}\\
&Sts. Euphemia, Lucy, \& Geminian, Martyrs, \textit{Memorial}\\
17&\\
18&\\
19&St. Januarius \& Companions, Martyrs, \textit{Double}\\
&St. Theodore of Canterbury, Bishop \& Confessor, \textit{Memorial}\\
20&St. Eustace \& Companions, Martyrs, \textit{Double}\\
&Vigil of St. Matthew, \textit{Vigil}\\
21&{\color{RubricRed}St. Matthew, Apostle, \& Evangelist}, \textit{\nth{2} Double}\\
22&St. Maurice \& Companions, Martyrs, \textit{Memorial}\\
23&St. Linus, Pope \& Martyr, \textit{Memorial}\\
&St. Thecla, Virgin \& Martyr, \textit{Memorial}\\
24&\\
25&\\
26&Sts. Cyprian \& Justina, Martyrs, \textit{Memorial}\\
27&Sts. Cosmas \& Damian, Martyrs, \textit{Memorial}\\
28&St. Wenceslaus, Martyr, \textit{Memorial}\\
29&{\color{RubricRed}Dedication of St. Michael the Archangel}, \textit{\nth{1} Double}\\
30&St. Jerome, Priest, Confessor, \& Doctor, \textit{Greater Double}\\
&St. Gregory of Armenia, Bishop \& Confessor, \textit{Memorial}\\
\multicolumn{2}{l}{\textsc{October}}\\
1&St. Remigius, Bishop \& Confessor, \textit{Memorial}\\
2&{\color{RubricRed}Holy Guardian Angels}, \textit{Greater Double}\\
3&\\
4&\\
5&St. Placidus \& Companions, Martyrs, \textit{Double}\\
6&\\
7&{\color{RubricRed}The Holy Rosary of the B.V.M.}, \textit{Greater Double}\\
&St. Mark of Rome, Bishop \& Confessor, \textit{Memorial}\\
&Sts. Sergius, Bacchus, Marcellus, \& Apuleius, Martyrs, \textit{Memorial}\\
8&\\
9&Sts. Denys, Rusticus, \& Eleutherius, Martyrs, \textit{Memorial}\\
10&\\
11&{\color{RubricRed}Motherhood of the B.V.M.}, \textit{\nth{2} Double}\\
12&St. Wilfrid, Bishop \& Confessor, \textit{Memorial}\\
13&St. Edward, King \& Confessor, \textit{Memorial}\\
14&St. Callistus, Pope \& Martyr, \textit{Double}\\
15&{\color{RubricRed}Our Lady of Walsingham}, \textit{Double}\\
16&\\
17&\\
18&{\color{RubricRed}St. Luke, Evangelist}, \textit{\nth{2} Double}\\
19&St. Frideswide, Virgin, \textit{Memorial}\\
20&\\
21&St. Hilarion, Abbot, \textit{Memorial}\\
&Sts. Ursula \& Companions, Martyrs, \textit{Memorial}\\
22&\\
23&\\
24&{\color{RubricRed}St. Raphael the Archangel}, \textit{Greater Double}\\
25&Sts. Chrysanthus \& Daria, Martyrs, \textit{Memorial}\\
26&St. Evaristus, Pope \& Martyr, \textit{Memorial}\\
27&Vigil of Sts. Simon \& Jude, \textit{Vigil}\\
28&{\color{RubricRed}Sts. Simon \& Jude, Apostles}, \textit{\nth{2} Double}\\
29&\\
30&\\
31&Vigil of All Hallows, \textit{Vigil}\\
\multicolumn{2}{l}{\textsc{November}}\\
1&{\color{RubricRed}All Hallows}, \textit{\nth{1} Double}, Common Octave\\
2&{\color{RubricRed}All Souls}, \textit{Double}\\
&{Day II within the Octave of All Hallows}, \textit{Semidouble}\\
3&Day III within the Octave of All Hallows, \textit{Semidouble}\\
&St. Winifred, Virgin \& Martyr, \textit{Memorial}\\
4&Day IV within the Octave of All Hallows, \textit{Semidouble}\\
&Sts. Vitalis \& Agricola, Martyrs, \textit{Memorial}\\
5&Day V within the Octave of All Hallows, \textit{Semidouble}\\
&St. Elizabeth, Mother of St. John Baptist, \textit{Memorial}\\
6&Day VI within the Octave of All Hallows, \textit{Semidouble}\\
7&Day VII within the Octave of All Hallows, \textit{Semidouble}\\
&St. Willibrord, Bishop \& Confessor, \textit{Memorial}\\
8&Octave Day of All Hallows, \textit{Greater Double}\\
&Four Crowned Martyrs, \textit{Memorial}\\
9&Dedication of the Basilica of St. Saviour, \textit{Greater Double}\\
&St. Theodore Tyro, Martyr, \textit{Memorial}\\
10&Sts. Tryphon, Respicius, \& Nympha, Martyrs, \textit{Memorial}\\
11&St. Martin of Tours, Bishop \& Confessor, \textit{Greater Double}\\
&St. Mennas, Martyr, \textit{Memorial}\\
12&St. Martin I, Pope \& Martyr, \textit{Memorial}\\
13&St. Britius of Tours, Bishop \& Confessor, \textit{Memorial}\\
14&St. Gregory Palamas, Bishop \& Confessor, \textit{Memorial}\\
15&\\
16&\\
17&St. Hilda of Whitby, Virgin, \textit{Memorial}\\
&St. Gregory of the Wonder-worker, Bishop \& Confessor, \textit{Memorial}\\
18&Dedication of the Basilica of the Holy Apostles Peter \& Paul, \textit{Greater Double}\\
19&St. Pontianus, Pope \& Martyr, \textit{Memorial}\\
20&St. Edmund, King \& Martyr, \textit{Double}\\
21&Presentation of the B.V.M., \textit{Greater Double}\\
&St. Gelasius, Bishop \& Confessor, \textit{Memorial}\\
&St. Columbanus, Abbot, \textit{Memorial}\\
22&St. Cecilia, Virgin \& Martyr, \textit{Greater Double}\\
23&St. Clement, Pope \& Martyr, \textit{Double}\\
&St. Felicitas, Martyr, \textit{Memorial}\\
24&St. Chrysogonus, \textit{Memorial}\\
25&St. Catherine of Alexandria, Virgin \& Martyr, \textit{Double}\\
26&St. Peter of Alexandria, Bishop \& Martyr, \textit{Memorial}\\
27&\\
28&\\
29&Vigil of St. Andrew, \textit{Vigil}\\
&St. Saturninus, Bishop \& Martyr, \textit{Memorial}\\
30&{\color{RubricRed}St. Andrew, Apostle}, \textit{\nth{2} Double}\\
\multicolumn{2}{l}{\textsc{December}}\\
1&\\
2&St. Peter Chrysologus, Bishop, Confessor, \& Doctor, \textit{Double}\\
&St. Bibiana, Virgin \& Martyr, \textit{Memorial}\\
3&\\
4&St. Barbara, Virgin \& Martyr, \textit{Memorial}\\
5&St. Sabbas of Jud{\ae}a, Abbot	, \textit{Memorial}\\
6&St. Nicholas, Bishop \& Confessor, \textit{Double}\\
7&St. Ambrose of Milan, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
8&{\color{RubricRed}Conception of the B.V.M.}, \textit{\nth{2} Double}, Common Octave\\
9&Day II within the Octave of the Conception, \textit{Semidouble}\\
10&Day III within the Octave of the Conception, \textit{Semidouble}\\
&St. Melchiades, Pope \& Martyr, \textit{Memorial}\\
11&Day IV within the Octave of the Conception, \textit{Semidouble}\\
&St. Damasus, Bishop \& Confessor, \textit{Memorial}\\
12&Day V within the Octave of the Conception, \textit{Semidouble}\\
13&St. Lucy, Virgin \& Martyr, \textit{Greater Double}\\
&Day VI within the Octave of the Conception, \textit{Semidouble}\\
&St. Herman of Alaska, Priest \& Confessor, \textit{Memorial}\\
14&Day VII within the Octave of the Conception, \textit{Semidouble}\\
15&Octave Day of the Conception of the B.V.M., \textit{Greater Double}\\
16&St. Eusebius of Vercelli, Bishop \& Martyr, \textit{Memorial}\\
17&\\
18&Expectation of the B.V.M., \textit{Double}\\
19&\\
20&Vigil of St. Thomas, \textit{Vigil}\\
21&{\color{RubricRed}St. Thomas, Apostle}, \textit{\nth{2} Double}\\
22&\\
23&\\
24&{\color{RubricRed}Vigil of the Nativity of Our Lord}, \textit{\nth{1} Vigil}\\
25&{\color{RubricRed}Nativity of Our Lord}, \textit{\nth{1} Double}, \nth{3} Octave\\
&St. Anastasia, Virgin \& Martyr, \textit{Commemoration}\\
26&{\color{RubricRed}St. Stephen, Protomartyr}, \textit{\nth{2} Double}, Simple Octave\\
&Day II within the Nativity Octave, \textit{Semidouble}\\
27&{\color{RubricRed}St. John, Apostle \& Evangelist}, \textit{\nth{2} Double}, Simple Octave\\
&Day III within the Nativity Octave, \textit{Semidouble}\\
28&{\color{RubricRed}Holy Innocents, Martyrs}, \textit{\nth{2} Double}, Simple Octave\\
&Day IV within the Nativity Octave, \textit{Semidouble}\\
29&Day V within the Nativity Octave, \textit{Semidouble}\\
30&Day VI within the Nativity Octave, \textit{Semidouble}\\
31&St. Sylvester, Bishop \& Confessor, \textit{Double}\\
&Day VII within the Nativity Octave, \textit{Semidouble}\\
\end{longtable}