\section{Aspersion of the People}
\fancyhead[RO,LE]{\textit{Aspersion}}
\subby{Asperges, Me}
%Rubrics from the Anglican Missal. Replace with English Missal rubrics?
%\begin{rubric}
%    The following order for the aspersion of holy water is to be observed on all Sundays (outside of Eastertide) before the principal Mass, even though it be not sung.
%\end{rubric}
\begin{rubric}
    Having blest the water, the Priest who is going to celebrate, vested in a cope of the colour of the day, proceeds to the Altar. And there kneeling, even in Eastertide, with the ministers below the steps, he receives the aspersorium from the Deacon, and first asperses the Altar thrice, then himself, and, standing up, the ministers, beginning the Antiphon \emph{Thou shalt purge me.} And the Quire continues: \emph{O Lord, with hyssop,} etc., as below. Meanwhile the Celebrant asperses the Clergy, and then the people, saying with a low voice with the ministers the Psalm \emph{Have mercy upon me, O God.}
\end{rubric}
\gregorioscore{resources/gabc/Asperges.gabc}
\begin{rubric}
    The following \emph{Gloria Patri} is omitted in Passiontide, the psalm simply being repeated.
\end{rubric}
\gregorioscore{resources/gabc/GloriaPatri.gabc}
%\elcol{\textit{Ant.} Thou shalt purge me, * O Lord, with hyssop, and I shall be clean: thou shalt wash me, and I shall be whiter than snow. \textit{Ps.} Have mercy upon me, O God, after thy great goodness.}{\textit{Ant.} Asperges me, Domine, hyssopo et mundabor, * Lavabis me, et super nivem dealbabor. \textit{Ps.} Miserere mei, Deus, * secundum magnam misericordiam tuam.}
%\elcol{\textsc{Glory} be to the Father, and to the Son, and to the Holy Ghost: * as it was in the beginning, is now and ever shall be, world without end. Amen.}{\textsc{Glória} Patri, et Fílio, et Spirítui Sancto. Sicut erat in princípio, et nunc, et semper: et in sǽcula s{\ae}culórum. Amen.}
%\elcol{\textit{Ant.} Thou shalt purge me, * O Lord, with hyssop, and I shall be clean: thou shalt wash me, and I shall be whiter than snow. \textit{Ps.} Have mercy upon me, O God, after thy great goodness.}{\textit{Ant.} Asperges me, Domine, hyssopo et mundabor, * Lavabis me, et super nivem dealbabor. \textit{Ps.} Miserere mei, Deus, * secundum magnam misericordiam tuam.}
\subby{Vidi Aquam}
\begin{inhead}
	{During Eastertide}
\end{inhead}
\elcol{\textit{Ant.} I beheld water * issuing out from the temple, on the right side, alleluia: and all to whom that water came were saved, and they shall say: alleluia, alleluia.}{\textit{Ant.} Vidi aquam egredientem de templo, * a latere dextro, allelúja: * Et omnes, ad quos pervenit aqua ista, * salvi facti sunt et dicent, allelúja, allelúja.}
\elcol{\textsc{O give} thanks unto the Lord, for he is gracious: because his mercy endureth for ever.}{\textsc{Confitémini} Domino, quoniam bonus: * quoniam in saeculum misericordia ejus.}
\elcol{\textsc{Glory} be to the Father, and to the Son, and to the Holy Ghost: * as it was in the beginning, is now and ever shall be, world without end. Amen.}{\textsc{Glória} Patri, et Fílio, et Spirítui Sancto. Sicut erat in princípio, et nunc, et semper: et in sǽcula s{\ae}culórum. Amen.}
\elcol{\textit{Ant.} I beheld water * issuing out from the temple, on the right side, alleluia: and all to whom that water came were saved, and they shall say: alleluia, alleluia.}{\textit{Ant.} Vidi aquam egredientem de templo, * a latere dextro, allelúja: * Et omnes, ad quos pervenit aqua ista, * salvi facti sunt et dicent, allelúja, allelúja.}