\fancyhead[C]{{\LARGE Eucharistic Devotions}}
\fancyhead[RO,LE]{\textit{Benediction}}
\section{Benediction of the Blessed Sacrament}
\begin{rubric}
The Priest brings the Sacrament out of the tabernacle with all kneeling, singing:
\end{rubric}
\subby{O Salutaris Hostia}
\elcol{\lett{O}{saving} Victim op'ning wide * The Gate of heav'n to man below,\par
\secondline{Our foes press on from ev'ry side, * Thine aid supply, thy strength bestow.}
\thirdline{All praise and thanks to thee ascend * For evermore, blest One in Three;}
O grant us life that shall not end, * In our true native land with thee. Amen.
}{\lett{O}{salut\smash{á}ris} Hóstia, * Qu{\ae} caeli pandis óstium:\par
\secondline{Bella premunt hostília, * Da robur, fer auxílium.}
\thirdline{Uni trinóque Dómino * Sit sempitérna glória,}
Qui vitam sine término * Nobis donet in pátria. Amen.
}
\begin{rubric}
After a time for silence, the following hymn is sung:
\end{rubric}
\subby{Tantum Ergo Sacramentum}
\elcol{\lett{T}{herefore} we, before him bending, * This great Sacrament revere;\par
\secondline{Types and shadows have their ending, * For the newer rite is here;}
\thirdline{Faith, our outward sense befriending, * Makes our inward vision clear.\\}
\par
Glory let us give and blessing * To the Father and the Son,\par
Honour, thanks, and praise addressing, * While eternal ages run;\par
Ever too his love confessing * Who from One with Both is One. Amen.\\
}{\lett{T}{\smash{á}ntum} ergo Sacraméntum * Venerémur cérnui,\par
\secondline{Et antíquum documéntum, * Nóvo cédat rítui,}
\thirdline{Prǽstet fídes suppleméntum * Sénsuum deféctui.\\}
\par
Genitóri, Genitóque, * Laus et jubilátio;\par
Sálus, hónor, vírtus quoque * Sit et benedíctio.\par
Procedénti ab unóque * Cómpar sit laudátio. Amen.\\}
\elcol{℣. Thou gavest them bread from heaven. (Alleluia)}{℣. Panem de c{\ae}lis pr{\ae}stitísti eis. (Allelúja.)}
\elcol{℟. Containing in itself all sweetness. (Alleluia)\\}{℟. Omne delectaméntum in se habéntem. (Allelúja.)\\}
\oremus
\elcoldent{\lett{O}{God,} who in this wonderful Sacrament hast left us a perpetual Memorial of Thy Passion: Grant us, we beseech thee, so to venerate the Sacred Mysteries of thy Body and Blood, that we may ever perceive within ourselves the fruit of thy redemption; who livest and reignest, world without end. \textit{Amen.}}{\lett{D}{eus,} qui nobis sub sacraménto mirábili, passiónis tu{\ae} memóriam reliquísti: tríbue, quǽsumus, ita nos córporis et sánguinis tui sacra mystéria venerári, ut redemptiónis tu{\ae} fructum in nobis iúgiter sentiámus. Qui vivis et regnas in s{\ae}cula s{\ae}culorum. \textit{Amen.}}
\begin{rubric}
All kneeling, the Priest blesses the People with the Blessed Sacrament, then placing it upon the altar and saying the Divine Praises.
\end{rubric}
\subby{Divine Praises}
\elcol{Blessed be God.}{Benedíctus Deus.}
\elcol{Blessed be His Holy Name.}{Benedíctum Nomen Sanctum ejus.}
\elcol{Blessed be Jesus Christ, true God and true Man.}{Benedíctus Jesus Christus, verus Deus et verus homo.}
\elcol{Blessed be the Name of Jesus.}{Benedíctum Nomen Jesu.}
\elcol{Blessed be Jesus in the Most Holy Sacrament of the Altar.}{Benedíctus Jesus in sanctíssimo altáris Sacraménto.}
\elcol{Blessed be God the Holy Ghost, the Comforter.}{Benedíctus Sanctus Spíritus, Paraclítus.}
\elcol{Blessed be the Great Mother of God, Mary most Holy.}{Benedícta magna Mater Dei, María sanctíssima.}
\elcol{Blessed be the name of Mary, Virgin and Mother.}{Benedíctum nomen Marí{\ae}, Vírginis et Matris.}
\elcol{Blessed be Saint Joseph, her most chaste spouse.}{Benedíctus sanctus Joseph, ejus castíssimus Sponsus.}
\elcol{Blessed be God in His Angels and in His Saints.}{Benedíctus Deus in Angelis suis, et in Sanctis suis.}

\subby{Psalm 117. \textit{Laudate Dominum}}
\begin{rubric}
The Blessed Sacrament is replaced in the Tabernacle. All stand and pray the following,
\end{rubric}\par\noindent
\elcol{\antiphon{}{Let us forever adore the Most Holy Sacrament. (Alleluia.)}\\}{\antiphon{}{Adorémus in {\ae}térnum Sanctissimum Sacramentum. (Allelúja.)}\\}
\elcol{\lett{O}{praise} the \textsc{Lord}, all ye nations; * praise him, all ye peoples.}{\lett{L}{aud\smash{á}te} Dóminum, omnes gentes: * laudáte eum, omnes pópuli:}
\elcol{For his merciful kindness is ever more and more toward us; *
and the truth of the \textsc{Lord} endureth for ever. Praise the \textsc{Lord}.}{Quóniam confirmáta est super nos misericórdia ejus: * et véritas Dómini manet in {\ae}térnum.}
\elcol{Glory be to the Father, and to the Son, *
    and to the Holy Ghost:}{Glória Patri, et Fílio, * et Spirítui Sancto.}
\elcol{As it was in the beginning, is now, and ever shall be, *
    world without end. Amen.}{Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}\par\noindent
\elcol{\textit{Ant.} Let us forever adore the Most Holy Sacrament. (Alleluia.)}{\textit{Ant.} Adorémus in {\ae}térnum Sanctíssimum Sacraméntum. (Allelúja.)}