\fancyhead[RO,LE]{\textit{Thanksgiving}}
\fancyhead[RE,LO]{}
\section{Thanksgiving for Holy Communion}\label{CommunionThanksgiving}%From 1933 English Missal:
\begin{multicols}{2}\par\noindent
\antiphon{}{Let us sing {\dag} the song of the three children, which the Saints sang in the furnace of fire, blessing the Lord (Alleluia).}
\begin{inhead}
	Benedicite
\end{inhead}
\lett{O}{all} ye Works of the Lord, bless ye the Lord: * praise him, and magnify him for ever.\par
\secondline{O ye Angels of the Lord, bless ye the Lord: * O ye heavens, bless ye the Lord.}
\thirdline{O ye waters that be above the firmament, bless ye the Lord: * O all ye powers of the Lord, bless ye the Lord.}

    O ye sun and moon, bless ye the Lord: * O ye stars of heaven, bless ye the Lord.
    
    O ye showers and dew, bless ye the Lord: * O ye winds of God, bless ye the Lord.
    
    O ye fire and heat, bless ye the Lord: * O ye winter and summer, bless ye the Lord.
    
    O ye dews and frosts, bless ye the Lord: * O ye frost and cold, bless ye the Lord.
    
    O ye ice and snow, bless ye the Lord: * O ye nights and days, bless ye the Lord.
    
    O ye light and darkness, bless ye the Lord: * O ye lightnings and clouds, bless ye the Lord.

\columnbreak
\lett{O}{let} the earth bless the Lord: yea, let it praise him, and magnify him for ever.\par
\secondline{O ye mountains and hills, bless ye the Lord: * O all ye green things upon the earth, bless ye the Lord.}
\thirdline{O ye wells, bless ye the Lord: * O ye Seas and Floods, bless ye the Lord.}
    O ye whales, and all that move in the waters, bless ye the Lord: * O all ye fowls of the air, bless ye the Lord.
    
    O all ye beasts and cattle, bless ye the Lord: * O ye children of men, bless ye the Lord.
    
\lett{O}{let} Israel bless the Lord: * praise him, and magnify him for ever.\par
\secondline{O ye priests of the Lord, bless ye the Lord: * O ye servants of the Lord, bless ye the Lord.}
\thirdline{O ye spirits and souls of the righteous, bless ye the Lord: * O ye holy and humble men of heart, bless ye the Lord.}
O Ananias, Azarias, and Misael, bless ye the Lord: * praise him, and magnify him for ever.

    \lett{L}{et} us bless the Father and the Son with the Holy Ghost: * let us praise him and magnify him for ever.\par
    \secondline{Blessed art thou, O Lord, in the firmament of heaven: * and to be praised and glorified and exalted above all for ever.}

\begin{rubric}
	Here neither \emph{Glory be} nor \emph{Amen} is said.
\end{rubric}

\begin{inhead}
	Psalm 150
\end{inhead}

\lett{O}{praise} God in his holiness : * praise him in the firmament of his power.\par
\secondlinepsalm{Praise him in his noble acts : * praise him according to his excellent greatness.}
\thirdlinepsalm{Praise him in the sound of the trumpet : * praise him upon the lute and harp.}
\psanum{4}Praise him in the cymbals and dances : * praise him upon the strings and pipe.\par
\psanum{5}Praise him upon the well-tuned cymbals : * praise him upon the loud cymbals.\par
\psanum{6}Let every thing that hath breath * praise the Lord.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\antiphon{}{Let us sing the song of the three children, which the Saints sang in the furnace of fire, blessing the Lord (Alleluia).}\\

℣. Lord, have mercy upon us.

℟. Christ, have mercy upon us.

℣. Lord, have mercy upon us.

\begin{rubric}
    {The Our Father is here said, in secret, kneeling, ending with,}
\end{rubric}

℣. And lead us not into temptation.

℟. But deliver us from evil.

℣. All thy works praise thee, O Lord.

℟. And thy Saints give thanks unto thee.

℣. Let thy Saints be joyful in glory,

℟. Let them rejoice in their beds.

℣. Not unto us, O Lord, not unto us,

℟. But unto thy name give the praise.

℣. O Lord, hear my prayer.

℟. And let my cry come unto thee.

℣. The Lord be with you.

℟. And with thy spirit.

\letuspray
\lett{O}{God,} who to the three children didst assuage the flames of fire: mercifully grant; that the flames of sin may not kindle upon thy servants.
\needspace{4\baselineskip}
\lett{P}{revent} us, O Lord, in all our doings with thy most gracious favour, and further us with thy continual help: that all our prayer and work may be begun, continued, and ended in thee.
\needspace{4\baselineskip}
\lett{G}{rant} to us, we beseech thee, O Lord, that we may quench the flames of our sins: as thou didst enable blessed Laurence to overcome the fires of his torments. Through Christ, our Lord. \textit{Amen.}
\end{multicols}