\fancyhead[RO,LE]{}
\fancyhead[RE,LO]{}
\fancyhead[C]{{\LARGE Prayers}}\label{prayers}
\phantomsection
\addcontentsline{toc}{section}{Prayers}
%Uniquely 1928 Prayers are mostly omitted.

%This section provides an interesting problem. We want to supply and abundance of prayers and thanksgivings for these United States and Canada. However, so many become overwhelming. Therefore, some are omitted or adjusted, in consultation with the Russian Observances (http://anglicanhistory.org/alcuin/tract12.html).
%\begin{rubric}
%To be used before the Prayer for all Conditions of Men, or, when that is not said, before the final Prayer of Thanksgiving or of Blessing, or before the Grace.
%\end{rubric}
%\begin{multicols}{2}
\subby{Prayers for the United States of America}
%Prayer for the President of the United States is moved out of the Offices, into the Prayers section, together with the Prayers for the King, since it is not certain which will be needful.
\subbysub{For the President of the United States, and all in Civil Authority}
\lett{O}{Lord}, our heavenly Father, the high and mighty Ruler of the universe, who dost from thy throne behold all the dwellers upon earth; Most heartily we beseech thee, with thy favour to behold and bless thy servant \textsc{the President of the United States}, and all others in authority; and so replenish them with the grace of thy Holy Spirit, that they may always incline to thy will, and walk in thy way. Endue them plenteously with heavenly gifts; grant them in health and prosperity long to live; and finally, after this life, to attain everlasting joy and felicity; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For Congress}
\begin{rubric}
To be used during their Session.
\end{rubric}
\lett{M}{ost} gracious God, we humbly beseech thee, as for the people of these United States in general, so especially for their Senate and Representatives in Congress assembled; that thou wouldest be pleased to direct and prosper all their consultations, to the advancement of thy glory, the good of thy Church, the safety, honour, and welfare of thy people; that all things may be so ordered and settled by their endeavours, upon the best and surest foundations, that peace and happiness, truth and justice, religion and piety, may be establish among us for all generations. These and all other necessaries, for them, for us, and thy whole Church, we humbly beg in the Name and mediation of Jesus Christ, our most blessed Lord and Saviour. \textit{Amen.}

%\subbysub{For a State Legislature}
%\lett{O}{God,} the fountain of wisdom, whose statutes are good and gracious and whose law is truth; We beseech thee so to guide and bless the Legislature of this State, that it may ordain for our governance only such things as please thee, to the glory of thy Name and the welfare of the people; through Jesus Christ, thy Son, our Lord. \textit{Amen.}

%\subbysub{For Courts of Justice}
%\lett{A}{lmighty} God, who sittest in the throne judging right; We humbly beseech thee to bless the courts of justice and the magistrates in all this land; and give unto them the spirit of wisdom and understanding, that they may discern the truth and impartially administer the law in the fear of thee alone; through him who shall come to be our judge, thy Son, our Saviour, Jesus Christ. \textit{Amen.}

%Army changed to Military to broaden it to all forces.
\subbysub{For the Military}
\lett{O}{Lord} God of Hosts, stretch forth, we pray thee, thine almighty arm to strengthen and protect the soldiers of our country; Support them in the day of battle, and in the time of peace keep them safe from all evil; endue them with courage and loyalty; and grant that in all things they may serve without reproach; through Jesus Christ our Lord. \textit{Amen.}

%Omitted, because there is the prayer for those at sea.
%\subbysub{For the Navy}
%\lett{O}{Eternal} Lord God, who alone spreadest out the heavens, and rulest the raging of the sea; Vouchsafe to take into thy almighty and most gracious protection our country's Navy, and all who serve therein. Preserve them from the dangers of the sea, and from the violence of the enemy; that they may be a safeguard unto the United States of America, and a security for such as pass on the seas upon their lawful occasions; that the inhabitants of our land may in peace and quietness serve thee our God, to the glory of thy Name; through Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For Our Country}
%\lett{A}{lmighty} God, who hast given us this good land for our heritage; We humbly beseech thee that we may always prove ourselves a people mindful of thy favour and glad to do thy will. Bless our land with honourable industry, sound learning, and pure manners. Save us from violence, discord, and confusion; from pride and arrogancy, and from every evil way. Defend our liberties, and fashion into one united people the multitudes brought hither out of many kindreds and tongues. Endue with the spirit of wisdom those to whom in thy Name we entrust the authority of government, that there may be justice and peace at home, and that, through obedience to thy law, we may show forth thy praise among the nations of the earth. In the time of prosperity, fill our hearts with thankfulness, and in the day of trouble, suffer not our trust in thee to fail; all which we ask through Jesus Christ our Lord. \textit{Amen.}

\subby{Prayers for Canada \& Commonwealth}
\subbysub{For the King's Majesty}
\lett{O}{Lord}, our heavenly Father, high and mighty, King of kings, Lord of lords, the only Ruler of princes, who dost from thy throne behold all the dwellers upon earth: Most heartily we beseech thee with thy favour to behold our most gracious Sovereign Lord, King \textsc{\textit{Charles}}; and so replenish him with the grace of thy Holy Spirit, that he may alway incline to thy will, and walk in thy way. Endue him plenteously with heavenly gifts; grant him in health and wealth long to live; strengthen him that he may vanquish and overcome all his enemies; and finally after this life he may attain everlasting joy and felicity; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For the Royal Family}
\lett{A}{lmighty} God, the fountain of all goodness, we humbly beseech thee to bless Queen \textit{Camilla}, \textit{William} Prince of Wales, the Princess of Wales, and all the Royal Family: Endue them with thy Holy Spirit; enrich them with thy heavenly grace; prosper them with all happiness; and bring them to thine everlasting kingdom; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For Parliament}
\lett{M}{ost} gracious God, we humbly beseech thee, as for this Empire in general, so especially for the High Court of Parliament, \textit{the Parliament of this Dominion, and the Legislature of this Province,} under our most religious and gracious King at this time assembled: That thou wouldest be pleased to direct and prosper all their consultations to the advancement of thy glory, the good of thy Church, the safety, honour, and welfare of our Sovereign and his Dominions; that all things may be so ordered and settled by their endeavours, upon the best and surest foundations, that peace and happiness, truth and justice, religion and piety, may be established among us for all generations. These and all other necessaries, for them, for us, and thy whole Church, we humbly beg in the Name and Mediation of Jesus Christ our most blessed Lord and Saviour. \textit{Amen.}

\subbysub{For the Forces of the King}
\lett{O}{Lord} of Hosts, stretch forth, we pray thee, thine Almighty arm to strengthen and protect the forces of our King in every peril of sea, and land, and air; shelter them in the day of battle, and in time of peace keep them safe from all evil; endue them ever with loyalty and courage; and grant that in all things they may serve as seeing Thee who art invisible; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For the Commonwealth}
%Supplied from the English 1928's Prayer for the British Empire. I would love to use the 1962's Prayer, but I fear it may be in copyright still.
\lett{A}{lmighty} God, who rulest in the kingdom of men, and hast given to our Sovereign Lord, King \textsc{\textit{Charles}}, a great dominion in all parts of the earth: Draw together, we pray thee, in true fellowship the men of divers races, languages, and customs, who dwell therein, that, bearing one another's burdens, and working together in brotherly concord, they may fulfil the purpose of thy providence, and set forward thy everlasting kingdom. Pardon, we beseech thee our sins and shortcomings: keep far from us an selfishness and pride: and give us grace to employ thy good gifts of order and freedom to thy glory and the welfare of mankind; through Jesus Christ thy Son our Lord, to whom with thee and the Holy Ghost be all glory and dominion, world without end. \textit{Amen.}

\subby{Prayers for the Church}
\subbysub{For the Unity of God's People}
\lett{O}{God,} the Father of our Lord Jesus Christ, our only Saviour, the Prince of Peace; Give us grace seriously to lay to heart the great dangers we are in by our unhappy divisions. Take away all hatred and prejudice, and whatsoever else may hinder us from godly union and concord: that as there is but one Body and one Spirit, and one hope of our calling, one Lord, one Faith, one Baptism, one God and Father of us all, so we may be all of one heart and of one soul, united in one holy bond of truth and peace, of faith and charity, and may with one mind and one mouth glorify thee; through Jesus Christ our Lord. \textit{Amen.}
\subbysub{For those who are to be admitted into Holy Orders}
\begin{rubric}
To be used in the Weeks preceding the stated Times of Ordination.
\end{rubric}
\lett{A}{lmighty} God, our heavenly Father, who hast purchased to thyself an universal Church by the precious blood of thy dear Son; Mercifully look upon the same, and at this time so guide and govern the minds of thy servants the Bishops and Pastors of thy flock, that they may lay hands suddenly on no man, but faithfully and wisely make choice of fit persons, to serve in the sacred Ministry of thy Church, And to those who shall be ordained to any holy function, give thy grace and heavenly benediction; that both by their life and doctrine they may show forth thy glory, and set forward the salvation of all men; through Jesus Christ our Lord. \textit{Amen.}
%\begin{inhead}
%or,
%\end{inhead}
%\lett{A}{lmighty} God, the giver of all good gifts, who of thy divine providence hast appointed divers Orders in thy Church; Give thy grace, we humbly beseech thee, to all those who are to be called to any office and administration in the same; and so replenish them with the truth of thy doctrine, and endue them with innocency of life, that they may faithfully serve before thee, to the glory of thy great Name, and the benefit of thy holy Church; through Jesus Christ our Lord. \textit{Amen.}
\subbysub{For the Increase of the Ministry}
\lett{O}{almighty} God, look mercifully upon the world which thou hast redeemed by the blood of thy dear Son, and incline the hearts of many to dedicate themselves to the sacred ministry of thy Church; through the same thy Son Jesus Christ our Lord. \textit{Amen.}
%The 1928 unnecessarily softens the collect here. The 1928 is removed in preference for the 1892.
\subbysub{For Missions}
\lett{O}{God,} who hast made of one blood all nations of men for to dwell on the face of the whole earth, and didst send thy blessed Son to preach peace to them that are far off and to them that are nigh; Grant that all men everywhere may seek after thee and find thee. Bring the nations into thy fold, and add the heathen to thine inheritance. And we pray thee shortly to accomplish the number of thine elect, and to hasten thy kingdom; through the same Jesus Christ our Lord. \textit{Amen.}

\subby{Prayers in Times of Illness}

\subbysub{In Time of Great Sickness and Mortality}
%The 1928 shows the American disregard for wrath and repentance. Reverted to the 1892.
\lett{O}{almighty} God, the Lord of life and death, of sickness and health Regard our supplications, we humbly beseech thee; and, thou hast thought fit to visit us for our sins with great sickness and mortality, in the midst of thy judgment, O Lord, remember mercy. Have pity upon us miserable sinners, and withdraw from us the grievous sickness with which we are afflicted. May this thy fatherly correction have its due influence upon us, by leading us to consider how frail and uncertain our life is; that we may apply our hearts unto that heavenly wisdom which in the end will bring us to everlasting life; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For a Sick Person}
\lett{O}{Father} of mercies and God of all comfort, our only help in time of need; We humbly beseech thee to behold, visit, and relieve thy sick \textit{servant} \textit{N.} for whom our prayers are desired. Look upon \textit{him} with the eyes of thy mercy; comfort \textit{him} with a sense of thy goodness; preserve \textit{him} from the temptations of the enemy; and give \textit{him} patience under \textit{his} affliction. In thy good time, restore \textit{him} to health, and enable \textit{him} to lead the residue of \textit{his} life in thy fear, and to thy glory; and grant that finally \textit{he} may dwell with thee in life everlasting; through Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For a Sick Child}
%\lett{O}{heavenly} Father, watch with us, we pray thee, over the sick child for whom our prayers are offered, and grant that \textit{he} may be restored to that perfect health which it is thine alone to give; through Jesus Christ our Lord. \textit{Amen.}

\subby{Prayers in Times of Other Necessities}

\subbysub{For Fruitful Seasons}
\begin{rubric}
	{To be used on Rogation Sunday and the Rogation Days}
\end{rubric}
\lett{A}{lmighty} God, who hast blessed the earth that it should be fruitful and bring forth whatsoever is needful for the life of man, and hast commanded us to work with quietness, and eat our own bread; Bless the labours of the husbandman, and grant such seasonable weather that we may gather in the fruits of the earth, and ever rejoice in thy goodness, to the praise of thy holy Name; through Jesus Christ our Lord. \textit{Amen.}
%\begin{rubric}
%Or this.
%\end{rubric}
%\lett{O}{gracious} Father, who openest thine hand and fillest all things living with plenteousness; We beseech thee of thine infinite goodness to hear us, who now make our prayers and supplications unto thee. Remember not our sins, but thy promises of mercy. Vouchsafe to bless the lands and multiply the harvests of the world. Let thy breath go forth that it may renew the face of the earth. Show thy loving-kindness, that our land may give her increase; and so fill us with good things that the poor and needy may give thanks unto thy Name; through Christ our Lord. \textit{Amen.}

\subbysub{For Rain}
\lett{O}{God,} heavenly Father, who by thy Son Jesus Christ hast promised to all those who seek thy kingdom, and the righteousness thereof, all things necessary to their bodily sustenance; Send us, we beseech thee, in this our necessity, such moderate rain and showers, that we may receive the fruits of the earth to our comfort, and to thy honour; through Jesus Christ our Lord. \textit{Amen.}

%The American tendency to shy away from God's wrath has corrupted this prayer. The 1549 form is restored.
\subbysub{For Fair Weather}
\lett{O}{Lord} God, which for the sin of man didst once drown all the world, except eight persons, and afterward of thy great mercy, didst promise never to destroy it so again: We humbly beseech thee, that although we for our iniquities have worthily deserved this plague of rain and waters, yet, upon our true repentance, thou wilt send us such weather whereby we may receive the fruits of the earth in due season, and learn both by the punishment to amend our lives, and by the granting of our petition to give thee praise and glory: Through Jesu Christ our Lord. \textit{Amen.}

\subbysub{In Time of Dearth and Famine}
\lett{O}{God,} heavenly Father, whose gift it is that the rain doth fall, and the earth bring forth her increase; Behold, we beseech thee, the afflictions of thy people; increase the fruits of the earth by thy heavenly benediction; and grant that the scarcity and dearth, which we now most justly suffer for our sins, may, through thy goodness, be mercifully turned into plenty; for the love of Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all honour and glory, now and for ever. \textit{Amen.}

\subbysub{In Time of War and Tumults}
%The 1928 shows an American discomfort for condemnation. For greater conformity with the Canadian, reverted to 1662.
\lett{O}{almighty} God, King of all kings, and Governor of all things, whose power no creature is able to resist, to whom it belongeth justly to punish sinners, and to be merciful to those who truly repent; Save and deliver us, we humbly beseech thee, from the hands of our enemies; abate their pride, assuage their malice, and confound their devices; that we, being armed with thy defence, may be preserved evermore from all perils, to glorify thee, who art the only giver of all victory; through the merits of thy Son, Jesus Christ our Lord. \textit{Amen.}

\subbysub{For a Person under Affliction}
\lett{O}{merciful} God, and heavenly Father, who hast taught us in thy holy Word that thou dost not willingly afflict or grieve the children of men; Look with pity, we beseech thee, upon the sorrows of thy \textit{servant} for whom our prayers are offered. Remember \textit{him}, O Lord, in mercy; endue \textit{his} soul with patience; comfort \textit{him} with a sense of thy goodness; lift up thy countenance upon \textit{him}, and give \textit{him} peace; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For a Person, or Persons, going to Sea}
\lett{O}{eternal} God, who alone spreadest out the heavens, and rulest the raging of the sea; We commend to thy almighty protection, thy \textit{servant}, for whose preservation on the great deep our prayers are desired. Guard \textit{him}, we beseech thee, from the dangers of the sea, from sickness, from the violence of enemies, and from every evil to which \textit{he} may be exposed. Conduct \textit{him} in safety to the haven where \textit{he} would be, with a grateful sense of thy mercies; through Jesus Christ our Lord. \textit{Amen.}

\subbysub{For Prisoners}
\lett{O}{God,} who sparest when we deserve punishment, and in thy wrath rememberest mercy; We humbly beseech thee, of thy goodness, to comfort and succour all prisoners (especially those who are condemned to die). Give them a right understanding of themselves, and of thy promises; that, trusting wholly in thy mercy, they may not place their confidence anywhere but in thee. Relieve the distressed, protect the innocent, awaken the guilty; and forasmuch as thou alone bringest light out of darkness, and good out of evil, grant to these thy servants, that by the power of thy Holy Spirit they may be set free from the chains of sin, and may be brought to newness of life; through Jesus Christ our Lord. \textit{Amen.}
\begin{inhead}
    or,
\end{inhead}
%Supplied from the 1892 for those about to receive the Death Penalty.
\lett{O}{most} gracious and merciful God, we earnestly beseech thee to have pity and compassion upon these persons recommended to our prayers, who now lie under the sentence of the law and are appointed to die. Visit them, O Lord, with thy mercy and salvation convince them of the miserable condition they are in, by their sins and wickedness; and let thy powerful grace produce in them such a godly sorrow, and sincere repentance, as thou wilt be pleased to accept. Give them a strong and lively faith in thy Son, our blessed Saviour and make it effectual to the salvation of their souls. O Lord, in judgment remember mercy and whatever sufferings they are to endure in this world, yet deliver them, O God, from the bitter pains of eternal death. Pardon their sins, and save their souls, for the sake and merits of thy dear Son, our blessed Saviour and Redeemer. \textit{Amen.}

%\subbysub{In Time of Calamity}
%\lett{O}{God,} merciful and compassionate, who art ever ready to hear the prayers of those who put their trust in thee; Graciously hearken to us who call upon thee, and grant us thy help in this our need; through Jesus Christ our Lord. \textit{Amen.}

%\subby{Prayers for the Christian Life}
%\subbysub{For Schools, Colleges, and Universities}
%\lett{A}{lmighty} God, we beseech thee, with thy gracious favour, to behold our universities, colleges, and schools, that knowledge may be increased among us, and all good learning flourish and abound; bless all who teach and all who learn; and grant that in humility of heart they may ever look unto thee, who art the fountain of all wisdom; through Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For Religious Education}
%\lett{A}{lmighty} God, our Heavenly Father, who hast committed to thy Holy Church the care and nurture of thy children; Enlighten with thy wisdom those who teach and those who learn; that, rejoicing in the knowledge of thy truth, they may worship thee and serve thee from generation to generation; through Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For Children}
%\lett{O}{lord} Jesus Christ, who dost embrace children with the arms of thy mercy, and dost make them living members of thy Church; Give them grace, we pray thee, to stand fast in thy faith, to obey thy word, and to abide in thy love; that being made strong by thy Holy Spirit they may resist temptation and overcome evil; and may rejoice in the life that now is, and dwell with thee in the life that is to come; through thy merits, O merciful Saviour, who with the Father and the Holy Ghost livest and reignest one God, world without end. \textit{Amen.}

%\subbysub{For Those About to be Confirmed}
%\lett{O}{God,} who through the teaching of thy Son Jesus Christ didst prepare the disciples for the coming of the Comforter; Make ready, we beseech thee, the hearts and minds of thy servants who at this time are seeking to be strengthened by the gift of the Holy Spirit through the laying on of hands, that, drawing near with penitent and faithful hearts, they may evermore be filled with the power of his divine indwelling; through the same Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For Christian Service}
%\lett{O}{Lord,} our heavenly Father, whose blessed Son came not to be ministered unto, but to minister; We beseech thee to bless all who, following in his steps, give themselves to the service of their fellow men. Endue them with wisdom, patience, and courage, that they may strengthen the weak and raise up those who fall; and, being inspired by thy love, may worthily minister in thy Name to the suffering, the friendless, and the needy; for the sake of him who laid down his life for us, the same thy Son, our Saviour, Jesus Christ. \textit{Amen.}

%\subbysub{For Social Justice}
%\lett{A}{lmighty} God, who hast created man in thine own image; Grant us grace fearlessly to contend against evil, and to make no peace with oppression; and, that we may reverently use our freedom, help us to employ it in the maintenance of justice among men and nations, to the glory of thy holy Name; through Jesus Christ our Lord. \textit{Amen.}

%\subbysub{For Every Man in his Work}
%\lett{A}{lmighty} God, our heavenly Father, who declarest thy glory and showest forth thy handiwork in the heavens and in the earth; Deliver us, we beseech thee, in our several callings, from the service of mammon, that we may do the work which thou givest us to do, in truth, in beauty, and in righteousness, with singleness of heart as thy servants, and to the benefit of our fellow men; for the sake of him who came among us as one that serveth, thy Son, Jesus Christ our Lord. \textit{Amen.}

\subby{Prayers for the Dead}\label{dead}
\subbysub{For All the Faithful Departed}
\lett{O}{God,} the Creator and Redeemer of all the faithful: grant unto the souls of thy servants and handmaids the remission of all their sins; that through devout supplications they may obtain the pardon they have always desired. Who with God the Father in the unity of the Holy Spirit livest and reignest God, world without end. \textit{Amen.}
\subbysub{On the Day of Burial}
\lett{A}{bsolve}, O Lord, we beseech thee, the soul of thy \textit{servant N.}, that being dead to the world \textit{he} may live unto thee: and whatsoever \textit{he} hath done amiss in \textit{his} earthly conversation through frailty of the flesh, do thou wipe away by the pardon of thy merciful goodness. Through.
\subbysub{For a Priest Departed}
\lett{O}{God,} who hast made \textit{thy servant N.} to flourish among the Ministers of Apostolic Succession in the honourable office of a Priest: grant, we beseech thee, that \textit{he} may also be joined with them in a perpetual fellowship. Through.
\subbysub{Another Collect for a Priest Departed}
\lett{G}{rant}, we beseech thee, O Lord: that the soul of thy servant the Priest \textit{N.}, whom while dwelling in this world thou didst adorn with holy gifts, may ever rejoice with glory in the heavenly mansions. Through.
\subbysub{For a Father and Mother}
\lett{O}{God,} who hast bidden us to honour our father and mother: of thy mercy have compassion on the \textit{souls} of \textit{my father and mother}; forgive \textit{their} sins, and grant that \textit{I} may behold \textit{them} in the joy of eternal brightness. Through.
\subbysub{For Many Deceased}
\lett{O}{God,} whose nature and property is ever to have mercy and to forgive: have compassion on the souls of thy servants and handmaids, and grant them the remission of all their sins; that being delivered from the bonds of mortality, they may be worthy to pass over into life. Through.
\subbysub{Another Collect for Many Deceased}
\lett{G}{rant}, we beseech thee, O Lord, to the souls of thy servants and handmaids thy perpetual mercy: and let it profit them in eternity that they hoped and believed in thee. Through.
\subbysub{For a Man Departed}
\lett{I}{ncline} thine ear, O Lord, unto our prayers, wherein we humbly entreat thy mercy: that thou wouldest appoint unto the soul of thy servant \textit{N.}, which thou hast bidden to depart out of this world, a place in the land of life and peace; and wouldest make him a partaker with thy Saints. Through.
\subbysub{For a Woman Departed}
\lett{W}{e} beseech thee, O Lord, of thy loving kindness have mercy upon the soul of thine handmaid \textit{N.}: and now that she is released from the contagion of mortality, do thou restore her portion in everlasting salvation. Through.
\subbysub{Through the Year}
\lett{O}{God,} who hast made thy servants to flourish among the Ministers of Apostolic Succession in the honourable offices of Bishopric and Priesthood: grant, we beseech thee, that they may also be joined with them in a perpetual fellowship.
\needspace{4\baselineskip}
\lett{O}{God,} the Giver of pardon and the Author of man's salvation: we humbly beseech thy mercy to grant that the \textit{brethren}, kinsfolk, and benefactors of our Congregation who have departed out of this world, blessed Mary ever Virgin and all thy Saints praying for them, may attain to the fellowship of everlasting blessedness.
\needspace{4\baselineskip}
\lett{O}{God,} the Creator and Redeemer of all the faithful: grant unto the souls of thy servants and handmaids the remission of all their sins; that through devout supplications they may obtain the pardon they have always desired. Who with God the Father, in the unity of the Holy Spirit, livest and reignest God, world without end. \textit{Amen.}

\subby{Additional Collects}\label{AdditionalCollects}
\begin{rubric}
    These Collects are said for the conclusion of Antecommunion. They may be said after the Collects of Morning and Evening Prayer, at the discretion of the Minister.
\end{rubric}
\lett{O}{Lord} Jesus Christ, who saidst unto thine Apostles, Peace I leave with you, my peace I give unto you; Regard not our sins, but the faith of thy Church; and grant to it that peace and unity which is according to thy will, who livest and reignest with the Father and the Holy Ghost, one God, world without end. \textit{Amen.}
\needspace{4\baselineskip}
\lett{A}{ssist} us mercifully, O Lord, in these our supplications and prayers, and dispose the way of thy servants towards the attainment of everlasting salvation; that, among all the changes and chances of this mortal life, they may ever be defended by thy most gracious and ready help; through Jesus Christ our Lord. \textit{Amen.}
\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, Almighty God, that the words which we have heard this day with our outward ears, may, through thy grace, be so grafted inwardly in our hearts, that they may bring forth in us the fruit of good living, to the honour and praise of thy Name; through Jesus Christ our Lord. \textit{Amen.}
\needspace{4\baselineskip}
\lett{D}{irect} us, O Lord, in all our doings, with thy most gracious favour, and further us with thy continual help; that in all our works begun, continued, and ended in thee, we may glorify thy holy Name, and finally, by thy mercy, obtain everlasting life; through Jesus Christ our Lord. \textit{Amen.}
\needspace{4\baselineskip}
\lett{A}{lmighty} God, the fountain of all wisdom, who knowest our necessities before we ask, and our ignorance in asking; We beseech thee to have compassion upon our infirmities; and those things which for our unworthiness we dare not, and for our blindness we cannot ask, vouchsafe to give us, for the worthiness of thy Son Jesus Christ our Lord. \textit{Amen.}
\needspace{4\baselineskip}
\lett{A}{lmighty} God, who hast promised to hear the petitions of those who ask in thy Son's Name; We beseech thee mercifully to incline thine ears to us who have now made our prayers and supplications unto thee; and grant that those things which we have faithfully asked according to thy will, may effectually be obtained, to the relief of our necessity, and to the setting forth of thy glory; through Jesus Christ our Lord. \textit{Amen.}
%\end{multicols}
%Inspired by the English Office Noted. Translation from the 1933 English Missal.
%\lett{D}{efend} us, O Lord, we beseech thee, from all dangers of mind and body: and at the intercession of the blessed and glorious Mother of God, Mary ever Virgin, with blessed Joseph, thy blessed Apostles Peter and Paul, (blessed \textit{N.},) and all the Saints, graciously grant us health and peace; that all adversities and errors being done away, thy Church may serve thee in freedom and quietness. Through the same Jesus Christ thy Son, our Lord, who liveth and reigneth with thee and the Holy Ghost, ever one God, world without end. \textit{Amen.}

%Since the 1928 BCP has a Bidding Prayer unique to itself, clearly incorporating elements of Laudian and Andrewian Branch Theory (as happens in the Prayer for the Church excluded from this Book), the following from King Henry VIII's reign is here provided, along with further bidding for the dead from the Manuale (https://www.google.com/books/edition/Forms_of_Bidding_Prayer_with_Introductio/ptkDAAAAQAAJ?hl=en&gbpv=1&dq=Forms+of+Bidding+Prayer&printsec=frontcover).
\subby{A Bidding Prayer}
\begin{rubric}
	{To be used to preface Sermons, or on Special Occasions.}
\end{rubric}
\lett{A}{fter} a laudable custom of our Mother holy Church, ye shall kneel down, moving your hearts unto Almighty God, and making your special prayers for the three estates, concerning all Christian people, i. e. for the Spirituality, the Temporality, and the souls %being in the pains of Purgatory.
departed.
\needspace{4\baselineskip}
\lett{F}{irst} for all archbishops, and bishops, and in special for my Lord Archbishop of \textit{these United States}, your Metropolitan, and also for my Lord Bishop of this \textit{Diocese}, and in general for all parsons, vicars, and parish priests having cure of souls, with the ministers of Christ's Church, as well religious as not religious.
\needspace{4\baselineskip}
\lett{S}{econdly,} ye shall pray for the unity and peace of all Christian Realms. And for the noble Realm of \textit{America}, and for \textit{the President of these United States}; that Almighty God may send them grace, so to govern and rule the land, that it may be pleasing unto Almighty God, wealth and profit to the land, and salvation to their souls. Also ye shall pray for all those that have honoured the Church with light, lamp, vestment, or bell, or with any other ornaments, by which the service of Almighty God is the better maintained and kept. Furthermore ye shall pray for all true travellers and tillers of the earth, that truly and duly done their duty to God and holy Church, as they be bound to do.

Also ye shall pray for all manner of fruits that be done upon the ground, or shall be that Almighty God of his great pity and mercy may send such weather, that it may come to the sustenance of man. Ye shall pray also for all those that be in debt or deadly sin, that Almighty God may give them grace to come out thereof, and the sooner by our prayer. Also ye shall pray for all those that be sick or diseased, either in body or in soul, that the Almighty would send them the thing that is most profitable as well bodily as ghostly. Also ye shall pray for all pilgrims, and palmers, that have taken the way to Rome, to Saint James of Jerusalem, or to any other place: that Almighty God may give them grace to go safe, and to come safe, and give us grace to have part of their prayers, and they part of ours. Also ye shall pray for the holy Cross, that is in possession and hands of unrightful people; that God Almighty may send it into the hands of Christian people when it pleaseth him. Furthermore I commit unto your devout prayers, all women that be in our ladies' bonds: that Almighty God may send them grace, the child to receive the Sacrament of Baptism, and the mother purification. Also ye shall pray for the good man and woman, that this day giveth bread to make the holy loaf, and for all those that first began it, and them that longest continue. For these and for all true Christian people, every man and woman say the Lord's Prayer and the Angelic Salutation.
\begin{rubric}
    The Priest and the People say,
\end{rubric}
\needspace{3\baselineskip}
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil: For thine is the kingdom, and the power, and the glory, for ever and ever. Amen.
\par\noindent
\needspace{3\baselineskip}
\lett{H}{ail} Mary, full of grace; The Lord is with thee; Blessed art thou amongst women, And blessed is the fruit of thy womb, Jesus. Holy Mary, Mother of God, Pray for us sinners, now and at the hour of our death. Amen.
\begin{rubric}
    The Priest then says,
\end{rubric}
\needspace{3\baselineskip}
\lett{Y}{e} shall make a special prayer for your fathers' souls, for your mothers' souls, godfathers' souls and godmothers' souls, brothers' souls and sisters' souls, and for all your elders' souls, and for all the souls, that ye or I be bound to pray for, and specially for all the souls whose bones are buried in this Church or in this Churchyard, or in any other holy place, and especially for all the souls that bid the great mercy of Almighty God, %in the bytter peyues of Purgatory, 
that God for his great mercy release them of their pain, if it be his blessed will, and that our prayers may sum what stand them in stead; every man and woman of your charity help them with the Lord's Prayer and the Angelic Salutation.
\begin{rubric}
    The Priest and the People say the Lord's Prayer and Angelic Salutation once more.
\end{rubric}
%\needspace{3\baselineskip}
%\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil: For thine is the kingdom, and the power, and the glory, for ever and ever. Amen.
%\par\noindent
%\needspace{3\baselineskip}
%\lett{H}{ail} Mary, full of grace; The Lord is with thee; Blessed art thou amongst women, And blessed is the fruit of thy womb, Jesus. Holy Mary, Mother of God, Pray for us sinners, now and at the hour of our death. Amen.

\clearpage
\fancyhead[RO,LE]{}
\fancyhead[RE,LO]{}
\fancyhead[C]{{\LARGE Thanksgivings}}\label{thanksgiving}
\phantomsection
\addcontentsline{toc}{section}{Thanksgivings}
%\begin{rubric}
%	To be used after the General Thanksgiving, or, when that is not said, before the final Prayer of Blessing or the Benediction.
%\end{rubric}
%\begin{multicols}{2}
\subby{A Thanksgiving to Almighty God for the Fruits of the Earth and all the other Blessings of his merciful Providence}
\lett{M}{ost} gracious God, by whose knowledge the depths are broken up, and the clouds drop down the dew; We yield thee unfeigned thanks and praise for the return of seed-time and harvest, for the increase of the ground and the gathering in of the fruits thereof, and for all the other blessings of thy merciful providence bestowed upon this nation and people. And, we beseech thee, give us a just sense of these great mercies; such as may appear in our lives by an humble, holy, and obedient walking before thee all our days; through Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all glory and honour, world without end. \textit{Amen.}
\subby{The Thanksgiving of Women after Child-birth}
\begin{rubric}
	To be said when any Woman, being present in Church, shall have desired to return thanks to Almighty God for her safe deliverance.
\end{rubric}%Changed servant to handmaid
\lett{O}{almighty} God, we give thee humble thanks for that thou hast been graciously pleased to preserve, through the great pain and peril of childbirth, this woman, thy handmaid, who desireth now to offer her praises and thanksgivings unto thee. Grant, we beseech thee, most merciful Father, that she, through thy help, may both faithfully live and walk according to thy will in this life present, and also may be partaker of everlasting glory in the life to come; through Jesus Christ our Lord. \textit{Amen.}
\subby{For Rain}
\lett{O}{God,} our heavenly Father, by whose gracious providence the former and the latter rain descend upon the earth, that it may bring forth fruit for the use of man; We give thee humble thanks that it hath pleased thee to send us rain to our great comfort, and to the glory of thy holy Name; through Jesus Christ our Lord. \textit{Amen.}
%MANUAL ADJUSTMENT
\subsec{For Fair Weather}
\lett{O}{Lord} God, who hast justly humbled us by thy late visitation of us with immoderate rain and waters, and in thy mercy hast relieved and comforted our souls by this seasonable and blessed change of weather; We praise and glorify thy holy Name for this thy mercy, and will always declare thy loving-kindness from generation to generation; through Jesus Christ our Lord. \textit{Amen.}
\subby{For Peace, and Deliverance from our Enemies}
\lett{O}{almighty} God, who art a strong tower of defence unto thy servants against the face of their enemies; We yield thee praise and thanksgiving for our deliverance from those great and apparent dangers wherewith we were compassed. We acknowledge it thy goodness that we were not delivered over as a prey unto them; beseeching thee still to continue such thy mercies towards us, that all the world may know that thou art our Saviour and mighty Deliverer; through Jesus Christ our Lord. \textit{Amen.}
\subby{For Restoring Public Peace at Home}
\lett{O}{eternal} God, our heavenly Father, who alone makest men to be of one mind in a house, and stillest the outrage of a violent and unruly people; We bless thy holy Name, that it hath pleased thee to appease the seditious tumults which have been lately raised up amongst us; most humbly beseeching thee to grant to all of us grace, that we may henceforth obediently walk in thy holy commandments; and, leading a quiet and peaceable life in all godliness and honesty, may continually offer unto thee our sacrifice of praise and thanksgiving for these thy mercies towards us; through Jesus Christ our Lord. \textit{Amen.}
\subby{For Deliverance from great Sickness and Mortality}
\lett{O}{Lord} God, who hast wounded us for our sins, and consumed us for our transgressions, by thy late heavy and dreadful visitation and now, in the midst of judgment remembering mercy, hast redeemed our souls from the jaws of death; We offer unto thy fatherly goodness ourselves, our souls and bodies which thou hast delivered, to be a living sacrifice unto thee, always praising and magnifying thy mercies in the midst of thy Church; through Jesus Christ our Lord. \textit{Amen.}
\subby{For Recovery from Sickness}
\lett{O}{God,} who art the giver of life, of health, and of safety; We bless thy Name, that thou hast been pleased to deliver from \textit{his} bodily sickness this thy \textit{servant}, who now desireth to return thanks unto thee, in the presence of all thy people. Gracious art thou, O Lord, and full of compassion to the children of men. May \textit{his} heart be duly impressed with a sense of thy merciful goodness, and may \textit{he} devote the residue of \textit{his} days to an humble, holy, and obedient walking before thee; through Jesus Christ our Lord. \textit{Amen.}
\subby{For a Child's Recovery from Sickness}
\lett{A}{lmighty} God and heavenly Father, we give thee humble thanks for that thou hast been graciously pleased to deliver from \textit{his} bodily sickness the child in whose behalf we bless and praise thy Name, in the presence of all thy people. Grant, we beseech thee, O gracious Father, that \textit{he}, through thy help, may both faithfully live in this world according to thy will, and also may be partaker of everlasting glory in the life to come; through Jesus Christ our Lord. \textit{Amen.}
\subby{For a Safe Return from a Journey}
\lett{M}{ost} gracious Lord, whose mercy is over all thy works; We praise thy holy Name that thou hast been pleased to conduct in safety, through the perils of the great deep, (\textit{his} way,) this thy \textit{servant}, who now desireth to return \textit{his} thanks unto thee in thy holy Church. May \textit{he} be duly sensible of thy merciful providence towards \textit{him}, and ever express \textit{his} thankfulness by a holy trust in thee, and obedience to thy laws; through Jesus Christ our Lord. \textit{Amen.}
%\end{multicols}