\fancyhead[C]{{\LARGE Psalter}}
\psalmsec{Day 1. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 1. Morning Prayer}}

\subbypsalm{Psalm 1. \textit{Beatus vir, qui non abiit, \&c.}}
\fancyhead[RE,LO]{Psalm 1}

\lett{B}{lessed} is the man that hath not walked in the counsel of the ungodly, nor stood in the way of sinners : and hath not sat in the seat of the scornful.\par
\secondlinepsalm{But his delight is in the law of the Lord : and in his law will he exercise himself day and night.}
\thirdlinepsalm{And he shall be like a tree planted by the water-side : that will bring forth his fruit in due season.}
\psanum{4}His leaf also shall not wither : and look, whatsoever he doeth, it shall prosper.\par
\psanum{5}As for the ungodly, it is not so with them : but they are like the chaff, which the wind scattereth away from the face of the earth.\par
\psanum{6}Therefore the ungodly shall not be able to stand in the judgement : neither the sinners in the congregation of the righteous.\par
\psanum{7}But the Lord knoweth the way of the righteous : and the way of the ungodly shall perish.\par

\subbypsalm{Psalm 2. \textit{Quare fremuerunt gentes?}}
\fancyhead[RE,LO]{Psalm 2}

\lett{W}{hy} do the heathen so furiously rage together : and why do the people imagine a vain thing?\par
\secondlinepsalm{The kings of the earth stand up, and the rulers take counsel together : against the Lord, and against his Anointed.}
\thirdlinepsalm{Let us break their bonds asunder : and cast away their cords from us.}
\psanum{4}He that dwelleth in heaven shall laugh them to scorn : the Lord shall have them in derision.\par
\psanum{5}Then shall he speak unto them in his wrath : and vex them in his sore displeasure.\par
\psanum{6}Yet have I set my King : upon my holy hill of Sion.\par
\psanum{7}I will preach the law, whereof the Lord hath said unto me : Thou art my Son, this day have I begotten thee.\par
\psanum{8}Desire of me, and I shall give thee the heathen for thine inheritance : and the utmost parts of the earth for thy possession.\par
\psanum{9}Thou shalt bruise them with a rod of iron : and break them in pieces like a potter's vessel.\par
\psanum{10}Be wise now therefore, O ye kings : be learned, ye that are judges of the earth.\par
\psanum{11}Serve the Lord in fear : and rejoice unto him with reverence.\par
\psanum{12}Kiss the Son, lest he be angry, and so ye perish from the right way : if his wrath be kindled, (yea, but a little,) blessed are all they that put their trust in him.\par

\subbypsalm{Psalm 3. \textit{Domine, quid multiplicati,}}
\fancyhead[RE,LO]{Psalm 3}

\lett{L}{ord,} how are they increased that trouble me : many are they that rise against me.\par
\secondlinepsalm{Many one there be that say of my soul : There is no help for him in his God.}
\thirdlinepsalm{But thou, O Lord, art my defender : thou art my worship, and the lifter up of my head.}
\psanum{4}I did call upon the Lord with my voice : and he heard me out of his holy hill.\par
\psanum{5}I laid me down and slept, and rose up again : for the Lord sustained me.\par
\psanum{6}I will not be afraid for ten thousands of the people : that have set themselves against me round about.\par
\psanum{7}Up, Lord, and help me, O my God : for thou smitest all mine enemies upon the cheekbone; thou hast broken the teeth of the ungodly.\par
\psanum{8}Salvation belongeth unto the Lord : and thy blessing is upon thy people.\par

\subbypsalm{Psalm 4. \textit{Cum invocarem}}
\fancyhead[RE,LO]{Psalm 4}

\lett{H}{ear} me when I call, O God of my righteousness : thou hast set me at liberty when I was in trouble; have mercy upon me, and hearken unto my prayer.\par
\secondlinepsalm{O ye sons of men, how long will ye blaspheme mine honour : and have such pleasure in vanity, and seek after leasing?}
\thirdlinepsalm{Know this also, that the Lord hath chosen to himself the man that is godly : when I call upon the Lord, he will hear me.}
\psanum{4}Stand in awe, and sin not : commune with your own heart, and in your chamber, and be still.\par
\psanum{5}Offer the sacrifice of righteousness : and put your trust in the Lord.\par
\psanum{6}There be many that say : Who will shew us any good?\par
\psanum{7}Lord, lift thou up : the light of thy countenance upon us.\par
\psanum{8}Thou hast put gladness in my heart : since the time that their corn and wine and oil increased.\par
\psanum{9}I will lay me down in peace, and take my rest : for it is thou, Lord, only, that makest me dwell in safety.\par

\subbypsalm{Psalm 5. \textit{Verba mea auribus.}}
\fancyhead[RE,LO]{Psalm 5}

\lett{P}{onder} my words, O Lord : consider my meditation\par
\secondlinepsalm{O hearken thou unto the voice of my calling, my King, and my God : for unto thee will I make my prayer.}
\thirdlinepsalm{My voice shalt thou hear betimes, O Lord : early in the morning will I direct my prayer unto thee, and will look up.}
\psanum{4}For thou art the God that hast no pleasure in wickedness : neither shall any evil dwell with thee.\par
\psanum{5}Such as be foolish shall not stand in thy sight : for thou hatest all them that work vanity.\par
\psanum{6}Thou shalt destroy them that speak leasing : the Lord will abhor both the blood-thirsty and deceitful man.\par
\psanum{7}But as for me, I will come into thine house, even upon the multitude of thy mercy : and in thy fear will I worship toward thy holy temple.\par
\psanum{8}Lead me, O Lord, in thy righteousness, because of mine enemies : make thy way plain before my face.\par
\psanum{9}For there is no faithfulness in his mouth : their inward parts are very wickedness.\par
\psanum{10}Their throat is an open sepulchre : they flatter with their tongue.\par
\psanum{11}Destroy thou them, O God; let them perish through their own imaginations : cast them out in the multitude of their ungodliness; for they have rebelled against thee.\par
\psanum{12}And let all them that put their trust in thee rejoice : they shall ever be giving of thanks, because thou defendest them; they that love thy Name shall be joyful in thee;\par
\psanum{13}For thou, Lord, wilt give thy blessing unto the righteous : and with thy favourable kindness wilt thou defend him as with a shield.\par

\psalmsec{Day 1. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 1. Evening Prayer}}

\subbypsalm{Psalm 6. \textit{Domine, ne in furore}}
\fancyhead[RE,LO]{Psalm 6}

\lett{O}{Lord,} rebuke me not in thine indignation : neither chasten me in thy displeasure.\par
\secondlinepsalm{Have mercy upon me, O Lord, for I am weak : O Lord, heal me, for my bones are vexed.}
\thirdlinepsalm{My soul also is sore troubled : but, Lord, how long wilt thou punish me?}
\psanum{4}Turn thee, O Lord, and deliver my soul : O save me for thy mercy's sake.\par
\psanum{5}For in death no man remembereth thee : and who will give thee thanks in the pit?\par
\psanum{6}I am weary of my groaning; every night wash I my bed : and water my couch with my tears.\par
\psanum{7}My beauty is gone for very trouble : and worn away because of all mine enemies.\par
\psanum{8}Away from me, all ye that work vanity : for the Lord hath heard the voice of my weeping.\par
\psanum{9}The Lord hath heard my petition : the Lord will receive my prayer.\par
\psanum{10}All mine enemies shall be confounded, and sore vexed : they shall be turned back, and put to shame suddenly.\par

\subbypsalm{Psalm 7. \textit{Domine, Deus meus}}
\fancyhead[RE,LO]{Psalm 7}

\lett{O}{Lord} my God, in thee have I put my trust : save me from all them that persecute me, and deliver me;\par
\secondlinepsalm{Lest he devour my soul, like a lion, and tear it in pieces : while there is none to help.}
\thirdlinepsalm{O Lord my God, if I have done any such thing : or if there be any wickedness in my hands;}
\psanum{4}If I have rewarded evil unto him that dealt friendly with me : yea, I have delivered him that without any cause is mine enemy,\par
\psanum{5}Then let mine enemy persecute my soul, and take me : yea, let him tread my life down upon the earth, and lay mine honour in the dust.\par
\psanum{6}Stand up, O Lord, in thy wrath, and lift up thyself, because of the indignation of mine enemies : arise up for me in the judgement that thou hast commanded.\par
\psanum{7}And so shall the congregation of the people come about thee : for their sakes therefore lift up thyself again.\par
\psanum{8}The Lord shall judge the people; give sentence with me, O Lord : according to my righteousness, and according to the innocency that is in me.\par
\psanum{9}O let the wickedness of the ungodly come to an end : but guide thou the just.\par
\psanum{10}For the righteous God : trieth the very hearts and reins.\par
\psanum{11}My help cometh of God : who preserveth them that are true of heart.\par
\psanum{12}God is a righteous Judge, strong and patient : and God is provoked every day.\par
\psanum{13}If a man will not turn, he will whet his sword : he hath bent his bow, and made it ready.\par
\psanum{14}He hath prepared for him the instruments of death : he ordaineth his arrows against the persecutors.\par
\psanum{15}Behold, he travaileth with mischief : he hath conceived sorrow, and brought forth ungodliness.\par
\psanum{16}He hath graven and digged up a pit : and is fallen on himself into the destruction that he made for other.\par
\psanum{17}For his travail shall come upon his own head : and his wickedness shall fall on his own pate.\par
\psanum{18}I will give thanks unto the Lord, according to his righteousness : and I will praise the Name of the Lord most High.\par

\subbypsalm{Psalm 8. \textit{Domine, Dominus noster}}
\fancyhead[RE,LO]{Psalm 8}

\lett{O}{Lord} our Governor, how excellent is thy Name in all the world : thou that hast set thy glory above the heavens!\par
\secondlinepsalm{Out of the mouth of very babes and sucklings hast thou ordained strength, because of thine enemies : that thou mightest still the enemy and the avenger.}
\thirdlinepsalm{For I will consider thy heavens, even the works of thy fingers : the moon and the stars, which thou hast ordained.}
\psanum{4}What is man, that thou art mindful of him : and the son of man, that thou visitest him?\par
\psanum{5}Thou madest him lower than the angels : to crown him with glory and worship.\par
\psanum{6}Thou makest him to have dominion of the works of thy hands : and thou hast put all things in subjection under his feet;\par
\psanum{7}All sheep and oxen : yea, and the beasts of the field;\par
\psanum{8}The fowls of the air, and the fishes of the sea : and whatsoever walketh through the paths of the seas.\par
\psanum{9}O Lord our Governor : how excellent is thy Name in all the world!\par

\psalmsec{Day 2. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 2. Morning Prayer}}

%MANUAL ADJUSTMENT:
\subsec{Psalm 9. \textit{Confitebor tibi}}
\fancyhead[RE,LO]{Psalm 9}

\lett{I}{will} give thanks unto thee, O Lord, with my whole heart : I will speak of all thy marvellous works.\par
\secondlinepsalm{I will be glad and rejoice in thee : yea, my songs will I make of thy Name, O thou most Highest.}
\thirdlinepsalm{While mine enemies are driven back : they shall fall and perish at thy presence.}
\psanum{4}For thou hast maintained my right and my cause : thou art set in the throne that judgest right.\par
\psanum{5}Thou hast rebuked the heathen, and destroyed the ungodly : thou hast put out their name for ever and ever.\par
\psanum{6}O thou enemy, destructions are come to a perpetual end : even as the cities which thou hast destroyed, their memorial is perished with them.\par
\psanum{7}But the Lord shall endure for ever : he hath also prepared his seat for judgement.\par
\psanum{8}For he shall judge the world in righteousness : and minister true judgement unto the people.\par
\psanum{9}The Lord also will be a defence for the oppressed : even a refuge in due time of trouble.\par
\psanum{10}And they that know thy Name will put their trust in thee : for thou, Lord, hast never failed them that seek thee.\par
\psanum{11}O praise the Lord which dwelleth in Sion : shew the people of his doings.\par
\psanum{12}For when he maketh inquisition for blood, he remembereth them : and forgetteth not the complaint of the poor.\par
\psanum{13}Have mercy upon me, O Lord; consider the trouble which I suffer of them that hate me : thou that liftest me up from the gates of death.\par
\psanum{14}That I may shew all thy praises within the ports of the daughter of Sion : I will rejoice in thy salvation.\par
\psanum{15}The heathen are sunk down in the pit that they made : in the same net which they hid privily, is their foot taken.\par
\psanum{16}The Lord is known to execute judgement : the ungodly is trapped in the work of his own hands.\par
\psanum{17}The wicked shall be turned into hell : and all the people that forget God.\par
\psanum{18}For the poor shall not alway be forgotten : the patient abiding of the meek shall not perish for ever.\par
\psanum{19}Up, Lord, and let not man have the upper hand : let the heathen be judged in thy sight.\par
\psanum{20}Put them in fear, O Lord : that the heathen may know themselves to be but men.\par

\subbypsalm{Psalm 10. \textit{Ut quid, Domine?}}

\fancyhead[RE,LO]{Psalm 10}

\lett{W}{hy} standest thou so far off, O Lord : and hidest thy face in the needful time of trouble?\par
\secondlinepsalm{The ungodly for his own lust doth persecute the poor : let them be taken in the crafty wiliness that they have imagined.}
\thirdlinepsalm{For the ungodly hath made boast of his own heart's desire : and speaketh good of the covetous, whom God abhorreth.}
\psanum{4}The ungodly is so proud, that he careth not for God : neither is God in all his thoughts.\par
\psanum{5}His ways are alway grievous : thy judgements are far above out of his sight, and therefore defieth he all his enemies.\par
\psanum{6}For he hath said in his heart, Tush, I shall never be cast down : there shall no harm happen unto me.\par
\psanum{7}His mouth is full of cursing, deceit, and fraud : under his tongue is ungodliness and vanity.\par
\psanum{8}He sitteth lurking in the thievish corners of the streets : and privily in his lurking dens doth he murder the innocent; his eyes are set against the poor.\par
\psanum{9}For he lieth waiting secretly, even as a lion lurketh he in his den : that he may ravish the poor.\par
\psanum{10}He doth ravish the poor : when he getteth him into his net.\par
\psanum{11}He falleth down, and humbleth himself : that the congregation of the poor may fall into the hands of his captains.\par
\psanum{12}He hath said in his heart, Tush, God hath forgotten : he hideth away his face, and he will never see it.\par
\psanum{13}Arise, O Lord God, and lift up thine hand : forget not the poor.\par
\psanum{14}Wherefore should the wicked blaspheme God : while he doth say in his heart, Tush, thou God carest not for it.\par
\psanum{15}Surely thou hast seen it : for thou beholdest ungodliness and wrong.\par
\psanum{16}That thou mayest take the matter into thy hand : the poor committeth himself unto thee; for thou art the helper of the friendless.\par
\psanum{17}Break thou the power of the ungodly and malicious : take away his ungodliness, and thou shalt find none.\par
\psanum{18}The Lord is King for ever and ever : and the heathen are perished out of the land.\par
\psanum{19}Lord, thou hast heard the desire of the poor : thou preparest their heart, and thine ear hearkeneth thereto;\par
\psanum{20}To help the fatherless and poor unto their right : that the man of the earth be no more exalted against them.\par

\subbypsalm{Psalm 11. \textit{In Domino confido}}
\fancyhead[RE,LO]{Psalm 11}

\lett{I}{n} the Lord put I my trust : how say ye then to my soul, that she should flee as a bird unto the hill?\par
\secondlinepsalm{For lo, the ungodly bend their bow, and make ready their arrows within the quiver : that they may privily shoot at them which are true of heart.}
\thirdlinepsalm{For the foundations will be cast down : and what hath the righteous done?}
\psanum{4}The Lord is in his holy temple : the Lord's seat is in heaven.\par
\psanum{5}His eyes consider the poor : and his eye-lids try the children of men.\par
\psanum{6}The Lord alloweth the righteous : but the ungodly, and him that delighteth in wickedness, doth his soul abhor.\par
\psanum{7}Upon the ungodly he shall rain snares, fire and brimstone, storm and tempest : this shall be their portion to drink.\par
\psanum{8}For the righteous Lord loveth righteousness : his countenance will behold the thing that is just.\par

\psalmsec{Day 2. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 2. Evening Prayer}}

\subbypsalm{Psalm 12. \textit{Salvum me fac}}
\fancyhead[RE,LO]{Psalm 12}

\lett{H}{elp} me, Lord, for there is not one godly man left : for the faithful are minished from among the children of men,\par
\secondlinepsalm{They talk of vanity every one with his neighbour : they do but flatter with their lips, and dissemble in their double heart.}
\thirdlinepsalm{The Lord shall root out all deceitful lips : and the tongue that speaketh proud things;}
\psanum{4}Which have said, With our tongue will we prevail : we are they that ought to speak, who is lord over us?\par
\psanum{5}Now for the comfortless trouble's sake of the needy : and because of the deep sighing of the poor,\par
\psanum{6}I will up, saith the Lord : and will help every one from him that swelleth against him, and will set him at rest.\par
\psanum{7}The words of the Lord are pure words : even as the silver, which from the earth is tried, and purified seven times in the fire.\par
\psanum{8}Thou shalt keep them, O Lord : thou shalt preserve him from this generation for ever.\par
\psanum{9}The ungodly walk on every side : when they are exalted, the children of men are put to rebuke.\par

\subbypsalm{Psalm 13. \textit{Usque quo, Domine?}}
\fancyhead[RE,LO]{Psalm 13}

\lett{H}{ow} long wilt thou forget me, O Lord, for ever : how long wilt thou hide thy face from me?\par
\secondlinepsalm{How long shall I seek counsel in my soul, and be so vexed in my heart : how long shall mine enemies triumph over me?}
\thirdlinepsalm{Consider, and hear me, O Lord my God : lighten mine eyes, that I sleep not in death.}
\psanum{4}Lest mine enemy say, I have prevailed against him : for if I be cast down, they that trouble me will rejoice at it.\par
\psanum{5}But my trust is in thy mercy : and my heart is joyful in thy salvation.\par
\psanum{6}I will sing of the Lord, because he hath dealt so lovingly with me : yea, I will praise the Name of the Lord most Highest.\par

\subbypsalm{Psalm 14. \textit{Dixit insipiens}}
\fancyhead[RE,LO]{Psalm 14}

\lett{T}{he} fool hath said in his heart : There is no God.\par
\secondlinepsalm{They are corrupt, and become abominable in their doings : there is none that doeth good, no not one.}
\thirdlinepsalm{The Lord looked down from heaven upon the children of men : to see if there were any that would understand, and seek after God.}
\psanum{4}But they are all gone out of the way, they are altogether become abominable : there is none that doeth good, no not one.\par
\psanum{5}Their throat is an open sepulchre, with their tongues have they deceived : the poison of asps is under their lips.\par
\psanum{6}Their mouth is full of cursing and bitterness : their feet are swift to shed blood.\par
\psanum{7}Destruction and unhappiness is in their ways, and the way of peace have they not known : there is no fear of God before their eyes.\par
\psanum{8}Have they no knowledge, that they are all such workers of mischief : eating up my people as it were bread, and call not upon the Lord?\par
\psanum{9}There were they brought in great fear, even where no fear was : for God is in the generation of the righteous.\par
\psanum{10}As for you, ye have made a mock at the counsel of the poor : because he putteth his trust in the Lord.\par
\psanum{11}Who shall give salvation unto Israel out of Sion? When the Lord turneth the captivity of his people : then shall Jacob rejoice, and Israel shall be glad.\par

\psalmsec{Day 3. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 3. Morning Prayer}}

\subbypsalm{Psalm 15. \textit{Domine, quis habitabit?}}
\fancyhead[RE,LO]{Psalm 15}

\lett{L}{ord,} who shall dwell in thy tabernacle : or who shall rest upon thy holy hill?\par
\secondlinepsalm{Even he that leadeth an uncorrupt life : and doeth the thing which is right, and speaketh the truth from his heart.}
\thirdlinepsalm{He that hath used no deceit in his tongue, nor done evil to his neighbour : and hath not slandered his neighbour.}
\psanum{4}He that setteth not by himself, but is lowly in his own eyes : and maketh much of them that fear the Lord.\par
\psanum{5}He that sweareth unto his neighbour, and disappointeth him not : though it were to his own hindrance.\par
\psanum{6}He that hath not given his money upon usury : nor taken reward against the innocent.\par
\psanum{7}Whoso doeth these things : shall never fall.\par

\subbypsalm{Psalm 16. \textit{Conserva me, Domine}}
\fancyhead[RE,LO]{Psalm 16}

\lett{P}{reserve} me, O God : for in thee have I put my trust.\par
\secondlinepsalm{O my soul, thou hast said unto the Lord : Thou art my God, my goods are nothing unto thee.}
\thirdlinepsalm{All my delight is upon the saints, that are in the earth : and upon such as excel in virtue.}
\psanum{4}But they that run after another god : shall have great trouble.\par
\psanum{5}Their drink-offerings of blood will I not offer : neither make mention of their names within my lips.\par
\psanum{6}The Lord himself is the portion of mine inheritance, and of my cup : thou shalt maintain my lot.\par
\psanum{7}The lot is fallen unto me in a fair ground : yea, I have a goodly heritage.\par
\psanum{8}I will thank the Lord for giving me warning : my reins also chasten me in the night-season.\par
\psanum{9}I have set God always before me : for he is on my right hand, therefore I shall not fall.\par
\psanum{10}Wherefore my heart was glad, and my glory rejoiced : my flesh also shall rest in hope.\par
\psanum{11}For why? thou shalt not leave my soul in hell : neither shalt thou suffer thy Holy One to see corruption.\par
\psanum{12}Thou shalt shew me the path of life; in thy presence is the fulness of joy : and at thy right hand there is pleasure for evermore.\par

\subbypsalm{Psalm 17. \textit{Exaudi, Domine}}
\fancyhead[RE,LO]{Psalm 17}

\lett{H}{ear} the right, O Lord, consider my complaint : and hearken unto my prayer, that goeth not out of feigned lips.\par
\secondlinepsalm{Let my sentence come forth from thy presence : and let thine eyes look upon the thing that is equal.}
\thirdlinepsalm{Thou hast proved and visited mine heart in the night-season; thou hast tried me, and shalt find no wickedness in me : for I am utterly purposed that my mouth shall not offend.}
\psanum{4}Because of men's works, that are done against the words of thy lips : I have kept me from the ways of the destroyer.\par
\psanum{5}O hold thou up my goings in thy paths : that my footsteps slip not.\par
\psanum{6}I have called upon thee, O God, for thou shalt hear me : incline thine ear to me, and hearken unto my words.\par
\psanum{7}Shew thy marvellous loving-kindness, thou that art the Saviour of them which put their trust in thee : from such as resist thy right hand.\par
\psanum{8}Keep me as the apple of an eye : hide me under the shadow of thy wings.\par
\psanum{9}From the ungodly that trouble me : mine enemies compass me round about to take away my soul.\par
\psanum{10}They are inclosed in their own fat : and their mouth speaketh proud things.\par
\psanum{11}They lie waiting in our way on every side : turning their eyes down to the ground.\par
\psanum{12}Like as a lion that is greedy of his prey : and as it were a lion's whelp, lurking in secret places.\par
\psanum{13}Up, Lord, disappoint him, and cast him down : deliver my soul from the ungodly, which is a sword of thine;\par
\psanum{14}From the men of thy hand, O Lord, from the men, I say, and from the evil world : which have their portion in this life, whose bellies thou fillest with thy hid treasure.\par
\psanum{15}They have children at their desire : and leave the rest of their substance for their babes.\par
\psanum{16}But as for me, I will behold thy presence in righteousness : and when I awake up after thy likeness, I shall be satisfied with it.\par

\psalmsec{Day 3. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 3. Evening Prayer}}

\subbypsalm{Psalm 18. \textit{Diligam te, Domine}}
\fancyhead[RE,LO]{Psalm 18}

\lett{I}{will} love thee, O Lord, my strength; the Lord is my stony rock, and my defence : my saviour, my God, and my might, in whom I will trust, my buckler, the horn also of my salvation, and my refuge.\par
\secondlinepsalm{I will call upon the Lord, which is worthy to be praised : so shall I be safe from mine enemies.}
\thirdlinepsalm{The sorrows of death compassed me : and the overflowings of ungodliness made me afraid.}
\psanum{4}The pains of hell came about me : the snares of death overtook me.\par
\psanum{5}In my trouble I will call upon the Lord : and complain unto my God.\par
\psanum{6}So shall he hear my voice out of his holy temple : and my complaint shall come before him, it shall enter even into his ears.\par
\psanum{7}The earth trembled and quaked : the very foundations also of the hills shook, and were removed, because he was wroth.\par
\psanum{8}There went a smoke out in his presence : and a consuming fire out of his mouth, so that coals were kindled at it.\par
\psanum{9}He bowed the heavens also, and came down : and it was dark under his feet.\par
\psanum{10}He rode upon the cherubins, and did fly : he came flying upon the wings of the wind.\par
\psanum{11}He made darkness his secret place : his pavilion round about him, with dark water and thick clouds to cover him.\par
\psanum{12}At the brightness of his presence his clouds removed : hail-stones, and coals of fire.\par
\psanum{13}The Lord also thundered out of heaven, and the Highest gave his thunder : hail-stones, and coals of fire.\par
\psanum{14}He sent out his arrows, and scattered them : he cast forth lightnings, and destroyed them.\par
\psanum{15}The springs of water were seen, and the foundations of the round world were discovered, at thy chiding, O Lord : at the blasting of the breath of thy displeasure.\par
\psanum{16}He shall send down from on high to fetch me : and shall take me out of many waters.\par
\psanum{17}He shall deliver me from my strongest enemy, and from them which hate me : for they are too mighty for me.\par
\psanum{18}They prevented me in the day of my trouble : but the Lord was my upholder.\par
\psanum{19}He brought me forth also into a place of liberty : he brought me forth, even because he had a favour unto me.\par
\psanum{20}The Lord shall reward me after my righteous dealing : according to the cleanness of my hands shall he recompense me.\par
\psanum{21}Because I have kept the ways of the Lord : and have not forsaken my God, as the wicked doth.\par
\psanum{22}For I have an eye unto all his laws : and will not cast out his commandments from me.\par
\psanum{23}I was also uncorrupt before him : and eschewed mine own wickedness.\par
\psanum{24}Therefore shall the Lord reward me after my righteous dealing : and according unto the cleanness of my hands in his eye-sight.\par
\psanum{25}With the holy thou shalt be holy : and with a perfect man thou shalt be perfect.\par
\psanum{26}With the clean thou shalt be clean : and with the froward thou shalt learn frowardness.\par
\psanum{27}For thou shalt save the people that are in adversity : and shalt bring down the high looks of the proud.\par
\psanum{28}Thou also shalt light my candle : the Lord my God shall make my darkness to be light.\par
\psanum{29}For in thee I shall discomfit an host of men : and with the help of my God I shall leap over the wall.\par
\psanum{30}The way of God is an undefiled way : the word of the Lord also is tried in the fire; he is the defender of all them that put their trust in him.\par
\psanum{31}For who is God, but the Lord : or who hath any strength, except our God?\par
\psanum{32}It is God, that girdeth me with strength of war : and maketh my way perfect.\par
\psanum{33}He maketh my feet like harts' feet : and setteth me up on high.\par
\psanum{34}He teacheth mine hands to fight : and mine arms shall break even a bow of steel.\par
\psanum{35}Thou hast given me the defence of thy salvation : thy right hand also shall hold me up, and thy loving correction shall make me great.\par
\psanum{36}Thou shalt make room enough under me for to go : that my footsteps shall not slide.\par
\psanum{37}I will follow upon mine enemies, and overtake them : neither will I turn again till I have destroyed them.\par
\psanum{38}I will smite them, that they shall not be able to stand : but fall under my feet.\par
\psanum{39}Thou hast girded me with strength unto the battle : thou shalt throw down mine enemies under me.\par
\psanum{40}Thou hast made mine enemies also to turn their backs upon me : and I shall destroy them that hate me.\par
\psanum{41}They shall cry, but there shall be none to help them : yea, even unto the Lord shall they cry, but he shall not hear them.\par
\psanum{42}I will beat them as small as the dust before the wind : I will cast them out as the clay in the streets.\par
\psanum{43}Thou shalt deliver me from the strivings of the people : and thou shalt make me the head of the heathen.\par
\psanum{44}A people whom I have not known : shall serve me.\par
\psanum{45}As soon as they hear of me, they shall obey me : but the strange children shall dissemble with me.\par
\psanum{46}The strange children shall fail : and be afraid out of their prisons.\par
\psanum{47}The Lord liveth, and blessed be my strong helper : and praised be the Lord of my salvation;\par
\psanum{48}Even the God that seeth that I be avenged : and subdueth the people unto me.\par
\psanum{49}It is he that delivereth me from my cruel enemies, and setteth me up above mine adversaries : thou shalt rid me from the wicked man.\par
\psanum{50}For this cause will I give thanks unto thee, O Lord, among the Gentiles : and sing praises unto thy Name.\par
\psanum{51}Great prosperity giveth he unto his King : and sheweth loving-kindness unto David his Anointed, and unto his seed for evermore.\par

\psalmsec{Day 4. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 4. Morning Prayer}}

\subbypsalm{Psalm 19. \textit{Caeli enarrant}}
\fancyhead[RE,LO]{Psalm 19}

\lett{T}{he} heavens declare the glory of God : and the firmament sheweth his handywork.\par
\secondlinepsalm{One day telleth another : and one night certifieth another.}
\thirdlinepsalm{There is neither speech nor language : but their voices are heard among them.}
\psanum{4}Their sound is gone out into all lands : and their words into the ends of the world.\par
\psanum{5}In them hath he set a tabernacle for the sun : which cometh forth as a bridegroom out of his chamber, and rejoiceth as a giant to run his course.\par
\psanum{6}It goeth forth from the uttermost part of the heaven, and runneth about unto the end of it again : and there is nothing hid from the heat thereof.\par
\psanum{7}The law of the Lord is an undefiled law, converting the soul : the testimony of the Lord is sure, and giveth wisdom unto the simple.\par
\psanum{8}The statutes of the Lord are right, and rejoice the heart : the commandment of the Lord is pure, and giveth light unto the eyes.\par
\psanum{9}The fear of the Lord is clean, and endureth for ever : the judgements of the Lord are true, and righteous altogether.\par
\psanum{10}More to be desired are they than gold, yea, than much fine gold : sweeter also than honey, and the honey-comb.\par
\psanum{11}Moreover, by them is thy servant taught : and in keeping of them there is great reward.\par
\psanum{12}Who can tell how oft he offendeth : O cleanse thou me from my secret faults.\par
\psanum{13}Keep thy servant also from presumptuous sins, lest they get the dominion over me : so shall I be undefiled, and innocent from the great offence.\par
\psanum{14}Let the words of my mouth, and the meditation of my heart : be alway acceptable in thy sight,\par
\psanum{15}O Lord : my strength, and my redeemer.\par

\subbypsalm{Psalm 20. \textit{Exaudiat te Dominus}}
\fancyhead[RE,LO]{Psalm 20}

\lett{T}{he} Lord hear thee in the day of trouble : the Name of the God of Jacob defend thee;\par
\secondlinepsalm{Send thee help from the sanctuary : and strengthen thee out of Sion;}
\thirdlinepsalm{Remember all thy offerings : and accept thy burnt-sacrifice;}
\psanum{4}Grant thee thy heart's desire : and fulfil all thy mind.\par
\psanum{5}We will rejoice in thy salvation, and triumph in the Name of the Lord our God : the Lord perform all thy petitions.\par
\psanum{6}Now know I that the Lord helpeth his Anointed, and will hear him from his holy heaven : even with the wholesome strength of his right hand.\par
\psanum{7}Some put their trust in chariots, and some in horses : but we will remember the Name of the Lord our God.\par
\psanum{8}They are brought down, and fallen : but we are risen, and stand upright.\par
\psanum{9}Save, Lord, and hear us, O King of heaven : when we call upon thee.\par

\subbypsalm{Psalm 21. \textit{Domine, in virtute tua}}
\fancyhead[RE,LO]{Psalm 21}

\lett{T}{he} King shall rejoice in thy strength, O Lord : exceeding glad shall he be of thy salvation.\par
\secondlinepsalm{Thou hast given him his heart's desire : and hast not denied him the request of his lips.}
\thirdlinepsalm{For thou shalt prevent him with the blessings of goodness : and shalt set a crown of pure gold upon his head.}
\psanum{4}He asked life of thee, and thou gavest him a long life : even for ever and ever.\par
\psanum{5}His honour is great in thy salvation : glory and great worship shalt thou lay upon him.\par
\psanum{6}For thou shalt give him everlasting felicity : and make him glad with the joy of thy countenance.\par
\psanum{7}And why? because the King putteth his trust in the Lord : and in the mercy of the most Highest he shall not miscarry.\par
\psanum{8}All thine enemies shall feel thine hand : thy right hand shall find out them that hate thee.\par
\psanum{9}Thou shalt make them like a fiery oven in time of thy wrath : the Lord shall destroy them in his displeasure, and the fire shall consume them.\par
\psanum{10}Their fruit shalt thou root out of the earth : and their seed from among the children of men.\par
\psanum{11}For they intended mischief against thee : and imagined such a device as they are not able to perform.\par
\psanum{12}Therefore shalt thou put them to flight : and the strings of thy bow shalt thou make ready against the face of them.\par
\psanum{13}Be thou exalted, Lord, in thine own strength : so we will sing, and praise thy power.\par

\psalmsec{Day 4. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 4. Evening Prayer}}

\subbypsalm{Psalm 22. \textit{Deus, Deus meus}}
\fancyhead[RE,LO]{Psalm 22}

\lett{M}{y} God, my God, look upon me; why hast thou forsaken me : and art so far from my health, and from the words of my complaint?\par
\secondlinepsalm{O my God, I cry in the day-time, but thou hearest not : and in the night-season also I take no rest.}
\thirdlinepsalm{And thou continuest holy : O thou worship of Israel.}
\psanum{4}Our fathers hoped in thee : they trusted in thee, and thou didst deliver them.\par
\psanum{5}They called upon thee, and were holpen : they put their trust in thee, and were not confounded.\par
\psanum{6}But as for me, I am a worm, and no man : a very scorn of men, and the outcast of the people.\par
\psanum{7}All they that see me laugh me to scorn : they shoot our their lips, and shake their heads, saying,\par
\psanum{8}He trusted in God, that he would deliver him : let him deliver him, if he will have him.\par
\psanum{9}But thou art he that took me out of my mother's womb : thou wast my hope, when I hanged yet upon my mother's breasts.\par
\psanum{10}I have been left unto thee ever since I was born : thou art my God, even from my mother's womb.\par
\psanum{11}O go not from me, for trouble is hard at hand : and there is none to help me.\par
\psanum{12}Many oxen are come about me : fat bulls of Basan close me in on every side.\par
\psanum{13}They gape upon me with their mouths : as it were a ramping and a roaring lion.\par
\psanum{14}I am poured out like water, and all my bones are out of joint : my heart also in the midst of my body is even like melting wax.\par
\psanum{15}My strength is dried up like a potsherd, and my tongue cleaveth to my gums : and thou shalt bring me into the dust of death.\par
\psanum{16}For many dogs are come about me : and the council of the wicked layeth siege against me.\par
\psanum{17}They pierced my hands and my feet; I may tell all my bones : they stand staring and looking upon me.\par
\psanum{18}They part my garments among them : and casts lots upon my vesture.\par
\psanum{19}But be not thou far from me, O Lord : thou art my succour, haste thee to help me.\par
\psanum{20}Deliver my soul from the sword : my darling from the power of the dog.\par
\psanum{21}Save me from the lion's mouth : thou hast heard me also from among the horns of the unicorns.\par
\psanum{22}I will declare thy Name unto my brethren : in the midst of the congregation will I praise thee.\par
\psanum{23}O praise the Lord, ye that fear him : magnify him, all ye of the seed of Jacob, and fear him, all ye seed of Israel.\par
\psanum{24}For he hath not despised, nor abhorred, the low estate of the poor : he hath not hid his face from him, but when he called unto him he heard him.\par
\psanum{25}My praise is of thee in the great congregation : my vows will I perform in the sight of them that fear him.\par
\psanum{26}The poor shall eat and be satisfied : they that seek after the Lord shall praise him; your heart shall live for ever.\par
\psanum{27}All the ends of the world shall remember themselves, and be turned unto the Lord : and all the kindreds of the nations shall worship before him.\par
\psanum{28}For the kingdom is the Lord's : and he is the Governor among the people.\par
\psanum{29}All such as be fat upon earth : have eaten and worshipped.\par
\psanum{30}All they that go down into the dust shall kneel before him : and no man hath quickened his own soul.\par
\psanum{31}My seed shall serve him : they shall be counted unto the Lord for a generation.\par
\psanum{32}They shall come, and the heavens shall declare his righteousness : unto a people that shall be born, whom the Lord hath made.\par

\subbypsalm{Psalm 23. \textit{Dominus regit me.}}
\fancyhead[RE,LO]{Psalm 23}

\lett{T}{he} Lord is my shepherd : therefore can I lack nothing.\par
\secondlinepsalm{He shall feed me in a green pasture : and lead me forth beside the waters of comfort.}
\thirdlinepsalm{He shall convert my soul : and bring me forth in the paths of righteousness, for his Name's sake.}
\psanum{4}Yea, though I walk through the valley of the shadow of death, I will fear no evil : for thou art with me; thy rod and thy staff comfort me.\par
\psanum{5}Thou shalt prepare a table before me against them that trouble me : thou hast anointed my head with oil, and my cup shall be full.\par
\psanum{6}But thy loving-kindness and mercy shall follow me all the days of my life : and I will dwell in the house of the Lord for ever.\par

\psalmsec{Day 5. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 5. Morning Prayer}}

\subbypsalm{Psalm 24. \textit{Domini est terra}}
\fancyhead[RE,LO]{Psalm 24}

\lett{T}{he} earth is the Lord's, and all that therein is : the compass of the world, and they that dwell therein.\par
\secondlinepsalm{For he hath founded it upon the seas : and prepared it upon the floods.}
\thirdlinepsalm{Who shall ascend into the hill of the Lord : or who shall rise up in his holy place?}
\psanum{4}Even he that hath clean hands, and a pure heart : and that hath not lift up his mind unto vanity, nor sworn to deceive his neighbour.\par
\psanum{5}He shall receive the blessing from the Lord : and righteousness from the God of his salvation.\par
\psanum{6}This is the generation of them that seek him : even of them that seek thy face, O Jacob.\par
\psanum{7}Lift up your heads, O ye gates, and be ye lift up, ye everlasting doors : and the King of glory shall come in.\par
\psanum{8}Who is the King of glory : it is the Lord strong and mighty, even the Lord mighty in battle.\par
\psanum{9}Lift up your heads, O ye gates, and be ye lift up, ye everlasting doors : and the King of glory shall come in.\par
\psanum{10}Who is the King of glory : even the Lord of hosts, he is the King of glory.\par

\subbypsalm{Psalm 25. \textit{Ad te, Domine, levavi}}
\fancyhead[RE,LO]{Psalm 25}

\lett{U}{nto} thee, O Lord, will I lift up my soul; my God, I have put my trust in thee : O let me not be confounded, neither let mine enemies triumph over me.\par
\secondlinepsalm{For all they that hope in thee shall not be ashamed : but such as transgress without a cause shall be put to confusion.}
\thirdlinepsalm{Shew me thy ways, O Lord : and teach me thy paths.}
\psanum{4}Lead me forth in thy truth, and learn me : for thou art the God of my salvation; in thee hath been my hope all the day long.\par
\psanum{5}Call to remembrance, O Lord, thy tender mercies : and thy loving-kindnesses, which have been ever of old.\par
\psanum{6}O remember not the sins and offences of my youth : but according to thy mercy think thou upon me, O Lord, for thy goodness.\par
\psanum{7}Gracious and righteous is the Lord : therefore will he teach sinners in the way.\par
\psanum{8}Them that are meek shall he guide in judgement : and such as are gentle, them shall he learn his way.\par
\psanum{9}All the paths of the Lord are mercy and truth : unto such as keep his covenant and his testimonies.\par
\psanum{10}For thy Name's sake, O Lord : be merciful unto my sin, for it is great.\par
\psanum{11}What man is he that feareth the Lord : him shall he teach in the way that he shall choose.\par
\psanum{12}His soul shall dwell at ease : and his seed shall inherit the land.\par
\psanum{13}The secret of the Lord is among them that fear him : and he will shew them his covenant.\par
\psanum{14}Mine eyes are ever looking unto the Lord : for he shall pluck my feet out of the net.\par
\psanum{15}Turn thee unto me, and have mercy upon me : for I am desolate and in misery.\par
\psanum{16}The sorrows of my heart are enlarged : O bring thou me out of my troubles.\par
\psanum{17}Look upon my adversity and misery : and forgive me all my sin.\par
\psanum{18}Consider mine enemies, how many they are : and they bear a tyrannous hate against me.\par
\psanum{19}O keep my soul, and deliver me : let me not be confounded, for I have put my trust in thee.\par
\psanum{20}Let perfectness and righteous dealing wait upon me : for my hope hath been in thee.\par
\psanum{21}Deliver Israel, O God : out of all his troubles.\par

\subbypsalm{Psalm 26. \textit{Judica me, Domine}}
\fancyhead[RE,LO]{Psalm 26}

\lett{B}{e} thou my judge, O Lord, for I have walked innocently : my trust hath been also in the Lord, therefore shall I not fall.\par
\secondlinepsalm{Examine me, O Lord, and prove me : try out my reins and my heart.}
\thirdlinepsalm{For thy loving-kindness is ever before mine eyes : and I will walk in thy truth.}
\psanum{4}I have not dwelt with vain persons : neither will I have fellowship with the deceitful.\par
\psanum{5}I have hated the congregation of the wicked : and will not sit among the ungodly.\par
\psanum{6}I will wash my hands in innocency, O Lord : and so will I go to thine altar.\par
\psanum{7}That I may shew the voice of thanksgiving : and tell of all thy wondrous works.\par
\psanum{8}Lord, I have loved the habitation of thy house : and the place where thine honour dwelleth.\par
\psanum{9}O shut not up my soul with the sinners : nor my life with the blood-thirsty.\par
\psanum{10}In whose hands is wickedness : and their right hand is full of gifts.\par
\psanum{11}But as for me, I will walk innocently : O deliver me, and be merciful unto me.\par
\psanum{12}My foot standeth right : I will praise the Lord in the congregations.\par

\psalmsec{Day 5. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 5. Evening Prayer}}

\subbypsalm{Psalm 27. \textit{Dominus illuminatio}}
\fancyhead[RE,LO]{Psalm 27}

\lett{T}{he} Lord is my light and my salvation; whom then shall I fear : the Lord is the strength of my life; of whom then shall I be afraid?\par
\secondlinepsalm{When the wicked, even mine enemies and my foes, came upon me to eat up my flesh : they stumbled and fell.}
\thirdlinepsalm{Though an host of men were laid against me, yet shall not my heart be afraid : and though there rose up war against me, yet will I put my trust in him.}
\psanum{4}One thing have I desired of the Lord, which I will require : even that I may dwell in the house of the Lord all the days of my life, to behold the fair beauty of the Lord, and to visit his temple.\par
\psanum{5}For in the time of trouble he shall hide me in his tabernacle : yea, in the secret place of his dwelling shall he hide me, and set me up upon a rock of stone.\par
\psanum{6}And now shall he lift up mine head : above mine enemies round about me.\par
\psanum{7}Therefore will I offer in his dwelling an oblation with great gladness : I will sing, and speak praises unto the Lord.\par
\psanum{8}Hearken unto my voice, O Lord, when I cry unto thee : have mercy upon me, and hear me.\par
\psanum{9}My heart hath talked of thee, Seek ye my face : Thy face, Lord, will I seek.\par
\psanum{10}O hide not thou thy face from me : nor cast thy servant away in displeasure.\par
\psanum{11}Thou hast been my succour : leave me not, neither forsake me, O God of my salvation.\par
\psanum{12}When my father and my mother forsake me : the Lord taketh me up.\par
\psanum{13}Teach me thy way, O Lord : and lead me in the right way, because of mine enemies.\par
\psanum{14}Deliver me not over into the will of mine adversaries : for there are false witnesses risen up against me, and such as speak wrong.\par
\psanum{15}I should utterly have fainted : but that I believe verily to see the goodness of the Lord in the land of the living.\par
\psanum{16}O tarry thou the Lord's leisure : be strong, and he shall comfort thine heart; and put thou thy trust in the Lord.\par

\subbypsalm{Psalm 28. \textit{Ad te, Domine}}
\fancyhead[RE,LO]{Psalm 28}

\lett{U}{nto} thee will I cry, O Lord my strength : think no scorn of me; lest, if thou make as though thou hearest not, I become like them that go down into the pit.\par
\secondlinepsalm{Hear the voice of my humble petitions, when I cry unto thee : when I hold up my hands towards the mercy-seat of thy holy temple.}
\thirdlinepsalm{O pluck me not away, neither destroy me, with the ungodly and wicked doers : which speak friendly to their neighbours, but imagine mischief in their hearts.}
\psanum{4}Reward them according to their deeds : and according to the wickedness of their own inventions.\par
\psanum{5}Recompense them after the work of their hands : pay them that they have deserved.\par
\psanum{6}For they regard not in their mind the works of the Lord, nor the operation of his hands : therefore shall he break them down, and not build them up.\par
\psanum{7}Praised be the Lord : for he hath heard the voice of my humble petitions.\par
\psanum{8}The Lord is my strength and my shield; my heart hath trusted in him, and I am helped : therefore my heart danceth for joy, and in my song will I praise him.\par
\psanum{9}The Lord is my strength : and he is the wholesome defence of his Anointed.\par
\psanum{10}O save thy people, and give thy blessing unto thine inheritance : feed them, and set them up for ever.\par

\subbypsalm{Psalm 29. \textit{Afferte Domino}}
\fancyhead[RE,LO]{Psalm 29}

\lett{B}{ring} unto the Lord, O ye mighty, bring young rams unto the Lord : ascribe unto the Lord worship and strength.\par
\secondlinepsalm{Give the Lord the honour due unto his Name : worship the Lord with holy worship.}
\thirdlinepsalm{It is the Lord that commandeth the waters : it is the glorious God that maketh the thunder.}
\psanum{4}It is the Lord that ruleth the sea; the voice of the Lord is mighty in operation : the voice of the Lord is a glorious voice.\par
\psanum{5}The voice of the Lord breaketh the cedar-trees : yea, the Lord breaketh the cedars of Libanus.\par
\psanum{6}He maketh them also to skip like a calf : Libanus also, and Sirion, like a young unicorn.\par
\psanum{7}The voice of the Lord divideth the flames of fire; the voice of the Lord shaketh the wilderness : yea, the Lord shaketh the wilderness of Cades.\par
\psanum{8}The voice of the Lord maketh the hinds to bring forth young, and discovereth the thick bushes : in his temple doth every man speak of his honour.\par
\psanum{9}The Lord sitteth above the water-flood : and the Lord remaineth a King for ever.\par
\psanum{10}The Lord shall give strength unto his people : the Lord shall give his people the blessing of peace.\par

\psalmsec{Day 6. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 6. Morning Prayer}}

%MANUAL ADJUSTMENT:
\subsec{Psalm 30. \textit{Exaltabo te, Domine}}
\fancyhead[RE,LO]{Psalm 30}

\lett{I}{will} magnify thee, O Lord, for thou hast set me up : and not made my foes to triumph over me.\par
\secondlinepsalm{O Lord my God, I cried unto thee : and thou hast healed me.}
\thirdlinepsalm{Thou, Lord, hast brought my soul out of hell : thou hast kept my life from them that go down to the pit.}
\psanum{4}Sing praises unto the Lord, O ye saints of his : and give thanks unto him for a remembrance of his holiness.\par
\psanum{5}For his wrath endureth but the twinkling of an eye, and in his pleasure is life : heaviness may endure for a night, but joy cometh in the morning.\par
\psanum{6}And in my prosperity I said, I shall never be removed : thou, Lord, of thy goodness hast made my hill so strong.\par
\psanum{7}Thou didst turn thy face from me : and I was troubled.\par
\psanum{8}Then cried I unto thee, O Lord : and gat me to my Lord right humbly.\par
\psanum{9}What profit is there in my blood : when I go down to the pit?\par
\psanum{10}Shall the dust give thanks unto thee : or shall it declare thy truth?\par
\psanum{11}Hear, O Lord, and have mercy upon me : Lord, be thou my helper.\par
\psanum{12}Thou hast turned my heaviness into joy : thou hast put off my sackcloth, and girded me with gladness.\par
\psanum{13}Therefore shall every good man sing of thy praise without ceasing : O my God, I will give thanks unto thee for ever.\par

\subbypsalm{Psalm 31. \textit{In te, Domine, speravi}}
\fancyhead[RE,LO]{Psalm 31}

\lett{I}{n} thee, O Lord, have I put my trust : let me never be put to confusion, deliver me in thy righteousness.\par
\secondlinepsalm{Bow down thine ear to me : make haste to deliver me.}
\thirdlinepsalm{And be thou my strong rock, and house of defence : that thou mayest save me.}
\psanum{4}For thou art my strong rock, and my castle : be thou also my guide, and lead me for thy Name's sake.\par
\psanum{5}Draw me out of the net that they have laid privily for me : for thou art my strength.\par
\psanum{6}Into thy hands I commend my spirit : for thou hast redeemed me, O Lord, thou God of truth.\par
\psanum{7}I have hated them that hold of superstitious vanities : and my trust hath been in the Lord.\par
\psanum{8}I will be glad and rejoice in thy mercy : for thou hast considered my trouble, and hast known my soul in adversities.\par
\psanum{9}Thou hast not shut me up into the hand of the enemy : but hast set my feet in a large room.\par
\psanum{10}Have mercy upon me, O Lord, for I am in trouble : and mine eye is consumed for very heaviness; yea, my soul and my body.\par
\psanum{11}For my life is waxen old with heaviness : and my years with mourning.\par
\psanum{12}My strength faileth me, because of mine iniquity : and my bones are consumed.\par
\psanum{13}I became a reproof among all mine enemies, but especially among my neighbours : and they of mine acquaintance were afraid of me; and they that did see me without conveyed themselves from me.\par
\psanum{14}I am clean forgotten, as a dead man out of mind : I am become like a broken vessel.\par
\psanum{15}For I have heard the blasphemy of the multitude : and fear is on every side, while they conspire together against me, and take their counsel to take away my life.\par
\psanum{16}But my hope hath been in thee, O Lord : I have said, Thou art my God.\par
\psanum{17}My time is in thy hand; deliver me from the hand of mine enemies : and from them that persecute me.\par
\psanum{18}Shew thy servant the light of thy countenance : and save me for thy mercy's sake.\par
\psanum{19}Let me not be confounded, O Lord, for I have called upon thee : let the ungodly be put to confusion, and be put to silence in the grave.\par
\psanum{20}Let the lying lips be put to silence : which cruelly, disdainfully, and despitefully, speak against the righteous.\par
\psanum{21}O how plentiful is thy goodness, which thou hast laid up for them that fear thee : and that thou hast prepared for them that put their trust in thee, even before the sons of men!\par
\psanum{22}Thou shalt hide them privily by thine own presence from the provoking of all men : thou shalt keep them secretly in thy tabernacle from the strife of tongues.\par
\psanum{23}Thanks be to the Lord : for he hath shewed me marvellous great kindness in a strong city.\par
\psanum{24}And when I made haste, I said : I am cast out of the sight of thine eyes.\par
\psanum{25}Nevertheless, thou heardest the voice of my prayer : when I cried unto thee.\par
\psanum{26}O love the Lord, all ye his saints : for the Lord preserveth them that are faithful, and plenteously rewardeth the proud doer.\par
\psanum{27}Be strong, and he shall establish your heart : all ye that put your trust in the Lord.\par

\psalmsec{Day 6. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 6. Evening Prayer}}

\subbypsalm{Psalm 32. \textit{Beati, quorum}}
\fancyhead[RE,LO]{Psalm 32}

\lett{B}{lessed} is he whose unrighteousness is forgiven : and whose sin is covered.\par
\secondlinepsalm{Blessed is the man unto whom the Lord imputeth no sin : and in whose spirit there is no guile.}
\thirdlinepsalm{For while I held my tongue : my bones consumed away through my daily complaining.}
\psanum{4}For thy hand is heavy upon me day and night : and my moisture is like the drought in summer.\par
\psanum{5}I will acknowledge my sin unto thee : and mine unrighteousness have I not hid.\par
\psanum{6}I said, I will confess my sins unto the Lord : and so thou forgavest the wickedness of my sin.\par
\psanum{7}For this shall every one that is godly make his prayer unto thee, in a time when thou mayest be found : but in the great water-floods they shall not come nigh him.\par
\psanum{8}Thou art a place to hide me in, thou shalt preserve me from trouble : thou shalt compass me about with songs of deliverance.\par
\psanum{9}I will inform thee, and teach thee in the way wherein thou shalt go : and I will guide thee with mine eye.\par
\psanum{10}Be ye not like to horse and mule, which have no understanding : whose mouths must be held with bit and bridle, lest they fall upon thee.\par
\psanum{11}Great plagues remain for the ungodly : but whoso putteth his trust in the Lord, mercy embraceth him on every side.\par
\psanum{12}Be glad, O ye righteous, and rejoice in the Lord : and be joyful, all ye that are true of heart.\par

\subbypsalm{Psalm 33. \textit{Exultate, justi}}
\fancyhead[RE,LO]{Psalm 33}

\lett{R}{ejoice} in the Lord, O ye righteous : for it becometh well the just to be thankful.\par
\secondlinepsalm{Praise the Lord with harp : sing praises unto him with the lute, and instrument of ten strings.}
\thirdlinepsalm{Sing unto the Lord a new song : sing praises lustily unto him with a good courage.}
\psanum{4}For the word of the Lord is true : and all his works are faithful.\par
\psanum{5}He loveth righteousness and judgement : the earth is full of the goodness of the Lord.\par
\psanum{6}By the word of the Lord were the heavens made : and all the hosts of them by the breath of his mouth.\par
\psanum{7}He gathereth the waters of the sea together, as it were upon an heap : and layeth up the deep, as in a treasure-house.\par
\psanum{8}Let all the earth fear the Lord : stand in awe of him, all ye that dwell in the world.\par
\psanum{9}For he spake, and it was done : he commanded, and it stood fast.\par
\psanum{10}The Lord bringeth the counsel of the heathen to nought : and maketh the devices of the people to be of none effect, and casteth out the counsels of princes.\par
\psanum{11}The counsel of the Lord shall endure for ever : and the thoughts of his heart from generation to generation.\par
\psanum{12}Blessed are the people, whose God is the Lord \textsc{Jehovah} : and blessed are the folk, that he hath chosen to him to be his inheritance.\par
\psanum{13}The Lord looked down from heaven, and beheld all the children of men : from the habitation of his dwelling he considereth all them that dwell on the earth.\par
\psanum{14}He fashioneth all the hearts of them : and understandeth all their works.\par
\psanum{15}There is no king that can be saved by the multitude of an host : neither is any mighty man delivered by much strength.\par
\psanum{16}A horse is counted but a vain thing to save a man : neither shall he deliver any man by his great strength.\par
\psanum{17}Behold, the eye of the Lord is upon them that fear him : and upon them that put their trust in his mercy;\par
\psanum{18}To deliver their soul from death : and to feed them in the time of dearth.\par
\psanum{19}Our soul hath patiently tarried for the Lord : for he is our help and our shield.\par
\psanum{20}For our heart shall rejoice in him : because we have hoped in his holy Name.\par
\psanum{21}Let thy merciful kindness, O Lord, be upon us : like as we do put our trust in thee.\par

\subbypsalm{Psalm 34. \textit{Benedicam Domino}}
\fancyhead[RE,LO]{Psalm 34}

\lett{I}{will} alway give thanks unto the Lord : his praise shall ever be in my mouth.\par
\secondlinepsalm{My soul shall make her boast in the Lord : the humble shall hear thereof, and be glad.}
\thirdlinepsalm{O praise the Lord with me : and let us magnify his Name together.}
\psanum{4}I sought the Lord, and he heard me : yea, he delivered me out of all my fear.\par
\psanum{5}They had an eye unto him, and were lightened : and their faces were not ashamed.\par
\psanum{6}Lo, the poor crieth, and the Lord heareth him : yea, and saveth him out of all his troubles.\par
\psanum{7}The angel of the Lord tarrieth round about them that fear him : and delivereth them.\par
\psanum{8}O taste, and see, how gracious the Lord is : blessed is the man that trusteth in him.\par
\psanum{9}O fear the Lord, ye that are his saints : for they that fear him lack nothing.\par
\psanum{10}The lions do lack, and suffer hunger : but they who seek the Lord shall want no manner of thing that is good.\par
\psanum{11}Come, ye children, and hearken unto me : I will teach you the fear of the Lord.\par
\psanum{12}What man is he that lusteth to live : and would fain see good days?\par
\psanum{13}Keep thy tongue from evil : and thy lips, that they speak no guile.\par
\psanum{14}Eschew evil, and do good : seek peace, and ensue it.\par
\psanum{15}The eyes of the Lord are over the righteous : and his ears are open unto their prayers.\par
\psanum{16}The countenance of the Lord is against them that do evil : to root out the remembrance of them from the earth.\par
\psanum{17}The righteous cry, and the Lord heareth them : and delivereth them out of all their troubles.\par
\psanum{18}The Lord is nigh unto them that are of a contrite heart : and will save such as be of an humble spirit.\par
\psanum{19}Great are the troubles of the righteous : but the Lord delivereth him out of all.\par
\psanum{20}He keepeth all his bones : so that not one of them is broken.\par
\psanum{21}But misfortune shall slay the ungodly : and they that hate the righteous shall be desolate.\par
\psanum{22}The Lord delivereth the souls of his servants : and all they that put their trust in him shall not be destitute.\par

\psalmsec{Day 7. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 7. Morning Prayer}}

\subbypsalm{Psalm 35. \textit{Judica, Domine}}
\fancyhead[RE,LO]{Psalm 35}

\lett{P}{lead} thou my cause, O Lord, with them that strive with me : and fight thou against them that fight against me.\par
\secondlinepsalm{Lay hand upon the shield and buckler : and stand up to help me.}
\thirdlinepsalm{Bring forth the spear, and stop the way against them that persecute me : say unto my soul, I am thy salvation.}
\psanum{4}Let them be confounded and put to shame, that seek after my soul : let them be turned back and brought to confusion, that imagine mischief for me.\par
\psanum{5}Let them be as the dust before the wind : and the angel of the Lord scattering them.\par
\psanum{6}Let their way be dark and slippery : and let the angel of the Lord persecute them.\par
\psanum{7}For they have privily laid their net to destroy me without a cause : yea, even without a cause have they made a pit for my soul.\par
\psanum{8}Let a sudden destruction come upon him unawares, and his net, that he hath laid privily, catch himself : that he may fall into his own mischief.\par
\psanum{9}And, my soul, be joyful in the Lord : it shall rejoice in his salvation.\par
\psanum{10}All my bones shall say, Lord, who is like unto thee, who deliverest the poor from him that is too strong for him : yea, the poor, and him that is in misery, from him that spoileth him?\par
\psanum{11}False witnesses did rise up : they laid to my charge things that I knew not.\par
\psanum{12}They rewarded me evil for good : to the great discomfort of my soul.\par
\psanum{13}Nevertheless, when they were sick, I put on sackcloth, and humbled my soul with fasting : and my prayer shall turn into mine own bosom.\par
\psanum{14}I behaved myself as though it had been my friend or my brother : I went heavily, as one that mourneth for his mother.\par
\psanum{15}But in mine adversity they rejoiced, and gathered themselves together : yea, the very abjects came together against me unawares, making mouths at me, and ceased not.\par
\psanum{16}With the flatterers were busy mockers : who gnashed upon me with their teeth.\par
\psanum{17}Lord, how long wilt thou look upon this : O deliver my soul from the calamities which they bring on me, and my darling from the lions.\par
\psanum{18}So will I give thee thanks in the great congregation : I will praise thee among much people.\par
\psanum{19}O let not them that are mine enemies triumph over me ungodly : neither let them wink with their eyes that hate me without a cause.\par
\psanum{20}And why? their communing is not for peace : but they imagine deceitful words against them that are quiet in the land.\par
\psanum{21}They gaped upon me with their mouths, and said : Fie on thee, fie on thee, we saw it with our eyes.\par
\psanum{22}This thou hast seen, O Lord : hold not thy tongue then, go not far from me, O Lord.\par
\psanum{23}Awake, and stand up to judge my quarrel : avenge thou my cause, my God, and my Lord.\par
\psanum{24}Judge me, O Lord my God, according to thy righteousness : and let them not triumph over me.\par
\psanum{25}Let them not say in their hearts, There, there, so would we have it : neither let them say, We have devoured him.\par
\psanum{26}Let them be put to confusion and shame together, that rejoice at my trouble : let them be clothed with rebuke and dishonour, that boast themselves against me.\par
\psanum{27}Let them be glad and rejoice, that favour my righteous dealing : yea, let them say alway, Blessed be the Lord, who hath pleasure in the prosperity of his servant.\par
\psanum{28}And as for my tongue, it shall be talking of thy righteousness : and of thy praise all the day long.\par

\subbypsalm{Psalm 36. \textit{Dixit injustus}}
\fancyhead[RE,LO]{Psalm 36}

\lett{M}{y} heart sheweth me the wickedness of the ungodly : that there is no fear of God before his eyes.\par
\secondlinepsalm{For he flattereth himself in his own sight : until his abominable sin be found out.}
\thirdlinepsalm{The words of his mouth are unrighteous, and full of deceit : he hath left off to behave himself wisely, and to do good.}
\psanum{4}He imagineth mischief upon his bed, and hath set himself in no good way : neither doth he abhor any thing that is evil.\par
\psanum{5}Thy mercy, O Lord, reacheth unto the heavens : and thy faithfulness unto the clouds.\par
\psanum{6}Thy righteousness standeth like the strong mountains : thy judgements are like the great deep.\par
\psanum{7}Thou, Lord, shalt save both man and beast; How excellent is thy mercy, O God : and the children of men shall put their trust under the shadow of thy wings.\par
\psanum{8}They shall be satisfied with the plenteousness of thy house : and thou shalt give them drink of thy pleasures, as out of the river.\par
\psanum{9}For with thee is the well of life : and in thy light shall we see light.\par
\psanum{10}O continue forth thy loving-kindness unto them that know thee : and thy righteousness unto them that are true of heart.\par
\psanum{11}O let not the foot of pride come against me : and let not the hand of the ungodly cast me down.\par
\psanum{12}There are they fallen, all that work wickedness : they are cast down, and shall not be able to stand.\par

\psalmsec{Day 7. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 7. Evening Prayer}}

\subbypsalm{Psalm 37. \textit{Noli aemulari}}
\fancyhead[RE,LO]{Psalm 37}

\lett{F}{ret} not thyself because of the ungodly : neither be thou envious against the evil-doers.\par
\secondlinepsalm{For they shall soon be cut down like the grass : and be withered even as the green herb.}
\thirdlinepsalm{Put thou thy trust in the Lord, and be doing good : dwell in the land, and verily thou shalt be fed.}
\psanum{4}Delight thou in the Lord : and he shall give thee thy heart's desire.\par
\psanum{5}Commit thy way unto the Lord, and put thy trust in him : and he shall bring it to pass.\par
\psanum{6}He shall make thy righteousness as clear as the light : and thy just dealing as the noon-day.\par
\psanum{7}Hold thee still in the Lord, and abide patiently upon him : but grieve not thyself at him whose way doth prosper, against the man that doeth after evil counsels.\par
\psanum{8}Leave off from wrath, and let go displeasure : fret not thyself, else shalt thou be moved to do evil.\par
\psanum{9}Wicked doers shall be rooted out : and they that patiently abide the Lord, those shall inherit the land.\par
\psanum{10}Yet a little while, and the ungodly shall be clean gone : thou shalt look after his place, and he shall be away.\par
\psanum{11}But the meek-spirited shall possess the earth : and shall be refreshed in the multitude of peace.\par
\psanum{12}The ungodly seeketh counsel against the just : and gnasheth upon him with his teeth.\par
\psanum{13}The Lord shall laugh him to scorn : for he hath seen that his day is coming.\par
\psanum{14}The ungodly have drawn out the sword, and have bent their bow : to cast down the poor and needy, and to slay such as are of a right conversation.\par
\psanum{15}Their sword shall go through their own heart : and their bow shall be broken.\par
\psanum{16}A small thing that the righteous hath : is better than great riches of the ungodly.\par
\psanum{17}For the arms of the ungodly shall be broken : and the Lord upholdeth the righteous.\par
\psanum{18}The Lord knoweth the days of the godly : and their inheritance shall endure for ever.\par
\psanum{19}They shall not be confounded in the perilous time : and in the days of dearth they shall have enough.\par
\psanum{20}As for the ungodly, they shall perish; and the enemies of the Lord shall consume as the fat of lambs : yea, even as the smoke shall they consume away.\par
\psanum{21}The ungodly borroweth, and payeth not again : but the righteous is merciful and liberal.\par
\psanum{22}Such as are blessed of God shall possess the land : and they that are cursed of him shall be rooted out.\par
\psanum{23}The Lord ordereth a good man's going : and maketh his way acceptable to himself.\par
\psanum{24}Though he fall, he shall not be cast away : for the Lord upholdeth him with his hand.\par
\psanum{25}I have been young, and now am old : and yet saw I never the righteous forsaken, nor his seed begging their bread.\par
\psanum{26}The righteous is ever merciful, and lendeth : and his seed is blessed.\par
\psanum{27}Flee from evil, and do the thing that is good : and dwell for evermore.\par
\psanum{28}For the Lord loveth the thing that is right : he forsaketh not his that be godly, but they are preserved for ever.\par
\psanum{29}The unrighteous shall be punished : as for the seed of the ungodly, it shall be rooted out.\par
\psanum{30}The righteous shall inherit the land : and dwell therein for ever.\par
\psanum{31}The mouth of the righteous is exercised in wisdom : and his tongue will be talking of judgement.\par
\psanum{32}The law of his God is in his heart : and his goings shall not slide.\par
\psanum{33}The ungodly seeth the righteous : and seeketh occasion to slay him.\par
\psanum{34}The Lord will not leave him in his hand : nor condemn him when he is judged.\par
\psanum{35}Hope thou in the Lord, and keep his way, and he shall promote thee, that thou shalt possess the land : when the ungodly shall perish, thou shalt see it.\par
\psanum{36}I myself have seen the ungodly in great power : and flourishing like a green bay-tree.\par
\psanum{37}I went by, and lo, he was gone : I sought him, but his place could no where be found.\par
\psanum{38}Keep innocency, and take heed unto the thing that is right : for that shall bring a man peace at the last.\par
\psanum{39}As for the transgressors, they shall perish together : and the end of the ungodly is, they shall be rooted out at the last.\par
\psanum{40}But the salvation of the righteous cometh of the Lord : who is also their strength in the time of trouble.\par
\psanum{41}And the Lord shall stand by them, and save them : he shall deliver them from the ungodly, and shall save them, because they put their trust in him.\par


\psalmsec{Day 8. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 8. Morning Prayer}}

\subbypsalm{Psalm 38. \textit{Domine, ne in furore}}
\fancyhead[RE,LO]{Psalm 38}

\lett{P}{ut} me not to rebuke, O Lord, in thine anger : neither chasten me in thy heavy displeasure.\par
\secondlinepsalm{For thine arrows stick fast in me : and thy hand presseth me sore.}
\thirdlinepsalm{There is no health in my flesh, because of thy displeasure : neither is there any rest in my bones, by reason of my sin.}
\psanum{4}For my wickednesses are gone over my head : and are like a sore burden, too heavy for me to bear.\par
\psanum{5}My wounds stink, and are corrupt : through my foolishness.\par
\psanum{6}I am brought into so great trouble and misery : that I go mourning all the day long.\par
\psanum{7}For my loins are filled with a sore disease : and there is no whole part in my body.\par
\psanum{8}I am feeble, and sore smitten : I have roared for the very disquietness of my heart.\par
\psanum{9}Lord, thou knowest all my desire : and my groaning is not hid from thee.\par
\psanum{10}My heart panteth, my strength hath failed me : and the sight of mine eyes is gone from me.\par
\psanum{11}My lovers and my neighbours did stand looking upon my trouble : and my kinsmen stood afar off.\par
\psanum{12}They also that sought after my life laid snares for me : and they that went about to do me evil talked of wickedness, and imagined deceit all the day long.\par
\psanum{13}As for me, I was like a deaf man, and heard not : and as one that is dumb, who doth not open his mouth.\par
\psanum{14}I became even as a man that heareth not : and in whose mouth are no reproofs.\par
\psanum{15}For in thee, O Lord, have I put my trust : thou shalt answer for me, O Lord my God.\par
\psanum{16}I have required that they, even mine enemies, should not triumph over me : for when my foot slipped, they rejoiced greatly against me.\par
\psanum{17}And I, truly, am set in the plague : and my heaviness is ever in my sight.\par
\psanum{18}For I will confess my wickedness : and be sorry for my sin.\par
\psanum{19}But mine enemies live, and are mighty : and they that hate me wrongfully are many in number.\par
\psanum{20}They also that reward evil for good are against me : because I follow the thing that good is.\par
\psanum{21}Forsake me not, O Lord my God : be not thou far from me.\par
\psanum{22}Haste thee to help me : O Lord God of my salvation.\par

\subbypsalm{Psalm 39. \textit{Dixi, Custodiam}}
\fancyhead[RE,LO]{Psalm 39}

\lett{I}{said,} I will take heed to my ways : that I offend not in my tongue.\par
\secondlinepsalm{I will keep my mouth as it were with a bridle : while the ungodly is in my sight.}
\thirdlinepsalm{I held my tongue, and spake nothing : I kept silence, yea, even from good words; but it was pain and grief to me.}
\psanum{4}My heart was hot within me, and while I was thus musing the fire kindled : and at the last I spake with my tongue;\par
\psanum{5}Lord, let me know mine end, and the number of my days : that I may be certified how long I have to live.\par
\psanum{6}Behold, thou hast made my days as it were a span long : and mine age is even as nothing in respect of thee; and verily every man living is altogether vanity.\par
\psanum{7}For man walketh in a vain shadow, and disquieteth himself in vain : he heapeth up riches, and cannot tell who shall gather them.\par
\psanum{8}And now, Lord, what is my hope : truly my hope is even in thee.\par
\psanum{9}Deliver me from all mine offences : and make me not a rebuke unto the foolish.\par
\psanum{10}I became dumb, and opened not my mouth : for it was thy doing.\par
\psanum{11}Take thy plague away from me : I am even consumed by the means of thy heavy hand.\par
\psanum{12}When thou with rebukes dost chasten man for sin, thou makest his beauty to consume away, like as it were a moth fretting a garment : every man therefore is but vanity.\par
\psanum{13}Hear my prayer, O Lord, and with thine ears consider my calling : hold not thy peace at my tears.\par
\psanum{14}For I am a stranger with thee : and a sojourner, as all my fathers were.\par
\psanum{15}O spare me a little, that I may recover my strength : before I go hence, and be no more seen.\par

\subbypsalm{Psalm 40. \textit{Expectans expectavi}}
\fancyhead[RE,LO]{Psalm 40}

\lett{I}{waited} patiently for the Lord : and he inclined unto me, and heard my calling.\par
\secondlinepsalm{He brought me also out of the horrible pit, out of the mire and clay : and set my feet upon the rock, and ordered my goings.}
\thirdlinepsalm{And he hath put a new song in my mouth : even a thanksgiving unto our God.}
\psanum{4}Many shall see it, and fear : and shall put their trust in the Lord.\par
\psanum{5}Blessed is the man that hath set his hope in the Lord : and turned not unto the proud, and to such as go about with lies.\par
\psanum{6}O Lord my God, great are the wondrous works which thou hast done, like as be also thy thoughts which are to us-ward : and yet there is no man that ordereth them unto thee:\par
\psanum{7}If I should declare them, and speak of them : they should be more than I am able to express.\par
\psanum{8}Sacrifice and meat-offering thou wouldest not : but mine ears hast thou opened.\par
\psanum{9}Burnt-offerings, and sacrifice for sin, hast thou not required : then said I, Lo, I come,\par
\psanum{10}In the volume of the book it is written of me, that I should fulfil thy will, O my God : I am content to do it; yea, thy law is within my heart.\par
\psanum{11}I have declared thy righteousness in the great congregation : lo, I will not refrain my lips, O Lord, and that thou knowest.\par
\psanum{12}I have not hid thy righteousness within my heart : my talk hath been of thy truth and of thy salvation.\par
\psanum{13}I have not kept back thy loving mercy and truth : from the great congregation.\par
\psanum{14}Withdraw not thou thy mercy from me, O Lord : let thy loving-kindness and thy truth alway preserve me.\par
\psanum{15}For innumerable troubles are come about me; my sins have taken such hold upon me that I am not able to look up : yea, they are more in number than the hairs of my head, and my heart hath failed me.\par
\psanum{16}O Lord, let it be thy pleasure to deliver me : make haste, O Lord, to help me.\par
\psanum{17}Let them be ashamed and confounded together, that seek after my soul to destroy it : let them be driven backward and put to rebuke, that wish me evil.\par
\psanum{18}Let them be desolate, and rewarded with shame : that say unto me, Fie upon thee, fie upon thee.\par
\psanum{19}Let all those that seek thee be joyful and glad in thee : and let such as love thy salvation say alway, The Lord be praised.\par
\psanum{20}As for me, I am poor and needy : but the Lord careth for me.\par
\psanum{21}Thou art my helper and redeemer : make no long tarrying, O my God.\par

\psalmsec{Day 8. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 8. Evening Prayer}}

\subbypsalm{Psalm 41. \textit{Beatus qui intelligit}}
\fancyhead[RE,LO]{Psalm 41}

\lett{B}{lessed} is he that considereth the poor and needy : the Lord shall deliver him in the time of trouble.\par
\secondlinepsalm{The Lord preserve him, and keep him alive, that he may be blessed upon earth : and deliver not thou him into the will of his enemies.}
\thirdlinepsalm{The Lord comfort him, when he lieth sick upon his bed : make thou all his bed in his sickness.}
\psanum{4}I said, Lord, be merciful unto me : heal my soul, for I have sinned against thee.\par
\psanum{5}Mine enemies speak evil of me : When shall he die, and his name perish?\par
\psanum{6}And if he come to see me, he speaketh vanity : and his heart conceiveth falsehood within himself, and when he cometh forth he telleth it.\par
\psanum{7}All mine enemies whisper together against me : even against me do they imagine this evil.\par
\psanum{8}Let the sentence of guiltiness proceed against him : and now that he lieth, let him rise up no more.\par
\psanum{9}Yea, even mine own familiar friend, whom I trusted : who did also eat of my bread, hath laid great wait for me.\par
\psanum{10}But be thou merciful unto me, O Lord : raise thou me up again, and I shall reward them.\par
\psanum{11}By this I know thou favourest me : that mine enemy doth not triumph against me.\par
\psanum{12}And when I am in my health, thou upholdest me : and shalt set me before thy face for ever.\par
\psanum{13}Blessed be the Lord God of Israel : world without end. Amen.\par

\subbypsalm{Psalm 42. \textit{Quemadmodum}}
\fancyhead[RE,LO]{Psalm 42}

\lett{L}{ike} as the hart desireth the water-brooks : so longeth my soul after thee, O God.\par
\secondlinepsalm{My soul is athirst for God, yea, even for the living God : when shall I come to appear before the presence of God?}
\thirdlinepsalm{My tears have been my meat day and night : while they daily say unto me, Where is now thy God?}
\psanum{4}Now when I think thereupon, I pour out my heart by myself : for I went with the multitude, and brought them forth into the house of God;\par
\psanum{5}In the voice of praise and thanksgiving : among such as keep holy-day.\par
\psanum{6}Why art thou so full of heaviness, O my soul : and why art thou so disquieted within me?\par
\psanum{7}Put thy trust in God : for I will yet give him thanks for the help of his countenance.\par
\psanum{8}My God, my soul is vexed within me : therefore will I remember thee concerning the land of Jordan, and the little hill of Hermon.\par
\psanum{9}One deep calleth another, because of the noise of the water-pipes : all thy waves and storms are gone over me.\par
\psanum{10}The Lord hath granted his loving-kindness in the day-time : and in the night-season did I sing of him, and made my prayer unto the God of my life.\par
\psanum{11}I will say unto the God of my strength, Why hast thou forgotten me : why go I thus heavily, while the enemy oppresseth me?\par
\psanum{12}My bones are smitten asunder as with a sword : while mine enemies that trouble me cast me in the teeth;\par
\psanum{13}Namely, while they say daily unto me : Where is now thy God?\par
\psanum{14}Why art thou so vexed, O my soul : and why art thou so disquieted within me?\par
\psanum{15}O put thy trust in God : for I will yet thank him, which is the help of my countenance, and my God.\par

\subbypsalm{Psalm 43. \textit{Judica me, Deus}}
\fancyhead[RE,LO]{Psalm 43}

\lett{G}{ive} sentence with me, O God, and defend my cause against the ungodly people : O deliver me from the deceitful and wicked man.\par
\secondlinepsalm{For thou art the God of my strength, why hast thou put me from thee : and why go I so heavily, while the enemy oppresseth me?}
\thirdlinepsalm{O send out thy light and thy truth, that they may lead me : and bring me unto thy holy hill, and to thy dwelling.}
\psanum{4}And that I may go unto the altar of God, even unto the God of my joy and gladness : and upon the harp will I give thanks unto thee, O God, my God.\par
\psanum{5}Why art thou so heavy, O my soul : and why art thou so disquieted within me?\par
\psanum{6}O put thy trust in God : for I will yet give him thanks, which is the help of my countenance, and my God.\par

\psalmsec{Day 9. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 9. Morning Prayer}}

\subbypsalm{Psalm 44. \textit{Deus, auribus}}
\fancyhead[RE,LO]{Psalm 44}

\lett{W}{e} have heard with our ears, O God, our fathers have told us : what thou hast done in their time of old;\par
\secondlinepsalm{How thou hast driven out the heathen with thy hand, and planted them in : how thou hast destroyed the nations and cast them out.}
\thirdlinepsalm{For they gat not the land in possession through their own sword : neither was it their own arm that helped them;}
\psanum{4}But thy right hand, and thine arm, and the light of thy countenance : because thou hadst a favour unto them.\par
\psanum{5}Thou art my King, O God : send help unto Jacob.\par
\psanum{6}Through thee will we overthrow our enemies : and in thy Name will we tread them under, that rise up against us.\par
\psanum{7}For I will not trust in my bow : it is not my sword that shall help me;\par
\psanum{8}But it is thou that savest us from our enemies : and puttest them to confusion that hate us.\par
\psanum{9}We make our boast of God all day long : and will praise thy Name for ever.\par
\psanum{10}But now thou art far off, and puttest us to confusion : and goest not forth with our armies.\par
\psanum{11}Thou makest us to turn our backs upon our enemies : so that they which hate us spoil our goods.\par
\psanum{12}Thou lettest us be eaten up like sheep : and hast scattered us among the heathen.\par
\psanum{13}Thou sellest thy people for nought : and takest no money for them.\par
\psanum{14}Thou makest us to be rebuked of our neighbours : to be laughed to scorn, and had in derision of them that are round about us.\par
\psanum{15}Thou makest us to be a by-word among the heathen : and that the people shake their heads at us.\par
\psanum{16}My confusion is daily before me : and the shame of my face hath covered me;\par
\psanum{17}For the voice of the slanderer and blasphemer : for the enemy and avenger.\par
\psanum{18}And though all this be come upon us, yet do we not forget thee : nor behave ourselves frowardly in thy covenant.\par
\psanum{19}Our heart is not turned back : neither our steps gone out of thy way;\par
\psanum{20}No, not when thou hast smitten us into the place of dragons : and covered us with the shadow of death.\par
\psanum{21}If we have forgotten the Name of our God, and holden up our hands to any strange god : shall not God search it out? for he knoweth the very secrets of the heart.\par
\psanum{22}For thy sake also are we killed all the day long : and are counted as sheep appointed to be slain.\par
\psanum{23}Up, Lord, why sleepest thou : awake, and be not absent from us for ever.\par
\psanum{24}Wherefore hidest thou thy face : and forgettest our misery and trouble?\par
\psanum{25}For our soul is brought low, even unto the dust : our belly cleaveth unto the ground.\par
\psanum{26}Arise, and help us : and deliver us for thy mercy's sake.\par

\subbypsalm{Psalm 45. \textit{Eructavit cor meum}}
\fancyhead[RE,LO]{Psalm 45}

\lett{M}{y} heart is inditing of a good matter : I speak of the things which I have made unto the King.\par
\secondlinepsalm{My tongue is the pen : of a ready writer.}
\thirdlinepsalm{Thou art fairer than the children of men : full of grace are thy lips, because God hath blessed thee for ever.}
\psanum{4}Gird thee with thy sword upon thy thigh, O thou most Mighty : according to thy worship and renown.\par
\psanum{5}Good luck have thou with thine honour : ride on, because of the word of truth, of meekness, and righteousness; and thy right hand shall teach thee terrible things.\par
\psanum{6}Thy arrows are very sharp, and the people shall be subdued unto thee : even in the midst among the King's enemies.\par
\psanum{7}Thy seat, O God, endureth for ever : the sceptre of thy kingdom is a right sceptre.\par
\psanum{8}Thou hast loved righteousness, and hated iniquity : wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows.\par
\psanum{9}All thy garments smell of myrrh, aloes, and cassia : out of the ivory palaces, whereby they have made thee glad.\par
\psanum{10}Kings' daughters were among thy honourable women : upon thy right hand did stand the queen in a vesture of gold, wrought about with divers colours.\par
\psanum{11}Hearken, O daughter, and consider, incline thine ear : forget also thine own people, and thy father's house.\par
\psanum{12}So shall the King have pleasure in thy beauty : for he is thy Lord God, and worship thou him.\par
\psanum{13}And the daughter of Tyre shall be there with a gift : like as the rich also among the people shall make their supplication before thee.\par
\psanum{14}The King's daughter is all glorious within : her clothing is of wrought gold.\par
\psanum{15}She shall be brought unto the King in raiment of needle-work : the virgins that be her fellows shall bear her company, and shall be brought unto thee.\par
\psanum{16}With joy and gladness shall they be brought : and shall enter into the King's palace.\par
\psanum{17}Instead of thy fathers thou shalt have children : whom thou mayest make princes in all lands.\par
\psanum{18}I will remember thy Name from one generation to another : therefore shall the people give thanks unto thee, world without end.\par

\subbypsalm{Psalm 46. \textit{Deus noster refugium}}
\fancyhead[RE,LO]{Psalm 46}

\lett{G}{od} is our hope and strength : a very present help in trouble.\par
\secondlinepsalm{Therefore will we not fear, though the earth be moved : and though the hills be carried into the midst of the sea;}
\thirdlinepsalm{Though the waters thereof rage and swell : and though the mountains shake at the tempest of the same.}
\psanum{4}The rivers of the flood thereof shall make glad the city of God : the holy place of the tabernacle of the most Highest.\par
\psanum{5}God is in the midst of her, therefore shall she not be removed : God shall help her, and that right early.\par
\psanum{6}The heathen make much ado, and the kingdoms are moved : but God hath shewed his voice, and the earth shall melt away.\par
\psanum{7}The Lord of hosts is with us : the God of Jacob is our refuge.\par
\psanum{8}O come hither, and behold the works of the Lord : what destruction he hath brought upon the earth.\par
\psanum{9}He maketh wars to cease in all the world : he breaketh the bow, and knappeth the spear in sunder, and burneth the chariots in the fire.\par
\psanum{10}Be still then, and know that I am God : I will be exalted among the heathen, and I will be exalted in the earth.\par
\psanum{11}The Lord of hosts is with us : the God of Jacob is our refuge.\par

\psalmsec{Day 9. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 9. Evening Prayer}}

\subbypsalm{Psalm 47. \textit{Omnes gentes, plaudite}}
\fancyhead[RE,LO]{Psalm 47}

\lett{O}{clap} your hands together, all ye people : O sing unto God with the voice of melody.\par
\secondlinepsalm{For the Lord is high, and to be feared : he is the great King upon all the earth.}
\thirdlinepsalm{He shall subdue the people under us : and the nations under our feet.}
\psanum{4}He shall choose out an heritage for us : even the worship of Jacob, whom he loved.\par
\psanum{5}God is gone up with a merry noise : and the Lord with the sound of the trump.\par
\psanum{6}O sing praises, sing praises unto our God : O sing praises, sing praises unto our King.\par
\psanum{7}For God is the King of all the earth : sing ye praises with understanding.\par
\psanum{8}God reigneth over the heathen : God sitteth upon his holy seat.\par
\psanum{9}The princes of the people are joined unto the people of the God of Abraham : for God, which is very high exalted, doth defend the earth, as it were with a shield.\par

\subbypsalm{Psalm 48. \textit{Magnus Dominus}}
\fancyhead[RE,LO]{Psalm 48}

\lett{G}{reat} is the Lord, and highly to be praised : in the city of our God, even upon his holy hill.\par
\secondlinepsalm{The hill of Sion is a fair place, and the joy of the whole earth : upon the north-side lieth the city of the great King; God is well known in her palaces as a sure refuge.}
\thirdlinepsalm{For lo, the kings of the earth : are gathered, and gone by together.}
\psanum{4}They marvelled to see such things : they were astonished, and suddenly cast down.\par
\psanum{5}Fear came there upon them, and sorrow : as upon a woman in her travail.\par
\psanum{6}Thou shalt break the ships of the sea : through the east-wind.\par
\psanum{7}Like as we have heard, so have we seen in the city of the Lord of hosts, in the city of our God : God upholdeth the same for ever.\par
\psanum{8}We wait for thy loving-kindness, O God : in the midst of thy temple.\par
\psanum{9}O God, according to thy Name, so is thy praise unto the world's end : thy right hand is full of righteousness.\par
\psanum{10}Let the mount Sion rejoice, and the daughters of Judah be glad : because of thy judgements.\par
\psanum{11}Walk about Sion, and go round about her : and tell the towers thereof.\par
\psanum{12}Mark well her bulwarks, set up her houses : that ye may tell them that come after.\par
\psanum{13}For this God is our God for ever and ever : he shall be our guide unto death.\par

\subbypsalm{Psalm 49. \textit{Audite haec, omnes}}
\fancyhead[RE,LO]{Psalm 49}

\lett{O}{hear} ye this, all ye people : ponder it with your ears, all ye that dwell in the world;\par
\secondlinepsalm{High and low, rich and poor : one with another.}
\thirdlinepsalm{My mouth shall speak of wisdom : and my heart shall muse of understanding.}
\psanum{4}I will incline mine ear to the parable : and shew my dark speech upon the harp.\par
\psanum{5}Wherefore should I fear in the days of wickedness : and when the wickedness of my heels compasseth me round about?\par
\psanum{6}There be some that put their trust in their goods : and boast themselves in the multitude of their riches.\par
\psanum{7}But no man may deliver his brother : nor make agreement unto God for him;\par
\psanum{8}For it cost more to redeem their souls : so that he must let that alone for ever;\par
\psanum{9}Yea, though he live long : and see not the grave.\par
\psanum{10}For he seeth that wise men also die, and perish together : as well as the ignorant and foolish, and leave their riches for other.\par
\psanum{11}And yet they think that their houses shall continue for ever : and that their dwelling-places shall endure from one generation to another; and call the lands after their own names.\par
\psanum{12}Nevertheless, man will not abide in honour : seeing he may be compared unto the beasts that perish; this is the way of them.\par
\psanum{13}This is their foolishness : and their posterity praise their saying.\par
\psanum{14}They lie in the hell like sheep, death gnaweth upon them, and the righteous shall have domination over them in the morning : their beauty shall consume in the sepulchre out of their dwelling.\par
\psanum{15}But God hath delivered my soul from the place of hell : for he shall receive me.\par
\psanum{16}Be not thou afraid, though one be made rich : or if the glory of his house be increased;\par
\psanum{17}For he shall carry nothing away with him when he dieth : neither shall his pomp follow him.\par
\psanum{18}For while he lived, he counted himself an happy man : and so long as thou doest well unto thyself, men will speak good of thee.\par
\psanum{19}He shall follow the generation of his fathers : and shall never see light.\par
\psanum{20}Man being in honour hath no understanding : but is compared unto the beasts that perish.\par

\psalmsec{Day 10. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 10. Morning Prayer}}

\subbypsalm{Psalm 50. \textit{Deus deorum}}
\fancyhead[RE,LO]{Psalm 50}

\lett{T}{he} Lord, even the most mighty God, hath spoken : and called the world, from the rising up of the sun unto the going down thereof.\par
\secondlinepsalm{Out of Sion hath God appeared : in perfect beauty.}
\thirdlinepsalm{Our God shall come, and shall not keep silence : there shall go before him a consuming fire, and a mighty tempest shall be stirred up round about him.}
\psanum{4}He shall call the heaven from above : and the earth, that he may judge his people.\par
\psanum{5}Gather my saints together unto me : those that have made a covenant with me with sacrifice.\par
\psanum{6}And the heavens shall declare his righteousness : for God is Judge himself.\par
\psanum{7}Hear, O my people, and I will speak : I myself will testify against thee, O Israel; for I am God, even thy God.\par
\psanum{8}I will not reprove thee because of thy sacrifices, or for thy burnt-offerings : because they were not alway before me.\par
\psanum{9}I will take no bullock out of thine house : nor he-goat out of thy folds.\par
\psanum{10}For all the beasts of the forest are mine : and so are the cattle upon a thousand hills.\par
\psanum{11}I know all the fowls upon the mountains : and the wild beasts of the field are in my sight.\par
\psanum{12}If I be hungry, I will not tell thee : for the whole world is mine, and all that is therein.\par
\psanum{13}Thinkest thou that I will eat bulls' flesh : and drink the blood of goats?\par
\psanum{14}Offer unto God thanksgiving : and pay thy vows unto the most Highest.\par
\psanum{15}And call upon me in the time of trouble : so will I hear thee, and thou shalt praise me.\par
\psanum{16}But unto the ungodly said God : Why dost thou preach my laws, and takest my covenant in thy mouth;\par
\psanum{17}Whereas thou hatest to be reformed : and hast cast my words behind thee?\par
\psanum{18}When thou sawest a thief, thou consentedst unto him : and hast been partaker with the adulterers.\par
\psanum{19}Thou hast let thy mouth speak wickedness : and with thy tongue thou hast set forth deceit.\par
\psanum{20}Thou satest, and spakest against thy brother : yea, and hast slandered thine own mother's son.\par
\psanum{21}These things hast thou done, and I held my tongue, and thou thoughtest wickedly, that I am even such a one as thyself : but I will reprove thee, and set before thee the things that thou hast done.\par
\psanum{22}O consider this, ye that forget God : lest I pluck you away, and there be none to deliver you.\par
\psanum{23}Whoso offereth me thanks and praise, he honoureth me : and to him that ordereth his conversation right will I shew the salvation of God.\par

\subbypsalm{Psalm 51. \textit{Miserere mei, Deus}}
\fancyhead[RE,LO]{Psalm 51}

\lett{H}{ave} mercy upon me, O God, after thy great goodness : according to the multitude of thy mercies do away mine offences.\par
\secondlinepsalm{Wash me throughly from my wickedness : and cleanse me from my sin.}
\thirdlinepsalm{For I acknowledge my faults : and my sin is ever before me.}
\psanum{4}Against thee only have I sinned, and done this evil in thy sight : that thou mightest be justified in thy saying, and clear when thou art judged.\par
\psanum{5}Behold, I was shapen in wickedness : and in sin hath my mother conceived me.\par
\psanum{6}But lo, thou requirest truth in the inward parts : and shalt make me to understand wisdom secretly.\par
\psanum{7}Thou shalt purge me with hyssop, and I shall be clean : thou shalt wash me, and I shall be whiter than snow.\par
\psanum{8}Thou shalt make me hear of joy and gladness : that the bones which thou hast broken may rejoice.\par
\psanum{9}Turn thy face from my sins : and put out all my misdeeds.\par
\psanum{10}Make me a clean heart, O God : and renew a right spirit within me.\par
\psanum{11}Cast me not away from thy presence : and take not thy holy Spirit from me.\par
\psanum{12}O give me the comfort of thy help again : and stablish me with thy free Spirit.\par
\psanum{13}Then shall I teach thy ways unto the wicked : and sinners shall be converted unto thee.\par
\psanum{14}Deliver me from blood-guiltiness, O God, thou that art the God of my health : and my tongue shall sing of thy righteousness.\par
\psanum{15}Thou shalt open my lips, O Lord : and my mouth shall shew thy praise.\par
\psanum{16}For thou desirest no sacrifice, else would I give it thee : but thou delightest not in burnt-offerings.\par
\psanum{17}The sacrifice of God is a troubled spirit : a broken and contrite heart, O God, shalt thou not despise.\par
\psanum{18}O be favourable and gracious unto Sion : build thou the walls of Jerusalem.\par
\psanum{19}Then shalt thou be pleased with the sacrifice of righteousness, with the burnt-offerings and oblations : then shall they offer young bullocks upon thine altar.\par

\subbypsalm{Psalm 52. \textit{Quid gloriaris?}}
\fancyhead[RE,LO]{Psalm 52}

\lett{W}{hy} boastest thou thyself, thou tyrant : that thou canst do mischief;\par
\secondlinepsalm{Whereas the goodness of God : endureth yet daily?}
\thirdlinepsalm{Thy tongue imagineth wickedness : and with lies thou cuttest like a sharp rasor.}
\psanum{4}Thou hast loved unrighteousness more than goodness : and to talk of lies more than righteousness.\par
\psanum{5}Thou hast loved to speak all words that may do hurt : O thou false tongue.\par
\psanum{6}Therefore shall God destroy thee for ever : he shall take thee, and pluck thee out of thy dwelling, and root thee out of the land of the living.\par
\psanum{7}The righteous also shall see this, and fear : and shall laugh him to scorn;\par
\psanum{8}Lo, this is the man that took not God for his strength : but trusted unto the multitude of his riches, and strengthened himself in his wickedness.\par
\psanum{9}As for me, I am like a green olive-tree in the house of God : my trust is in the tender mercy of God for ever and ever.\par
\psanum{10}I will always give thanks unto thee for that thou hast done : and I will hope in thy Name, for thy saints like it well.\par

\psalmsec{Day 10. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 10. Evening Prayer}}

\subbypsalm{Psalm 53. \textit{Dixit insipiens}}
\fancyhead[RE,LO]{Psalm 53}

\lett{T}{he} foolish body hath said in his heart : There is no God.\par
\secondlinepsalm{Corrupt are they, and become abominable in their wickedness : there is none that doeth good.}
\thirdlinepsalm{God looked down from heaven upon the children of men : to see if there were any that would understand, and seek after God.}
\psanum{4}But they are all gone out of the way, they are altogether become abominable : there is also none that doeth good, no not one.\par
\psanum{5}Are not they without understanding, that work wickedness : eating up my people as if they would eat bread? they have not called upon God.\par
\psanum{6}They were afraid where no fear was : for God hath broken the bones of him that besieged thee; thou hast put them to confusion, because God hath despised them.\par
\psanum{7}O that the salvation were given unto Israel out of Sion : O that the Lord would deliver his people out of captivity!\par
\psanum{8}Then should Jacob rejoice : and Israel should be right glad.\par

\subbypsalm{Psalm 54. \textit{Deus, in nomine}}
\fancyhead[RE,LO]{Psalm 54}

\lett{S}{ave} me, O God, for thy Name's sake : and avenge me in thy strength.\par
\secondlinepsalm{Hear my prayer, O God : and hearken unto the words of my mouth.}
\thirdlinepsalm{For strangers are risen up against me : and tyrants, which have not God before their eyes, seek after my soul.}
\psanum{4}Behold, God is my helper : the Lord is with them that uphold my soul.\par
\psanum{5}He shall reward evil unto mine enemies : destroy thou them in thy truth.\par
\psanum{6}An offering of a free heart will I give thee, and praise thy Name, O Lord : because it is so comfortable.\par
\psanum{7}For he hath delivered me out of all my trouble : and mine eye hath seen his desire upon mine enemies.\par

\subbypsalm{Psalm 55. \textit{Exaudi, Deus}}
\fancyhead[RE,LO]{Psalm 55}

\lett{H}{ear} my prayer, O God : and hide not thyself from my petition.\par
\secondlinepsalm{Take heed unto me, and hear me : how I mourn in my prayer, and am vexed.}
\thirdlinepsalm{The enemy crieth so, and the ungodly cometh on so fast : for they are minded to do me some mischief; so maliciously are they set against me.}
\psanum{4}My heart is disquieted within me : and the fear of death is fallen upon me.\par
\psanum{5}Fearfulness and trembling are come upon me : and an horrible dread hath overwhelmed me.\par
\psanum{6}And I said, O that I had wings like a dove : for then would I flee away, and be at rest.\par
\psanum{7}Lo, then would I get me away far off : and remain in the wilderness.\par
\psanum{8}I would make haste to escape : because of the stormy wind and tempest.\par
\psanum{9}Destroy their tongues, O Lord, and divide them : for I have spied unrighteousness and strife in the city.\par
\psanum{10}Day and night they go about within the walls thereof : mischief also and sorrow are in the midst of it.\par
\psanum{11}Wickedness is therein : deceit and guile go not out of their streets.\par
\psanum{12}For it is not an open enemy, that hath done me this dishonour : for then I could have borne it.\par
\psanum{13}Neither was it mine adversary, that did magnify himself against me : for then peradventure I would have hid myself from him.\par
\psanum{14}But it was even thou, my companion : my guide, and mine own familiar friend.\par
\psanum{15}We took sweet counsel together : and walked in the house of God as friends.\par
\psanum{16}Let death come hastily upon them, and let them go down quick into hell : for wickedness is in their dwellings, and among them.\par
\psanum{17}As for me, I will call upon God : and the Lord shall save me.\par
\psanum{18}In the evening, and morning, and at noonday will I pray, and that instantly : and he shall hear my voice.\par
\psanum{19}It is he that hath delivered my soul in peace from the battle that was against me : for there were many with me.\par
\psanum{20}Yea, even God, that endureth for ever, shall hear me, and bring them down : for they will not turn, nor fear God.\par
\psanum{21}He laid his hands upon such as be at peace with him : and he brake his covenant.\par
\psanum{22}The words of his mouth were softer than butter, having war in his heart : his words were smoother than oil, and yet be they very swords.\par
\psanum{23}O cast thy burden upon the Lord, and he shall nourish thee : and shall not suffer the righteous to fall for ever.\par
\psanum{24}And as for them : thou, O God, shalt bring them into the pit of destruction.\par
\psanum{25}The blood-thirsty and deceitful men shall not live out half their days : nevertheless, my trust shall be in thee, O Lord.\par

\psalmsec{Day 11. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 11. Morning Prayer}}

\subbypsalm{Psalm 56. \textit{Miserere mei, Deus}}
\fancyhead[RE,LO]{Psalm 56}

\lett{B}{e} merciful unto me, O God, for man goeth about to devour me : he is daily fighting, and troubling me.\par
\secondlinepsalm{Mine enemies are daily in hand to swallow me up : for they be many that fight against me, O thou most Highest.}
\thirdlinepsalm{Nevertheless, though I am sometime afraid : yet put I my trust in thee.}
\psanum{4}I will praise God, because of his word : I have put my trust in God, and will not fear what flesh can do unto me.\par
\psanum{5}They daily mistake my words : all that they imagine is to do me evil.\par
\psanum{6}They hold all together, and keep themselves close : and mark my steps, when they lay wait for my soul.\par
\psanum{7}Shall they escape for their wickedness : thou, O God, in thy displeasure shalt cast them down.\par
\psanum{8}Thou tellest my flittings; put my tears into thy bottle : are not these things noted in thy book?\par
\psanum{9}Whensoever I call upon thee, then shall mine enemies be put to flight : this I know; for God is on my side.\par
\psanum{10}In God's word I will rejoice : in the Lord's word will I comfort me.\par
\psanum{11}Yea, in God have I put my trust : I will not be afraid what man can do unto me.\par
\psanum{12}Unto thee, O God, will I pay my vows : unto thee will I give thanks.\par
\psanum{13}For thou hast delivered my soul from death, and my feet from falling : that I may walk before God in the light of the living.\par

\subbypsalm{Psalm 57. \textit{Miserere mei, Deus}}
\fancyhead[RE,LO]{Psalm 57}

\lett{B}{e} merciful unto me, O God, be merciful unto me, for my soul trusteth in thee : and under the shadow of thy wings shall be my refuge, until this tyranny be over-past.\par
\secondlinepsalm{I will call unto the most high God : even unto the God that shall perform the cause which I have in hand.}
\thirdlinepsalm{He shall send from heaven : and save me from the reproof of him that would eat me up.}
\psanum{4}God shall send forth his mercy and truth : my soul is among lions.\par
\psanum{5}And I lie even among the children of men, that are set on fire : whose teeth are spears and arrows, and their tongue a sharp sword.\par
\psanum{6}Set up thyself, O God, above the heavens : and thy glory above all the earth.\par
\psanum{7}They have laid a net for my feet, and pressed down my soul : they have digged a pit before me, and are fallen into the midst of it themselves.\par
\psanum{8}My heart is fixed, O God, my heart is fixed : I will sing, and give praise.\par
\psanum{9}Awake up, my glory; awake, lute and harp : I myself will awake right early.\par
\psanum{10}I will give thanks unto thee, O Lord, among the people : and I will sing unto thee among the nations.\par
\psanum{11}For the greatness of thy mercy reacheth unto the heavens : and thy truth unto the clouds.\par
\psanum{12}Set up thyself, O God, above the heavens : and thy glory above all the earth.\par

\subbypsalm{Psalm 58. \textit{Si vere utique}}
\fancyhead[RE,LO]{Psalm 58}

\lett{A}{re} your minds set upon righteousness, O ye congregation : and do ye judge the thing that is right, O ye sons of men?\par
\secondlinepsalm{Yea, ye imagine mischief in your heart upon the earth : and your hands deal with wickedness.}
\thirdlinepsalm{The ungodly are froward, even from their mother's womb : as soon as they are born, they go astray, and speak lies.}
\psanum{4}They are as venomous as the poison of a serpent : even like the deaf adder that stoppeth her ears;\par
\psanum{5}Which refuseth to hear the voice of the charmer : charm he never so wisely.\par
\psanum{6}Break their teeth, O God, in their mouths; smite the jaw-bones of the lions, O Lord : let them fall away like water that runneth apace; and when they shoot their arrows let them be rooted out.\par
\psanum{7}Let them consume away like a snail, and be like the untimely fruit of a woman : and let them not see the sun.\par
\psanum{8}Or ever your pots be made hot with thorns : so let indignation vex him, even as a thing that is raw.\par
\psanum{9}The righteous shall rejoice when he seeth the vengeance : he shall wash his footsteps in the blood of the ungodly.\par
\psanum{10}So that a man shall say, Verily there is a reward for the righteous : doubtless there is a God that judgeth the earth.\par

\psalmsec{Day 11. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 11. Evening Prayer}}

\subbypsalm{Psalm 59. \textit{Eripe me de inimicis}}
\fancyhead[RE,LO]{Psalm 59}

\lett{D}{eliver} me from mine enemies, O God : defend me from them that rise up against me.\par
\secondlinepsalm{O deliver me from the wicked doers : and save me from the blood-thirsty men.}
\thirdlinepsalm{For lo, they lie waiting for my soul : the mighty men are gathered against me, without any offence or fault of me, O Lord.}
\psanum{4}They run and prepare themselves without my fault : arise thou therefore to help me, and behold.\par
\psanum{5}Stand up, O Lord God of hosts, thou God of Israel, to visit all the heathen : and be not merciful unto them that offend of malicious wickedness.\par
\psanum{6}They go to and fro in the evening : they grin like a dog, and run about through the city.\par
\psanum{7}Behold, they speak with their mouth, and swords are in their lips : for who doth hear?\par
\psanum{8}But thou, O Lord, shalt have them in derision : and thou shalt laugh all the heathen to scorn.\par
\psanum{9}My strength will I ascribe unto thee : for thou art the God of my refuge.\par
\psanum{10}God sheweth me his goodness plenteously : and God shall let me see my desire upon mine enemies.\par
\psanum{11}Slay them not, lest my people forget it : but scatter them abroad among the people, and put them down, O Lord, our defence.\par
\psanum{12}For the sin of their mouth, and for the words of their lips, they shall be taken in their pride : and why? their preaching is of cursing and lies.\par
\psanum{13}Consume them in thy wrath, consume them, that they may perish : and know that it is God that ruleth in Jacob, and unto the ends of the world.\par
\psanum{14}And in the evening they will return : grin like a dog, and will go about the city.\par
\psanum{15}They will run here and there for meat : and grudge if they be not satisfied.\par
\psanum{16}As for me, I will sing of thy power, and will praise thy mercy betimes in the morning : for thou hast been my defence and refuge in the day of my trouble.\par
\psanum{17}Unto thee, O my strength, will I sing : for thou, O God, art my refuge, and my merciful God.\par

\subbypsalm{Psalm 60. \textit{Deus, repulisti nos}}
\fancyhead[RE,LO]{Psalm 60}

\lett{O}{God,} thou hast cast us out, and scattered us abroad : thou hast also been displeased; O turn thee unto us again.\par
\secondlinepsalm{Thou hast moved the land, and divided it : heal the sores thereof, for it shaketh.}
\thirdlinepsalm{Thou hast shewed thy people heavy things : thou hast given us a drink of deadly wine.}
\psanum{4}Thou hast given a token for such as fear thee : that they may triumph because of the truth.\par
\psanum{5}Therefore were thy beloved delivered : help me with thy right hand, and hear me.\par
\psanum{6}God hath spoken in his holiness, I will rejoice, and divide Sichem : and mete out the valley of Succoth.\par
\psanum{7}Gilead is mine, and Manasses is mine : Ephraim also is the strength of my head; Judah is my law-giver;\par
\psanum{8}Moab is my wash-pot; over Edom will I cast out my shoe : Philistia, be thou glad of me.\par
\psanum{9}Who will lead me into the strong city : who will bring me into Edom?\par
\psanum{10}Hast not thou cast us out, O God : wilt not thou, O God, go out with our hosts?\par
\psanum{11}O be thou our help in trouble : for vain is the help of man.\par
\psanum{12}Through God will we do great acts : for it is he that shall tread down our enemies.\par

\subbypsalm{Psalm 61. \textit{Exaudi, Deus}}
\fancyhead[RE,LO]{Psalm 61}

\lett{H}{ear} my crying, O God : give ear unto my prayer.\par
\secondlinepsalm{From the ends of the earth will I call upon thee : when my heart is in heaviness.}
\thirdlinepsalm{O set me up upon the rock that is higher than I : for thou hast been my hope, and a strong tower for me against the enemy.}
\psanum{4}I will dwell in thy tabernacle for ever : and my trust shall be under the covering of thy wings.\par
\psanum{5}For thou, O Lord, hast heard my desires : and hast given an heritage unto those that fear thy Name.\par
\psanum{6}Thou shalt grant the King a long life : that his years may endure throughout all generations.\par
\psanum{7}He shall dwell before God for ever : O prepare thy loving mercy and faithfulness, that they may preserve him.\par
\psanum{8}So will I always sing praise unto thy Name : that I may daily perform my vows.\par

\psalmsec{Day 12. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 12. Morning Prayer}}

\subbypsalm{Psalm 62. \textit{Nonne Deo?}}
\fancyhead[RE,LO]{Psalm 62}

\lett{M}{y} soul truly waiteth still upon God : for of him cometh my salvation.\par
\secondlinepsalm{He verily is my strength and my salvation : he is my defence, so that I shall not greatly fall.}
\thirdlinepsalm{How long will ye imagine mischief against every man : ye shall be slain all the sort of you; yea, as a tottering wall shall ye be, and like a broken hedge.}
\psanum{4}Their device is only how to put him out whom God will exalt : their delight is in lies; they give good words with their mouth, but curse with their heart.\par
\psanum{5}Nevertheless, my soul, wait thou still upon God : for my hope is in him.\par
\psanum{6}He truly is my strength and my salvation : he is my defence, so that I shall not fall.\par
\psanum{7}In God is my health, and my glory : the rock of my might, and in God is my trust.\par
\psanum{8}O put your trust in him alway, ye people : pour out your hearts before him, for God is our hope.\par
\psanum{9}As for the children of men, they are but vanity : the children of men are deceitful upon the weights, they are altogether lighter than vanity itself.\par
\psanum{10}O trust not in wrong and robbery, give not yourselves unto vanity : if riches increase, set not your heart upon them.\par
\psanum{11}God spake once, and twice I have also heard the same : that power belongeth unto God;\par
\psanum{12}And that thou, Lord, art merciful : for thou rewardest every man according to his work.\par

\subbypsalm{Psalm 63. \textit{Deus, Deus meus}}
\fancyhead[RE,LO]{Psalm 63}

\lett{O}{God,} thou art my God : early will I seek thee.\par
\secondlinepsalm{My soul thirsteth for thee, my flesh also longeth after thee : in a barren and dry land where no water is.}
\thirdlinepsalm{Thus have I looked for thee in holiness : that I might behold thy power and glory.}
\psanum{4}For thy loving-kindness is better than the life itself : my lips shall praise thee.\par
\psanum{5}As long as I live will I magnify thee on this manner : and lift up my hands in thy Name.\par
\psanum{6}My soul shall be satisfied, even as it were with marrow and fatness : when my mouth praiseth thee with joyful lips.\par
\psanum{7}Have I not remembered thee in my bed : and thought upon thee when I was waking?\par
\psanum{8}Because thou hast been my helper : therefore under the shadow of thy wings will I rejoice.\par
\psanum{9}My soul hangeth upon thee : thy right hand hath upholden me.\par
\psanum{10}These also that seek the hurt of my soul : they shall go under the earth.\par
\psanum{11}Let them fall upon the edge of the sword : that they may be a portion for foxes.\par
\psanum{12}But the King shall rejoice in God; all they also that swear by him shall be commended : for the mouth of them that speak lies shall be stopped.\par

\subbypsalm{Psalm 64. \textit{Exaudi, Deus}}
\fancyhead[RE,LO]{Psalm 64}

\lett{H}{ear} my voice, O God, in my prayer : preserve my life from fear of the enemy.\par
\secondlinepsalm{Hide me from the gathering together of the froward : and from the insurrection of wicked doers;}
\thirdlinepsalm{Who have whet their tongue like a sword : and shoot out their arrows, even bitter words;}
\psanum{4}That they may privily shoot at him that is perfect : suddenly do they hit him, and fear not.\par
\psanum{5}They encourage themselves in mischief : and commune among themselves how they may lay snares, and say that no man shall see them.\par
\psanum{6}They imagine wickedness, and practise it : that they keep secret among themselves, every man in the deep of his heart.\par
\psanum{7}But God shall suddenly shoot at them with a swift arrow : that they shall be wounded.\par
\psanum{8}Yea, their own tongues shall make them fall : insomuch that whoso seeth them shall laugh them to scorn.\par
\psanum{9}And all men that see it shall say, This hath God done : for they shall perceive that it is his work.\par
\psanum{10}The righteous shall rejoice in the Lord, and put his trust in him : and all they that are true of heart shall be glad.\par

\psalmsec{Day 12. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 12. Evening Prayer}}

\subbypsalm{Psalm 65. \textit{Te decet hymnus}}
\fancyhead[RE,LO]{Psalm 65}

\lett{T}{hou}, O God, art praised in Sion : and unto thee shall the vow be performed in Jerusalem.\par
\secondlinepsalm{Thou that hearest the prayer : unto thee shall all flesh come.}
\thirdlinepsalm{My misdeeds prevail against me : O be thou merciful unto our sins.}
\psanum{4}Blessed is the man whom thou choosest, and receivest unto thee : he shall dwell in thy court, and shall be satisfied with the pleasures of thy house, even of thy holy temple.\par
\psanum{5}Thou shalt shew us wonderful things in thy righteousness, O God of our salvation : thou that art the hope of all the ends of the earth, and of them that remain in the broad sea.\par
\psanum{6}Who in his strength setteth fast the mountains : and is girded about with power.\par
\psanum{7}Who stilleth the raging of the sea : and the noise of his waves, and the madness of the people.\par
\psanum{8}They also that dwell in the uttermost parts of the earth shall be afraid at thy tokens : thou that makest the outgoings of the morning and evening to praise thee.\par
\psanum{9}Thou visitest the earth, and blessest it : thou makest it very plenteous.\par
\psanum{10}The river of God is full of water : thou preparest their corn, for so thou providest for the earth.\par
\psanum{11}Thou waterest her furrows, thou sendest rain into the little valleys thereof : thou makest it soft with the drops of rain, and blessest the increase of it.\par
\psanum{12}Thou crownest the year with thy goodness : and thy clouds drop fatness.\par
\psanum{13}They shall drop upon the dwellings of the wilderness : and the little hills shall rejoice on every side.\par
\psanum{14}The folds shall be full of sheep : the valleys also shall stand so thick with corn, that they shall laugh and sing.\par

\subbypsalm{Psalm 66. \textit{Jubilate Deo}}
\fancyhead[RE,LO]{Psalm 66}

\lett{O}{be} joyful in God, all ye lands : sing praises unto the honour of his Name, make his praise to be glorious.\par
\secondlinepsalm{Say unto God, O how wonderful art thou in thy works : through the greatness of thy power shall thine enemies be found liars unto thee.}
\thirdlinepsalm{For all the world shall worship thee : sing of thee, and praise thy Name.}
\psanum{4}O come hither, and behold the works of God : how wonderful he is in his doing toward the children of men.\par
\psanum{5}He turned the sea into dry land : so that they went through the water on foot; there did we rejoice thereof.\par
\psanum{6}He ruleth with his power for ever; his eyes behold the people : and such as will not believe shall not be able to exalt themselves.\par
\psanum{7}O praise our God, ye people : and make the voice of his praise to be heard;\par
\psanum{8}Who holdeth our soul in life : and suffereth not our feet to slip.\par
\psanum{9}For thou, O God, hast proved us : thou also hast tried us, like as silver is tried.\par
\psanum{10}Thou broughtest us into the snare : and laidest trouble upon our loins.\par
\psanum{11}Thou sufferedst men to ride over our heads : we went through fire and water, and thou broughtest us out into a wealthy place.\par
\psanum{12}I will go into thine house with burnt-offerings : and will pay thee my vows, which I promised with my lips, and spake with my mouth, when I was in trouble.\par
\psanum{13}I will offer unto thee fat burnt-sacrifices, with the incense of rams : I will offer bullocks and goats.\par
\psanum{14}O come hither, and hearken, all ye that fear God : and I will tell you what he hath done for my soul.\par
\psanum{15}I called unto him with my mouth : and gave him praises with my tongue.\par
\psanum{16}If I incline unto wickedness with mine heart : the Lord will not hear me.\par
\psanum{17}But God hath heard me : and considered the voice of my prayer.\par
\psanum{18}Praised be God, who hath not cast out my prayer : nor turned his mercy from me.\par

\subbypsalm{Psalm 67. \textit{Deus misereatur}}
\fancyhead[RE,LO]{Psalm 67}

\lett{G}{od} be merciful unto us, and bless us : and shew us the light of his countenance, and be merciful unto us;\par
\secondlinepsalm{That thy way may be known upon earth : thy saving health among all nations.}
\thirdlinepsalm{Let the people praise thee, O God : yea, let all the people praise thee.}
\psanum{4}O let the nations rejoice and be glad : for thou shalt judge the folk righteously, and govern the nations upon earth.\par
\psanum{5}Let the people praise thee, O God : let all the people praise thee.\par
\psanum{6}Then shall the earth bring forth her increase : and God, even our own God, shall give us his blessing.\par
\psanum{7}God shall bless us : and all the ends of the world shall fear him.\par

\psalmsec{Day 13. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 13. Morning Prayer}}

\subbypsalm{Psalm 68. \textit{Exurgat Deus}}
\fancyhead[RE,LO]{Psalm 68}

\lett{L}{et} God arise, and let his enemies be scattered : let them also that hate him flee before him.\par
\secondlinepsalm{Like as the smoke vanisheth, so shalt thou drive them away : and like as wax melteth at the fire, so let the ungodly perish at the presence of God.}
\thirdlinepsalm{But let the righteous be glad and rejoice before God : let them also be merry and joyful.}
\psanum{4}O sing unto God, and sing praises unto his Name : magnify him that rideth upon the heavens, as it were upon an horse; praise him in his Name \textsc{Jah}, and rejoice before him.\par
\psanum{5}He is a father of the fatherless, and defendeth the cause of the widows : even God in his holy habitation.\par
\psanum{6}He is the God that maketh men to be of one mind in an house, and bringeth the prisoners out of captivity : but letteth the runagates continue in scarceness.\par
\psanum{7}O God, when thou wentest forth before the people : when thou wentest through the wilderness;\par
\psanum{8}The earth shook, and the heavens dropped at the presence of God : even as Sinai also was moved at the presence of God, who is the God of Israel.\par
\psanum{9}Thou, O God, sentest a gracious rain upon thine inheritance : and refreshedst it when it was weary.\par
\psanum{10}Thy congregation shall dwell therein : for thou, O God, hast of thy goodness prepared for the poor.\par
\psanum{11}The Lord gave the word : great was the company of the preachers.\par
\psanum{12}Kings with their armies did flee, and were discomfited : and they of the household divided the spoil.\par
\psanum{13}Though ye have lien among the pots, yet shall ye be as the wings of a dove : that is covered with silver wings, and her feathers like gold.\par
\psanum{14}When the Almighty scattered kings for their sake : then were they as white as snow in Salmon.\par
\psanum{15}As the hill of Basan, so is God's hill : even an high hill, as the hill of Basan.\par
\psanum{16}Why hop ye so, ye high hills? this is God's hill, in the which it pleaseth him to dwell : yea, the Lord will abide in it for ever.\par
\psanum{17}The chariots of God are twenty thousand, even thousands of angels : and the Lord is among them, as in the holy place of Sinai.\par
\psanum{18}Thou art gone up on high, thou hast led captivity captive, and received gifts for men : yea, even for thine enemies, that the Lord God might dwell among them.\par
\psanum{19}Praised be the Lord daily : even the God who helpeth us, and poureth his benefits upon us.\par
\psanum{20}He is our God, even the God of whom cometh salvation : God is the Lord, by whom we escape death.\par
\psanum{21}God shall wound the head of his enemies : and the hairy scalp of such a one as goeth on still in his wickedness.\par
\psanum{22}The Lord hath said, I will bring my people again, as I did from Basan : mine own will I bring again, as I did sometime from the deep of the sea.\par
\psanum{23}That thy foot may be dipped in the blood of thine enemies : and that the tongue of thy dogs may be red through the same.\par
\psanum{24}It is well seen, O God, how thou goest : how thou, my God and King, goest in the sanctuary.\par
\psanum{25}The singers go before, the minstrels follow after : in the midst are the damsels playing with the timbrels.\par
\psanum{26}Give thanks, O Israel, unto God the Lord in the congregations : from the ground of the heart.\par
\psanum{27}There is little Benjamin their ruler, and the princes of Judah their counsel : the princes of Zabulon, and the princes of Nephthali.\par
\psanum{28}Thy God hath sent forth strength for thee : stablish the thing, O God, that thou hast wrought in us,\par
\psanum{29}For thy temple's sake at Jerusalem : so shall kings bring presents unto thee.\par
\psanum{30}When the company of the spear-men, and multitude of the mighty are scattered abroad among the beasts of the people, so that they humbly bring pieces of silver : and when he hath scattered the people that delight in war;\par
\psanum{31}Then shall the princes come out of Egypt : the Morians' land shall soon stretch our her hands unto God.\par
\psanum{32}Sing unto God, O ye kingdoms of the earth : O sing praises unto the Lord;\par
\psanum{33}Who sitteth in the heavens over all from the beginning : lo, he doth send out his voice, yea, and that a mighty voice.\par
\psanum{34}Ascribe ye the power to God over Israel : his worship and strength is in the clouds.\par
\psanum{35}O God, wonderful art thou in thy holy places : even the God of Israel, he will give strength and power unto his people; blessed be God.\par

\psalmsec{Day 13. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 13. Evening Prayer}}

\subbypsalm{Psalm 69. \textit{Salvum me fac}}
\fancyhead[RE,LO]{Psalm 69}

\lett{S}{ave} me, O God : for the waters are come in, even unto my soul.\par
\secondlinepsalm{I stick fast in the deep mire, where no ground is : I am come into deep waters, so that the floods run over me.}
\thirdlinepsalm{I am weary of crying; my throat is dry : my sight faileth me for waiting so long upon my God.}
\psanum{4}They that hate me without a cause are more than the hairs of my head : they that are mine enemies, and would destroy me guiltless, are mighty.\par
\psanum{5}I paid them the things that I never took : God, thou knowest my simpleness, and my faults are not hid from thee.\par
\psanum{6}Let not them that trust in thee, O Lord God of hosts, be ashamed for my cause : let not those that seek thee be confounded through me, O Lord God of Israel.\par
\psanum{7}And why? for thy sake have I suffered reproof : shame hath covered my face.\par
\psanum{8}I am become a stranger unto my brethren : even an alien unto my mother's children.\par
\psanum{9}For the zeal of thine house hath even eaten me : and the rebukes of them that rebuked thee are fallen upon me.\par
\psanum{10}I wept, and chastened myself with fasting : and that was turned to my reproof.\par
\psanum{11}I put on sackcloth also : and they jested upon me.\par
\psanum{12}They that sit in the gate speak against me : and the drunkards make songs upon me.\par
\psanum{13}But, Lord, I make my prayer unto thee : in an acceptable time.\par
\psanum{14}Hear me, O God, in the multitude of thy mercy : even in the truth of thy salvation.\par
\psanum{15}Take me out of the mire, that I sink not : O let me be delivered from them that hate me, and out of the deep waters.\par
\psanum{16}Let not the water-flood drown me, neither let the deep swallow me up : and let not the pit shut her mouth upon me.\par
\psanum{17}Hear me, O Lord, for thy loving-kindness is comfortable : turn thee unto me according to the multitude of thy mercies.\par
\psanum{18}And hide not thy face from thy servant, for I am in trouble : O haste thee, and hear me.\par
\psanum{19}Draw nigh unto my soul, and save it : O deliver me, because of mine enemies.\par
\psanum{20}Thou hast known my reproof, my shame, and my dishonour : mine adversaries are all in thy sight.\par
\psanum{21}Thy rebuke hath broken my heart; I am full of heaviness : I looked for some to have pity on me, but there was no man, neither found I any to comfort me.\par
\psanum{22}They gave me gall to eat : and when I was thirsty they gave me vinegar to drink.\par
\psanum{23}Let their table be made a snare to take themselves withal : and let the things that should have been for their wealth be unto them an occasion of falling.\par
\psanum{24}Let their eyes be blinded, that they see not : and ever bow thou down their backs.\par
\psanum{25}Pour out thine indignation upon them : and let thy wrathful displeasure take hold of them.\par
\psanum{26}Let their habitation be void : and no man to dwell in their tents.\par
\psanum{27}For they persecute him whom thou hast smitten : and they talk how they may vex them whom thou hast wounded.\par
\psanum{28}Let them fall from one wickedness to another : and not come into thy righteousness.\par
\psanum{29}Let them be wiped out of the book of the living : and not be written among the righteous.\par
\psanum{30}As for me, when I am poor and in heaviness : thy help, O God, shall lift me up.\par
\psanum{31}I will praise the Name of God with a song : and magnify it with thanksgiving.\par
\psanum{32}This also shall please the Lord : better than a bullock that hath horns and hoofs.\par
\psanum{33}The humble shall consider this, and be glad : seek ye after God, and your soul shall live.\par
\psanum{34}For the Lord heareth the poor : and despiseth not his prisoners.\par
\psanum{35}Let heaven and earth praise him : the sea, and all that moveth therein.\par
\psanum{36}For God will save Sion, and build the cities of Judah : that men may dwell there, and have it in possession.\par
\psanum{37}The posterity also of his servants shall inherit it : and they that love his Name shall dwell therein.\par

\subbypsalm{Psalm 70. \textit{Deus, in adjutorium}}
\fancyhead[RE,LO]{Psalm 70}

\lett{H}{aste} thee, O God, to deliver me : make haste to help me, O Lord.\par
\secondlinepsalm{Let them be ashamed and confounded that seek after my soul : let them be turned backward and put to confusion that wish me evil.}
\thirdlinepsalm{Let them for their reward be soon brought to shame : that cry over me, There, there.}
\psanum{4}But let all those that seek thee be joyful and glad in thee : and let all such as delight in thy salvation say alway, The Lord be praised.\par
\psanum{5}As for me, I am poor and in misery : haste thee unto me, O God.\par
\psanum{6}Thou art my helper and my redeemer : O Lord, make no long tarrying.\par

\psalmsec{Day 14. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 14. Morning Prayer}}

\subbypsalm{Psalm 71. \textit{In te, Domine, speravi}}
\fancyhead[RE,LO]{Psalm 71}

\lett{I}{n} thee, O Lord, have I put my trust, let me never be put to confusion : but rid me and deliver me in thy righteousness, incline thine ear unto me, and save me.\par
\secondlinepsalm{Be thou my strong hold, whereunto I may alway resort : thou hast promised to help me, for thou art my house of defence and my castle.}
\thirdlinepsalm{Deliver me, O my God, out of the hand of the ungodly : out of the hand of the unrighteous and cruel man.}
\psanum{4}For thou, O Lord God, art the thing that I long for : thou art my hope, even from my youth.\par
\psanum{5}Through thee have I been holden up ever since I was born : thou art he that took me out of my mother's womb; my praise shall be always of thee.\par
\psanum{6}I am become as it were a monster unto many : but my sure trust is in thee.\par
\psanum{7}O let my mouth be filled with thy praise : that I may sing of thy glory and honour all the day long.\par
\psanum{8}Cast me not away in the time of age : forsake me not when my strength faileth me.\par
\psanum{9}For mine enemies speak against me, and they that lay wait for my soul take their counsel together, saying : God hath forsaken him; persecute him, and take him, for there is none to deliver him.\par
\psanum{10}Go not far from me, O God : my God, haste thee to help me.\par
\psanum{11}Let them be confounded and perish that are against my soul : let them be covered with shame and dishonour that seek to do me evil.\par
\psanum{12}As for me, I will patiently abide alway : and will praise thee more and more.\par
\psanum{13}My mouth shall daily speak of thy righteousness and salvation : for I know no end thereof.\par
\psanum{14}I will go forth in the strength of the Lord God : and will make mention of thy righteousness only.\par
\psanum{15}Thou, O God, hast taught me from my youth up until now : therefore will I tell of thy wondrous works.\par
\psanum{16}Forsake me not, O God, in mine old age, when I am gray-headed : until I have shewed thy strength unto this generation, and thy power to all them that are yet for to come.\par
\psanum{17}Thy righteousness, O God, is very high : and great things are they that thou hast done; O God, who is like unto thee?\par
\psanum{18}O what great troubles and adversities hast thou shewed me, and yet didst thou turn and refresh me : yea, and broughtest me from the deep of the earth again.\par
\psanum{19}Thou hast brought me to great honour : and comforted me on every side.\par
\psanum{20}Therefore will I praise thee and thy faithfulness, O God, playing upon an instrument of musick : unto thee will I sing upon the harp, O thou Holy One of Israel.\par
\psanum{21}My lips will be fain when I sing unto thee : and so will my soul whom thou hast delivered.\par
\psanum{22}My tongue also shall talk of thy righteousness all the day long : for they are confounded and brought unto shame that seek to do me evil.\par

\subbypsalm{Psalm 72. \textit{Deus, judicium}}
\fancyhead[RE,LO]{Psalm 72}

\lett{G}{ive} the King thy judgements, O God : and thy righteousness unto the King's son.\par
\secondlinepsalm{Then shall he judge thy people according unto right : and defend the poor.}
\thirdlinepsalm{The mountains also shall bring peace : and the little hills righteousness unto the people.}
\psanum{4}He shall keep the simple folk by their right : defend the children of the poor, and punish the wrong-doer.\par
\psanum{5}They shall fear thee, as long as the sun and moon endureth : from one generation to another.\par
\psanum{6}He shall come down like the rain into a fleece of wool : even as the drops that water the earth.\par
\psanum{7}In his time shall the righteous flourish : yea, and abundance of peace, so long as the moon endureth.\par
\psanum{8}His dominion shall be also from the one sea to the other : and from the flood unto the world's end.\par
\psanum{9}They that dwell in the wilderness shall kneel before him : his enemies shall lick the dust.\par
\psanum{10}The kings of Tharsis and of the isles shall give presents : the kings of Arabia and Saba shall bring gifts.\par
\psanum{11}All kings shall fall down before him : all nations shall do him service.\par
\psanum{12}For he shall deliver the poor when he crieth : the needy also, and him that hath no helper.\par
\psanum{13}He shall be favourable to the simple and needy : and shall preserve the souls of the poor.\par
\psanum{14}He shall deliver their souls from falsehood and wrong : and dear shall their blood be in his sight.\par
\psanum{15}He shall live, and unto him shall be given of the gold of Arabia : prayer shall be made ever unto him, and daily shall he be praised.\par
\psanum{16}There shall be an heap of corn in the earth, high upon the hills : his fruit shall shake like Libanus, and shall be green in the city like grass upon the earth.\par
\psanum{17}His Name shall endure for ever; his Name shall remain under the sun among the posterities : which shall be blessed through him; and all the heathen shall praise him.\par
\psanum{18}Blessed be the Lord God, even the God of Israel : which only doeth wondrous things;\par
\psanum{19}And blessed be the Name of his majesty for ever : and all the earth shall be filled with his majesty. Amen, Amen.\par

\psalmsec{Day 14. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 14. Evening Prayer}}

\subbypsalm{Psalm 73. \textit{Quam bonus Israel!}}
\fancyhead[RE,LO]{Psalm 73}

\lett{T}{ruly} God is loving unto Israel : even unto such as are of a clean heart.\par
\secondlinepsalm{Nevertheless, my feet were almost gone : my treadings had well-nigh slipt.}
\thirdlinepsalm{And why? I was grieved at the wicked : I do also see the ungodly in such prosperity.}
\psanum{4}For they are in no peril of death : but are lusty and strong.\par
\psanum{5}They come in no misfortune like other folk : neither are they plagued like other men.\par
\psanum{6}And this is the cause that they are so holden with pride : and overwhelmed with cruelty.\par
\psanum{7}Their eyes swell with fatness : and they do even what they lust.\par
\psanum{8}They corrupt other, and speak of wicked blasphemy : their talking is against the most High.\par
\psanum{9}For they stretch forth their mouth unto the heaven : and their tongue goeth through the world.\par
\psanum{10}Therefore fall the people unto them : and thereout suck they no small advantage.\par
\psanum{11}Tush, say they, how should God perceive it : is there knowledge in the most High?\par
\psanum{12}Lo, these are the ungodly, these prosper in the world, and these have riches in possession : and I said, Then have I cleansed my heart in vain, and washed mine hands in innocency.\par
\psanum{13}All the day long have I been punished : and chastened every morning.\par
\psanum{14}Yea, and I had almost said even as they : but lo, then I should have condemned the generation of thy children.\par
\psanum{15}Then thought I to understand this : but it was too hard for me,\par
\psanum{16}Until I went into the sanctuary of God : then understood I the end of these men;\par
\psanum{17}Namely, how thou dost set them in slippery places : and castest them down, and destroyest them.\par
\psanum{18}O how suddenly do they consume : perish, and come to a fearful end!\par
\psanum{19}Yea, even like as a dream when one awaketh : so shalt thou make their image to vanish out of the city.\par
\psanum{20}Thus my heart was grieved : and it went even through my reins.\par
\psanum{21}So foolish was I, and ignorant : even as it were a beast before thee.\par
\psanum{22}Nevertheless, I am alway by thee : for thou hast holden me by my right hand.\par
\psanum{23}Thou shalt guide me with thy counsel : and after that receive me with glory.\par
\psanum{24}Whom have I in heaven but thee : and there is none upon earth that I desire in comparison of thee.\par
\psanum{25}My flesh and my heart faileth : but God is the strength of my heart, and my portion for ever.\par
\psanum{26}For lo, they that forsake thee shall perish : thou hast destroyed all them that commit fornication against thee.\par
\psanum{27}But it is good for me to hold me fast by God, to put my trust in the Lord God : and to speak of all thy works in the gates of the daughter of Sion.\par

\subbypsalm{Psalm 74. \textit{Ut quid, Deus?}}
\fancyhead[RE,LO]{Psalm 74}

\lett{O}{God,} wherefore art thou absent from us so long : why is thy wrath so hot against the sheep of thy pasture?\par
\secondlinepsalm{O think upon thy congregation : whom thou hast purchased and redeemed of old.}
\thirdlinepsalm{Think upon the tribe of thine inheritance : and mount Sion, wherein thou hast dwelt.}
\psanum{4}Lift up thy feet, that thou mayest utterly destroy every enemy : which hath done evil in thy sanctuary.\par
\psanum{5}Thine adversaries roar in the midst of thy congregations : and set up their banners for tokens.\par
\psanum{6}He that hewed timber afore out of the thick trees : was known to bring it to an excellent work.\par
\psanum{7}But now they break down all the carved work thereof : with axes and hammers.\par
\psanum{8}They have set fire upon thy holy places : and have defiled the dwelling-place of thy Name, even unto the ground.\par
\psanum{9}Yea, they said in their hearts, Let us make havock of them altogether : thus have they burnt up all the houses of God in the land.\par
\psanum{10}We see not our tokens, there is not one prophet more : no, not one is there among us, that understandeth any more.\par
\psanum{11}O God, how long shall the adversary do this dishonour : how long shall the enemy blaspheme thy Name, for ever?\par
\psanum{12}Why withdrawest thou thy hand : why pluckest thou not thy right hand out of thy bosom to consume the enemy?\par
\psanum{13}For God is my King of old : the help that is done upon earth he doeth it himself.\par
\psanum{14}Thou didst divide the sea through thy power : thou brakest the heads of the dragons in the waters.\par
\psanum{15}Thou smotest the heads of Leviathan in pieces : and gavest him to be meat for the people in the wilderness.\par
\psanum{16}Thou broughtest out fountains and waters out of the hard rocks : thou driedst up mighty waters.\par
\psanum{17}The day is thine, and the night is thine : thou hast prepared the light and the sun.\par
\psanum{18}Thou hast set all the borders of the earth : thou hast made summer and winter.\par
\psanum{19}Remember this, O Lord, how the enemy hath rebuked : and how the foolish people hath blasphemed thy Name.\par
\psanum{20}O deliver not the soul of thy turtle-dove unto the multitude of the enemies : and forget not the congregation of the poor for ever.\par
\psanum{21}Look upon the covenant : for all the earth is full of darkness and cruel habitations.\par
\psanum{22}O let not the simple go away ashamed : but let the poor and needy give praise unto thy Name.\par
\psanum{23}Arise, O God, maintain thine own cause : remember how the foolish man blasphemeth thee daily.\par
\psanum{24}Forget not the voice of thine enemies : the presumption of them that hate thee increaseth ever more and more.\par

\psalmsec{Day 15. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 15. Morning Prayer}}

\subbypsalm{Psalm 75. \textit{Confitebimur tibi}}
\fancyhead[RE,LO]{Psalm 75}

\lett{U}{nto} thee, O God, do we give thanks : yea, unto thee do we give thanks.\par
\secondlinepsalm{Thy Name also is so nigh : and that do thy wondrous works declare.}
\thirdlinepsalm{When I receive the congregation : I shall judge according unto right.}
\psanum{4}The earth is weak, and all the inhabiters thereof : I bear up the pillars of it.\par
\psanum{5}I said unto the fools, Deal not so madly : and to the ungodly, Set not up your horn.\par
\psanum{6}Set not up your horn on high : and speak not with a stiff neck.\par
\psanum{7}For promotion cometh neither from the east, nor from the west : nor yet from the south.\par
\psanum{8}And why? God is the Judge : he putteth down one, and setteth up another.\par
\psanum{9}For in the hand of the Lord there is a cup, and the wine is red : it is full mixed, and he poureth out of the same.\par
\psanum{10}As for the dregs thereof : all the ungodly of the earth shall drink them, and suck them out.\par
\psanum{11}But I will talk of the God of Jacob : and praise him for ever.\par
\psanum{12}All the horns of the ungodly also will I break : and the horns of the righteous shall be exalted.\par

\subbypsalm{Psalm 76. \textit{Notus in Judaea}}
\fancyhead[RE,LO]{Psalm 76}

\lett{I}{n} Jewry is God known : his Name is great in Israel.\par
\secondlinepsalm{At Salem is his tabernacle : and his dwelling in Sion.}
\thirdlinepsalm{There brake he the arrows of the bow : the shield, the sword, and the battle.}
\psanum{4}Thou art of more honour and might : than the hills of the robbers.\par
\psanum{5}The proud are robbed, they have slept their sleep : and all the men whose hands were mighty have found nothing.\par
\psanum{6}At thy rebuke, O God of Jacob : both the chariot and horse are fallen.\par
\psanum{7}Thou, even thou art to be feared : and who may stand in thy sight when thou art angry?\par
\psanum{8}Thou didst cause thy judgement to be heard from heaven : the earth trembled, and was still;\par
\psanum{9}When God arose to judgement : and to help all the meek upon earth.\par
\psanum{10}The fierceness of man shall turn to thy praise : and the fierceness of them shalt thou refrain.\par
\psanum{11}Promise unto the Lord your God, and keep it, all ye that are round about him : bring presents unto him that ought to be feared.\par
\psanum{12}He shall refrain the spirit of princes : and is wonderful among the kings of the earth.\par

\subbypsalm{Psalm 77. \textit{Voce mea ad Dominum}}
\fancyhead[RE,LO]{Psalm 77}

\lett{I}{will} cry unto God with my voice : even unto God will I cry with my voice, and he shall hearken unto me.\par
\secondlinepsalm{In the time of my trouble I sought the Lord : my sore ran and ceased not in the night-season; my soul refused comfort.}
\thirdlinepsalm{When I am in heaviness, I will think upon God : when my heart is vexed, I will complain.}
\psanum{4}Thou holdest mine eyes waking : I am so feeble, that I cannot speak.\par
\psanum{5}I have considered the days of old : and the years that are past.\par
\psanum{6}I call to remembrance my song : and in the night I commune with mine own heart, and search out my spirits.\par
\psanum{7}Will the Lord absent himself for ever : and will he be no more intreated?\par
\psanum{8}Is his mercy clean gone for ever : and is his promise come utterly to an end for evermore?\par
\psanum{9}Hath God forgotten to be gracious : and will he shut up his loving-kindness in displeasure?\par
\psanum{10}And I said, It is mine own infirmity : but I will remember the years of the right hand of the most Highest.\par
\psanum{11}I will remember the works of the Lord : and call to mind thy wonders of old time.\par
\psanum{12}I will think also of all thy works : and my talking shall be of thy doings.\par
\psanum{13}Thy way, O God, is holy : who is so great a God as our God?\par
\psanum{14}Thou art the God that doeth wonders : and hast declared thy power among the people.\par
\psanum{15}Thou hast mightily delivered thy people : even the sons of Jacob and Joseph.\par
\psanum{16}The waters saw thee, O God, the waters saw thee, and were afraid : the depths also were troubled.\par
\psanum{17}The clouds poured out water, the air thundered : and thine arrows went abroad.\par
\psanum{18}The voice of thy thunder was heard round about : the lightnings shone upon the ground; the earth was moved, and shook withal.\par
\psanum{19}Thy way is in the sea, and thy paths in the great waters : and thy footsteps are not known.\par
\psanum{20}Thou leddest thy people like sheep : by the hand of Moses and Aaron.\par

\psalmsec{Day 15. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 15. Evening Prayer}}

\subbypsalm{Psalm 78. \textit{Attendite, popule}}
\fancyhead[RE,LO]{Psalm 78}

\lett{H}{ear} my law, O my people : incline your ears unto the words of my mouth.\par
\secondlinepsalm{I will open my mouth in a parable : I will declare hard sentences of old;}
\thirdlinepsalm{Which we have heard and known : and such as our fathers have told us;}
\psanum{4}That we should not hide them from the children of the generations to come : but to shew the honour of the Lord, his mighty and wonderful works that he hath done.\par
\psanum{5}He made a covenant with Jacob, and gave Israel a law : which he commanded our forefathers to teach their children;\par
\psanum{6}That their posterity might know it : and the children which were yet unborn;\par
\psanum{7}To the intent that when they came up : they might shew their children the same;\par
\psanum{8}That they might put their trust in God : and not to forget the works of God, but to keep his commandments;\par
\psanum{9}And not to be as their forefathers, a faithless and stubborn generation : a generation that set not their heart aright, and whose spirit cleaveth not stedfastly unto God;\par
\psanum{10}Like as the children of Ephraim : who being harnessed, and carrying bows, turned themselves back in the day of battle.\par
\psanum{11}They kept not the covenant of God : and would not walk in his law;\par
\psanum{12}But forgat what he had done : and the wonderful works that he had shewed for them.\par
\psanum{13}Marvellous things did he in the sight of our forefathers, in the land of Egypt : even in the field of Zoan.\par
\psanum{14}He divided the sea, and let them go through : he made the waters to stand on an heap.\par
\psanum{15}In the day-time also he led them with a cloud : and all the night through with a light of fire.\par
\psanum{16}He clave the hard rocks in the wilderness : and gave them drink thereof, as it had been out of the great depth.\par
\psanum{17}He brought waters out of the stony rock : so that it gushed out like the rivers.\par
\psanum{18}Yet for all this they sinned more against him : and provoked the most Highest in the wilderness.\par
\psanum{19}They tempted God in their hearts : and required meat for their lust.\par
\psanum{20}They spake against God also, saying : Shall God prepare a table in the wilderness?\par
\psanum{21}He smote the stony rock indeed, that the waters gushed out, and the streams flowed withal : but can he give bread also, or provide flesh for his people?\par
\psanum{22}When the Lord heard this, he was wroth : so the fire was kindled in Jacob, and there came up heavy displeasure against Israel;\par
\psanum{23}Because they believed not in God : and put not their trust in his help.\par
\psanum{24}So he commanded the clouds above : and opened the doors of heaven.\par
\psanum{25}He rained down manna also upon them for to eat : and gave them food from heaven.\par
\psanum{26}So man did eat angels' food : for he sent them meat enough.\par
\psanum{27}He caused the east-wind to blow under heaven : and through his power he brought in the south-west-wind.\par
\psanum{28}He rained flesh upon them as thick as dust : and feathered fowls like as the sand of the sea.\par
\psanum{29}He let it fall among their tents : even round about their habitation.\par
\psanum{30}So they did eat and were well filled, for he gave them their own desire : they were not disappointed of their lust.\par
\psanum{31}But while the meat was yet in their mouths, the heavy wrath of God came upon them, and slew the wealthiest of them : yea, and smote down the chosen men that were in Israel.\par
\psanum{32}But for all this they sinned yet more : and believed not his wondrous works.\par
\psanum{33}Therefore their days did he consume in vanity : and their years in trouble.\par
\psanum{34}When he slew them, they sought him : and turned them early, and inquired after God.\par
\psanum{35}And they remembered that God was their strength : and that the high God was their redeemer.\par
\psanum{36}Nevertheless, they did but flatter him with their mouth : and dissembled with him in their tongue.\par
\psanum{37}For their heart was not whole with him : neither continued they stedfast in his covenant.\par
\psanum{38}But he was so merciful, that he forgave their misdeeds : and destroyed them not.\par
\psanum{39}Yea, many a time turned he his wrath away : and would not suffer his whole displeasure to arise.\par
\psanum{40}For he considered that they were but flesh : and that they were even a wind that passeth away, and cometh not again.\par
\psanum{41}Many a time did they provoke him in the wilderness : and grieved him in the desert.\par
\psanum{42}They turned back, and tempted God : and moved the Holy One in Israel.\par
\psanum{43}They thought not of his hand : and of the day when he delivered them from the hand of the enemy;\par
\psanum{44}How he had wrought his miracles in Egypt : and his wonders in the field of Zoan.\par
\psanum{45}He turned their waters into blood : so that they might not drink of the rivers.\par
\psanum{46}He sent lice among them, and devoured them up : and frogs to destroy them.\par
\psanum{47}He gave their fruit unto the caterpillar : and their labour unto the grasshopper.\par
\psanum{48}He destroyed their vines with hail-stones : and their mulberry-trees with the frost.\par
\psanum{49}He smote their cattle also with hail-stones : and their flocks with hot thunderbolts.\par
\psanum{50}He cast upon them the furiousness of his wrath, anger, displeasure, and trouble : and sent evil angels among them.\par
\psanum{51}He made a way to his indignation, and spared not their soul from death : but gave their life over to the pestilence;\par
\psanum{52}And smote all the first-born in Egypt : the most principal and mightiest in the dwellings of Ham.\par
\psanum{53}But as for his own people, he led them forth like sheep : and carried them in the wilderness like a flock.\par
\psanum{54}He brought them out safely, that they should not fear : and overwhelmed their enemies with the sea.\par
\psanum{55}And brought them within the borders of his sanctuary : even to his mountain which he purchased with his right hand.\par
\psanum{56}He cast out the heathen also before them : caused their land to be divided among them for an heritage, and made the tribes of Israel to dwell in their tents.\par
\psanum{57}So they tempted and displeased the most high God : and kept not his testimonies;\par
\psanum{58}But turned their backs, and fell away like their forefathers : starting aside like a broken bow.\par
\psanum{59}For they grieved him with their hill-altars : and provoked him to displeasure with their images.\par
\psanum{60}When God heard this, he was wroth : and took sore displeasure at Israel.\par
\psanum{61}So that he forsook the tabernacle in Silo : even the tent that he had pitched among men.\par
\psanum{62}He delivered their power into captivity : and their beauty into the enemy's hand.\par
\psanum{63}He gave his people over also unto the sword : and was wroth with his inheritance.\par
\psanum{64}The fire consumed their young men : and their maidens were not given to marriage.\par
\psanum{65}Their priests were slain with the sword : and there were no widows to make lamentation.\par
\psanum{66}So the Lord awaked as one out of sleep : and like a giant refreshed with wine.\par
\psanum{67}He smote his enemies in the hinder parts : and put them to a perpetual shame.\par
\psanum{68}He refused the tabernacle of Joseph : and chose not the tribe of Ephraim;\par
\psanum{69}But chose the tribe of Judah : even the hill of Sion which he loved.\par
\psanum{70}And there he built his temple on high : and laid the foundation of it like the ground which he hath made continually.\par
\psanum{71}He chose David also his servant : and took him away from the sheep-folds.\par
\psanum{72}As he was following the ewes great with young ones he took him : that he might feed Jacob his people, and Israel his inheritance.\par
\psanum{73}So he fed them with a faithful and true heart : and ruled them prudently with all his power.\par

\psalmsec{Day 16. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 16. Morning Prayer}}

\subbypsalm{Psalm 79. \textit{Deus, venerunt}}
\fancyhead[RE,LO]{Psalm 79}

\lett{O}{God,} the heathen are come into thine inheritance : thy holy temple have they defiled, and made Jerusalem an heap of stones.\par
\secondlinepsalm{The dead bodies of thy servants have they given to be meat unto the fowls of the air : and the flesh of thy saints unto the beasts of the land.}
\thirdlinepsalm{Their blood have they shed like water on every side of Jerusalem : and there was no man to bury them.}
\psanum{4}We are become an open shame to our enemies : a very scorn and derision unto them that are round about us.\par
\psanum{5}Lord, how long wilt thou be angry : shall thy jealousy burn like fire for ever?\par
\psanum{6}Pour out thine indignation upon the heathen that have not known thee : and upon the kingdoms that have not called upon thy Name.\par
\psanum{7}For they have devoured Jacob : and laid waste his dwelling-place.\par
\psanum{8}O remember not our old sins, but have mercy upon us, and that soon : for we are come to great misery.\par
\psanum{9}Help us, O God of our salvation, for the glory of thy Name : O deliver us, and be merciful unto our sins, for thy Name's sake.\par
\psanum{10}Wherefore do the heathen say : Where is now their God?\par
\psanum{11}O let the vengeance of thy servants' blood that is shed : be openly shewed upon the heathen in our sight.\par
\psanum{12}O let the sorrowful sighing of the prisoners come before thee : according to the greatness of thy power, preserve thou those that are appointed to die.\par
\psanum{13}And for the blasphemy wherewith our neighbours have blasphemed thee : reward thou them, O Lord, seven-fold into their bosom.\par
\psanum{14}So we, that are thy people, and sheep of thy pasture, shall give thee thanks for ever : and will alway be shewing forth thy praise from generation to generation.\par

\subbypsalm{Psalm 80. \textit{Qui regis Israel}}
\fancyhead[RE,LO]{Psalm 80}

\lett{H}{ear,} O thou Shepherd of Israel, thou that leadest Joseph like a sheep : shew thyself also, thou that sittest upon the cherubims.\par
\secondlinepsalm{Before Ephraim, Benjamin, and Manasses : stir up thy strength, and come, and help us.}
\thirdlinepsalm{Turn us again, O God : shew the light of thy countenance, and we shall be whole.}
\psanum{4}O Lord God of hosts : how long wilt thou be angry with thy people that prayeth?\par
\psanum{5}Thou feedest them with the bread of tears : and givest them plenteousness of tears to drink.\par
\psanum{6}Thou hast made us a very strife unto our neighbours : and our enemies laugh us to scorn.\par
\psanum{7}Turn us again, thou God of hosts : shew the light of thy countenance, and we shall be whole.\par
\psanum{8}Thou hast brought a vine out of Egypt : thou hast cast out the heathen, and planted it.\par
\psanum{9}Thou madest room for it : and when it had taken root it filled the land.\par
\psanum{10}The hills were covered with the shadow of it : and the boughs thereof were like the goodly cedar-trees.\par
\psanum{11}She stretched out her branches unto the sea : and her boughs unto the river.\par
\psanum{12}Why hast thou then broken down her hedge : that all they that go by pluck off her grapes?\par
\psanum{13}The wild boar out of the wood doth root it up : and the wild beasts of the field devour it.\par
\psanum{14}Turn thee again, thou God of hosts, look down from heaven : behold, and visit this vine;\par
\psanum{15}And the place of the vineyard that thy right hand hath planted : and the branch that thou madest so strong for thyself.\par
\psanum{16}It is burnt with fire, and cut down : and they shall perish at the rebuke of thy countenance.\par
\psanum{17}Let thy hand be upon the man of thy right hand : and upon the son of man, whom thou madest so strong for thine own self.\par
\psanum{18}And so will not we go back from thee : O let us live, and we shall call upon thy Name.\par
\psanum{19}Turn us again, O Lord God of hosts : shew the light of thy countenance, and we shall be whole.\par

\subbypsalm{Psalm 81. \textit{Exultate Deo}}
\fancyhead[RE,LO]{Psalm 81}

\lett{S}{ing} we merrily unto God our strength : make a cheerful noise unto the God of Jacob.\par
\secondlinepsalm{Take the psalm, bring hither the tabret : the merry harp with the lute.}
\thirdlinepsalm{Blow up the trumpet in the new-moon : even in the time appointed, and upon our solemn feast-day.}
\psanum{4}For this was made a statute for Israel : and a law of the God of Jacob.\par
\psanum{5}This he ordained in Joseph for a testimony : when he came out of the land of Egypt, and had heard a strange language.\par
\psanum{6}I eased his shoulder from the burden : and his hands were delivered from making the pots.\par
\psanum{7}Thou calledst upon me in troubles, and I delivered thee : and heard thee what time as the storm fell upon thee.\par
\psanum{8}I proved thee also : at the waters of strife.\par
\psanum{9}Hear, O my people, and I will assure thee, O Israel : if thou wilt hearken unto me,\par
\psanum{10}There shall no strange god be in thee : neither shalt thou worship any other god.\par
\psanum{11}I am the Lord thy God, who brought thee out of the land of Egypt : open thy mouth wide, and I shall fill it.\par
\psanum{12}But my people would not hear my voice : and Israel would not obey me.\par
\psanum{13}So I gave them up unto their own hearts' lusts : and let them follow their own imaginations.\par
\psanum{14}O that my people would have hearkened unto me : for if Israel had walked in my ways,\par
\psanum{15}I should soon have put down their enemies : and turned my hand against their adversaries.\par
\psanum{16}The haters of the Lord should have been found liars : but their time should have endured for ever.\par
\psanum{17}He should have fed them also with the finest wheat-flour : and with honey out of the stony rock should I have satisfied thee.\par

\psalmsec{Day 16. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 16. Evening Prayer}}

\subbypsalm{Psalm 82. \textit{Deus stetit}}
\fancyhead[RE,LO]{Psalm 82}

\lett{G}{od} standeth in the congregation of princes : he is a Judge among gods.\par
\secondlinepsalm{How long will ye give wrong judgement : and accept the persons of the ungodly?}
\thirdlinepsalm{Defend the poor and fatherless : see that such as are in need and necessity have right.}
\psanum{4}Deliver the outcast and poor : save them from the hand of the ungodly.\par
\psanum{5}They will not be learned nor understand, but walk on still in darkness : all the foundations of the earth are out of course.\par
\psanum{6}I have said, Ye are gods : and ye are all the children of the most Highest.\par
\psanum{7}But ye shall die like men : and fall like one of the princes.\par
\psanum{8}Arise, O God, and judge thou the earth : for thou shalt take all heathen to thine inheritance.\par

\subbypsalm{Psalm 83. \textit{Deus, quis similis?}}
\fancyhead[RE,LO]{Psalm 83}

\lett{H}{old} not thy tongue, O God, keep not still silence : refrain not thyself, O God.\par
\secondlinepsalm{For lo, thine enemies make a murmuring : and they that hate thee have lift up their head.}
\thirdlinepsalm{They have imagined craftily against thy people : and taken counsel against thy secret ones.}
\psanum{4}They have said, Come, and let us root them out, that they be no more a people : and that the name of Israel may be no more in remembrance.\par
\psanum{5}For they have cast their heads together with one consent : and are confederate against thee;\par
\psanum{6}The tabernacles of the Edomites, and the Ismaelites : the Moabites and Hagarenes;\par
\psanum{7}Gebal, and Ammon, and Amalek : the Philistines, with them that dwell at Tyre.\par
\psanum{8}Assur also is joined with them : and have holpen the children of Lot.\par
\psanum{9}But do thou to them as unto the Madianites : unto Sisera, and unto Jabin at the brook of Kison;\par
\psanum{10}Who perished at Endor : and became as the dung of the earth.\par
\psanum{11}Make them and their princes like Oreb and Zeb : yea, make all their princes like as Zeba and Salmana;\par
\psanum{12}Who say, Let us take to ourselves : the houses of God in possession.\par
\psanum{13}O my God, make them like unto a wheel : and as the stubble before the wind;\par
\psanum{14}Like as the fire that burneth up the wood : and as the flame that consumeth the mountains.\par
\psanum{15}Persecute them even so with thy tempest : and make them afraid with thy storm.\par
\psanum{16}Make their faces ashamed, O Lord : that they may seek thy Name.\par
\psanum{17}Let them be confounded and vexed ever more and more : let them be put to shame, and perish.\par
\psanum{18}And they shall know that thou, whose Name is \textsc{Jehovah} : art only the most Highest over all the earth.\par

\subbypsalm{Psalm 84. \textit{Quam dilecta!}}
\fancyhead[RE,LO]{Psalm 84}

\lett{O}{how} amiable are thy dwellings : thou Lord of hosts!\par
\secondlinepsalm{My soul hath a desire and longing to enter into the courts of the Lord : my heart and my flesh rejoice in the living God.}
\thirdlinepsalm{Yea, the sparrow hath found her an house, and the swallow a nest where she may lay her young : even thy altars, O Lord of hosts, my King and my God.}
\psanum{4}Blessed are they that dwell in thy house : they will be alway praising thee.\par
\psanum{5}Blessed is the man whose strength is in thee : in whose heart are thy ways.\par
\psanum{6}Who going through the vale of misery use it for a well : and the pools are filled with water.\par
\psanum{7}They will go from strength to strength : and unto the God of gods appeareth every one of them in Sion.\par
\psanum{8}O Lord God of hosts, hear my prayer : hearken, O God of Jacob.\par
\psanum{9}Behold, O God our defender : and look upon the face of thine Anointed.\par
\psanum{10}For one day in thy courts : is better than a thousand.\par
\psanum{11}I had rather be a door-keeper in the house of my God : than to dwell in the tents of ungodliness.\par
\psanum{12}For the Lord God is a light and defence : the Lord will give grace and worship, and no good thing shall he withhold from them that live a godly life.\par
\psanum{13}O Lord God of hosts : blessed is the man that putteth his trust in thee.\par

%MANUAL ADJUSTMENT:
\subsec{Psalm 85. \textit{Benedixisti, Domine}}

\fancyhead[RE,LO]{Psalm 85}

\lett{L}{ord,} thou art become gracious unto thy land : thou hast turned away the captivity of Jacob.\par
\secondlinepsalm{Thou hast forgiven the offence of thy people : and covered all their sins.}
\thirdlinepsalm{Thou hast taken away all thy displeasure : and turned thyself from thy wrathful indignation.}
\psanum{4}Turn us then, O God our Saviour : and let thine anger cease from us.\par
\psanum{5}Wilt thou be displeased at us for ever : and wilt thou stretch out thy wrath from one generation to another?\par
\psanum{6}Wilt thou not turn again, and quicken us : that thy people may rejoice in thee?\par
\psanum{7}Shew us thy mercy, O Lord : and grant us thy salvation.\par
\psanum{8}I will hearken what the Lord God will say concerning me : for he shall speak peace unto his people, and to his saints, that they turn not again.\par
\psanum{9}For his salvation is nigh them that fear him : that glory may dwell in our land.\par
\psanum{10}Mercy and truth are met together : righteousness and peace have kissed each other.\par
\psanum{11}Truth shall flourish out of the earth : and righteousness hath looked down from heaven.\par
\psanum{12}Yea, the Lord shall shew loving-kindness : and our land shall give her increase.\par
\psanum{13}Righteousness shall go before him : and he shall direct his going in the way.\par

\psalmsec{Day 17. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 17. Morning Prayer}}

\subbypsalm{Psalm 86. \textit{Inclina, Domine}}
\fancyhead[RE,LO]{Psalm 86}

\lett{B}{ow} down thine ear, O Lord, and hear me : for I am poor, and in misery.\par
\secondlinepsalm{Preserve thou my soul, for I am holy : my God, save thy servant that putteth his trust in thee.}
\thirdlinepsalm{Be merciful unto me, O Lord : for I will call daily upon thee.}
\psanum{4}Comfort the soul of thy servant : for unto thee, O Lord, do I lift up my soul.\par
\psanum{5}For thou, Lord, art good and gracious : and of great mercy unto all them that call upon thee.\par
\psanum{6}Give ear, Lord, unto my prayer : and ponder the voice of my humble desires.\par
\psanum{7}In the time of my trouble I will call upon thee : for thou hearest me.\par
\psanum{8}Among the gods there is none like unto thee, O Lord : there is not one that can do as thou doest.\par
\psanum{9}All nations whom thou hadst made shall come and worship thee, O Lord : and shall glorify thy Name.\par
\psanum{10}For thou art great, and doest wondrous things : thou art God alone.\par
\psanum{11}Teach me thy way, O Lord, and I will walk in thy truth : O knit my heart unto thee, that I may fear thy Name.\par
\psanum{12}I will thank thee, O Lord my God, with all my heart : and will praise thy Name for evermore.\par
\psanum{13}For great is thy mercy toward me : and thou hast delivered my soul from the nethermost hell.\par
\psanum{14}O God, the proud are risen against me : and the congregations of naughty men have sought after my soul, and have not set thee before their eyes.\par
\psanum{15}But thou, O Lord God, art full of compassion and mercy : long-suffering, plenteous in goodness and truth.\par
\psanum{16}O turn thee then unto me, and have mercy upon me : give thy strength unto thy servant, and help the son of thine handmaid.\par
\psanum{17}Shew some token upon me for good, that they who hate me may see it and be ashamed : because thou, Lord, hast holpen me and comforted me.\par

\subbypsalm{Psalm 87. \textit{Fundamenta ejus}}
\fancyhead[RE,LO]{Psalm 87}

\lett{H}{er} foundations are upon the holy hills : the Lord loveth the gates of Sion more than all the dwellings of Jacob.\par
\secondlinepsalm{Very excellent things are spoken of thee : thou city of God.}
\thirdlinepsalm{I will think upon Rahab and Babylon : with them that know me.}
\psanum{4}Behold ye the Philistines also : and they of Tyre, with the Morians; lo, there was he born.\par
\psanum{5}And of Sion it shall be reported that he was born in her : and the most High shall stablish her.\par
\psanum{6}The Lord shall rehearse it when he writeth up the people : that he was born there.\par
\psanum{7}The singers also and trumpeters shall he rehearse : All my fresh springs shall be in thee.\par

\subbypsalm{Psalm 88. \textit{Domine Deus}}
\fancyhead[RE,LO]{Psalm 88}

\lett{O}{Lord} God of my salvation, I have cried day and night before thee : O let my prayer enter into thy presence, incline thine ear unto my calling.\par
\secondlinepsalm{For my soul is full of trouble : and my life draweth nigh unto hell.}
\thirdlinepsalm{I am counted as one of them that go down into the pit : and I have been even as a man that hath no strength.}
\psanum{4}Free among the dead, like unto them that are wounded, and lie in the grave : who are out of remembrance, and are cut away from thy hand.\par
\psanum{5}Thou hast laid me in the lowest pit : in a place of darkness, and in the deep.\par
\psanum{6}Thine indignation lieth hard upon me : and thou hast vexed me with all thy storms.\par
\psanum{7}Thou hast put away mine acquaintance far from me : and made me to be abhorred of them.\par
\psanum{8}I am so fast in prison : that I cannot get forth.\par
\psanum{9}My sight faileth for very trouble : Lord, I have called daily upon thee, I have stretched forth my hands unto thee.\par
\psanum{10}Dost thou shew wonders among the dead : or shall the dead rise up again, and praise thee?\par
\psanum{11}Shall thy loving-kindness be shewed in the grave : or thy faithfulness in destruction?\par
\psanum{12}Shall thy wondrous works be known in the dark : and thy righteousness in the land where all things are forgotten?\par
\psanum{13}Unto thee have I cried, O Lord : and early shall my prayer come before thee.\par
\psanum{14}Lord, why abhorrest thou my soul : and hidest thou thy face from me?\par
\psanum{15}I am in misery, and like unto him that is at the point to die : even from my youth up thy terrors have I suffered with a troubled mind.\par
\psanum{16}Thy wrathful displeasure goeth over me : and the fear of thee hath undone me.\par
\psanum{17}They came round about me daily like water : and compassed me together on every side.\par
\psanum{18}My lovers and friends hast thou put away from me : and hid mine acquaintance out of my sight.\par

\psalmsec{Day 17. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 17. Evening Prayer}}

\subbypsalm{Psalm 89. \textit{Misericordias Domini}}
\fancyhead[RE,LO]{Psalm 89}

\lett{M}{y} song shall be alway of the loving-kindness of the Lord : with my mouth will I ever be shewing thy truth from one generation to another.\par
\secondlinepsalm{For I have said, Mercy shall be set up for ever : thy truth shalt thou stablish in the heavens.}
\thirdlinepsalm{I have made a covenant with my chosen : I have sworn unto David my servant;}
\psanum{4}Thy seed will I stablish for ever : and set up thy throne from one generation to another.\par
\psanum{5}O Lord, the very heavens shall praise thy wondrous works : and thy truth in the congregation of the saints.\par
\psanum{6}For who is he among the clouds : that shall be compared unto the Lord?\par
\psanum{7}And what is he among the gods : that shall be like unto the Lord?\par
\psanum{8}God is very greatly to be feared in the council of the saints : and to be had in reverence of all them that are round about him.\par
\psanum{9}O Lord God of hosts, who is like unto thee : thy truth, most mighty Lord, is on every side.\par
\psanum{10}Thou rulest the raging of the sea : thou stillest the waves thereof when they arise.\par
\psanum{11}Thou hast subdued Egypt, and destroyed it : thou hast scattered thine enemies abroad with thy mighty arm.\par
\psanum{12}The heavens are thine, the earth also is thine : thou hast laid the foundation of the round world, and all that therein is.\par
\psanum{13}Thou hast made the north and the south : Tabor and Hermon shall rejoice in thy Name.\par
\psanum{14}Thou hast a mighty arm : strong is thy hand, and high is thy right hand.\par
\psanum{15}Righteousness and equity are the habitation of thy seat : mercy and truth shall go before thy face.\par
\psanum{16}Blessed is the people, O Lord, that can rejoice in thee : they shall walk in the light of thy countenance.\par
\psanum{17}Their delight shall be daily in thy Name : and in thy righteousness shall they make their boast.\par
\psanum{18}For thou art the glory of their strength : and in thy loving-kindness thou shalt lift up our horns.\par
\psanum{19}For the Lord is our defence : the Holy One of Israel is our King.\par
\psanum{20}Thou spakest sometime in visions unto thy saints, and saidst : I have laid help upon one that is mighty; I have exalted one chosen out of the people.\par
\psanum{21}I have found David my servant : with my holy oil have I anointed him.\par
\psanum{22}My hand shall hold him fast : and my arm shall strengthen him.\par
\psanum{23}The enemy shall not be able to do him violence : the son of wickedness shall not hurt him.\par
\psanum{24}I will smite down his foes before his face : and plague them that hate him.\par
\psanum{25}My truth also and my mercy shall be with him : and in my Name shall his horn be exalted.\par
\psanum{26}I will set his dominion also in the sea : and his right hand in the floods.\par
\psanum{27}He shall call me, Thou art my Father : my God, and my strong salvation.\par
\psanum{28}And I will make him my first-born : higher than the kings of the earth.\par
\psanum{29}My mercy will I keep for him for evermore : and my covenant shall stand fast with him.\par
\psanum{30}His seed also will I make to endure for ever : and his throne as the days of heaven.\par
\psanum{31}But if his children forsake my law : and walk not in my judgements;\par
\psanum{32}If they break my statutes, and keep not my commandments : I will visit their offences with the rod, and their sin with scourges.\par
\psanum{33}Nevertheless, my loving-kindness will I not utterly take from him : nor suffer my truth to fail.\par
\psanum{34}My covenant I will not break, nor alter the thing that is gone out of my lips : I have sworn once by my holiness, that I will not fail David.\par
\psanum{35}His seed shall endure for ever : and his seat is like as the sun before me.\par
\psanum{36}He shall stand fast for evermore as the moon : and as the faithful witness in heaven.\par
\psanum{37}But thou hast abhorred and forsaken thine Anointed : and art displeased at him.\par
\psanum{38}Thou hast broken the covenant of thy servant : and cast his crown to the ground.\par
\psanum{39}Thou hast overthrown all his hedges : and broken down his strong holds.\par
\psanum{40}All they that go by spoil him : and he is become a reproach to his neighbours.\par
\psanum{41}Thou hast set up the right hand of his enemies : and made all his adversaries to rejoice.\par
\psanum{42}Thou hast taken away the edge of his sword : and givest him not victory in the battle.\par
\psanum{43}Thou hast put out his glory : and cast his throne down to the ground.\par
\psanum{44}The days of his youth hast thou shortened : and covered him with dishonour.\par
\psanum{45}Lord, how long wilt thou hide thyself, for ever : and shall thy wrath burn like fire?\par
\psanum{46}O remember how short my time is : wherefore hast thou made all men for nought?\par
\psanum{47}What man is he that liveth, and shall not see death : and shall he deliver his soul from the hand of hell?\par
\psanum{48}Lord, where are thy old loving-kindnesses : which thou swarest unto David in thy truth?\par
\psanum{49}Remember, Lord, the rebuke that thy servants have : and how I do bear in my bosom the rebukes of many people.\par
\psanum{50}Wherewith thine enemies have blasphemed thee, and slandered the footsteps of thine Anointed : Praised be the Lord for evermore. Amen, and Amen.\par

\psalmsec{Day 18. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 18. Morning Prayer}}

\subbypsalm{Psalm 90. \textit{Domine, refugium}}
\fancyhead[RE,LO]{Psalm 90}

\lett{L}{ord,} thou hast been our refuge : from one generation to another.\par
%Third line too long:
\secondlinepsalm{Before the mountains were brought forth, or ever the earth and the world were made : thou art God from everlasting, and world without}
\noindent
end.\par
\psanum{3}Thou turnest man to destruction : again thou sayest, Come again, ye children of men.\par
\psanum{4}For a thousand years in thy sight are but as yesterday : seeing that is past as a watch in the night.\par
\psanum{5}As soon as thou scatterest them they are even as a sleep : and fade away suddenly like the grass.\par
\psanum{6}In the morning it is green, and groweth up : but in the evening it is cut down, dried up, and withered.\par
\psanum{7}For we consume away in thy displeasure : and are afraid at thy wrathful indignation.\par
\psanum{8}Thou hast set our misdeeds before thee : and our secret sins in the light of thy countenance.\par
\psanum{9}For when thou art angry all our days are gone : we bring our years to an end, as it were a tale that is told.\par
\psanum{10}The days of our age are threescore years and ten; and though men be so strong that they come to fourscore years : yet is their strength then but labour and sorrow; so soon passeth it away, and we are gone.\par
\psanum{11}But who regardeth the power of thy wrath : for even thereafter as a man feareth, so is thy displeasure.\par
\psanum{12}So teach us to number our days : that we may apply our hearts unto wisdom.\par
\psanum{13}Turn thee again, O Lord, at the last : and be gracious unto thy servants.\par
\psanum{14}O satisfy us with thy mercy, and that soon : so shall we rejoice and be glad all the days of our life.\par
\psanum{15}Comfort us again now after the time that thou hast plagued us : and for the years wherein we have suffered adversity.\par
\psanum{16}Shew thy servants thy work : and their children thy glory.\par
\psanum{17}And the glorious majesty of the Lord our God be upon us : prosper thou the work of our hands upon us, O prosper thou our handywork.\par

\subbypsalm{Psalm 91. \textit{Qui habitat}}
\fancyhead[RE,LO]{Psalm 91}

\lett{W}{hoso} dwelleth under the defence of the most High : shall abide under the shadow of the Almighty.\par
\secondlinepsalm{I will say unto the Lord, Thou art my hope, and my strong hold : my God, in him will I trust.}
\thirdlinepsalm{For he shall deliver thee from the snare of the hunter : and from the noisome pestilence.}
\psanum{4}He shall defend thee under his wings, and thou shalt be safe under his feathers : his faithfulness and truth shall be thy shield and buckler.\par
\psanum{5}Thou shalt not be afraid for any terror by night : nor for the arrow that flieth by day;\par
\psanum{6}For the pestilence that walketh in darkness : nor for the sickness that destroyeth in the noon-day.\par
\psanum{7}A thousand shall fall beside thee, and ten thousand at thy right hand : but it shall not come nigh thee.\par
\psanum{8}Yea, with thine eyes shalt thou behold : and see the reward of the ungodly.\par
\psanum{9}For thou, Lord, art my hope : thou hast set thine house of defence very high.\par
\psanum{10}There shall no evil happen unto thee : neither shall any plague come nigh thy dwelling.\par
\psanum{11}For he shall give his angels charge over thee : to keep thee in all thy ways.\par
\psanum{12}They shall bear thee in their hands : that thou hurt not thy foot against a stone.\par
\psanum{13}Thou shalt go upon the lion and adder : the young lion and the dragon shalt thou tread under thy feet.\par
\psanum{14}Because he hath set his love upon me, therefore will I deliver him : I will set him up, because he hath known my Name.\par
\psanum{15}He shall call upon me, and I will hear him : yea, I am with him in trouble; I will deliver him, and bring him to honour.\par
\psanum{16}With long life will I satisfy him : and shew him my salvation.\par

\subbypsalm{Psalm 92. \textit{Bonum est confiteri}}
\fancyhead[RE,LO]{Psalm 92}

\lett{I}{t} is a good thing to give thanks unto the Lord : and to sing praises unto thy Name, O most Highest;\par
\secondlinepsalm{To tell of thy loving-kindness early in the morning : and of thy truth in the night-season;}
\thirdlinepsalm{Upon an instrument of ten strings, and upon the lute : upon a loud instrument, and upon the harp.}
\psanum{4}For thou, Lord, hast made me glad through thy works : and I will rejoice in giving praise for the operations of thy hands.\par
\psanum{5}O Lord, how glorious are thy works : thy thoughts are very deep.\par
\psanum{6}An unwise man doth not well consider this : and a fool doth not understand it.\par
\psanum{7}When the ungodly are green as the grass, and when all the workers of wickedness do flourish : then shall they be destroyed for ever; but thou, Lord, art the most Highest for evermore.\par
\psanum{8}For lo, thine enemies, O Lord, lo, thine enemies shall perish : and all the workers of wickedness shall be destroyed.\par
\psanum{9}But mine horn shall be exalted like the horn of an unicorn : for I am anointed with fresh oil.\par
\psanum{10}Mine eye also shall see his lust of mine enemies : and mine ear shall hear his desire of the wicked that arise up against me.\par
\psanum{11}The righteous shall flourish like a palm-tree : and shall spread abroad like a cedar in Libanus.\par
\psanum{12}Such as are planted in the house of the Lord : shall flourish in the courts of the house of our God.\par
\psanum{13}They also shall bring forth more fruit in their age : and shall be fat and well-liking.\par
\psanum{14}That they may shew how true the Lord my strength is : and that there is no unrighteousness in him.\par

\psalmsec{Day 18. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 18. Evening Prayer}}

\subbypsalm{Psalm 93. \textit{Dominus regnavit}}
\fancyhead[RE,LO]{Psalm 93}

\lett{T}{he} Lord is King, and hath put on glorious apparel : the Lord hath put on his apparel, and girded himself with strength.\par
\secondlinepsalm{He hath made the round world so sure : that it cannot be moved.}
\thirdlinepsalm{Ever since the world began hath thy seat been prepared : thou art from everlasting.}
\psanum{4}The floods are risen, O Lord, the floods have lift up their voice : the floods lift up their waves.\par
\psanum{5}The waves of the sea are mighty, and rage horribly : but yet the Lord, who dwelleth on high, is mightier.\par
\psanum{6}Thy testimonies, O Lord, are very sure : holiness becometh thine house for ever.\par

\subbypsalm{Psalm 94. \textit{Deus ultionum}}
\fancyhead[RE,LO]{Psalm 94}

\lett{O}{Lord} God, to whom vengeance belongeth : thou God, to whom vengeance belongeth, shew thyself.\par
\secondlinepsalm{Arise, thou Judge of the world : and reward the proud after their deserving.}
\thirdlinepsalm{Lord, how long shall the ungodly : how long shall the ungodly triumph?}
\psanum{4}How long shall all wicked doers speak so disdainfully : and make such proud boasting?\par
\psanum{5}They smite down thy people, O Lord : and trouble thine heritage.\par
\psanum{6}They murder the widow and the stranger : and put the fatherless to death.\par
\psanum{7}And yet they say, Tush, the Lord shall not see : neither shall the God of Jacob regard it.\par
\psanum{8}Take heed, ye unwise among the people : O ye fools, when will ye understand?\par
\psanum{9}He that planted the ear, shall he not hear : or he that made the eye, shall he not see?\par
\psanum{10}Or he that nurtureth the heathen : it is he that teacheth man knowledge, shall not he punish?\par
\psanum{11}The Lord knoweth the thoughts of man : that they are but vain.\par
\psanum{12}Blessed is the man whom thou chastenest, O Lord : and teachest him in thy law;\par
\psanum{13}That thou mayest give him patience in time of adversity : until the pit be digged up for the ungodly.\par
\psanum{14}For the Lord will not fail his people : neither will he forsake his inheritance;\par
\psanum{15}Until righteousness turn again unto judgement : all such as are true in heart shall follow it.\par
\psanum{16}Who will rise up with me against the wicked : or who will take my part against the evil-doers?\par
\psanum{17}If the Lord had not helped me : it had not failed but my soul had been put to silence.\par
\psanum{18}But when I said, My foot hath slipt : thy mercy, O Lord, held me up.\par
\psanum{19}In the multitude of the sorrows that I had in my heart : thy comforts have refreshed my soul.\par
\psanum{20}Wilt thou have any thing to do with the stool of wickedness : which imagineth mischief as a law?\par
\psanum{21}They gather them together against the soul of the righteous : and condemn the innocent blood.\par
\psanum{22}But the Lord is my refuge : and my God is the strength of my confidence.\par
\psanum{23}He shall recompense them their wickedness, and destroy them in their own malice : yea, the Lord our God shall destroy them.\par

\psalmsec{Day 19. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 19. Morning Prayer}}

\subbypsalm{Psalm 95. \textit{Venite, exultemus}}
\fancyhead[RE,LO]{Psalm 95}

\lett{O}{come,} let us sing unto the Lord : let us heartily rejoice in the strength of our salvation.\par
\secondlinepsalm{Let us come before his presence with thanksgiving : and shew ourselves glad in him with psalms.}
\thirdlinepsalm{For the Lord is a great God : and a great King above all gods.}
\psanum{4}In his hand are all the corners of the earth : and the strength of the hills is his also.\par
\psanum{5}The sea is his, and he made it : and his hands prepared the dry land.\par
\psanum{6}O come, let us worship and fall down : and kneel before the Lord our Maker.\par
\psanum{7}For he is the Lord our God : and we are the people of his pasture, and the sheep of his hand.\par
\psanum{8}To-day if ye will hear his voice, harden not your hearts : as in the provocation, and as in the day of temptation in the wilderness.\par
\psanum{9}When your fathers tempted me : proved me, and saw my works.\par
\psanum{10}Forty years long was I grieved with this generation, and said : It is a people that do err in their hearts, for they have not known my ways;\par
\psanum{11}Unto whom I sware in my wrath : that they should not enter into my rest.\par

\subbypsalm{Psalm 96. \textit{Cantate Domino}}
\fancyhead[RE,LO]{Psalm 96}

\lett{O}{sing} unto the Lord a new song : sing unto the Lord, all the whole earth.\par
\secondlinepsalm{Sing unto the Lord, and praise his Name : be telling of his salvation from day to day.}
\thirdlinepsalm{Declare his honour unto the heathen : and his wonders unto all people.}
\psanum{4}For the Lord is great, and cannot worthily be praised : he is more to be feared than all gods.\par
\psanum{5}As for all the gods of the heathen, they are but idols : but it is the Lord that made the heavens.\par
\psanum{6}Glory and worship are before him : power and honour are in his sanctuary.\par
\psanum{7}Ascribe unto the Lord, O ye kindreds of the people : ascribe unto the Lord worship and power.\par
\psanum{8}Ascribe unto the Lord the honour due unto his Name : bring presents, and come into his courts.\par
\psanum{9}O worship the Lord in the beauty of holiness : let the whole earth stand in awe of him.\par
\psanum{10}Tell it out among the heathen that the Lord is King : and that it is he who hath made the round world so fast that it cannot be moved; and how that he shall judge the people righteously.\par
\psanum{11}Let the heavens rejoice, and let the earth be glad : let the sea make a noise, and all that therein is.\par
\psanum{12}Let the field be joyful, and all that is in it : then shall all the trees of the wood rejoice before the Lord.\par
\psanum{13}For he cometh, for he cometh to judge the earth : and with righteousness to judge the world, and the people with his truth.\par

\subbypsalm{Psalm 97. \textit{Dominus regnavit}}
\fancyhead[RE,LO]{Psalm 97}

\lett{T}{he} Lord is King, the earth may be glad thereof : yea, the multitude of the isles may be glad thereof.\par
\secondlinepsalm{Clouds and darkness are round about him : righteousness and judgement are the habitation of his seat.}
\thirdlinepsalm{There shall go a fire before him : and burn up his enemies on every side.}
\psanum{4}His lightnings gave shine unto the world : the earth saw it, and was afraid.\par
\psanum{5}The hills melted like wax at the presence of the Lord : at the presence of the Lord of the whole earth.\par
\psanum{6}The heavens have declared his righteousness : and all the people have seen his glory.\par
\psanum{7}Confounded be all they that worship carved images, and that delight in vain gods : worship him, all ye gods.\par
\psanum{8}Sion heard of it, and rejoiced : and the daughters of Judah were glad, because of thy judgements, O Lord.\par
\psanum{9}For thou, Lord, art higher than all that are in the earth : thou art exalted far above all gods.\par
\psanum{10}O ye that love the Lord, see that ye hate the thing which is evil : the Lord preserveth the souls of his saints; he shall deliver them from the hand of the ungodly.\par
\psanum{11}There is sprung up a light for the righteous : and joyful gladness for such as are true-hearted.\par
\psanum{12}Rejoice in the Lord, ye righteous : and give thanks for a remembrance of his holiness.\par

\psalmsec{Day 19. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 19. Evening Prayer}}

\subbypsalm{Psalm 98. \textit{Cantate Domino}}
\fancyhead[RE,LO]{Psalm 98}

\lett{O}{sing} unto the Lord a new song : for he hath done marvellous things.\par
\secondlinepsalm{With his own right hand, and with his holy arm : hath he gotten himself the victory.}
\thirdlinepsalm{The Lord declared his salvation : his righteousness hath he openly shewed in the sight of the heathen.}
\psanum{4}He hath remembered his mercy and truth toward the house of Israel : and all the ends of the world have seen the salvation of our God.\par
\psanum{5}Shew yourselves joyful unto the Lord, all ye lands : sing, rejoice, and give thanks.\par
\psanum{6}Praise the Lord upon the harp : sing to the harp with a psalm of thanksgiving.\par
\psanum{7}With trumpets also and shawms : O shew yourselves joyful before the Lord the King.\par
\psanum{8}Let the sea make a noise, and all that therein is : the round world, and they that dwell therein.\par
\psanum{9}Let the floods clap their hands, and let the hills be joyful together before the Lord : for he is come to judge the earth.\par
\psanum{10}With righteousness shall he judge the world : and the people with equity.\par

\subbypsalm{Psalm 99. \textit{Dominus regnavit}}
\fancyhead[RE,LO]{Psalm 99}

\lett{T}{he} Lord is King, be the people never so unpatient : he sitteth between the cherubims, be the earth never so unquiet.\par
\secondlinepsalm{The Lord is great in Sion : and high above all people.}
\thirdlinepsalm{They shall give thanks unto thy Name : which is great, wonderful, and holy.}
\psanum{4}The King's power loveth judgement; thou hast prepared equity : thou hast executed judgement and righteousness in Jacob.\par
\psanum{5}O magnify the Lord our God : and fall down before his footstool, for he is holy.\par
\psanum{6}Moses and Aaron among his priests, and Samuel among such as call upon his Name : these called upon the Lord, and he heard them.\par
\psanum{7}He spake unto them out of the cloudy pillar : for they kept his testimonies, and the law that he gave them.\par
\psanum{8}Thou heardest them, O Lord our God : thou forgavest them, O God, and punishedst their own inventions.\par
\psanum{9}O magnify the Lord our God, and worship him upon his holy hill : for the Lord our God is holy.\par

\subbypsalm{Psalm 100. \textit{Jubilate Deo}}
\fancyhead[RE,LO]{Psalm 100}

\lett{O}{be} joyful in the Lord, all ye lands : serve the Lord with gladness, and come before his presence with a song.\par
\secondlinepsalm{Be ye sure that the Lord he is God : it is he that hath made us, and not we ourselves; we are his people, and the sheep of his pasture.}
\thirdlinepsalm{O go your way into his gates with thanksgiving, and into his courts with praise : be thankful unto him, and speak good of his Name.}
\psanum{4}For the Lord is gracious, his mercy is everlasting : and his truth endureth from generation to generation.\par

\subbypsalm{Psalm 101. \textit{Misericordiam et judicium}}
\fancyhead[RE,LO]{Psalm 101}

\lett{M}{y} song shall be of mercy and judgement : unto thee, O Lord, will I sing.\par
\secondlinepsalm{O let me have understanding : in the way of godliness.}
\thirdlinepsalm{When wilt thou come unto me : I will walk in my house with a perfect heart.}
\psanum{4}I will take no wicked thing in hand; I hate the sins of unfaithfulness : there shall no such cleave unto me.\par
\psanum{5}A froward heart shall depart from me : I will not know a wicked person.\par
\psanum{6}Whoso privily slandereth his neighbour : him will I destroy.\par
\psanum{7}Whoso hath also a proud look and high stomach : I will not suffer him.\par
\psanum{8}Mine eyes look upon such as are faithful in the land : that they may dwell with me.\par
\psanum{9}Whoso leadeth a godly life : he shall be my servant.\par
\psanum{10}There shall no deceitful person dwell in my house : he that telleth lies shall not tarry in my sight.\par
\psanum{11}I shall soon destroy all the ungodly that are in the land : that I may root out all wicked doers from the city of the Lord.\par

\psalmsec{Day 20. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 20. Morning Prayer}}

\subbypsalm{Psalm 102. \textit{Domine, exaudi}}
\fancyhead[RE,LO]{Psalm 102}

\lett{H}{ear} my prayer, O Lord : and let my crying come unto thee.\par
\secondlinepsalm{Hide not thy face from me in the time of my trouble : incline thine ear unto me when I call; O hear me, and that right soon.}
\thirdlinepsalm{For my days are consumed away like smoke : and my bones are burnt up as it were a firebrand.}
\psanum{4}My heart is smitten down, and withered liked grass : so that I forget to eat my bread.\par
\psanum{5}For the voice of my groaning : my bones will scarce cleave to my flesh.\par
\psanum{6}I am become like a pelican in the wilderness : and like an owl that is in the desert.\par
\psanum{7}I have watched, and am even as it were a sparrow : that sitteth alone upon the house-top.\par
\psanum{8}Mine enemies revile me all the day long : and they that are mad upon me are sworn together against me.\par
\psanum{9}For I have eaten ashes as it were bread : and mingled my drink with weeping;\par
\psanum{10}And that because of thine indignation and wrath : for thou hast taken me up, and cast me down.\par
\psanum{11}My days are gone like a shadow : and I am withered like grass.\par
\psanum{12}But thou, O Lord, shalt endure for ever : and thy remembrance throughout all generations.\par
\psanum{13}Thou shalt arise, and have mercy upon Sion : for it is time that thou have mercy upon her, yea, the time is come.\par
\psanum{14}And why? thy servants think upon her stones : and it pitieth them to see her in the dust.\par
\psanum{15}The heathen shall fear thy Name, O Lord : and all the kings of the earth thy majesty;\par
\psanum{16}When the Lord shall build up Sion : and when his glory shall appear;\par
\psanum{17}When he turneth him unto the prayer of the poor destitute : and despiseth not their desire.\par
\psanum{18}This shall be written for those that come after : and the people which shall be born shall praise the Lord.\par
\psanum{19}For he hath looked down from his sanctuary : out of the heaven did the Lord behold the earth;\par
\psanum{20}That he might hear the mournings of such as are in captivity : and deliver the children appointed unto death;\par
\psanum{21}That they may declare the Name of the Lord in Sion : and his worship at Jerusalem;\par
\psanum{22}When the people are gathered together : and the kingdoms also, to serve the Lord.\par
\psanum{23}He brought down my strength in my journey : and shortened my days.\par
\psanum{24}But I said, O my God, take me not away in the midst of mine age : as for thy years, they endure throughout all generations.\par
\psanum{25}Thou, Lord, in the beginning hast laid the foundation of the earth : and the heavens are the work of thy hands.\par
\psanum{26}They shall perish, but thou shalt endure : they all shall wax old as doth a garment;\par
\psanum{27}And as a vesture shalt thou change them, and they shall be changed : but thou art the same, and thy years shall not fail.\par
\psanum{28}The children of thy servants shall continue : and their seed shall stand fast in thy sight.\par

\subbypsalm{Psalm 103. \textit{Benedic, anima mea}}
\fancyhead[RE,LO]{Psalm 103}

\lett{P}{raise} the Lord, O my soul : and all that is within me praise his holy Name.\par
\secondlinepsalm{Praise the Lord, O my soul : and forget not all his benefits;}
\thirdlinepsalm{Who forgiveth all thy sin : and healeth all thine infirmities;}
\psanum{4}Who saveth thy life from destruction : and crowneth thee with mercy and loving-kindness;\par
\psanum{5}Who satisfieth thy mouth with good things : making thee young and lusty as an eagle.\par
\psanum{6}The Lord executeth righteousness and judgement : for all them that are oppressed with wrong.\par
\psanum{7}He shewed his ways unto Moses : his works unto the children of Israel.\par
\psanum{8}The Lord is full of compassion and mercy : long-suffering, and of great goodness.\par
\psanum{9}He will not alway be chiding : neither keepeth he his anger for ever.\par
\psanum{10}He hath not dealt with us after our sins : nor rewarded us according to our wickednesses.\par
\psanum{11}For look how high the heaven is in comparison of the earth : so great is his mercy also toward them that fear him.\par
\psanum{12}Look how wide also the east is from the west : so far hath he set our sins from us.\par
\psanum{13}Yea, like as a father pitieth his own children : even so is the Lord merciful unto them that fear him.\par
\psanum{14}For he knoweth whereof we are made : he remembereth that we are but dust.\par
\psanum{15}The days of man are but as grass : for he flourisheth as a flower of the field.\par
\psanum{16}For as soon as the wind goeth over it, it is gone : and the place thereof shall know it no more.\par
\psanum{17}But the merciful goodness of the Lord endureth for ever and ever upon them that fear him : and his righteousness upon children's children;\par
\psanum{18}Even upon such as keep his covenant : and think upon his commandments to do them.\par
\psanum{19}The Lord hath prepared his seat in heaven : and his kingdom ruleth over all.\par
\psanum{20}O praise the Lord, ye angels of his, ye that excel in strength : ye that fulfil his commandment, and hearken unto the voice of his words.\par
\psanum{21}O praise the Lord, all ye his hosts : ye servants of his that do his pleasure.\par
\psanum{22}O speak good of the Lord, all ye works of his, in all places of his dominion : praise thou the Lord, O my soul.\par

\psalmsec{Day 20. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 20. Evening Prayer}}

\subbypsalm{Psalm 104. \textit{Benedic, anima mea}}
\fancyhead[RE,LO]{Psalm 104}

\lett{P}{raise} the Lord, O my soul : O Lord my God, thou art become exceeding glorious; thou art clothed with majesty and honour.\par
\secondlinepsalm{Thou deckest thyself with light as it were with a garment : and spreadest out the heavens like a curtain.}
\thirdlinepsalm{Who layeth the beams of his chambers in the waters : and maketh the clouds his chariot, and walketh upon the wings of the wind.}
\psanum{4}He maketh his angels spirits : and his ministers a flaming fire.\par
\psanum{5}He laid the foundations of the earth : that it never should move at any time.\par
\psanum{6}Thou coveredst it with the deep like as with a garment : the waters stand in the hills.\par
\psanum{7}At thy rebuke they flee : at the voice of thy thunder they are afraid.\par
\psanum{8}They go up as high as the hills, and down to the valleys beneath : even unto the place which thou hast appointed for them.\par
\psanum{9}Thou hast set them their bounds which they shall not pass : neither turn again to cover the earth.\par
\psanum{10}He sendeth the springs into the rivers : which run among the hills.\par
\psanum{11}All beasts of the field drink thereof : and the wild asses quench their thirst.\par
\psanum{12}Beside them shall the fowls of the air have their habitation : and sing among the branches.\par
\psanum{13}He watereth the hills from above : the earth is filled with the fruit of thy works.\par
\psanum{14}He bringeth forth grass for the cattle : and green herb for the service of men;\par
\psanum{15}That he may bring food out of the earth, and wine that maketh glad the heart of man : and oil to make him a cheerful countenance, and bread to strengthen man's heart.\par
\psanum{16}The trees of the Lord also are full of sap : even the cedars of Libanus which he hath planted;\par
\psanum{17}Wherein the birds make their nests : and the fir-trees are a dwelling for the stork.\par
\psanum{18}The high hills are a refuge for the wild goats : and so are the stony rocks for the conies.\par
\psanum{19}He appointed the moon for certain seasons : and the sun knoweth his going down.\par
\psanum{20}Thou makest darkness that it may be night : wherein all the beasts of the forest do move.\par
\psanum{21}The lions roaring after their prey : do seek their meat from God.\par
\psanum{22}The sun ariseth, and they get them away together : and lay them down in their dens.\par
\psanum{23}Man goeth forth to his work, and to his labour : until the evening.\par
\psanum{24}O Lord, how manifold are thy works : in wisdom hast thou made them all; the earth is full of thy riches.\par
\psanum{25}So is the great and wide sea also : wherein are things creeping innumerable, both small and great beasts.\par
\psanum{26}There go the ships, and there is that Leviathan : whom thou hast made to take his pastime therein.\par
\psanum{27}These wait all upon thee : that thou mayest give them meat in due season.\par
\psanum{28}When thou givest it them they gather it : and when thou openest thy hand they are filled with good.\par
\psanum{29}When thou hidest thy face they are troubled : when thou takest away their breath they die, and are turned again to their dust.\par
\psanum{30}When thou lettest thy breath go forth they shall be made : and thou shalt renew the face of the earth.\par
\psanum{31}The glorious majesty of the Lord shall endure for ever : the Lord shall rejoice in his works.\par
\psanum{32}The earth shall tremble at the look of him : if he do but touch the hills, they shall smoke.\par
\psanum{33}I will sing unto the Lord as long as I live : I will praise my God while I have my being.\par
\psanum{34}And so shall my words please him : my joy shall be in the Lord.\par
\psanum{35}As for sinners, they shall be consumed out of the earth, and the ungodly shall come to an end : praise thou the Lord, O my soul, praise the Lord.\par

\psalmsec{Day 21. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 21. Morning Prayer}}

\subbypsalm{Psalm 105. \textit{Confitemini Domino}}
\fancyhead[RE,LO]{Psalm 105}

\lett{O}{give} thanks unto the Lord, and call upon his Name : tell the people what things he hath done.\par
\secondlinepsalm{O let your songs be of him, and praise him : and let your talking be of all his wondrous works.}
\thirdlinepsalm{Rejoice in his holy Name : let the heart of them rejoice that seek the Lord.}
\psanum{4}Seek the Lord and his strength : seek his face evermore.\par
\psanum{5}Remember the marvellous works that he hath done : his wonders, and the judgements of his mouth.\par
\psanum{6}O ye seed of Abraham his servant : ye children of Jacob his chosen.\par
\psanum{7}He is the Lord our God : his judgements are in all the world.\par
\psanum{8}He hath been alway mindful of his covenant and promise : that he made to a thousand generations;\par
\psanum{9}Even the covenant that he made with Abraham : and the oath that he sware unto Isaac;\par
\psanum{10}And appointed the same unto Jacob for a law : and to Israel for an everlasting testament;\par
\psanum{11}Saying, Unto thee will I give the land of Canaan : the lot of your inheritance;\par
\psanum{12}When there were yet but a few of them : and they strangers in the land;\par
\psanum{13}What time as they went from one nation to another : from one kingdom to another people;\par
\psanum{14}He suffered no man to do them wrong : but reproved even kings for their sakes;\par
\psanum{15}Touch not mine Anointed : and do my prophets no harm.\par
\psanum{16}Moreover, he called for a dearth upon the land : and destroyed all the provision of bread.\par
\psanum{17}But he had sent a man before them : even Joseph, who was sold to be a bond-servant;\par
\psanum{18}Whose feet they hurt in the stocks : the iron entered into his soul;\par
\psanum{19}Until the time came that his cause was known : the word of the Lord tried him.\par
\psanum{20}The king sent, and delivered him : the prince of the people let him go free.\par
\psanum{21}He made him lord also of his house : and ruler of all his substance;\par
\psanum{22}That he might inform his princes after his will : and teach his senators wisdom.\par
\psanum{23}Israel also came into Egypt : and Jacob was a stranger in the land of Ham.\par
\psanum{24}And he increased his people exceedingly : and made them stronger than their enemies;\par
\psanum{25}Whose heart turned, so that they hated his people : and dealt untruly with his servants.\par
\psanum{26}Then sent he Moses his servant : and Aaron whom he had chosen.\par
\psanum{27}And these shewed his tokens among them : and wonders in the land of Ham.\par
\psanum{28}He sent darkness, and it was dark : and they were not obedient unto his word.\par
\psanum{29}He turned their waters into blood : and slew their fish.\par
\psanum{30}Their land brought forth frogs : yea, even in their kings' chambers.\par
\psanum{31}He spake the word, and there came all manner of flies : and lice in all their quarters.\par
\psanum{32}He gave them hail-stones for rain : and flames of fire in their land.\par
\psanum{33}He smote their vines also and fig-trees : and destroyed the trees that were in their coasts.\par
\psanum{34}He spake the word, and the grasshoppers came, and caterpillars innumerable : and did eat up all the grass in their land, and devoured the fruit of their ground.\par
\psanum{35}He smote all the first-born in their land : even the chief of all their strength.\par
\psanum{36}He brought them forth also with silver and gold : there was not one feeble person among their tribes.\par
\psanum{37}Egypt was glad at their departing : for they were afraid of them.\par
\psanum{38}He spread out a cloud to be a covering : and fire to give light in the night-season.\par
\psanum{39}At their desire he brought quails : and he filled them with the bread of heaven.\par
\psanum{40}He opened the rock of stone, and the waters flowed out : so that rivers ran in the dry places.\par
\psanum{41}For why? he remembered his holy promise : and Abraham his servant.\par
\psanum{42}And he brought forth his people with joy : and his chosen with gladness;\par
\psanum{43}And gave them the lands of the heathen : and they took the labours of the people in possession;\par
\psanum{44}That they might keep his statutes : and observe his laws.\par

\psalmsec{Day 21. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 21. Evening Prayer}}

\subbypsalm{Psalm 106. \textit{Confitemini Domino}}
\fancyhead[RE,LO]{Psalm 106}

\lett{O}{give} thanks unto the Lord, for he is gracious : and his mercy endureth for ever.\par
\secondlinepsalm{Who can express the noble acts of the Lord : or shew forth all his praise?}
\thirdlinepsalm{Blessed are they that alway keep judgement : and do righteousness.}
\psanum{4}Remember me, O Lord, according to the favour that thou bearest unto thy people : O visit me with thy salvation;\par
\psanum{5}That I may see the felicity of thy chosen : and rejoice in the gladness of thy people, and give thanks with thine inheritance.\par
\psanum{6}We have sinned with our fathers : we have done amiss, and dealt wickedly.\par
\psanum{7}Our fathers regarded not thy wonders in Egypt, neither kept they thy great goodness in remembrance : but were disobedient at the sea, even at the Red sea.\par
\psanum{8}Nevertheless, he helped them for his Name's sake : that he might make his power to be known.\par
\psanum{9}He rebuked the Red sea also, and it was dried up : so he led them through the deep, as through a wilderness.\par
\psanum{10}And he saved them from the adversaries' hand : and delivered them from the hand of the enemy.\par
\psanum{11}As for those that troubled them, the waters overwhelmed them : there was not one of them left.\par
\psanum{12}Then believed they his words : and sang praise unto him.\par
\psanum{13}But within a while they forgat his works : and would not abide his counsel.\par
\psanum{14}But lust came upon them in the wilderness : and they tempted God in the desert.\par
\psanum{15}And he gave them their desire : and sent leanness withal into their soul.\par
\psanum{16}They angered Moses also in the tents : and Aaron the saint of the Lord.\par
\psanum{17}So the earth opened, and swallowed up Dathan : and covered the congregation of Abiram.\par
\psanum{18}And the fire was kindled in their company : the flame burnt up the ungodly.\par
\psanum{19}They made a calf in Horeb : and worshipped the molten image.\par
\psanum{20}Thus they turned their glory : into the similitude of a calf that eateth hay.\par
\psanum{21}And they forgat God their Saviour : who had done so great things in Egypt;\par
\psanum{22}Wondrous works in the land of Ham : and fearful things by the Red sea.\par
\psanum{23}So he said, he would have destroyed them, had not Moses his chosen stood before him in the gap : to turn away his wrathful indignation, lest he should destroy them.\par
\psanum{24}Yea, they thought scorn of that pleasant land : and gave no credence unto his word;\par
\psanum{25}But murmured in their tents : and hearkened not unto the voice of the Lord.\par
\psanum{26}Then lift he up his hand against them : to overthrow them in the wilderness;\par
\psanum{27}To cast out their seed among the nations : and to scatter them in the lands.\par
\psanum{28}They joined themselves unto Baal-peor : and ate the offerings of the dead.\par
\psanum{29}Thus they provoked him to anger with their own inventions : and the plague was great among them.\par
\psanum{30}Then stood up Phinees and prayed : and so the plague ceased.\par
\psanum{31}And that was counted unto him for righteousness : among all posterities for evermore.\par
\psanum{32}They angered him also at the waters of strife : so that he punished Moses for their sakes;\par
\psanum{33}Because they provoked his spirit : so that he spake unadvisedly with his lips.\par
\psanum{34}Neither destroyed they the heathen : as the Lord commanded them;\par
\psanum{35}But were mingled among the heathen : and learned their works.\par
\psanum{36}Insomuch that they worshipped their idols, which turned to their own decay : yea, they offered their sons and their daughters unto devils;\par
\psanum{37}And shed innocent blood, even the blood of their sons and of their daughters : whom they had offered unto the idols of Canaan; and the land was defiled with blood.\par
\psanum{38}Thus were they stained with their own works : and went a whoring with their own inventions.\par
\psanum{39}Therefore was the wrath of the Lord kindled against his people : insomuch that he abhorred his own inheritance.\par
\psanum{40}And he gave them over into the hands of the heathen : and they that hated them were lords over them.\par
\psanum{41}Their enemies oppressed them : and had them in subjection.\par
\psanum{42}Many a time did he deliver them : but they rebelled against him with their own inventions, and were brought down in their wickedness.\par
\psanum{43}Nevertheless, when he saw their adversity : he heard their complaint.\par
\psanum{44}He thought upon his covenant, and pitied them according unto the multitude of his mercies : yea, he made all those that led them away captive to pity them.\par
\psanum{45}Deliver us, O Lord our God, and gather us from among the heathen : that we may give thanks unto thy holy Name, and make our boast of thy praise.\par
\psanum{46}Blessed be the Lord God of Israel from everlasting and world without end : and let all the people say, Amen.\par

\psalmsec{Day 22. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 22. Morning Prayer}}

\subbypsalm{Psalm 107. \textit{Confitemini Domino}}
\fancyhead[RE,LO]{Psalm 107}

\lett{O}{give} thanks unto the Lord, for he is gracious : and his mercy endureth for ever.\par
\secondlinepsalm{Let them give thanks whom the Lord hath redeemed : and delivered from the hand of the enemy;}
\thirdlinepsalm{And gathered them out of the lands, from the east and from the west : from the north and from the south.}
\psanum{4}They went astray in the wilderness out of the way : and found no city to dwell in;\par
\psanum{5}Hungry and thirsty : their soul fainted in them.\par
\psanum{6}So they cried unto the Lord in their trouble : and he delivered them from their distress.\par
\psanum{7}He led them forth by the right way : that they might go to the city where they dwelt.\par
\psanum{8}O that men would therefore praise the Lord for his goodness : and declare the wonders that he doeth for the children of men!\par
\psanum{9}For he satisfieth the empty soul : and filleth the hungry soul with goodness.\par
\psanum{10}Such as sit in darkness, and in the shadow of death : being fast bound in misery and iron;\par
\psanum{11}Because they rebelled against the words of the Lord : and lightly regarded the counsel of the most Highest;\par
\psanum{12}He also brought down their heart through heaviness : they fell down, and there was none to help them.\par
\psanum{13}So when they cried unto the Lord in their trouble : he delivered them out of their distress.\par
\psanum{14}For he brought them out of darkness, and out of the shadow of death : and brake their bonds in sunder.\par
\psanum{15}O that men would therefore praise the Lord for his goodness : and declare the wonders that he doeth for the children of men!\par
\psanum{16}For he hath broken the gates of brass : and smitten the bars of iron in sunder.\par
\psanum{17}Foolish men are plagued for their offence : and because of their wickedness.\par
\psanum{18}Their soul abhorred all manner of meat : and they were even hard at death's door.\par
\psanum{19}So when they cried unto the Lord in their trouble : he delivered them out of their distress.\par
\psanum{20}He sent his word, and healed them : and they were saved from their destruction.\par
\psanum{21}O that men would therefore praise the Lord for his goodness : and declare the wonders that he doeth for the children of men!\par
\psanum{22}That they would offer unto him the sacrifice of thanksgiving : and tell out his works with gladness!\par
\psanum{23}They that go down to the sea in ships : and occupy their business in great waters;\par
\psanum{24}These men see the works of the Lord : and his wonders in the deep.\par
\psanum{25}For at his word the stormy wind ariseth : which lifteth up the waves thereof.\par
\psanum{26}They are carried up to the heaven, and down again to the deep : their soul melteth away because of the trouble.\par
\psanum{27}They reel to and fro, and stagger like a drunken man : and are at their wits' end.\par
\psanum{28}So when they cry unto the Lord in their trouble : he delivereth them out of their distress.\par
\psanum{29}For he maketh the storm to cease : so that the waves thereof are still.\par
\psanum{30}Then are they glad, because they are at rest : and so he bringeth them unto the haven where they would be.\par
\psanum{31}O that men would therefore praise the Lord for his goodness : and declare the wonders that he doeth for the children of men!\par
\psanum{32}That they would exalt him also in the congregation of the people : and praise him in the seat of the elders!\par
\psanum{33}Who turneth the floods into a wilderness : and drieth up the water-springs.\par
\psanum{34}A fruitful land maketh he barren : for the wickedness of them that dwell therein.\par
\psanum{35}Again, he maketh the wilderness a standing water : and water-springs of a dry ground.\par
\psanum{36}And there he setteth the hungry : that they may build them a city to dwell in;\par
\psanum{37}That they may sow their land, and plant vineyards : to yield them fruits of increase.\par
\psanum{38}He blesseth them so that they multiply exceedingly : and suffereth not their cattle to decrease.\par
\psanum{39}And again, when they are minished and brought low : through oppression, through any plague or trouble;\par
\psanum{40}Though he suffer them to be evil intreated through tyrants : and let them wander out of the way in the wilderness;\par
\psanum{41}Yet helpeth he the poor out of misery : and maketh him households like a flock of sheep.\par
\psanum{42}The righteous will consider this, and rejoice : and the mouth of all wickedness shall be stopped.\par
\psanum{43}Whoso is wise will ponder these things : and they shall understand the loving-kindness of the Lord.\par

\psalmsec{Day 22. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 22. Evening Prayer}}

\subbypsalm{Psalm 108. \textit{Paratum cor meum}}
\fancyhead[RE,LO]{Psalm 108}

\lett{O}{God,} my heart is ready, my heart is ready : I will sing and give praise with the best member that I have.\par
\secondlinepsalm{Awake, thou lute, and harp : I myself will awake right early.}
\thirdlinepsalm{I will give thanks unto thee, O Lord, among the people : I will sing praises unto thee among the nations.}
\psanum{4}For thy mercy is greater than the heavens : and thy truth reacheth unto the clouds.\par
\psanum{5}Set up thyself, O God, above the heavens : and thy glory above all the earth.\par
\psanum{6}That thy beloved may be delivered : let thy right hand save them, and hear thou me.\par
\psanum{7}God hath spoken in his holiness : I will rejoice therefore, and divide Sichem, and mete out the valley of Succoth.\par
\psanum{8}Gilead is mine, and Manasses is mine : Ephraim also is the strength of my head.\par
\psanum{9}Judah is my law-giver, Moab is my wash-pot : over Edom will I cast out my shoe, upon Philistia will I triumph.\par
\psanum{10}Who will lead me into the strong city : and who will bring me into Edom?\par
\psanum{11}Hast not thou forsaken us, O God : and wilt not thou, O God, go forth with our hosts?\par
\psanum{12}O help us against the enemy : for vain is the help of man.\par
\psanum{13}Through God we shall do great acts : and it is he that shall tread down our enemies.\par

\subbypsalm{Psalm 109. \textit{Deus, laudem}}
\fancyhead[RE,LO]{Psalm 109}

\lett{H}{old} not thy tongue, O God of my praise : for the mouth of the ungodly, yea, the mouth of the deceitful is opened upon me.\par
\secondlinepsalm{And they have spoken against me with false tongues : they compassed me about also with words of hatred, and fought against me without a cause.}
\thirdlinepsalm{For the love that I had unto them, lo, they take now my contrary part : but I give myself unto prayer.}
\psanum{4}Thus have they rewarded me evil for good : and hatred for my good will.\par
\psanum{5}Set thou an ungodly man to be ruler over him : and let Satan stand at his right hand.\par
\psanum{6}When sentence is given upon him, let him be condemned : and let his prayer be turned into sin.\par
\psanum{7}Let his days be few : and let another take his office.\par
\psanum{8}Let his children be fatherless : and his wife a widow.\par
\psanum{9}Let his children be vagabonds, and beg their bread : let them seek it also out of desolate places.\par
\psanum{10}Let the extortioner consume all that he hath : and let the stranger spoil his labour.\par
\psanum{11}Let there be no man to pity him : nor to have compassion upon his fatherless children.\par
\psanum{12}Let his posterity be destroyed : and in the next generation let his name be clean put out.\par
\psanum{13}Let the wickedness of his fathers be had in remembrance in the sight of the Lord : and let not the sin of his mother be done away.\par
\psanum{14}Let them alway be before the Lord : that he may root out the memorial of them from off the earth.\par
\psanum{15}And that, because his mind was not to do good : but persecuted the poor helpless man, that he might slay him that was vexed at the heart.\par
\psanum{16}His delight was in cursing, and it shall happen unto him : he loved not blessing, therefore shall it be far from him.\par
\psanum{17}He clothed himself with cursing, like as with a raiment : and it shall come into his bowels like water, and like oil into his bones.\par
\psanum{18}Let it be unto him as the cloke that he hath upon him : and as the girdle that he is alway girded withal.\par
\psanum{19}Let it thus happen from the Lord unto mine enemies : and to those that speak evil against my soul.\par
\psanum{20}But deal thou with me, O Lord God, according unto thy Name : for sweet is thy mercy.\par
\psanum{21}O deliver me, for I am helpless and poor : and my heart is wounded within me.\par
\psanum{22}I go hence like the shadow that departeth : and am driven away as the grasshopper.\par
\psanum{23}My knees are weak through fasting : my flesh is dried up for want of fatness.\par
\psanum{24}I became also a reproach unto them : they that looked upon me shaked their heads.\par
\psanum{25}Help me, O Lord my God : O save me according to thy mercy.\par
\psanum{26}And they shall know, how that this is thy hand : and that thou, Lord, hast done it.\par
\psanum{27}Though they curse, yet bless thou : and let them be confounded that rise up against me; but let thy servant rejoice.\par
\psanum{28}Let mine adversaries be clothed with shame : and let them cover themselves with their own confusion, as with a cloke.\par
\psanum{29}As for me, I will give great thanks unto the Lord with my mouth : and praise him among the multitude.\par
\psanum{30}For he shall stand at the right hand of the poor : to save his soul from the unrighteous judges.\par

\psalmsec{Day 23. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 23. Morning Prayer}}

\subbypsalm{Psalm 110. \textit{Dixit Dominus}}
\fancyhead[RE,LO]{Psalm 110}

\lett{T}{he} Lord said unto my Lord : Sit thou on my right hand, until I make thine enemies thy footstool.\par
\secondlinepsalm{The Lord shall send the rod of thy power out of Sion : be thou ruler, even in the midst among thine enemies.}
\thirdlinepsalm{In the day of thy power shall the people offer thee free-will offerings with an holy worship : the dew of thy birth is of the womb of the morning.}
\psanum{4}The Lord sware, and will not repent : Thou art a priest for ever after the order of Melchisedech.\par
\psanum{5}The Lord upon thy right hand : shall wound even kings in the day of his wrath.\par
\psanum{6}He shall judge among the heathen; he shall fill the places with the dead bodies : and smite in sunder the heads over divers countries.\par
\psanum{7}He shall drink of the brook in the way : therefore shall he lift up his head.\par

\subbypsalm{Psalm 111. \textit{Confitebor tibi}}
\fancyhead[RE,LO]{Psalm 111}

\lett{I}{will} give thanks unto the Lord with my whole heart : secretly among the faithful, and in the congregation.\par
\secondlinepsalm{The works of the Lord are great : sought out of all them that have pleasure therein.}
\thirdlinepsalm{His work is worthy to be praised and had in honour : and his righteousness endureth for ever.}
\psanum{4}The merciful and gracious Lord hath so done his marvellous works : that they ought to be had in remembrance.\par
\psanum{5}He hath given meat unto them that fear him : he shall ever be mindful of his covenant.\par
\psanum{6}He hath shewed his people the power of his works : that he may give them the heritage of the heathen.\par
\psanum{7}The works of his hands are verity and judgement : all his commandments are true.\par
\psanum{8}They stand fast for ever and ever : and are done in truth and equity.\par
\psanum{9}He sent redemption unto his people : he hath commanded his covenant for ever; holy and reverend is his Name.\par
\psanum{10}The fear of the Lord is the beginning of wisdom : a good understanding have all they that do thereafter; the praise of it endureth for ever.\par

\subbypsalm{Psalm 112. \textit{Beatus vir}}
\fancyhead[RE,LO]{Psalm 112}

\lett{B}{lessed} is the man that feareth the Lord : he hath great delight in his commandments.\par
\secondlinepsalm{His seed shall be mighty upon earth : the generation of the faithful shall be blessed.}
\thirdlinepsalm{Riches and plenteousness shall be in his house : and his righteousness endureth for ever.}
\psanum{4}Unto the godly there ariseth up light in the darkness : he is merciful, loving, and righteous.\par
\psanum{5}A good man is merciful, and lendeth : and will guide his words with discretion.\par
\psanum{6}For he shall never be moved : and the righteous shall be had in everlasting remembrance.\par
\psanum{7}He will not be afraid of any evil tidings : for his heart standeth fast, and believeth in the Lord.\par
\psanum{8}His heart is established, and will not shrink : until he see his desire upon his enemies.\par
\psanum{9}He hath dispersed abroad, and given to the poor : and his righteousness remaineth for ever; his horn shall be exalted with honour.\par
\psanum{10}The ungodly shall see it, and it shall grieve him : he shall gnash with his teeth, and consume away; the desire of the ungodly shall perish.\par

\subbypsalm{Psalm 113. \textit{Laudate, pueri}}
\fancyhead[RE,LO]{Psalm 113}

\lett{P}{raise} the Lord, ye servants : O praise the Name of the Lord.\par
\secondlinepsalm{Blessed be the Name of the Lord : from this time forth for evermore.}
\thirdlinepsalm{The Lord's Name is praised : from the rising up of the sun unto the going down of the same.}
\psanum{4}The Lord is high above all heathen : and his glory above the heavens.\par
\psanum{5}Who is like unto the Lord our God, that hath his dwelling so high : and yet humbleth himself to behold the things that are in heaven and earth?\par
\psanum{6}He taketh up the simple out of the dust : and lifteth the poor out of the mire;\par
\psanum{7}That he may set him with the princes : even with the princes of his people.\par
\psanum{8}He maketh the barren woman to keep house : and to be a joyful mother of children.\par

\psalmsec{Day 23. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 23. Evening Prayer}}

\subbypsalm{Psalm 114. \textit{In exitu Israel}}
\fancyhead[RE,LO]{Psalm 114}

\lett{W}{hen} Israel came out of Egypt : and the house of Jacob from among the strange people,\par
\secondlinepsalm{Judah was his sanctuary : and Israel his dominion.}
\thirdlinepsalm{The sea saw that, and fled : Jordan was driven back.}
\psanum{4}The mountains skipped like rams : and the little hills like young sheep.\par
\psanum{5}What aileth thee, O thou sea, that thou fleddest : and thou Jordan, that thou wast driven back?\par
\psanum{6}Ye mountains, that ye skipped like rams : and ye little hills, like young sheep?\par
\psanum{7}Tremble, thou earth, at the presence of the Lord : at the presence of the God of Jacob;\par
\psanum{8}Who turned the hard rock into a standing water : and the flint-stone into a springing well.\par

\subbypsalm{Psalm 115. \textit{Non nobis, Domine}}
\fancyhead[RE,LO]{Psalm 115}

\lett{N}{ot} unto us, O Lord, not unto us, but unto thy Name give the praise : for thy loving mercy and for thy truth's sake.\par
\secondlinepsalm{Wherefore shall the heathen say : Where is now their God?}
\thirdlinepsalm{As for our God, he is in heaven : he hath done whatsoever pleased him.}
\psanum{4}Their idols are silver and gold : even the work of men's hands.\par
\psanum{5}They have mouths, and speak not : eyes have they, and see not.\par
\psanum{6}They have ears, and hear not : noses have they, and smell not.\par
\psanum{7}They have hands, and handle not; feet have they, and walk not : neither speak they through their throat.\par
\psanum{8}They that make them are like unto them : and so are all such as put their trust in them.\par
\psanum{9}But thou, house of Israel, trust thou in the Lord : he is their succour and defence.\par
\psanum{10}Ye house of Aaron, put your trust in the Lord : he is their helper and defender.\par
\psanum{11}Ye that fear the Lord, put your trust in the Lord : he is their helper and defender.\par
\psanum{12}The Lord hath been mindful of us, and he shall bless us : even he shall bless the house of Israel, he shall bless the house of Aaron.\par
\psanum{13}He shall bless them that fear the Lord : both small and great.\par
\psanum{14}The Lord shall increase you more and more : you and your children.\par
\psanum{15}Ye are the blessed of the Lord : who made heaven and earth.\par
\psanum{16}All the whole heavens are the Lord's : the earth hath he given to the children of men.\par
\psanum{17}The dead praise not thee, O Lord : neither all they that go down into silence.\par
\psanum{18}But we will praise the Lord : from this time forth for evermore. Praise the Lord.\par

\psalmsec{Day 24. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 24. Morning Prayer}}

%MANUAL ADJUSTMENT:
\subsec{Psalm 116. \textit{Dilexi, quoniam}}
\fancyhead[RE,LO]{Psalm 116}

\lett{I}{am} well pleased : that the Lord hath heard the voice of my prayer;\par
\secondlinepsalm{That he hath inclined his ear unto me : therefore will I call upon him as long as I live.}
\thirdlinepsalm{The snares of death compassed me round about : and the pains of hell gat hold upon me.}
\psanum{4}I shall find trouble and heaviness, and I will call upon the Name of the Lord : O Lord, I beseech thee, deliver my soul.\par
\psanum{5}Gracious is the Lord, and righteous : yea, our God is merciful.\par
\psanum{6}The Lord preserveth the simple : I was in misery, and he helped me.\par
\psanum{7}Turn again then unto thy rest, O my soul : for the Lord hath rewarded thee.\par
\psanum{8}And why? thou hast delivered my soul from death : mine eyes from tears, and my feet from falling.\par
\psanum{9}I will walk before the Lord : in the land of the living.\par
\psanum{10}I believed, and therefore will I speak; but I was sore troubled : I said in my haste, All men are liars.\par
\psanum{11}What reward shall I give unto the Lord : for all the benefits that he hath done unto me?\par
\psanum{12}I will receive the cup of salvation : and call upon the Name of the Lord.\par
\psanum{13}I will pay my vows now in the presence of all his people : right dear in the sight of the Lord is the death of his saints.\par
\psanum{14}Behold, O Lord, how that I am thy servant : I am thy servant, and the son of thine handmaid; thou hast broken my bonds in sunder.\par
\psanum{15}I will offer to thee the sacrifice of thanksgiving : and will call upon the Name of the Lord.\par
\psanum{16}I will pay my vows unto the Lord, in the sight of all his people : in the courts of the Lord's house, even in the midst of thee, O Jerusalem. Praise the Lord.\par

\subbypsalm{Psalm 117. \textit{Laudate Dominum}}
\fancyhead[RE,LO]{Psalm 117}

\lett{O}{praise} the Lord, all ye heathen : praise him, all ye nations.\par
\secondlinepsalm{For his merciful kindness is ever more and more towards us : and the truth of the Lord endureth for ever. Praise the Lord.}

\subbypsalm{Psalm 118. \textit{Confitemini Domino}}
\fancyhead[RE,LO]{Psalm 118}

\lett{O}{give} thanks unto the Lord, for he is gracious : because his mercy endureth for ever.\par
\secondlinepsalm{Let Israel now confess that he is gracious : and that his mercy endureth for ever.}
\thirdlinepsalm{Let the house of Aaron now confess : that his mercy endureth for ever.}
\psanum{4}Yea, let them now that fear the Lord confess : that his mercy endureth for ever.\par
\psanum{5}I called upon the Lord in trouble : and the Lord heard me at large.\par
\psanum{6}The Lord is on my side : I will not fear what man doeth unto me.\par
\psanum{7}The Lord taketh my part with them that help me : therefore shall I see my desire upon mine enemies.\par
\psanum{8}It is better to trust in the Lord : than to put any confidence in man.\par
\psanum{9}It is better to trust in the Lord : than to put any confidence in princes.\par
\psanum{10}All nations compassed me round about : but in the Name of the Lord will I destroy them.\par
\psanum{11}They kept me in on every side, they kept me in, I say, on every side : but in the Name of the Lord will I destroy them.\par
\psanum{12}They came about me like bees, and are extinct even as the fire among the thorns : for in the Name of the Lord I will destroy them.\par
\psanum{13}Thou hast thrust sore at me, that I might fall : but the Lord was my help.\par
\psanum{14}The Lord is my strength, and my song : and is become my salvation.\par
\psanum{15}The voice of joy and health is in the dwellings of the righteous : the right hand of the Lord bringeth mighty things to pass.\par
\psanum{16}The right hand of the Lord hath the pre-eminence : the right hand of the Lord bringeth mighty things to pass.\par
\psanum{17}I shall not die, but live : and declare the works of the Lord.\par
\psanum{18}The Lord hath chastened and corrected me : but he hath not given me over unto death.\par
\psanum{19}Open me the gates of righteousness : that I may go into them, and give thanks unto the Lord.\par
\psanum{20}This is the gate of the Lord : the righteous shall enter into it.\par
\psanum{21}I will thank thee, for thou hast heard me : and art become my salvation.\par
\psanum{22}The same stone which the builders refused : is become the head-stone in the corner.\par
\psanum{23}This is the Lord's doing : and it is marvellous in our eyes.\par
\psanum{24}This is the day which the Lord hath made : we will rejoice and be glad in it.\par
\psanum{25}Help me now, O Lord : O Lord, send us now prosperity.\par
\psanum{26}Blessed be he that cometh in the Name of the Lord : we have wished you good luck, ye that are of the house of the Lord.\par
\psanum{27}God is the Lord who hath shewed us light : bind the sacrifice with cords, yea, even unto the horns of the altar.\par
\psanum{28}Thou art my God, and I will thank thee : thou art my God, and I will praise thee.\par
\psanum{29}O give thanks unto the Lord, for he is gracious : and his mercy endureth for ever.\par

\psalmsec{Day 24. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 24. Evening Prayer}}

\subbypsalm{Psalm 119. \textit{Beati immaculati}}
\fancyhead[RE,LO]{Psalm 119}

\lett{B}{lessed} are those that are undefiled in the way : and walk in the law of the Lord.\par
\secondlinepsalm{Blessed are they that keep his testimonies : and seek him with their whole heart.}
\thirdlinepsalm{For they who do no wickedness : walk in his ways.}
\psanum{4}Thou hast charged : that we shall diligently keep thy commandments.\par
\psanum{5}O that my ways were made so direct : that I might keep thy statutes!\par
\psanum{6}So shall I not be confounded : while I have respect unto all thy commandments.\par
\psanum{7}I will thank thee with an unfeigned heart : when I shall have learned the judgements of thy righteousness.\par
\psanum{8}I will keep thy ceremonies : O forsake me not utterly.\par

\subbypsalm{In quo corriget?}
\fancyhead[RE,LO]{Psalm 119:9}
\lett{W}{herewithal} shall a young man cleanse his way : even by ruling himself after thy word.\par
\secondlinepsalmonenineteen{With my whole heart have I sought thee : O let me not go wrong out of thy commandments.}{10}
\thirdlinepsalmonenineteen{Thy words have I hid within my heart : that I should not sin against thee.}{11}
\psanum{12}Blessed art thou, O Lord : O teach me thy statutes.\par
\psanum{13}With my lips have I been telling : of all the judgements of thy mouth.\par
\psanum{14}I have had as great delight in the way of thy testimonies : as in all manner of riches.\par
\psanum{15}I will talk of thy commandments : and have respect unto thy ways.\par
\psanum{16}My delight shall be in thy statutes : and I will not forget thy word.\par

\subbypsalm{Retribue servo tuo}
\fancyhead[RE,LO]{Psalm 119:17}
\lett{O}{do} well unto thy servant : that I may live, and keep thy word.\par
\secondlinepsalmonenineteen{Open thou mine eyes : that I may see the wondrous things of thy law.}{18}
\thirdlinepsalmonenineteen{I am a stranger upon earth : O hide not thy commandments from me.}{19}
\psanum{20}My soul breaketh out for the very fervent desire : that it hath alway unto thy judgements.\par
\psanum{21}Thou hast rebuked the proud : and cursed are they that do err from thy commandments.\par
\psanum{22}O turn from me shame and rebuke : for I have kept thy testimonies.\par
\psanum{23}Princes also did sit and speak against me : but thy servant is occupied in thy statutes.\par
\psanum{24}For thy testimonies are my delight : and my counsellors.\par

\subbypsalm{Adh{\ae}sit pavimento}
\fancyhead[RE,LO]{Psalm 119:25}
\lett{M}{y} soul cleaveth to the dust : O quicken thou me, according to thy word.\par
\secondlinepsalmonenineteen{I have acknowledged my ways, and thou heardest me : O teach me thy statutes.}{26}
\thirdlinepsalmonenineteen{Make me to understand the way of thy commandments : and so shall I talk of thy wondrous works.}{27}
\psanum{28}My soul melteth away for very heaviness : comfort thou me according unto thy word.\par
\psanum{29}Take from me the way of lying : and cause thou me to make much of thy law.\par
\psanum{30}I have chosen the way of truth : and thy judgements have I laid before me.\par
\psanum{31}I have stuck unto thy testimonies : O Lord, confound me not.\par
\psanum{32}I will run the way of thy commandments : when thou hast set my heart at liberty.\par

\psalmsec{Day 25. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 25. Morning Prayer}}

\subbypsalm{Legem pone}
\fancyhead[RE,LO]{Psalm 119:33}
\lett{T}{each} me, O Lord, the way of thy statutes : and I shall keep it unto the end.\par
\secondlinepsalmonenineteen{Give me understanding, and I shall keep thy law : yea, I shall keep it with my whole heart.}{34}
\thirdlinepsalmonenineteen{Make me to go in the path of thy commandments : for therein is my desire.}{35}
\psanum{36}Incline my heart unto thy testimonies : and not to covetousness.\par
\psanum{37}O turn away mine eyes, lest they behold vanity : and quicken thou me in thy way.\par
\psanum{38}O stablish thy word in thy servant : that I may fear thee.\par
\psanum{39}Take away the rebuke that I am afraid of : for thy judgements are good.\par
\psanum{40}Behold, my delight is in thy commandments : O quicken me in thy righteousness.\par

\subbypsalm{Et veniat super me}
\fancyhead[RE,LO]{Psalm 119:41}
\lett{L}{et} thy loving mercy come also unto me, O Lord : even thy salvation, according unto thy word.\par
\secondlinepsalmonenineteen{So shall I make answer unto my blasphemers : for my trust is in thy word.}{42}
\thirdlinepsalmonenineteen{O take not the word of thy truth utterly out of my mouth : for my hope is in thy judgements.}{43}
\psanum{44}So shall I alway keep thy law : yea, for ever and ever.\par
\psanum{45}And I will walk at liberty : for I seek thy commandments.\par
\psanum{46}I will speak of thy testimonies also, even before kings : and will not be ashamed.\par
\psanum{47}And my delight shall be in thy commandments : which I have loved.\par
\psanum{48}My hands also will I lift up unto thy commandments, which I have loved : and my study shall be in thy statutes.\par

\subbypsalm{Memor esto servi tui}
\fancyhead[RE,LO]{Psalm 119:49}
\lett{O}{think} upon thy servant, as concerning thy word : wherein thou hast caused me to put my trust.\par
\secondlinepsalmonenineteen{The same is my comfort in my trouble : for thy word hath quickened me.}{50}
\thirdlinepsalmonenineteen{The proud have had me exceedingly in derision : yet have I not shrinked from thy law.}{51}
\psanum{52}For I remembered thine everlasting judgements, O Lord : and received comfort.\par
\psanum{53}I am horribly afraid : for the ungodly that forsake thy law.\par
\psanum{54}Thy statutes have been my songs : in the house of my pilgrimage.\par
\psanum{55}I have thought upon thy Name, O Lord, in the night-season : and have kept thy law.\par
\psanum{56}This I had : because I kept thy commandments.\par

\subbypsalm{Portio mea, Domine}
\fancyhead[RE,LO]{Psalm 119:57}
\lett{T}{hou} art my portion, O Lord : I have promised to keep thy law.\par
\secondlinepsalmonenineteen{I made my humble petition in thy presence with my whole heart : O be merciful unto me, according to thy word.}{58}
\thirdlinepsalmonenineteen{I called mine own ways to remembrance : and turned my feet unto thy testimonies.}{59}
\psanum{60}I made haste, and prolonged not the time : to keep thy commandments.\par
\psanum{61}The congregations of the ungodly have robbed me : but I have not forgotten thy law.\par
\psanum{62}At midnight I will rise to give thanks unto thee : because of thy righteous judgements.\par
\psanum{63}I am a companion of all them that fear thee : and keep thy commandments.\par
\psanum{64}The earth, O Lord, is full of thy mercy : O teach me thy statutes.\par

\subbypsalm{Bonitatem fecisti}
\fancyhead[RE,LO]{Psalm 119:65}
\lett{O}{Lord,} thou hast dealt graciously with thy servant : according unto thy word.\par
\secondlinepsalmonenineteen{O learn me true understanding and knowledge : for I have believed thy commandments.}{66}
\thirdlinepsalmonenineteen{Before I was troubled, I went wrong : but now have I kept thy word.}{67}
\psanum{68}Thou art good and gracious : O teach me thy statutes.\par
\psanum{69}The proud have imagined a lie against me : but I will keep thy commandments with my whole heart.\par
\psanum{70}Their heart is as fat as brawn : but my delight hath been in thy law.\par
\psanum{71}It is good for me that I have been in trouble : that I may learn thy statutes.\par
\psanum{72}The law of thy mouth is dearer unto me : than thousands of gold and silver.\par

\psalmsec{Day 25. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 25. Evening Prayer}}

\subbypsalm{Manus tuae fecerunt me}
\fancyhead[RE,LO]{Psalm 119:73}
\lett{T}{hy} hands have made me and fashioned me : O give me understanding, that I may learn thy commandments.\par
\secondlinepsalmonenineteen{They that fear thee will be glad when they see me : because I have put my trust in thy word.}{74}
\thirdlinepsalmonenineteen{I know, O Lord, that thy judgements are right : and that thou of very faithfulness hast caused me to be troubled.}{75}
\psanum{76}O let thy merciful kindness be my comfort : according to thy word unto thy servant.\par
\psanum{77}O let thy loving mercies come unto me, that I may live : for thy law is my delight.\par
\psanum{78}Let the proud be confounded, for they go wickedly about to destroy me : but I will be occupied in thy commandments.\par
\psanum{79}Let such as fear thee, and have known thy testimonies : be turned unto me.\par
\psanum{80}O let my heart be sound in thy statutes : that I be not ashamed.\par

\subbypsalm{Defecit anima mea}
\fancyhead[RE,LO]{Psalm 119:81}
\lett{M}{y} soul hath longed for thy salvation : and I have a good hope because of thy word.\par
\secondlinepsalmonenineteen{Mine eyes long sore for thy word : saying, O when wilt thou comfort me?}{82}
\thirdlinepsalmonenineteen{For I am become like a bottle in the smoke : yet do I not forget thy statutes.}{83}
\psanum{84}How many are the days of thy servant : when wilt thou be avenged of them that persecute me?\par
\psanum{85}The proud have digged pits for me : which are not after thy law.\par
\psanum{86}All thy commandments are true : they persecute me falsely; O be thou my help.\par
\psanum{87}They had almost made an end of me upon earth : but I forsook not thy commandments.\par
\psanum{88}O quicken me after thy loving-kindness : and so shall I keep the testimonies of thy mouth.\par

\subbypsalm{In aeternum. Domine}
\fancyhead[RE,LO]{Psalm 119:89}
\lett{O}{Lord,} thy word : endureth for ever in heaven.\par
\secondlinepsalmonenineteen{Thy truth also remaineth from one generation to another : thou hast laid the foundation of the earth, and it abideth.}{90}
\thirdlinepsalmonenineteen{They continue this day according to thine ordinance : for all things serve thee.}{91}
\psanum{92}If my delight had not been in thy law : I should have perished in my trouble.\par
\psanum{93}I will never forget thy commandments : for with them thou hast quickened me.\par
\psanum{94}I am thine, O save me : for I have sought thy commandments.\par
\psanum{95}The ungodly laid wait for me to destroy me : but I will consider thy testimonies.\par
\psanum{96}I see that all things come to an end : but thy commandment is exceeding broad.\par

\subbypsalm{Quomodo dilexi!}
\fancyhead[RE,LO]{Psalm 119:97}
\lett{L}{ord,} what love have I unto thy law : all the day long is my study in it.\par
\secondlinepsalmonenineteen{Thou through thy commandments hast made me wiser than mine enemies : for they are ever with me.}{98}
\thirdlinepsalmonenineteen{I have more understanding than my teachers : for thy testimonies are my study.}{99}
\psanum{100}I am wiser than the aged : because I keep thy commandments.\par
\psanum{101}I have refrained my feet from every evil way : that I may keep thy word.\par
\psanum{102}I have not shrunk from thy judgements : for thou teachest me.\par
\psanum{103}O how sweet are thy words unto my throat : yea, sweeter than honey unto my mouth.\par
\psanum{104}Through thy commandments I get understanding : therefore I hate all evil ways.\par

\psalmsec{Day 26. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 26. Morning Prayer}}

\subbypsalm{Lucerna pedibus meis}
\fancyhead[RE,LO]{Psalm 119:105}
\lett{T}{hy} word is a lantern unto my feet : and a light unto my paths.\par
\secondlinepsalmonenineteen{I have sworn, and am stedfastly purposed : to keep thy righteous judgements.}{106}
\thirdlinepsalmonenineteen{I am troubled above measure : quicken me, O Lord, according to thy word.}{107}
\psanum{108}Let the free-will offerings of my mouth please thee, O Lord : and teach me thy judgements.\par
\psanum{109}My soul is alway in my hand : yet do I not forget thy law.\par
\psanum{110}The ungodly have laid a snare for me : but yet I swerved not from thy commandments.\par
\psanum{111}Thy testimonies have I claimed as mine heritage for ever : and why? they are the very joy of my heart.\par
\psanum{112}I have applied my heart to fulfil thy statutes alway : even unto the end.\par

\subbypsalm{Iniquos odio habui}
\fancyhead[RE,LO]{Psalm 119:113}
\lett{I}{hate} them that imagine evil things : but thy law do I love.\par
\secondlinepsalmonenineteen{Thou art my defence and shield : and my trust is in thy word.}{114}
\thirdlinepsalmonenineteen{Away from me, ye wicked : I will keep the commandments of my God.}{115}
\psanum{116}O stablish me according to thy word, that I may live : and let me not be disappointed of my hope.\par
\psanum{117}Hold thou me up, and I shall be safe : yea, my delight shall be ever in thy statutes.\par
\psanum{118}Thou hast trodden down all them that depart from thy statutes : for they imagine but deceit.\par
\psanum{119}Thou puttest away all the ungodly of the earth like dross : therefore I love thy testimonies.\par
\psanum{120}My flesh trembleth for fear of thee : and I am afraid of thy judgements.\par

\subbypsalm{Feci judicium}
\fancyhead[RE,LO]{Psalm 119:121}
\lett{I}{deal} with the thing that is lawful and right : O give me not over unto mine oppressors.\par
\secondlinepsalmonenineteen{Make thou thy servant to delight in that which is good : that the proud do me no wrong.}{122}
\thirdlinepsalmonenineteen{Mine eyes are wasted away with looking for thy health : and for the word of thy righteousness.}{123}
\psanum{124}O deal with thy servant according unto thy loving mercy : and teach me thy statutes.\par
\psanum{125}I am thy servant, O grant me understanding : that I may know thy testimonies.\par
\psanum{126}It is time for thee, Lord, to lay to thine hand : for they have destroyed thy law.\par
\psanum{127}For I love thy commandments : above gold and precious stone.\par
\psanum{128}Therefore hold I straight all thy commandments : and all false ways I utterly abhor.\par

\subbypsalm{Mirabilia}
\fancyhead[RE,LO]{Psalm 119:129}
\lett{T}{hy} testimonies are wonderful : therefore doth my soul keep them.\par
\secondlinepsalmonenineteen{When thy word goeth forth : it giveth light and understanding unto the simple.}{130}
\thirdlinepsalmonenineteen{I opened my mouth, and drew in my breath : for my delight was in thy commandments.}{131}
\psanum{132}O look thou upon me, and be merciful unto me : as thou usest to do unto those that love thy Name.\par
\psanum{133}Order my steps in thy word : and so shall no wickedness have dominion over me.\par
\psanum{134}O deliver me from the wrongful dealings of men : and so shall I keep thy commandments.\par
\psanum{135}Shew the light of thy countenance upon thy servant : and teach me thy statutes.\par
\psanum{136}Mine eyes gush out with water : because men keep not thy law.\par

\subbypsalm{Justus es, Domine}
\fancyhead[RE,LO]{Psalm 119:137}
\lett{R}{ighteous} art thou, O Lord : and true is thy judgement.\par
\secondlinepsalmonenineteen{The testimonies that thou hast commanded : are exceeding righteous and true.}{138}
\thirdlinepsalmonenineteen{My zeal hath even consumed me : because mine enemies have forgotten thy words.}{139}
\psanum{140}Thy word is tried to the uttermost : and thy servant loveth it.\par
\psanum{141}I am small, and of no reputation : yet do I not forget thy commandments.\par
\psanum{142}Thy righteousness is an everlasting righteousness : and thy law is the truth.\par
\psanum{143}Trouble and heaviness have taken hold upon me : yet is my delight in thy commandments.\par
\psanum{144}The righteousness of thy testimonies is everlasting : O grant me understanding, and I shall live.\par

\psalmsec{Day 26. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 26. Evening Prayer}}

\subbypsalm{Clamavi in toto corde meo}
\fancyhead[RE,LO]{Psalm 119:145}
\lett{I}{call} with my whole heart : hear me, O Lord, I will keep thy statutes.\par
\secondlinepsalmonenineteen{Yea, even unto thee do I call : help me, and I shall keep thy testimonies.}{146}
\thirdlinepsalmonenineteen{Early in the morning do I cry unto thee : for in thy word is my trust.}{147}
\psanum{148}Mine eyes prevent the night-watches : that I might be occupied in thy words.\par
\psanum{149}Hear my voice, O Lord, according unto thy loving-kindness : quicken me, according as thou art wont.\par
\psanum{150}They draw nigh that of malice persecute me : and are far from thy law.\par
\psanum{151}Be thou nigh at hand, O Lord : for all thy commandments are true.\par
\psanum{152}As concerning thy testimonies, I have known long since : that thou hast grounded them for ever.\par

\subbypsalm{Vide humilitatem}
\fancyhead[RE,LO]{Psalm 119:153}
\lett{O}{consider} mine adversity, and deliver me : for I do not forget thy law.\par
\secondlinepsalmonenineteen{Avenge thou my cause, and deliver me : quicken me, according to thy word.}{154}
\thirdlinepsalmonenineteen{Health is far from the ungodly : for they regard not thy statutes.}{155}
\psanum{156}Great is thy mercy, O Lord : quicken me, as thou art wont.\par
\psanum{157}Many there are that trouble me, and persecute me : yet do I not swerve from thy testimonies.\par
\psanum{158}It grieveth me when I see the transgressors : because they keep not thy law.\par
\psanum{159}Consider, O Lord, how I love thy commandments : O quicken me, according to thy loving-kindness.\par
\psanum{160}Thy word is true from everlasting : all the judgements of thy righteousness endure for evermore.\par

\subbypsalm{Principes persecuti sunt}
\fancyhead[RE,LO]{Psalm 119:161}
\lett{P}{rinces} have persecuted me without a cause : but my heart standeth in awe of thy word.\par
\secondlinepsalmonenineteen{I am as glad of thy word : as one that findeth great spoils.}{162}
\thirdlinepsalmonenineteen{As for lies, I hate and abhor them : but thy law do I love.}{163}
\psanum{164}Seven times a day do I praise thee : because of thy righteous judgements.\par
\psanum{165}Great is the peace that they have who love thy law : and they are not offended at it.\par
\psanum{166}Lord, I have looked for thy saving health : and done after thy commandments.\par
\psanum{167}My soul hath kept thy testimonies : and loved them exceedingly.\par
\psanum{168}I have kept thy commandments and testimonies : for all my ways are before thee.\par

\subbypsalm{Appropinquet deprecatio}
\fancyhead[RE,LO]{Psalm 119:169}
\lett{L}{et} my complaint come before thee, O Lord : give me understanding, according to thy word.\par
\secondlinepsalmonenineteen{Let my supplication come before thee : deliver me, according to thy word.}{170}
\thirdlinepsalmonenineteen{My lips shall speak of thy praise : when thou hast taught me thy statutes.}{171}
\psanum{172}Yea, my tongue shall sing of thy word : for all thy commandments are righteous.\par
\psanum{173}Let thine hand help me : for I have chosen thy commandments.\par
\psanum{174}I have longed for thy saving health, O Lord : and in thy law is my delight.\par
\psanum{175}O let my soul live, and it shall praise thee : and thy judgements shall help me.\par
\psanum{176}I have gone astray like a sheep that is lost : O seek thy servant, for I do not forget thy commandments.\par

\psalmsec{Day 27. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 27. Morning Prayer}}

\subbypsalm{Psalm 120. \textit{Ad Dominum}}
\fancyhead[RE,LO]{Psalm 120}

\lett{W}{hen} I was in trouble I called upon the Lord : and he heard me.\par
\secondlinepsalm{Deliver my soul, O Lord, from lying lips : and from a deceitful tongue.}
\thirdlinepsalm{What reward shall be given or done unto thee, thou false tongue : even mighty and sharp arrows, with hot burning coals.}
\psanum{4}Woe is me, that I am constrained to dwell with Mesech : and to have my habitation among the tents of Kedar.\par
\psanum{5}My soul hath long dwelt among them : that are enemies unto peace.\par
\psanum{6}I labour for peace, but when I speak unto them thereof : they make them ready to battle.\par

\subbypsalm{Psalm 121. \textit{Levavi oculus}}
\fancyhead[RE,LO]{Psalm 121}

\lett{I}{will} lift up mine eyes unto the hills : from whence cometh my help.\par
\secondlinepsalm{My help cometh even from the Lord : who hath made heaven and earth.}
\thirdlinepsalm{He will not suffer thy foot to be moved : and he that keepeth thee will not sleep.}
\psanum{4}Behold, he that keepeth Israel : shall neither slumber nor sleep.\par
\psanum{5}The Lord himself is thy keeper : the Lord is thy defence upon thy right hand;\par
\psanum{6}So that the sun shall not burn thee by day : neither the moon by night.\par
\psanum{7}The Lord shall preserve thee from all evil : yea, it is even he that shall keep thy soul.\par
\psanum{8}The Lord shall preserve thy going out, and thy coming in : from this time forth for evermore.\par

\subbypsalm{Psalm 122. \textit{Laetatus sum}}
\fancyhead[RE,LO]{Psalm 122}

\lett{I}{was} glad when they said unto me : We will go into the house of the Lord.\par
\secondlinepsalm{Our feet shall stand in thy gates : O Jerusalem.}
\thirdlinepsalm{Jerusalem is built as a city : that is at unity in itself.}
\psanum{4}For thither the tribes go up, even the tribes of the Lord : to testify unto Israel, to give thanks unto the Name of the Lord.\par
\psanum{5}For there is the seat of judgement : even the seat of the house of David.\par
\psanum{6}O pray for the peace of Jerusalem : they shall prosper that love thee.\par
\psanum{7}Peace be within thy walls : and plenteousness within thy palaces.\par
\psanum{8}For my brethren and companions' sakes : I will wish thee prosperity.\par
\psanum{9}Yea, because of the house of the Lord our God : I will seek to do thee good.\par

\subbypsalm{Psalm 123. \textit{Ad te levavi oculos meos}}
\fancyhead[RE,LO]{Psalm 123}

\lett{U}{nto} thee lift I up mine eyes : O thou that dwellest in the heavens.\par
%SECOND LINE TOO LONG
\secondlinepsalm{Behold, even as the eyes of servants look unto the hand of their masters, and as the eyes of a maiden unto the hand of her mistress : even}
\noindent
so our eyes wait upon the Lord our God, until he have mercy upon us.\par
\psanum{3}Have mercy upon us, O Lord, have mercy upon us : for we are utterly despised.\par
\psanum{4}Our soul is filled with the scornful reproof of the wealthy : and with the despitefulness of the proud.\par

\subbypsalm{Psalm 124. \textit{Nisi quia Dominus}}
\fancyhead[RE,LO]{Psalm 124}

\lett{I}{f} the Lord himself had not been on our side, now may Israel say : if the Lord himself had not been on our side, when men rose up against us;\par
\secondlinepsalm{They had swallowed us up quick : when they were so wrathfully displeased at us.}
\thirdlinepsalm{Yea, the waters had drowned us : and the stream had gone over our soul.}
\psanum{4}The deep waters of the proud : had gone even over our soul.\par
\psanum{5}But praised be the Lord : who hath not given us over for a prey unto their teeth.\par
\psanum{6}Our soul is escaped even as a bird out of the snare of the fowler : the snare is broken, and we are delivered.\par
\psanum{7}Our help standeth in the Name of the Lord : who hath made heaven and earth.\par

\subbypsalm{Psalm 125. \textit{Qui confidunt}}
\fancyhead[RE,LO]{Psalm 125}

\lett{T}{hey} that put their trust in the Lord shall be even as the mount Sion : which may not be removed, but standeth fast for ever.\par
\secondlinepsalm{The hills stand about Jerusalem : even so standeth the Lord round about his people, from this time forth for evermore.}
\thirdlinepsalm{For the rod of the ungodly cometh not into the lot of the righteous : lest the righteous put their hand unto wickedness.}
\psanum{4}Do well, O Lord : unto those that are good and true of heart.\par
\psanum{5}As for such as turn back unto their own wickedness : the Lord shall lead them forth with the evil-doers; but peace shall be upon Israel.\par

\psalmsec{Day 27. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 27. Evening Prayer}}

\subbypsalm{Psalm 126. \textit{In convertendo}}
\fancyhead[RE,LO]{Psalm 126}

\lett{W}{hen} the Lord turned again the captivity of Sion : then were we like unto them that dream.\par
\secondlinepsalm{Then was our mouth filled with laughter : and our tongue with joy.}
\thirdlinepsalm{Then said they among the heathen : The Lord hath done great things for them.}
\psanum{4}Yea, the Lord hath done great things for us already : whereof we rejoice.\par
\psanum{5}Turn our captivity, O Lord : as the rivers in the south.\par
\psanum{6}They that sow in tears : shall reap in joy.\par
\psanum{7}He that now goeth on his way weeping, and beareth forth good seed : shall doubtless come again with joy, and bring his sheaves with him.\par

\subbypsalm{Psalm 127. \textit{Nisi Dominus}}
\fancyhead[RE,LO]{Psalm 127}

\lett{E}{xcept} the Lord build the house : their labour is but lost that build it.\par
\secondlinepsalm{Except the Lord keep the city : the watchman waketh but in vain.}
\thirdlinepsalm{It is but lost labour that ye haste to rise up early, and so late take rest, and eat the bread of carefulness : for so he giveth his beloved sleep.}
\psanum{4}Lo, children and the fruit of the womb : are an heritage and gift that cometh of the Lord.\par
\psanum{5}Like as the arrows in the hand of the giant : even so are the young children.\par
\psanum{6}Happy is the man that hath his quiver full of them : they shall not be ashamed when they speak with their enemies in the gate.\par

\subbypsalm{Psalm 128. \textit{Beati omnes}}
\fancyhead[RE,LO]{Psalm 128}

\lett{B}{lessed} are all they that fear the Lord : and walk in his ways.\par
\secondlinepsalm{For thou shalt eat the labours of thine hands : O well is thee, and happy shalt thou be.}
\thirdlinepsalm{Thy wife shall be as the fruitful vine : upon the walls of thine house.}
\psanum{4}Thy children like the olive-branches : round about thy table.\par
\psanum{5}Lo, thus shall the man be blessed : that feareth the Lord.\par
\psanum{6}The Lord from out of Sion shall so bless thee : that thou shalt see Jerusalem in prosperity all thy life long.\par
\psanum{7}Yea, that thou shalt see thy children's children : and peace upon Israel.\par

\subbypsalm{Psalm 129. \textit{Saepe expugnaverunt}}
\fancyhead[RE,LO]{Psalm 129}

\lett{M}{any} a time have they fought against me from my youth up : may Israel now say.\par
\secondlinepsalm{Yea, many a time have they vexed me from my youth up : but they have not prevailed against me.}
\thirdlinepsalm{The plowers plowed upon my back : and made long furrows.}
\psanum{4}But the righteous Lord : hath hewn the snares of the ungodly in pieces.\par
\psanum{5}Let them be confounded and turned backward : as many as have evil will at Sion.\par
\psanum{6}Let them be even as the grass growing upon the house-tops : which withereth afore it be plucked up;\par
\psanum{7}Whereof the mower filleth not his hand : neither he that bindeth up the sheaves his bosom.\par
\psanum{8}So that they who go by say not so much as, The Lord prosper you : we wish you good luck in the Name of the Lord.\par

\subbypsalm{Psalm 130. \textit{De profundis}}
\fancyhead[RE,LO]{Psalm 130}

\lett{O}{ut} of the deep have I called unto thee, O Lord : Lord, hear my voice.\par
\secondlinepsalm{O let thine ears consider well : the voice of my complaint.}
\thirdlinepsalm{If thou, Lord, wilt be extreme to mark what is done amiss : O Lord, who may abide it?}
\psanum{4}For there is mercy with thee : therefore shalt thou be feared.\par
\psanum{5}I look for the Lord; my soul doth wait for him : in his word is my trust.\par
\psanum{6}My soul fleeth unto the Lord : before the morning watch, I say, before the morning watch.\par
\psanum{7}O Israel, trust in the Lord, for with the Lord there is mercy : and with him is plenteous redemption.\par
\psanum{8}And he shall redeem Israel : from all his sins.\par

\subbypsalm{Psalm 131. \textit{Domine, non est}}
\fancyhead[RE,LO]{Psalm 131}

\lett{L}{ord,} I am not high-minded : I have no proud looks.\par
\secondlinepsalm{I do not exercise myself in great matters : which are too high for me.}
\thirdlinepsalm{But I refrain my soul, and keep it low, like as a child that is weaned from his mother : yea, my soul is even as a weaned child.}
\psanum{4}O Israel, trust in the Lord : from this time forth for evermore.\par

\psalmsec{Day 28. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 28. Morning Prayer}}

\subbypsalm{Psalm 132. \textit{Memento, Domine}}
\fancyhead[RE,LO]{Psalm 132}

\lett{L}{ord,} remember David : and all his trouble;\par
\secondlinepsalm{How he sware unto the Lord : and vowed a vow unto the Almighty God of Jacob;}
\thirdlinepsalm{I will not come within the tabernacle of mine house : nor climb up into my bed;}
\psanum{4}I will not suffer mine eyes to sleep, nor mine eye-lids to slumber : neither the temples of my head to take any rest;\par
\psanum{5}Until I find out a place for the temple of the Lord : an habitation for the mighty God of Jacob.\par
\psanum{6}Lo, we heard of the same at Ephrata : and found it in the wood.\par
\psanum{7}We will go into his tabernacle : and fall low on our knees before his footstool.\par
\psanum{8}Arise, O Lord, into thy resting-place : thou, and the ark of thy strength.\par
\psanum{9}Let thy priests be clothed with righteousness : and let thy saints sing with joyfulness.\par
\psanum{10}For thy servant David's sake : turn not away the presence of thine Anointed.\par
\psanum{11}The Lord hath made a faithful oath unto David : and he shall not shrink from it;\par
\psanum{12}Of the fruit of thy body : shall I set upon thy seat.\par
\psanum{13}If thy children will keep my covenant, and my testimonies that I shall learn them : their children also shall sit upon thy seat for evermore.\par
\psanum{14}For the Lord hath chosen Sion to be an habitation for himself : he hath longed for her.\par
\psanum{15}This shall be my rest for ever : here will I dwell, for I have a delight therein.\par
\psanum{16}I will bless her victuals with increase : and will satisfy her poor with bread.\par
\psanum{17}I will deck her priests with health : and her saints shall rejoice and sing.\par
\psanum{18}There shall I make the horn of David to flourish : I have ordained a lantern for mine Anointed.\par
\psanum{19}As for his enemies, I shall clothe them with shame : but upon himself shall his crown flourish.\par

\subbypsalm{Psalm 133. \textit{Ecce, quam bonum!}}
\fancyhead[RE,LO]{Psalm 133}

\lett{B}{ehold,} how good and joyful a thing it is : brethren, to dwell together in unity!\par
\secondlinepsalm{It is like the precious ointment upon the head, that ran down unto the beard : even unto Aaron's beard, and went down to the skirts of his clothing.}
\thirdlinepsalm{Like as the dew of Hermon : which fell upon the hill of Sion.}
\psanum{4}For there the Lord promised his blessing : and life for evermore.\par

\subbypsalm{Psalm 134. \textit{Ecce nunc}}
\fancyhead[RE,LO]{Psalm 134}

\lett{B}{ehold} now, praise the Lord : all ye servants of the Lord;\par
\secondlinepsalm{Ye that by night stand in the house of the Lord : even in the courts of the house of our God.}
\thirdlinepsalm{Lift up your hands in the sanctuary : and praise the Lord.}
\psanum{4}The Lord that made heaven and earth : give thee blessing out of Sion.\par

\subbypsalm{Psalm 135. \textit{Laudate Nomen}}
\fancyhead[RE,LO]{Psalm 135}

\lett{O}{praise} the Lord, laud ye the Name of the Lord : praise it, O ye servants of the Lord;\par
\secondlinepsalm{Ye that stand in the house of the Lord : in the courts of the house of our God.}
\thirdlinepsalm{O praise the Lord, for the Lord is gracious : O sing praises unto his Name, for it is lovely.}
\psanum{4}For why? the Lord hath chosen Jacob unto himself : and Israel for his own possession.\par
\psanum{5}For I know that the Lord is great : and that our Lord is above all gods.\par
\psanum{6}Whatsoever the Lord pleased, that did he in heaven and in earth : and in the sea, and in all deep places.\par
\psanum{7}He bringeth forth the clouds from the ends of the world : and sendeth forth lightnings with the rain, bringing the winds out of his treasures.\par
\psanum{8}He smote the first-born of Egypt : both of man and beast.\par
\psanum{9}He hath sent tokens and wonders into the midst of thee, O thou land of Egypt : upon Pharaoh, and all his servants.\par
\psanum{10}He smote divers nations : and slew mighty kings;\par
\psanum{11}Sehon king of the Amorites, and Og the king of Basan : and all the kingdoms of Canaan;\par
\psanum{12}And gave their land to be an heritage : even an heritage unto Israel his people.\par
\psanum{13}Thy Name, O Lord, endureth for ever : so doth thy memorial, O Lord, from one generation to another.\par
\psanum{14}For the Lord will avenge his people : and be gracious unto his servants.\par
\psanum{15}As for the images of the heathen, they are but silver and gold : the work of men's hands.\par
\psanum{16}They have mouths, and speak not : eyes have they, but they see not.\par
\psanum{17}They have ears, and yet they hear not : neither is there any breath in their mouths.\par
\psanum{18}They that make them are like unto them : and so are all they that put their trust in them.\par
\psanum{19}Praise the Lord, ye house of Israel : praise the Lord, ye house of Aaron.\par
\psanum{20}Praise the Lord, ye house of Levi : ye that fear the Lord, praise the Lord.\par
\psanum{21}Praised be the Lord out of Sion : who dwelleth at Jerusalem.\par

\psalmsec{Day 28. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 28. Evening Prayer}}

\subbypsalm{Psalm 136. \textit{Confitemini}}
\fancyhead[RE,LO]{Psalm 136}

\lett{O}{give} thanks unto the \textsc{Lord}, for he is gracious : and his mercy endureth for ever.\par
\secondlinepsalm{O give thanks unto the God of all gods : for his mercy endureth for ever.}
\thirdlinepsalm{O thank the Lord of all lords : for his mercy endureth for ever.}
\psanum{4}Who only doeth great wonders : for his mercy endureth for ever.\par
\psanum{5}Who by his excellent wisdom made the heavens : for his mercy endureth for ever.\par
\psanum{6}Who laid out the earth above the waters : for his mercy endureth for ever.\par
\psanum{7}Who hath made great lights : for his mercy endureth for ever;\par
\psanum{8}The sun to rule the day : for his mercy endureth for ever;\par
\psanum{9}The moon and the stars to govern the night : for his mercy endureth for ever.\par
\psanum{10}Who smote Egypt with their first-born : for his mercy endureth for ever;\par
\psanum{11}And brought out Israel from among them : for his mercy endureth for ever;\par
\psanum{12}With a mighty hand, and stretched out arm : for his mercy endureth for ever.\par
\psanum{13}Who divided the Red sea in two parts : for his mercy endureth for ever;\par
\psanum{14}And made Israel to go through the midst of it : for his mercy endureth for ever.\par
\psanum{15}But as for Pharaoh and his host, he overthrew them in the Red sea : for his mercy endureth for ever.\par
\psanum{16}Who led his people through the wilderness : for his mercy endureth for ever.\par
\psanum{17}Who smote great kings : for his mercy endureth for ever;\par
\psanum{18}Yea, and slew mighty kings : for his mercy endureth for ever;\par
\psanum{19}Sehon king of the Amorites : for his mercy endureth for ever;\par
\psanum{20}And Og the king of Basan : for his mercy endureth for ever;\par
\psanum{21}And gave away their land for an heritage : for his mercy endureth for ever;\par
\psanum{22}Even for an heritage unto Israel his servant : for his mercy endureth for ever.\par
\psanum{23}Who remembered us when we were in trouble : for his mercy endureth for ever;\par
\psanum{24}And hath delivered us from our enemies : for his mercy endureth for ever.\par
\psanum{25}Who giveth food to all flesh : for his mercy endureth for ever.\par
\psanum{26}O give thanks unto the God of heaven : for his mercy endureth for ever.\par
\psanum{27}O give thanks unto the Lord of lords : for his mercy endureth for ever.\par

\subbypsalm{Psalm 137. \textit{Super flumina}}
\fancyhead[RE,LO]{Psalm 137}

\lett{B}{y} the waters of Babylon we sat down and wept : when we remembered thee, O Sion.\par
\secondlinepsalm{As for our harps, we hanged them up : upon the trees that are therein.}
\thirdlinepsalm{For they that led us away captive required of us then a song, and melody in our heaviness : Sing us one of the songs of Sion.}
\psanum{4}How shall we sing the Lord's song : in a strange land?\par
\psanum{5}If I forget thee, O Jerusalem : let my right hand forget her cunning.\par
\psanum{6}If I do not remember thee, let my tongue cleave to the roof of my mouth : yea, if I prefer not Jerusalem in my mirth.\par
\psanum{7}Remember the children of Edom, O Lord, in the day of Jerusalem : how they said, Down with it, down with it, even to the ground.\par
\psanum{8}O daughter of Babylon, wasted with misery : yea, happy shall he be that rewardeth thee, as thou hast served us.\par
\psanum{9}Blessed shall he be that taketh thy children : and throweth them against the stones.\par

\subbypsalm{Psalm 138. \textit{Confitebor tibi}}
\fancyhead[RE,LO]{Psalm 138}

\lett{I}{will} give thanks unto thee, O Lord, with my whole heart : even before the gods will I sing praise unto thee.\par
\secondlinepsalm{I will worship toward thy holy temple, and praise thy Name, because of thy loving-kindness and truth : for thou hast magnified thy Name and thy word above all things.}
\thirdlinepsalm{When I called upon thee, thou heardest me : and enduedst my soul with much strength.}
\psanum{4}All the kings of the earth shall praise thee, O Lord : for they have heard the words of thy mouth.\par
\psanum{5}Yea, they shall sing in the ways of the Lord : that great is the glory of the Lord.\par
\psanum{6}For though the Lord be high, yet hath he respect unto the lowly : as for the proud, he beholdeth them afar off.\par
\psanum{7}Though I walk in the midst of trouble, yet shalt thou refresh me : thou shalt stretch forth thy hand upon the furiousness of mine enemies, and thy right hand shall save me.\par
\psanum{8}The Lord shall make good his loving-kindness toward me : yea, thy mercy, O Lord, endureth for ever; despise not then the works of thine own hands.\par

\psalmsec{Day 29. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 29. Morning Prayer}}

\subbypsalm{Psalm 139. \textit{Domine, probasti}}
\fancyhead[RE,LO]{Psalm 139}

\lett{O}{Lord,} thou hast searched me out and known me : thou knowest my down-sitting and mine up-rising, thou understandest my thoughts long before.\par
\secondlinepsalm{Thou art about my path, and about my bed : and spiest out all my ways.}
\thirdlinepsalm{For lo, there is not a word in my tongue : but thou, O Lord, knowest it altogether.}
\psanum{4}Thou hast fashioned me behind and before : and laid thine hand upon me.\par
\psanum{5}Such knowledge is too wonderful and excellent for me : I cannot attain unto it.\par
\psanum{6}Whither shall I go then from thy Spirit : or whither shall I go then from thy presence?\par
\psanum{7}If I climb up into heaven, thou art there : if I go down to hell, thou art there also.\par
\psanum{8}If I take the wings of the morning : and remain in the uttermost parts of the sea;\par
\psanum{9}Even there also shall thy hand lead me : and thy right hand shall hold me.\par
\psanum{10}If I say, Peradventure the darkness shall cover me : then shall my night be turned to day.\par
\psanum{11}Yea, the darkness is no darkness with thee, but the night is as clear as the day : the darkness and light to thee are both alike.\par
\psanum{12}For my reins are thine : thou hast covered me in my mother's womb.\par
\psanum{13}I will give thanks unto thee, for I am fearfully and wonderfully made : marvellous are thy works, and that my soul knoweth right well.\par
\psanum{14}My bones are not hid from thee : though I be made secretly, and fashioned beneath in the earth.\par
\psanum{15}Thine eyes did see my substance, yet being unperfect : and in thy book were all my members written;\par
\psanum{16}Which day by day were fashioned : when as yet there was none of them.\par
\psanum{17}How dear are thy counsels unto me, O God : O how great is the sum of them!\par
\psanum{18}If I tell them, they are more in number than the sand : when I wake up I am present with thee.\par
\psanum{19}Wilt thou not slay the wicked, O God : depart from me, ye blood-thirsty men.\par
\psanum{20}For they speak unrighteously against thee : and thine enemies take thy Name in vain.\par
\psanum{21}Do not I hate them, O Lord, that hate thee : and am not I grieved with those that rise up against thee?\par
\psanum{22}Yea, I hate them right sore : even as though they were mine enemies.\par
\psanum{23}Try me, O God, and seek the ground of my heart : prove me, and examine my thoughts.\par
\psanum{24}Look well if there be any way of wickedness in me : and lead me in the way everlasting.\par

\subbypsalm{Psalm 140. \textit{Eripe me, Domine}}
\fancyhead[RE,LO]{Psalm 140}

\lett{D}{eliver} me, O Lord, from the evil man : and preserve me from the wicked man.\par
\secondlinepsalm{Who imagine mischief in their hearts : and stir up strife all the day long.}
\thirdlinepsalm{They have sharpened their tongues like a serpent : adders' poison is under their lips.}
\psanum{4}Keep me, O Lord, from the hands of the ungodly : preserve me from the wicked men, who are purposed to overthrow my goings.\par
\psanum{5}The proud have laid a snare for me, and spread a net abroad with cords : yea, and set traps in my way.\par
\psanum{6}I said unto the Lord, Thou art my God : hear the voice of my prayers, O Lord.\par
\psanum{7}O Lord God, thou strength of my health : thou hast covered my head in the day of the battle.\par
\psanum{8}Let not the ungodly have his desire, O Lord : let not his mischievous imagination prosper, lest they be too proud.\par
\psanum{9}Let the mischief of their own lips fall upon the head of them : that compass me about.\par
\psanum{10}Let hot burning coals fall upon them : let them be cast into the fire and into the pit, that they never rise up again.\par
\psanum{11}A man full of words shall not prosper upon the earth : evil shall hunt the wicked person to overthrow him.\par
\psanum{12}Sure I am that the Lord will avenge the poor : and maintain the cause of the helpless.\par
\psanum{13}The righteous also shall give thanks unto thy Name : and the just shall continue in thy sight.\par

\subbypsalm{Psalm 141. \textit{Domine, clamavi}}
\fancyhead[RE,LO]{Psalm 141}

\lett{L}{ord,} I call upon thee, haste thee unto me : and consider my voice when I cry unto thee.\par
\secondlinepsalm{Let my prayer be set forth in thy sight as the incense : and let the lifting up of my hands be an evening sacrifice.}
\thirdlinepsalm{Set a watch, O Lord, before my mouth : and keep the door of my lips.}
\psanum{4}O let not mine heart be inclined to any evil thing : let me not be occupied in ungodly works with the men that work wickedness, lest I eat of such things as please them.\par
\psanum{5}Let the righteous rather smite me friendly : and reprove me.\par
\psanum{6}But let not their precious balms break my head : yea, I will pray yet against their wickedness.\par
\psanum{7}Let their judges be overthrown in stony places : that they may hear my words, for they are sweet.\par
\psanum{8}Our bones lie scattered before the pit : like as when one breaketh and heweth wood upon the earth.\par
\psanum{9}But mine eyes look unto thee, O Lord God : in thee is my trust, O cast not out my soul.\par
\psanum{10}Keep me from the snare that they have laid for me : and from the traps of the wicked doers.\par
\psanum{11}Let the ungodly fall into their own nets together : and let me ever escape them.\par

\psalmsec{Day 29. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 29. Evening Prayer}}

\subbypsalm{Psalm 142. \textit{Voce mea ad Dominum}}
\fancyhead[RE,LO]{Psalm 142}

\lett{I}{cried} unto the Lord with my voice : yea, even unto the Lord did I make my supplication.\par
\secondlinepsalm{I poured out my complaints before him : and shewed him of my trouble.}
\thirdlinepsalm{When my spirit was in heaviness thou knewest my path : in the way wherein I walked have they privily laid a snare for me.}
\psanum{4}I looked also upon my right hand : and saw there was no man that would know me.\par
\psanum{5}I had no place to flee unto : and no man cared for my soul.\par
\psanum{6}I cried unto thee, O Lord, and said : Thou art my hope, and my portion in the land of the living.\par
\psanum{7}Consider my complaint : for I am brought very low.\par
\psanum{8}O deliver me from my persecutors : for they are too strong for me.\par
\psanum{9}Bring my soul out of prison, that I may give thanks unto thy Name : which thing if thou wilt grant me, then shall the righteous resort unto my company.\par

\subbypsalm{Psalm 143. \textit{Domine, exaudi}}
\fancyhead[RE,LO]{Psalm 143}

\lett{H}{ear} my prayer, O Lord, and consider my desire : hearken unto me for thy truth and righteousness' sake.\par
\secondlinepsalm{And enter not into judgement with thy servant : for in thy sight shall no man living be justified.}
\thirdlinepsalm{For the enemy hath persecuted my soul; he hath smitten my life down to the ground : he hath laid me in the darkness, as the men that have been long dead.}
\psanum{4}Therefore is my spirit vexed within me : and my heart within me is desolate.\par
\psanum{5}Yet do I remember the time past; I muse upon all thy works : yea, I exercise myself in the works of thy hands.\par
\psanum{6}I stretch forth my hands unto thee : my soul gaspeth unto thee as a thirsty land.\par
\psanum{7}Hear me, O Lord, and that soon, for my spirit waxeth faint : hide not thy face from me, lest I be like unto them that go down into the pit.\par
\psanum{8}O let me hear thy loving-kindness betimes in the morning, for in thee is my trust : shew thou me the way that I should walk in, for I lift up my soul unto thee.\par
\psanum{9}Deliver me, O Lord, from mine enemies : for I flee unto thee to hide me.\par
\psanum{10}Teach me to do the thing that pleaseth thee, for thou art my God : let thy loving Spirit lead me forth into the land of righteousness.\par
\psanum{11}Quicken me, O Lord, for thy Name's sake : and for thy righteousness' sake bring my soul out of trouble.\par
\psanum{12}And of thy goodness slay mine enemies : and destroy all them that vex my soul; for I am thy servant.\par

\psalmsec{Day 30. Morning Prayer}
\fancyhead[RO,LE]{\textit{Day 30. Morning Prayer}}
\begin{secrubric}
    The Psalms for the \nth{30} Day are repeated on the \nth{31} Day, if there be one.
\end{secrubric}
\subbypsalm{Psalm 144. \textit{Benedictus Dominus}}
\fancyhead[RE,LO]{Psalm 144}

\lett{B}{lessed} be the Lord my strength : who teacheth my hands to war, and my fingers to fight;\par
\secondlinepsalm{My hope and my fortress, my castle and deliverer, my defender in whom I trust : who subdueth my people that is under me.}
\thirdlinepsalm{Lord, what is man, that thou hast such respect unto him : or the son of man, that thou so regardest him?}
\psanum{4}Man is like a thing of nought : his time passeth away like a shadow.\par
\psanum{5}Bow thy heavens, O Lord, and come down : touch the mountains, and they shall smoke.\par
\psanum{6}Cast forth thy lightning, and tear them : shoot out thine arrows, and consume them.\par
\psanum{7}Send down thine hand from above : deliver me, and take me out of the great waters, from the hand of strange children;\par
\psanum{8}Whose mouth talketh of vanity : and their right hand is a right hand of wickedness.\par
\psanum{9}I will sing a new song unto thee, O God : and sing praises unto thee upon a ten-stringed lute.\par
\psanum{10}Thou hast given victory unto kings : and hast delivered David thy servant from the peril of the sword.\par
\psanum{11}Save me, and deliver me from the hand of strange children : whose mouth talketh of vanity, and their right hand is a right hand of iniquity.\par
\psanum{12}That our sons may grow up as the young plants : and that our daughters may be as the polished corners of the temple.\par
\psanum{13}That our garners may be full and plenteous with all manner of store : that our sheep may bring forth thousands and ten thousands in our streets.\par
\psanum{14}That our oxen may be strong to labour, that there be no decay : no leading into captivity, and no complaining in our streets.\par
\psanum{15}Happy are the people that are in such a case : yea, blessed are the people who have the Lord for their God.\par

\subbypsalm{Psalm 145. \textit{Exaltabo te, Deus}}
\fancyhead[RE,LO]{Psalm 145}

\lett{I}{will} magnify thee, O God, my King : and I will praise thy Name for ever and ever.\par
\secondlinepsalm{Every day will I give thanks unto thee : and praise thy Name for ever and ever.}
\thirdlinepsalm{Great is the Lord, and marvellous worthy to be praised : there is no end of his greatness.}
\psanum{4}One generation shall praise thy works unto another : and declare thy power.\par
\psanum{5}As for me, I will be talking of thy worship : thy glory, thy praise, and wondrous works;\par
\psanum{6}So that men shall speak of the might of thy marvellous acts : and I will also tell of thy greatness.\par
\psanum{7}The memorial of thine abundant kindness shall be shewed : and men shall sing of thy righteousness.\par
\psanum{8}The Lord is gracious and merciful : long-suffering and of great goodness.\par
\psanum{9}The Lord is loving unto every man : and his mercy is over all his works.\par
\psanum{10}All thy works praise thee, O Lord : and thy saints give thanks unto thee.\par
\psanum{11}They shew the glory of thy kingdom : and talk of thy power;\par
\psanum{12}That thy power, thy glory, and mightiness of thy kingdom : might be known unto men.\par
\psanum{13}Thy kingdom is an everlasting kingdom : and thy dominion endureth throughout all ages.\par
\psanum{14}The Lord upholdeth all such as fall : and lifteth up all those that are down.\par
\psanum{15}The eyes of all wait upon thee, O Lord : and thou givest them their meat in due season.\par
\psanum{16}Thou openest thine hand : and fillest all things living with plenteousness.\par
\psanum{17}The Lord is righteous in all his ways : and holy in all his works.\par
\psanum{18}The Lord is nigh unto all them that call upon him : yea, all such as call upon him faithfully.\par
\psanum{19}He will fulfil the desire of them that fear him : he also will hear their cry, and will help them.\par
\psanum{20}The Lord preserveth all them that love him : but scattereth abroad all the ungodly.\par
\psanum{21}My mouth shall speak the praise of the Lord : and let all flesh give thanks unto his holy Name for ever and ever.\par

\subbypsalm{Psalm 146. \textit{Lauda, anima mea}}
\fancyhead[RE,LO]{Psalm 146}

\lett{P}{raise} the Lord, O my soul; while I live will I praise the Lord : yea, as long as I have any being, I will sing praises unto my God.\par
\secondlinepsalm{O put not your trust in princes, nor in any child of man : for there is no help in them.}
\thirdlinepsalm{For when the breath of man goeth forth he shall turn again to his earth : and then all his thoughts perish.}
\psanum{4}Blessed is he that hath the God of Jacob for his help : and whose hope is in the Lord his God;\par
\psanum{5}Who made heaven and earth, the sea, and all that therein is : who keepeth his promise for ever;\par
\psanum{6}Who helpeth them to right that suffer wrong : who feedeth the hungry.\par
\psanum{7}The Lord looseth men out of prison : the Lord giveth sight to the blind.\par
\psanum{8}The Lord helpeth them that are fallen : the Lord careth for the righteous.\par
\psanum{9}The Lord careth for the strangers, he defendeth the fatherless and widow : as for the way of the ungodly, he turneth it upside down.\par
\psanum{10}The Lord thy God, O Sion, shall be King for evermore : and throughout all generations.\par

\psalmsec{Day 30. Evening Prayer}
\fancyhead[RO,LE]{\textit{Day 30. Evening Prayer}}

%MANUAL ADJUSTMENT:
\subsec{Psalm 147. \textit{Laudate Dominum}}
\fancyhead[RE,LO]{Psalm 147}

\lett{O}{praise} the Lord, for it is a good thing to sing praises unto our God : yea, a joyful and pleasant thing it is to be thankful.\par
\secondlinepsalm{The Lord doth build up Jerusalem : and gather together the out-casts of Israel.}
\thirdlinepsalm{He healeth those that are broken in heart : and giveth medicine to heal their sickness.}
\psanum{4}He telleth the number of the stars : and calleth them all by their names.\par
\psanum{5}Great is our Lord, and great is his power : yea, and his wisdom is infinite.\par
\psanum{6}The Lord setteth up the meek : and bringeth the ungodly down to the ground.\par
\psanum{7}O sing unto the Lord with thanksgiving : sing praises upon the harp unto our God;\par
\psanum{8}Who covereth the heaven with clouds, and prepareth rain for the earth : and maketh the grass to grow upon the mountains, and herb for the use of men;\par
\psanum{9}Who giveth fodder unto the cattle : and feedeth the young ravens that call upon him.\par
\psanum{10}He hath no pleasure in the strength of an horse : neither delighteth he in any man's legs.\par
\psanum{11}But the Lord's delight is in them that fear him : and put their trust in his mercy.\par
\psanum{12}Praise the Lord, O Jerusalem : praise thy God, O Sion.\par
\psanum{13}For he hath made fast the bars of thy gates : and hath blessed thy children within thee.\par
\psanum{14}He maketh peace in thy borders : and filleth thee with the flour of wheat.\par
\psanum{15}He sendeth forth his commandment upon earth : and his word runneth very swiftly.\par
\psanum{16}He giveth snow like wool : and scattereth the hoar-frost like ashes.\par
\psanum{17}He casteth forth his ice like morsels : who is able to abide his frost?\par
\psanum{18}He sendeth out his word, and melteth them : he bloweth with his wind, and the waters flow.\par
\psanum{19}He sheweth his word unto Jacob : his statutes and ordinances unto Israel.\par
\psanum{20}He hath not dealt so with any nation : neither have the heathen knowledge of his laws.\par

\subbypsalm{Psalm 148. \textit{Laudate Dominum}}
\fancyhead[RE,LO]{Psalm 148}

\lett{O}{praise} the Lord of heaven : praise him in the height.\par
\secondlinepsalm{Praise him, all ye angels of his : praise him, all his host.}
\thirdlinepsalm{Praise him, sun and moon : praise him, all ye stars and light.}
\psanum{4}Praise him, all ye heavens : and ye waters that are above the heavens.\par
\psanum{5}Let them praise the Name of the Lord : for he spake the word, and they were made; he commanded, and they were created.\par
\psanum{6}He hath made them fast for ever and ever : he hath given them a law which shall not be broken.\par
\psanum{7}Praise the Lord upon earth : ye dragons, and all deeps;\par
\psanum{8}Fire and hail, snow and vapours : wind and storm, fulfilling his word;\par
\psanum{9}Mountains and all hills : fruitful trees and all cedars;\par
\psanum{10}Beasts and all cattle : worms and feathered fowls;\par
\psanum{11}Kings of the earth and all people : princes and all judges of the world;\par
\psanum{12}Young men and maidens, old men and children, praise the Name of the Lord : for his Name only is excellent, and his praise above heaven and earth.\par
\psanum{13}He shall exalt the horn of his people; all his saints shall praise him : even the children of Israel, even the people that serveth him.\par

\subbypsalm{Psalm 149. \textit{Cantate Domino}}
\fancyhead[RE,LO]{Psalm 149}

\lett{O}{sing} unto the Lord a new song : let the congregation of saints praise him.\par
\secondlinepsalm{Let Israel rejoice in him that made him : and let the children of Sion be joyful in their King.}
\thirdlinepsalm{Let them praise his Name in the dance : let them sing praises unto him with tabret and harp.}
\psanum{4}For the Lord hath pleasure in his people : and helpeth the meek-hearted.\par
\psanum{5}Let the saints be joyful with glory : let them rejoice in their beds.\par
\psanum{6}Let the praises of God be in their mouth : and a two-edged sword in their hands;\par
\psanum{7}To be avenged of the heathen : and to rebuke the people;\par
\psanum{8}To bind their kings in chains : and their nobles with links of iron.\par
\psanum{9}That they may be avenged of them, as it is written : Such honour have all his saints.\par

\subbypsalm{Psalm 150. \textit{Laudate Dominum}}
\fancyhead[RE,LO]{Psalm 150}

\lett{O}{praise} God in his holiness : praise him in the firmament of his power.\par
\secondlinepsalm{Praise him in his noble acts : praise him according to his excellent greatness.}
\thirdlinepsalm{Praise him in the sound of the trumpet : praise him upon the lute and harp.}
\psanum{4}Praise him in the cymbals and dances : praise him upon the strings and pipe.\par
\psanum{5}Praise him upon the well-tuned cymbals : praise him upon the loud cymbals.\par
\psanum{6}Let every thing that hath breath : praise the Lord.