   \fancyhead[C]{{\LARGE Votive Masses}}
   \fancyhead[RO,LE]{}
   \fancyhead[RE,LO]{}
\section{Our Lady on Saturday}\label{MarySaturday}
\fancyhead[RO,LE]{\textit{Our Lady on Saturday}}
\fancyhead[RE,LO]{Daily Office}
\subby{Daily Office Propers}
\subbysub{Through the Year}
\begin{inhead}
    Evensong
\end{inhead}\par\noindent
\antiphon{Mag.}{O blessed Mother {\dag} and spotless Virgin, thou glorious Queen of the world, intercede for us to the Lord.}

\needspace{4\baselineskip}
\begin{rubric}
	The Collect is said as below in Mattins. The following Commemoration is then said.
\end{rubric}
\lett{D}{efend} us, we beseech thee, O Lord, from all perils of mind and body: and at the intercession of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.} and of all Saints, graciously bestow upon us both peace and safety: that all adversity and error being done away, thy Church may serve thee in untroubled freedom. Through.
\begin{inhead}
    Mattins
\end{inhead}\par\noindent
\antiphon{Ben.}{O ever blessed Mother of God, {\dag} Mary ever Virgin, temple of the Godhead, hallowed shrine of the Holy Spirit, thou only, above all others, wast acceptable to our Lord Jesus Christ: pray for the people, entreat for the clergy, intercede for all womankind vowed unto God.}

%DOUBLE CHECK LATIN TO MAKE SURE RUBRIC IS CORRECT:
\begin{rubric}
	Collect as in St. Mary in Eastertide (p. \pageref{SPMaryInEaster}), followed by the above Commemoration.
\end{rubric}
%\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may rejoice in continual health of mind and body: and that by the glorious intercession of blessed Mary ever Virgin we may be delivered from this present heaviness, and have the fruition of eternal joy. Through.

\subbysub{After Christmas}
\begin{inhead}
    From Saturday after the Epiphany Octave through Saturday before Purification.
\end{inhead}
\antiphon{Mag.}{Great {\dag} is the mystery of the inheritance: the womb of her that knew not man is become the temple of the Godhead: by taking flesh of her, he was no way defiled: all the nations shall gather, saying: Glory be to thee, O Lord.}

\antiphon{Ben.}{A great and wondrous mystery {\dag} is made known to us this day; a new thing is wrought in both natures: God is made man; that which was, remained, and that which was not, he assumed; suffering no confusion, nor yet division.}

\begin{rubric}
	Collect as in St. Mary in Christmastide (p. \pageref{SPMaryPostChristmas}), followed by the above Commemoration.
\end{rubric}
%\lett{O}{God,} who by the fruitful virginity of blessed Mary hast bestowed upon mankind the reward of eternal salvation: grant, we beseech thee, that we may know the help of her intercession, through whom we have been accounted worthy to receive the Author of our life, Jesus Christ, thy Son our Lord: Who with thee.
\subsubsec{In Eastertide}
\antiphon{Mag. \& Ben.}{O Queen of heaven, {\dag} be joyful, alleluia; Because he whom so meetly thou barest, alleluia, Hath arisen, as he promised, alleluia: Pray for us to the Father, alleluia.}
\begin{rubric}
	Collect as in St. Mary in Eastertide (p. \pageref{SPMaryInEaster}). The following Commemoration is then said.
\end{rubric}
\needspace{4\baselineskip}
\lett{O}{God,} who didst will that thy Son should for us undergo the burden of the Cross, that he might deliver us from the power of the enemy: grant unto us thy servants that we may attain unto the joy of his resurrection. Through the same.
%\needspace{4\baselineskip}
%\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may rejoice in continual health of mind and body: and that by the glorious intercession of blessed Mary ever Virgin we may be delivered from this present heaviness, and have the fruition of eternal joy. Through.
\subby{Mass Propers}
\subbysub{Advent}
\fancyhead[RE,LO]{Advent}
\needspace{4\baselineskip}
\begin{inhead}
    Introit
\end{inhead}
\lett{D}{rop} down dew, ye heavens, from above, and let the skies pour down righteousness: let the earth open, and bring forth a Saviour. \textit{Ps.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob.
\begin{inhead}
    Collect
\end{inhead}
%\lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the Blessed Virgin Mary at the message of an Angel: grant that we thy humble servants, who believe her to be truly the Mother of God; may be aided by her intercession in thy sight. Through the same.
\begin{rubric}
    The Collects \emph{of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}), of the Advent Feria, and \emph{of the Holy Ghost} (p. \pageref{SPHolyGhost}) are here said.
\end{rubric}
\readingcitation{Epistle}{Isaiah 7:10}
\lett{I}{n those days:} The \divineName{Lord} spake again unto Ahaz, saying, Ask thee a sign of the \divineName{Lord} thy God; ask it either in the depth, or in the height above. But Ahaz said, I will not ask, neither will I tempt the \divineName{Lord}. And he said, Hear ye now, O house of David; Is it a small thing for you to weary men, but will ye weary my God also? Therefore the Lord himself shall give you a sign; Behold, a virgin shall conceive, and bear a son, and shall call his name Immanuel. Butter and honey shall he eat, that he may know to refuse the evil, and choose the good.
\gradall{Lift up your heads, O ye gates: and be ye lift up, ye everlasting doors: and the King of glory shall come in. ℣. Who shall ascend into the hill of the Lord? or who shall rise up in his holy place? Even he that hath clean hands, and a pure heart.}{Alleluia, alleluia. ℣. Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women. Alleluia.}
\readingcitation{Gospel}{Luke 1:26}
\lett{A}{t that time:} The Angel Gabriel was sent from God unto a city of Galilee, named Nazareth, To a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin's name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name {\textsc{Jesus}}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: And he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word.
\offertory{Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women, and blessed is the fruit of thy womb.}
\begin{inhead}
    Secret
\end{inhead}
%\lett{S}{tablish} in our hearts, we beseech thee, O Lord, the mysteries of the true faith: that like as we acknowledge thy Son conceived of a Virgin to be very God and very man; so by the power of his Resurrection we may be found worthy to attain unto everlasting gladness. Through the same.
\begin{rubric}
    The Secrets \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}), of the Advent Feria, and \emph{of the Holy Ghost} (p. \pageref{SPHolyGhost}) are here said.
\end{rubric}

\communion{Behold, a Virgin shall conceive, and bear a son: and shall call his name Immanuel.}
\begin{inhead}
    Postcommunion
\end{inhead}
%\lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts: that, as we have known the incarnation of thy Son Jesus Christ by the message of an Angel; so by his cross and passion we may be brought unto the glory of his resurrection. Through the same.
\begin{rubric}
    The Postcommunions \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}), of the Advent Feria, and \emph{of the Holy Ghost} (p. \pageref{SPHolyGhost}) are here said.
\end{rubric}

\subby{Epiphanytide}
\fancyhead[RE,LO]{Epiphanytide}
\begin{rubric}
    {From the Octave of the Epiphany (or in Votive Masses, from 26 December) until 1 February, inclusive.}
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Introit
\end{inhead}
\lett{A}{ll} the rich among the people shall make their supplication before thee: the Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy and gladness. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\needspace{4\baselineskip}
\begin{inhead}
    Collect
\end{inhead}
\lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son our Lord. Who liveth.
\readingcitation{Epistle}{Titus 3:4}
\lett{D}{early beloved:} The kindness and love of God our Saviour toward man appeared, Not by works of righteousness which we have done, but according to his mercy he saved us, by the washing of regeneration, and renewing of the Holy Ghost; Which he shed on us abundantly through Jesus Christ our Saviour; That being justified by his grace, we should be made heirs according to the hope of eternal life:
%1933 EM:
in Christ Jesus our Lord.
\gradall{Thou art fairer than the children of men: full of grace are thy lips. ℣. My heart is inditing of a good matter: I speak of the things which I have made unto the King: my tongue is the pen of a ready writer.}{Alleluia, alleluia. ℣. After childbirth, O Virgin, thou didst remain inviolate: Mother of God, intercede for us. Alleluia.}
\readingcitation{Gospel}{Luke 2:15}
\lett{A}{t that time:} The shepherds said one to another, Let us now go even unto Bethlehem, and see this thing which is come to pass, which the Lord hath made known unto us. And they came with haste, and found Mary, and Joseph, and the babe lying in a manger. And when they had seen it, they made known abroad the saying which was told them concerning this child. And all they that heard it wondered at those things which were told them by the shepherds. But Mary kept all these things, and pondered them in her heart. And the shepherds returned, glorifying and praising God for all the things that they had heard and seen, as it was told unto them.
\offertory{Happy indeed art thou, O sacred Virgin Mary, and most worthy of all praise: for out of thee hath arisen the sun of righteousness, Christ our God.}
\needspace{4\baselineskip}
\begin{inhead}
    Secret
\end{inhead}
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary, ever Virgin, may this oblation avail for our prosperity and peace, both now and for ever. Through.
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\needspace{4\baselineskip}
\begin{inhead}
    Postcommunion
\end{inhead}
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary, the Virgin Mother of God, make us partakers of thy heavenly healing. Through the same.


\subby{Septuagesimatide}
\fancyhead[RE,LO]{Septuagesimatide}
\begin{rubric}
    {From the Saturday after the Purification of the Blessed Virgin Mary to the Saturday after Sexagesima Sunday, or in Votive Masses till the Saturday after Passion Sunday, inclusive, according to the Rubrics.}
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Introit
\end{inhead}
\lett{H}{ail,} O Mother most holy, who in childbirth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\needspace{4\baselineskip}
\begin{inhead}
    Collect
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, by the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\begin{rubric}
    The Collects of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 24:9}
\lett{H}{e} created me from the beginning before the world, and I shall never fail. In the holy tabernacle I served before him; and so was I established in Sion. Likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance.
%Verse not in the KJV is not included.
\gradall{Blessed and venerable art thou O Virgin Mary: who without spot wast found Mother of the Saviour. ℣. Virgin Mother of God, he whom the whole world containeth not, being made man lay hid in thy womb.}{Alleluia, alleluia. ℣. Now hath blossomed Jesse's rod: a Virgin bears both man and God: God restoreth peace to men, high and low are one again. Alleluia.}
\begin{rubric}
    After Septuagesima Sunday, use the following Tract instead of the Alleluia.
\end{rubric}
\tract{Rejoice, O Virgin Mary, alone thou hast destroyed all heresies. ℣. Who didst believe the words of the Archangel Gabriel. ℣. Whilst a Virgin, thou didst bring forth him who is God and man: and after childbirth, O Virgin, inviolate didst remain. ℣. Mother of God, intercede for us.}
\readingcitation{Gospel}{Luke 11:27}
\lett{A}{t that time:} As Jesus spake to the multitudes, a certain woman of the company lifted up her voice, and said unto him, Blessed is the womb that bare thee, and the paps which thou hast sucked. But he said, Yea rather, blessed are they that hear the word of God, and keep it.
\offertory{Happy indeed art thou, O sacred Virgin Mary, and most worthy of all praise: for out of thee hath arisen the sun of righteousness, Christ our God.}
\needspace{4\baselineskip}
\begin{inhead}
    Secret
\end{inhead}
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and ever. Through.
\begin{rubric}
    The Secrets of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\needspace{4\baselineskip}
\begin{inhead}
    Postcommunion
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord: that we, who have received these aids to our salvation, may at all times and in all places be protected by the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. Through.
\begin{rubric}
    The Postcommunions of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}

\subby{Eastertide}
\fancyhead[RE,LO]{Eastertide}
\begin{rubric}
    {From Saturday after Low Sunday until Saturday after the Fourth Sunday after Easter; or in Votive Masses from Wednesday after Easter till Saturday in the Octave of Pentecost, inclusive, according to the Rubrics.}
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Introit
\end{inhead}
\lett{H}{ail,} O Mother most holy, who in childbirth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever, alleluia, alleluia. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\begin{rubric}
    On Saturday, \emph{Gloria in excelsis} is said.
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Collect
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\begin{rubric}
    The Collects of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 24:9}
\lett{H}{e} created me from the beginning before the world, and I shall never fail. In the holy tabernacle I served before him; and so was I established in Sion. Likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance.
%Verse not in the KJV is not included.
\alleluia{Alleluia, alleluia. ℣. Now hath blossomed Jesse's rod: a Virgin bears both man and God: God restoreth peace to men, high and low are one again. Alleluia. ℣. Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women. Alleluia.}
\readingcitation{Gospel}{John 19:25}
\lett{A}{t that time:} There stood by the cross of Jesus his mother, and his mother's sister, Mary the wife of Cleophas, and Mary Magdalene. When Jesus therefore saw his mother, and the disciple standing by, whom he loved, he saith unto his mother, Woman, behold thy son! Then saith he to the disciple, Behold thy mother! And from that hour that disciple took her unto his own home.
\offertory{Blessed art thou, O Virgin Mary, who didst bear the Creator of all things: thou broughtest forth him who made thee, and for ever remainest a Virgin, alleluia.}
\needspace{4\baselineskip}
\begin{inhead}
    Secret
\end{inhead}
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and ever. Through.
\begin{rubric}
    The Secrets of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father, alleluia.}
\needspace{4\baselineskip}
\begin{inhead}
    Postcommunion
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord: that we who have received these aids to our salvation, may at all times and in all places be protected by the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. Through.
\begin{rubric}
    The Postcommunions of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}

\subby{Trinitytide}
\fancyhead[RE,LO]{Trinitytide}
\begin{rubric}
    {From the Saturday after the Divine Compassion Octave, or in Votives from the Monday after Trinity Sunday, until the Saturday before the First Sunday in Advent, inclusive.}
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Introit
\end{inhead}
\lett{H}{ail,} O Mother most holy, who in childbirth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\begin{rubric}
    On Saturday, \emph{Gloria in excelsis} is said.
\end{rubric}
\needspace{4\baselineskip}
\begin{inhead}
    Collect
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, by the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\begin{rubric}
    The Collects of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 24:9}
\lett{H}{e} created me from the beginning before the world, and I shall never fail. In the holy tabernacle I served before him; and so was I established in Sion. Likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance.
%Verse not in the KJV is not included.
\gradall{Blessed and venerable art thou, O Virgin Mary: who without spot wast found Mother of the Saviour. ℣. Virgin Mother of God, he whom the whole world containeth not, being made man lay hid in thy womb.}{Alleluia, alleluia. ℣. After childbirth, O Virgin, thou didst remain inviolate : Mother of God, intercede for us. Alleluia.}
\readingcitation{Gospel}{Luke 11:27}
\lett{A}{t that time:} As Jesus spake to the multitudes, a certain woman of the company lifted up her voice, and said unto him, Blessed is the womb that bare thee, and the paps which thou hast sucked. But he said, Yea rather, blessed are they that hear the word of God, and keep it.
\offertory{Hall, Mary, full of grace; the Lord is with thee: blessed art thou among women, and blessed is the fruit of thy womb.}
\needspace{4\baselineskip}
\begin{inhead}
    Secret
\end{inhead}
\lett{T}{hrough} thy mercy, Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and ever. Through.
\begin{rubric}
    The Secrets of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\needspace{4\baselineskip}
\begin{inhead}
    Postcommunion
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord: that we who have received these aids to our salvation may at all times and in all places be protected by the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. Through.
\begin{rubric}
    The Postcommunions of the Holy Ghost (p. \pageref{SPHolyGhost}) and Against the Persecutors of the Church (p. \pageref{SPAgainst}), or for the Chief Bishop (p. \pageref{SPChiefBishop}), are then said.
\end{rubric}
\clearpage

\section{Mass of the Angels}\label{MassOfTheAngels}
\fancyhead[RO,LE]{\textit{Angels}}
\fancyhead[RE,LO]{}
\begin{rubric}
    It is custom, when rubrics permit, to celebrate a Votive Mass for the Angels on Tuesday and for the death of a departed baptised Child.
\end{rubric}
\subby{Introit}
\lett{O}{praise} the Lord, ye Angels of his: ye that excel in strength, ye that fulfil his commandment, and hearken unto the voice of his words (Alleluia, alleluia.) \textit{Ps.} Praise the Lord, O my soul: and all that is within me praise his holy name.

\begin{rubric}
    \emph{Gloria in excelsis} is always said.
\end{rubric}

\subby{Collect}
\lett{O}{everlasting} God, who hast ordained and constituted the services of Angels and men in a wonderful order: mercifully grant; that as thy holy Angels alway do thee service in heaven, so by thy appointment they may succour and defend us on earth. Through.

\readingcitation{Epistle}{Revelation 5:11}
\lett{I}{n those days:} I heard the voice of many Angels round about the throne and the beasts and the elders: and the number of them was ten thousand times ten thousand, and thousands of thousands; Saying with a loud voice, Worthy is the Lamb that was slain to receive power, and riches, and wisdom, and strength, and honour, and glory, and blessing. And every creature which is in heaven, and on the earth, and under the earth, and such as are in the sea, and all that are in them, heard I saying, Blessing, and honour, and glory, and power, be unto him that sitteth upon the throne, and unto the Lamb for ever and ever. And the four beasts said, Amen. And the four and twenty elders fell down and worshipped him that liveth for ever and ever.

\gradall{O praise the Lord of heaven: praise in him in the height. ℣. Praise him, all ye Angels of his: praise him, all his host.}{Alleluia, alleluia. ℣. In the presence of the Angels will I sing praise unto thee: I will worship toward thy holy temple, and praise thy name. Alleluia.}

\begin{rubric}
    After Septuagesima, the following Tract is said instead of the Gradual \& Alleluia Verse.
\end{rubric}

\tract{O praise the Lord, ye Angels of his: ye that excel in strength, ye that fulfil his commandment. ℣. O praise the Lord, all ye his hosts: ye servants of his that do his pleasure. ℣. O speak good of the Lord, all ye works of his: in all places of his dominion, praise thou the Lord, O my soul.}

\begin{rubric}
    In Eastertide, the following Greater Alleluia is said instead of the Gradual \& the Lesser Alleluia Verse.
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. In the presence of the Angels will I sing praise unto thee: I will worship toward thy holy temple, and praise thy name. Alleluia. ℣. The Angel of the Lord descended from heaven, and came and rolled back the stone, and sat upon it. Alleluia.}

\readingcitation{Gospel}{John 1:47}
\lett{A}{t that time:} Jesus saw Nathanael coming to him, and saith of him, Behold an Israelite indeed, in whom is no guile! Nathanael saith unto him, Whence knowest thou me? Jesus answered and said unto him, Before that Philip called thee, when thou wast under the fig tree, I saw thee. Nathanael answered and saith unto him, Rabbi, thou art the Son of God; thou art the King of Israel. Jesus answered and said unto him, Because I said unto thee, I saw thee under the fig tree, believest thou? thou shalt see greater things than these. And he saith unto him, Verily, verily, I say unto you, Hereafter ye shall see heaven open, and the angels of God ascending and descending upon the Son of man.

\offertory{An Angel stood at the altar of the temple, having a golden censer in his hand: and there was given unto him much incense: and the smoke of the incense ascended up before God. (Alleluia.)}

\subby{Secret}
\lett{W}{e} offer thee, O Lord, sacrifices of praise, humbly beseeching thee: that, by the prayers of the Angels interceding for us, thou wouldest both graciously accept the same, and grant that they may avail to our salvation. Through.

\communion{Angels, Archangels, Thrones and Dominations, Principalities and Powers, Virtues of heaven, Cherubim and Seraphim, bless ye the Lord for ever. (Alleluia.)}

\subby{Postcommunion}
\lett{O}{Lord,} who hast filled us with thy heavenly benediction, we humbly beseech thee: that through the assistance of thy holy Angels and Archangels we may feel the benefit of the mystery which in our weakness we celebrate. Through.

\clearpage

\section{Holy Matrimony}\label{MatrimonyMass}
\fancyhead[RO,LE]{\textit{Matrimony}}
\fancyhead[RE,LO]{}
\begin{rubric}
    On a Sunday or \nth{1} or \nth{2} Double, or Vigil of Whitsunday, or within the Octaves of Epiphany, Easter, Whitsunday, or Corpus Christi; Mass of the Day is said, with Creed and \emph{Gloria in excelsis}, if the Mass require them, with Commemoration of the following Mass at a wedding, and the other things that belong.\par
    On other days (excepting All Souls' Day, on which both the Votive Mass and the solemn Nuptial benediction are forbidden) the following Mass is said, even if a Greater Double or less occur.
\end{rubric}

\subby{Introit}
\lett{T}{he} God of Israel make you one: and may he be with you, even as he had mercy of two that were the only-begotten of their fathers: and now, O Lord, grant them to bless thee yet more abundantly. (Alleluia, alleluia.) \textit{Ps.} Blessed are all they that fear the Lord: and walk in his ways.

\subby{Collect}
\lett{O}{eternal} God, we humbly beseech thee, favourably to behold these thy servants now joined in wedlock according to thy holy ordinance; and grant that they, seeking first thy kingdom and thy righteousness, may obtain the manifold blessings of thy grace. Through.
 
\readingcitation{Epistle}{Ephesians 5:20}
\lett{B}{rethren:} Give thanks always for all things unto God and the Father in the name of our Lord Jesus Christ; submit-ting yourselves one to another in the fear of God. Wives, submit yourselves unto your own husbands, as unto the Lord. For the husband is the head of the wife, even as Christ is the head of the church: and he is the saviour of the body. Therefore as the church is subject unto Christ, so let the wives be to their own husbands in every thing. Husbands, love your wives, even as Christ also loved the church, and gave himself for it; that he might sanctify and cleanse it with the washing of water by the word, that he might present it to himself a glorious church, not having spot, or wrinkle, or any such thing; but that it should be holy and without blemish. So ought men to love their wives as their own bodies. He that loveth his wife loveth himself. For no man ever yet hated his own flesh; but nourisheth and cherisheth it, even as the Lord the church: for we are members of his body, of his flesh, and of his bones. For this cause shall a man leave his father and mother, and shall be joined unto his wife, and they two shall be one flesh. This is a great mystery: but I speak concerning Christ and the church. Nevertheless let every one of you in particular so love his wife even as himself; and the wife see that she reverence her husband.

\gradall{Thy wife shall be as the fruitful vine upon the walls of thine house. ℣. Thy children like the olive-branches round about thy table.}{Alleluia, alleluia. ℣. The Lord send you help from the sanctuary: and strengthen you out of Sion. Alleluia.}

\begin{rubric}
    After Septuagesima, the following Tract is said instead of the Gradual \& Alleluia Verse.
\end{rubric}

\tract{Lo, thus shall the man be blessed, that feareth the Lord. ℣. The Lord from out of Sion shall so bless thee: that thou shalt see Jerusalem in prosperity all thy life long. ℣. Yea, that thou shalt see thy children's children: and peace upon Israel.}

\begin{rubric}
    In Eastertide, the following Greater Alleluia is said instead of the Gradual \& the Lesser Alleluia Verse.
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. The Lord send you help from the sanctuary: and strengthen you out of Sion. Alleuia. ℣. The Lord that made heaven and earth: give you blessing out of Sion. Alleluia.}

\readingcitation{Gospel}{Matthew 19:4}
\lett{A}{t that time:} The Pharisees came unto Jesus, tempting him, and saying unto him: Is it lawful for a man to put away his wife for every cause? And he answered and said unto them, Have ye not read, that he which made them at the beginning made them male and female, and said, For this cause shall a man leave father and mother, and shall cleave to his wife: and they twain shall be one flesh? Wherefore they are no more twain, but one flesh. What therefore God hath joined together, let not man put asunder.

\offertory{My hope hath been in thee, O Lord: I have said: Thou art my God: my time is in thy hand. (Alleluia.)}

\subby{Secret}
\lett{O}{Lord,} we beseech thee to accept this gift which we present unto thee for the sacred ordinance of matrimony: that this work begun by the bounty of thy goodness may be disposed according to thy will. Through.

\begin{rubric}
    Once the Lord's Prayer is ended, the Priest immediately moves to the Epistle corner and turning towards the Bridegroom and Bride kneeling before the Altar, says the following prayers over them.
\end{rubric}
\lett{O}{merciful} Lord, and heavenly Father, by whose gracious gift mankind is increased: We beseech thee assist with thy blessing these two persons, that they may both be fruitful in procreation of children; and also live together so long in godly love and honesty, that they may see their children's children, unto the third and fourth generation, unto thy praise and honour: through Jesus Christ our Lord. \textit{Amen.}\par
\lett{O}{God} which by thy mighty power hast made all things of naught, which also after other things set in order didst appoint that out of man (created after thine own image and similitude) woman should take her beginning: and, knitting them together, didst teach, that it should never be lawful to put asunder those, whom thou by matrimony hadst made one: O God, which hast consecrated the state of matrimony to such an excellent mystery, that in it is signified and represented the spiritual marriage and unity betwixt Christ and his Church: Look mercifully upon these thy servants, that both this man may love his wife, according to thy Word, (as Christ did love his spouse the Church, who gave himself for it, loving and cherishing it even as his own flesh;) And also that this woman may be loving and amiable to her husband as Rachel, wise as Rebecca, faithful and obedient as Sara; And in all quietness, sobriety, and peace, be a follower of holy and godly matrons. O Lord, bless them both, and grant them to inherit thy everlasting kingdom, through Jesu Christ our Lord. \textit{Amen.}

\begin{rubric}
    The Priest then moving to the middle of the Altar, he continues with the \emph{Deliver us}.\par
    \textsc{Note,} during the Distribution of Communion, the Bridegroom and Bride should be communed first.
\end{rubric}

\communion{Lo, thus shall the man be blessed that feareth the Lord: yea, thou shalt see thy children's children: and peace upon Israel. (Alleluia.)}

\subby{Postcommunion}
\lett{W}{e} beseech thee, almighty God: let thy gracious favour accompany the institution of thy providence; that those whom thou hast joined together in lawful union thou wouldest preserve in lasting peace. Through.

\begin{rubric}
    Having said the Dismissal, the Priest, before blessing the People, turns to the Bridegroom and Bride and says:
\end{rubric}
\lett{A}{lmighty} God, which at the beginning did create our first parents Adam and Eve, and did sanctify and join them together in marriage: Pour upon you the riches of his grace, sanctify and {\ding{64}} bless you, that ye may please him both in body and soul; and live together in holy love unto your lives' end. \textit{Amen.}
\begin{rubric}
    The Priest shall solemnly charge them that they shall be faithful to each other; that they shall remain chaste at the time of prayer, and especially during fasts and solemnities; that they shall love each other, and remain in the fear of God. Then he sprinkles them with holy water, and having said \emph{Let this my bounden duty}, gives the blessing, and reads the Last Gospel as usual.
\end{rubric}

\clearpage
\section{Masses for the Dead}\label{RequiemMasses}
\fancyhead[RO,LE]{\textit{Requiem Masses}}
\fancyhead[RE,LO]{}
\begin{rubric}
    Instead of \emph{Gloria Patri} is said the \emph{Rest eternal}.
\end{rubric}
\vspace{-2ex}
\subby{On the Commemoration of all the Faithful Departed}
\begin{rubric}
    On this day, Masses are said of All Souls' Day (p. \pageref{AllSouls}).
\end{rubric}
\begin{rubric}
    For a departed Bishop or Priest, as well on the Day of burial as on the Anniversary, the first of the All Souls' Day Masses is said, with the special prayers in Various Prayers (p. \pageref{RequiemPrayers}).
\end{rubric}
\subby{On the Day of Death or Burial of a Departed Person}

\begin{rubric}
    The Mass is said of All Souls' Day's First Mass (p. \pageref{AllSouls}), except for the following.
\end{rubric}

\subbysub{Collect}
\lett{O}{God,} whose property is ever to have mercy and to forgive, we humbly entreat thee for the soul of thy \textit{servant} \textit{N.}, which thou hast this day commanded to depart from this world: deliver it not into the hands of the enemy, neither forget it at the last, but command that it be received by thy holy Angels, and brought unto the fatherland of paradise; that forasmuch as \textit{he} hoped and believed in thee, \textit{he} may not undergo the pains of hell, but may obtain everlasting felicity. Through.
\needspace{4\baselineskip}
\begin{rubric}
    On the \nth{3}, \nth{7}, and \nth{30} Day, the following Collect is instead said,
\end{rubric}
\lett{W}{e} beseech thee, O Lord, that thou wouldest vouchsafe to grant unto the soul of thy \textit{servant} \textit{N.}, whose burial \textit{three} days since, we now commemorate, the fellowship of thy Saints and elect: and that thou wouldest pour upon it the continual dew of thy mercy. Through.

\readingcitation{Epistle}{1 Thessalonians 4:13}
\lett{B}{rethren:} I would not have you to be ignorant concerning them which are asleep, that ye sorrow not, even as others which have no hope. For if we believe that Jesus died and rose again: even so them also which sleep in Jesus will God bring with him. For this we say unto you by the word of the Lord, that we which are alive and remain unto the coming of the Lord shall not prevent them which are asleep. For the Lord himself shall descend from heaven with a shout, with the voice of the Archangel, and with the trump of God: and the dead in Christ shall rise first. Then we which are alive and remain shall be caught up together with them in the clouds, to meet the Lord in the air, and so shall we ever be with the Lord. Wherefore comfort one another with these words.

\readingcitation{Gospel}{John 11:21}
\lett{A}{t that time:} Martha said unto Jesus: Lord, if thou hadst been here, my brother had not died: but I know, that even now, whatsoever thou wilt ask of God, God will give it thee. Jesus saith unto her: Thy brother shall rise again. Martha saith unto him: know that he shall rise again in the resurrection at the last day. Jesus said unto her: I am the resurrection, and the life: he that believeth in me, though he were dead, yet shall he live: and whosoever liveth and believeth in me shall never die. Believest thou this? She saith unto him: Yea, Lord, I believe that thou art the Christ, the Son of God, which should come into the world.

\subbysub{Secret}
\lett{W}{e} beseech thee, O Lord, to have mercy upon the soul of thy \textit{servant} \textit{N.}, for which we offer unto thee the sacrifice of praise, humbly entreating thy majesty: that by these offices of loving atonement, it may be worthy to attain unto everlasting rest. Through.

\begin{rubric}
    On the \nth{3}, \nth{7}, and \nth{30} Day, the following Secret is instead said,
\end{rubric}
\lett{W}{e} beseech thee, O Lord, graciously to regard the gifts which we offer unto thee for the soul of thy \textit{servant} \textit{N.}: that, being cleansed by heavenly remedies, it may rest in thy mercy. Through.

\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, almighty God: that the soul of thy \textit{servant} \textit{N.}, which hath this day passed from this world, being cleansed by these sacrifices and delivered from sins, may obtain thy pardon and everlasting rest. Through.

\begin{rubric}
    On the \nth{3}, \nth{7}, and \nth{30} Day, the following Secret is instead said,
\end{rubric}
\lett{R}{eceive,} O Lord, our prayers for the soul of thy \textit{servant} \textit{N.}: that whatsoever stains of earthly defilement cleave to it may by thy merciful pardon be done away. Through.

\subby{On the Anniversary of the Departed}
\begin{rubric}
    All is said of All Souls' Day's Second Mass (p. \pageref{AllSoulsII}), except the following.
\end{rubric}
\subbysub{Collect}
\lett{O}{God,} the Lord of mercies: grant unto the \textit{soul} of thy \textit{servant} \textit{N.}, the anniversary of whose burial we now commemorate, a place of refreshment, the blessedness of rest, and the brightness of thy light. Through.

\subbysub{Secret}
\lett{B}{e} favourable, O Lord, to our supplications for the \textit{soul} of thy \textit{servant} \textit{N.}, whose anniversary is kept this day: for which we offer unto thee this sacrifice of praise; that thou wouldest vouchsafe to join \textit{it} to the fellowship of thy Saints. Through.

\subbysub{Postcommunion}
\lett{G}{rant,} we beseech thee, O Lord: that the \textit{soul} of thy \textit{servant} \textit{N.}, the anniversary of whose burial we commemorate; being purged by these sacrifices, may obtain both pardon and everlasting rest. Through.

\subby{In Daily Masses of the Dead}
\begin{rubric}
    All is said of All Souls' Day's Third Mass (p. \pageref{AllSoulsIII}), except the following.
\end{rubric}
%MANUAL ADJUSTMENT:
\vspace{-0.5\baselineskip}
\subsubsec{Collect}
\begin{rubric}
    If the Mass is said for the all the faithful departed, then all three Collects are said.
\end{rubric}
\begin{rubric}
    If the Mass is said for specified persons, the \nth{1} Collect is for him (p. \pageref{RequiemPrayers}), or say the second Collect following if the name is unknown), the \nth{2} Collect is \emph{ad libitum}, \& the \nth{3} Collect is the third Collect following.
\end{rubric}
\lett{O}{God,} who didst cause thy servants to enjoy the dignity of Priest or Bishop in the apostolic Priesthood: grant, we beseech thee; that they may evermore be joined unto the fellowship of the same. Through.

\lett{O}{God,} the giver of pardon and lover of man's salvation: we beseech thee of thy mercy to grant; that the brethren, kinsfolk, and benefactors of our congregation who have passed out of this world, may, at the intercession of blessed Mary ever Virgin and of all thy Saints, attain unto the fellowship of eternal blessedness. (Through.)
\needspace{4\baselineskip}
\lett{O}{God,} the Creator and Redeemer of all the faithful: grant unto the souls of thy servants and handmaids the remission of all their sins; that through devout supplications they may obtain the pardon which they have alway desired. Who livest and reignest.

\subbysub{Secret}
\lett{A}{ccept,} O Lord, we beseech thee, the sacrifices which we offer to thee for the souls of thy servants, Bishops or Priests: that forasmuch as in this life thou didst bestow on them the dignity of Bishop or Priest, so thou wouldest bid them be joined unto the fellowship of thy Saints in the heavenly kingdom. Through.
\needspace{4\baselineskip}
\lett{O}{God,} whose mercies are without number, graciously receive our humble prayers: and through these sacraments of our salvation grant unto the souls of our brethren, kinsfolk, and benefactors, to whom thou hast given the confession of thy name, the remission of all their sins. (Through.)
\needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord, look favourably upon the sacrifices which we offer unto thee for the souls of thy servants and handmaids: that as thou hast bestowed on them the merit of the Christian faith, thou wouldest grant them the reward of the same. Through.

\subbysub{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that thy merciful loving kindness, which we implore, may be profitable for the souls of thy servants, Bishops or Priests: that by thy mercy they may attain unto the eternal fellowship of him in whom they have hoped and believed. Through.
\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty and merciful God: that the souls of our brethren, kinsfolk, and benefactors, for whom we have offered unto thy majesty this sacrifice of praise; may by the power of this sacrament be cleansed from all their sins, and by thy mercy obtain the blessedness of everlasting light. (Through.)
\needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord, that the prayer of thy suppliant people may be profitable for the souls of thy servants and handmaids: that thou wouldest deliver them from all their sins, and make them to be partakers of thy redemption. Who livest.