\fancyhead[RO,LE]{\textit{Athanasian Creed}}
\fancyhead[RE,LO]{}
%The 1928 English BCP's revised translation of the Athanasian Creed is first seen in the 1918 and then 1962 Canadian BCPs, so it deserves use here.
%Relevant notes for consideration found here: https://www.ccel.org/ccel/schaff/creeds2.iv.i.iv.html
%Usher text: 
 \section{Athanasian Creed}\label{Ath}
\elcol{\lett{W}{hosoever} would be saved : needeth before all things to hold fast the Catholick Faith.\par
\secondlinepsalm{Which Faith except a man keep whole and undefiled : without doubt he will perish eternally.}
\thirdlinepsalm{Now the Catholick Faith is this : that we worship one God in Trinity, and the Trinity in Unity;}
    \psanum{4}Neither confusing the Persons : nor dividing the substance.\par
    \psanum{5}For there is one Person of the Father, another of the Son : another of the Holy Ghost;\par
    \psanum{6}But the Godhead of the Father, and of the Son, and of the Holy Ghost is all one : the glory equal, the majesty co-eternal.\par
    \psanum{7}Such as the Father is, such is the Son : and such is the Holy Ghost.\par
    \psanum{8}The Father uncreated, the Son uncreated : the Holy Ghost uncreated;\par
    \psanum{9}The Father infinite, the Son infinite : the Holy Ghost infinite;\par
    \psanum{10}The Father eternal, the Son eternal : the Holy Ghost eternal.\par
    \psanum{11}And yet there are not three eternals : but one eternal\par
    \psanum{12}As also there are not three uncreated, nor three infinites : but one infinite, and one uncreated.\par
    \psanum{13}So likewise the Father is almighty, the Son almighty : the Holy Ghost almighty;\par
    \psanum{14}And yet there are not three almighties : but one almighty.\par
    \psanum{15}So the Father is God, the Son God : the Holy Ghost God;\par
    \psanum{16}And yet there are not three Gods : but one God.\par
    \psanum{17}So the Father is Lord, the Son Lord : the Holy Ghost Lord;\par
    \psanum{18}And yet there are not three Lords : but one Lord.\par
    \psanum{19}For like as we are compelled by the Christian verity : to confess each Person by himself to be both God and Lord;\par
    \psanum{20}So are we forbidden by the Catholick religion : to speak of three Gods or three Lords.\par
    \psanum{21}The Father is made of none : nor created, nor begotten.\par
    \psanum{22}The Son is of the Father : not made, nor created, but begotten.\par
    \psanum{23}The Holy Ghost is of the Father : not made, nor created, nor begotten, but proceeding.\par
    \psanum{24}There is therefore one Father, not three Fathers; one Son, not three Sons : one Holy Ghost, not three Holy Ghosts.\par
    \psanum{25}And in this Trinity there is no before or after : no greater or less;\par
    \psanum{26}But all three Persons are co-eternal together : and co-equal.\par
    \psanum{27}So that in all ways, as is aforesaid : both the Trinity is to be worshipped in Unity, and the Unity in Trinity.\par
    \psanum{28}He therefore that would be saved : let him thus think of the Trinity.\par

    \lett{F}{urthermore} it is necessary to eternal salvation : that he also believe faithfully the Incarnation of our Lord Jesus Christ.\par
    \secondlinepsalmonenineteen{Now the right faith is that we believe and confess : that our Lord Jesus Christ, the Son of God, is both God and man.}{30}
    \thirdlinepsalmonenineteen{He is God, of the substance of the Father, begotten before the worlds : and he is man, of the substance of his Mother, born in the world;}{31}
    \psanum{32}Perfect God : perfect man, of reasoning soul and human flesh subsisting;\par
    \psanum{33}Equal to the Father as touching his Godhead : less than the Father as touching his manhood.\par
    \psanum{34}Who although he be God and man : yet he is not two, but is one Christ ;\par
    \psanum{35}One however, not by conversion of Godhead into flesh : but by taking manhood into God;\par
    \psanum{36}One altogether : not by confusion of substance, but by unity of person.\par
    \psanum{37}For as reasoning soul and flesh is one man : so God and man is one Christ;\par
    \psanum{38}Who suffered for our salvation : descended into hell, rose again from the dead;\par
    \psanum{39}Ascended into heaven, sat down at the right hand of the Father : from whence he shall come to judge the quick and the dead.\par
    \psanum{40}At whose coming all men must rise again with their bodies : and shall give account for their own deeds.\par
    \psanum{41}And they that have done good will go into life eternal : they that have done evil into eternal fire.\par

    \lett{T}{his} is the Catholick Faith : which except a man do faithfully and steadfastly believe, he cannot be saved.\par
    \secondline{℣. Glory be to the Father, and to the Son : and to the Holy Ghost;}
    \thirdline{℟. As it was in the beginning, is now, and ever shall be : world without end. Amen.}}
{\lett{Q}{uic\smash{ú}mque} vult salvus esse, * ante ómnia opus est, ut téneat cathólicam fidem:\par
\secondline{Quam nisi quisque íntegram inviolatámque serváverit, * absque dúbio in {\ae}térnum períbit.}
\thirdline{Fides autem cathólica h{\ae}c est: * ut unum Deum in Trinitáte, et Trinitátem in unitáte venerémur.}
Neque confundéntes persónas, * neque substántiam separántes.\par
Alia est enim persóna Patris, ália Fílii, * ália Spíritus Sancti:\par
Sed Patris, et Fílii, et Spíritus Sancti una est divínitas, * {\ae}quális glória, co{\ae}térna majéstas.\par
Qualis Pater, talis Fílius, * talis Spíritus Sanctus.\par
Increátus Pater, increátus Fílius, * increátus Spíritus Sanctus.\par
Imménsus Pater, imménsus Fílius, * imménsus Spíritus Sanctus.\par
{\AE}térnus Pater, {\ae}térnus Fílius, * {\ae}térnus Spíritus Sanctus.\par
Et tamen non tres {\ae}térni, * sed unus {\ae}térnus.\par
Sicut non tres increáti, nec tres imménsi, * sed unus increátus, et unus imménsus.\par
Simíliter omnípotens Pater, omnípotens Fílius, * omnípotens Spíritus Sanctus.\par
Et tamen non tres omnipoténtes, * sed unus omnípotens.\par
Ita Deus Pater, Deus Fílius, * Deus Spíritus Sanctus.\par
Ut tamen non tres Dii, * sed unus est Deus.\par
Ita Dóminus Pater, Dóminus Fílius, * Dóminus Spíritus Sanctus.\par
Et tamen non tres Dómini, * sed unus est Dóminus.\par
Quia, sicut singillátim unamquámque persónam Deum ac Dóminum confitéri christiána veritáte compéllimur: * ita tres Deos aut Dóminos dícere cathólica religióne prohibémur.\par
Pater a nullo est factus: * nec creátus, nec génitus.\par
Fílius a Patre est: * non factus, nec creátus, sed génitus.\par
Spíritus Sanctus a Patre: * non factus, nec creátus, nec génitus, sed procédens.\par
%This was a very difficult decision to make. The Orthodox Church, from its very beginning at Pentecost, has never known a dual origin of the Spirit. In the West, the Filioque (or the `et Filio' in this place) did not express a dual hypostatic origin but the consubstantiality of the Spirit and the Son. Unfortunately, this phrasing is distorted and perverted in the West, leading to the unfortunate falling away which is being healed in the AWRV. Just like μια φθσιν had an acceptable sense under St. Cyril but was perverted under Severus, likewise the procession from the Son had an acceptable sense under St. Augustine but was perverted by the Franks (amongst others) and eventually the Popes of Rome. In light of this, and receiving the Faith of the Eighth Ecumenical Council, we follow Usher's Latin text which omits both the Son's `solo' and the Spirit's `et Filio'. This is not as much a concern about textual criticism as much as a recognition that the ancient Creed must change in order to avoid scandal.
Unus ergo Pater, non tres Patres: unus Fílius, non tres Fílii: * unus Spíritus Sanctus, non tres Spíritus Sancti.\par
Et in hac Trinitáte nihil prius aut postérius, nihil majus aut minus: * sed tot{\ae} tres persón{\ae} co{\ae}térn{\ae} sibi sunt et co{\ae}quáles.\par
Ita ut per ómnia, sicut jam supra dictum est, * et únitas in Trinitáte, et Trínitas in unitáte veneránda sit.\par
Qui vult ergo salvus esse, * ita de Trinitáte séntiat.
\par
\lett{S}{ed} necessárium est ad {\ae}térnam salútem, * ut Incarnatiónem quoque Dómini nostri Jesu Christi fidéliter credat.\par
Est ergo fides recta ut credámus et confiteámur, * quia Dóminus noster Jesus Christus, Dei Fílius, Deus et homo est.\par
Deus est ex substántia Patris ante sǽcula génitus: * et homo est ex substántia matris in sǽculo natus.\par
Perféctus Deus, perféctus homo: * ex ánima rationáli et humána carne subsístens.\par
{\AE}quális Patri secúndum divinitátem: * minor Patre secúndum humanitátem.\par
Qui licet Deus sit et homo, * non duo tamen, sed unus est Christus.\par
Unus autem non conversióne divinitátis in carnem, * sed assumptióne humanitátis in Deum.\par
Unus omníno, non confusióne substánti{\ae}, * sed unitáte persón{\ae}.\par
Nam sicut ánima rationális et caro unus est homo: * ita Deus et homo unus est Christus.\par
Qui passus est pro salúte nostra: descéndit ad ínferos: * tértia die resurréxit a mórtuis.\par
Ascéndit ad c{\ae}los, sedet ad déxteram Dei Patris omnipoténtis: * inde ventúrus est judicáre vivos et mórtuos.\par
Ad cujus advéntum omnes hómines resúrgere habent cum corpóribus suis; * et redditúri sunt de factis própriis ratiónem.\par
Et qui bona egérunt, ibunt in vitam {\ae}térnam: * qui vero mala, in ignem {\ae}térnum.
\par
\lett{H}{{\ae}c} est fides cathólica, * quam nisi quisque fidéliter firmitérque credíderit, salvus esse non póterit.\par
\secondline{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\thirdline{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}}
