\section{Commemoration of the Faithful Departed}
\fancyhead[RO,LE]{\textit{Commemoration}}
\fancyhead[RE,LO]{}
\begin{rubric}
    This Commemoration is a laudable custom, which may be said after any Hour.
\end{rubric}
\begin{multicols}{2}
℣. Let us commemorate our departed kinsfolk, neighbours, friends, and benefactors.

℟. May they rest in peace.

\subbypsalm{Psalm 130. \textit{De profundis}}

\lett{O}{ut} of the deep have I called unto thee, O Lord : Lord, hear my voice.\par
\secondlinepsalm{O let thine ears consider well : the voice of my complaint.}
\thirdlinepsalm{If thou, Lord, wilt be extreme to mark what is done amiss : O Lord, who may abide it?}
\psanum{4}For there is mercy with thee : therefore shalt thou be feared.\par
\psanum{5}I look for the Lord; my soul doth wait for him : in his word is my trust.\par
\psanum{6}My soul fleeth unto the Lord : before the morning watch, I say, before the morning watch.\par
\psanum{7}O Israel, trust in the Lord, for with the Lord there is mercy : and with him is plenteous redemption.\par
\psanum{8}And he shall redeem Israel : from all his sins.\par
℣. Rest eternal * grant unto them, O Lord.\par
℟. And let light perpetual * shine upon them.\\

℣. From the gate of hell.\par
℟. Deliver their souls, O Lord.

℣. May they rest in peace.\par
℟. Amen.

℣. O Lord hear my prayer.\par
℟. And let my cry come unto thee.\par
℣. The Lord be with you.\par
℟. And with thy spirit.\par
\letuspray
\lett{O}{God,} the Giver of pardon and the Author of man's salvation: we humbly beseech thy mercy to grant that our kinsfolk, neighbours, friends, and benefactors who have departed out of this world, blessed Mary ever Virgin and all thy Saints praying for them, may attain to the fellowship of everlasting blessedness. Through.

℣. Rest eternal grant unto them, O Lord.\par
℟. And let light perpetual shine upon them.\par
℣. May they rest in peace.\par
℟. Amen.
\end{multicols}