\fancyhead[RO,LE]{\textit{Compline}}
\fancyhead[RE,LO]{}
\section{Hour of Compline}
\vspace{-4ex}
\begin{rubric}
This Service may be used when Evensong has been previously said.
\end{rubric}
\begin{rubric}
    All standing up, the Minister shall say,
\end{rubric}
\elcol{\lett{T}{he} Lord Almighty grant us a quiet night and a perfect end.}{\lett{N}{octem} quiétam et finem perféctum concédat nobis Dóminus omnípotens.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{\lett{B}{rethren,} be sober, be vigilant; because your adversary the devil, as a roaring lion, walketh about, seeking whom he may devour : Whom resist, steadfast in the faith.%\vr{1 Peter 5:8-9.}
}{\lett{F}{ratres:} Sóbrii estóte, et vigiláte: quia adversárius vester diábolus tamquam leo rúgiens círcuit, qu{\ae}rens quem dévoret: cui resístite fortes in fide.}
\elcol{℣. But thou, O Lord, have mercy upon us;}{℣. Tu autem, Dómine, miserére nobis.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subsec{Introduction}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}
\subby{Psalm 4. \textit{Cum invocarem}}
\lett{H}{ear} me when I call, O God of my righteousness : thou hast set me at liberty when I was in trouble; have mercy upon me, and hearken unto my prayer.\par
\secondlinepsalm{O ye sons of men, how long will ye blaspheme mine honour : and have such pleasure in vanity, and seek after leasing?}
\thirdlinepsalm{Know this also, that the Lord hath chosen to himself the man that is godly : when I call upon the Lord, he will hear me.}
\psanum{4}Stand in awe, and sin not : commune with your own heart, and in your chamber, and be still.\par
\psanum{5}Offer the sacrifice of righteousness : and put your trust in the Lord.\par
\psanum{6}There be many that say : Who will shew us any good?\par
\psanum{7}Lord, lift thou up : the light of thy countenance upon us.\par
\psanum{8}Thou hast put gladness in my heart : since the time that their corn and wine and oil increased.\par
\psanum{9}I will lay me down in peace, and take my rest : for it is thou, Lord, only, that makest me dwell in safety.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
\par
\needspace{4\baselineskip}
\subsec{Psalm 31. \textit{In te, Domine, speravi}}
\lett{I}{n} thee, O Lord, have I put my trust : let me never be put to confusion, deliver me in thy righteousness.\par
\secondlinepsalm{Bow down thine ear to me : make haste to deliver me.}
\thirdlinepsalm{And be thou my strong rock, and house of defence : that thou mayest save me.}
\psanum{4}For thou art my strong rock, and my castle : be thou also my guide, and lead me for thy Name's sake.\par
\psanum{5}Draw me out of the net that they have laid privily for me : for thou art my strength.\par
\psanum{6}Into thy hands I commend my spirit : for thou hast redeemed me, O Lord, thou God of truth.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
\par
\needspace{4\baselineskip}
\subsec{Psalm 91. \textit{Qui habitat}}
\lett{W}{hoso} dwelleth under the defence of the most High : shall abide under the shadow of the Almighty.\par
\secondlinepsalm{I will say unto the Lord, Thou art my hope, and my strong hold : my God, in him will I trust.}
\thirdlinepsalm{For he shall deliver thee from the snare of the hunter : and from the noisome pestilence.}
\psanum{4}He shall defend thee under his wings, and thou shalt be safe under his feathers : his faithfulness and truth shall be thy shield and buckler.\par
\psanum{5}Thou shalt not be afraid for any terror by night : nor for the arrow that flieth by day;\par
\psanum{6}For the pestilence that walketh in darkness : nor for the sickness that destroyeth in the noon-day.\par
\psanum{7}A thousand shall fall beside thee, and ten thousand at thy right hand : but it shall not come nigh thee.\par
\psanum{8}Yea, with thine eyes shalt thou behold : and see the reward of the ungodly.\par
\psanum{9}For thou, Lord, art my hope : thou hast set thine house of defence very high.\par
\psanum{10}There shall no evil happen unto thee : neither shall any plague come nigh thy dwelling.\par
\psanum{11}For he shall give his angels charge over thee : to keep thee in all thy ways.\par
\psanum{12}They shall bear thee in their hands : that thou hurt not thy foot against a stone.\par
\psanum{13}Thou shalt go upon the lion and adder : the young lion and the dragon shalt thou tread under thy feet.\par
\psanum{14}Because he hath set his love upon me, therefore will I deliver him : I will set him up, because he hath known my Name.\par
\psanum{15}He shall call upon me, and I will hear him : yea, I am with him in trouble; I will deliver him, and bring him to honour.\par
\psanum{16}With long life will I satisfy him : and shew him my salvation.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
\par
\needspace{4\baselineskip}
\subsec{Psalm 134. \textit{Ecce nunc}}
\lett{B}{ehold} now, praise the Lord : all ye servants of the Lord;\par
\secondlinepsalm{Ye that by night stand in the house of the Lord : even in the courts of the house of our God.}
\thirdlinepsalm{Lift up your hands in the sanctuary : and praise the Lord.}
\psanum{4}The Lord that made heaven and earth : give thee blessing out of Sion.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
\par
\vspace{-1\baselineskip}
%\begin{rubric}
%	Then shall be read one of the following short Lessons.
%\end{rubric}
%I see no precedent or reason to expand the options.
\subby{Chapter}
\elcol{\lett{T}{hou,} O Lord, art in the midst of us, and we are called by thy name. Leave us not, O Lord our God.}{\lett{T}{u} autem in nobis es, Dómine, et nomen sanctum tuum invocátum est super nos: ne derelínquas nos, Dómine, Deus noster.}
%\begin{rubric}
%	or
%\end{rubric}
%\lett{C}{ome} unto me, all ye that labour and are heavy laden, and I will give you rest. Take my yoke upon you, and learn of me; for I am meek and lowly in heart: and ye shall find rest unto your souls. For my yoke is easy, and my burden is light.
%\begin{rubric}
%	or
%\end{rubric}
%\lett{N}{ow} the God of peace, that brought again from the dead our Lord Jesus, that great Shepherd of the sheep, through the blood of the Everlasting Covenant, make you perfect in every good work to do his will, working in you that which is well-pleasing in his sight; through Jesus Christ, to whom be glory for ever and ever. Amen.
%\par
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subby{Short Respond}
\elcol{℣. Into thy hands, O Lord, I commend my spirit;}{℣. In manus tuas, Dómine, * Comméndo spíritum meum.}

\elcol{℟. Into thy hands, O Lord, I commend my spirit;}{℟. In manus tuas, Dómine, * Comméndo spíritum meum.}

\elcol{℣. For thou hast redeemed me, O Lord, thou God of truth.}{℣. Redemísti nos, Dómine, Deus veritátis.}

\elcol{℟. I commend my spirit.}{℟. Comméndo spíritum meum.}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost}{℣. Glória Patri, et Fílio, * et Spirítui Sancto.}

\elcol{℟. Into thy hands, O Lord, I commend my spirit.}{℟. In manus tuas, Dómine, * Comméndo spíritum meum.}

\subby{Te lucis ante terminum.}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{B}{efore} the ending of the day,\par
\secondline{Creator of the world we pray,}
\thirdline{That with thy wonted favour thou}\par\noindent
Wouldst be our guard and keeper now.
\par\noindent
From all ill dreams defend our eyes,\par\noindent
From nightly fears and fantasies\par\noindent
Tread under foot our ghostly foe,\par\noindent
That no pollution we may know.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Doth live and reign eternally. Amen.\\
\end{hangparas}
℣. Keep me as the apple of an eye;

℟. Hide me under the shadow of thy wings.}{\lett{T}{e} lucis ante términum,\par
\secondline{Rerum Creátor, póscimus,}
\thirdline{Ut pro tua cleméntia}
Sis pr{\ae}sul et custódia.\\
\par
Procul recédant sómnia,\par
Et nóctium phantásmata;\par
Hostémque nóstrum cómprime,\par
Ne polluántur córpora.\\
\par
Jesu, tibi sit glória,\par
Qui apparuísti Géntibus,\par
Cum Patre et almo Spíritu,\par
In sempitérna sǽcula. Amen.\\

℣. Custódi nos, Dómine, ut pupíllam óculi.

℟. Sub umbra alárum tuárum prótege nos.}

\subby{Nunc Dimittis}
\elcol{\antiphon{}{Preserve us, O Lord, waking, and guard us sleeping, that we may watch with Christ, and rest in peace.}}{\antiphon{}{Salva nos, Dómine, vigilántes, custódi nos dormiéntes; ut vigilémus cum Christo, et requiescámus in pace.}}

\needspace{4\baselineskip}
\elcol{\lett{L}{ord,} {\ding{64}} now lettest thou thy servant depart in peace, * according to thy word.\par
    \secondline{For mine eyes have seen * thy salvation,}
    \thirdline{Which thou hast prepared * before the face of all people;}
    To be a light to lighten the Gentiles, * and to be the glory of thy people Israel.\par
	℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
	℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.
}{\lett{N}{unc} dimíttis {\ding{64}} servum tuum, Dómine, * secúndum verbum tuum in pace:\par
\secondline{Quia vidérunt óculi mei * salutáre tuum,}
\thirdline{Quod parásti * ante fáciem ómnium populórum,}
Lumen ad revelatiónem géntium, * et glóriam plebis tu{\ae} Isra{\"e}l.\par
℣. Glória Patri, et Fílio, * et Spirítui Sancto:\par
℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}

\elcol{\antiphon{}{Preserve us, O Lord, waking, and guard us sleeping, that we may watch with Christ, and rest in peace.}}{\antiphon{}{Salva nos, Dómine, vigilántes, custódi nos dormiéntes; ut vigilémus cum Christo, et requiescámus in pace.}}
%Apostle's Creed is move into Suffrages in accordance with Sarum.
\subby{Suffrages}
\begin{rubric}
	The Suffrages, said kneeling, are said on all Ferias outside of Eastertide. They are omitted on Sundays, Feasts, and within Principle, Primary, and Secondary Octaves.\par
	\textsc{Note,} If omitted, the Office continues with the Collect.
\end{rubric}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\subbysub{Lord's Prayer}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\subbysub{Apostles' Creed}
\elcol{\lett{I}{believe} in God the Father Almighty, Maker of heaven and earth:
    And in Jesus Christ his only Son our Lord: Who was conceived by the Holy Ghost, Born of the Virgin Mary: Suffered under Pontius Pilate, Was crucified, dead, and buried: He descended into hell; The third day he rose again from the dead: He ascended into heaven, And sitteth on the right hand of God the Father Almighty: From thence he shall come to judge the quick and the dead.\par
    I believe in the Holy Ghost: The holy Catholic Church; The Communion of Saints: The Forgiveness of sins: The Resurrection of the body: {\ding{64}} And the Life everlasting. Amen.
}{\lett{C}{redo} in Deum, Patrem omnipoténtem, Creatórem c{\ae}li et terr{\ae}.
     Et in Jesum Christum, Fílium ejus únicum, Dóminum nostrum: qui concéptus est de Spíritu Sancto, natus ex María Vírgine, passus sub Póntio Piláto, crucifíxus, mórtuus, et sepúltus: descéndit ad ínferos; tértia die resurréxit a mórtuis; ascéndit ad c{\ae}los; sedet ad déxteram Dei Patris omnipoténtis: inde ventúrus est judicáre vivos et mórtuos.\par
     Credo in Spíritum Sanctum, sanctam Ecclésiam cathólicam, Sanctórum communiónem, remissiónem peccatórum, carnis {\ding{64}} resurrectiónem, vitam {\ae}térnam. Amen.}
%\subby{Preces}
   \begin{rubric}
	Then shall follow these Versicles and Responses.
\end{rubric}
%From the Clementine Vulgate
\elcol{℣. Blessed art thou, Lord God of our fathers;}{℣. Benedíctus es, Dómine Deus patrum nostrórum,}

\elcol{℟. To be praised and glorified above all for ever.}{℟. Et laudábile, et gloriósum nomen tuum in s{\ae}cula.}

\elcol{℣. Let us bless the Father, the Son, and the Holy Ghost;}{℣. Benedicámus Patrem et Fílium cum Sancto Spíritu:}

\elcol{℟. Let us praise him and magnify him for ever.}{℟. Laudémus et superexaltémus eum in sǽcula.}

\elcol{℣. Blessed art thou, O Lord, in the firmament of heaven;}{℣. Benedíctus es, Dómine, in firmaménto c{\ae}li:}

\elcol{℟. To be praised and glorified above all for ever.}{℟. Et laudábilis, et gloriósus, et superexaltátus in sǽcula.}

\elcol{℣. The Almighty and most merciful Lord guard us and give us his blessing.\par
℟. Amen.}{℣. Benedícat et custódiat nos omnípotens et miséricors Dóminus.\par
℟. Amen.}


\subbysub{Confession}
\begin{rubric}
	Then shall the Minister and People say together the Confession following,
\end{rubric}
\needspace{4\baselineskip}
%Replaced with Sarum. Text from the Monastic Diurnal.
\elcol{\lett{I}{confess} to God, to blessed Mary, to all the saints, and to you: that I have sinned exceedingly in thought, word, and deed, by my own fault. I beg holy Mary, all the Saints of God, and you, to pray for me.}{\lett{C}{onf\smash{í}teor} Deo: beát{\ae} Marí{\ae}, ómnibus sanctis, et vobis: peccávi nimis: cogitatióne, locutióne, et ópere, mea culpa. Precor sanctam Maríam, omnes sanctos Dei, et vos oráre pro me.}

\begin{rubric}
	The Minister then stands and says,
\end{rubric}
%1928 English Translation
\needspace{4\baselineskip}
\elcol{\lett{A}{lmighty} God, have mercy upon us, forgive us all our sins and deliver us from all evil, confirm and strengthen us in all goodness, and bring us to life everlasting.\par
℟. Amen.}{\lett{M}{isere\smash{á}tur} nostri omnípotens Deus et dimíttat nobis ómnia peccáta nostra: líberet nos ab omni malo : consérvet et confírmet in bono et ad vitam perdúcat {\ae}térnam.\par
℟. Amen.}
\begin{rubric}
	If a Priest be present, he shall stand and pronounce the following Absolution,
\end{rubric}
%1928 English Translation
\needspace{4\baselineskip}
\elcol{\lett{M}{ay} the almighty and merciful Lord grant unto you pardon {\ding{64}} and remission of all your sins, time for amendment of life, and the grace and comfort of the Holy Spirit.\par
℟. Amen.}{\lett{A}{bsoluti\smash{ó}nem} {\ding{64}} et remissiónem ómnium peccatórum vestrórum, spácium vere peniténti{\ae}, emendatiónem vit{\ae}, grátiam et consolatiónem Sancti Spíritus: tríbuat vobis omnípotens et miséricors Dóminus.\par
℟. Amen.}

%\subby{Preces}
\elcol{℣. Wilt thou not turn again and quicken us, O God?}{℣. Deus, tu convérsus vivificábis nos?}
\elcol{℟. That thy people may rejoice in thee.}{℟. Et plebs tua l{\ae}tábitur in te.}
\elcol{℣. O Lord, show thy mercy upon us.}{℣. Osténde nobis, Dómine, misericórdiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutáre tuum da nobis.}
\elcol{℣. Vouchsafe, O Lord, to keep us this day without sin;}{℣. Dignáre, Dómine, die isto sine peccáto nos custodíre. }
\elcol{℟. O Lord, have mercy upon us, have mercy upon us.}{℟. Miserére nostri, Dómine, miserére nostri.}

\subby{Collect}
\elcol{℣. O Lord, hear our prayer.}{℣. Dómine, exáudi oratiónem nostram.}
\elcol{℟. And let our cry come unto thee.}{℟. Et clamor noster ad te véniat.}

\oremus
\elcol{\lett{V}{isit,} we beseech thee, O Lord, this place, and drive from it all the snares of the enemy; let thy holy angels dwell herein to preserve us in peace; and may thy blessing be upon us evermore; through Jesus Christ our Lord. \textit{Amen.}}{\lett{V}{\smash{í}sita,} quǽsumus, Dómine, habitatiónem istam, et omnes insídias inimíci ab ea lónge repélle: Ángeli tui sancti hábitent in ea, qui nos in pace custódiant; et benedíctio tua sit super nos semper. Per Dóminum nostrum Jesum Christum. \textit{Amen.}}

%\lett{L}{ighten} our darkness, we beseech thee, O Lord; and by thy great mercy defend us from all perils and dangers of this night; for the love of thy only Son, our Saviour, Jesus Christ. Amen.
\begin{inhead}
	or,
\end{inhead}
\needspace{4\baselineskip}
\lett{L}{ord} Jesus Christ, Son of the living God, who at this evening hour didst rest in the sepulchre, and didst thereby sanctify the grave to be a bed of hope to thy people; Make us so to abound in sorrow for our sins, which were the cause of thy passion, that when our bodies lie in the dust, our souls may live with thee; who livest and reignest with the Father and the Holy Ghost, one God world without end. Amen.

%\lett{L}{ook} down, O Lord, from thy heavenly throne, illuminate the darkness of this night with thy celestial brightness, and from the sons of light banish the deeds of darkness; through Jesus Christ our Lord. Amen.

%\lett{B}{e} present, O merciful God, and protect us through the silent hours of this night, so that we who are wearied by the changes and chances of this fleeting world, may repose upon thy eternal changelessness; through Jesus Christ our Lord. Amen. 
\subby{Conclusion}
%Taken from Psalm 4:8. Latin supplied from Vulgate Psalter juxta Hebrorum.
\elcol{℣. We will lay us down in peace and take our rest.\par
℟. For it is thou, Lord, only that makest us dwell in safety.\par
℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.\par
℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.\par
℣. Let us bless the Lord. (alleluia, alleluia.)\par
℟. Thanks be to God. (alleluia, alleluia.)}{℣. In pace simul requiéscam et dormiam.\par
℟. Quia tu Dómine speciáliter secúrum habitáre fecísti me.\par
℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.\par
℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.\par
℣. Benedicámus Dómino (allelúja, allelúja).\par
℟. Deo grátias (allelúja, allelúja).}
%\elcol{℣. We will lay us down in peace and take our rest;}{℣. In pace simul requiéscam et dormiam.}
%\elcol{℟. For it is thou, Lord, only that makest us dwell in safety.}{℟. Quia tu Dómine speciáliter secúrum habitáre fecísti me.}
%\elcol{℣. The Lord be with you.\par
%\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
%\textit{vel,} Dómine, exáudi oratiónem nostram.}
%\elcol{℟. And with thy spirit.\par
%\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
%\textit{vel,} Et clamor noster ad te véniat.}
%\elcol{℣. Let us bless the Lord. (alleluia, alleluia.)}{℣. Benedicámus Dómino (allelúja, allelúja).}
%\elcol{℟. Thanks be to God. (alleluia, alleluia.)}{℟. Deo grátias (allelúja, allelúja).}

\needspace{4\baselineskip}
\elcol{\lett{T}{he} Almighty and merciful Lord, the Father, {\ding{64}} the Son, and the Holy Ghost; bless and preserve us. \textit{Amen.}}{\lett{B}{ened\smash{í}cat} et custódiat nos omnípotens et miséricors Dóminus, {\ding{64}} Pater, et Fílius, et Spíritus Sanctus. \textit{Amen.}}
\begin{rubric}
	{The appropriate Marian Anthem (p. \pageref{mary}) shall be said or sung here.}
 \end{rubric}
\begin{center}
\textsc{Here Endeth the Order of Compline.}
\end{center}