%Translation from the Monastic Diurnal
\section{Marian Anthems}\label{mary}
\fancyhead[RO,LE]{\textit{Marian Anthems}}
\fancyhead[RE,LO]{Alma Redemptoris Mater}
\subby{Alma Redemptoris Mater}
\vspace{-1ex}
\begin{inhead}
From I Evensong of Advent I until II Evensong of the Purification, inclusive.
\end{inhead}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{G}{racious} Mother of our Redeemer, for ever abiding\par
\secondline{Heaven's gateway, and star of ocean,}
%\thirdline{O succour the people, who, though falling, strive to rise again.}
\par\noindent
O succour the people, who, though falling, strive to rise again.\par\noindent
Thou Maiden who barest thy holy Creator, to the wonder of all nature;\par\noindent
Ever Virgin, after, as before thou receivedst that Ave\par\noindent
From the mouth of Gabriel; have compassion on us sinners.
\end{hangparas}
\begin{inhead}
	In Advent 
\end{inhead}
℣. The Angel of the Lord announced unto Mary.\par
℟. And she conceived by the Holy Ghost.\par
\letuspray
\lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts: that, as we have known the incarnation of thy Son Jesus Christ by the message of an angel, so by his Cross and Passion we may be brought unto the glory of his Resurrection. Through the same Christ our Lord. \textit{Amen.}\par
\begin{inhead}
	In Christmastide
\end{inhead}
℣. After Childbearing, O Virgin, thou didst remain inviolate.\par
℟. Intercede for us, O Mother of God.\par
\letuspray
\lett{O}{God,} who by the fruitful virginity of Blessed Mary hast bestowed upon mankind the reward of eternal salvation: Grant, we beseech thee, that we may know the help of her intercession through whom we have been accounted worthy to receive the Author of our life, Jesus Christ thy Son our Lord. \textit{Amen.}}{\lett{A}{lma} Redemptóris Mater, qu{\ae} pérvia c{\ae}li porta manes,\par
\secondline{Et stella maris, succúrre cadénti,}
\thirdline{Súrgere qui curat, pópulo: tu qu{\ae} genuísti,}
Natúra miránte, tuum sanctum Genitórem,\par
Virgo prius ac postérius, Gabriélis ab ore\par
Sumens illud Ave, peccatórum miserére.\par
\begin{inhead}
	In Adventu
\end{inhead}
℣. Ángelus Dómini nuntiávit Marí{\ae}.\par
℟. Et concépit de Spíritu Sancto.\par
\oremuslatin
\lett{G}{r\smash{á}tiam} tuam, quǽsumus, Dómine, méntibus nostris infúnde: ut, qui, Ángelo nuntiánte, Christi Fílii tui incarnatiónem cognóvimus; per passiónem ejus et crucem, ad resurrectiónis glóriam perducámur. Per eúmdem Christum Dóminum nóstrum. \textit{Amen.}\par
\begin{inhead}
	De Nativitate
\end{inhead}
℣. Post partum, Virgo, invioláta permansísti.\par
℟. Dei Génitrix, intercéde pro nobis.\par
\oremuslatin
\lett{D}{eus,} qui salútis {\ae}térn{\ae}, beát{\ae} Marí{\ae} virginitáte fecúnda, humáno géneri prǽmia pr{\ae}stitísti: tríbue, quǽsumus; ut ipsam pro nobis intercédere sentiámus, per quam merúimus auctórem vit{\ae} suscípere, Dóminum nóstrum Jesum Christum Fílium tuum. \textit{Amen.}}
\vspace{-2ex}
\subby{Ave, Regina C{\ae}lorum}
\fancyhead[RE,LO]{Ave, Regina C{\ae}lorum}
\vspace{-1ex}
\begin{inhead}
From Compline of the Purification until Compline of Holy Wednesday, inclusive.
\end{inhead}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{Q}{ueen} of the heavens, we hail thee,\par
\secondline{Hail thee, Lady of all the Angels;}
%\thirdline{Thou the dawn, the door of morning}
\par\noindent
Thou the dawn, the door of morning\par\noindent
Whence the world's true light is risen:\\
\par\noindent
Joy to thee, O Virgin glorious,\par\noindent
Beautiful beyond all other;\par\noindent
Hail and farewell, O most gracious,\par\noindent
Intercede for us alway to Jesus.\\
\end{hangparas}
\par 
℣. Vouchsafe that I may praise thee, O holy Virgin.\par
℟. Give me strength against thine enemies.\par
\centerline{\textit{Let us pray.}}

\lett{G}{rant} us, O merciful God, protection in our weakness: that we who celebrate the memory of the holy Mother of God may, through the aid of her intercession, rise again from our sins. Through the same Christ our Lord. \textit{Amen.}}
{\lett{A}{ve}, Regína c{\ae}lórum,\par
\secondline{Ave, Dómina Angelórum:}
\thirdline{Salve radix, salve porta,}
Ex qua mundo lux est orta:\\

Gaude, Virgo gloriósa,\par
Super omnes speciósa,\par
Vale, o valde decóra,\par
Et pro nobis Christum exóra.\\
\par
℣. Dignáre me laudáre te, Virgo sacráta.\par
℟. Da mihi virtútem contra hostes tuos.\par
\oremuslatin
\lett{C}{onc\smash{é}de,} miséricors Deus, fragilitáti nostr{\ae} pr{\ae}sídium; ut, qui sanct{\ae} Dei Genetrícis memóriam ágimus; intercessiónis ejus auxílio, a nostris iniquitátibus resurgámus. Per eúmdem Christum Dóminum nóstrum. \textit{Amen.}}

\subby{Regina C{\ae}li}
\fancyhead[RE,LO]{Regina C{\ae}li}
\vspace{-1ex}
\begin{inhead}
From Compline of Holy Saturday until I Evensong of Trinity Sunday, exclusive.
\end{inhead}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{O}{Queen} of heaven, be joyful, alleluia;\par
\secondline{Because he whom so meetly thou barest, alleluia,}
%\thirdline{Hath arisen, as he promised, alleluia:}
\par\noindent
Hath arisen, as he promised, alleluia:\par\noindent
Pray for us to the Father, alleluia.\\
\end{hangparas}
\par
℣. Rejoice and be glad, O Virgin Mary, alleluia.\par
℟. For the Lord is risen indeed, alleluia.\par
\letuspray
\lett{O}{God,} who, by the resurrection of thy Son Jesus Christ didst vouchsafe to give gladness unto the world: Grant, we beseech thee, that we, being holpen by the Virgin Mary, his Mother, may attain unto the joys of everlasting life. Through the same Christ our Lord. \textit{Amen.}}
{\lett{R}{eg\smash{í}na} c{\ae}li, l{\ae}táre, allelúja;\par
\secondline{Quia quem meruísti portáre, allelúja,}
\thirdline{Resurréxit, sicut dixit, allelúja:}
Ora pro nobis Deum, allelúja.\\
\par
℣. Gaude et l{\ae}táre, Virgo María, allelúja.\par
℟. Quia surréxit Dóminus vere, allelúja.\par
\oremuslatin
\lett{D}{eus,} qui per resurrectiónem Fílii tui, Dómini nostri Jesu Christi, mundum l{\ae}tificáre dignátus es: pr{\ae}sta, quǽsumus; ut, per eíus Genetrícem Vírginem Maríam, perpétu{\ae} capiámus gáudia vit{\ae}. Per eúmdem Christum Dóminum nóstrum. \textit{Amen.}}
\vspace{-1ex}
\subby{Salve Regina}
\fancyhead[RE,LO]{Salve Regina}
\vspace{-1ex}
\begin{inhead}
From I Evensong of Trinity Sunday until I Evensong of Advent I, exclusive.
\end{inhead}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{M}{ary,} we hail thee, Mother and Queen compassionate;\par
\secondline{Mary, our comfort, life and hope, we hail thee.}
%\thirdline{To thee we exiles, children of Eve, lift our crying.}
\par\noindent
To thee we exiles, children of Eve, lift our crying.\par\noindent
To thee we are sighing, as mournful and weeping,\par\noindent
We pass through this vale of sorrow.\\
\par\noindent
Turn thou therefore, O our intercessor,\par\noindent
Those thine eyes of pity and loving-kindness upon us sinners.\par\noindent
Hereafter, when our earthly exile shall be ended,\par\noindent
Show us Jesus, the blessed fruit of thy womb.\par\noindent
O gentle, O tender, O gracious Virgin Mary.\\
\end{hangparas}
\par
℣. Pray for us, O holy Mother of God.\par
℟. That we may be made worthy of the promises of Christ.\par
\letuspray
\lett{A}{lmighty} and everlasting God, who by the cooperation of the Holy Ghost, didst prepare the body and soul of the glorious Virgin Mother Mary to become a habitation meet for thy Son: Grant that as we rejoice in her commemoration, we may be delivered by her loving intercession from our present evils and from eternal death. Through the same Christ our Lord. \textit{Amen.}}
{\lett{S}{alve}, Regína, mater misericórdi{\ae};\par
\secondline{Vita, dulcédo et spes nóstra, salve.}
\thirdline{Ad te clamámus éxsules fílii Hev{\ae}.}
Ad te suspirámus geméntes et flentes\par
In hac lacrimárum valle.\\
\par
Eja ergo, advocáta nostra,\par
Illos tuos misericórdes óculos ad nos convérte.\par
Et Jesum, benedíctum fructum ventris tui,\par
Nobis post hoc exsílium osténde.\par
O clemens, o pia, o dulcis Virgo María.\\
\par
℣. Ora pro nobis, sancta Dei Génitrix.\par
℟. Ut digni efficiámur promissiónibus Christi.\par
\oremuslatin
\lett{O}{mn\smash{í}potens} sempitérne Deus, qui gloriós{\ae} Vírginis Matris Marí{\ae} corpus et ánimam, ut dignum Fílii tui habitáculum éffici mererétur, Spíritu Sancto cooperánte, pr{\ae}parásti: da, ut, cujus commemoratióne l{\ae}támur, ejus pia intercessióne, ab instántibus malis et a morte perpétua liberémur. Per eúmdem Christum Dóminum nóstrum. \textit{Amen.}}