\fancyhead[RO,LE]{\textit{None}}
\fancyhead[RE,LO]{}
\section{Hour of None}

\begin{multicols}{2}

\fakecol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\fakecol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\fakecol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\fakecol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\fakecol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}

\subby{Rerum Deus}
\fakecol{\begin{hangparas}{1.25em}{1}
\lett{O}{God,} creation’s secret force,\par
\secondline{Thyself unmoved, all motion's source,}
%\thirdline{Who from the morn till evening ray,}
\par\noindent
Who from the morn till evening ray,\par\noindent
Through all its changes guid'st the day:\\
\par\noindent
Grant us, when this short life is past,\par\noindent
The glorious evening that shall last;\par\noindent
That, by a holy death attained,\par\noindent
Eternal glory may be gained.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Doth live and reign eternally. Amen.\\
\end{hangparas}}
{\lett{R}{erum,} Deus, tenax vigor,\par
\secondline{Immótus in te pérmanens,}
\thirdline{Lucis diúrnæ témpora}
Succéssibus detérminans:\\
\par
Largíre lumen véspere,\par
Quo vita nusquam décidat,\par
Sed prǽmium mortis sacr{\ae}\par
Perénnis instet glória.\\
\par
Præsta, Pater piíssime,\par
Patríque compar Únice,\par
Cum Spíritu Paráclito\par
Regnans per omne. Amen.\\}

\subsec{Mirabilia}
\lett{T}{hy} testimonies are wonderful : therefore doth my soul keep them.\par
\secondline{\psanum{130}When thy word goeth forth : it giveth light and understanding unto the simple.}
\thirdline{\psanum{131}I opened my mouth, and drew in my breath : for my delight was in thy commandments.}
\psanum{132}O look thou upon me, and be merciful unto me : as thou usest to do unto those that love thy Name.\par
\psanum{133}Order my steps in thy word : and so shall no wickedness have dominion over me.\par
\psanum{134}O deliver me from the wrongful dealings of men : and so shall I keep thy commandments.\par
\psanum{135}Shew the light of thy countenance upon thy servant : and teach me thy statutes.\par
\psanum{136}Mine eyes gush out with water : because men keep not thy law.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Justus es, Domine}
\lett{R}{ighteous} art thou, O Lord : and true is thy judgement.\par
\secondline{\psanum{138}The testimonies that thou hast commanded : are exceeding righteous and true.}
\thirdline{\psanum{139}My zeal hath even consumed me : because mine enemies have forgotten thy words.}
\psanum{140}Thy word is tried to the uttermost : and thy servant loveth it.\par
\psanum{141}I am small, and of no reputation : yet do I not forget thy commandments.\par
\psanum{142}Thy righteousness is an everlasting righteousness : and thy law is the truth.\par
\psanum{143}Trouble and heaviness have taken hold upon me : yet is my delight in thy commandments.\par
\psanum{144}The righteousness of thy testimonies is everlasting : O grant me understanding, and I shall live.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Clamavi in toto corde meo}
\lett{I}{call} with my whole heart : hear me, O Lord, I will keep thy statutes.\par
\secondline{\psanum{146}Yea, even unto thee do I call : help me, and I shall keep thy testimonies.}
\thirdline{\psanum{147}Early in the morning do I cry unto thee : for in thy word is my trust.}
\psanum{148}Mine eyes prevent the night-watches : that I might be occupied in thy words.\par
\psanum{149}Hear my voice, O Lord, according unto thy loving-kindness : quicken me, according as thou art wont.\par
\psanum{150}They draw nigh that of malice persecute me : and are far from thy law.\par
\psanum{151}Be thou nigh at hand, O Lord : for all thy commandments are true.\par
\psanum{152}As concerning thy testimonies, I have known long since : that thou hast grounded them for ever.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Vide humilitatem}
\lett{O}{consider} mine adversity, and deliver me : for I do not forget thy law.\par
\secondline{\psanum{154}Avenge thou my cause, and deliver me : quicken me, according to thy word.}
\thirdline{\psanum{155}Health is far from the ungodly : for they regard not thy statutes.}
\psanum{156}Great is thy mercy, O Lord : quicken me, as thou art wont.\par
\psanum{157}Many there are that trouble me, and persecute me : yet do I not swerve from thy testimonies.\par
\psanum{158}It grieveth me when I see the transgressors : because they keep not thy law.\par
\psanum{159}Consider, O Lord, how I love thy commandments : O quicken me, according to thy loving-kindness.\par
\psanum{160}Thy word is true from everlasting : all the judgements of thy righteousness endure for evermore.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%MANUAL ADJUSTMENT:
\subby{Principes persecuti sunt}
\lett{P}{rinces} have persecuted me without a cause : but my heart standeth in awe of thy word.\par
\secondline{\psanum{162}I am as glad of thy word : as one that findeth great spoils.}
\thirdline{\psanum{163}As for lies, I hate and abhor them : but thy law do I love.}
\psanum{164}Seven times a day do I praise thee : because of thy righteous judgements.\par
\psanum{165}Great is the peace that they have who love thy law : and they are not offended at it.\par
\psanum{166}Lord, I have looked for thy saving health : and done after thy commandments.\par
\psanum{167}My soul hath kept thy testimonies : and loved them exceedingly.\par
\psanum{168}I have kept thy commandments and testimonies : for all my ways are before thee.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Appropinquet deprecatio}
\lett{L}{et} my complaint come before thee, O Lord : give me understanding, according to thy word.\par
\secondline{\psanum{170}Let my supplication come before thee : deliver me, according to thy word.}
\thirdline{\psanum{171}My lips shall speak of thy praise : when thou hast taught me thy statutes.}
\psanum{172}Yea, my tongue shall sing of thy word : for all thy commandments are righteous.\par
\psanum{173}Let thine hand help me : for I have chosen thy commandments.\par
\psanum{174}I have longed for thy saving health, O Lord : and in thy law is my delight.\par
\psanum{175}O let my soul live, and it shall praise thee : and thy judgements shall help me.\par
\psanum{176}I have gone astray like a sheep that is lost : O seek thy servant, for I do not forget thy commandments.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subby{Chapter}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}\par\noindent
Bear ye one another's burdens, and so fulfil the law of Christ.

\begin{inhead}
	Sundays \& Feast Days
\end{inhead}\par\noindent
One Lord, one Faith, one Baptism, one God and Father of all, who is above all, and through all, and in you all.

\begin{inhead}
	Eastertide
\end{inhead}\par\noindent
There are three that bear witness in earth: the Spirit, and the Water, and the Blood: and these three agree in one.

\fakecol{℟. Thanks be to God.}{℟. Deo grátias.}

\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\fakecol{℣. Deliver me, O Lord, and be merciful unto me.}{℣. Rédime me, et miserére mei.}

\fakecol{℟. Deliver me, O Lord, and be merciful unto me.}{℟. Rédime me, et miserére mei.}

\fakecol{℣. My foot standeth right; I will praise the Lord in the congregations.}{℣. Pes meus stetit in dirécto: in ecclésiis benedícam te, Dómine.}

\fakecol{℟. And be merciful unto me.}{℟. Et miserére mei.}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\fakecol{℟. Deliver me, O Lord, and be merciful unto me.}{℟. Rédime me, et miserére mei.}

\fakecol{℣. I will praise thy Name, O Lord.}{℣. Et confitébor nómini tuo, Dómine.}

\fakecol{℟. For thou hast delivered me out of all my trouble.}{℟. Quóniam ex omni tribulatióne eripuísti me.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\fakecol{℣. I call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℣. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\fakecol{℟. I call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℟. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\fakecol{℣. I will keep thy statutes.}{℣. Justificatiónes tuas requíram.}

\fakecol{℟. Hear me, O Lord (alleluia, alleluia).}{℟. Exáudi me, Dómine (allelúja, allelúja).}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}

\fakecol{℟. I will call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℟. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\fakecol{℣. Cleanse me, O Lord, from my secret faults (alleluia).}{℣. Ab occúltis meis munda me (allelúja).}

\fakecol{℟. Keep thy servant also from presumptuous sins (alleluia).}{℟. Et ab aliénis parce servo tuo (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages are said, kneeling, on all Ferias outside of Eastertide.
\end{rubric}

\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\fakecol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\fakecol{℣. Cast me not away in the time of age.}{℣. Ne proícias me in témpore senectútis.}

\fakecol{℟. Forsake me not when my strength faileth me.}{℟. Cum defécerit virtus mea, ne derelínquas me.}

\fakecol{℣. Hide not thy face from me.}{℣. Non avértas fáciem tuam a me.}

\fakecol{℟. Lest I be like unto them that go down into the pit.}{℟. Et símilis ero descendéntibus in lacum.}

\fakecol{℣. Quicken me, O Lord, for thy Name's sake.}{℣. Propter nomen tuum, Dómine, vivificábis me.}

\fakecol{℟. And for thy righteousness' sake bring my soul out of trouble.}{℟. In {\ae}quitáte tua edúces de tribulatióne ánimam meam}

\subby{Collects}
\fakecol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\fakecol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\fakecol{℣. Let us pray.}{℣. Orémus.}

\needspace{4\baselineskip}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}

\lett{L}{ord} Jesus Christ, who for our sakes didst tread the paths of death; make known to us the way of life; that as thou wast reckoned with the transgressors in thy death, and with the rich in thy burial, so we, who are dead in trespasses and sins, may be raised up by thee to the land of true riches. Who livest and reignest with the Father and the Holy Ghost, ever one God, world without end. \textit{Amen.}\\

\fakecol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\fakecol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\fakecol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\fakecol{℟. Thanks be to God.}{℟. Deo grátias}
\fakecol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\needspace{4\baselineskip}
\lett{N}{ow} unto him that is able to keep us from falling and to present us faultless before the presence of his glory with exceeding joy; to the only wise God our Saviour be glory and majesty, dominion, and power, both now and ever. \textit{Amen.}

\end{multicols}

\begin{center}
	{\textsc{Here endeth the Order of None.}}
\end{center}