\fancyhead[RO,LE]{\textit{Old Testament Canticles}}
\section{Old Testament Canticles}\label{OT}
\begin{rubric}
    The Festal Canticle is prayed on Feast Days Double and higher, on Octave Days, and the Saturday Office of St. Mary.
\end{rubric}
\begin{multicols}{2}
\subby{Confitebor tibi}
\fancyhead[RE,LO]{Confitebor tibi}
\begin{inhead}
Monday Ferial Canticle
\end{inhead}
\lett{O}{Lord,} I will praise thee, though thou wast angry with me; * thine anger is turned away, and thou comfortedst me.\par
\secondline{Behold, God is my salvation; * I will trust, and not be afraid:}
\thirdline{For the Lord Jehovah is my strength and my song; * he also is become my salvation.}
Therefore with joy shall ye draw water out of the wells of salvation: * and in that day shall ye say, Praise the Lord, call upon his Name,\par
Declare his doings among the people, * make mention that his Name is exalted.\par
Sing unto the Lord; for he hath done excellent things: * this is known in all the earth.\par
Cry out and shout, thou inhabitant of Sion: * for great is the Holy One of Israel in the midst of thee.
\subby{Benedictus es, Domine}
\fancyhead[RE,LO]{Benedictus es, Domine}
\begin{inhead}
Monday Festal Canticle
\end{inhead}
\lett{B}{lessed} be thou, Lord God of Israel * our father, for ever and ever.\par
\secondline{Thine, O Lord is the greatness, and the power, and the glory, {\dag} and the victory, and the majesty: * for all that is in the heaven and in the earth is thine;}
\thirdline{Thine is the kingdom, O Lord, * and thou art exalted as head above all.}
Both riches and honour come of thee, * and thou reignest over all;\par
And in thine hand is power and might; * and in thine hand it is to make great, and to give strength unto all.\par
Now therefore, our God, we thank thee, and praise thy glorious name.
\subby{Ego dixi}
\fancyhead[RE,LO]{Ego dixi}
\begin{inhead}
Tuesday Ferial Canticle
\end{inhead}
\lett{I}{said} in the cutting off of my days, {\dag} I shall go to the gates of the grave: * I am deprived of the residue of my years.\par
\secondline{I said, I shall not see the Lord, {\dag} even the Lord, in the land of the living: * I shall behold man no more with the inhabitants of the world.}
\thirdline{Mine age is departed, * and is removed from me as a shepherd's tent;}
I have cut off like a weaver my life: {\dag} he will cut me off with pining sickness: * from day even to night wilt thou make an end of me.\par
I reckoned till morning, {\dag} that, as a lion, so will he break all my bones: * from day even to night wilt thou make an end of me.\par
Like a crane or a swallow, so did I chatter: * I did mourn as a dove.\par
Mine eyes fail with looking upward: * O Lord, I am oppressed; undertake for me.\par
What shall I say? {\dag} he hath both spoken unto me, and himself hath done it: * I shall go softly all my years in the bitterness of my soul.\par
O Lord, by these things men live, {\dag} and in all these things is the life of my spirit: * so wilt thou recover me, and make me to live.\par
Behold, for peace I had great bitterness: {\dag} but thou hast in love to my soul delivered it from the pit of corruption: * for thou hast cast all my sins behind thy back.\par
For the grave cannot praise thee, {\dag} death can not celebrate thee: * they that go down into the pit cannot hope for thy truth.\par
The living, the living, he shall praise thee, as I do this day: * the father to the children shall make known thy truth.\par
The Lord was ready to save me: * therefore we will sing my songs to the stringed instruments all the days of our life in the house of the Lord.
\subby{Magnus es, Domine}
\fancyhead[RE,LO]{Magnus es, Domine}
\begin{inhead}
Tuesday Festal Canticle
\end{inhead}
\lett{B}{lessed} be God that liveth for ever, * and blessed be his kingdom.\par
\secondline{For he doth scourge, and hath mercy: {\dag} he leadeth down to hell, and bringeth up again: * neither is there any that can avoid his hand.}
\thirdline{Confess him before the Gentiles, ye children of Israel: * for he hath scattered us among them.}
There declare his greatness, and extol him before all the living: * for he is our Lord, and he is the God our Father for ever.\par
And he will scourge us for our iniquities, * and will have mercy again, and will gather us out of all nations, among whom he hath scattered us.\par
If ye turn to him with your whole heart, and with your whole mind, {\dag} and deal uprightly before him, * then will he turn unto you, and will not hide his face from you.\par
Therefore see what he will do with you, and confess him with your whole mouth, and praise the Lord of might, * and extol the everlasting King.\par
In the land of my captivity do I praise him, * and declare his might and majesty to a sinful nation.\par
O ye sinners, turn and do justice before him: * who can tell if he will accept you, and have mercy on you?\par
I will extol my God, * and my soul shall praise the King of heaven, and shall rejoice in his greatness.
\subby{Exsultavit cor meum}
\fancyhead[RE,LO]{Exsultavit cor meum}
\begin{inhead}
Wednesday Ferial Canticle
\end{inhead}
\lett{M}{y} heart rejoiceth in the Lord,* mine horn is exalted in the Lord:\par
\secondline{My mouth is enlarged over mine enemies; * because I rejoice in thy salvation.}
\thirdline{There is none holy as the Lord: {\dag} for there is none beside thee: * neither is there any rock like our God.}
Talk no more so exceeding proudly; * let not arrogancy come out of your mouth:\par
For the Lord is a God of knowledge, * and by him actions are weighed.\par
The bows of the mighty men are broken, * and they that stumbled are girded with strength.\par
They that were full have hired out themselves for bread; * and they that were hungry ceased:\par
So that the barren hath born seven; * and she that hath many children is waxed feeble.\par
The Lord killeth, and maketh alive: * he bringeth down to the grave, and bringeth up.\par
The Lord maketh poor, and maketh rich: * he bringeth low, and lifteth up.\par
He raiseth up the poor out of the dust, * and lifteth up the beggar from the dunghill,\par
To set them among princes, * and to make them inherit the throne of glory:\par
For the pillars of the earth are the Lord's, * and he hath set the world upon them.\par
He will keep the feet of his saints, {\dag} and the wicked shall be silent in darkness; * for by strength shall no man prevail.\par
The adversaries of the Lord shall be broken to pieces; * out of heaven shall he thunder upon them:\par
The Lord shall judge the ends of the earth; {\dag} and he shall give strength unto his King, * and exalt the horn of his Anointed.
\subby{Hymnum cantemus}
\fancyhead[RE,LO]{Hymnum cantemus}
\begin{inhead}
Wednesday Festal Canticle
\end{inhead}
\lett{I}{will} sing unto the Lord a new song: * O Lord, thou art great and glorious, wonderful in strength, and invincible.\par
\secondline{Let all creatures serve thee: * for thou spakest, and they were made,}
\thirdline{Thou didst send forth thy spirit, and it created them, * and there is none that can resist thy voice.}
For the mountains shall be moved from their foundations with the waters, {\dag} the rocks shall melt as wax at thy presence: * yet thou art merciful to them that fear thee.\par
For all sacrifice is too little for a sweet savour unto thee: * but he that feareth the Lord is great at all times.\par
Woe to the nations that rise up against my kindred! * the Lord Almighty will take vengeance of them in the day of judgment.
\subby{Cantemus Domino}
\fancyhead[RE,LO]{Cantemus Domino}
\begin{inhead}
Thursday Ferial Canticle
\end{inhead}
\lett{I}{will} sing unto the Lord, for he hath triumphed gloriously: * the horse and his rider hath he thrown into the sea.\par
\secondline{The Lord is my strength and song, * and he is become my salvation:}
\thirdline{He is my God, and I will prepare him an habitation; * my father’s God, and I will exalt him.}
The Lord is a man of war: * the Lord is his Name.\par
Pharaoh’s chariots and his host hath he cast into the sea: * his chosen captains also are drowned in the Red Sea.\par
The depths have covered them: * they sank into the bottom as a stone.\par
Thy right hand, O Lord, is become glorious in power: * thy right hand, O Lord, hath dashed in pieces the enemy.\par
And in the greatness of thine excellency {\dag} thou hast overthrown them that rose up against thee: * thou sentest forth thy wrath, which consumed them as stubble.\par
And with the blast of thy nostrils the waters were gathered together, * the floods stood upright as an heap, and the depths were congealed in the heart of the sea.\par
The enemy said, {\dag} I will pursue, I will overtake, I will divide the spoil; * my lust shall be satisfied upon them;\par
I will draw my sword, * my hand shall destroy them.\par
Thou didst blow with thy wind, the sea covered them: * they sank as lead in the mighty waters.\par
Who is like unto thee, O Lord, among the gods? * who is like thee, glorious in holiness, fearful in praises, doing wonders?\par
Thou stretchedst out thy right hand, * the earth swallowed them.\par
Thou in thy mercy hast led forth the people which thou hast redeemed: * thou hast guided them in thy strength unto thy holy habitation.\par
The people shall hear, and be afraid: * sorrow shall take hold on the inhabitants of Palestina.\par
Then the dukes of Edom shall be amazed; * the mighty men of Moab, trembling shall take hold upon them; all the inhabitants of Canaan shall melt away.\par
Fear and dread shall fall upon them; {\dag} by the greatness of thine arm they shall be as still as a stone; * till thy people pass over, O Lord, till the people pass over, which thou hast purchased.\par
Thou shalt bring them in, and plant them in the mountain of thine inheritance, * in the place, O Lord, which thou hast made for thee to dwell in, in the Sanctuary, O Lord, which thy hands have established.\par
The Lord shall reign * for ever and ever.\par
For the horse of Pharaoh went in with his chariots and with his horsemen into the sea, * and the Lord brought again the waters of the sea upon them.\par
But the children of Israel went on dry land * in the midst of the sea.
\subby{Audite verbum}
\fancyhead[RE,LO]{Audite verbum}
\begin{inhead}
Thursday Festal Canticle
\end{inhead}
\lett{H}{ear} the word of the Lord, O ye nations, * and declare it in the isles afar off,\par
\secondline{And say, He that scattered Israel will gather him, * and keep him, as a shepherd doth his flock.}
\thirdline{For the Lord hath redeemed Jacob, * and ransomed him from the hand of him that was stronger than he.}
Therefore they shall come and sing in the height of Sion, * and shall flow together to the goodness of the Lord,\par
For wheat, and for wine, and for oil, * and for the young of the flock and of the herd:\par
And their soul shall be as a watered garden; * and they shall not sorrow any more at all.\par
Then shall the virgin rejoice in the dance, * both young men and old together:\par
For I will turn their mourning into joy, * and will comfort them, and make them rejoice from their sorrow.\par
And I will satiate the soul of the priests with fatness, * and my people shall be satisfied with my goodness, saith the Lord.
\subby{Domine audivi}
\fancyhead[RE,LO]{Domine audivi}
\begin{inhead}
Friday Ferial Canticle
\end{inhead}
\lett{O}{lord,} I have heard thy speech, * and was afraid:\par
\secondline{O Lord, revive thy work in the midst of the years, {\dag} in the midst of the years make known; * in wrath remember mercy.}
\thirdline{God came from Teman, * and the Holy One from mount Paran.}
His glory covered the heavens, * and the earth was full of his praise.\par
And his brightness was as the light; {\dag} he had horns coming out of his hand: * and there was the hiding of his power.\par
Before him went the pestilence, * and burning coals went forth at his feet.\par
He stood, and measured the earth: * he beheld, and drove asunder the nations;\par
And the everlasting mountains were scattered, {\dag} the perpetual hills did bow: * his ways are everlasting.\par
I saw the tents of Cushan in affliction: * and the curtains of the land of Midian did tremble.\par
Was the Lord displeased against the rivers? * was thine anger against the rivers?\par
Was thy wrath against the sea, * that thou didst ride upon thine horses and thy chariots of salvation?\par
Thy bow was made quite naked, * according to the oaths of the tribes, even thy word.\par
Thou didst cleave the earth with rivers: * the mountains saw thee, and they trembled:\par
The overflowing of the water passed by: * the deep uttered his voice, and lifted up his hands on high.\par
The sun and moon stood still in their habitation: * at the light of thine arrows they went, and at the shining of thy glittering spear.\par
Thou didst march through the land in indignation, * thou didst thresh the heathen in anger.\par
Thou wentest forth for the salvation of thy people, * even for salvation with thine Anointed;\par
Thou woundedst the head out of the house of the wicked, * by discovering the foundation unto the neck.\par
Thou didst strike through with his staves the head of his villages: * they came out as a whirlwind to scatter me: their rejoicing was as to devour the poor secretly.\par
Thou didst walk through the sea with thine horses, * through the heap of great waters.\par
When I heard, my belly trembled; * my lips quivered at the voice:\par
Rottenness entered into my bones, * and I trembled in myself, that I might rest in the day of trouble:\par
When he cometh up unto the people, * he will invade them with his troops.\par
Although the fig tree shall not blossom, * neither shall fruit be in the vines;\par
The labour of the olive shall fail, * and the fields shall yield no meat;\par
The flock shall be cut off from the fold, * and there shall be no herd in the stalls:\par
Yet I will rejoice in the Lord, * I will joy in the God of my salvation.\par
The Lord God is my strength, and he will make my feet like hinds' feet, * and he will make me to walk upon mine high places.
\subby{Vere tu es Deus absconditus}
\fancyhead[RE,LO]{Vere tu es Deus}
\begin{inhead}
Friday Festal Canticle
\end{inhead}
\lett{V}{erily} thou art a God that hidest thyself, * O God of Israel, the Saviour.\par
\secondline{They shall be ashamed, and also confounded, all of them: * they shall go to confusion together that are makers of idols.}
\thirdline{But Israel shall be saved in the Lord with an everlasting salvation: * ye shall not be ashamed nor confounded world without end.}
For thus saith the Lord that created the heavens; * God himself that formed the earth and made it;\par
He hath established it, he created it not in vain, * he formed it to be inhabited:\par
I am the Lord; * and there is none else.\par
I have not spoken in secret, * in a dark place of the earth:\par
I said not unto the seed of Jacob, Seek ye me in vain: * I the Lord speak righteousness, I declare things that are right.\par
Assemble yourselves and come; * draw near together, ye that are escaped of the nations:\par
They have no knowledge that set up the wood of their graven image, * and pray unto a god that cannot save.\par
Tell ye, and bring them near; {\dag} yea, let them take counsel together: * who hath declared this from ancient time? who hath told it from that time?\par
Have not I the Lord? and there is no God else beside me; * a just God and a Saviour; there is none beside me.\par
Look unto me, and be ye saved, all the ends of the earth: * for I am God, and there is none else.\par
I have sworn by myself, * the word is gone out of my mouth in righteousness, and shall not return,\par
That unto me every knee shall bow, * every tongue shall swear.\par
Surely, shall one say, in the Lord have I righteousness and strength: * even to him shall men come; and all that are incensed against him shall be ashamed.\par
In the Lord shall all the seed of Israel be justified, * and shall glory.
\subby{Audite c{\ae}li}
\fancyhead[RE,LO]{Audite c{\ae}li}
\begin{inhead}
Saturday Ferial Canticle
\end{inhead}
\lett{G}{ive} ear, O ye heavens, and I will speak; * and hear, O earth, the words of my mouth.\par
\secondline{My doctrine shall drop as the rain, {\dag} my speech shall distil as the dew, * as the small rain upon the tender herb, and as the showers upon the grass:}
\thirdline{Because I will publish the Name of the Lord: * ascribe ye greatness unto our God.}
He is the Rock, his work is perfect: * for all his ways are judgment:\par
A God of truth and without iniquity, * just and right is he.\par
They have corrupted themselves, {\dag} their spot is not the spot of his children: * they are a perverse and crooked generation.\par
Do ye thus requite the Lord, O foolish people and unwise? * is not he thy father that hath bought thee? hath he not made thee, and established thee?\par
Remember the days of old, * consider the years of many generations:\par
Ask thy father, and he will shew thee; * thy elders, and they will tell thee.\par
When the Most High divided to the nations their inheritance, * when he separated the sons of Adam,\par
He set the bounds of the people * according to the number of the children of Israel.\par
For the Lord's portion is his people; * Jacob is the lot of his inheritance.\par
He found him in a desert land, * and in the waste howling wilderness;\par
He led him about, he instructed him, * he kept him as the apple of his eye.\par
As an eagle stirreth up her nest, fluttereth over her young, * spreadeth abroad her wings, taketh them, beareth them on her wings:\par
So the Lord alone did lead him, * and there was no strange god with him.\par
He made him ride on the high places of the earth, * that he might eat the increase of the fields;\par
And he made him to suck honey out of the rock, * and oil out of the flinty rock;\par
Butter of kine, and milk of sheep, * with fat of lambs, and rams of the breed of Bashan,\par
And goats, with the fat of kidneys of wheat; * and thou didst drink the pure blood of the grape.\par
But Jeshurun waxed fat, and kicked: * thou art waxen fat, thou art grown thick, thou art covered with fatness;\par
Then he forsook God which made him, * and lightly esteemed the Rock of his salvation.\par
They provoked him to jealousy with strange gods, * with abominations provoked they him to anger.\par
They sacrificed unto devils, not to God; * to gods whom they knew not, to new gods that came newly up, whom your fathers feared not.\par
Of the Rock that begat thee thou art unmindful, * and hast forgotten God that formed thee.\par
And when the Lord saw it, he abhorred them, * because of the provoking of his sons, and of his daughters.\par
And he said, I will hide my face from them, {\dag} I will see what their end shall be: * for they are a very froward generation, children in whom is no faith.\par
They have moved me to jealousy with that which is not God; * they have provoked me to anger with their vanities:\par
And I will move them to jealousy with those which are not a people; * I will provoke them to anger with a foolish nation.
\begin{inhead}
Ignis succensus est
\end{inhead}
\lett{F}{or} a fire is kindled in mine anger, {\dag} and shall burn unto the lowest hell, * and shall consume the earth with her increase, and set on fire the foundations of the mountains.\par
\secondline{I will heap mischiefs upon them; * I will spend mine arrows upon them.}
\thirdline{They shall be burnt with hunger, * and devoured with burning heat, and with bitter destruction:}
I will also send the teeth of beasts upon them, * with the poison of serpents of the dust.\par
The sword without, and terror within, {\dag} shall destroy both the young man and the virgin, * the suckling also with the man of gray hairs.\par
I said, I would scatter them into corners, * I would make the remembrance of them to cease from among men:\par
Were it not that I feared the wrath of the enemy, * lest their adversaries should behave themselves strangely,\par
And lest they should say, Our hand is high, * and the Lord hath not done all this.\par
For they are a nation void of counsel, * neither is there any understanding in them.\par
O that they were wise, that they understood this, * that they would consider their latter end!\par
How should one chase a thousand, {\dag} and two put ten thousand to flight, * except their Rock had sold them, and the Lord had shut them up?\par
For their rock is not as our Rock, * even our enemies themselves being judges.\par
For their vine is of the vine of Sodom, * and of the fields of Gomorrah:\par
Their grapes are grapes of gall, * their clusters are bitter:\par
Their wine is the poison of dragons, * and the cruel venom of asps.\par
Is not this laid up in store with me, {\dag} and sealed up among my treasures? to me belongeth vengeance, and recompence;\par
Their foot shall slide in due time: * for the day of their calamity is at hand, and the things that shall come upon them make haste.\par
For the Lord shall judge his people, * and repent himself for his servants,\par
When he seeth that their power is gone, * and there is none shut up, or left.\par
And he shall say, Where are their gods, * their rock in whom they trusted,\par
Which did eat the fat of their sacrifices, * and drank the wine of their drink offerings?\par
Let them rise up and help you, *and be your protection.\par
See now that I, even I, am he, * and there is no god with me:\par
I kill, and I make alive; {\dag} I wound, and I heal: * neither is there any that can deliver out of my hand.\par
For I lift up my hand to heaven, and say, * I live for ever.\par
If I whet my glittering sword, and mine hand take hold on judgment; * I will render vengeance to mine enemies, and will reward them that hate me.\par
I will make mine arrows drunk with blood, {\dag} and my sword shall devour flesh; * and that with the blood of the slain and of the captives, from the beginning of revenges upon the enemy.\par
Rejoice, O ye nations, with his people: * for he will avenge the blood of his servants,\par
And will render vengeance to his adversaries, * and will be merciful unto his land, and to his people.
\subby{Miserere nostri, Deus}
\fancyhead[RE,LO]{Miserere nostri, Deus}
\begin{inhead}
Saturday Festal Canticle
\end{inhead}
\begin{rubric}
    This Canticle is used when Sunday Mattins is anticipated on Saturday evening.
\end{rubric}
\lett{H}{ave} mercy upon us, O Lord God of all, and behold us: * and send thy fear upon all the nations that seek not after thee.\par
\secondline{Lift up thy hand against the strange nations, * and let them see thy power.}
\thirdline{As thou wast sanctified in us before them: * so be thou magnified among them before us.}
And let them know thee, as we have known thee, * that there is no God but only thou, O God.\par
Shew new signs, and repeat thy wonders: * glorify thy hand and thy right arm, that they may set forth thy wondrous works.\par
Raise up indignation, and pour out wrath: * take away the adversary, and destroy the enemy.\par
Make the time short, remember the covenant, * and let them declare thy wonderful works.\par
Let him that escapeth be consumed by the rage of the fire; * and let them perish that oppress the people.\par
Smite in sunder the heads of the rulers of the heathen, * that say, There is none other but we.\par
Gather all the tribes of Jacob together, * and inherit thou them, as from the beginning.
\end{multicols}