\section{Pretiosa}
\fancyhead[RO,LE]{\textit{Pretiosa}}
\fancyhead[RE,LO]{}
\begin{rubric}
    The \emph{Pretiosa} is a laudable custom and may be added to \emph{Prime}. The Martyrology reading for the Day may here be read. The following is then said.
\end{rubric}
\begin{multicols}{2}
\fakecol{℣. Right dear in the sight of the Lord.}{℣. Preciósa est in conspéctu Dómini.}
\fakecol{℟. Is the death of his Saints.\\}{℟. Mors sanctórum ejus.\\}

\needspace{4\baselineskip}
\fakecol{\lett{M}{ay} holy Mary and all the Saints intercede for us to the Lord: that we may be worthy to obtain from him help and salvation, who liveth and reigneth for ever and ever. \textit{Amen.}\\}{\lett{S}{ancta} María mater Dómini nostri Jesu Christi: atque omnes sancti justi et elécti Dei intercédant et orent pro nobis peccatóribus ad Dóminum Deum nostrum: ut nos mereámur ab eo adjuvári et salvári: qui in Trinitáte perfécta vivit et regnat Deus. Per ómnia sǽcula s{\ae}culórum. \textit{Amen.}\\}

\fakecol{℣. O God, {\ding{64}} make speed to save me.}{℣. Deus, {\ding{64}} in adiutórium meum inténde.}
\fakecol{℟. O Lord, make haste to help me.}{℟. Dómine, ad adiuvándum me festína.}
\fakecol{℣. O God, {\ding{64}} make speed to save me.}{℣. Deus, {\ding{64}} in adiutórium meum inténde.}
\fakecol{℟. O Lord, make haste to help me.}{℟. Dómine, ad adiuvándum me festína.}
\fakecol{℣. O God, {\ding{64}} make speed to save me.}{℣. Deus, {\ding{64}} in adiutórium meum inténde.}
\fakecol{℟. O Lord, make haste to help me.\\}{℟. Dómine, ad adiuvándum me festína.\\}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\fakecol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}

\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.\\}{℣. Kýrie, eléison.\\}

\needspace{4\baselineskip}
\fakecol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\fakecol{℣. Let thy loving mercy come also unto us, O Lord.}{℣. Et véniat super nos misericórdia tua, Dómine.}
\fakecol{℟. Even thy salvation, according unto thy word.}{℟. Salutáre tuum secúndum elóquium tuum.}
\fakecol{℣. And show thy servants thy work.}{℣. Et réspice in servos tuos et in ópera tua.}
\fakecol{℟. And their children thy glory.}{℟. Et dírige fílios eórum.}
\fakecol{℣. And the glorious Majesty of the Lord our God be upon us.}{℣. Et sit splendor Dómini Dei nostri super nos}
\fakecol{℟. And prosper thou the work of our hands upon us: O prosper thou our handy-work.}{℟. Et ópera mánuum nostrárum dírige super nos: et opus mánuum nostrárum dírige.}
\letuspray

\fakecol{\lett{O}{almighty} Lord and everlasting God, vouchsafe, we beseech thee, to direct, sanctify, and govern, both our hearts and bodies in the ways of thy laws and in the works of thy commandments: that through thy most mighty protection, both here and ever, we may be preserved in body and soul; through our Lord and Saviour Jesus Christ. \textit{Amen.}}{\lett{D}{ir\smash{í}gere} et sanctificáre et régere dignáre Dómine Deus quǽsumus corda et córpora nostra in lege tua : et in opéribus mandatórum tuórum : ut hic et in {\ae}térnum te auxiliánte semper sani et salvi esse mereámur. Per Dóminum nostrum Jesum Christum Filium tuum. \textit{Amen.}}

\fakecol{℣. May the Divine Assistance remain always with us.}{℣. Divínum auxílium máneat semper nobíscum.}
\fakecol{℟. Amen.}{℟. Amen.}
\end{multicols}