\fancyhead[RO,LE]{\textit{Prime}}
\fancyhead[RE,LO]{}
\section{Hour of Prime}
\begin{secrubric}
	This service may be said at any convenient time before or after the saying of Morning Prayer, but not in substitution therefor.
\end{secrubric}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}
\subby{Jam Lucis Ordo Sidere}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{N}{ow} that the daylight fills the sky,\par
\secondline{We lift our hearts to God on high,}
%\thirdline{That he, in all we do or say,}
\par\noindent
That he, in all we do or say,\par\noindent
Would keep us free from harm to-day:\\
\par\noindent
Would guard our hearts and tongues from strife;\par\noindent
From anger's din would hide our life;\par\noindent
From all ill sights would turn our eyes,\par\noindent
Would close our ears from vanities:\\
\par\noindent
Would keep our inmost conscience pure;\par\noindent
Our souls from folly would secure;\par\noindent
Would bid us check the pride of sense\par\noindent
With due and holy abstinence.\\
\par\noindent
So we, when this new day is gone,\par\noindent
And night in turn is drawing on,\par\noindent
With conscience by the world unstained\par\noindent
Shall praise his name for victory gained.\\
\par\noindent
All laud to God the Father be;\par\noindent
All praise, eternal Son, to thee;\par\noindent
All glory, as is ever meet,\par\noindent
To God the holy Paraclete. Amen.\\
\end{hangparas}}
{\lett{J}{am} lucis orto sídere,\par
\secondline{Deum precémur súpplices,}
\thirdline{Ut in diúrnis áctibus}
Nos servet a nocéntibus.\\

Linguam refrénans témperet,\par
Ne litis horror ínsonet:\par
Visum fovéndo cóntegat,\par
Ne vanitátes háuriat.\\

Sint pura cordis íntima,\par
Absístat et vecórdia;\par
Carnis terat supérbiam\par
Potus cibíque párcitas.\\

Ut, cum dies abscésserit,\par
Noctémque sors redúxerit,\par
Mundi per abstinéntiam\par
Ipsi canámus glóriam.\\

Deo Patri sit glória,\par
Ejúsque soli Fílio,\par
Cum Spíritu Paráclito,\par
Nunc et per omne sǽculum. Amen.\\}

\subby{Psalm 54. \textit{Deus, in nomine}}
\lett{S}{ave} me, O God, for thy Name's sake : and avenge me in thy strength.\par
\secondlinepsalm{Hear my prayer, O God : and hearken unto the words of my mouth.}
\thirdlinepsalm{For strangers are risen up against me : and tyrants, which have not God before their eyes, seek after my soul.}
\psanum{4}Behold, God is my helper : the Lord is with them that uphold my soul.\par
\psanum{5}He shall reward evil unto mine enemies : destroy thou them in thy truth.\par
\psanum{6}An offering of a free heart will I give thee, and praise thy Name, O Lord : because it is so comfortable.\par
\psanum{7}For he hath delivered me out of all my trouble : and mine eye hath seen his desire upon mine enemies.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subby{Psalm 119. \textit{Beati immaculati}}
\lett{B}{lessed} are those that are undefiled in the way : and walk in the law of the Lord.\par
\secondlinepsalm{Blessed are they that keep his testimonies : and seek him with their whole heart.}
\thirdlinepsalm{For they who do no wickedness : walk in his ways.}
\psanum{4}Thou hast charged : that we shall diligently keep thy commandments.\par
\psanum{5}O that my ways were made so direct : that I might keep thy statutes!\par
\psanum{6}So shall I not be confounded : while I have respect unto all thy commandments.\par
\psanum{7}I will thank thee with an unfeigned heart : when I shall have learned the judgements of thy righteousness.\par
\psanum{8}I will keep thy ceremonies : O forsake me not utterly.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%MANUAL ADJUSTMENT:
\vspace{-1.625ex}
\subsec{In quo corriget?}
\lett{W}{herewithal} shall a young man cleanse his way : even by ruling himself after thy word.\par
\secondlinepsalmonenineteen{With my whole heart have I sought thee : O let me not go wrong out of thy commandments.}{10}
\thirdlinepsalmonenineteen{Thy words have I hid within my heart : that I should not sin against thee.}{11}
\psanum{12}Blessed art thou, O Lord : O teach me thy statutes.\par
\psanum{13}With my lips have I been telling : of all the judgements of thy mouth.\par
\psanum{14}I have had as great delight in the way of thy testimonies : as in all manner of riches.\par
\psanum{15}I will talk of thy commandments : and have respect unto thy ways.\par
\psanum{16}My delight shall be in thy statutes : and I will not forget thy word.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%MANUAL ADJUSTMENT:
\vspace{-1.625ex}
\subsec{Retribue servo tuo}
\lett{O}{do} well unto thy servant : that I may live, and keep thy word.\par
\secondlinepsalmonenineteen{Open thou mine eyes : that I may see the wondrous things of thy law.}{18}
\thirdlinepsalmonenineteen{I am a stranger upon earth : O hide not thy commandments from me.}{19}
\psanum{20}My soul breaketh out for the very fervent desire : that it hath alway unto thy judgements.\par
\psanum{21}Thou hast rebuked the proud : and cursed are they that do err from thy commandments.\par
\psanum{22}O turn from me shame and rebuke : for I have kept thy testimonies.\par
\psanum{23}Princes also did sit and speak against me : but thy servant is occupied in thy statutes.\par
\psanum{24}For thy testimonies are my delight : and my counsellors.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%\vspace{-3ex}
\subsec{Adhaesit pavimento}
\lett{M}{y} soul cleaveth to the dust : O quicken thou me, according to thy word.\par
\secondlinepsalmonenineteen{I have acknowledged my ways, and thou heardest me : O teach me thy statutes.}{26}
\thirdlinepsalmonenineteen{Make me to understand the way of thy commandments : and so shall I talk of thy wondrous works.}{27}
\psanum{28}My soul melteth away for very heaviness : comfort thou me according unto thy word.\par
\psanum{29}Take from me the way of lying : and cause thou me to make much of thy law.\par
\psanum{30}I have chosen the way of truth : and thy judgements have I laid before me.\par
\psanum{31}I have stuck unto thy testimonies : O Lord, confound me not.\par
\psanum{32}I will run the way of thy commandments : when thou hast set my heart at liberty.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\begin{rubric}
    Here, according to the ancient use of Prime, may be said the Athanasian Creed (p. \pageref{Ath}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-2ex}
\subby{Chapter}

%MANUAL ADJUSTMENT:
\vspace{-1ex}
%From Sarum:
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcoldent{Love the truth and peace, saith the Lord of hosts.}{Pacem et veritátem dilígite, ait Dóminus omnípotens.}

\begin{inhead}
	Sundays and Feast Days
\end{inhead}
\elcoldent{Now unto the King eternal, immortal, invisible, the only wise God, be honour and glory for ever and ever.}{Regi s{\ae}culórum immortáli et invisíbili, soli Deo honor et glória in sǽcula s{\ae}culórum.}

\begin{inhead}
	Eastertide
\end{inhead}
\elcoldent{O Lord, be gracious unto us; we have waited for thee: be thou our arm every morning, our salvation also in the time of trouble.}{Dómine, miserére nostri: te enim expectávimus esto bráchium nostrum in mane: et salus nostra in témpore tribulatiónis.}
\begin{rubric}
	After the Chapter is said,
\end{rubric}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subby{Short Respond}
\elcol{℣. Jesu Christ, Son of the living God, have mercy upon us;}{℣. Jesu Christe, Fili Dei vivi, miserére nobis.}

\elcol{℟. Jesu Christ, Son of the living God, have mercy upon us.}{℟. Jesu Christe, Fili Dei vivi, miserére nobis.}

\elcol{℣. Thou that sittest at the right hand of the Father;}{℣. Qui sedes ad déxteram Patris.}
    
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. Jesu Christ, Son of the living God, have mercy upon us.\\}{℟. Jesu Christe, Fili Dei vivi, miserére nobis.\\}

\elcol{℣. O Lord, arise, help us;}{℣. Exsúrge, Christe, ádjuva nos.}

\elcol{℟. And deliver us for thy name’s sake.}{℟. Et líbera nos propter nomen tuum.}
\subby{Suffrages}
\begin{rubric}
	The Suffrages, said kneeling, are said on all Ferias outside of Eastertide. They are omitted on Sundays, Feasts, and within Privileged Octaves.\par
	\textsc{Note,} If omitted, the Office continues with the Collect.
\end{rubric}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\subbysub{Lord's Prayer}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}
\begin{rubric}
	Then shall follow these Versicles and Responses.
\end{rubric}
\elcol{℣. O let my mouth be filled with thy praise;}{℣. Repleátur os meum laude.}

\elcol{℟. That I may sing of thy glory and honour all the day long.}{℟. Ut cantem glóriam tuam, tota die magnitúdinem tuam.}

\elcol{℣. Turn thy face from my sins, O Lord;}{℣. Dómine, avérte fáciem tuam a peccátis meis.}

\elcol{℟. And put out all my misdeeds.}{℟. Et omnes iniquitátes meas dele.}

\elcol{℣. Make me a clean heart, O God;}{℣. Cor mundum crea in me, Deus.}

\elcol{℟. And renew a right spirit within me.}{℟. Et spíritum rectum ínnova in viscéribus meis.}

\elcol{℣. Cast me not away from thy presence;}{℣. Ne projícias me a fácie tua.}

\elcol{℟. And take not thy Holy Spirit from me.}{℟. Et spíritum sanctum tuum ne áuferas a me.}

\elcol{℣. O give me the comfort of thy help again;}{℣. Redde mihi l{\ae}títiam salutáris tui.}

\elcol{℟. And stablish me with thy free spirit.}{℟. Et spíritu principáli confírma me.}

\subbysub{Confession}
\begin{rubric}
	Then shall the Minister and People say together the Confession following,
\end{rubric}
%Replaced with Sarum. Text from the Monastic Diurnal.
\elcol{\lett{I}{confess} to God, to blessed Mary, to all the saints, and to you: that I have sinned exceedingly in thought, word, and deed, by my own fault. I beg holy Mary, all the Saints of God, and you, to pray for me.}{\lett{C}{onf\smash{í}teor} Deo: beát{\ae} Marí{\ae}, ómnibus sanctis, et vobis: peccávi nimis: cogitatióne, locutióne, et ópere, mea culpa. Precor sanctam Maríam, omnes sanctos Dei, et vos oráre pro me.}
\begin{rubric}
	The Minister then stands and says,
\end{rubric}
\par
\needspace{4\baselineskip}
%1928 English Translation
\elcol{\lett{A}{lmighty} God, have mercy upon us, forgive us all our sins and deliver us from all evil, confirm and strengthen us in all goodness, and bring us to life everlasting.}{\lett{M}{isere\smash{á}tur} nostri omnípotens Deus et dimíttat nobis ómnia peccáta nostra: líberet nos ab omni malo : consérvet et confírmet in bono et ad vitam perdúcat {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}

\begin{rubric}
	If a Priest be present, he shall stand and pronounce the following Absolution,
\end{rubric}
\needspace{4\baselineskip}
%1928 English Translation
%\elcol{\lett{M}{ay} the almighty and merciful Lord grant unto you pardon {\ding{64}} and remission of all your sins, time for amendment of life, and the grace and comfort of the Holy Spirit.}{\lett{A}{bsoluti\smash{ó}nem} {\ding{64}} et remissiónem ómnium peccatórum vestrórum, spácium vere p{\ae}niténti{\ae}, emendatiónem vit{\ae}, grátiam et consolatiónem Sancti Spíritus: tríbuat vobis omnípotens et miséricors Dóminus.}
%\elcol{℟. Amen.}{℟. Amen.}
\lett{M}{ay} the almighty and merciful Lord grant unto you pardon {\ding{64}} and remission of all your sins, time for amendment of life, and the grace and comfort of the Holy Spirit. \textit{Amen.}

\elcol{℣. Wilt thou not turn again and quicken us, O God?}{℣. Deus, tu convérsus vivificábis nos?}
\elcol{℟. That thy people may rejoice in thee.}{℟. Et plebs tua l{\ae}tábitur in te.}
\elcol{℣. O Lord, show thy mercy upon us.}{℣. Osténde nobis, Dómine, misericórdiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutáre tuum da nobis.}

\elcol{℣. Vouchsafe, O Lord, to keep us this day without sin;}{℣. Dignáre, Dómine, die isto sine peccáto nos custodíre.}

\elcol{℟. O Lord, have mercy upon us, have mercy upon us.}{℟. Miserére nostri, Dómine, miserére nostri.}

%MANUAL ADJUSTMENT:
\vspace{-2ex}
\subby{Collect}
\elcol{℣. O Lord, hear our prayer.}{℣. Dómine, exáudi oratiónem nostram.}

\elcol{℟. And let our cry come unto thee.}{℟. Et clamor noster ad te véniat.}

\oremus
%\lett{S}{atisfy} us with thy mercy, O Lord, at the beginning of this day; that in fulness of joy, we may find true delight in singing thy praises all the day long. Through.
%From Sarum:
\elcol{\lett{I}{n} this hour of this day, fill us, O Lord, with thy mercy; that going forth in thy strength, we may make our boast of thee all the day long. Through.}{\lett{I}{n} hac hora hujus diéi tua nos Dómine reple misericórdia: ut per totum diem exultántes: in tuis láudibus delectémur. Per Dóminum.}

\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}

%\elcol{\lett{T}{he} Lord bless us, and preserve us from all evil, and bring us to everlasting life; and may the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace. \textit{Amen.}}{\lett{D}{\smash{ó}minus} nos benedícat, et ab omni malo deféndat, et ad vitam perdúcat {\ae}térnam. Et fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace. \textit{Amen.}}
\elcol{\lett{T}{he} Lord bless us, and preserve us from all evil, and bring us to everlasting life; and may the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace. \textit{Amen.}}{\lett{D}{\smash{ó}minus} nos benedícat, {\ding{64}} et ab omni malo deféndat, et ad vitam perdúcat ætérnam. Et fidélium ánimæ per misericórdiam Dei requiéscant in pace. \textit{Amen.}}
\begin{rubric}
	The Office of Prime may end here, or the following devotions may be added.
\end{rubric}