\fancyhead[RO,LE]{\textit{Sext}}
\fancyhead[RE,LO]{}
\section{Hour of Sext}

\begin{multicols}{2}

\fakecol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\fakecol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\fakecol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\fakecol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\fakecol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}

\subby{Rector Potens}
\fakecol{\begin{hangparas}{1.25em}{1}
\lett{O}{God} of truth, O Lord of might,\par
\secondline{Who ord'rest time and change aright,}
\par\noindent
And send'st the early morning ray,\par\noindent
And light'st the glow of perfect day.\\
\par\noindent
Extinguish thou each sinful fire,\par\noindent
And banish every ill desire:\par\noindent
And while thou keep'st the body whole,\par\noindent
Shed forth thy peace upon the soul.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Doth live and reign eternally. Amen.\\
\end{hangparas}}
{\lett{R}{ector} potens, verax Deus,\par
\secondline{Qui témperas rerum vices,}
\thirdline{Splendóre mane illúminas,}
Et ígnibus merídiem:\\
\par
Exstíngue flammas lítium,\par
Aufer calórem nóxium,\par
Confer salútem córporum,\par
Verámque pacem córdium.\\
\par
Pr{\ae}sta, Pater piíssime,\par
Patríque compar Únice,\par
Cum Spíritu Paráclito\par
Regnans per omne sǽculum. Amen.\\}

\subsec{Defecit anima mea}
\lett{M}{y} soul hath longed for thy salvation : and I have a good hope because of thy word.\par
\secondline{\psanum{82}Mine eyes long sore for thy word : saying, O when wilt thou comfort me?}
\thirdline{\psanum{83}For I am become like a bottle in the smoke : yet do I not forget thy statutes.}
\psanum{84}How many are the days of thy servant : when wilt thou be avenged of them that persecute me?\par
\psanum{85}The proud have digged pits for me : which are not after thy law.\par
\psanum{86}All thy commandments are true : they persecute me falsely; O be thou my help.\par
\psanum{87}They had almost made an end of me upon earth : but I forsook not thy commandments.\par
\psanum{88}O quicken me after thy loving-kindness : and so shall I keep the testimonies of thy mouth.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{In aeternum. Domine}
\lett{O}{Lord,} thy word : endureth for ever in heaven.\par
\secondline{\psanum{90}Thy truth also remaineth from one generation to another : thou hast laid the foundation of the earth, and it abideth.}
\thirdline{\psanum{91}They continue this day according to thine ordinance : for all things serve thee.}
\psanum{92}If my delight had not been in thy law : I should have perished in my trouble.\par
\psanum{93}I will never forget thy commandments : for with them thou hast quickened me.\par
\psanum{94}I am thine, O save me : for I have sought thy commandments.\par
\psanum{95}The ungodly laid wait for me to destroy me : but I will consider thy testimonies.\par
\psanum{96}I see that all things come to an end : but thy commandment is exceeding broad.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Quomodo dilexi!}
\lett{L}{ord,} what love have I unto thy law : all the day long is my study in it.\par
\secondline{\psanum{98}Thou through thy commandments hast made me wiser than mine enemies : for they are ever with me.}
\thirdline{\psanum{99}I have more understanding than my teachers : for thy testimonies are my study.}
\psanum{100}I am wiser than the aged : because I keep thy commandments.\par
\psanum{101}I have refrained my feet from every evil way : that I may keep thy word.\par
\psanum{102}I have not shrunk from thy judgements : for thou teachest me.\par
\psanum{103}O how sweet are thy words unto my throat : yea, sweeter than honey unto my mouth.\par
\psanum{104}Through thy commandments I get understanding : therefore I hate all evil ways.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Lucerna pedibus meis}
\lett{T}{hy} word is a lantern unto my feet : and a light unto my paths.\par
\secondline{\psanum{106}I have sworn, and am stedfastly purposed : to keep thy righteous judgements.}
\thirdline{\psanum{107}I am troubled above measure : quicken me, O Lord, according to thy word.}
\psanum{108}Let the free-will offerings of my mouth please thee, O Lord : and teach me thy judgements.\par
\psanum{109}My soul is alway in my hand : yet do I not forget thy law.\par
\psanum{110}The ungodly have laid a snare for me : but yet I swerved not from thy commandments.\par
\psanum{111}Thy testimonies have I claimed as mine heritage for ever : and why? they are the very joy of my heart.\par
\psanum{112}I have applied my heart to fulfil thy statutes alway : even unto the end.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Iniquos odio habui}
\lett{I}{hate} them that imagine evil things : but thy law do I love.\par
\secondline{\psanum{114}Thou art my defence and shield : and my trust is in thy word.}
\thirdline{\psanum{115}Away from me, ye wicked : I will keep the commandments of my God.}
\psanum{116}O stablish me according to thy word, that I may live : and let me not be disappointed of my hope.\par
\psanum{117}Hold thou me up, and I shall be safe : yea, my delight shall be ever in thy statutes.\par
\psanum{118}Thou hast trodden down all them that depart from thy statutes : for they imagine but deceit.\par
\psanum{119}Thou puttest away all the ungodly of the earth like dross : therefore I love thy testimonies.\par
\psanum{120}My flesh trembleth for fear of thee : and I am afraid of thy judgements.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Feci judicium}
\lett{I}{deal} with the thing that is lawful and right : O give me not over unto mine oppressors.\par
\secondline{\psanum{122}Make thou thy servant to delight in that which is good : that the proud do me no wrong.}
\thirdline{\psanum{123}Mine eyes are wasted away with looking for thy health : and for the word of thy righteousness.}
\psanum{124}O deal with thy servant according unto thy loving mercy : and teach me thy statutes.\par
\psanum{125}I am thy servant, O grant me understanding : that I may know thy testimonies.\par
\psanum{126}It is time for thee, Lord, to lay to thine hand : for they have destroyed thy law.\par
\psanum{127}For I love thy commandments : above gold and precious stone.\par
\psanum{128}Therefore hold I straight all thy commandments : and all false ways I utterly abhor.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subby{Chapter}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}\par\noindent
Prove all things, hold fast to that which is good: abstain from all appearance of evil.

\begin{inhead}
	Sundays \& Feast Days
\end{inhead}\par\noindent
There are three that bear record in heaven: The Father, the Word, and the Holy Ghost: and these three are one.

\begin{inhead}
	Eastertide
\end{inhead}\par\noindent
Who is he that overcometh the world, but he that believeth that Jesus is the Son of God? This is he that came by water and blood, even Jesus Christ.

\fakecol{℟. Thanks be to God.}{℟. Deo grátias.}

\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\fakecol{℣. I will bless the Lord at all times.}{℣. Benedícam Dóminum in omni témpore.}

\fakecol{℟. I will bless the Lord at all times.}{℟. Benedícam Dóminum in omni témpore.}

\fakecol{℣. His praise shall ever be in my mouth.}{℣. Semper laus ejus in ore meo.}
    
\fakecol{℟. At all times.}{℟. In omni témpore.}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\fakecol{℟. I will bless the Lord at all times.\\}{℟. Benedícam Dóminum in omni témpore.\\}

\fakecol{℣. Behold, God is my helper.}{℣. Ecce enim, Deus ádjuvat me.}

\fakecol{℟. The Lord is with them that uphold my soul.}{℟. Et Dóminus suscéptor est ánim{\ae} me{\ae}.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\fakecol{℣. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℣. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\fakecol{℟. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℟. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\fakecol{℣. Thy truth also remaineth from one generation to another.}{℣. In generatiónem et generatiónem véritas tua.}

\fakecol{℟. For ever in heaven (alleluia, alleluia).}{℟. Pérmanet in c{\ae}lo (allelúja, allelúja).}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\fakecol{℟. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℟. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\fakecol{℣. The Lord is my shepherd, therefore can I lack nothing (alleluia).}{℣. Dóminus regit me, et nihil mihi déerit.}

\fakecol{℟. He shall feed me in a green pasture (alleluia).}{℟. In loco páscu{\ae} ibi me collocávit (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages are said, kneeling, on all Ferias outside of Eastertide.
\end{rubric}

\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\fakecol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\fakecol{℣. Hide not thy face from thy servant, for I am in trouble.}{℣. Et ne avértas fáciem tuam a púero tuo: quóniam tríbulor.}

\fakecol{℟. O haste thee and hear me.}{℟. Velóciter exáudi me.}

\fakecol{℣. Draw nigh unto my soul and save it.}{℣. Inténde ánim{\ae} me{\ae}, et líbera eam.}

\fakecol{℟. O deliver me, because of mine enemies.}{℟. Propter inimícos meos éripe me.}

\fakecol{℣. Arise, O Christ, and help us.}{℣. Exsúrge, Christe, ádjuva nos.}

\fakecol{℟. And deliver us for thy Name's sake.}{℟. Et líbera nos propter nomen tuum.}

\subby{Collects}
\fakecol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\fakecol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\fakecol{℣. Let us pray.}{℣. Orémus.}

\needspace{4\baselineskip}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}

\lett{L}{ord} Jesus, who didst stretch out thine arms of love on the hard wood of the Cross, that all men might come within the reach of thy saving embrace; clothe us in thy Spirit, that we, stretching forth our hands in loving labour for others, may bring those who know thee not to the knowledge and love of thee, who with the Father and the Holy Ghost livest and reignest ever one God, world without end. \textit{Amen.}\\

\fakecol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\fakecol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\fakecol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\fakecol{℟. Thanks be to God.}{℟. Deo grátias}
\fakecol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\needspace{4\baselineskip}
\lett{N}{ow} unto the King eternal, immortal, invisible, the only wise God, be honour and glory, for ever and ever. \textit{Amen.}

\end{multicols}

\begin{center}
	{\textsc{Here endeth the Order of Sext.}}
\end{center}