\fancyhead[RO,LE]{\textit{Terce}}
\fancyhead[RE,LO]{}
\section{Hour of Terce}

\begin{multicols}{2}

\fakecol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\fakecol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\fakecol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\fakecol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\fakecol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}

\subby{Nunc Sancte}
\begin{hangparas}{1.25em}{1}
\lett{C}{ome} Holy Ghost, with God the Son,\par
\secondline{And God the Father, ever one:}\par\noindent
Shed forth thy grace within our breast,\par\noindent
And dwell with us, a ready Guest.\\
\par\noindent
By every power, by heart and tongue,\par\noindent
By act and deed, thy praise be sung;\par\noindent
Inflame with perfect love each sense,\par\noindent
That others' souls may kindle thence.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Shall live and reign eternally. Amen.\\
\end{hangparas}

\subsec{Legem pone}
\lett{T}{each} me, O Lord, the way of thy statutes : and I shall keep it unto the end.\par
\secondline{\psanum{34}Give me understanding, and I shall keep thy law : yea, I shall keep it with my whole heart.}
\thirdline{\psanum{35}Make me to go in the path of thy commandments : for therein is my desire.}
\psanum{36}Incline my heart unto thy testimonies : and not to covetousness.\par
\psanum{37}O turn away mine eyes, lest they behold vanity : and quicken thou me in thy way.\par
\psanum{38}O stablish thy word in thy servant : that I may fear thee.\par
\psanum{39}Take away the rebuke that I am afraid of : for thy judgements are good.\par
\psanum{40}Behold, my delight is in thy commandments : O quicken me in thy righteousness.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Et veniat super me}
\lett{L}{et} thy loving mercy come also unto me, O Lord : even thy salvation, according unto thy word.\par
\secondline{\psanum{42}So shall I make answer unto my blasphemers : for my trust is in thy word.}
\thirdline{\psanum{43}O take not the word of thy truth utterly out of my mouth : for my hope is in thy judgements.}
\psanum{44}So shall I alway keep thy law : yea, for ever and ever.\par
\psanum{45}And I will walk at liberty : for I seek thy commandments.\par
\psanum{46}I will speak of thy testimonies also, even before kings : and will not be ashamed.\par
\psanum{47}And my delight shall be in thy commandments : which I have loved.\par
\psanum{48}My hands also will I lift up unto thy commandments, which I have loved : and my study shall be in thy statutes.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Memor esto servi tui}
\lett{O}{think} upon thy servant, as concerning thy word : wherein thou hast caused me to put my trust.\par
\secondline{\psanum{50}The same is my comfort in my trouble : for thy word hath quickened me.}
\thirdline{\psanum{51}The proud have had me exceedingly in derision : yet have I not shrinked from thy law.}
\psanum{52}For I remembered thine everlasting judgements, O Lord : and received comfort.\par
\psanum{53}I am horribly afraid : for the ungodly that forsake thy law.\par
\psanum{54}Thy statutes have been my songs : in the house of my pilgrimage.\par
\psanum{55}I have thought upon thy Name, O Lord, in the night-season : and have kept thy law.\par
\psanum{56}This I had : because I kept thy commandments.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Portio mea, Domine}
\lett{T}{hou} art my portion, O Lord : I have promised to keep thy law.\par
\secondline{\psanum{58}I made my humble petition in thy presence with my whole heart : O be merciful unto me, according to thy word.}
\thirdline{\psanum{59}I called mine own ways to remembrance : and turned my feet unto thy testimonies.}
\psanum{60}I made haste, and prolonged not the time : to keep thy commandments.\par
\psanum{61}The congregations of the ungodly have robbed me : but I have not forgotten thy law.\par
\psanum{62}At midnight I will rise to give thanks unto thee : because of thy righteous judgements.\par
\psanum{63}I am a companion of all them that fear thee : and keep thy commandments.\par
\psanum{64}The earth, O Lord, is full of thy mercy : O teach me thy statutes.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subsec{Bonitatem fecisti}
\lett{O}{Lord,} thou hast dealt graciously with thy servant : according unto thy word.\par
\secondline{\psanum{66}O learn me true understanding and knowledge : for I have believed thy commandments.}
\thirdline{\psanum{67}Before I was troubled, I went wrong : but now have I kept thy word.}
\psanum{68}Thou art good and gracious : O teach me thy statutes.\par
\psanum{69}The proud have imagined a lie against me : but I will keep thy commandments with my whole heart.\par
\psanum{70}Their heart is as fat as brawn : but my delight hath been in thy law.\par
\psanum{71}It is good for me that I have been in trouble : that I may learn thy statutes.\par
\psanum{72}The law of thy mouth is dearer unto me : than thousands of gold and silver.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\vspace{-2ex}

\subsec{Manus tuae fecerunt me}
\lett{T}{hy} hands have made me and fashioned me : O give me understanding, that I may learn thy commandments.\par
\secondline{\psanum{74}They that fear thee will be glad when they see me : because I have put my trust in thy word.}
\thirdline{\psanum{75}I know, O Lord, that thy judgements are right : and that thou of very faithfulness hast caused me to be troubled.}
\psanum{76}O let thy merciful kindness be my comfort : according to thy word unto thy servant.\par
\psanum{77}O let thy loving mercies come unto me, that I may live : for thy law is my delight.\par
\psanum{78}Let the proud be confounded, for they go wickedly about to destroy me : but I will be occupied in thy commandments.\par
\psanum{79}Let such as fear thee, and have known thy testimonies : be turned unto me.\par
\psanum{80}O let my heart be sound in thy statutes : that I be not ashamed.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\vspace{-2ex}

\subby{Chapter}
\begin{inhead}
	Feria Days Throughout the Year
\end{inhead}\par\noindent
Heal me, O Lord, and I shall be healed; save me, and I shall be saved: for thou art my praise.

\begin{inhead}
	Sundays \& Feast Days
\end{inhead}\par\noindent
O the depth of the riches both of the wisdom and knowledge of God! How unsearchable are his judgements, and his ways past finding out!

\begin{inhead}
	Eastertide
\end{inhead}\par\noindent
For whatsoever is born of God overcometh the world: and this is the victory that overcometh the world, even our faith.

℟. Thanks be to God.

\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\fakecol{℣. Heal my soul, for I have sinned against thee.}{℣. Sana ánimam meam, Quia peccávi tibi.}

\fakecol{℟. Heal my soul, for I have sinned against thee.}{℟. Sana ánimam meam, Quia peccávi tibi.}

\fakecol{℣. I said, Lord, be merciful unto me.}{℣. Ego dixi: Dómine, miserére mei.}
    
\fakecol{℟. For I have sinned against thee.}{℟. Quia peccávi tibi.}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\fakecol{℟. Heal my soul, for I have sinned against thee.\\}{℟. Sana ánimam meam, Quia peccávi tibi.\\}

\fakecol{℣. Thou hast been my succour.}{℣. Adjútor meus, esto.}

\fakecol{℟. Leave me not, neither forsake me, O God of my salvation.}{℟. Ne derelínquas me, Neque despícias me, Deus, salutáris meus.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\fakecol{℣. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).}{℣.  Inclína cor meum in testimónia tua (allelúja, allelúja).}

\fakecol{℟. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).}{℟. Inclína cor meum in testimónia tua (allelúja, allelúja).}

\fakecol{℣. Turn away mine eyes, lest they behold vanity.}{℣. Avérte óculos meos ne vídeant vanitátem.}
    
\fakecol{℟. Unto thy testimonies (alleluia, alleluia).}{℟. In testimónia tua (allelúja, allelúja).}

\fakecol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\fakecol{℟. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).\\}{℟. Inclína cor meum in testimónia tua (allelúja, allelúja).\\}

\fakecol{℣. I said, Lord, be merciful unto me (alleluia).}{℣. Ego dixi: Dómine, miserére mei (allelúja).}

\fakecol{℟. For I have sinned against thee (alleluia).}{℟. Quia peccávi tibi (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages are said, kneeling, on all Ferias outside of Eastertide.
\end{rubric}

\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\fakecol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\fakecol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.

\fakecol{℣. Send out thy light and thy truth, that they may lead me.}{℣. Emítte lucem tuam et veritátem tuam: ipsa me deduxérunt.}

\fakecol{℟. And bring me unto thy holy hill, and to thy dwelling.}{℟. Et adduxérunt in montem sanctum tuum et in tabernácula tua.}

\fakecol{℣. O let us live.}{℣. Vivificábis nos.}

\fakecol{℟. And we shall call upon thy Name.}{℟. Et nomen tuum invocábimus.}

\fakecol{℣. Turn us again, O Lord God of hosts.}{℣. Dómine, Deus virtútum, convérte nos.}

\fakecol{℟. Show the light of thy countenance, and we shall be whole.}{℟. Et osténde fáciem tuam, et salvi érimus.}

\subby{Collects}
\fakecol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\fakecol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\fakecol{℣. Let us pray.}{℣. Orémus.}

\needspace{4\baselineskip}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}

\lett{H}{elp} us this day, O God, to serve thee devoutly and the world busily. May we do our work wisely, give succour secretly, go to our meat appetitely, sit thereat discreetly, arise temperately, please our friends duly, go to our bed merrily, and sleep surely; for the joy of our Lord Jesus Christ. \textit{Amen.}\\

℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.\par
℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.\par
℣. Let us bless the Lord.\par
℟. Thanks be to God.\par
℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.

\needspace{4\baselineskip}
\lett{U}{nto} him that loved us, and washed us from our sins in his own blood, and hath made us kings and priests unto God and his Father, to him be glory and dominion for ever and ever. \textit{Amen.}


%\elcol{\lett{U}{nto} him that loved us, and washed us from our sins in his own blood, and hath made us kings and priests unto God and his Father, to him be glory and dominion for ever and ever. Amen.}{\lett{Q}{ui} dilexit nos, et lavit nos a peccatis nostris in sanguine suo, 6 et fecit nos regnum, et sacerdotes Deo et Patri suo: ipsi gloria et imperium in s{\ae}cula s{\ae}culorum. Amen.}

\end{multicols}

\begin{center}
	{\textsc{Here endeth the Order of Terce.}}
\end{center}