\fancyhead[RO,LE]{\textit{Prefaces}}
\fancyhead[RE,LO]{}
\section{Prefaces for the Mass}\label{prefaces}

\begin{multicols}{2}
\subsection{Advent}%From the Ritual of the Altar
\lett{T}{hrough} Jesus Christ our Lord; whom thou, the merciful and faithful God, didst promise as a Saviour to lost mankind; by whose truth the unlearned should be instructed, by whose holiness the ungodly should be justified, by whose power the weak should be holpen. Now, therefore, when the time in which thou shouldest send him draweth near, and now that the day of our deliverance already shineth; trusting in these thy promises, with holy joy we do rejoice.

\subsection{Christmas}%1928 BCP:
\noindent
\lett{B}{ecause} thou didst give Jesus Christ, thine only Son, to be born (as at this time) for us; who, by the operation of the Holy Ghost, was made very man, of the substance of the Virgin Mary his mother; and that without spot of sin, to make us clean from all sin.

\subsection{Epiphany}%1928 BCP:
\noindent
\lett{T}{hrough} Jesus Christ our Lord, who, in substance of our mortal flesh, manifested forth his glory; that he might bring us out of darkness into his own glorious light.

\subsection{Lent}%1933 English Missal:
\noindent
\lett{W}{ho} by bodily fasting dost overcome vice, dost raise the mind, and dost bestow on us virtue and heavenly rewards: through Christ, our Lord.

\subsection{Holy Rood}%1933 English Missal:
\noindent
\lett{W}{ho} by the tree of the Cross didst give salvation unto mankind: that whence death arose, thence life might rise again: and that he, who by a tree overcame, might also by a tree be overcome: through Christ our Lord.

\subsection{Easter} %1928 BCP:
\noindent
\lett{B}{ut} chiefly are we bound to praise thee for the glorious Resurrection of thy Son Jesus Christ our Lord: for he is the very Paschal Lamb, which was offered for us, and hath taken away the sin of the world; who by his death hath destroyed death, and by his rising to life again hath restored to us everlasting life.

\subsection{Ascension}%1928 BCP:
\noindent
\lett{T}{hrough} thy most dearly beloved Son Jesus Christ our Lord; who, after his most glorious Resurrection, manifestly appeared to all his Apostles, and in their sight ascended up into heaven, to prepare a place for us; that where he is, thither we might also ascend, and reign with him in glory.

\subsection{Holy Ghost}%1928 BCP Whitsun Preface:
\noindent
\lett{T}{hrough} Jesus Christ our Lord; according to whose most true promise, the Holy Ghost came down (as at this time) from heaven, lighting upon the Apostles, to teach them, and to lead them into all truth; giving them boldness with fervent zeal constantly to preach the Gospel unto all nations; whereby we have been brought out of darkness and error into the clear light and true knowledge of thee, and of thy Son Jesus Christ.

\subsection{Most Holy Trinity}%1928 BCP:
\noindent
\lett{W}{ho,} with thine only-begotten Son, and the Holy Ghost, art one God, one Lord, in Trinity of Persons and in Unity of Substance. For that which we believe of thy glory, O Father, the same we believe of the Son, and of the Holy Ghost, without any difference of inequality.

\subsection{Compassion of Our Lord}%1933 English Missal:
\noindent
\lett{W}{ho} didst will that thine only-begotten Son, as he hung upon the Cross, should be pierced by the soldier's spear: that his Heart thus opened, the shrine of the divine bounty, might pour forth upon us streams of mercy and of grace: that as it never ceased to burn with love for us, so it might be a rest for the godly, and an open refuge of salvation for the penitent.

\subsection{Christ the King}%1933 English Missal:
\noindent
\lett{W}{ho} didst anoint thine only-begotten Son, our Lord Jesus Christ, with the oil of gladness, to be a Priest for ever and the King of all the world: that, offering himself an unspotted sacrifice of peace upon the altar of the cross, he might accomplish the sacrament of the redemption of mankind: and making all creatures subject to his governance, might deliver up to thine infinite Majesty an eternal and universal kingdom. A kingdom of truth and life: a kingdom of sanctity and grace: a kingdom of justice, love, and peace.

\subsection{All Saints}%1928 BCP:
\noindent
\lett{W}{ho,} in the multitude of thy saints, hast compassed us about with so great a cloud of witnesses that we, rejoicing in their fellowship, may run with patience the race that is set before us, and together with them may receive the crown of glory that fadeth not away.

%MANUAL ADJUSTMENT:
\needspace{6\baselineskip}
\subsection{Of Our Lord}
\noindent
\lett{B}{ecause} in the Mystery of the Word made flesh, thou hast caused a new light to shine in our hearts, to give the knowledge of thy glory in the face of thy Son, Jesus Christ our Lord.

\subsection{Blessed Virgin Mary}
\noindent
\lett{A}{nd} that in the \textit{N.} of Blessed Mary ever Virgin we should praise, bless, and magnify thee. In that by the overshadowing of the Holy Ghost she conceived thine only-begotten Son: and, the glory of her maidenhood yet abiding, shed forth upon the world the light eternal, Jesus Christ, our Lord.

\subsection{St. Joseph, Spouse of the Blessed Virgin}
\noindent
\lett{A}{nd} that in the Festivity (Veneration) of blessed Joseph, we should magnify, bless, and glorify thee with worthy praises. Who being a just man was given by thee for a Spouse to the Virgin Mother of God: and, being a faithful and wise servant, was made ruler over thy Household: that in the office of a father he should guard thine only-begotten Son, conceived by the overshadowing of the Holy Ghost, even Jesus Christ our Lord.

\subsection{Apostles \& Evangelists}
\noindent
\lett{T}{hat} we, O Lord, should humbly entreat thee, that thou, the everlasting Shepherd, wouldst not forsake thy flock: but through thy blessed Apostles keep it by thy continual protection. That it may be governed by those same rulers, whom in thy stead thou hast appointed for thy work, as shepherds of thy people.

\subsection{Requiem}
\noindent
\lett{T}{hrough} Christ, our Lord. In whom hath shone forth unto us the hope of a blessed resurrection, that they who bewail the certain condition of their mortality may be consoled by the promise of immortality to come. For to thy faithful people, O Lord, life is changed not taken away: and at the dissolution of the tabernacle of this earthly sojourning a dwelling-place eternal is made ready in the heavens.
\end{multicols}