   \fancyhead[C]{{\LARGE Proper of Season}}
	\fancyhead[RO,LE]{}
	\fancyhead[RE,LO]{}
	
\phantomsection
\addcontentsline{toc}{section}{Proper of the Season}

   \subby{First Sunday of Advent}\label{AdventI}
   \feastday{Advent I}
\begin{inhead}
    {First Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{U}{nto} thee lift I up my soul: my God, in thee have I trusted, let me not be confounded: neither let mine enemies triumph over me: for all they that look for thee shall not be ashamed. \textit{Ps.} Shew me thy ways, O Lord; and teach me thy paths.
\collect\label{AdventICollect}
\lett{A}{lmighty} God, give us grace that we may cast away the works of darkness, and put upon us the armour of light, now in the time of this mortal life, in which thy Son Jesus Christ came to visit us in great humility; that in the last day, when he shall come again in his glorious majesty to judge both the quick and the dead, we may rise to the life immortal, through him who liveth and reigneth with thee and the Holy Ghost, now and ever. \textit{Amen.}
\begin{rubric}
    This Collect is to be repeated every day, with the other Collects in Advent, until Christmas Day.
\end{rubric}
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 13:8}
\lett{B}{rethren:} Owe no man any thing, but to love one another: for he that loveth another hath fulfilled the law. For this, Thou shalt not commit adultery, Thou shalt not kill, Thou shalt not steal, Thou shalt not bear false witness, Thou shalt not covet; and if there be any other commandment, it is briefly comprehended in this saying, namely, Thou shalt love thy neighbour as thyself. Love worketh no ill to his neighbour: therefore love is the fulfilling of the law. And that, knowing the time, that now it is high time to awake out of sleep: for now is our salvation nearer than when we believed. The night is far spent, the day is at hand: let us therefore cast off the works of darkness, and let us put on the armour of light. Let us walk honestly, as in the day; not in rioting and drunkenness, not in chambering and wantonness, not in strife and envying. But put ye on the Lord Jesus Christ, and make not provision for the flesh, to fulfil the lusts thereof.
\gradall{All they that look for thee shall not be ashamed, O Lord. ℣. Shew me thy ways, O Lord; and teach me thy paths.}{Alleluia, alleluia. ℣. Shew us thy mercy, O Lord: and grant us thy salvation. Alleluia.}
\begin{rubric}
    On Ferias in Advent, when during the week the Sunday Mass is resumed, Alleluia is not said, but only the Gradual.
\end{rubric}

\readingcitation{Gospel}{Matthew 21:1}
\lett{A}{t that time:} When they drew nigh unto Jerusalem, and were come to Bethphage, unto the mount of Olives, then sent Jesus two disciples, Saying unto them, Go into the village over against you, and straightway ye shall find an ass tied, and a colt with her: loose them, and bring them unto me. And if any man say ought unto you, ye shall say, The Lord hath need of them; and straightway he will send them. All this was done, that it might be fulfilled which was spoken by the prophet, saying, Tell ye the daughter of Sion, Behold, thy King cometh unto thee, meek, and sitting upon an ass, and a colt the foal of an ass. And the disciples went, and did as Jesus commanded them, And brought the ass, and the colt, and put on them their clothes, and they set him thereon. And a very great multitude spread their garments in the way; others cut down branches from the trees, and strawed them in the way. And the multitudes that went before, and that followed, cried, saying, Hosanna to the Son of David: Blessed is he that cometh in the name of the Lord; Hosanna in the highest. And when he was come into Jerusalem, all the city was moved, saying, Who is this? And the multitude said, This is Jesus the prophet of Nazareth of Galilee. And Jesus went into the temple of God, and cast out all them that sold and bought in the temple, and overthrew the tables of the moneychangers, and the seats of them that sold doves, And said unto them, It is written, My house shall be called the house of prayer; but ye have made it a den of thieves.

\offertory{Unto thee lift I up my soul: my God in thee have I trusted, let me not be confounded: neither let mine enemies triumph over me: for all they that look for thee shall not be ashamed.}
%MANUAL ADJUSTMENT:
\clearpage
\secret\label{AdventISecret}
\lett{M}{ay} these sacred mysteries, O Lord, so cleanse us by their mighty power, that growing in purity, we may attain unto thee who art the author of the same. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{The Lord shall shew loving-kindness; and our land shall give her increase.}
\postcommunion\label{AdventIPostcommunion}
\lett{G}{rant,} O Lord, that we may so wait for thy loving-kindness in the midst of thy temple; that we may with worthy reverence prevent the coming festival of our redemption. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Second Sunday of Advent}
\feastday{Advent II}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{people} of Sion, behold the Lord is nigh at hand to redeem the nations: and in the gladness of your heart the Lord shall cause his glorious voice to be heard. \textit{Ps.} Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep.
\collect
\lett{B}{lessed} Lord, who hast caused all holy Scriptures to be written for our learning; Grant that we may in such wise hear them, read, mark, learn, and inwardly digest them, that by patience and comfort of thy holy Word, we may embrace, and ever hold fast, the blessed hope of everlasting life, which thou hast given us in our Saviour Jesus Christ. Who liveth.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 15:4}
\lett{B}{rethren:} Whatsoever things were written aforetime were written for our learning, that we through patience and comfort of the scriptures might have hope. Now the God of patience and consolation grant you to be likeminded one toward another according to Christ Jesus: That ye may with one mind and one mouth glorify God, even the Father of our Lord Jesus Christ. Wherefore receive ye one another, as Christ also received us to the glory of God. Now I say that Jesus Christ was a minister of the circumcision for the truth of God, to confirm the promises made unto the fathers: And that the Gentiles might glorify God for his mercy; as it is written, For this cause I will confess to thee among the Gentiles, and sing unto thy name. And again he saith, Rejoice, ye Gentiles, with his people. And again, Praise the Lord, all ye Gentiles; and laud him, all ye people. And again, Esaias saith, There shall be a root of Jesse, and he that shall rise to reign over the Gentiles; in him shall the Gentiles trust. Now the God of hope fill you with all joy and peace in believing, that ye may abound in hope, through the power of the Holy Ghost.

%MANUAL ADJUSTMENT:
\gradual{Out of Sion hath God appeared; in perfect beauty, ℣. Gather my saints together unto me, those that have made a covenant with me with sacrifice.}

\alleluia{Alleluia, alleluia. ℣. I was glad when they said unto me: we will go into the house of the Lord. Alleluia.}

\vspace{-1ex}

\readingcitation{Gospel}{Luke 21:25}
\lett{A}{t that time:} Jesus said unto his disciples: There shall be signs in the sun, and in the moon, and in the stars; and upon the earth distress of nations, with perplexity; the sea and the waves roaring; Men's hearts failing them for fear, and for looking after those things which are coming on the earth: for the powers of heaven shall be shaken. And then shall they see the Son of man coming in a cloud with power and great glory. And when these things begin to come to pass, then look up, and lift up your heads; for your redemption draweth nigh. And he spake to them a parable; Behold the fig tree, and all the trees; When they now shoot forth, ye see and know of your own selves that summer is now nigh at hand. So likewise ye, when ye see these things come to pass, know ye that the kingdom of God is nigh at hand. Verily I say unto you, This generation shall not pass away, till all be fulfilled. Heaven and earth shall pass away: but my words shall not pass away.

\offertory{Wilt not thou turn again, O God, and quicken us, that thy people may rejoice in thee; shew us thy mercy, O Lord, and grant us thy salvation.}
\secret
\lett{W}{e} beseech thee, O Lord, mercifully to accept the prayers and sacrifices of thy humble servants: that we, who cannot of our own prayers and merits help ourselves may be defended by thy succour. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\communion{Arise, O Jerusalem, and stand on high, and behold the joy that cometh unto thee from thy God.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\postcommunion
\lett{O}{Lord,} who hast satisfied us with spiritual food and sustenance, we humbly beseech thee: that through the partaking of this mystery, we may learn to despise things earthly, and to love the things that are heavenly. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Third Sunday of Advent}
\feastday{Advent III}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{R}{ejoice} in the Lord alway: and again I say, rejoice. Let your moderation be known unto all men: for the Lord is at hand. Be careful for nothing; but in all things by prayer let your requests be made known unto God. \textit{Ps.} Lord, thou art become gracious unto thy land; thou hast turned away the captivity of Jacob.

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\collect
\lett{O}{Lord} Jesus Christ, who at thy first coming didst send thy messenger to prepare thy way before thee; Grant that the ministers and stewards of thy mysteries may likewise so prepare and make ready thy way, by turning the hearts of the disobedient to the wisdom of the just, that at thy second coming to judge the world we may be found an acceptable people in thy sight, who livest and reignest with the Father and the Holy Spirit ever one God, world without end. \textit{Amen.}
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{1 Corinthians 4:1}
\lett{B}{rethren:} Let a man so account of us, as of the ministers of Christ, and stewards of the mysteries of God. Moreover it is required in stewards, that a man be found faithful. But with me it is a very small thing that I should be judged of you, or of man's judgment: yea, I judge not mine own self. For I know nothing by myself; yet am I not hereby justified: but he that judgeth me is the Lord. Therefore judge nothing before the time, until the Lord come, who both will bring to light the hidden things of darkness, and will make manifest the counsels of the hearts: and then shall every man have praise of God.

\gradall{O Lord, thou that sittest upon the Cherubim, stir up thy strength and come. ℣. Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep.}{Alleluia, alleluia. ℣. Stir up thy strength, O Lord, and come and help us. Alleluia.}

\readingcitation{Gospel}{Matthew 11:2}
\lett{A}{t that time:} When John had heard in the prison the works of Christ, he sent two of his disciples, And said unto him, Art thou he that should come, or do we look for another? Jesus answered and said unto them, Go and shew John again those things which ye do hear and see: The blind receive their sight, and the lame walk, the lepers are cleansed, and the deaf hear, the dead are raised up, and the poor have the gospel preached to them. And blessed is he, whosoever shall not be offended in me. And as they departed, Jesus began to say unto the multitudes concerning John, What went ye out into the wilderness to see? A reed shaken with the wind? But what went ye out for to see? A man clothed in soft raiment? behold, they that wear soft clothing are in kings' houses. But what went ye out for to see? A prophet? yea, I say unto you, and more than a prophet. For this is he, of whom it is written, Behold, I send my messenger before thy face, which shall prepare thy way before thee.

\offertory{O Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob: thou hast forgiven the offence of thy people.}
\secret
\lett{W}{e} beseech thee, O Lord, that the continual offering of this sacrifice of our devotion may fulfil thine institution of this sacred mystery and may accomplish in us the wondrous work of thy salvation. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{Say to them that are of a fearful heart: be strong, fear not: behold, our God will come and save us.}

\postcommunion
\lett{W}{e} entreat, O Lord, thy mercy: that these means of heavenly grace may so cleanse us from our sins, that we may be made ready for thy coming festival. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Ember Wednesday in Advent}
\feastday{Advent Emberday}
\begin{inhead}
{Greater Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Wednesday}

\properantiphonfix

\introit
\lett{D}{rop} down, ye heavens, from above, and let the skies pour down justice: let the earth open, and bring forth a Saviour. \textit{Ps.} The heavens declare the glory of God: and the firmament sheweth his handywork.
\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that the coming solemnity of our redemption may bestow on us thy succour in this present life, and win for us the rewards of everlasting blessedness. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Isaiah 2:2}
\lett{I}{n those days:} Said Isaiah the Prophet: And it shall come to pass in the last days, that the mountain of the \divineName{Lord}'s house shall be established in the top of the mountains, and shall be exalted above the hills; and all nations shall flow unto it. And many people shall go and say, Come ye, and let us go up to the mountain of the \divineName{Lord}, to the house of the God of Jacob; and he will teach us of his ways, and we will walk in his paths: for out of Zion shall go forth the law, and the word of the \divineName{Lord} from Jerusalem. And he shall judge among the nations, and shall rebuke many people: and they shall beat their swords into plowshares, and their spears into pruninghooks: nation shall not lift up sword against nation, neither shall they learn war any more. O house of Jacob, come ye, and let us walk in the light of the \divineName{Lord} our God. %`Our God' added according to English Missal
\gradual{Lift up your heads, O ye gates: and be ye lift up, ye everlasting doors: and the King of glory shall come in, ℣. Who shall ascend into the hill of the Lord? or who shall rise up in his holy place? Even he that hath clean hands and a pure heart.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\collect
\lett{O}{Lord,} make haste, we pray thee, and delay not to bestow upon us the assistance of thy heavenly might: that they who put their confidence in thy mercy may be succoured by the consolations of thine advent. Who livest.

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\readingcitation{Epistle}{Isaiah 7:10}
\lett{I}{n those days:} The \divineName{Lord} spake unto Ahaz, saying, Ask thee a sign of the \divineName{Lord} thy God; ask it either in the depth, or in the height above. But Ahaz said, I will not ask, neither will I tempt the \divineName{Lord}. And he said, Hear ye now, O house of David; Is it a small thing for you to weary men, but will ye weary my God also? Therefore the Lord himself shall give you a sign; Behold, a virgin shall conceive, and bear a son, and shall call his name Immanuel. Butter and honey shall he eat, that he may know to refuse the evil, and choose the good.

\gradual{The Lord is nigh unto all them that call upon him: yea all such as call upon him faithfully. ℣. My mouth shall speak the praise of the Lord: and let all flesh give thanks unto his holy name.}
\readingcitation{Gospel}{Luke 1:26}
\lett{A}{t that time:} The angel Gabriel was sent from God unto a city of Galilee, named Nazareth, To a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin's name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name \textsc{Jesus}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: And he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word.
\offertory{Be strong, fear not: behold, our God will come with vengeance, even God with a recompence: he will come and save us.}

\secret
\lett{W}{e} beseech thee, O Lord, that these our fasts may be acceptable in thy sight: that we, being cleansed thereby and made worthy of thy grace, may be brought unto thine everlasting promises. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{Behold, a Virgin shall conceive, and bear a son: and shall call his name Immanuel.}

%MANUAL ADJUSTMENT:
\clearpage

\postcommunion
\lett{O}{Lord,} who hast satisfied us with the gift of thy salvation, we humbly beseech thee: that as we joyfully taste thereof, so we may effectually be renewed by the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Ember Friday in Advent}
\feastday{Advent Emberday}
\begin{inhead}
{Greater Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Friday}

\properantiphonfix

\introit
\lett{B}{e} thou nigh at hand, O Lord, for all thy commandments are true: as concerning thy testimonies, I have known long since, that thou art from everlasting. \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.
\collect
\lett{S}{tir} up, we beseech thee, O Lord, thy power, and come among us: that they who put their confidence in thy goodness may speedily be delivered from all adversity. Who livest.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Isaiah 11:1}
\lett{T}{hus} saith the Lord God: There shall come forth a rod out of the stem of Jesse, and a Branch shall grow out of his roots: And the spirit of the \divineName{Lord} shall rest upon him, the spirit of wisdom and understanding, the spirit of counsel and might, the spirit of knowledge and of the fear of the \divineName{Lord}; And shall make him of quick understanding in the fear of the \divineName{Lord}: and he shall not judge after the sight of his eyes, neither reprove after the hearing of his ears: But with righteousness shall he judge the poor, and reprove with equity for the meek of the earth: and he shall smite the earth with the rod of his mouth, and with the breath of his lips shall he slay the wicked. And righteousness shall be the girdle of his loins, and faithfulness the girdle of his reins.
\gradual{Shew us thy mercy, O Lord: and grant us thy salvation. ℣. Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob.}
\readingcitation{Gospel}{Luke 1:39}
\lett{A}{t that time:} Mary arose, and went into the hill country with haste, into a city of Juda; And entered into the house of Zacharias, and saluted Elisabeth. And it came to pass, that, when Elisabeth heard the salutation of Mary, the babe leaped in her womb; and Elisabeth was filled with the Holy Ghost: And she spake out with a loud voice, and said, Blessed art thou among women, and blessed is the fruit of thy womb. And whence is this to me, that the mother of my Lord should come to me? For, lo, as soon as the voice of thy salutation sounded in mine ears, the babe leaped in my womb for joy. And blessed is she that believed: for there shall be a performance of those things which were told her from the Lord. And Mary said, My soul doth magnify the Lord, And my spirit hath rejoiced in God my Saviour.
\offertory{Wilt thou not turn again, O God, and quicken us, that thy people may rejoice in thee: shew us thy mercy, O Lord, and grant us thy salvation.}
\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us, whom thou dost cleanse by thy heavenly mysteries. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{Behold, the Lord shall come, and all his Saints with him: and in that day there shall be a great light.}
\postcommunion
\lett{M}{ay} the holy partaking of thy sacrament, O Lord, in such wise restore us: that we, being cleansed from our former nature, may enter into the fellowship of this saving mystery. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Ember Saturday in Advent}
\feastday{Advent Emberday}
\begin{inhead}
{Greater Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Saturday}

\properantiphonfix

\introit
\lett{C}{ome} and shew us the light of thy countenance, O Lord, that sittest upon the Cherubim: and we shall be whole. \textit{Ps.} Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep.
\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{O}{God,} who seest us that we are afflicted by reason of our iniquity: mercifully grant; that we may be comforted by thy visitation. Who livest.

\readingcitation{Epistle}{Isaiah 19:20}
\lett{I}{n those days:} They shall cry unto the \divineName{Lord} because of the oppressors, and he shall send them a saviour, and a great one, and he shall deliver them. And the \divineName{Lord} shall be known to Egypt, and the Egyptians shall know the \divineName{Lord} in that day, and shall do sacrifice and oblation; yea, they shall vow a vow unto the \divineName{Lord}, and perform it. And the \divineName{Lord} shall smite Egypt: he shall smite and heal it: and they shall return even to the \divineName{Lord}, and he shall be intreated of them, and shall heal them,
%Added according to EM.
the Lord our God.

\gradual{He goeth forth from the uttermost part of the heaven; and runneth about unto the end of it again. ℣. The heavens declare the glory of God; and the firmament sheweth his handy-work.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that, whereas through our ancient bondage we are bowed down beneath the yoke of sin; the new birth of thine only-begotten Son which we await may be for our deliverance. Who liveth.
\readingcitation{Epistle}{Isaiah 35:1}
\lett{T}{hus saith the Lord:} The wilderness and the solitary place shall be glad for them; and the desert shall rejoice, and blossom as the rose. It shall blossom abundantly, and rejoice even with joy and singing: the glory of Lebanon shall be given unto it, the excellency of Carmel and Sharon, they shall see the glory of the \divineName{Lord}, and the excellency of our God. Strengthen ye the weak hands, and confirm the feeble knees. Say to them that are of a fearful heart, Be strong, fear not: behold, your God will come with vengeance, even God with a recompence; he will come and save you. Then the eyes of the blind shall be opened, and the ears of the deaf shall be unstopped. Then shall the lame man leap as an hart, and the tongue of the dumb sing: for in the wilderness shall waters break out, and streams in the desert. And the parched ground shall become a pool, and the thirsty land springs of water: %Added according to EM.
saith the Lord almighty.

\gradual{He hath set his tabernacle in the sun: and he cometh forth as a bridegroom out of his chamber. ℣. He goeth forth from the uttermost part of the heaven: and runneth about unto the end of it again.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{O}{Lord,} we pray thee, that we thine unworthy servants, who now bewail the guilt of our misdeeds, may be gladdened by the advent of thy only-begotten Son. Who liveth.
\readingcitation{Epistle}{Isaiah 40:9}
\lett{T}{hus saith the Lord:} O Zion, that bringest good tidings, get thee up into the high mountain; O Jerusalem, that bringest good tidings, lift up thy voice with strength; lift it up, be not afraid; say unto the cities of Judah, Behold your God! Behold, the Lord \divineName{God} will come with strong hand, and his arm shall rule for him: behold, his reward is with him, and his work before him. He shall feed his flock like a shepherd: he shall gather the lambs with his arm, and carry them in his bosom,
%Added according to EM.
the Lord our God.
\gradual{Turn us again, O Lord God of hosts: shew the light of thy countenance, and we shall be whole. ℣. Stir up thy strength, O Lord, and come, and save us.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that the coming solemnity of thy Son may bestow upon us healing in this present life, and obtain for us everlasting rewards. Through the same.
\readingcitation{Epistle}{Isaiah 45:1}
\lett{T}{hus} saith the \divineName{Lord} to his anointed, to Cyrus, whose right hand I have holden, to subdue nations before him; and I will loose the loins of kings, to open before him the two leaved gates; and the gates shall not be shut; I will go before thee, and make the crooked places straight: I will break in pieces the gates of brass, and cut in sunder the bars of iron: And I will give thee the treasures of darkness, and hidden riches of secret places, that thou mayest know that I, the \divineName{Lord}, which call thee by thy name, am the God of Israel. For Jacob my servant's sake, and Israel mine elect, I have even called thee by thy name: I have surnamed thee, though thou hast not known me. I am the \divineName{Lord}, and there is none else, there is no God beside me: I girded thee, though thou hast not known me: That they may know from the rising of the sun, and from the west, that there is none beside me. I am the \divineName{Lord}, and there is none else. I form the light, and create darkness: I make peace, and create evil: I the \divineName{Lord} do all these things. Drop down, ye heavens, from above, and let the skies pour down righteousness: let the earth open, and let them bring forth salvation, and let righteousness spring up together; I the \divineName{Lord} have created it.

\gradual{Stir up, O Lord, thy strength, and come, and save us. ℣. Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep: shew thyself, thou that sittest upon the Cherubim, before Ephraim, Benjamin, and Manasses.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{O}{Lord,} we beseech thee favourably to hear the prayers of thy people: that we, who are justly punished for our offences, may be comforted by the visitation of thy mercy. Who livest.
\readingcitation{Epistle}{Song of the Three Young Men 26}
\lett{I}{n those days:} The angel of the Lord came down into the oven together with Azarias and his fellows, and smote the flame of the fire out of the oven; And made the midst of the furnace as it had been a moist whistling wind, so that the fire touched them not at all, neither hurt nor troubled them. Then the three, as out of one mouth, praised, glorified, and blessed, God in the furnace, saying,
\begin{rubric}
    Here the response \emph{Thanks be to God}, is not made.
\end{rubric}
\subbysub{Hymn}
\lett{B}{lessed} art thou, O Lord God of our fathers. And worthy to be praised and glorious for ever.\par
\secondline{And blessed is the name of thy glory, which is holy. And worthy to be praised and glorious for ever.}
\thirdline{Blessed art thou in the holy temple of thy glory. And worthy to be praised and glorious for ever.}

Blessed art thou on the holy throne of thy kingdom. And worthy to be praised and glorious for ever.

Blessed art thou in the sceptre of thy Godhead. And worthy to be praised and glorious for ever.

Blessed art thou that sittest upon the Cherubim, and beholdest the depths. And worthy to be praised and glorious for ever.

Blessed art thou that walkest on the wings of the winds and on the waves of the sea. And worthy to be praised and glorious for ever.

Let all thine Angels and Saints bless thee. And let them praise thee and glorify thee for ever.

Let the heavens, the earth, the sea, and all that in them is bless thee. And let them praise thee and glorify thee for ever.

Glory be to the Father, and to the Son, and to the Holy Ghost. And worthy to be praised and glorious for ever.

As it was in the beginning, is now, and ever shall be: world without end. Amen. And worthy to be praised and glorious for ever.

Blessed art thou, O Lord God of our fathers. And worthy to be praised and glorious for ever.
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\collect
\lett{O}{God,} who to the three children didst assuage the flames of fire: mercifully grant; that the flames of sin may not kindle upon us thy servants. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{2 Thessalonians 2:1}
\lett{B}{rethren:} We beseech you, by the coming of our Lord Jesus Christ, and by our gathering together unto him, That ye be not soon shaken in mind, or be troubled, neither by spirit, nor by word, nor by letter as from us, as that the day of Christ is at hand. Let no man deceive you by any means: for that day shall not come, except there come a falling away first, and that man of sin be revealed, the son of perdition; Who opposeth and exalteth himself above all that is called God, or that is worshipped; so that he as God sitteth in the temple of God, shewing himself that he is God. Remember ye not, that, when I was yet with you, I told you these things? And now ye know what withholdeth that he might be revealed in his time. For the mystery of iniquity doth already work: only he who now letteth will let, until he be taken out of the way. And then shall that Wicked be revealed, whom the Lord shall consume with the spirit of his mouth, and shall destroy with the brightness of his coming.
\tract{Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep, ℣. Thou that sittest upon the Cherubim, shew thyself before Ephraim, Benjamin, and Manasses. ℣. Stir up thy strength, O Lord, and come, and save us.}
\readingcitation{Gospel}{Luke 3:1}
\lett{I}{n} the fifteenth year of the reign of Tiberius Caesar, Pontius Pilate being governor of Judaea, and Herod being tetrarch of Galilee, and his brother Philip tetrarch of Ituraea and of the region of Trachonitis, and Lysanias the tetrarch of Abilene, Annas and Caiaphas being the high priests, the word of God came unto John the son of Zacharias in the wilderness. And he came into all the country about Jordan, preaching the baptism of repentance for the remission of sins; As it is written in the book of the words of Esaias the prophet, saying, The voice of one crying in the wilderness, Prepare ye the way of the Lord, make his paths straight. Every valley shall be filled, and every mountain and hill shall be brought low; and the crooked shall be made straight, and the rough ways shall be made smooth; And all flesh shall see the salvation of God.
\offertory{Rejoice greatly, O daughter of Zion, shout aloud, O daughter of Jerusalem: behold, thy King cometh to thee, he is just and having salvation.}
\secret
\lett{W}{e} beseech thee, O Lord, mercifully to have respect unto these our sacrifices: that they may increase our devotion, and set forward our salvation. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{He rejoiceth as a giant to run his course: he goeth forth from the uttermost part of the heaven, and runneth about unto the end of it again.}
\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that these holy mysteries which thou hast given to us for the assurance of our redemption; may be for our healing both in this life and in that which is to come. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Fourth Sunday of Advent}
\feastday{Advent IV}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}
\fancyhead[RE,LO]{}

\properantiphonfix

\introit
\lett{D}{rop} down, ye heavens, from above, and let the skies pour down justice: let the earth open, and bring forth a Saviour. \textit{Ps.} The heavens declare the glory of God: and the firmament sheweth his handywork.

%MANUAL ADJUSTMENT:
\subsubsection{Collect}
\lett{O}{Lord,} raise up, we pray thee, thy power, and come among us, and with great might succour us; that whereas, through our sins and wickedness, we are sore let and hindered in running the race that is set before us, thy bountiful grace and mercy may speedily help and deliver us; through the satisfaction of thy Son our Lord, to whom, with thee and the Holy Ghost, be honour and glory, world without end. \textit{Amen.}
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{Philippians 4:4}
\lett{B}{rethren:} Rejoice in the Lord alway: and again I say, Rejoice. Let your moderation be known unto all men. The Lord is at hand. Be careful for nothing; but in every thing by prayer and supplication with thanksgiving let your requests be made known unto God. And the peace of God, which passeth all understanding, shall keep your hearts and minds through Christ Jesus our Lord.

\gradual{The Lord is nigh unto all them that call upon him: yea, all such as call upon him faithfully. ℣. My mouth shall speak the praise of the Lord: and let all flesh give thanks unto his holy name.}

\alleluia{Alleluia, alleluia. ℣. Come, O Lord, and tarry not: forgive the misdeeds of thy people Israel. Alleluia.}

\readingcitation{Gospel}{John 1:19}
\lett{T}{his is the record of John:} The Jews sent priests and Levites from Jerusalem to ask him, Who art thou? And he confessed, and denied not; but confessed, I am not the Christ. And they asked him, What then? Art thou Elias? And he saith, I am not. Art thou that prophet? And he answered, No. Then said they unto him, Who art thou? that we may give an answer to them that sent us. What sayest thou of thyself? He said, I am the voice of one crying in the wilderness, Make straight the way of the Lord, as said the prophet Esaias. And they which were sent were of the Pharisees. And they asked him, and said unto him, Why baptizest thou then, if thou be not that Christ, nor Elias, neither that prophet? John answered them, saying, I baptize with water: but there standeth one among you, whom ye know not; He it is, who coming after me is preferred before me, whose shoe's latchet I am not worthy to unloose. These things were done in Bethabara beyond Jordan, where John was baptizing.

\offertory{Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women, and blessed is the fruit of thy womb.}

\secret
\lett{W}{e} beseech thee, O Lord, mercifully to have respect unto these our sacrifices: that they may increase our devotion and set forward our salvation. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{Behold, a Virgin shall conceive, and bear a son: and his name shall be called Immanuel.}

\postcommunion
\lett{W}{e} beseech thee, O Lord: that as we have now received thy gifts, so, continually drawing near to this mystery, we may set forward the work of our salvation. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Vigil of the Nativity of Our Lord}
\feastday{Christmas Eve}
\begin{inhead}
{First Class Vigil}
\end{inhead}
\par\noindent
\textit{Opening Sentence.} To-morrow shall the wickedness of the earth be done away. And the Saviour of the world shall be King over us.\par

\begin{rubric}
    If the Vigil of the Nativity fall on a Sunday, the whole Office of the Mass is of the Vigil, with commemoration of the Sunday, but its Gospel is not read at the end.
\end{rubric}

\introit
\lett{T}{o-day} shall ye know that the Lord will come to deliver us: and in the morning shall ye behold his glory. \textit{Ps.} The earth is the Lord's, and all that therein is; the compass of the world, and they that dwell therein.
\collect
\lett{O}{God,} who makest us glad with the yearly expectation of our redemption: vouchsafe; that as we joyfully receive thine Only-begotten Son for our Redeemer, so we may with sure confidence behold him when he shall come to be our judge, even Jesus Christ thy Son our Lord. Who liveth.
\readingcitation{Epistle}{Romans 1:1}
\lett{P}{aul,} a servant of Jesus Christ, called to be an apostle, separated unto the gospel of God, (Which he had promised afore by his prophets in the holy scriptures,) Concerning his Son Jesus Christ our Lord, which was made of the seed of David according to the flesh; And declared to be the Son of God with power, according to the spirit of holiness, by the resurrection from the dead: By whom we have received grace and apostleship, for obedience to the faith among all nations, for his name: Among whom are ye also the called of Jesus Christ our Lord.

\gradual{To-day shall ye know that the Lord will come to deliver us: and in the morning shall ye behold his glory. ℣. Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep: thou that sittest upon the Cherubim, shew thyself before Ephraim, Benjamin, and Manasses.}
\begin{rubric}
The Alleluia Verse is not said unless the Vigil fall on a Sunday.
\end{rubric}
\alleluia{Alleluia, alleluia. ℣. On the morrow the iniquity of the earth shall be blotted out: and the Saviour of the world shall reign over us. Alleluia.}
\readingcitation{Gospel}{Matthew 1:18}
\lett{W}{hen} as Mary the Mother of Jesus was espoused to Joseph, before they came together, she was found with child of the Holy Ghost. Then Joseph her husband, being a just man, and not willing to make her a publick example, was minded to put her away privily. But while he thought on these things, behold, the angel of the Lord appeared unto him in a dream, saying, Joseph, thou son of David, fear not to take unto thee Mary thy wife: for that which is conceived in her is of the Holy Ghost. And she shall bring forth a son, and thou shalt call his name \textsc{Jesus}: for he shall save his people from their sins.
\offertory{Lift up your heads, O ye gates: and be ye lift up, ye everlasting doors, and the King of glory shall come in.}
\secret
\lett{G}{rant} to us, we beseech thee, almighty God: that like as we do prevent the feast of the adorable Birth of thy Son, so we may joyfully receive his everlasting gifts. Who with thee liveth.
\communion{The glory of the Lord shall be revealed: and all flesh shall see the salvation of our God.}
\postcommunion
\lett{G}{rant} to us, we beseech thee, O Lord: that like as we do eat and drink of the heavenly mystery of thine only begotten Son; so we may be renewed by the remembrance of his birth. Through the same.


\subby{Nativity of Our Lord}
\feastday{Christmas}
\begin{inhead}
{First Class Double, Third Class Octave}
\end{inhead}
\centerline{\small{(The First Mass at Night)}}

\introit
\lett{T}{he} Lord hath said unto me: Thou art my Son, this day have I begotten thee. \textit{Ps.} Why do the heathen so furiously rage together: and why do the people imagine a vain thing?
\collect
\lett{O}{God,} who makest us glad with the yearly remembrance of the birth of thine only Son Jesus Christ; Grant that as we joyfully receive him for our Redeemer, so we may with sure confidence behold him when he shall come to be our Judge, who liveth and reigneth with thee and the Holy Ghost, one God, world without end. \textit{Amen.}
\readingcitation{Epistle}{Titus 2:11}
\lett{D}{early beloved:} The grace of God that bringeth salvation hath appeared to all men, Teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly, in this present world; Looking for that blessed hope, and the glorious appearing of the great God and our Saviour Jesus Christ; Who gave himself for us, that he might redeem us from all iniquity, and purify unto himself a peculiar people, zealous of good works. These things speak, and exhort: in Christ Jesus our Lord.
\gradall{In the day of thy power shall the people offer thee free-will offerings with an holy worship: the dew of thy birth is of the womb of the morning. ℣. The Lord said unto my Lord: Sit thou on my right hand: until I make thine enemies thy footstool.}{Alleluia, alleluia. ℣. The Lord hath said unto me: Thou art my Son, this day have I begotten thee. Alleluia.}
\readingcitation{Gospel}{Luke 2:1}
\lett{A}{t that time:} There went out a decree from Caesar Augustus, that all the world should be taxed. (And this taxing was first made when Cyrenius was governor of Syria.) And all went to be taxed, every one into his own city. And Joseph also went up from Galilee, out of the city of Nazareth, into Judaea, unto the city of David, which is called Bethlehem; (because he was of the house and lineage of David:) To be taxed with Mary his espoused wife, being great with child. And so it was, that, while they were there, the days were accomplished that she should be delivered. And she brought forth her firstborn son, and wrapped him in swaddling clothes, and laid him in a manger; because there was no room for them in the inn. And there were in the same country shepherds abiding in the field, keeping watch over their flock by night. And, lo, the angel of the Lord came upon them, and the glory of the Lord shone round about them: and they were sore afraid. And the angel said unto them, Fear not: for, behold, I bring you good tidings of great joy, which shall be to all people. For unto you is born this day in the city of David a Saviour, which is Christ the Lord. And this shall be a sign unto you; Ye shall find the babe wrapped in swaddling clothes, lying in a manger. And suddenly there was with the angel a multitude of the heavenly host praising God, and saying, Glory to God in the highest, and on earth peace, good will toward men.
\offertory{Let the heavens rejoice, and let the earth be glad before the Lord: for he is come.}
\secret
\lett{W}{e} beseech thee, O Lord, that the oblation of this day's festival may be acceptable unto thee: that by thy bountiful grace we may, through this holy communion, be found in the likeness of him in whom our substance is united unto thee. Who liveth.
\communion{With an holy worship, the dew of thy birth is of the womb of the morning.}

\postcommunion
\lett{G}{rant,} to us, we beseech thee, O Lord our God, that we who in these mysteries draw near with gladness unto the Nativity of Jesus Christ our Lord; may by godly conversation be found worthy to attain unto his fellowship. Who liveth.


\subby{Nativity of Our Lord}
\centerline{\small{(The Second Mass at Dawn)}}
\feastday{Christmas}
%The Second Mass of the 1928 BCP is identical to the third, so the Roman is presented for the second, and the English of the second Mass is provided in the third.

\introit
\lett{L}{ight} shall shine upon us to-day: for unto us the Lord is born: and he shall be called Wonderful, God, the Prince of Peace, Father of the world to come: of whose kingdom there shall be no end. \textit{Ps.} The Lord is King, and hath put on glorious apparel: the Lord hath put on his apparel, and girded himself with strength.
\collect
\lett{G}{rant,} to us, we beseech thee, almighty God: that as thou dost pour forth on us the new light of thine incarnate Word; so he, who doth illuminate our minds by faith, may likewise in our works shew forth his brightness. Through the same.
\Needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty God: that we who devoutly keep the festival of thy blessed Martyr Anastasia; may know her to be our advocate with thee. Through.
\readingcitation{Epistle}{Titus 3:4}
\lett{D}{early beloved:} The kindness and love of God our Saviour toward man appeared, Not by works of righteousness which we have done, but according to his mercy he saved us, by the washing of regeneration, and renewing of the Holy Ghost; Which he shed on us abundantly through Jesus Christ our Saviour; That being justified by his grace, we should be made heirs according to the hope of eternal life: %
in Christ Jesus our Lord.
\gradall{Blessed is he that cometh in the name of the Lord: God is the Lord who hath shewed us light. ℣. This is the Lord's doing: and it is marvellous in our eyes.}{Alleluia, alleluia. ℣. The Lord is King, and hath put on glorious apparel: the Lord hath put on his apparel, and girded himself with strength. Alleluia.}
\readingcitation{Gospel}{Luke 2:15}
\lett{A}{t that time:} The shepherds said one to another, Let us now go even unto Bethlehem, and see this thing which is come to pass, which the Lord hath made known unto us. And they came with haste, and found Mary, and Joseph, and the babe lying in a manger. And when they had seen it, they made known abroad the saying which was told them concerning this child. And all they that heard it wondered at those things which were told them by the shepherds. But Mary kept all these things, and pondered them in her heart. And the shepherds returned, glorifying and praising God for all the things that they had heard and seen, as it was told unto them.
\offertory{God hath made the round world so sure, that it cannot be moved: ever since the world began, hath thy seat, O God, been prepared, thou art from everlasting.}
\secret
\lett{W}{e} beseech thee, O Lord, that these our gifts may be worthy of the mysteries of this day's Nativity, and may evermore shed forth thy peace upon us: that even as he who was born in our manhood did shew forth therein the glory of the Godhead, so we in these thy earthly creatures may be made partakers of that which is heavenly. Through the same.
\Needspace{4\baselineskip}
\lett{A}{ccept,} we beseech thee, O Lord, the gifts which we duly offer: that, by the merits and intercession of blessed Anastasia thy Martyr, they may be profitable to the advancement of our salvation. Through.
\communion{Rejoice greatly, O daughter of Zion, shout, O daughter of Jerusalem: behold, thy King cometh, the holy one and the Saviour of the world.}
\postcommunion
\lett{O}{Lord,} who by the wondrous Birth of thy Son hast put away the old nature of our manhood: grant that by this sacrament of his Nativity we may be restored unto newness of life. Through the same.
\Needspace{4\baselineskip}
\lett{O}{Lord,} who hast satisfied thy family with sacred gifts: we beseech thee, that we may at all times be comforted by the intercession of her whose festival we celebrate. Through.

\subby{Nativity of Our Lord}\label{NativityMassIII}
\feastday{Christmas}
\centerline{\small{(The Third Mass during the Day)}}
\introit
\lett{U}{nto} us a child is born, unto us a son is given: and the government shall be upon his shoulder: and his name shall be called, Angel of mighty counsel. \textit{Ps.} O sing unto the Lord a new song, for he hath done marvellous things.
\collect\label{NativityMassIIICollect}
\lett{A}{lmighty} God, who hast given us thy only begotten Son to take our nature upon him, and as at this time to be born of a Holy Virgin; Grant that we being regenerate, and made thy children by adoption and grace, may daily be renewed by thy holy Spirit; through the same our Lord Jesus Christ, who liveth and reigneth with thee and the same Spirit ever, one God, world without end. \textit{Amen.}
\readingcitation{Epistle}{Hebrews 1:1}
\lett{G}{od,} who at sundry times and in divers manners spake in time past unto the fathers by the prophets, Hath in these last days spoken unto us by his Son, whom he hath appointed heir of all things, by whom also he made the worlds; Who being the brightness of his glory, and the express image of his person, and upholding all things by the word of his power, when he had by himself purged our sins, sat down on the right hand of the Majesty on high; Being made so much better than the angels, as he hath by inheritance obtained a more excellent name than they. For unto which of the angels said he at any time, Thou art my Son, this day have I begotten thee? And again, I will be to him a Father, and he shall be to me a Son? And again, when he bringeth in the firstbegotten into the world, he saith, And let all the angels of God worship him. And of the angels he saith, Who maketh his angels spirits, and his ministers a flame of fire. But unto the Son he saith, Thy throne, O God, is for ever and ever: a sceptre of righteousness is the sceptre of thy kingdom. Thou hast loved righteousness, and hated iniquity; therefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows. And, Thou, Lord, in the beginning hast laid the foundation of the earth; and the heavens are the works of thine hands: They shall perish; but thou remainest; and they all shall wax old as doth a garment; And as a vesture shalt thou fold them up, and they shall be changed: but thou art the same, and thy years shall not fail.
\gradall{All the ends of the earth have seen the salvation of our God: O be joyful in God, all ye lands. ℣. The Lord hath declared his salvation : in the sight of the heathen hath he openly shewed his righteousness.}{Alleluia, alleluia. ℣. A hallowed day hath dawned upon us: come, ye nations, and worship the Lord: for on this day a great light hath descended upon the earth. Alleluia.}
\readingcitation{Gospel}{John 1:1}
\lett{I}{n} the beginning was the Word, and the Word was with God, and the Word was God. The same was in the beginning with God. All things were made by him; and without him was not any thing made that was made. In him was life; and the life was the light of men. And the light shineth in darkness; and the darkness comprehended it not. There was a man sent from God, whose name was John. The same came for a witness, to bear witness of the Light, that all men through him might believe. He was not that Light, but was sent to bear witness of that Light. That was the true Light, which lighteth every man that cometh into the world. He was in the world, and the world was made by him, and the world knew him not. He came unto his own, and his own received him not. But as many as received him, to them gave he power to become the sons of God, even to them that believe on his name: Which were born, not of blood, nor of the will of the flesh, nor of the will of man, but of God. \inrub{Here Genuflect.} And the Word was made flesh, and dwelt among us, (and we beheld his glory, the glory as of the only begotten of the Father,) full of grace and truth.
\offertory{The heavens are thine, the earth also is thine: thou hast laid the foundation of the round world, and all that therein is: righteousness and equity are the habitation of thy seat.}
\secret\label{NativityMassIIISecret}
\lett{S}{anctify,} O Lord, by the new Birth of thine only-begotten Son, the gifts which we offer: and cleanse us from the defilements of our iniquities. Through the same.
\communion{All the ends of the world have seen the salvation of our God.}
\postcommunion\label{NativityMassIIIPostcommunion}
\lett{G}{rant,} we beseech thee, almighty God: that as he who was born this day the Saviour of the world is the author of our heavenly birth; so he may likewise bestow on us the gift of everlasting life. Who liveth.

\readingcitation{Last Gospel}{Matthew 2:1}
\lett{W}{hen} Jesus was born in Bethlehem of Jud{\ae}a in the days of Herod the king, behold, there came wise men from the east to Jerusalem, Saying, Where is he that is born King of the Jews? for we have seen his star in the east, and are come to worship him. When Herod the king had heard these things, he was troubled, and all Jerusalem with him. And when he had gathered all the chief priests and scribes of the people together, he demanded of them where Christ should be born. And they said unto him, In Bethlehem of Judaea: for thus it is written by the prophet, And thou Bethlehem, in the land of Juda, art not the least among the princes of Juda: for out of thee shall come a Governor, that shall rule my people Israel. Then Herod, when he had privily called the wise men, enquired of them diligently what time the star appeared. And he sent them to Bethlehem, and said, Go and search diligently for the young child; and when ye have found him, bring me word again, that I may come and worship him also. When they had heard the king, they departed; and, lo, the star, which they saw in the east, went before them, till it came and stood over where the young child was. When they saw the star, they rejoiced with exceeding great joy. And when they were come into the house, they saw the young child with Mary his mother, and fell down, and worshipped him: and when they had opened their treasures, they presented unto him gifts; gold, and frankincense, and myrrh. And being warned of God in a dream that they should not return to Herod, they departed into their own country another way.

%MANUAL ADJUSTMENT:
\clearpage

\subby{St. Stephen}
\feastday{St. Stephen}
\begin{inhead}
{26 December}\par
{Second Class Double, Simple Octave}
\end{inhead}
\par\noindent
\textit{Opening Sentence.} The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus.\vr{Ps. 92:11}

\properantiphonfix

\introit
\lett{P}{rinces} did sit, and speak against me, and the wicked persecuted me: help me, O Lord my God, for thy servant is occupied in thy commandments. \textit{Ps.} Blessed are those that are undefiled in the way, and walk in the law of the Lord.
\collect
\lett{G}{rant,} O Lord, that, in all our sufferings here upon earth for the testimony of thy truth, we may stedfastly look up to heaven, and by faith behold the glory that shall be revealed; and, being filled with the Holy Ghost, may learn to love and bless our persecutors by the example of thy first Martyr Saint Stephen, who prayed for his murderers to thee, O blessed Jesus, who standest at the right hand of God to succour all those who suffer for thee, our only Mediator and Advocate. \textit{Amen.}
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIICollect}).
\end{rubric}

%\needspace{4\baselineskip}
\begin{rubric}
    Collect on the Octave Day of St. Stephen,
\end{rubric}
\lett{A}{lmighty} and everlasting God, who, in the blood of the blessed Levite Stephen, didst consecrate the first-fruits of the Martyrs: grant, we beseech thee; that he may intercede for us, who prayed even for his persecutors to our Lord Jesus Christ, thy Son. Who liveth and reigneth with thee.

\readingcitation{Epistle}{Acts 7:55}
\lett{I}{n those days:} Stephen, being full of the Holy Ghost, looked up stedfastly into heaven, and saw the glory of God, and Jesus standing on the right hand of God, and said, Behold, I see the heavens opened, and the Son of man standing on the right hand of God. Then they cried out with a loud voice, and stopped their ears, and ran upon him with one accord, and cast him out of the city, and stoned him: and the witnesses laid down their clothes at a young man’s feet, whose name was Saul. And they stoned Stephen, calling upon God, and saying, Lord Jesus, receive my spirit. And he kneeled down, and cried with a loud voice, Lord, lay not this sin to their charge. And when he had said this, he fell asleep.

\gradall{Princes did sit and speak against me; and the wicked persecuted me. ℣. Help me, O Lord my God: save me for thy mercy's sake.}{Alleluia, alleluia. ℣. I see the heavens opened, and Jesus standing on the right hand of the power of God. Alleluia.}

\readingcitation{Gospel}{Matthew 23:34}
\lett{A}{t that time:} Jesus said unto the scribes and Pharisees: Behold, I send unto you prophets, and wise men, and scribes: and some of them ye shall kill and crucify; and some of them shall ye scourge in your synagogues, and persecute them from city to city: That upon you may come all the righteous blood shed upon the earth, from the blood of righteous Abel unto the blood of Zacharias son of Barachias, whom ye slew between the temple and the altar. Verily I say unto you, All these things shall come upon this generation. O Jerusalem, Jerusalem, thou that killest the prophets, and stonest them which are sent unto thee, how often would I have gathered thy children together, even as a hen gathereth her chickens under her wings, and ye would not! Behold, your house is left unto you desolate. For I say unto you, Ye shall not see me henceforth, till ye shall say, Blessed is he that cometh in the name of the Lord.
\offertory{The Apostles chose Stephen the Levite, a man full of faith and of the Holy Ghost: whom the Jews stoned as he prayed, saying: Lord Jesus, receive my spirit, alleluia.}
\secret
\lett{A}{ccept,} O Lord, these gifts for the commemoration of thy Saints: that like as their passion hath raised them to glory; so our devotion may lead us to innocency of life. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIISecret}).
\end{rubric}
\communion{I see the heavens opened, and Jesus standing on the tight hand of the power of God: O Lord Jesus, receive my spirit, and lay not this sin to their charge.}
\postcommunion
\lett{M}{ay} the mysteries which we have received be for our help, O Lord: and at the intercession of blessed Stephen thy Martyr, may they stablish us with thine everlasting protection. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIIPostcommunion}).
\end{rubric}


\subby{St. John}
\feastday{St. John}
\begin{inhead}
{27 December}\par
{Second Class Double, Simple Octave}
\end{inhead}
\par\noindent
\textit{Opening Sentence.} Greatly to be had in honour is blessed John. For he leaned on the Lord's bosom at Supper.

\properantiphonfix

\introit
\lett{I}{n} the midst of the Church he opened his mouth: and the Lord filled him with the spirit of wisdom and understanding: in a robe of glory he arrayed him. \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.
\collect
\lett{M}{erciful} Lord, we beseech thee to cast thy bright beams of light upon thy Church, that it, being illumined by the doctrine of thy blessed Apostle and Evangelist Saint John, may so walk in the light of thy truth, that it may at length attain to life everlasting. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIICollect}).
\end{rubric}
\readingcitation{Epistle}{1 John 1:1}
\lett{T}{hat} which was from the beginning, which we have heard, which we have seen with our eyes, which we have looked upon, and our hands have handled, of the Word of life; (For the life was manifested, and we have seen it, and bear witness, and shew unto you that eternal life, which was with the Father, and was manifested unto us;) That which we have seen and heard declare we unto you, that ye also may have fellowship with us: and truly our fellowship is with the Father, and with his Son Jesus Christ. And these things write we unto you, that your joy may be full. This then is the message which we have heard of him, and declare unto you, that God is light, and in him is no darkness at all. If we say that we have fellowship with him, and walk in darkness, we lie, and do not the truth: But if we walk in the light, as he is in the light, we have fellowship one with another, and the blood of Jesus Christ his Son cleanseth us from all sin. If we say that we have no sin, we deceive ourselves, and the truth is not in us. If we confess our sins, he is faithful and just to forgive us our sins, and to cleanse us from all unrighteousness. If we say that we have not sinned, we make him a liar, and his word is not in us.

%MANUAL ADJUSTMENT:

\gradall{Then went this saying abroad among the brethren: That that disciple should not die: yet Jesus said not: He shall not die. ℣. But: If I will that he tarry till I come: follow thou me.}{Alleluia, alleluia. ℣. This is the disciple which testifieth of these things: and we know that his testimony is true. Alleluia.}

\readingcitation{Gospel}{John 21:19}
\lett{A}{t that time:} Jesus said unto Peter: Follow me. Then Peter, turning about, seeth the disciple whom Jesus loved following; which also leaned on his breast at supper, and said, Lord, which is he that betrayeth thee? Peter seeing him saith to Jesus, Lord, and what shall this man do? Jesus saith unto him, If I will that he tarry till I come, what is that to thee? follow thou me. Then went this saying abroad among the brethren, that that disciple should not die: yet Jesus said not unto him, He shall not die; but, If I will that he tarry till I come, what is that to thee? This is the disciple which testifieth of these things, and wrote these things: and we know that his testimony is true. And there are also many other things which Jesus did, the which, if they should be written every one, I suppose that even the world itself could not contain the books that should be written.
\offertory{The just shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\secret
\lett{R}{eceive,} O Lord, the gifts which we offer unto thee on the solemnity (commemoration) of him, in whose advocacy we trust for deliverance. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIISecret}).
\end{rubric}
\communion{Then went abroad this saying among the brethren, that that disciple should not die: yet Jesus said not: He shall not die; but: If I will that he tarry till I come.}
\postcommunion
\lett{O}{God,} who hast refreshed us with heavenly meat and drink, we humbly beseech thee: that we may be defended by the prayers of him, in whose memory we have received the same. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIIPostcommunion}).
\end{rubric}


\subby{Holy Innocents}
\feastday{Holy Innocents}
\begin{inhead}
{28 December}\par
{Second Class Double, Simple Octave}
\end{inhead}
\par\noindent
\textit{Opening Sentence.} The righteous live for evermore. Their reward also is with the Lord.

\properantiphonfix

\introit
\lett{O}{ut} of the mouths of babes, O God, and of sucklings, hast thou perfected praise because of thine adversaries. \textit{Ps.} O Lord our Governor: how excellent is thy name in all the world.
\begin{rubric}
    \emph{Gloria in excelsis} is not said, nor \emph{Alleluia}, nor \emph{Ite, Missa est}, unless this Feast falls on a Sunday, or enjoys the rite of a \nth{1} Double; but on the Octave Day they are always said.
\end{rubric}
\collect
\lett{O}{almighty} God, who out of the mouths of babes and sucklings hast ordained strength, and madest infants to glorify thee by their deaths: Mortify and kill all vices in us, and so strengthen us by thy grace, that by the innocency of our lives, and constancy of our faith even unto death, we may glorify thy holy Name. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIICollect}).
\end{rubric}
\readingcitation{Epistle}{Revelation 14:1}
\lett{I}{n those days:} I looked, and, lo, a Lamb stood on the mount Zion, and with him an hundred forty and four thousand, having his Father's name written in their foreheads. And I heard a voice from heaven, as the voice of many waters, and as the voice of a great thunder: and I heard the voice of harpers harping with their harps: And they sung as it were a new song before the throne, and before the four beasts, and the elders: and no man could learn that song but the hundred and forty and four thousand, which were redeemed from the earth. These are they which were not defiled with women; for they are virgins. These are they which follow the Lamb whithersoever he goeth. These were redeemed from among men, being the firstfruits unto God and to the Lamb. And in their mouth was found no guile: for they are without fault before the throne of God.
\gradual{Our soul is escaped, even as a bird, out of the snare of the fowler. ℣. The snare is broken, and we are delivered: our help standeth in the name of the Lord, who hath made heaven and earth.}
\begin{rubric}
     The Tract is not said when the Alleluia is said.
\end{rubric}
\tract{The blood of the Saints have they shed like water on every side of Jerusalem. ℣. And there was no man to bury them. ℣. Avenge, O Lord, the blood of thy Saints, that is shed upon the earth.}
\begin{rubric}
    The following Alleluia is said when this Feast falls on Sunday or enjoys the rite of a \nth{1} Double: but on the Octave Day it is always said, as also in Votive Masses before Septuagesima or after Pentecost.
\end{rubric}
\alleluia{Alleluia, alleluia. ℣. Praise the Lord, ye children: O praise the name of the Lord. Alleluia.}
\readingcitation{Gospel}{Matthew 2:13}
\lett{A}{t that time:} the angel of the Lord appeareth to Joseph in a dream, saying, Arise, and take the young child and his mother, and flee into Egypt, and be thou there until I bring thee word: for Herod will seek the young child to destroy him. When he arose, he took the young child and his mother by night, and departed into Egypt: And was there until the death of Herod: that it might be fulfilled which was spoken of the Lord by the prophet, saying, Out of Egypt have I called my son. Then Herod, when he saw that he was mocked of the wise men, was exceeding wroth, and sent forth, and slew all the children that were in Bethlehem, and in all the coasts thereof, from two years old and under, according to the time which he had diligently enquired of the wise men. Then was fulfilled that which was spoken by Jeremy the prophet, saying, In Rama was there a voice heard, lamentation, and weeping, and great mourning, Rachel weeping for her children, and would not be comforted, because they are not.
\offertory{Our soul is escaped, even as a bird out of the snare of the fowler: the snare is broken, and we are delivered.}
\secret
\lett{M}{ay} the devout prayers of thy Saints never fail us, O Lord: that they may render our gifts acceptable unto thee, and ever obtain for us thy pardon. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIISecret}).
\end{rubric}
\communion{In Rama, a voice was heard, lamentation and mourning: Rachel weeping for her children, and would not be comforted, because they are not.}
\postcommunion
\lett{W}{e} beseech thee, O Lord: that the gifts, which we have offered and received, may through the prayers of thy Saints effectually avail for our succour both in this life and in that which is to come. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIIPostcommunion}).
\end{rubric}


\subby{Sunday within the Nativity Octave}
\feastday{Christmas I}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{hile} all things were in quiet silence, and night was in the midst of her swift course, thine almighty Word, O Lord, leaped down from heaven out of thy royal throne. \textit{Ps.} The Lord is King, and hath put on glorious apparel: the Lord hath put on his apparel, and girded himself with strength.
\collect
\lett{A}{lmighty} God, who hast given us thy only-begotten Son to take our nature upon him, and as at this time to be born of a pure virgin: Grant that we being regenerate, and made thy children by adoption and grace, may daily be renewed by thy Holy Spirit; through the same our Lord Jesus Christ, who liveth and reigneth with thee and the same Spirit ever, one God, world without end. \textit{Amen.}
\Needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty God: that we, who through our ancient bondage are held beneath the yoke of sin; may by the new Birth of thy only-begotten Son in the flesh obtain deliverance. Through the same.
\readingcitation{Epistle}{Galatians 4:1}
\lett{B}{rethren:} The heir, as long as he is a child, differeth nothing from a servant, though he be lord of all; But is under tutors and governors until the time appointed of the father. Even so we, when we were children, were in bondage under the elements of the world: But when the fulness of the time was come, God sent forth his Son, made of a woman, made under the law, To redeem them that were under the law, that we might receive the adoption of sons. And because ye are sons, God hath sent forth the Spirit of his Son into your hearts, crying, Abba, Father. Wherefore thou art no more a servant, but a son; and if a son, then an heir of God through Christ.
\gradall{Thou art fairer than the children of men: full of grace are thy lips. ℣. My heart is inditing of a good matter, I speak of the things which I have made unto the King: my tongue is the pen of a ready writer.}{Alleluia, alleluia. ℣. The Lord is King, and hath put on glorious apparel: the Lord hath put on his apparel, and girded himself with strength. Alleluia.}

\readingcitation{Gospel}{Matthew 1:18}
\lett{T}{he} birth of Jesus Christ was on this wise: When as his mother Mary was espoused to Joseph, before they came together, she was found with child of the Holy Ghost. Then Joseph her husband, being a just man, and not willing to make her a publick example, was minded to put her away privily. But while he thought on these things, behold, the angel of the Lord appeared unto him in a dream, saying, Joseph, thou son of David, fear not to take unto thee Mary thy wife: for that which is conceived in her is of the Holy Ghost. And she shall bring forth a son, and thou shalt call his name \divineName{Jesus}: for he shall save his people from their sins. Now all this was done, that it might be fulfilled which was spoken of the Lord by the prophet, saying, Behold, a virgin shall be with child, and shall bring forth a son, and they shall call his name Emmanuel, which being interpreted is, God with us. Then Joseph being raised from sleep did as the angel of the Lord had bidden him, and took unto him his wife: And knew her not till she had brought forth her firstborn son: and he called his name \divineName{Jesus}.

\offertory{God hath made the round world so sure, that it cannot be moved: ever since the world began hath thy seat, O God, been prepared, thou art from everlasting.}
\secret
\lett{G}{rant,} we beseech thee, almighty God: that the gift which we offer in the sight of thy majesty, may obtain for us grace to serve thee with all godliness, and bring us in the end to everlasting felicity. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIISecret}).
\end{rubric}
\communion{Take the young Child and his Mother, and go into the land of Israel: for they are dead which sought the young Child's life.}
\postcommunion
\lett{M}{ay} the operation of this mystery, O Lord, avail for the cleansing of our sins, and for the fulfilment of our godly desires. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIIIPostcommunion}).
\end{rubric}


\subby{Circumcision of Our Lord}
\feastday{Circumcision}
\begin{inhead}
{1 January}\par
{Second Class Double}
\end{inhead}
\par\noindent
\textit{Opening Sentence.} The Word was made flesh and dwelt among us. Alleluia.

\properantiphonfix

\introit
\lett{U}{nto} us a Child is born, unto us a son is given; and the government shall be upon his shoulder: and his name shall be called Angel of mighty counsel. \textit{Ps.} O sing unto the Lord a new song: for he hath done marvellous things.
\collect
\lett{A}{lmighty} God, who madest thy blessed Son to be circumcised, and obedient to the law for man; Grant us the true circumcision of the Spirit; that, our hearts, and all our members, being mortified from all worldly and carnal lusts, we may in all things obey thy blessed will; through the same thy Son Jesus Christ our Lord. Who liveth.
%1928 Reading is reverted to the Prayer Book Tradition. The 1928 innovation does not seem fitting when a Feast of the Holy Name (which it did not have) is about to be celebrated.
\readingcitation{Epistle}{Romans 4:8}
\lett{B}{rethren:} Blessed is the man to whom the Lord will not impute sin. Cometh this blessedness then upon the circumcision only, or upon the uncircumcision also? for we say that faith was reckoned to Abraham for righteousness. How was it then reckoned? when he was in circumcision, or in uncircumcision? Not in circumcision, but in uncircumcision. And he received the sign of circumcision, a seal of the righteousness of the faith which he had yet being uncircumcised: that he might be the father of all them that believe, though they be not circumcised; that righteousness might be imputed unto them also: And the father of circumcision to them who are not of the circumcision only, but who also walk in the steps of that faith of our father Abraham, which he had being yet uncircumcised. For the promise, that he should be the heir of the world, was not to Abraham, or to his seed, through the law, but through the righteousness of faith. For if they which are of the law be heirs, faith is made void, and the promise made of none effect.

%MANUAL ADJUSTMENT:

\gradual{All the ends of the earth have seen the salvation of our God: O be joyful in God, all ye lands. ℣. The Lord declared his salvation: his righteousness hath he shewed in the sight of the heathen.}

\alleluia{Alleluia, alleluia. ℣. God, who at sundry times spake in time past unto the fathers by the Prophets, hath in these last days spoken unto us by his Son. Alleluia.}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\readingcitation{Gospel}{Luke 2:15}
\lett{A}{t that time:} It came to pass, as the angels were gone away from them into heaven, the shepherds said one to another, Let us now go even unto Bethlehem, and see this thing which is come to pass, which the Lord hath made known unto us. And they came with haste, and found Mary, and Joseph, and the babe lying in a manger. And when they had seen it, they made known abroad the saying which was told them concerning this child. And all they that heard it wondered at those things which were told them by the shepherds. But Mary kept all these things, and pondered them in her heart. And the shepherds returned, glorifying and praising God for all the things that they had heard and seen, as it was told unto them. And when eight days were accomplished for the circumcising of the child, his name was called \divineName{Jesus}, which was so named of the angel before he was conceived in the womb.

\offertory{The heavens are thine, the earth also is thine: thou hast laid the foundation of the round world, and all that therein is: righteousness and equity are the habitation of thy seat.}
\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us, whom thou dost cleanse by thy heavenly mysteries. Through.
\communion{All the ends of the world have seen the salvation of our God.}
\postcommunion
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary the Virgin Mother of God, make us partakers of thy heavenly healing. Through the same.

\begin{rubric}
    The same Collect, Epistle, and Gospel shall serve for every day after unto the Epiphany.
\end{rubric}

\subby{Most Holy Name of Jesus}\label{MostHolyName}
\feastday{Most Holy Name}
\begin{inhead}
{2 January}\par
{Second Class Double}
\end{inhead}
\par\noindent
%From the 7 August Feast in Sarum:
\textit{Opening Sentence.} Our help is in the Name of the Lord who hath made heaven and earth.\vr{Ps. 124:8}

\introit
\lett{A}{t} the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth: and every tongue should confess that Jesus Christ is Lord, to the glory of God the Father. \textit{Ps.} O Lord our Governor, how excellent is thy name in all the world.
\collect
\lett{O}{God,} who didst appoint thine only-begotten Son to be the Saviour of mankind, and didst command that he should be called \divineName{Jesus}: mercifully grant; that as we venerate his holy name on earth, so we may rejoice to behold him in heaven. Through the same.
\readingcitation{Epistle}{Acts 4:8}
\lett{I}{n those days:} Peter, filled with the Holy Ghost, said unto them, Ye rulers of the people, and elders of Israel, If we this day be examined of the good deed done to the impotent man, by what means he is made whole; Be it known unto you all, and to all the people of Israel, that by the name of Jesus Christ of Nazareth, whom ye crucified, whom God raised from the dead, even by him doth this man stand here before you whole. This is the stone which was set at nought of you builders, which is become the head of the corner. Neither is there salvation in any other: for there is none other name under heaven given among men, whereby we must be saved.
\gradall{Deliver us, O Lord our God, and gather us from among the heathen: that we may give thanks unto thy holy name, and make our boast of thy praise. ℣. Thou, O Lord, art our Father, our Redeemer: thy name is from everlasting.}{Alleluia, alleluia. ℣. My mouth shall speak the praise of the Lord: and let all flesh give thanks unto his holy name. Alleluia.}
\readingcitation{Gospel}{Luke 2:21}
\lett{A}{t that time:} When eight days were accomplished for the circumcising of the child, his name was called \divineName{Jesus}, which was so named of the angel before he was conceived in the womb.
\offertory{I will thank thee, O Lord my God, with all my heart, and will praise thy name for evermore: for thou, Lord, art good and gracious: and of great mercy unto all them that call upon thee, alleluia.}
\secret
\lett{W}{e} beseech thee, most merciful God, that thy benediction, whereby all created things are sustained, may sanctify this our sacrifice, which we offer unto thee to the glory of the name of thy Son Jesus Christ our Lord: that it may be acceptable for the praise of thy majesty, and profitable for our salvation. Through the same.
\communion{All nations whom thou hast made shall come and worship thee, O Lord, and shall glorify thy name: for thou art great, and doest wondrous things: thou art God alone, alleluia.}
\postcommunion
\lett{A}{lmighty} and everlasting God, our Creator and Redeemer, mercifully regard our prayers: and vouchsafe to receive with favourable and gracious countenance the sacrifice of the saving victim, which we have offered to thy majesty; in honour of the name of thy Son Jesus Christ our Lord; that we, being fulfilled with thy grace, may rejoice that our names are written in heaven beneath the glorious name of Jesus, in the book of everlasting predestination. Through the same.

%It happens that there is a second Sunday after Christmas but before Epiphany. The traditional Roman solution is simply to celebrate the given feast day (the Second Sunday will always be an Octave Day of St. John, Holy Innocents, or St. Stephen; or it will be the Vigil of Epiphany). Modern Roman practice is to celebrate the Feast of the Holy Name (early or late).
%The English tradition has been to celebrate the Feast of the Circumcision again. However, the English and American 1928 Prayer Books provide propers for this Sunday, where the previous practice was to use the propers from the Circumcision. Given that the AWRV generally uses the Roman Rite as it existed in 1950, we simply provide the American 1928 here.
%There is the additional problem of which Minor Propers to use. We simply use the Circumcision propers, in line with Prayer Book Tradition and as suggested by the 1933 English Missal.
\subby{Second Sunday after Nativity}
\feastday{Christmas II}
\begin{inhead}
{The Sunday between Circumcision \& Epiphany, if there be one.}\par
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{U}{nto} us a Child is born, unto us a son is given; and the government shall be upon his shoulder: and his name shall be called Angel of mighty counsel. \textit{Ps.} O sing unto the Lord a new song: for he hath done marvellous things.

\collect
\lett{A}{lmighty} God, who hast poured upon us the new light of thine incarnate Word; Grant that the same light enkindled in our hearts may shine forth in our lives. Through.

\readingcitation{Epistle}{Isaiah 61:1}
\lett{T}{he} Spirit of the Lord \divineName{God} is upon me; because the \divineName{Lord} hath anointed me to preach good tidings unto the meek; he hath sent me to bind up the brokenhearted, to proclaim liberty to the captives, and the opening of the prison to them that are bound; To proclaim the acceptable year of the \divineName{Lord}, and the day of vengeance of our God; to comfort all that mourn; To appoint unto them that mourn in Zion, to give unto them beauty for ashes, the oil of joy for mourning, the garment of praise for the spirit of heaviness; that they might be called trees of righteousness, the planting of the \divineName{Lord}, that he might be glorified.

%MANUAL ADJUSTMENT:
\gradual{All the ends of the earth have seen the salvation of our God: O be joyful in God, all ye lands. ℣. The Lord declared his salvation: his righteousness hath he shewed in the sight of the heathen.}

\alleluia{Alleluia, alleluia. ℣. God, who at sundry times spake in time past unto the fathers by the Prophets, hath in these last days spoken unto us by his Son. Alleluia.}

\readingcitation{Gospel}{Matthew 2:19}
\lett{A}{t that time:} When Herod was dead, behold, an angel of the Lord appeareth in a dream to Joseph in Egypt, Saying, Arise, and take the young child and his mother, and go into the land of Israel: for they are dead which sought the young child's life. And he arose, and took the young child and his mother, and came into the land of Israel. But when he heard that Archelaus did reign in Judaea in the room of his father Herod, he was afraid to go thither: notwithstanding, being warned of God in a dream, he turned aside into the parts of Galilee: And he came and dwelt in a city called Nazareth: that it might be fulfilled which was spoken by the prophets, He shall be called a Nazarene.
\offertory{The heavens are thine, the earth also is thine: thou hast laid the foundation of the round world, and all that therein is: righteousness and equity are the habitation of thy seat.}
\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us, whom thou dost cleanse by thy heavenly mysteries. Through.
\communion{All the ends of the world have seen the salvation of our God.}
\postcommunion
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary the Virgin Mother of God, make us partakers of thy heavenly healing. Through the same.

%MANUAL ADJUSTMENT:
\clearpage

\subby{Epiphany of Our Lord}\label{EpiphanyMass}
\feastday{Epiphany}
\begin{inhead}
{First Class Double, Second Class Octave}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{ehold,} the Lord the ruler is come: and in his hand the kingdom, and power, and dominion. \textit{Ps.} Give the King thy judgments, O God: and thy righteousness unto the King's Son.
\collect\label{EpiphanyMassCollect}
\lett{O}{God,} who by the leading of a star didst manifest thy only-begotten Son to the Gentiles; Mercifully grant that we, who know thee now by faith, may after this life have the fruition of thy glorious Godhead. Through the same.
\begin{rubric}
    This Collect to be said daily throughout the Octave.
\end{rubric}
\readingcitation{Epistle}{Ephesians 3:1}
\lett{B}{rethren:} For this cause I Paul, the prisoner of Jesus Christ for you Gentiles, If ye have heard of the dispensation of the grace of God which is given me to you-ward: How that by revelation he made known unto me the mystery; (as I wrote afore in few words, Whereby, when ye read, ye may understand my knowledge in the mystery of Christ) Which in other ages was not made known unto the sons of men, as it is now revealed unto his holy apostles and prophets by the Spirit; That the Gentiles should be fellowheirs, and of the same body, and partakers of his promise in Christ by the gospel: Whereof I was made a minister, according to the gift of the grace of God given unto me by the effectual working of his power. Unto me, who am less than the least of all saints, is this grace given, that I should preach among the Gentiles the unsearchable riches of Christ; And to make all men see what is the fellowship of the mystery, which from the beginning of the world hath been hid in God, who created all things by Jesus Christ: To the intent that now unto the principalities and powers in heavenly places might be known by the church the manifold wisdom of God, According to the eternal purpose which he purposed in Christ Jesus our Lord: In whom we have boldness and access with confidence by the faith of him.

%MANUAL ADJUSTMENT:
\gradual{All they from Saba shall come, bringing gold and incense, and shall shew forth the praises of the Lord. ℣. Arise and shine, O Jerusalem: for the glory of the Lord is risen upon thee.}

\alleluia{Alleluia, alleluia. ℣. We have seen his star in the East: and are come with gifts to worship the Lord. Alleluia.}

\readingcitation{Gospel}{Matthew 2:1}
\lett{W}{hen} Jesus was born in Bethlehem of Jud{\ae}a in the days of Herod the king, behold, there came wise men from the east to Jerusalem, Saying, Where is he that is born King of the Jews? for we have seen his star in the east, and are come to worship him. When Herod the king had heard these things, he was troubled, and all Jerusalem with him. And when he had gathered all the chief priests and scribes of the people together, he demanded of them where Christ should be born. And they said unto him, In Bethlehem of Jud{\ae}a: for thus it is written by the prophet, And thou Bethlehem, in the land of Juda, art not the least among the princes of Juda: for out of thee shall come a Governor, that shall rule my people Israel. Then Herod, when he had privily called the wise men, enquired of them diligently what time the star appeared. And he sent them to Bethlehem, and said, Go and search diligently for the young child; and when ye have found him, bring me word again, that I may come and worship him also. When they had heard the king, they departed; and, lo, the star, which they saw in the east, went before them, till it came and stood over where the young child was. When they saw the star, they rejoiced with exceeding great joy. And when they were come into the house, they saw the young child with Mary his mother, \inrub{Genuflect here.} and fell down, and worshipped him: and when they had opened their treasures, they presented unto him gifts; gold, and frankincense, and myrrh. And being warned of God in a dream that they should not return to Herod, they departed into their own country another way.

\offertory{The kings of Tharsis and of the isles shall give presents: the kings of Arabia and Saba shall bring gifts: all kings shall fall down before him, all nations shall do him service.}
\secret\label{EpiphanyMassSecret}
\lett{W}{e} beseech thee, O Lord, mercifully to look upon the gifts of thy Church: wherein no longer offering gold and frankincense and myrrh; we sacrifice and receive him who by those gifts was signified, even Jesus Christ thy Son our Lord; Who liveth and reigneth with thee.
\communion{We have seen his star in the East, and are come with gifts to worship the Lord.}
\postcommunion\label{EpiphanyMassPostcommunion}
\lett{G}{rant,} we beseech thee, almighty God: that those things which we celebrate in solemn worship, we may attain in pureness of mind and understanding. Through.
\begin{rubric}
    Within the Octave, Mass is said as on the Feast. The \nth{2} Prayer \emph{of St. Mary} (p. \pageref{SPMaryPostChristmas}), and \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}), or, \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}

\subsec{Sunday within the Octave of the Epiphany}
\feastday{Epiphany I}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{n} a throne exalted I beheld a man sitting, whom a multitude of Angels worship, singing together: behold him, the name of whose empire is for everlasting. \textit{Ps.} O be joyful in the Lord, all ye lands: serve the Lord with gladness.
\collect
%1928 same as Roman
\lett{O}{Lord,} we beseech thee mercifully to receive the prayers of thy people who call upon thee: and grant that they may both perceive and know what things they ought to do, and also may have grace and power faithfully to fulfil the same. Through.
\begin{rubric}
    Commemoration is made of the Epiphany (p. \pageref{EpiphanyMassCollect}).
\end{rubric}
\readingcitation{Epistle}{Romans 12:1}
\lett{B}{rethren:} I beseech you, by the mercies of God, that ye present your bodies a living sacrifice, holy, acceptable unto God, which is your reasonable service. And be not conformed to this world: but be ye transformed by the renewing of your mind, that ye may prove what is that good, and acceptable, and perfect, will of God. For I say, through the grace given unto me, to every man that is among you, not to think of himself more highly than he ought to think; but to think soberly, according as God hath dealt to every man the measure of faith. For as we have many members in one body, and all members have not the same office: So we, being many, are one body in Christ, and every one members one of another.

%MANUAL ADJUSTMENT:
\gradual{Blessed be the Lord, even the God of Israel, which only doeth wondrous things from the beginning. ℣. The mountains shall bring peace, and the little hills righteousness unto thy people.}

\alleluia{Alleluia, alleluia. ℣. O be joyful in the Lord, all ye lands: serve the Lord with gladness. Alleluia.}

%\vspace{-1ex}
\readingcitation{Gospel}{Luke 2:41}
\lett{N}{ow} Jesus' parents went to Jerusalem every year at the feast of the passover. And when he was twelve years old, they went up to Jerusalem after the custom of the feast. And when they had fulfilled the days, as they returned, the child Jesus tarried behind in Jerusalem; and Joseph and his mother knew not of it. But they, supposing him to have been in the company, went a day's journey; and they sought him among their kinsfolk and acquaintance. And when they found him not, they turned back again to Jerusalem, seeking him. And it came to pass, that after three days they found him in the temple, sitting in the midst of the doctors, both hearing them, and asking them questions. And all that heard him were astonished at his understanding and answers. And when they saw him, they were amazed: and his mother said unto him, Son, why hast thou thus dealt with us? behold, thy father and I have sought thee sorrowing. And he said unto them, How is it that ye sought me? wist ye not that I must be about my Father's business? And they understood not the saying which he spake unto them. And he went down with them, and came to Nazareth, and was subject unto them: but his mother kept all these sayings in her heart. And Jesus increased in wisdom and stature, and in favour with God and man.

\offertory{O be joyful in the Lord, all ye lands, serve the Lord with gladness: and come before his presence with a song: for the Lord he is God.}
\secret
\lett{M}{ay} the sacrifice which we offer unto thee, O Lord, evermore quicken and defend us. Through.
\begin{rubric}
    Commemoration is made of the Epiphany (p. \pageref{EpiphanyMassSecret}).
\end{rubric}
\communion{Son, why hast thou thus dealt with us? behold, thy father and I have sought thee sorrowing. And how is it that ye sought me? wist ye not that I must be about my Father's business?}
\postcommunion
\lett{W}{e} humbly beseech thee, almighty God: that as thou dost refresh us with thy sacraments, so thou wouldest vouchsafe unto us to do thee worthy and acceptable service. Through.
\begin{rubric}
    Commemoration is made of the Epiphany (p. \pageref{EpiphanyMassPostcommunion}).
\end{rubric}

\subby{Baptism of Our Lord}

\centerline{\small{(Octave Day of the Epiphany)}}

\feastday{Baptism of Our Lord}
\begin{inhead}
{Greater Double}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{ehold,} the Lord the ruler is come: and in his hand the kingdom, and power, and dominion. \textit{Ps.} Give the King thy judgements, O God: and thy righteousness unto the King's Son.
\collect
\lett{O}{God,} whose only-begotten Son hath been made manifest in substance of our flesh: grant, we beseech thee; that, as we have known him after the fashion of our outward likeness, so through him we may be made worthy inwardly to be renewed. Who liveth.

\readingcitation{Epistle}{Isaiah 60:1}
\lett{A}{rise,} shine; for thy light is come, and the glory of the \divineName{Lord} is risen upon thee. For, behold, the darkness shall cover the earth, and gross darkness the people: but the \divineName{Lord} shall arise upon thee, and his glory shall be seen upon thee. And the Gentiles shall come to thy light, and kings to the brightness of thy rising. Lift up thine eyes round about, and see: all they gather themselves together, they come to thee: thy sons shall come from far, and thy daughters shall be nursed at thy side. Then thou shalt see, and flow together, and thine heart shall fear, and be enlarged; because the abundance of the sea shall be converted unto thee, the forces of the Gentiles shall come unto thee. The multitude of camels shall cover thee, the dromedaries of Midian and Ephah; all they from Sheba shall come: they shall bring gold and incense; and they shall shew forth the praises of the \divineName{Lord}.

%MANUAL ADJUSTMENT:
\gradual{All they from Saba shall come, bringing gold and incense, and shall shew forth the praise of the Lord. ℣. Arise and shine, O Jerusalem: for the glory of the Lord is risen upon thee.}

\alleluia{Alleluia, alleluia. ℣. We have seen his star in the East, and are come with gifts to worship the Lord. Alleluia.}

\readingcitation{Gospel}{John 1:29}
\lett{A}{t that time:} John seeth Jesus coming unto him, and saith, Behold the Lamb of God, which taketh away the sin of the world. This is he of whom I said, After me cometh a man which is preferred before me: for he was before me. And I knew him not: but that he should be made manifest to Israel, therefore am I come baptizing with water. And John bare record, saying, I saw the Spirit descending from heaven like a dove, and it abode upon him. And I knew him not: but he that sent me to baptize with water, the same said unto me, Upon whom thou shalt see the Spirit descending, and remaining on him, the same is he which baptizeth with the Holy Ghost. And I saw, and bare record that this is the Son of God.

\offertory{The Kings of Tharsis and of the isles shall give presents: the Kings of Arabia and Saba shall bring gifts: all Kings shall fall down before him, all nations shall do him service.}
\secret
\lett{W}{e} offer our sacrifices unto thee, O Lord, in remembrance of the manifestation of thy new-born Son, humbly beseeching thee: that like as he is the author of our gifts, so he may also mercifully receive the same, even Jesus Christ our Lord. Who liveth.
\communion{We have seen his star in the East, and are come with gifts to worship the Lord.}
\postcommunion
\lett{P}{revent} us, O Lord, we beseech thee, at all times and in all places with thy heavenly light: that as thou hast vouchsafed to make us partakers of this mystery, so we may understand it aright in purity of heart, and receive it with worthy affection. Through.

%MANUAL ADJUSTMENT:
\clearpage

%The 1928 breaks the unity of the Epiphany events---the Magi, Baptism, and Wedding at Cana---by forcing the Baptism propers on the Sunday after the Octave Day. It also divorces the Mass from the Office. It must be assuming the Baptism will never be celebrated. Reverted to 1892 BCP.
\subby{Second Sunday after Epiphany}\label{EpiphanyII}
\feastday{Epiphany II}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{A}{ll} the earth shall worship thee, O God, and sing of thee: they shall sing praise unto thy name, O thou most Highest. \textit{Ps.} O be joyful in God, all ye lands: sing praises unto the honour of his name: make his praise to be glorious.
\collect\label{EpiphanyIICollect}
\lett{A}{lmighty} and everlasting God, who dost govern all things in heaven and earth; Mercifully hear the supplications of thy people, and grant us thy peace all the days of our life. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Collect is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 12:6}
\lett{B}{rethren:} Having then gifts differing according to the grace that is given to us, whether prophecy, let us prophesy according to the proportion of faith; Or ministry, let us wait on our ministering: or he that teacheth, on teaching; Or he that exhorteth, on exhortation: he that giveth, let him do it with simplicity; he that ruleth, with diligence; he that sheweth mercy, with cheerfulness. Let love be without dissimulation. Abhor that which is evil; cleave to that which is good. Be kindly affectioned one to another with brotherly love; in honour preferring one another; Not slothful in business; fervent in spirit; serving the Lord; Rejoicing in hope; patient in tribulation; continuing instant in prayer; Distributing to the necessity of saints; given to hospitality Bless them which persecute you: bless, and curse not. Rejoice with them that do rejoice, and weep with them that weep. Be of the same mind one toward another. Mind not high things, but condescend to men of low estate.
\gradall{The Lord sent his word, and healed them: and they were saved from their destruction. ℣. O that men would therefore praise the Lord for his goodness: and declare the wonders that he doeth for the children of men.}{Alleluia, alleluia. ℣. Praise the Lord, all ye Angels of his: praise him, all his hosts. Alleluia.}

\readingcitation{Gospel}{John 2:1}
\lett{A}{t that time:} The third day there was a marriage in Cana of Galilee; and the mother of Jesus was there: And both Jesus was called, and his disciples, to the marriage. And when they wanted wine, the mother of Jesus saith unto him, They have no wine. Jesus saith unto her, Woman, what have I to do with thee? mine hour is not yet come. His mother saith unto the servants, Whatsoever he saith unto you, do it. And there were set there six waterpots of stone, after the manner of the purifying of the Jews, containing two or three firkins apiece. Jesus saith unto them, Fill the waterpots with water. And they filled them up to the brim. And he saith unto them, Draw out now, and bear unto the governor of the feast. And they bare it. When the ruler of the feast had tasted the water that was made wine, and knew not whence it was: (but the servants which drew the water knew;) the governor of the feast called the bridegroom, And saith unto him, Every man at the beginning doth set forth good wine; and when men have well drunk, then that which is worse: but thou hast kept the good wine until now. This beginning of miracles did Jesus in Cana of Galilee, and manifested forth his glory; and his disciples believed on him.

\offertory{O be joyful in God, all ye lands : sing praises unto the honour of his name: O come hither and hearken, all ye that fear God, and I will tell you what the Lord hath done for my soul, alleluia.}
\secret\label{EpiphanyIISecret}
\lett{S}{anctify,} O Lord, the gifts which we offer; and cleanse us from the defilements of our iniquities. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Secret is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{The Lord saith: Fill the water-pots with water, and bear unto the governor of the feast. When the ruler of the feast had tasted the water that was made wine, he saith unto the bridegroom: Thou hast kept the good wine until now. This beginning of miracles did Jesus before his disciples.}
\postcommunion\label{EpiphanyIIPostcommunion}
\lett{I}{ncrease} in us, O Lord, we beseech thee, the operation of thy power: that we, whom thou hast quickened with divine sacraments, may by thy bounty be made ready to receive the promises of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Postcommunion is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Third Sunday after Epiphany}\label{epiphany}
\feastday{Epiphany III}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{orship} God, all ye Angels of his: Sion heard, and rejoiced: and the daughters of Judah were glad \textit{Ps.} The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof.
\collect
\lett{A}{lmighty} and everlasting God, mercifully look upon our infirmities, and in all our dangers and necessities stretch forth thy right hand to help and defend us. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Collect is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 12:16}
\lett{B}{rethren:} Be not wise in your own conceits. Recompense to no man evil for evil. Provide things honest in the sight of all men. If it be possible, as much as lieth in you, live peaceably with all men. Dearly beloved, avenge not yourselves, but rather give place unto wrath: for it is written, Vengeance is mine; I will repay, saith the Lord. Therefore if thine enemy hunger, feed him; if he thirst, give him drink: for in so doing thou shalt heap coals of fire on his head. Be not overcome of evil, but overcome evil with good.
\gradall{The heathen shall fear thy name, O Lord, and all the kings of the earth thy majesty. ℣. When the Lord shall build up Sion, and when his glory shall appear.}{Alleluia, alleluia. ℣. The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof. Alleluia.}
\readingcitation{Gospel}{Matthew 8:1}
\lett{A}{t that time:} When Jesus was come down from the mountain, great multitudes followed him. And, behold, there came a leper and worshipped him, saying, Lord, if thou wilt, thou canst make me clean. And Jesus put forth his hand, and touched him, saying, I will; be thou clean. And immediately his leprosy was cleansed. And Jesus saith unto him, See thou tell no man; but go thy way, shew thyself to the priest, and offer the gift that Moses commanded, for a testimony unto them. And when Jesus was entered into Capernaum, there came unto him a centurion, beseeching him, And saying, Lord, my servant lieth at home sick of the palsy, grievously tormented. And Jesus saith unto him, I will come and heal him. The centurion answered and said, Lord, I am not worthy that thou shouldest come under my roof: but speak the word only, and my servant shall be healed. For I am a man under authority, having soldiers under me: and I say to this man, Go, and he goeth; and to another, Come, and he cometh; and to my servant, Do this, and he doeth it. When Jesus heard it, he marvelled, and said to them that followed, Verily I say unto you, I have not found so great faith, no, not in Israel And I say unto you, That many shall come from the east and west, and shall sit down with Abraham, and Isaac, and Jacob, in the kingdom of heaven. But the children of the kingdom shall be cast out into outer darkness: there shall be weeping and gnashing of teeth. And Jesus said unto the centurion, Go thy way; and as thou hast believed, so be it done unto thee. And his servant was healed in the selfsame hour.

\offertory{The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}
\secret
\lett{W}{e} beseech thee, O Lord, that this oblation may cleanse us from our sins: and sanctify thy servants both in body and mind for the celebration of this sacrifice. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Secret is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{All wondered at the gracious words which proceeded out of the mouth of God.}
\postcommunion
\lett{O}{Lord,} who dost suffer us to be made partakers of these great mysteries: we beseech thee; that thou wouldest vouchsafe to render us worthy to obtain the benefits of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Postcommunion is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Fourth Sunday after Epiphany}
\feastday{Epiphany IV}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{orship} God, all ye Angels of his: Sion heard, and rejoiced: and the daughters of Judah were glad \textit{Ps.} The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof.
\collect
\lett{O}{God,} who knowest us to be set in the midst of so many and great dangers, that by reason of the frailty of our nature we cannot always stand upright; Grant to us such strength and protection, as may support us in all dangers, and carry us through all temptations. Through.
\begin{rubric}
    If it be after 2 February, \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.\par
    Otherwise, \nth{2} Collect is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Collect is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 13:1}
\lett{B}{rethren:} Let every soul be subject unto the higher powers. For there is no power but of God: the powers that be are ordained of God. Whosoever therefore resisteth the power, resisteth the ordinance of God: and they that resist shall receive to themselves damnation. For rulers are not a terror to good works, but to the evil. Wilt thou then not be afraid of the power? do that which is good, and thou shalt have praise of the same: For he is the minister of God to thee for good. But if thou do that which is evil, be afraid; for he beareth not the sword in vain: for he is the minister of God, a revenger to execute wrath upon him that doeth evil. Wherefore ye must needs be subject, not only for wrath, but also for conscience sake. For for this cause pay ye tribute also: for they are God's ministers, attending continually upon this very thing. Render therefore to all their dues: tribute to whom tribute is due; custom to whom custom; fear to whom fear; honour to whom honour.

\gradall{The heathen shall fear thy name, O Lord, and all the kings of the earth thy majesty. ℣. When the Lord shall build up Sion, and when his glory shall appear.}{Alleluia, alleluia. ℣. The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof. Alleluia.}

\readingcitation{Gospel}{Matthew 8:23}
\lett{A}{t that time:} When Jesus was entered into a ship, his disciples followed him. And, behold, there arose a great tempest in the sea, insomuch that the ship was covered with the waves: but he was asleep. And his disciples came to him, and awoke him, saying, Lord, save us: we perish. And he saith unto them, Why are ye fearful, O ye of little faith? Then he arose, and rebuked the winds and the sea; and there was a great calm. But the men marvelled, saying, What manner of man is this, that even the winds and the sea obey him! And when he was come to the other side into the country of the Gergesenes, there met him two possessed with devils, coming out of the tombs, exceeding fierce, so that no man might pass by that way. And, behold, they cried out, saying, What have we to do with thee, Jesus, thou Son of God? art thou come hither to torment us before the time? And there was a good way off from them an herd of many swine feeding. So the devils besought him, saying, If thou cast us out, suffer us to go away into the herd of swine. And he said unto them, Go. And when they were come out, they went into the herd of swine: and, behold, the whole herd of swine ran violently down a steep place into the sea, and perished in the waters. And they that kept them fled, and went their ways into the city, and told every thing, and what was befallen to the possessed of the devils. And, behold, the whole city came out to meet Jesus: and when they saw him, they besought him that he would depart out of their coasts.

\offertory{The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}

%MANUAL ADJUSTMENT:
\clearpage

\secret
\lett{G}{rant,} we beseech thee, almighty God: that the gift now offered in this sacrifice may cleanse our frailty from all evil and evermore defend us. Through.
\begin{rubric}
    If it be after 2 February, \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.\par
    Otherwise, \nth{2} Secret is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Secret is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{All wondered at the gracious words which proceeded out of the mouth of God.}
\postcommunion
\lett{M}{ay} these thy gifts, O God, deliver us from earthly desires: and ever renew us with heavenly nourishment. Through.
\begin{rubric}
    If it be after 2 February, \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.\par
    Otherwise, \nth{2} Postcommunion is \emph{Of St. Mary after Christmas} (p. \pageref{SPMaryPostChristmas}) \& \nth{3} Postcommunion is \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Fifth Sunday after Epiphany}\label{EpiphanyV}
\feastday{Epiphany V}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{orship} God, all ye Angels of his: Sion heard, and rejoiced: and the daughters of Judah were glad \textit{Ps.} The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof.
\collect\label{EpiphanyVCollect}
\lett{O}{Lord,} we beseech thee to keep thy Church and household continually in thy true religion; that they who do lean only upon the hope of thy heavenly grace may evermore be defended by thy mighty power. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\readingcitation{Epistle}{Colossians 3:12}
\lett{B}{rethren:} Put on therefore, as the elect of God, holy and beloved, bowels of mercies, kindness, humbleness of mind, meekness, longsuffering; Forbearing one another, and forgiving one another, if any man have a quarrel against any: even as Christ forgave you, so also do ye. And above all these things put on charity, which is the bond of perfectness. And let the peace of God rule in your hearts, to the which also ye are called in one body; and be ye thankful. Let the word of Christ dwell in you richly in all wisdom; teaching and admonishing one another in psalms and hymns and spiritual songs, singing with grace in your hearts to the Lord. And whatsoever ye do in word or deed, do all in the name of the Lord Jesus, giving thanks to God and the Father by Our Lord Jesus Christ.

%MANUAL ADJUSTMENT:

\gradual{The heathen shall fear thy name, O Lord, and all the kings of the earth thy majesty. ℣. When the Lord shall build up Sion, and when his glory shall appear.}

\alleluia{Alleluia, alleluia. ℣. The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof. Alleluia.}

\readingcitation{Gospel}{Matthew 13:24}
\lett{A}{t that time:} Jesus spake this parable unto the crowds: The kingdom of heaven is likened unto a man which sowed good seed in his field: But while men slept, his enemy came and sowed tares among the wheat, and went his way. But when the blade was sprung up, and brought forth fruit, then appeared the tares also. So the servants of the householder came and said unto him, Sir, didst not thou sow good seed in thy field? from whence then hath it tares? He said unto them, An enemy hath done this. The servants said unto him, Wilt thou then that we go and gather them up? But he said, Nay; lest while ye gather up the tares, ye root up also the wheat with them. Let both grow together until the harvest: and in the time of harvest I will say to the reapers, Gather ye together first the tares, and bind them in bundles to burn them: but gather the wheat into my barn.

\offertory{The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}
\secret\label{EpiphanyVSecret}
\lett{W}{e} offer unto thee, O Lord, sacrifices of propitiation: that thou wouldest mercifully absolve our offences and thyself direct the hearts of those that go astray. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\communion{All wondered at the gracious words which proceeded out of the mouth of God.}
\postcommunion\label{EpiphanyVPostcommunion}
\lett{W}{e} beseech thee, almighty God: that as in these mysteries we have received the pledge of our salvation, so we may be brought unto the fulfilment of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subby{Sixth Sunday after Epiphany}
\feastday{Epiphany VI}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{orship} God, all ye Angels of his: Sion heard, and rejoiced: and the daughters of Judah were glad \textit{Ps.} The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof.

\collect
\lett{O}{God,} whose blessed Son was manifested that he might destroy the works of the devil, and make us the sons of God, and heirs of eternal life; Grant us, we beseech thee, that, having this hope, we may purify ourselves, even as he is pure; that, when he shall appear again with power and great glory, we may be made like unto him in his eternal and glorious kingdom; where with thee, O Father, and thee, O Holy Ghost, he liveth and reigneth ever, one God, world without end. \textit{Amen.}
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 John 3:1}
\lett{D}{early beloved:} Behold, what manner of love the Father hath bestowed upon us, that we should be called the sons of God: therefore the world knoweth us not, because it knew him not. Beloved, now are we the sons of God, and it doth not yet appear what we shall be: but we know that, when he shall appear, we shall be like him; for we shall see him as he is. And every man that hath this hope in him purifieth himself, even as he is pure. Whosoever committeth sin transgresseth also the law: for sin is the transgression of the law. And ye know that he was manifested to take away our sins; and in him is no sin. Whosoever abideth in him sinneth not: whosoever sinneth hath not seen him, neither known him. Little children, let no man deceive you: he that doeth righteousness is righteous, even as he is righteous. He that committeth sin is of the devil; for the devil sinneth from the beginning. For this purpose the Son of God was manifested, that he might destroy the works of the devil.

\gradall{The heathen shall fear thy name, Lord, and all the kings of the earth thy majesty. ℣. When the Lord shall build up Sion, and when his glory shall appear.}{Alleluia, alleluia. ℣. The Lord is King, the earth may be glad thereof: yea, the multitude of the isles may be glad thereof. Alleluia.}

\readingcitation{Gospel}{Matthew 24:23}
\lett{A}{t that time:} Jesus spake unto his disciples: If any man shall say unto you, Lo, here is Christ, or there; believe it not. For there shall arise false Christs, and false prophets, and shall shew great signs and wonders; insomuch that, if it were possible, they shall deceive the very elect. Behold, I have told you before. Wherefore if they shall say unto you, Behold, he is in the desert; go not forth: behold, he is in the secret chambers; believe it not. For as the lightning cometh out of the east, and shineth even unto the west; so shall also the coming of the Son of man be. For wheresoever the carcase is, there will the eagles be gathered together. Immediately after the tribulation of those days shall the sun be darkened, and the moon shall not give her light, and the stars shall fall from heaven, and the powers of the heavens shall be shaken: And then shall appear the sign of the Son of man in heaven: and then shall all the tribes of the earth mourn, and they shall see the Son of man coming in the clouds of heaven with power and great glory. And he shall send his angels with a great sound of a trumpet, and they shall gather together his elect from the four winds, from one end of heaven to the other.

\offertory{The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}
\secret
\lett{O}{God,} we pray thee, that this oblation may cleanse and regenerate, govern, and defend us. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\communion{All wondered at the gracious words which proceeded out of the mouth of God.}
\postcommunion
\lett{O}{Lord,} who hast fulfilled us with thy heavenly delights: we beseech thee; that we may ever earnestly seek after those things whereby we truly live. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\begin{rubric}
    If there be more than six Sundays of Epiphany before Septuagesima, then the upcoming unused Sundays after Trinity (p. \pageref{trinity}) are used, the Introit, Gradual, Alleluia, Offertory, and Communion being of Epiphany VI.
\end{rubric}

\subby{Septuagesima Sunday}
\feastday{Septuagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} sorrows of death compassed me, the pains of hell came about me: and in my tribulation I called upon the Lord, and he heard my voice out of his holy temple. \textit{Ps.} I will love thee, O Lord, my strength: the Lord is my stony rock, my fortress, and my saviour.
\begin{rubric}
    \emph{Gloria in excelsis} is not said in Masses of the Season from this Sunday until Wednesday in Holy Week, inclusive, either on Sundays or week-days.
\end{rubric}
\collect
\lett{O}{Lord,} we beseech thee favourably to hear the prayers of thy people; that we, who are justly punished for our offences, may be mercifully delivered by thy goodness, for the glory of thy Name; through Jesus Christ our Saviour, who liveth and reigneth with thee and the Holy Ghost ever, one God, world without end. \textit{Amen.}
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\readingcitation{Epistle}{1 Corinthians 9:24}
\lett{B}{rethren:} Know ye not that they which run in a race run all, but one receiveth the prize? So run, that ye may obtain. And every man that striveth for the mastery is temperate in all things. Now they do it to obtain a corruptible crown; but we an incorruptible. I therefore so run, not as uncertainly; so fight I, not as one that beateth the air: But I keep under my body, and bring it into subjection: lest that by any means, when I have preached to others, I myself should be a castaway.

\gradtr{A refuge in due time of trouble: they that know thee will put their trust in thee: for thou, Lord, never failest them that seek thee. ℣. For the poor shall not alway be forgotten: the patient abiding of the meek shall not perish for ever: up, Lord, and let not man have the upper hand.}{Out of the deep have I called unto thee, O Lord: Lord, hear my voice. ℣. O let thine ears consider well the prayer of thy servant. ℣. If thou, Lord, wilt be extreme to mark what is done amiss: O Lord, who may abide it? ℣. For there is mercy with thee, and because of thy law, I have waited for thee, O Lord.}
\begin{rubric}
    From Septuagesima until the Tuesday after Quinquagesima Sunday inclusive, when the Sunday Mass is resumed on week-days, the Tract is not said, but only the Gradual.
\end{rubric}

\readingcitation{Gospel}{Matthew 20:1}
\lett{A}{t that time:} Jesus spake unto his disciples this parable: The kingdom of heaven is like unto a man that is an householder, which went out early in the morning to hire labourers into his vineyard. And when he had agreed with the labourers for a penny a day, he sent them into his vineyard. And he went out about the third hour, and saw others standing idle in the marketplace, And said unto them; Go ye also into the vineyard, and whatsoever is right I will give you. And they went their way. Again he went out about the sixth and ninth hour, and did likewise. And about the eleventh hour he went out, and found others standing idle, and saith unto them, Why stand ye here all the day idle? They say unto him, Because no man hath hired us. He saith unto them, Go ye also into the vineyard; and whatsoever is right, that shall ye receive. So when even was come, the lord of the vineyard saith unto his steward, Call the labourers, and give them their hire, beginning from the last unto the first. And when they came that were hired about the eleventh hour, they received every man a penny. But when the first came, they supposed that they should have received more; and they likewise received every man a penny. And when they had received it, they murmured against the goodman of the house, Saying, These last have wrought but one hour, and thou hast made them equal unto us, which have borne the burden and heat of the day. But he answered one of them, and said, Friend, I do thee no wrong: didst not thou agree with me for a penny? Take that thine is, and go thy way: I will give unto this last, even as unto thee. Is it not lawful for me to do what I will with mine own? Is thine eye evil, because I am good? So the last shall be first, and the first last: for many be called, but few chosen.
\offertory{It is a good thing to give thanks unto the Lord, and to sing praises unto thy name, O most Highest.}
\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us, whom thou dost cleanse by thy heavenly mysteries. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\communion{Shew thy servant the light of thy countenance, and save me for thy mercy's sake: let me not be confounded, O Lord, for I have called upon thee.}
\postcommunion
\lett{M}{ay} thy faithful people, O God, be strengthened by thy gifts: that they receiving the same may seek them the more, and seeking them may obtain them everlastingly. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Sexagesima Sunday}
\feastday{Sexagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{A}{rise,} O Lord, wherefore sleepest thou? Awake, and cast us not away for ever: wherefore hidest thou thy countenance, and forgettest our trouble? Our belly cleaveth unto the ground: arise, O Lord, help us, and deliver us. \textit{Ps.} O God, we have heard with our ears: our fathers have told us.

\collect
\lett{O}{God,} who seest that we put not our trust in any thing that we do: mercifully grant; that by the protection of the Doctor of the Gentiles we may be defended against all adversity. Through.
%\lett{O}{Lord} God, who seest that we put not our trust in any thing that we do; Mercifully grant that by thy power we may be defended against all adversity. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{2 Corinthians 11:19}
\lett{B}{rethren:} Ye suffer fools gladly, seeing ye yourselves are wise. For ye suffer, if a man bring you into bondage, if a man devour you, if a man take of you, if a man exalt himself, if a man smite you on the face. I speak as concerning reproach, as though we had been weak. Howbeit whereinsoever any is bold, (I speak foolishly,) I am bold also. Are they Hebrews? so am I. Are they Israelites? so am I. Are they the seed of Abraham? so am I. Are they ministers of Christ? (I speak as a fool) I am more; in labours more abundant, in stripes above measure, in prisons more frequent, in deaths oft. Of the Jews five times received I forty stripes save one. Thrice was I beaten with rods, once was I stoned, thrice I suffered shipwreck, a night and a day I have been in the deep; In journeyings often, in perils of waters, in perils of robbers, in perils by mine own countrymen, in perils by the heathen, in perils in the city, in perils in the wilderness, in perils in the sea, in perils among false brethren; In weariness and painfulness, in watchings often, in hunger and thirst, in fastings often, in cold and nakedness. Beside those things that are without, that which cometh upon me daily, the care of all the churches. Who is weak, and I am not weak? who is offended, and I burn not? If I must needs glory, I will glory of the things which concern mine infirmities. The God and Father of our Lord Jesus Christ, which is blessed for evermore, knoweth that I lie not.

\gradtr{Let the nations know that thou, whose name is Jehovah: art only the Most Highest over all the earth. ℣. O my God, make them like unto a wheel, and as the stubble before the wind.}{Thou hast moved the land, O Lord, and divided it. ℣. Heal the sores thereof, for it shaketh. ℣. That they may triumph because of the truth: that thy beloved may be delivered.}
\readingcitation{Gospel}{Luke 8:4}
\lett{A}{t that time:} When much people were gathered together, and were come to Jesus out of every city, he spake by a parable: A sower went out to sow his seed: and as he sowed, some fell by the way side; and it was trodden down, and the fowls of the air devoured it. And some fell upon a rock; and as soon as it was sprung up, it withered away, because it lacked moisture. And some fell among thorns; and the thorns sprang up with it, and choked it. And other fell on good ground, and sprang up, and bare fruit an hundredfold. And when he had said these things, he cried, He that hath ears to hear, let him hear. And his disciples asked him, saying, What might this parable be? And he said, Unto you it is given to know the mysteries of the kingdom of God: but to others in parables; that seeing they might not see, and hearing they might not understand. Now the parable is this: The seed is the word of God. Those by the way side are they that hear; then cometh the devil, and taketh away the word out of their hearts, lest they should believe and be saved. They on the rock are they, which, when they hear, receive the word with joy; and these have no root, which for a while believe, and in time of temptation fall away. And that which fell among thorns are they, which, when they have heard, go forth, and are choked with cares and riches and pleasures of this life, and bring no fruit to perfection. But that on the good ground are they, which in an honest and good heart, having heard the word, keep it, and bring forth fruit with patience.
\offertory{O hold thou up my goings in thy paths, that my footsteps slip not: incline thine ear to me, and hearken unto my words: shew thy marvellous loving-kindness, O Lord, thou that art the Saviour of them which put their trust in thee.}

\secret
\lett{M}{ay} the sacrifice which we offer unto thee, O Lord, evermore quicken and defend us. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{I will go unto the altar of God, even unto the God of my joy and gladness.}

\postcommunion
\lett{W}{e} humbly beseech thee, almighty God: that as thou dost refresh us with thy sacraments, so thou wouldest vouchsafe unto us to do thee worthy and acceptable service. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Quinquagesima Sunday}
\feastday{Quinquagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{e} thou my strong rock, and house of defence, that thou mayest save me: for thou art my strong rock, and my castle: be thou also my guide, and lead me for thy name's sake. \textit{Ps.} In thee, O Lord, have I put my trust, let me never be put to confusion: deliver me in thy righteousness and save me.
\collect
\lett{O}{Lord,} who hast taught us that all our doings without charity are nothing worth; Send thy Holy Ghost, and pour into our hearts that most excellent gift of charity, the very bond of peace and of all virtues, without which whosoever liveth is counted dead before thee. Grant this for thine only Son Jesus Christ's sake. Who liveth.
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 13:1}
\lett{B}{rethren:} Though I speak with the tongues of men and of angels, and have not charity, I am become as sounding brass, or a tinkling cymbal. And though I have the gift of prophecy, and understand all mysteries, and all knowledge; and though I have all faith, so that I could remove mountains, and have not charity, I am nothing. And though I bestow all my goods to feed the poor, and though I give my body to be burned, and have not charity, it profiteth me nothing. Charity suffereth long, and is kind; charity envieth not; charity vaunteth not itself, is not puffed up, Doth not behave itself unseemly, seeketh not her own, is not easily provoked, thinketh no evil; Rejoiceth not in iniquity, but rejoiceth in the truth; Beareth all things, believeth all things, hopeth all things, endureth all things. Charity never faileth: but whether there be prophecies, they shall fail; whether there be tongues, they shall cease; whether there be knowledge, it shall vanish away. For we know in part, and we prophesy in part. But when that which is perfect is come, then that which is in part shall be done away. When I was a child, I spake as a child, I understood as a child, I thought as a child: but when I became a man, I put away childish things. For now we see through a glass, darkly; but then face to face: now I know in part; but then shall I know even as also I am known. And now abideth faith, hope, charity, these three; but the greatest of these is charity.
\gradtr{Thou art the God that only doeth wonders: and hast declared thy power among the people. ℣. Thou hast mightily delivered thy people, even the sons of Jacob and Joseph.}{O be joyful in the Lord, all ye lands: serve the Lord with gladness. ℣. Come before his presence with a song: be ye sure that the Lord he is God. ℣. It is he that hath made us, and not we ourselves: we are his people, and the sheep of his pasture.}
\readingcitation{Gospel}{Luke 18:31}
\lett{A}{t that time:} Jesus took unto him the twelve, and said unto them, Behold, we go up to Jerusalem, and all things that are written by the prophets concerning the Son of man shall be accomplished. For he shall be delivered unto the Gentiles, and shall be mocked, and spitefully entreated, and spitted on: And they shall scourge him, and put him to death: and the third day he shall rise again. And they understood none of these things: and this saying was hid from them, neither knew they the things which were spoken. And it came to pass, that as he was come nigh unto Jericho, a certain blind man sat by the way side begging: And hearing the multitude pass by, he asked what it meant. And they told him, that Jesus of Nazareth passeth by. And he cried, saying, Jesus, thou Son of David, have mercy on me. And they which went before rebuked him, that he should hold his peace: but he cried so much the more, Thou Son of David, have mercy on me. And Jesus stood, and commanded him to be brought unto him: and when he was come near, he asked him, Saying, What wilt thou that I shall do unto thee? And he said, Lord, that I may receive my sight. And Jesus said unto him, Receive thy sight: thy faith hath saved thee. And immediately he received his sight, and followed him, glorifying God: and all the people, when they saw it, gave praise unto God.

\offertory{Blessed art thou, O Lord, O teach me thy statutes: with my lips have I been telling of all the judgments of thy mouth.}
\secret
\lett{W}{e} beseech thee, O Lord, that this oblation may cleanse us from our sins: and sanctify thy servants both in body and mind for the celebration of this sacrifice. Through.
\begin{rubric}
    \nth{2} Secret is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\communion{They did eat, and were well filled, for the Lord gave them the own desire: they were not disappointed of their lust.}
\postcommunion
\lett{W}{e} beseech thee, almighty God: that we who have received this heavenly food, may be thereby defended against all adversities. Through.
\begin{rubric}
    \nth{2} Postcommunion is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subby{Ash Wednesday}
\feastday{Ash Wednesday}
\begin{inhead}
{First Class Feria}
\end{inhead}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\begin{rubric}
    On Ash Wednesday, during the Order of Morning Prayer, the Great Litany (p. \pageref{litany}) should said in its appropriate place after the Third Collect followed by A Penitential Office (p. \pageref{penitential}).
\end{rubric}

\subbysub{Rite of Ashes}
\begin{rubric}
    Before Mass, ashes prepared from the branches of olive or other trees blessed the year before are blessed, as followeth.
\end{rubric}
\begin{rubric}
%None changed to Mattins:
Mattins being ended, the Priest vested in a violet cope, or without a chasuble, accompanied by the Ministers similarly vested, proceeds to bless ashes placed in some vessel on the Altar. And first the following Antiphon is sung by the choir, akin to the Introit.
\end{rubric}
\lett{H}{ear} us, O Lord, for thy loving-kindness is comfortable: turn thee unto us, O Lord, according to the multitude of thy mercies. \textit{Ps.} Save me, O God: for the waters are come in, even unto my soul.
\begin{rubric}
    Then the Priest at the Epistle corner, without turning to the people or disjoining his bands (which rule of joined hands holds in the Prayers of all benedictions), says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{A}{lmighty} and everlasting God, spare them that are penitent, be favourable to them that call upon thee: and vouchsafe to send thy holy Angel from heaven, to bl {\ding{64}} ess, and sanc {\ding{64}} tify these ashes, that they may be a wholesome medicine to all who humbly call upon thy holy name, who in their consciences by sin are accused, who in the sight of thy divine compassion bewail their faults, and earnestly and meekly implore thy most gracious loving-kindness: and grant that, by the invocation of thy most holy name; all who are sprinkled therewith for the redemption of their transgressions may receive healing of body and protection of soul. Through Christ, our Lord. \textit{Amen.}
\letuspray
\lett{O}{God,} who wouldest not the death of a sinner, but rather that he should repent: mercifully look upon the frailty of our mortal nature: and of thy goodness vouchsafe to bl {\ding{64}} ess these ashes which we purpose to have set upon our heads for the increase of humility, and the meriting of pardon: that we, acknowledging that we are but ashes, and that by reason of our vileness unto dust we shall return; may through thy mercy be found meet to receive forgiveness of all our sins and the rewards which thou hast promised to them that are penitent. Through Christ, our Lord. \textit{Amen.}
\letuspray
\lett{O}{God,} who dost turn unto them that abase themselves, and art favourable unto them that make satisfaction: incline thy merciful ear to our prayers; and graciously pour forth upon the heads of thy servants now to be sprinkled with these ashes the grace of thy benediction: that thou wouldest fill them with the spirit of contrition, and grant that those things which they have asked rightly they may obtain effectually; that those things which thou dost bestow may by thine appointment remain constant and unchanged for evermore. Through Christ, our Lord. \textit{Amen.}
\letuspray
\lett{A}{lmighty} and everlasting God, who upon the people of Nineveh, repenting in sackcloth and ashes, didst bestow the healing of thy pardon: mercifully grant; that we may so imitate them in outward fashion that we may be made like unto them in the obtaining of thy forgiveness. Through.
\begin{rubric}
    The Celebrant after putting incense into the thurible, thrice sprinkles the ashes with holy water, saying the Antiphon, \emph{Thou shalt purge me}, without singing and without the Psalm, and then thrice censes them.
\end{rubric}
\begin{rubric}
    The senior Priest amongst the Clergy approaches the Altar, and puts ashes on the bead of the Celebrant, who does not kneel. If there be no other Priest present, the Celebrant, turning to the Altar, puts ashes upon his own head, saying nothing, and forthwith the Choir sings the Antiphon,
\end{rubric}
\lett{L}{et} us change our raiment for sackcloth and ashes: let us fast and mourn before the Lord; for our God is exceedingly merciful to forgive us our sins.
\begin{inhead}
    or,
\end{inhead}
\lett{L}{et} the priests, the ministers of the Lord, weep between the porch and the altar, and let them say: Spare, O Lord, spare thy people: and shut not up, Lord, the mouths of them that praise thee.
\begin{rubric}
    The Responsory follows.
\end{rubric}
\lett{L}{et} us amend the sins that in our ignorance we have committed: lest the day of death come suddenly upon us, and we find no place for repentance, though we seek it.\par
\secondline{℟. Hear, O Lord, and have mercy: for we have sinned against thee.}
\thirdline{℣. Help us, O God of our salvation: and for the glory of thy name deliver us, O Lord.}
℟. Hear, O Lord, and have mercy: for we have sinned against thee.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. Hear, O Lord, and have mercy: for we have sinned against thee.\par
\begin{rubric}
    While the Antiphons and Responsory are in singing, the Priest with uncovered head first puts ashes on the senior Priest from whom be has received them, and then on the Ministers being vested and kneeling before the Altar, saying:
\end{rubric}
\begin{center}
    \textsc{Remember, O man, that dust thou art, and unto dust shalt thou return.}
\end{center}
\begin{rubric}
    Then come the others, first the Clergy in order then the people: and kneeling before the Altar receive the ashes severally from the Priest, as has been said of the Ministers.
\end{rubric}
\begin{rubric}
    The imposition of ashes being finished, the Priest says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{G}{rant} to us, O Lord, to put on the armour of our Christian warfare with holy fastings: that we who are about to fight against all spiritual wickedness, may be defended by the power of abstinence. Through Christ, our Lord. \textit{Amen.}
\begin{rubric}
    The Mass is then said.
\end{rubric}
\introit\label{AshWednesdayMass}
\lett{T}{hou} hast mercy upon all, O Lord, and hatest nothing that thou hast created: and winkest at the sins of men, because they should amend, and sparest them: for thou art the Lord our God. \textit{Ps.} Be merciful unto me, O God, be merciful unto me: for my soul trusteth in thee.
\collect\label{AshWednesdayMassCollect}
\lett{A}{lmighty} and everlasting God, who hatest nothing that thou hast made, and dost forgive the sins of all those who are penitent; Create and make in us new and contrite hearts, that we, worthily lamenting our sins and acknowledging our wretchedness, may obtain of thee, the God of all mercy, perfect remission and forgiveness. Through.
\begin{rubric}
%BCP Rubric:
This Collect is to be said every day in Lent, after the Collect appointed for the day, until Palm Sunday.
\end{rubric}
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{Joel 2:12}
\lett{T}{hus} saith the Lord: Turn ye even to me with all your heart, and with fasting, and with weeping, and with mourning: And rend your heart, and not your garments, and turn unto the \divineName{Lord} your God: for he is gracious and merciful, slow to anger, and of great kindness, and repenteth him of the evil. Who knoweth if he will return and repent, and leave a blessing behind him; even a meat offering and a drink offering unto the \divineName{Lord} your God? Blow the trumpet in Zion, sanctify a fast, call a solemn assembly: Gather the people, sanctify the congregation, assemble the elders, gather the children, and those that suck the breasts: let the bridegroom go forth of his chamber, and the bride out of her closet. Let the priests, the ministers of the \divineName{Lord}, weep between the porch and the altar, and let them say, Spare thy people, O \divineName{Lord}, and give not thine heritage to reproach, that the heathen should rule over them: wherefore should they say among the people, Where is their God?
\gradual{Be merciful unto me, O God, be merciful unto me: for my soul trusteth in thee. ℣. He shall send from heaven, and save me from the reproof of him that would eat me up.}

\begin{rubric}
    The following Tract is said in the Masses of Monday, Wednesday, and Friday---up to Monday in Holy Week, inclusive---except on Ember Wednesday.
\end{rubric}

\tract{O Lord, deal not with us after our sins: nor reward us according to our wickednesses. ℣. O Lord, remember not our old sins: but have mercy upon us, and that soon, for we are come to great misery. \inrub{Here genuflect.} ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us: and be merciful unto our sins, for thy name's sake.}
\readingcitation{Gospel}{Matthew 6:16}
\lett{A}{t that time:} Jesus spake unto his disciples: When ye fast, be not, as the hypocrites, of a sad countenance: for they disfigure their faces, that they may appear unto men to fast. Verily I say unto you, They have their reward. But thou, when thou fastest, anoint thine head, and wash thy face; That thou appear not unto men to fast, but unto thy Father which is in secret: and thy Father, which seeth in secret, shall reward thee openly. Lay not up for yourselves treasures upon earth, where moth and rust doth corrupt, and where thieves break through and steal: But lay up for yourselves treasures in heaven, where neither moth nor rust doth corrupt, and where thieves do not break through nor steal: For where your treasure is, there will your heart be also.
\offertory{I will magnify thee, O Lord, for thou hast set me up, and not made my foes to triumph over me: O Lord, I cried unto thee, and thou hast healed me.}
\secret\label{AshWednesdayMassSecret}
\lett{G}{rant,} we beseech thee, O Lord, that like as we by these our gifts do celebrate the institution of this venerable sacrament: so we may by thee be enabled worthily to offer the same. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{He who doth meditate on the law of the Lord day and night, will bring forth his fruit in due season.}
\postcommunion\label{AshWednesdayMassPostcommunion}
\lett{L}{et} the sacraments, O Lord, which we have received, avail for our succour: that these our fasts may be acceptable unto thee, and profitable unto us for our healing. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\begin{rubric}
    The Priest then prays the following:
\end{rubric}
\lett{L}{ook} down in mercy, O Lord, upon those who bow before thy majesty: that they who have been refreshed by the divine gift, may evermore be nourished by thy heavenly succour. Through.
\begin{rubric}
This manner of saying the Prayer over the people is observed in Masses of the Feria only until the Wednesday in Holy Week inclusive.
\end{rubric}
%This rubric is from the 1892 and 1928 American Prayer Books, reflecting their simpler lectionary. It is fitting for the simplicity which marks the English tradition. However, given this is a great deviation from the Missal which has propers for those days, it is the opinion of the editor that this rubric will probably not make it into the final edition.
\begin{rubric}
    The Propers of Ash Wednesday shall serve for every day after, unto the next Sunday, except upon a Feast Day.
\end{rubric}

\subby{First Sunday of Lent}
\feastday{Lent I}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{e} shall call upon me, and I will hearken unto him: I will deliver him, and bring him to honour: with length of days will I satisfy him. \textit{Ps.} Whoso dwelleth under the defence of the most High, shall abide under the shadow of the Almighty.
\collect
\lett{O}{Lord,} who for our sake didst fast forty days and forty nights; Give us grace to use such abstinence, that, our flesh being subdued to the Spirit, we may ever obey thy godly motions in righteousness, and true holiness, to thy honour and glory, who livest and reignest with the Father and the Holy Ghost, one God, world without end. \textit{Amen.}
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{2 Corinthians 6:1}
\lett{B}{rethren:} We then, as workers together with him, beseech you also that ye receive not the grace of God in vain. (For he saith, I have heard thee in a time accepted, and in the day of salvation have I succoured thee: behold, now is the accepted time; behold, now is the day of salvation.) Giving no offence in any thing, that the ministry be not blamed: But in all things approving ourselves as the ministers of God, in much patience, in afflictions, in necessities, in distresses, In stripes, in imprisonments, in tumults, in labours, in watchings, in fastings; By pureness, by knowledge, by longsuffering, by kindness, by the Holy Ghost, by love unfeigned, By the word of truth, by the power of God, by the armour of righteousness on the right hand and on the left, By honour and dishonour, by evil report and good report: as deceivers, and yet true; As unknown, and yet well known; as dying, and, behold, we live; as chastened, and not killed; As sorrowful, yet alway rejoicing; as poor, yet making many rich; as having nothing, and yet possessing all things.
\gradtr{God shall give his Angels charge over thee, to keep thee in all thy ways. ℣. They shall bear thee in their hands, that thou hurt not thy foot against a stone.}{Whoso dwelleth under the defence of the most High, shall abide under the shadow of the Almighty. ℣. I will say unto the Lord: Thou art my hope and my stronghold: my God, in him will I trust. ℣. For he shall deliver thee from the snare of the hunter, and from the noisome pestilence. ℣. He shall defend thee under his wings, and thou shalt be safe under his feathers. ℣. His faithfulness and truth shall be thy shield and buckler: thou shalt not be afraid for any terror by night. ℣. Nor for the arrow that flieth by day, for the pestilence that walketh in darkness, nor for the sickness that destroyeth in the noon-day. ℣. A thousand shall fall beside thee, and ten thousand at thy right hand: but it shall not come nigh thee. ℣. For he shall give his Angels charge over thee, to keep thee in all thy ways. ℣. They shall bear thee in their hands: that thou hurt not thy foot against a stone. ℣. Thou shalt go upon the lion and adder, the young lion and the dragon shalt thou tread under thy feet. ℣. Because he hath set his love upon me, therefore will I deliver him: I will set him up, because he hath known my name. ℣. He shall call upon me, and I will hear him: yea, I am with him in trouble. ℣. I will deliver him and bring him to honour: with long life will I satisfy him, and shew him my salvation.}
\readingcitation{Gospel}{Matthew 4:1}
\lett{A}{t that time:} Jesus was led up of the Spirit into the wilderness to be tempted of the devil. And when he had fasted forty days and forty nights, he was afterward an hungred. And when the tempter came to him, he said, If thou be the Son of God, command that these stones be made bread. But he answered and said, It is written, Man shall not live by bread alone, but by every word that proceedeth out of the mouth of God. Then the devil taketh him up into the holy city, and setteth him on a pinnacle of the temple, And saith unto him, If thou be the Son of God, cast thyself down: for it is written, He shall give his angels charge concerning thee: and in their hands they shall bear thee up, lest at any time thou dash thy foot against a stone. Jesus said unto him, It is written again, Thou shalt not tempt the Lord thy God. Again, the devil taketh him up into an exceeding high mountain, and sheweth him all the kingdoms of the world, and the glory of them; And saith unto him, All these things will I give thee, if thou wilt fall down and worship me. Then saith Jesus unto him, Get thee hence, Satan: for it is written, Thou shalt worship the Lord thy God, and him only shalt thou serve. Then the devil leaveth him, and, behold, angels came and ministered unto him.
\offertory{The Lord shall defend thee under his wings, and thou shalt be safe under his feathers: his faithfulness and truth shall be thy shield and buckler.}
\secret
\lett{W}{e} solemnly offer unto thee, O Lord, this sacrifice in the beginning of Lent, beseeching thee: that, while we abstain from carnal feastings, we may likewise refrain from hurtful pleasures. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{The Lord shall defend thee under his wings, and thou shalt be safe under his feathers: his faithfulness and truth shall be thy shield and buckler.}
\postcommunion
\lett{M}{ay} the holy partaking of thy sacrament, O Lord, in such wise restore us: that we, being cleansed from our former nature, may enter into the fellowship of this saving mystery. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\subby{Ember Wednesday in Lent}
\feastday{Lenten Emberday}
\begin{inhead}
{Second Class Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Wednesday}

\properantiphonfix

\introit
\lett{C}{all} to remembrance thy tender mercies, O Lord, and thy loving kindnesses, which have been ever of old: neither let our enemies triumph over us: deliver us, O God of Israel, out of all our troubles. \textit{Ps.} Unto thee, O Lord, will I lift up my soul: my God, I have put my trust in thee, O let me not be confounded.
\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{W}{e} beseech thee, O Lord, mercifully hear our prayers: and stretch forth the right hand of thy majesty to be our defence against all things that may hurt us. Through.
\readingcitation{Epistle}{Exodus 24:12}
\lett{I}{n those days:} The \divineName{Lord} said unto Moses, Come up to me into the mount, and be there: and I will give thee tables of stone, and a law, and commandments which I have written; that thou mayest teach them. And Moses rose up, and his minister Joshua: and Moses went up into the mount of God. And he said unto the elders, Tarry ye here for us, until we come again unto you: and, behold, Aaron and Hur are with you: if any man have any matters to do, let him come unto them. And Moses went up into the mount, and a cloud covered the mount. And the glory of the \divineName{Lord} abode upon mount Sinai, and the cloud covered it six days: and the seventh day he called unto Moses out of the midst of the cloud. And the sight of the glory of the \divineName{Lord} was like devouring fire on the top of the mount in the eyes of the children of Israel. And Moses went into the midst of the cloud, and gat him up into the mount: and Moses was in the mount forty days and forty nights.
\gradual{The sorrows of my heart are enlarged: O Lord, bring thou me out of my troubles. ℣. Look upon my adversity and misery: and forgive me all my sin.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{W}{e} beseech thee, O Lord, graciously to regard the devout prayers of thy people: that they who by abstinence do mortify the body, may in their minds be refreshed by the fruit of good works. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{1 Kings 19:3}
\lett{I}{n those days:} Elijah came to Beersheba, which belongeth to Judah, and left his servant there. But he himself went a day's journey into the wilderness, and came and sat down under a juniper tree: and he requested for himself that he might die; and said, It is enough; now, O \divineName{Lord}, take away my life; for I am not better than my fathers. And as he lay and slept under a juniper tree, behold, then an angel touched him, and said unto him, Arise and eat. And he looked, and, behold, there was a cake baken on the coals, and a cruse of water at his head. And he did eat and drink, and laid him down again. And the angel of the \divineName{Lord} came again the second time, and touched him, and said, Arise and eat; because the journey is too great for thee. And he arose, and did eat and drink, and went in the strength of that meat forty days and forty nights unto Horeb the mount of God.
\tract{O Lord, bring thou me out of my troubles: look upon my adversity and misery: and forgive me all my sin. ℣. Unto thee, O Lord, will I lift up my soul: my God, I have put my trust in thee: O let me not be confounded: neither let mine enemies triumph over me. ℣. For all they that hope in thee shall not be ashamed: but such as transgress without a cause shall be put to confusion.}
\readingcitation{Gospel}{Matthew 12:38}
\lett{A}{t that time:} Certain of the scribes and of the Pharisees answered Jesus, saying, Master, we would see a sign from thee. But he answered and said unto them, An evil and adulterous generation seeketh after a sign; and there shall no sign be given to it, but the sign of the prophet Jonas: For as Jonas was three days and three nights in the whale's belly; so shall the Son of man be three days and three nights in the heart of the earth. The men of Nineveh shall rise in judgment with this generation, and shall condemn it: because they repented at the preaching of Jonas; and, behold, a greater than Jonas is here. The queen of the south shall rise up in the judgment with this generation, and shall condemn it: for she came from the uttermost parts of the earth to hear the wisdom of Solomon; and, behold, a greater than Solomon is here. When the unclean spirit is gone out of a man, he walketh through dry places, seeking rest, and findeth none. Then he saith, I will return into my house from whence I came out; and when he is come, he findeth it empty, swept, and garnished. Then goeth he, and taketh with himself seven other spirits more wicked than himself, and they enter in and dwell there: and the last state of that man is worse than the first. Even so shall it be also unto this wicked generation. While he yet talked to the people, behold, his mother and his brethren stood without, desiring to speak with him. Then one said unto him, Behold, thy mother and thy brethren stand without, desiring to speak with thee. But he answered and said unto him that told him, Who is my mother? and who are my brethren? And he stretched forth his hand toward his disciples, and said, Behold my mother and my brethren! For whosoever shall do the will of my Father which is in heaven, the same is my brother, and sister, and mother.
\offertory{My delight shall be in thy commandments, which I have loved exceedingly: my hand also will I lift up to thy commandments which I have loved.}
\secret
\lett{W}{e} offer unto thee, O Lord, sacrifices of propitation: that thou wouldest mercifully absolve our offences, and thyself direct the hearts of them that go astray. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{Consider my meditation: O hearken thou unto the voice of my calling, my King, and my God: for unto thee, O Lord, will I make my prayer.}
\postcommunion
\lett{O}{Lord,} let the partaking of thy sacrament both cleanse us from our secret faults; and deliver us from the snares of our enemies. Through.

\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\lett{O}{Lord,} we beseech thee, illumine our minds with the light of thy brightness: that we may be able to perceive those things which we ought to do; and have strength to do those things which be right. Through.

%MANUAL ADJUSTMENT:
%\clearpage

\subby{Ember Friday in Lent}
\feastday{Lenten Emberday}
\begin{inhead}
{Second Class Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Friday}

\properantiphonfix

\introit
\lett{O}{Lord,} bring thou me out of my troubles: look upon my adversity and misery, and forgive me all my sin. \textit{Ps.} Unto thee, O Lord, will I lift up my soul; my God, I have put my trust in thee, O let me not be confounded.
\collect
\lett{B}{e} favourable, O Lord, to thy people: and mercifully comfort with thy gracious help those whom thou dost make to do thee godly service. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{Ezekiel 18:20}
\lett{T}{hus saith the Lord God:} The soul that sinneth, it shall die. The son shall not bear the iniquity of the father, neither shall the father bear the iniquity of the son: the righteousness of the righteous shall be upon him, and the wickedness of the wicked shall be upon him. But if the wicked will turn from all his sins that he hath committed, and keep all my statutes, and do that which is lawful and right, he shall surely live, he shall not die. All his transgressions that he hath committed, they shall not be mentioned unto him: in his righteousness that he hath done he shall live. Have I any pleasure at all that the wicked should die? saith the Lord \divineName{God}: and not that he should return from his ways, and live? But when the righteous turneth away from his righteousness, and committeth iniquity, and doeth according to all the abominations that the wicked man doeth, shall he live? All his righteousness that he hath done shall not be mentioned: in his trespass that he hath trespassed, and in his sin that he hath sinned, in them shall he die. Yet ye say, The way of the Lord is not equal. Hear now, O house of Israel; Is not my way equal? are not your ways unequal? When a righteous man turneth away from his righteousness, and committeth iniquity, and dieth in them; for his iniquity that he hath done shall he die. Again, when the wicked man turneth away from his wickedness that he hath committed, and doeth that which is lawful and right, he shall save his soul alive. Because he considereth, and turneth away from all his transgressions that he hath committed, he shall surely live, he shall not die, saith the Lord almighty.
\gradtr{My God, save thy servant that putteth his trust in thee. ℣. Give ear, Lord, unto my prayer.}{O Lord, deal not with us after our sins: nor reward us according to our wickednesses. ℣. O Lord, remember not our old sins: but have mercy upon us, and that soon, for we are come to great misery. \inrub{Here genuflect} ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us: and be merciful unto our sins, for thy name's sake.}
\readingcitation{Gospel}{John 5:1}
\lett{A}{t that time:} There was a feast of the Jews; and Jesus went up to Jerusalem. Now there is at Jerusalem by the sheep market a pool, which is called in the Hebrew tongue Bethesda, having five porches. In these lay a great multitude of impotent folk, of blind, halt, withered, waiting for the moving of the water. For an angel went down at a certain season into the pool, and troubled the water: whosoever then first after the troubling of the water stepped in was made whole of whatsoever disease he had. And a certain man was there, which had an infirmity thirty and eight years. When Jesus saw him lie, and knew that he had been now a long time in that case, he saith unto him, Wilt thou be made whole? The impotent man answered him, Sir, I have no man, when the water is troubled, to put me into the pool: but while I am coming, another steppeth down before me. Jesus saith unto him, Rise, take up thy bed, and walk. And immediately the man was made whole, and took up his bed, and walked: and on the same day was the sabbath. The Jews therefore said unto him that was cured, It is the sabbath day: it is not lawful for thee to carry thy bed. He answered them, He that made me whole, the same said unto me, Take up thy bed, and walk. Then asked they him, What man is that which said unto thee, Take up thy bed, and walk? And he that was healed wist not who it was: for Jesus had conveyed himself away, a multitude being in that place. Afterward Jesus findeth him in the temple, and said unto him, Behold, thou art made whole: sin no more, lest a worse thing come unto thee. The man departed, and told the Jews that it was Jesus, which had made him whole.
\offertory{Praise the Lord, O my soul, and forget not all his benefits: who maketh thee young and lusty as an eagle.}
\secret
\lett{A}{ccept,} O Lord, we pray thee, these oblations of our humble service: and graciously sanctify these thy gifts. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{All mine enemies shall be confounded, and sore vexed: they shall be turned back, and put to shame suddenly.}
\postcommunion
\lett{M}{ay} the operation of this mystery, O Lord, avail for the cleansing of our sins, and for the fulfilment of our godly desires. Through.\\
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\lett{G}{raciously} hear us, O merciful God: and shew forth in our hearts the light of thy grace. Through.

\subby{Ember Saturday in Lent}
\feastday{Lenten Emberday}
\begin{inhead}
{Second Class Feria}
\end{inhead}
\fancyhead[RE,LO]{Ember Saturday}

\properantiphonfix

\introit
\lett{O}{let} my prayer enter into thy presence: incline thine ear, O Lord, unto my calling. \textit{Ps.} O Lord God of my salvation: I have cried day and night before thee.
\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{W}{e} beseech thee, O Lord, graciously look upon thy people: and mercifully turn away from them the scourges of thy wrath. Through.
\readingcitation{Epistle}{Deuteronomy 26:12}
\lett{I}{n those days:} Spake Moses unto the people saying: When thou hast made an end of tithing all the tithes of thine increase the third year, which is the year of tithing, and hast given it unto the Levite, the stranger, the fatherless, and the widow, that they may eat within thy gates, and be filled; Then thou shalt say before the \divineName{Lord} thy God, I have brought away the hallowed things out of mine house, and also have given them unto the Levite, and unto the stranger, to the fatherless, and to the widow, according to all thy commandments which thou hast commanded me: I have not transgressed thy commandments, neither have I forgotten them: I have not eaten thereof in my mourning, neither have I taken away ought thereof for any unclean use, nor given ought thereof for the dead: but I have hearkened to the voice of the \divineName{Lord} my God, and have done according to all that thou hast commanded me. Look down from thy holy habitation, from heaven, and bless thy people Israel, and the land which thou hast given us, as thou swarest unto our fathers, a land that floweth with milk and honey. This day the \divineName{Lord} thy God hath commanded thee to do these statutes and judgments: thou shalt therefore keep and do them with all thine heart, and with all thy soul. Thou hast avouched the \divineName{Lord} this day to be thy God, and to walk in his ways, and to keep his statutes, and his commandments, and his judgments, and to hearken unto his voice: And the \divineName{Lord} hath avouched thee this day to be his peculiar people, as he hath promised thee, and that thou shouldest keep all his commandments; And to make thee high above all nations which he hath made, in praise, and in name, and in honour; and that thou mayest be an holy people unto the \divineName{Lord} thy God, as he hath spoken.
\gradual{Be merciful, O Lord, unto our sins: wherefore do the heathen say: Where is now their God? ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{B}{ehold,} O God our defender: that we, who are oppressed by the weight of our misdeeds, may through the obtaining of thy mercy serve thee in freedom of spirit. Through.
\readingcitation{Epistle}{Deuteronomy 11:22}
\lett{I}{n those days:} Moses said unto the children of Israel: If ye shall diligently keep all these commandments which I command you, to do them, to love the \divineName{Lord} your God, to walk in all his ways, and to cleave unto him; Then will the \divineName{Lord} drive out all these nations from before you, and ye shall possess greater nations and mightier than yourselves. Every place whereon the soles of your feet shall tread shall be yours: from the wilderness and Lebanon, from the river, the river Euphrates, even unto the uttermost sea shall your coast be. There shall no man be able to stand before you: for the \divineName{Lord} your God shall lay the fear of you and the dread of you upon all the land that ye shall tread upon, as he hath said unto you, the Lord your God.
\gradual{Behold, O God our defender, and look upon thy servants. ℣. O Lord God of hosts, hear the prayers of thy servants.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{A}{ssist} us, O Lord, we beseech thee, in these our supplications: that by thy goodness we may be found worthy to be humble in prosperity and tranquil in adversity. Through.
\readingcitation{Epistle}{1 Maccabees 2:23}
\lett{I}{n those days:} The priests made a prayer whilst the sacrifice was consuming, I say, both the priests, and all the rest, Jonathan beginning, and the rest answering thereunto, as Neemias did. And the prayer was after this manner; O Lord, Lord God, Creator of all things, who art fearful and strong, and righteous, and merciful, and the only and gracious King, The only giver of all things, the only just, almighty, and everlasting, thou that deliverest Israel from all trouble, and didst choose the fathers, and sanctify them: Receive the sacrifice for thy whole people Israel, and preserve thine own portion, and sanctify it. Gather those together that are scattered from us, deliver them that serve among the heathen, look upon them that are despised and abhorred, and let the heathen know that thou art our God.
\gradual{Turn thee again, O Lord, at the last, and be gracious unto thy servants. ℣. Lord, thou hast been our refuge, from one generation to another.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{O}{Lord,} we beseech thee favourably to hear the prayers of thy people: that we, who are justly punished for our offences, may be mercifully delivered for the glory of thy name. Through.
\readingcitation{Epistle}{Ecclesiasticus 36:1}
%\lett{H}{ave} mercy upon us, O Lord God of all, and behold us: And send thy fear upon all the nations that seek not after thee. Lift up thy hand against the strange nations, and let them see thy power. As thou wast sanctified in us before them: so be thou magnified among them before us. And let them know thee, as we have known thee, that there is no God but only thou, O God. Shew new signs, and make other strange wonders: glorify thy hand and thy right arm, that they may set forth thy wondrous works. Raise up indignation, and pour out wrath: take away the adversary, and destroy the enemy. Make the time short, remember the covenant, and let them declare thy wonderful works, O Lord our God.
%RV:
\lett{H}{ave} mercy upon us, O Lord the God of all, and behold; And send thy fear upon all the nations: Lift up thy hand against the strange nations; And let them see thy mighty power. As thou wast sanctified in us before them, So be thou magnified in them before us. And let them know thee, as we also have known thee, That there is no God but only thou, O God. Shew new signs, and work divers wonders; Glorify thy hand and thy right arm. Raise up indignation, and pour out wrath; Take away the adversary, and destroy the enemy. Hasten the time, and remember the oath; And let them declare thy mighty works.

\gradual{Let my prayer be set forth in thy sight, O Lord, as the incense. ℣. Let the lifting up of my hands be an evening sacrifice.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{P}{revent} us, O Lord, we beseech thee, in all our doings with thy most gracious favour, and further us with thy continual help: that all our prayers and works may be begun, continued, and ended in thee. Through.
\readingcitation{Epistle}{Song of the Three Children 26}
\lett{I}{n those days:} The angel of the Lord came down into the oven together with Azarias and his fellows, and smote the flame of the fire out of the oven; And made the midst of the furnace as it had been a moist whistling wind, so that the fire touched them not at all, neither hurt nor troubled them. Then the three, as out of one mouth, praised, glorified, and blessed, God in the furnace, saying,
\begin{rubric}
    Here the response \emph{Thanks be to God}, is not made.
\end{rubric}
\subbysub{Hymn}
\lett{B}{lessed} art thou, O Lord God of our fathers. And worthy to be praised and glorious for ever.\par
\secondline{And blessed is the name of thy glory, which is holy. And worthy to be praised and glorious for ever.}
\thirdline{Blessed art thou in the holy temple of thy glory. And worthy to be praised and glorious for ever.}

Blessed art thou on the holy throne of thy kingdom. And worthy to be praised and glorious for ever.

Blessed art thou in the sceptre of thy Godhead. And worthy to be praised and glorious for ever.

Blessed art thou that sittest upon the Cherubim, and beholdest the depths. And worthy to be praised and glorious for ever.

Blessed art thou that walkest on the wings of the winds and on the waves of the sea. And worthy to be praised and glorious for ever.

Let all thine Angels and Saints bless thee. And let them praise thee and glorify thee for ever.

Let the heavens, the earth, the sea, and all that in them is bless thee. And let them praise thee and glorify thee for ever.

Glory be to the Father, and to the Son, and to the Holy Ghost. And worthy to be praised and glorious for ever.

As it was in the beginning, is now, and ever shall be: world without end. Amen. And worthy to be praised and glorious for ever.
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{O}{God,} who to the three children didst assuage the flames of fire: mercifully grant; that the flames of sin may not kindle upon us thy servants. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{1 Thessalonians 5:14}
\lett{B}{rethren:} We exhort you, brethren, warn them that are unruly, comfort the feebleminded, support the weak, be patient toward all men. See that none render evil for evil unto any man; but ever follow that which is good, both among yourselves, and to all men. Rejoice evermore. Pray without ceasing. In every thing give thanks: for this is the will of God in Christ Jesus concerning you. Quench not the Spirit. Despise not prophesyings. Prove all things; hold fast that which is good. Abstain from all appearance of evil. And the very God of peace sanctify you wholly; and I pray God your whole spirit and soul and body be preserved blameless unto the coming of our Lord Jesus Christ.
\tract{O praise the Lord all ye heathen: praise him, all ye nations. ℣. For his merciful kindness is ever more and more towards us: and the truth of the Lord endureth for ever.}
\readingcitation{Gospel}{Matthew 17:1}
\lett{A}{t that time:} Jesus taketh Peter, James, and John his brother, and bringeth them up into an high mountain apart, And was transfigured before them: and his face did shine as the sun, and his raiment was white as the light. And, behold, there appeared unto them Moses and Elias talking with him. Then answered Peter, and said unto Jesus, Lord, it is good for us to be here: if thou wilt, let us make here three tabernacles; one for thee, and one for Moses, and one for Elias. While he yet spake, behold, a bright cloud overshadowed them: and behold a voice out of the cloud, which said, This is my beloved Son, in whom I am well pleased; hear ye him. And when the disciples heard it, they fell on their face, and were sore afraid. And Jesus came and touched them, and said, Arise, and be not afraid. And when they had lifted up their eyes, they saw no man, save Jesus only. And as they came down from the mountain, Jesus charged them, saying, Tell the vision to no man, until the Son of man be risen again from the dead.
\offertory{O Lord God of my salvation, I have cried day and night before thee: let my prayer enter into thy presence, O Lord.}
\secret
\lett{S}{anctify,} we beseech thee, O Lord, our fasts by these present sacrifices: that those things, which our observance outwardly shews forth, it may inwardly effect. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{O Lord my God, in thee have I put my trust: save me from all them that persecute me, and deliver me.}
\postcommunion
\lett{A}{lmighty} God, let thy holy mysteries both heal our vices, and bestow upon us everlasting remedies. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\lett{M}{ay} thy faithful people, O God, be strengthened by the blessing which they desire: that they may never turn aside from thy will, and may evermore rejoice in thy bountiful goodness. Through.

\subby{Second Sunday of Lent}
\feastday{Lent II}
\begin{inhead}
{First Class Semidouble}
\end{inhead}
\fancyhead[RE,LO]{}

\properantiphonfix

\introit
\lett{C}{all} to remembrance thy tender mercies, O Lord, and thy loving kindnesses, which have been ever of old: neither let our enemies triumph over us: deliver us, O God of Israel, out of all our troubles. \textit{Ps.} Unto thee, O Lord, do I lift up my soul: my God, in thee have I trusted, let me not be confounded.
\collect
\lett{A}{lmighty} God, who seest that we have no power of ourselves to help ourselves; Keep us both outwardly in our bodies, and inwardly in our souls; that we may be defended from all adversities which may happen to the body, and from all evil thoughts which may assault and hurt the soul. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{1 Thessalonians 4:1}
\lett{B}{rethren:} we beseech you, and exhort you by the Lord Jesus, that as ye have received of us how ye ought to walk and to please God, so ye would abound more and more. For ye know what commandments we gave you by the Lord Jesus. For this is the will of God, even your sanctification, that ye should abstain from fornication: That every one of you should know how to possess his vessel in sanctification and honour; Not in the lust of concupiscence, even as the Gentiles which know not God That no man go beyond and defraud his brother in any matter: because that the Lord is the avenger of all such, as we also have forewarned you and testified. For God hath not called us unto uncleanness, but unto holiness. He therefore that despiseth, despiseth not man, but God, who hath also given unto us his holy Spirit.

\gradtr{The sorrows of my heart are enlarged: O bring thou me out of my troubles, O Lord. ℣. Look upon my adversity and misery: and forgive me all my sin.}{O give thanks unto the Lord, for he is gracious: and his mercy endureth for ever. ℣. Who can express the noble acts of the Lord: or shew forth all his praise? ℣. Blessed are they that alway keep judgment, and do righteousness. ℣. Remember us, O Lord, according to the favour that thou bearest unto thy people: O visit us with thy salvation.}

\readingcitation{Gospel}{Matthew 15:21}
\lett{A}{t that time:} Jesus went thence, and departed into the coasts of Tyre and Sidon. And, behold, a woman of Canaan came out of the same coasts, and cried unto him, saying, Have mercy on me, O Lord, thou Son of David; my daughter is grievously vexed with a devil. But he answered her not a word. And his disciples came and besought him, saying, Send her away; for she crieth after us. But he answered and said, I am not sent but unto the lost sheep of the house of Israel. Then came she and worshipped him, saying, Lord, help me. But he answered and said, It is not meet to take the children's bread, and to cast it to dogs. And she said, Truth, Lord: yet the dogs eat of the crumbs which fall from their masters' table. Then Jesus answered and said unto her, O woman, great is thy faith: be it unto thee even as thou wilt. And her daughter was made whole from that very hour.

\offertory{My delight shall be in thy commandments, which I have loved exceedingly: my hands also will I lift up unto thy commandments, which I have loved.}
\secret
\lett{W}{e} beseech thee, O Lord, mercifully to have respect unto these our sacrifices: that they may increase our devotion and set forward our salvation. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\vspace{-1ex}

\communion{Consider my meditation: O hearken thou unto the voice of my calling, my King and my God: for unto thee will I make my prayer, O Lord.}
\postcommunion
\lett{W}{e} humbly beseech thee, almighty God: that as thou dost refresh us with thy sacraments, so thou wouldest vouchsafe unto us to do thee worthy and acceptable service. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\subby{Third Sunday of Lent}
\feastday{Lent III}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{M}{ine} eyes are ever looking unto the Lord: for he shall pluck my feet out of the net: look thou upon me and have mercy upon me, for I am desolate, and in misery. \textit{Ps.} Unto thee, O Lord, do I lift up my soul; my God, in thee have I trusted: let me not be confounded.
\collect
\lett{W}{e} beseech thee, Almighty God, look upon the hearty desires of thy humble servants, and stretch forth the right hand of thy Majesty, to be our defence against all our enemies. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{Ephesians 5:1}
\lett{B}{rethren:} Be ye therefore followers of God, as dear children; And walk in love, as Christ also hath loved us, and hath given himself for us an offering and a sacrifice to God for a sweetsmelling savour. But fornication, and all uncleanness, or covetousness, let it not be once named among you, as becometh saints; Neither filthiness, nor foolish talking, nor jesting, which are not convenient: but rather giving of thanks. For this ye know, that no whoremonger, nor unclean person, nor covetous man, who is an idolater, hath any inheritance in the kingdom of Christ and of God. Let no man deceive you with vain words: for because of these things cometh the wrath of God upon the children of disobedience. Be not ye therefore partakers with them. For ye were sometimes darkness, but now are ye light in the Lord: walk as children of light: (For the fruit of the Spirit is in all goodness and righteousness and truth;) Proving what is acceptable unto the Lord. And have no fellowship with the unfruitful works of darkness, but rather reprove them. For it is a shame even to speak of those things which are done of them in secret. But all things that are reproved are made manifest by the light: for whatsoever doth make manifest is light. Wherefore he saith, Awake thou that sleepest, and arise from the dead, and Christ shall give thee light.

\gradual{Up, Lord, and let not man have the upper hand: let the heathen be judged in thy sight. ℣. While mine enemies are driven back: they shall fall and perish at thy presence.}

\tract{Unto thee lift I up mine eyes: O thou that dwellest in the heavens. ℣. Behold, even as the eyes of servants: look unto the hand of their masters. ℣. And as the eyes of a maiden unto the hand of her mistress: even so our eyes wait upon the Lord our God, until he have mercy upon us. ℣. Have mercy upon us, O Lord: have mercy upon us.}

\readingcitation{Gospel}{Luke 11:14}
\lett{A}{t that time:} Jesus was casting out a devil, and it was dumb. And it came to pass, when the devil was gone out, the dumb spake; and the people wondered. But some of them said, He casteth out devils through Beelzebub the chief of the devils. And others, tempting him, sought of him a sign from heaven. But he, knowing their thoughts, said unto them, Every kingdom divided against itself is brought to desolation; and a house divided against a house falleth. If Satan also be divided against himself, how shall his kingdom stand? because ye say that I cast out devils through Beelzebub. And if I by Beelzebub cast out devils, by whom do your sons cast them out? therefore shall they be your judges. But if I with the finger of God cast out devils, no doubt the kingdom of God is come upon you. When a strong man armed keepeth his palace, his goods are in peace: But when a stronger than he shall come upon him, and overcome him, he taketh from him all his armour wherein he trusted, and divideth his spoils. He that is not with me is against me: and he that gathereth not with me scattereth. When the unclean spirit is gone out of a man, he walketh through dry places, seeking rest; and finding none, he saith, I will return unto my house whence I came out. And when he cometh, he findeth it swept and garnished Then goeth he, and taketh to him seven other spirits more wicked than himself; and they enter in, and dwell there: and the last state of that man is worse than the first. And it came to pass, as he spake these things, a certain woman of the company lifted up her voice, and said unto him, Blessed is the womb that bare thee, and the paps which thou hast sucked. But he said, Yea rather, blessed are they that hear the word of God, and keep it.
\offertory{The statutes of the Lord are right, and rejoice the heart; his judgments also are sweeter than honey and the honey-comb: moreover, thy servant keepeth them.}
\secret
\lett{W}{e} beseech thee, O Lord, that this oblation may cleanse us from our sins: and sanctify thy servants both in body and mind for the celebration of this sacrifice. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{The sparrow hath found her an house, and the swallow a nest, where she may lay her young : even thy altars, O Lord of hosts, my King and my God: blessed are they that dwell in thy house: they will be alway praising thee.}
\postcommunion
\lett{O}{Lord,} who sufferest us to be made partakers of so great a mystery: we beseech thee, that by thy mercy we may be delivered from all perils and iniquities. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\subsec{Fourth Sunday of Lent}
\feastday{Lent IV}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{R}{ejoice,} O Jerusalem, and come together, all ye that love her; rejoice for joy, all ye that have mourned: that ye may be glad, and be satisfied with the breasts of your consolation. \textit{Ps.} I was glad when they said unto me: We will go into the house of the Lord.
\collect
\lett{G}{rant,} we beseech thee, Almighty God, that we, who for our evil deeds do worthily deserve to be punished, by the comfort of thy grace may mercifully be relieved; through our Lord and Saviour Jesus Christ. Who liveth.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\readingcitation{Epistle}{Galatians 4:21}
\lett{B}{rethren:} tell me, ye that desire to be under the law, do ye not hear the law? For it is written, that Abraham had two sons, the one by a bondmaid, the other by a freewoman. But he who was of the bondwoman was born after the flesh; but he of the freewoman was by promise. Which things are an allegory: for these are the two covenants; the one from the mount Sinai, which gendereth to bondage, which is Agar. For this Agar is mount Sinai in Arabia, and answereth to Jerusalem which now is, and is in bondage with her children. But Jerusalem which is above is free, which is the mother of us all. For it is written, Rejoice, thou barren that bearest not; break forth and cry, thou that travailest not: for the desolate hath many more children than she which hath an husband. Now we, brethren, as Isaac was, are the children of promise. But as then he that was born after the flesh persecuted him that was born after the Spirit, even so it is now. Nevertheless what saith the scripture? Cast out the bondwoman and her son: for the son of the bondwoman shall not be heir with the son of the freewoman. So then, brethren, we are not children of the bondwoman, but of the free.

\gradtr{I was glad when they said unto me: We will go into the house of the Lord. ℣. Peace be within thy walls: and plenteousness within thy palaces.}{They that put their trust in the Lord shall be even as the mount Sion: which may not be removed, but standeth fast for ever. ℣. The hills stand about Jerusalem: even so standeth the Lord round about his people, from this time forth for evermore.}

\readingcitation{Gospel}{John 6:1}
\lett{A}{t that time:} Jesus went over the sea of Galilee, which is the sea of Tiberias. And a great multitude followed him, because they saw his miracles which he did on them that were diseased. And Jesus went up into a mountain, and there he sat with his disciples. And the passover, a feast of the Jews, was nigh. When Jesus then lifted up his eyes, and saw a great company come unto him, he saith unto Philip, Whence shall we buy bread, that these may eat? And this he said to prove him: for he himself knew what he would do. Philip answered him, Two hundred pennyworth of bread is not sufficient for them, that every one of them may take a little. One of his disciples, Andrew, Simon Peter's brother, saith unto him, There is a lad here, which hath five barley loaves, and two small fishes: but what are they among so many? And Jesus said, Make the men sit down. Now there was much grass in the place. So the men sat down, in number about five thousand. And Jesus took the loaves; and when he had given thanks, he distributed to the disciples, and the disciples to them that were set down; and likewise of the fishes as much as they would. When they were filled, he said unto his disciples, Gather up the fragments that remain, that nothing be lost. Therefore they gathered them together, and filled twelve baskets with the fragments of the five barley loaves, which remained over and above unto them that had eaten. Then those men, when they had seen the miracle that Jesus did, said, This is of a truth that prophet that should come into the world.

\offertory{O praise the Lord, for he is gracious: O sing praises unto his name, for he is lovely: whatsoever he pleased, that did he in heaven, and in earth.}
\secret
\lett{W}{e} beseech thee, O Lord, mercifully to have respect unto these our sacrifices: that they may increase our devotion and set forward our salvation. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}
\communion{Jerusalem is built as a city, that is at unity in itself: for thither the tribes go up, even the tribes of the Lord, to give thanks unto thy name, O Lord.}
\postcommunion
\lett{O}{merciful} God, who never failest to fulfil us with thy holy mysteries: grant to us, we beseech thee, that we may ever approach them in lowliness and sincerity, and receive them in faithfulness of heart. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage

\subby{Passion Sunday}\label{PassionSunday}
\feastday{Passion Sunday}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\begin{rubric}
    From this Sunday until Maundy Thursday, inclusive, in Masses of the Season the Psalm, \emph{Give sentence}, is not said before the Confession nor \emph{Glory be} at the Introit and after the Psalm \emph{Lavabo}.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\introit
\lett{G}{ive} sentence with me, O God, and defend my cause against the ungodly people: O deliver me from the deceitful and wicked man: for thou art the God of my strength. \textit{Ps.} O send out thy light and thy truth: that they may lead me and bring me unto thy holy hill, and to thy dwelling.
\collect
\lett{W}{e} beseech thee, Almighty God, mercifully to look upon thy people; that by thy great goodness they may be governed and preserved evermore, both in body and soul. Through.
\begin{rubric}
    From today until Saturday before Low Sunday inclusive, after the Collect of the Day is said the Collect against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Hebrews 9:11}
\lett{B}{rethren:} Christ being come an high priest of good things to come, by a greater and more perfect tabernacle, not made with hands, that is to say, not of this building; Neither by the blood of goats and calves, but by his own blood he entered in once into the holy place, having obtained eternal redemption for us. For if the blood of bulls and of goats, and the ashes of an heifer sprinkling the unclean, sanctifieth to the purifying of the flesh: How much more shall the blood of Christ, who through the eternal Spirit offered himself without spot to God, purge your conscience from dead works to serve the living God? And for this cause he is the mediator of the new testament, that by means of death, for the redemption of the transgressions that were under the first testament, they which are called might receive the promise of eternal inheritance, in Christ Jesus our Lord.

%MANUAL ADJUSTMENT:
\gradual{Deliver me, O Lord, from mine enemies: teach me to do the thing that pleaseth thee. ℣. It is thou, Lord, that deliverest me from my cruel enemies: and settest me up above mine adversaries: thou shalt rid me from the wicked man.}

\tract{Many a time have they fought against me from my youth up. ℣. May Israel now say: yea, many a time have they vexed me from my youth up. ℣. But they have not prevailed against me: the plowers plowed upon my back. ℣. And made long furrows: but the righteous Lord hath hewn the snares of the ungodly in pieces.}

\readingcitation{Gospel}{John 8:46}
\lett{A}{t that time:} Jesus said unto the multitudes of the Jews: Which of you convinceth me of sin? And if I say the truth, why do ye not believe me? He that is of God heareth God's words: ye therefore hear them not, because ye are not of God. Then answered the Jews, and said unto him, Say we not well that thou art a Samaritan, and hast a devil? Jesus answered, I have not a devil; but I honour my Father, and ye do dishonour me. And I seek not mine own glory: there is one that seeketh and judgeth. Verily, verily, I say unto you, If a man keep my saying, he shall never see death. Then said the Jews unto him, Now we know that thou hast a devil. Abraham is dead, and the prophets; and thou sayest, If a man keep my saying, he shall never taste of death. Art thou greater than our father Abraham, which is dead? and the prophets are dead: whom makest thou thyself? Jesus answered, If I honour myself, my honour is nothing: it is my Father that honoureth me; of whom ye say, that he is your God: Yet ye have not known him; but I know him: and if I should say, I know him not, I shall be a liar like unto you: but I know him, and keep his saying. Your father Abraham rejoiced to see my day: and he saw it, and was glad. Then said the Jews unto him, Thou art not yet fifty years old, and hast thou seen Abraham? Jesus said unto them, Verily, verily, I say unto you, Before Abraham was, \textsc{I am}. Then took they up stones to cast at him: but Jesus hid himself, and went out of the temple.

\offertory{I will give thanks unto thee, O Lord, with my whole heart: O do well unto thy servant, that I may live, and keep thy word: quicken thou me, O Lord, according to thy word.}
\secret
\lett{W}{e} beseech thee, O Lord, that these our oblations may both loose the bonds of our iniquity, and obtain for us the gifts of thy loving kindness. Through.
\begin{rubric}
    The Secret against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}) is said.
\end{rubric}
\communion{This is my body, which is given for you: this cup is the new Testament in my blood, saith the Lord: this do ye, as oft as ye drink it, in remembrance of me.}
%\postcommunion

\subsubsec{Postcommunion}
\lett{A}{ssist} us, O Lord our God: that we, whom thou hast here refreshed with thy mysteries, may be defended by thy perpetual succour. Through.
\begin{rubric}
    The Postcommunion against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}) is said.
\end{rubric}

\subsec{Palm Sunday}\label{PalmSunday}
\feastday{Palm Sunday}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\subbysub{Blessing of Palms}
\begin{rubric}
In Choir, Mattins having been said, the sprinkling of water is performed in the usual manner. Then the Priest, vested in a violet Cope, or without the Chasuble, with the Ministers vested in like manner, proceeds to bless branches of palm and olive or other trees placed in the midst before the Altar, or at the Epistle corner. And first the Choir sing the Antiphon:
\end{rubric}
\antiphon{}{Hosanna to the Son of David: blessed is he that cometh in the name of the Lord. O King of Israel: Hosanna in the Highest.}
\begin{rubric}
    Then the Priest standing at the Epistle corner, without turning toward the people, sings with joined hands in the tone of the ferial Collect:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{O}{God,} whose love and service is our justice, multiply upon us the gifts of thy unspeakable grace: and like as by the death of thy Son thou hast given unto us the hope of our faith; so grant that by his resurrection we may be brought unto the country which we seek. Who liveth and reigneth with thee.
\begin{rubric}
    Then the Subdeacon in the accustomed place sings the following Lesson in the tone of the Epistle, and at the end kisses the hand of the Priest.
\end{rubric}
\readingcitation{Lesson}{Exodus 15:27}
\lett{I}{n those days:} The children of Israel came to Elim, where were twelve wells of water, and threescore and ten palm trees: and they encamped there by the waters. And they took their journey from Elim, and all the congregation of the children of Israel came unto the wilderness of Sin, which is between Elim and Sinai, on the fifteenth day of the second month after their departing out of the land of Egypt. And the whole congregation of the children of Israel murmured against Moses and Aaron in the wilderness: And the children of Israel said unto them, Would to God we had died by the hand of the \divineName{Lord} in the land of Egypt, when we sat by the flesh pots, and when we did eat bread to the full; for ye have brought us forth into this wilderness, to kill this whole assembly with hunger. Then said the \divineName{Lord} unto Moses, Behold, I will rain bread from heaven for you; and the people shall go out and gather a certain rate every day, that I may prove them, whether they will walk in my law, or no. And it shall come to pass, that on the sixth day they shall prepare that which they bring in; and it shall be twice as much as they gather daily. And Moses and Aaron said unto all the children of Israel, At even, then ye shall know that the \divineName{Lord} hath brought you out from the land of Egypt: And in the morning, then ye shall see the glory of the \divineName{Lord}.
\gradual{The chief priests and the Pharisees gathered a council, and said: What do we, for this man doeth many miracles? If we let him thus alone, all men will believe on him: * And the Romans shall come, and take away both our place and nation. ℟. And one of them named Caiaphas, being the high priest that same year, prophesied saying: It is expedient for us, that one man should die for the people, and that the whole nation perish not. Then from that day forth they took counsel together to put him to death, saying. * And the Romans shall come, and take away both our place and nation. ℟. On the Mount of Olives he prayed to his Father; Father, if it be possible, let this cup pass from me. * The spirit indeed is willing, but the flesh is weak; thy will be done. ℣. Watch and pray, that ye enter not into temptation. * The spirit indeed is willing, but the flesh is weak; thy will be done.}
\begin{rubric}
    During the singing of the \emph{℟.}, the Deacon places the Book of the Gospels on the Altar: and the Priest, the Deacon ministering the incense-boat, places incense in the censer. Then the Deacon says: \emph{Cleanse my heart}, takes the Book from the Altar, and asks a blessing from the Priest: then, while the Subdeacon holds the Book, he stands between two Acolytes holding lighted candles, signs the book and censes it, and sings the Gospel, as below, in the usual way: after which the Subdeacon carries the Book to be kissed by the Priest, who is censed by the Deacon.
\end{rubric}
\readingcitation{Gospel}{Matthew 21:1}
\lett{A}{t that time:} When Jesus drew nigh unto Jerusalem, and was come to Bethphage, unto the mount of Olives, then sent he two disciples, Saying unto them, Go into the village over against you, and straightway ye shall find an ass tied, and a colt with her: loose them, and bring them unto me. And if any man say ought unto you, ye shall say, The Lord hath need of them; and straightway he will send them. All this was done, that it might be fulfilled which was spoken by the prophet, saying, Tell ye the daughter of Sion, Behold, thy King cometh unto thee, meek, and sitting upon an ass, and a colt the foal of an ass. And the disciples went, and did as Jesus commanded them, And brought the ass, and the colt, and put on them their clothes, and they set him thereon. And a very great multitude spread their garments in the way; others cut down branches from the trees, and strawed them in the way. And the multitudes that went before, and that followed, cried, saying, Hosanna to the Son of David: Blessed is he that cometh in the name of the Lord.
\begin{rubric}
    After this the branches are blessed. The Priest, standing at the same Epistle corner, says in the tone of the ferial Collect:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{I}{ncrease,} O God, the faith of them that put their trust in thee, and graciously hear the prayers of thy humble servants; let thy manifold mercies descend upon us: and let these branches of palm or olive be ble {\ding{64}} ssed : and, as in a figure of the Church thou didst multiply Noah going forth from the ark, and Moses when he went out of Egypt with the children of Israel: so may we, bearing palms and branches of olive, go forth with good works to meet Christ: and through him enter into everlasting joy: Who liveth and reigneth with thee in the unity of the Holy Ghost, God: throughout all ages, world without end. \textit{Amen.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Lift up your hearts.}{℣. Sursum corda.}
\elcol{℟. We lift them up unto the Lord.}{℟. Habémus ad Dóminum.}
\elcol{℣. Let us give thanks unto our Lord God.}{℣. Grátias agámus Dómino, Deo nostro.}
\elcol{℟. It is meet and right so to do.}{℟. Dignum et justum est.}\par\noindent
It is very meet, right, and our bounden duty, that we should at all times, and in all places, give thanks unto thee: O Lord holy Father almighty, everlasting God: Who art glorified in the council of thy Saints. For thee thy creatures serve: and acknowledge thee their only maker and their God, and thee all thy works do praise, and thee thy Saints do bless. For with fearless voice, before the kings and rulers of this world, they confess the great name of thine only begotten Son. Before whom Angels and Archangels, Thrones and Dominations stand: and with all the host of the heavenly army, do sing the hymn of thy glory, evermore saying:
\begin{rubric}
    And the \emph{Sanctus} is sung by the Choir,
\end{rubric}
\elcol{\lett{H}{oly, Holy, Holy,} Lord God of hosts: Heaven and earth are full of thy glory. Hosanna in the highest. Blessed {\ding{64}} is he that cometh in the Name of the Lord. Hosanna in the Highest.}
{\lett{S}{anctus, Sanctus, Sanctus,} Dóminus, Deus Sábaoth. Pleni sunt c{\ae}li et terra glória tua. Hosánna in excélsis. Benedíctus, {\ding{64}} qui venit in nómine Dómini. Hosánna in excélsis.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{W}{e} beseech thee, O Lord holy Father almighty, everlasting God: that thou wouldest vouchsafe to bl {\ding{64}} ess and sanc {\ding{64}} tify this creature of olive, which thou hast bidden to spring from the wood of the tree, which likewise the dove, returning to the ark, did bear in her mouth: that all who shall receive thereof, may obtain for themselves protection both in body and soul: and let it be, O Lord, a remedy for our salvation, a sacrament
of thy grace. Through.
\letuspray
\lett{O}{God,} who dost gather together the things which are dispersed abroad, and dost preserve that which thou dost gather: who didst bless the people when they went forth bearing branches to meet Jesus: bl {\ding{64}} ess also these branches of palm and olive which thy servants receive with faith to the honour of thy name; that into whatsoever place they shall be brought, the dwellers in that place may obtain thy blessing: that all adversity being put to flight, thy right hand may protect those who are redeemed by Jesus Christ thy Son our Lord. Who liveth and reigneth with thee.
\letuspray
\lett{O}{God,} who by a wonderful order of thy providence wast pleased, even by means of things insensible, to shew forth the dispensation of our salvation: grant, we beseech thee; that the devout hearts of thy faithful people may profitably understand what is mystically signified by that deed when, on this day, the multitude, inspired by the heavenly light, went forth to meet the Redeemer, and strawed branches of palm and olive beneath his feet. The branches, then, of palm foreshadow his triumph over the prince of death; and the boughs of olive in a manner proclaim that the anointing of the Spirit is come. For the multitude rejoiced to know that even then it was prefigured: that our Redeemer, having compassion on the misery of mankind, was about to battle with the prince of death, and by his death to triumph. And therefore obediently they laid before him such things as should signify in him both the triumphs of his victory and the abundance of his mercy. Wherefore we likewise with sure faith, having in memory both the deed and the signification of the same, humbly beseech thee, O Lord holy Father almighty, everlasting God, through the same Jesus Christ our Lord: that in him and through him, whose members thou hast been pleased to make us, we may win the victory over the empire of death, and be found worthy to be partakers of his glorious resurrection. Who liveth and reigneth with thee.
\letuspray
\lett{O}{God,} who by an olive-branch didst command the dove to proclaim peace on earth: vouchsafe, we beseech thee; to sanctify with thy heavenly bless {\ding{64}} ing these branches of olive and other trees; that for all thy people they may be profitable unto salvation. Through Christ, our Lord. \textit{Amen.}
\letuspray
\lett{B}{l {\ding{64}} ess,} we beseech thee, O Lord, these branches of palm or olive: and grant; that as thy people on this day do outwardly perform these things to thy honour, so they, spiritually fulfilling the same with pure devotion, may win the victory over the enemy, and cleave steadfastly to every work of mercy. Through.
\begin{rubric}
    Then the Celebrant, after putting incense into the thurible, thrice sprinkles the boughs with holy water, saying the Antiphon, \emph{Thou shalt purge me}, without note and without the Psalm: and then thrice incenses them. Then he says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{O}{God,} who didst send forth thy Son, Jesus Christ our Lord, into the world for our salvation, that he might humble himself to be made like unto us, and call us back to thee: before whom, at his coming to Jerusalem for the fulfilling of the Scriptures, the multitude of the people that believed with faithful devotion strawed their raiment and palm branches in the way: grant, we beseech thee; that we may so prepare for him the way of faith, that, every stone of stumbling and rock of offence being done away, our works may blossom before thee with branches of righteousness: and that we may be found worthy to follow in his footsteps. Who liveth and reigneth with thee.
\begin{rubric}
    The blessing being completed, the senior of the Clergy goes up to the Altar and presents a blessed branch to the Celebrant, who does not genuflect, nor kiss the hand of him who gives it. Then the Celebrant, standing before the Altar, turns to the people and distributes the branches, first to the senior from whom he received his, then to the Deacon and Subdeacon vested, and to the other clergy severally in order, last of all to the laity: all genuflecting and kissing the branch and the hand of the Celebrant, except Prelates, if they are present. And when he begins to distribute, the Choir sings the following Antiphons:
\end{rubric}\par
\antiphon{}{The children of the Hebrews bearing branches of olive, went out to meet the Lord, crying out and saying: Hosanna in the highest.}\par
\antiphon{Another}{The children of the Hebrews spread their garments in the way and cried out, saying: Hosanna to the Son of David; blessed is he that cometh in the name of the Lord.}
\begin{rubric}
    If these are not sufficient, they are repeated until the distribution is finished. Then the Priest says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{A}{lmighty} and everlasting God, who didst ordain that our Lord Jesus Christ should ride upon the foal of an ass, and didst teach the multitudes of the people to straw in his way their raiment and branches of trees, and likewise to sing Hosanna in his praise: grant, we beseech thee; that we may by thee be enabled to follow their innocency, and be partakers of their merit. Through the same Christ, our Lord. \textit{Amen.}
\begin{rubric}
    Then the Procession is made. And first the Celebrant puts incense in the censer: and the Deacon, turning to the people, says: \emph{Let us proceed in peace}, and the Choir responds: \emph{In the name of Christ. Amen.}
\end{rubric}
\begin{rubric}
    The Thurifer goes first with the smoking censer: then the Subdeacon vested, carrying the Cross, between two Acolytes with lighted candles: the Clergy follow in order, and last of all the Celebrant with the Deacon on his left, all with branches in their hands: and the following Antiphons are sung, either all, or some, so long as the Procession lasts.
\end{rubric}\par
\antiphon{}{When the Lord drew nigh unto Jerusalem, he sent two of his disciples, saying: Go into the village over against you: and ye shall find the colt of an ass tied, whereon never man sat: loose him, and bring him unto me. If any man say ought unto you, ye shall say: The Lord hath need of him. They loose him and bring him to Jesus: and put on him their clothes, and set him thereon: some spread their garments in the way: others strawed branches from the trees: and they that followed, cried: Hosanna, blessed is he that cometh in the name of the Lord: blessed is the kingdom of our father David: Hosanna in the highest: Son of David, have mercy upon us.}\par
\antiphon{Another}{When the people heard that Jesus was coming to Jerusalem, they took branches of palm-trees: and went forth to meet him, and the children cried saying: This is he that should come for the salvation of the people. This is our salvation and the redemption of Israel. How great is he whom Thrones and Dominations go forth to meet! Fear not, daughter of Sion: behold thy King cometh to thee, sitting upon an ass' colt, as it is written. Hail, O King, Creator of the world, who art come to redeem us.}\par
\antiphon{Another}{Six days before the solemnity of the Passover, when the Lord came into the city of Jerusalem, the children went forth to meet him: and in their hands carried branches of palm trees, and cried with a loud voice, saying: Hosanna in the highest: blessed art thou, that art come in the multitude of thy mercies: Hosanna in the highest.}\par
\antiphon{Another}{The multitudes with flowers and palms go forth to meet the Redeemer: and render worthy homage to the triumphant conqueror: the Gentiles with their lips proclaim the Son of God: and in the praise of Christ their voices thunder through the skies: Hosanna in the highest.}\par
\antiphon{Another}{With the Angels and the children may we be found faithful, crying unto the vanquisher of death: Hosanna in the highest.}\par
\antiphon{Another}{A great multitude, that were come together to the feast, cried unto the Lord: Blessed is he that cometh in the name of the Lord: Hosanna in the highest.}
\begin{rubric}
    As the Procession turns back, two or four Cantors enter the church, and the door being shut, with their faces toward the Procession, sing the first verse of the following hymn, \emph{Gloria laus, etc.}; which verse is repeated by the Priest and others outside the church. Then those within sing the other verses which follow; either all or a part, as shall seem good: and those who are without respond to each verse: \emph{Gloria laus,} as at the ℣. at the beginning.
\end{rubric}
\needspace{4\baselineskip}
\lett{G}{lory} and honour and praise be to thee, our King and Redeemer,\par
\secondline{Christ, to whom children of old loved their Hosannas to raise. ℟. Glory and honour.}
\thirdline{Israel's Monarch art thou, and the glorious Offspring of David,}
Thou that approachest, a King, blest in the Name of the Lord. ℟. Glory and honour.\par
Glory to thee in the highest, the heavenly armies are singing: Glory to thee upon earth, man and creation reply. ℟. Glory and honour.\par
Met thee with palms in their hands that day the folk of the Hebrews,\par
We with our prayers and our hymns now to thy presence approach. ℟. Glory and honour.\par
They to thee proffered their praise for to herald thy dolorous Passion;\par
We to the King on his throne utter the jubilant hymn. ℟. Glory and honour.\par
They were then pleasing to thee, unto thee our devotion be pleasing;\par
Merciful King, kind King, who in all goodness art pleased. ℟. Glory and honour.\par
\begin{inhead}
    or
\end{inhead}
\lett{A}{ll} glory, laud, and honour To thee, Redeemer, King, To whom the lips of children Made sweet hosannas ring. ℟. All glory, laud.\par
\secondline{Thou art the King of Israel, Thou David's royal Son, Who in the Lord's name comest, The King and blessed One. ℟. All glory, laud.}
\thirdline{The company of Angels Are praising thee on high, And mortal men and all things Created make reply. ℟. All glory, laud.}
The people of the Hebrews With palms before thee went; Our praise and prayer and anthems Before thee we present. ℟. All glory, laud.\par
To thee before thy passion They sang their hymns of praise; To thee now high exalted Our melody we raise. ℟. All glory, laud.\par
Thou didst accept their praises, Accept the prayers we bring, Who in all good delightest, Thou good and gracious King. ℟. All glory, laud.\par
\begin{rubric}
    Then the Subdeacon with the foot of the Cross knocks at the door, which being straightway opened, the Procession enters the church singing:
\end{rubric}\par\noindent
℟. When the Lord entered the holy city, the children of the Hebrews, foretelling the resurrection of life, * With branches of palm cried: Hosanna in the highest. ℣. When the people heard that Jesus was coming to Jerusaiem, they went forth to meet him. With branches of palm cried: Hosanna in the highest.
\begin{rubric}
    Then Mass is celebrated, and branches are held in the hands, while the Passion and Gospel are sung only.
\end{rubric}
\introit
\lett{B}{e} not thou far from me, O Lord, thou art my succour, haste thee to help me: save me from the lion's mouth, thou hast heard me also from among the horns of the unicorns. \textit{Ps.} My God, my God, look upon me, why hast thou forsaken me: and art so far from my health, and from the words of my complaint? 
\collect
\lett{A}{lmighty} and everlasting God, who of thy tender love towards mankind, hast sent thy Son, our Saviour Jesus Christ, to take upon him our flesh, and to suffer death upon the cross, that all mankind should follow the example of his great humility; Mercifully grant, that we may both follow the example of his patience, and also be made partakers of his resurrection. Through the same.
\begin{rubric}
    This Collect only is said, even if a Commemoration has been made in the Office.
\end{rubric}
\begin{rubric}
    This Collect is to be said every day, after the Collect appointed for the day, until Good Friday.
\end{rubric}
\readingcitation{Epistle}{Philippians 2:5}
\lett{B}{rethren:} Let this mind be in you, which was also in Christ Jesus: Who, being in the form of God, thought it not robbery to be equal with God: But made himself of no reputation, and took upon him the form of a servant, and was made in the likeness of men: And being found in fashion as a man, he humbled himself, and became obedient unto death, even the death of the cross. Wherefore God also hath highly exalted him, and given him a name which is above every name: \inrub{genuflect} That at the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth; And that every tongue should confess that Jesus Christ is Lord, to the glory of God the Father.

\gradtr{Thou hast holden me by my right hand; thou shalt guide me with thy counsel: and after that receive me with glory. ℣. Truly God is loving unto Israel, even unto such as are of a clean heart; nevertheless, my feet were almost gone; my treadings had well nigh slipt: and why? I was grieved at the wicked, I do also see the ungodly in such prosperity.}{My God, my God, look upon me: why hast thou forsaken me? ℣. And art so far from my health, and from the words of my complaint? ℣. O my God, I cry in the daytime, but thou hearest not: and in the night season also I take no rest. ℣. And thou continuest holy, O thou worship of Israel. ℣. Our fathers hoped in thee: they trusted in thee, and thou didst deliver them. ℣. They called upon thee, and were holpen: they put their trust in thee, and were not confounded. ℣. But as for me, I am a worm, and no man: a very scorn of men, and the outcast of the people. ℣. All they that see me laugh me to scorn: they shoot out their lips, and shake their head, saying: ℣.He trusted in God that he would deliver him: let him deliver him, if he will have him. ℣. They stand staring and looking upon me: they part my garments among them, and cast lots upon my vesture. ℣. Save me from the lion's mouth: thou hast heard me also from among the horns of the unicorns. ℣. O praise the Lord, ye that fear him: magnify him, all ye of the seed of Jacob. ℣. They shall be counted unto the Lord for a generation: they shall come, and the heavens shall declare his righteousness. ℣. Unto a people that shall be born, whom the Lord hath made.}
\begin{rubric}
    The Passion begins at once: \emph{Cleanse my heart} is not said, nor is a blessing asked, and lights and incense are not carried; neither is \emph{The Lord be with you} said, nor the response, \emph{Glory be to thee, O Lord,} and the Celebrant or Deacon, while be announces The Passion of our Lord, does not sign the book nor himself. Which is observed likewise on other days, when the Passion is read.
\end{rubric}

\readingcitation{Gospel}{Matthew 26:1}
\lett{A}{t that time:} Jesus said unto his disciples: {\ding{64}} Ye know that after two days is the feast of the passover, and the Son of man is betrayed to be crucified. \textit{C.} Then assembled together the chief priests, and the scribes, and the elders of the people, unto the palace of the high priest, who was called Caiaphas, and consulted that they might take Jesus by subtilty, and kill him. But they said, \textit{S.} Not on the feast day, lest there be an uproar among the people.

\textit{C.} Now when Jesus was in Bethany, in the house of Simon the leper, there came unto him a woman having an alabaster box of very precious ointment, and poured it on his head, as he sat at meat. But when his disciples saw it, they had indignation, saying, \textit{S.} To what purpose is this waste? For this ointment might have been sold for much, and given to the poor. \textit{C.} When Jesus understood it, he said unto them, {\ding{64}} Why trouble ye the woman? for she hath wrought a good work upon me. For ye have the poor always with you; but me ye have not always. For in that she hath poured this ointment on my body, she did it for my burial. Verily I say unto you, Wheresoever this gospel shall be preached in the whole world, there shall also this, that this woman hath done, be told for a memorial of her.

\textit{C.} Then one of the twelve, called Judas Iscariot, went unto the chief priests, and said unto them, \textit{S.} What will ye give me, and I will deliver him unto you? \textit{C.} And they covenanted with him for thirty pieces of silver. And from that time he sought opportunity to betray him.

Now the first day of the feast of unleavened bread the disciples came to Jesus, saying unto him, \textit{S.} Where wilt thou that we prepare for thee to eat the passover? \textit{C.} And he said, {\ding{64}} Go into the city to such a man, and say unto him, The Master saith, My time is at hand; I will keep the passover at thy house with my disciples. \textit{C.} And the disciples did as Jesus had appointed them; and they made ready the passover.

Now when the even was come, he sat down with the twelve. And as they did eat, he said, {\ding{64}} Verily I say unto you, that one of you shall betray me. \textit{C.} And they were exceeding sorrowful, and began every one of them to say unto him, \textit{S.} Lord, is it I? \textit{C.} And he answered and said, {\ding{64}} He that dippeth his hand with me in the dish, the same shall betray me. The Son of man goeth as it is written of him: but woe unto that man by whom the Son of man is betrayed! it had been good for that man if he had not been born. \textit{C.} Then Judas, which betrayed him, answered and said, \textit{S.} Master, is it I? \textit{C.} He said unto him, {\ding{64}} Thou hast said.

\textit{C.} And as they were eating, Jesus took bread, and blessed it, and brake it, and gave it to the disciples, and said, {\ding{64}} Take, eat; this is my body. \textit{C.} And he took the cup, and gave thanks, and gave it to them, saying, {\ding{64}} Drink ye all of it; for this is my blood of the new testament, which is shed for many for the remission of sins. But I say unto you, I will not drink henceforth of this fruit of the vine, until that day when I drink it new with you in my Father’s kingdom.

\textit{C.} And when they had sung an hymn, they went out into the mount of Olives. Then saith Jesus unto them, {\ding{64}} All ye shall be offended because of me this night: for it is written, I will smite the shepherd, and the sheep of the flock shall be scattered abroad. But after I am risen again, I will go before you into Galilee. \textit{C.} Peter answered and said unto him, \textit{S.} Though all men shall be offended because of thee, yet will I never be offended. \textit{C.} Jesus said unto him, {\ding{64}} Verily I say unto thee, That this night, before the cock crow, thou shalt deny me thrice. \textit{C.} Peter said unto him, \textit{S.} Though I should die with thee, yet will I not deny thee. \textit{C.} Likewise also said all the disciples.

Then cometh Jesus with them unto a place called Gethsemane, and saith unto the disciples, {\ding{64}} Sit ye here, while I go and pray yonder. \textit{C.} And he took with him Peter and the two sons of Zebedee, and began to be sorrowful and very heavy. Then saith he unto them, {\ding{64}} My soul is exceeding sorrowful, even unto death: tarry ye here, and watch with me. \textit{C.} And he went a little further, and fell on his face, and prayed, saying, {\ding{64}} O my Father, if it be possible, let this cup pass from me: nevertheless not as I will, but as thou wilt.

\textit{C.} And he cometh unto the disciples, and findeth them asleep, and saith unto Peter, {\ding{64}} What, could ye not watch with me one hour? Watch and pray, that ye enter not into temptation: the spirit indeed is willing, but the flesh is weak.

\textit{C.} He went away again the second time, and prayed, saying, {\ding{64}} O my Father, if this cup may not pass away from me, except I drink it, thy will be done. \textit{C.} And he came and found them asleep again: for their eyes were heavy. And he left them, and went away again, and prayed the third time, saying the same words.

Then cometh he to his disciples, and saith unto them, {\ding{64}} Sleep on now, and take your rest: behold, the hour is at hand, and the Son of man is betrayed into the hands of sinners. Rise, let us be going: behold, he is at hand that doth betray me.

\textit{C.} And while he yet spake, lo, Judas, one of the twelve, came, and with him a great multitude with swords and staves, from the chief priests and elders of the people. Now he that betrayed him gave them a sign, saying, \textit{S.} Whomsoever I shall kiss, that same is he: hold him fast. \textit{C.} And forthwith he came to Jesus, and said, \textit{S.} Hail, master; \textit{C.} and kissed him. And Jesus said unto him, {\ding{64}} Friend, wherefore art thou come? \textit{C.} Then came they, and laid hands on Jesus, and took him.

And, behold, one of them which were with Jesus stretched out his hand, and drew his sword, and struck a servant of the high priest's, and smote off his ear. Then said Jesus unto him, {\ding{64}} Put up again thy sword into his place: for all they that take the sword shall perish with the sword. Thinkest thou that I cannot now pray to my Father, and he shall presently give me more than twelve legions of angels? But how then shall the scriptures be fulfilled, that thus it must be?

\textit{C.} In that same hour said Jesus to the multitudes, {\ding{64}} Are ye come out as against a thief with swords and staves for to take me? I sat daily with you teaching in the temple, and ye laid no hold on me. But all this was done, that the scriptures of the prophets might be fulfilled. \textit{C.} Then all the disciples forsook him, and fled.

And they that had laid hold on Jesus led him away to Caiaphas the high priest, where the scribes and the elders were assembled. But Peter followed him afar off unto the high priest's palace, and went in, and sat with the servants, to see the end. Now the chief priests, and elders, and all the council, sought false witness against Jesus, to put him to death; but found none: yea, though many false witnesses came, yet found they none. At the last came two false witnesses, and said, \textit{S.} This fellow said, I am able to destroy the temple of God, and to build it in three days. \textit{C.} And the high priest arose, and said unto him, \textit{S.} Answerest thou nothing? what is it which these witness against thee? \textit{C.} But Jesus held his peace. And the high priest answered and said unto him, \textit{S.} I adjure thee by the living God, that thou tell us whether thou be the Christ, the Son of God. \textit{C.} Jesus saith unto him, {\ding{64}} Thou hast said: nevertheless I say unto you, Hereafter shall ye see the Son of man sitting on the right hand of power, and coming in the clouds of heaven. \textit{C.} Then the high priest rent his clothes, saying, \textit{S.} He hath spoken blasphemy; what further need have we of witnesses? behold, now ye have heard his blasphemy. What think ye? \textit{C.} They answered and said, \textit{S.} He is guilty of death.

\textit{C.} Then did they spit in his face, and buffeted him; and others smote him with the palms of their hands, saying, \textit{S.} Prophesy unto us, thou Christ, Who is he that smote thee?

\textit{C.} Now Peter sat without in the palace: and a damsel came unto him, saying, \textit{S.} Thou also wast with Jesus of Galilee. \textit{C.} But he denied before them all, saying, \textit{S.} I know not what thou sayest. \textit{C.} And when he was gone out into the porch, another maid saw him, and said unto them that were there, \textit{S.} This fellow was also with Jesus of Nazareth. \textit{C.} And again he denied with an oath, \textit{S.} I do not know the man. \textit{C.} And after a while came unto him they that stood by, and said to Peter, \textit{S.} Surely thou also art one of them; for thy speech bewrayeth thee. \textit{C.} Then began he to curse and to swear, saying, \textit{S.} I know not the man. \textit{C.} And immediately the cock crew. And Peter remembered the word of Jesus, which said unto him, {\ding{64}} Before the cock crow, thou shalt deny me thrice. \textit{C.} And he went out, and wept bitterly.

When the morning was come, all the chief priests and elders of the people took counsel against Jesus to put him to death: and when they had bound him, they led him away, and delivered him to Pontius Pilate the governor.

Then Judas, which had betrayed him, when he saw that he was condemned, repented himself, and brought again the thirty pieces of silver to the chief priests and elders, saying, \textit{S.} I have sinned in that I have betrayed the innocent blood. \textit{C.} And they said, \textit{S.} What is that to us? see thou to that. \textit{C.} And he cast down the pieces of silver in the temple, and departed, and went and hanged himself. And the chief priests took the silver pieces, and said, \textit{S.} It is not lawful for to put them into the treasury, because it is the price of blood. \textit{C.} And they took counsel, and bought with them the potter's field, to bury strangers in. Wherefore that field was called, The field of blood, unto this day. Then was fulfilled that which was spoken by Jeremy the prophet, saying, And they took the thirty pieces of silver, the price of him that was valued, whom they of the children of Israel did value; and gave them for the potter's field, as the Lord appointed me.

And Jesus stood before the governor: and the governor asked him, saying, \textit{S.} Art thou the King of the Jews? \textit{C.} And Jesus said unto him, {\ding{64}} Thou sayest. \textit{C.} And when he was accused of the chief priests and elders, he answered nothing. Then said Pilate unto him, \textit{S.} Hearest thou not how many things they witness against thee? \textit{C.} And he answered him to never a word; insomuch that the governor marvelled greatly.

Now at that feast the governor was wont to release unto the people a prisoner, whom they would. And they had then a notable prisoner, called Barabbas. Therefore when they were gathered together, Pilate said unto them, \textit{S.} Whom will ye that I release unto you? Barabbas, or Jesus which is called Christ? \textit{C.} For he knew that for envy they had delivered him.

When he was set down on the judgment seat, his wife sent unto him, saying, \textit{S.} Have thou nothing to do with that just man: for I have suffered many things this day in a dream because of him. \textit{C.} But the chief priests and elders persuaded the multitude that they should ask Barabbas, and destroy Jesus. The governor answered and said unto them, \textit{S.} Whether of the twain will ye that I release unto you? \textit{C.} They said, \textit{S.} Barabbas. \textit{C.} Pilate saith unto them, \textit{S.} What shall I do then with Jesus which is called Christ? \textit{C.} They all say unto him, \textit{S.} Let him be crucified. \textit{C.} And the governor said, \textit{S.} Why, what evil hath he done? \textit{C.} But they cried out the more, saying, \textit{S.} Let him be crucified.

\textit{C.} When Pilate saw that he could prevail nothing, but that rather a tumult was made, he took water, and washed his hands before the multitude, saying, \textit{S.} I am innocent of the blood of this just person: see ye to it. \textit{C.} Then answered all the people, and said, \textit{S.} His blood be on us, and on our children. \textit{C.} Then released he Barabbas unto them: and when he had scourged Jesus, he delivered him to be crucified.

Then the soldiers of the governor took Jesus into the common hall, and gathered unto him the whole band of soldiers. And they stripped him, and put on him a scarlet robe. And when they had platted a crown of thorns, they put it upon his head, and a reed in his right hand: and they bowed the knee before him, and mocked him, saying, \textit{S.} Hail, King of the Jews! \textit{C.} And they spit upon him, and took the reed, and smote him on the head. And after that they had mocked him, they took the robe off from him, and put his own raiment on him, and led him away to crucify him. And as they came out, they found a man of Cyrene, Simon by name: him they compelled to bear his cross.

And when they were come unto a place called Golgotha, that is to say, a place of a skull, they gave him vinegar to drink mingled with gall: and when he had tasted thereof, he would not drink. And they crucified him, and parted his garments, casting lots: that it might be fulfilled which was spoken by the prophet, They parted my garments among them, and upon my vesture did they cast lots. And sitting down they watched him there; and set up over his head his accusation written, \textsc{This is Jesus the King of the Jews}. Then were there two thieves crucified with him, one on the right hand, and another on the left.

And they that passed by reviled him, wagging their heads, and saying, \textit{S.} Thou that destroyest the temple, and buildest it in three days, save thyself. If thou be the Son of God, come down from the cross. \textit{C.} Likewise also the chief priests mocking him, with the scribes and elders, said, \textit{S.} He saved others; himself he cannot save. If he be the King of Israel, let him now come down from the cross, and we will believe him. He trusted in God; let him deliver him now, if he will have him: for he said, I am the Son of God. \textit{C.} The thieves also, which were crucified with him, cast the same in his teeth.

Now from the sixth hour there was darkness over all the land unto the ninth hour. And about the ninth hour Jesus cried with a loud voice, saying, {\ding{64}} Eli, Eli, lama sabachthani? \textit{C.} that is to say, {\ding{64}} My God, my God, why hast thou forsaken me? \textit{C.} Some of them that stood there, when they heard that, said, \textit{S.} This man calleth for Elias. \textit{C.} And straightway one of them ran, and took a spunge, and filled it with vinegar, and put it on a reed, and gave him to drink. The rest said, \textit{S.} Let be, let us see whether Elias will come to save him.

\textit{C.} Jesus, when he had cried again with a loud voice, yielded up the ghost. \inrub{Here genuflect and pause awhile.} And, behold, the veil of the temple was rent in twain from the top to the bottom; and the earth did quake, and the rocks rent; and the graves were opened; and many bodies of the saints which slept arose, and came out of the graves after his resurrection, and went into the holy city, and appeared unto many. Now when the centurion, and they that were with him, watching Jesus, saw the earthquake, and those things that were done, they feared greatly, saying, \textit{S.} Truly this was the Son of God.

\textit{C.} And many women were there beholding afar off, which followed Jesus from Galilee, ministering unto him: among which was Mary Magdalene, and Mary the mother of James and Joses, and the mother of Zebedee's children.

When the even was come, there came a rich man of Arimath{\ae}a, named Joseph, who also himself was Jesus' disciple: he went to Pilate, and begged the body of Jesus. Then Pilate commanded the body to be delivered. And when Joseph had taken the body, he wrapped it in a clean linen cloth, and laid it in his own new tomb, which he had hewn out in the rock: and he rolled a great stone to the door of the sepulchre, and departed. And there was Mary Magdalene, and the other Mary, sitting over against the sepulchre.
\begin{rubric}
    Here \emph{Cleanse my heart} is said, a blessing is asked, incense is carried without lights, and the book is censed: \emph{The Lord be with you} is not said, and the Celebrant or Deacon does not sign the Book nor himself: and what follows is sung in the tone of the Gospel: at the conclusion of which the Celebrant kisses the Book and is censed. All of which is observed in the other Passions except on Good Friday.
\end{rubric}
Now the next day, that followed the day of the preparation, the chief priests and Pharisees came together unto Pilate, saying, Sir, we remember that that deceiver said, while he was yet alive, After three days I will rise again. Command therefore that the sepulchre be made sure until the third day, lest his disciples come by night, and steal him away, and say unto the people, He is risen from the dead: so the last error shall be worse than the first. Pilate said unto them, Ye have a watch: go your way, make it as sure as ye can. So they went, and made the sepulchre sure, sealing the stone, and setting a watch.
\offertory{Thy rebuke hath broken my heart; I am full of heaviness: I looked for some to have pity on me, but there was no man, neither found I any to comfort me: they gave me gall to eat: and when I was thirsty they gave me vinegar to drink.}

\secret
\lett{G}{rant,} we beseech thee, O God: that the gift which we offer in the sight of thy majesty may obtain for us grace to serve thee with devotion, and bring us in the end to everlasting felicity. Through
\communion{Father, if this cup may not pass away from me, except I drink it: thy will be done.}
\postcommunion
\lett{M}{ay} the operation of this mystery, O Lord: avail for the cleansing of our sins, and for the fulfilment of our godly desires. Through.

%\clearpage
\subby{Holy Week}
\feastday{Holy Week}
%\begin{rubric}
%    A supplement %The \emph{Book of Occasional Services} 
%    should be consulted for Holy Week. The Daily Office propers are here provided.
%\end{rubric}
%\begin{multicols}{2}
	\subsubsec{Holy Monday}
	\lett{A}{lmighty} God, whose most dear Son went not up to joy but first he suffered pain, and entered not into glory before he was crucified; Mercifully grant that we, walking in the way of the cross, may find it none other than the way of life and peace. Through the same.

	\subsubsec{Holy Tuesday}
	\lett{O}{Lord} God, whose blessed Son, our Saviour, gave his back to the smiters and hid not his face from shame; Grant us grace to take joyfully the sufferings of the present time, in full assurance of the glory that shall be revealed. Through the same.
	
	\subsubsec{Spy Wednesday}
    \lett{A}{ssist} us mercifully with thy help, O Lord God of our salvation; that we may enter with joy upon the meditation of those mighty acts, whereby thou hast given unto us life and immortality. Through.
	
    \subsubsec{Maundy Thursday}
    \lett{A}{lmighty} Father, whose dear Son, on the night before he suffered, did institute the Sacrament of his Body and Blood; Mercifully grant that we may thankfully receive the same in remembrance of him, who in these holy mysteries giveth us a pledge of life eternal; through the same thy Son Jesus Christ our Lord, who now liveth and reigneth with thee and the Holy Spirit ever, one God, world without end. \textit{Amen.}
    
    \subsubsec{Good Friday}
    \needspace{4\baselineskip}
    \lett{A}{lmighty} God, we beseech thee graciously to behold this thy family, for which our Lord Jesus Christ was contented to be betrayed and given up into the hands of wicked men, and to suffer death upon the cross; who now liveth and reigneth with thee and the Holy Ghost ever, one God, world without end. \textit{Amen.}
    \needspace{4\baselineskip}
    \lett{A}{lmighty} and everlasting God, by whose Spirit the whole body of the Church is governed and sanctified; Receive our supplications and prayers, which we offer before thee for all estates of men in thy holy Church, that every member of the same, in his vocation and ministry, may truly and godly serve thee. Through.
	\needspace{4\baselineskip}
    %1928 faithless revision reverted to 1892:
    \lett{O}{merciful} God, who hast made all men, and hatest nothing that thou hast made, nor desirest the death of a sinner, but rather that he should be converted and live; Have mercy upon all Jews, Turks, infidels, and heretics; and take from them all ignorance, hardness of heart, and contempt of thy Word; and so fetch them home, blessed Lord, to thy flock, that they may be saved among the remnant of the true Israelites, and be made one fold under one shepherd, Jesus Christ our Lord, who liveth and reigneth with thee and the Holy Spirit, one God, world without end. \textit{Amen.}
    
    %MANUAL ADJUSTMENT:
    \clearpage
    
    \subsubsec{Easter Even}
    \needspace{4\baselineskip}
    \lett{G}{rant,} O Lord, that as we are baptized into the death of thy blessed Son, our Saviour Jesus Christ, so by continual mortifying our corrupt affections we may be buried with him; and that through the grave, and gate of death, we may pass to our joyful resurrection; for his merits, who died, and was buried, and rose again for us, the same thy Son Jesus Christ our Lord. Who with thee.
%\end{multicols}

\subby{Easter Day}
\feastday{Easter Day}
\begin{inhead}
    {First Class Double, First Class Octave}
\end{inhead}

\subbysub{Pascha Nostrum}\label{pascha}
\begin{rubric}
    At Morning Prayer, instead of the \emph{Venite}, the following shall be said, and may be said throughout the Octave.
\end{rubric}
\needspace{4\baselineskip}
\lett{C}{hrist} our Passover is sacrificed for us : therefore let us keep the feast,
    
    \secondline{Not with old leaven, neither with the leaven of malice and wickedness : but with the unleavened bread of sincerity and truth.\vr{1 Cor. 5:7}}

\Needspace{4\baselineskip}
\lett{C}{hrist} being raised from the dead dieth no more : death hath no more dominion over him.
    
    \secondline{For in that he died, he died unto sin once : but in that he liveth, he liveth unto God.}
    \thirdline{Likewise reckon ye also yourselves to be dead indeed unto sin : but alive unto God through Jesus Christ our Lord.\vr{Rom. 6:9}}

\Needspace{4\baselineskip}
\lett{C}{hrist} is risen from the dead : and become the firstfruits of them that slept.

    \secondline{For since by man came death : by man came also the resurrection of the dead.}
    \thirdline{For as in Adam all die : even so in Christ shall all be made alive.\vr{1 Cor. 15:20}}
\par
℣. Glory be to the Father, and to the Son : and to the Holy Ghost;\par
℟. As it was in the beginning, is now, and ever shall be: world without end. Amen.

\introit
\lett{I}{am} risen, and am still with thee, alleluia: thou hast laid thine hand upon me, alleluia: thy knowledge is too wonderful, alleluia, alleluia. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine uprising.
\collect\label{EasterCollect}
\lett{A}{lmighty} God, who through thine only-begotten Son Jesus Christ hast overcome death, and opened unto us the gate of everlasting life; We humbly beseech thee that, as by thy special grace preventing us thou dost put into our minds good desires, so by thy continual help we may bring the same to good effect; through the same Jesus Christ our Lord, who liveth and reigneth with thee and the Holy Ghost ever, one God, world without end. \textit{Amen.}
\begin{rubric}
    This Collect is to be said daily throughout the Easter Octave.
\end{rubric}
%Given that the only prior Mass would be the Vigil, omitted from this Book, the 1928 propers for the second Mass are used.
\readingcitation{Epistle}{Colossians 3:1}
\lett{B}{rethren:} If ye then be risen with Christ, seek those things which are above, where Christ sitteth on the right hand of God. Set your affection on things above, not on things on the earth. For ye are dead, and your life is hid with Christ in God. When Christ, who is our life, shall appear, then shall ye also appear with him in glory. Mortify therefore your members which are upon the earth; fornication, uncleanness, inordinate affection, evil concupiscence, and covetousness, which is idolatry: For which things' sake the wrath of God cometh on the children of disobedience: In the which ye also walked some time, when ye lived in them.
\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. O give thanks unto the Lord, for he is gracious: and his mercy endureth for ever.}{Alleluia, alleluia. ℣. Christ our Passover is sacrificed for us.}
\subsubsec{Sequence}\label{PaschalSequence}
%\begin{multicols}{2}
\lett{C}{hristians,} to the Paschal Victim offer your thankful praises.\par
\secondline{A Lamb the sheep redeemeth: Christ, who only is sinless, reconcileth sinners to the Father.}
\thirdline{Death and life have contended in that combat stupendous: the prince of life, who died, reigns immortal.}
Speak, Mary, declaring what thou sawest wayfaring:\par
The tomb of Christ, who is living. The glory of Jesu's Resurrection;\par
Bright angels attesting, the shroud and napkin resting.\par
Yea, Christ my hope is arisen: to Galilee he goes before you.\par
Happy they who hear the witness, Mary's word believing above the tales of Jewry deceiving.\par
Christ indeed from death is risen, our new life obtaining: have mercy, victor King, ever reigning. Amen. Alleluia.
%\end{multicols}
\begin{rubric}
     The Sequence is said until Saturday before Low Sunday, inclusive.
\end{rubric}
\readingcitation{Gospel}{John 20:1}
\lett{A}{t that time:} On the first day of the week cometh Mary Magdalene early, when it was yet dark, unto the sepulchre, and seeth the stone taken away from the sepulchre. Then she runneth, and cometh to Simon Peter, and to the other disciple, whom Jesus loved, and saith unto them, They have taken away the Lord out of the sepulchre, and we know not where they have laid him. Peter therefore went forth, and that other disciple, and came to the sepulchre. So they ran both together: and the other disciple did outrun Peter, and came first to the sepulchre. And he stooping down, and looking in, saw the linen clothes lying; yet went he not in. Then cometh Simon Peter following him, and went into the sepulchre, and seeth the linen clothes lie, And the napkin, that was about his head, not lying with the linen clothes, but wrapped together in a place by itself. Then went in also that other disciple, which came first to the sepulchre, and he saw, and believed. For as yet they knew not the scripture, that he must rise again from the dead. Then the disciples went away again unto their own home.

\offertory{The earth trembled and was still, when God arose to judgment, alleluia.}

\secret\label{EasterSecret}
\lett{R}{eceive,} we beseech thee, O Lord, the prayers of thy people together with the offering of these sacrifices: that those things which we have begun in these paschal mysteries may, by thine operation, be profitable for our healing in eternity. Through.
\communion{Christ our Passover is sacrificed for us, alleluia: therefore let us keep the feast with the unleavened bread of sincerity and truth, alleluia, alleluia, alleluia.}
\postcommunion\label{EasterPostcommunion}
\lett{P}{our} forth upon us, O Lord, the Spirit of thy charity: that as thou hast fulfilled us with these paschal sacraments, so of thy goodness thou wouldest make us to be of one heart and mind. Through.


\subby{Easter Monday}
\feastday{Easter Monday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord hath brought you into a land flowing with milk and honey, alleluia: that the law of the Lord may alway be in your mouth, alleluia, alleluia. \textit{Ps.} O give thanks unto the Lord, and call upon his name: tell the people what things he hath done.

\collect
%\lett{O}{God,} whose blessed Son did manifest himself to his disciples in the breaking of bread; Open, we pray thee, the eyes of our faith, that we may behold thee in all thy works. Through the same.
\lett{O}{God,} who by the paschal solemnity hast bestowed healing upon the world: prosper, we beseech thee, thy people with thy heavenly gift; that they may be worthy to attain unto perfect freedom, and may likewise be profited unto everlasting life. Through.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}).
\end{rubric}

\readingcitation{Epistle}{Acts 10:34}
\lett{I}{n those days:} Peter opened his mouth, and said, Of a truth I perceive that God is no respecter of persons: But in every nation he that feareth him, and worketh righteousness, is accepted with him. The word which God sent unto the children of Israel, preaching peace by Jesus Christ: (he is Lord of all:) That word, I say, ye know, which was published throughout all Judaea, and began from Galilee, after the baptism which John preached; How God anointed Jesus of Nazareth with the Holy Ghost and with power: who went about doing good, and healing all that were oppressed of the devil; for God was with him. And we are witnesses of all things which he did both in the land of the Jews, and in Jerusalem; whom they slew and hanged on a tree: Him God raised up the third day, and shewed him openly; Not to all the people, but unto witnesses chosen before of God, even to us, who did eat and drink with him after he rose from the dead. And he commanded us to preach unto the people, and to testify that it is he which was ordained of God to be the Judge of quick and dead. To him give all the prophets witness, that through his name whosoever believeth in him shall receive remission of sins.

\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. Let Israel now confess, that he is gracious: and that his mercy endureth for ever.}{Alleluia, alleluia. ℣. The Angel of the Lord descended from heaven: and came and roiled back the stone and sat upon it.}
\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{Luke 24:13}
\lett{A}{t that time:} Behold, two of Jesus' disciples went that same day to a village called Emmaus, which was from Jerusalem about threescore furlongs. And they talked together of all these things which had happened. And it came to pass, that, while they communed together and reasoned, Jesus himself drew near, and went with them. But their eyes were holden that they should not know him. And he said unto them, What manner of communications are these that ye have one to another, as ye walk, and are sad? And the one of them, whose name was Cleopas, answering said unto him, Art thou only a stranger in Jerusalem, and hast not known the things which are come to pass there in these days? And he said unto them, What things? And they said unto him, Concerning Jesus of Nazareth, which was a prophet mighty in deed and word before God and all the people: And how the chief priests and our rulers delivered him to be condemned to death, and have crucified him. But we trusted that it had been he which should have redeemed Israel: and beside all this, to day is the third day since these things were done. Yea, and certain women also of our company made us astonished, which were early at the sepulchre; And when they found not his body, they came, saying, that they had also seen a vision of angels, which said that he was alive. And certain of them which were with us went to the sepulchre, and found it even so as the women had said: but him they saw not. Then he said unto them, O fools, and slow of heart to believe all that the prophets have spoken: Ought not Christ to have suffered these things, and to enter into his glory? And beginning at Moses and all the prophets, he expounded unto them in all the scriptures the things concerning himself. And they drew nigh unto the village, whither they went: and he made as though he would have gone further. But they constrained him, saying, Abide with us: for it is toward evening, and the day is far spent. And he went in to tarry with them. And it came to pass, as he sat at meat with them, he took bread, and blessed it, and brake, and gave to them. And their eyes were opened, and they knew him; and he vanished out of their sight. And they said one to another, Did not our heart burn within us, while he talked with us by the way, and while he opened to us the scriptures? And they rose up the same hour, and returned to Jerusalem, and found the eleven gathered together, and them that were with them Saying, The Lord is risen indeed, and hath appeared to Simon. And they told what things were done in the way, and how he was known of them in breaking of bread.

\offertory{The Angel of the Lord descended from heaven, and said unto the women: He whom ye seek is risen as he said, alleluia.}

\secret
\lett{R}{eceive,} we beseech thee, O Lord, the prayers of thy people together with the offering of these sacrifices; that those things which we have begun in these paschal mysteries may, by thine operation, be profitable for our healing in eternity. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}).
\end{rubric}

\communion{The Lord is risen, and hath appeared to Peter, alleluia.}

\postcommunion
\lett{P}{our} forth upon us, O Lord, the Spirit of thy charity: that as thou hast fulfilled us with these paschal sacraments, so of thy goodness thou wouldest make us to be of one heart and mind. Through ... in the unity of the same.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}).
\end{rubric}

\subby{Easter Tuesday}
\feastday{Easter Tuesday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\introit
\lett{T}{he} water of wisdom hath he given them to drink, alleluia; he shall be stayed upon them, and shall not be moved, alleluia: and he shall exalt them for ever, alleluia, alleluia. \textit{Ps.} O give thanks unto the Lord, and call upon his name: tell the people what things he hath done.

\collect
%\lett{G}{rant,} we beseech thee, Almighty God, that we who celebrate with reverence the Paschal feast, may be found worthy to attain to everlasting joys. Through.
\lett{O}{God,} who dost ever multiply thy Church with new offspring; grant unto thy servants; that they may hold fast in their lives the sacrament which by faith they have received. Through.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}).
\end{rubric}

\readingcitation{Epistle}{Acts 13:26}
\lett{I}{n those days:} Paul stood up, and beckoning with his hand said, Men of Israel, and ye that fear God, give audience. Men and brethren, children of the stock of Abraham, and whosoever among you feareth God, to you is the word of this salvation sent. For they that dwell at Jerusalem, and their rulers, because they knew him not, nor yet the voices of the prophets which are read every sabbath day, they have fulfilled them in condemning him. And though they found no cause of death in him, yet desired they Pilate that he should be slain. And when they had fulfilled all that was written of him, they took him down from the tree, and laid him in a sepulchre. But God raised him from the dead: And he was seen many days of them which came up with him from Galilee to Jerusalem, who are his witnesses unto the people. And we declare unto you glad tidings, how that the promise which was made unto the fathers, God hath fulfilled the same unto us their children, in that he hath raised up Jesus again; as it is also written in the second psalm, Thou art my Son, this day have I begotten thee. And as concerning that he raised him up from the dead, now no more to return to corruption, he said on this wise, I will give you the sure mercies of David. Wherefore he saith also in another psalm, Thou shalt not suffer thine Holy One to see corruption. For David, after he had served his own generation by the will of God, fell on sleep, and was laid unto his fathers, and saw corruption: But he, whom God raised again, saw no corruption. Be it known unto you therefore, men and brethren, that through this man is preached unto you the forgiveness of sins: And by him all that believe are justified from all things, from which ye could not be justified by the law of Moses. Beware therefore, lest that come upon you, which is spoken of in the prophets; Behold, ye despisers, and wonder, and perish: for I work a work in your days, a work which ye shall in no wise believe, though a man declare it unto you.

\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. Let them now give thanks whom the Lord hath redeemed: and delivered from the hand of the enemy, and gathered them out of the lands.}{Alleluia, alleluia. ℣. The Lord is risen from the tomb, who for us hung upon the tree.}
\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{Luke 24:36}
\lett{A}{t that time:} Jesus stood in the midst of his disciples, and saith unto them, Peace be unto you. But they were terrified and affrighted, and supposed that they had seen a spirit. And he said unto them, Why are ye troubled? and why do thoughts arise in your hearts? Behold my hands and my feet, that it is I myself: handle me, and see; for a spirit hath not flesh and bones, as ye see me have. And when he had thus spoken, he shewed them his hands and his feet. And while they yet believed not for joy, and wondered, he said unto them, Have ye here any meat? And they gave him a piece of a broiled fish, and of an honeycomb. And he took it, and did eat before them. And he said unto them, These are the words which I spake unto you, while I was yet with you, that all things must be fulfilled, which were written in the law of Moses, and in the prophets, and in the psalms, concerning me. Then opened he their understanding, that they might understand the scriptures, And said unto them, Thus it is written, and thus it behoved Christ to suffer, and to rise from the dead the third day: And that repentance and remission of sins should be preached in his name among all nations, beginning at Jerusalem. And ye are witnesses of these things.

\offertory{The Lord thundered out of heaven: and the Highest gave his thunder: and the springs of waters were seen, alleluia.}

\secret
\lett{R}{eceive,} O Lord, the prayers of thy faithful people, together with the offering of these sacrifices: that through these observances of our bounden devotion we may attain unto heavenly glory. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}).
\end{rubric}

\communion{If ye be risen with Christ, seek those things which are above, where Christ sitteth on the right hand of God, alleluia: set your affection on things above, alleluia.}

\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that the paschal sacrament, which we have received, may continually abide in our hearts. Through.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}).
\end{rubric}


\subby{Easter Wednesday}
\feastday{Easter Wednesday}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{C}{ome,} ye blessed of my Father, inherit the kingdom, alleluia: which hath been prepared for you from the foundation of the world, alleluia, alleluia, alleluia. \textit{Ps.} O sing unto the Lord a new song: sing unto the Lord, all the whole earth.

\collect
\lett{O}{God,} who dost gladden us with the yearly solemnity of the Resurrection of the Lord: mercifully grant; that through this temporal feast which we observe, we may be found worthy to attain unto everlasting joys. Through the same.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{Acts 3:12}
\lett{I}{n those days:} Peter opened his mouth and said: Ye men of Israel, and ye that fear God, hearken. The God of Abraham, and of Isaac, and of Jacob, the God of our fathers, hath glorified his Son Jesus; whom ye delivered up, and denied him in the presence of Pilate, when he was determined to let him go. But ye denied the Holy One and the Just, and desired a murderer to be granted unto you; And killed the Prince of life, whom God hath raised from the dead; whereof we are witnesses. And his name through faith in his name hath made this man strong, whom ye see and know: yea, the faith which is by him hath given him this perfect soundness in the presence of you all. And now, brethren, I wot that through ignorance ye did it, as did also your rulers. But those things, which God before had shewed by the mouth of all his prophets, that Christ should suffer, he hath so fulfilled. Repent ye therefore, and be converted, that your sins may be blotted out.

\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me.}{Alleluia, alleluia. ℣. The Lord is risen indeed: and and hath appeared to Peter.}

\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{John 21:1}
\lett{A}{t that time:} Jesus shewed himself again to the disciples at the sea of Tiberius. And on this wise shewed he himself. There were together Simon Peter, and Thomas called Didymus, and Nathanael of Cana in Galilee, and the sons of Zebedee, and two other of his disciples. Simon Peter saith unto them, I go a fishing. They say unto him, We also go with thee. They went forth, and entered into a ship immediately; and that night they caught nothing. But when the morning was now come, Jesus stood on the shore: but the disciples knew not that it was Jesus. Then Jesus saith unto them, Children, have ye any meat? They answered him, No. And he said unto them, Cast the net on the right side of the ship, and ye shall find. They cast therefore, and now they were not able to draw it for the multitude of fishes. Therefore that disciple whom Jesus loved saith unto Peter, It is the Lord. Now when Simon Peter heard that it was the Lord, he girt his fisher's coat unto him, (for he was naked,) and did cast himself into the sea. And the other disciples came in a little ship; (for they were not far from land, but as it were two hundred cubits,) dragging the net with fishes. As soon then as they were come to land, they saw a fire of coals there, and fish laid thereon, and bread. Jesus saith unto them, Bring of the fish which ye have now caught. Simon Peter went up, and drew the net to land full of great fishes, an hundred and fifty and three: and for all there were so many, yet was not the net broken. Jesus saith unto them, Come and dine. And none of the disciples durst ask him, Who art thou? knowing that it was the Lord. Jesus then cometh, and taketh bread, and giveth them, and fish likewise. This is now the third time that Jesus shewed himself to his disciples, after that he was risen from the dead.

\offertory{The Lord opened the doors of heaven: and rained down manna upon them for to eat: he gave them food from heaven: so man did eat Angels' food, alleluia.}

\secret
\lett{W}{e} offer unto thee, O Lord, the sacrifices of our paschal gladness: whereby thou dost wondrously both feed and sustain thy Church. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{Christ being raised from the dead, dieth no more, alleluia: death hath no more dominion over him, alleluia, alleluia.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we, being cleansed from all our former nature: may by the devout receiving of thy sacrament be transformed into a new creature. Who livest.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Easter Thursday}
\feastday{Easter Thursday}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{hy} victorious hand, O Lord, have they magnified, with one accord, alleluia: for wisdom hath opened the mouth of the dumb, and made eloquent the tongues of them that cannot speak, alleluia, alleluia. \textit{Ps.} O sing unto the Lord a new song: for he hath done marvellous things.

\collect
\lett{O}{God,} who hast united the diversity of nations in the confession of thy name: grant that they who are born again in the font of Baptism may agree in unity of faith and in godliness of conversation. Through.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{Acts 8:26}
\lett{I}{n those days:} The Angel of the Lord spake unto Philip, saying: Arise, and go toward the south unto the way that goeth down from Jerusalem unto Gaza, which is desert. And he arose and went: and, behold, a man of Ethiopia, an eunuch of great authority under Candace queen of the Ethiopians, who had the charge of all her treasure, and had come to Jerusalem for to worship, Was returning, and sitting in his chariot read Esaias the prophet. Then the Spirit said unto Philip, Go near, and join thyself to this chariot. And Philip ran thither to him, and heard him read the prophet Esaias, and said, Understandest thou what thou readest? And he said, How can I, except some man should guide me? And he desired Philip that he would come up and sit with him. The place of the scripture which he read was this, He was led as a sheep to the slaughter; and like a lamb dumb before his shearer, so opened he not his mouth: In his humiliation his judgment was taken away: and who shall declare his generation? for his life is taken from the earth. And the eunuch answered Philip, and said, I pray thee, of whom speaketh the prophet this? of himself, or of some other man? Then Philip opened his mouth, and began at the same scripture, and preached unto him Jesus. And as they went on their way, they came unto a certain water: and the eunuch said, See, here is water; what doth hinder me to be baptized? And Philip said, If thou believest with all thine heart, thou mayest. And he answered and said, I believe that Jesus Christ is the Son of God. And he commanded the chariot to stand still: and they went down both into the water, both Philip and the eunuch; and he baptized him. And when they were come up out of the water, the Spirit of the Lord caught away Philip, that the eunuch saw him no more: and he went on his way rejoicing. But Philip was found at Azotus: and passing through he preached in all the cities, till he came to Caesarea.

\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. The same stone, which the builders refused, is become the head-stone in the corner: this is the Lord's doing, and it is marvellous in our eyes.}{Alleluia, alleluia. ℣. Christ, who created all things, is risen: and he hath had compassion on mankind.}

\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{John 20:11}
\lett{A}{t that time:} Mary stood without at the sepulchre weeping. And as she wept, she stooped down, and looked into the sepulchre, And seeth two angels in white sitting, the one at the head, and the other at the feet, where the body of Jesus had lain. And they say unto her, Woman, why weepest thou? She saith unto them, Because they have taken away my Lord, and I know not where they have laid him. And when she had thus said, she turned herself back, and saw Jesus standing, and knew not that it was Jesus. Jesus saith unto her, Woman, why weepest thou? whom seekest thou? She, supposing him to be the gardener, saith unto him, Sir, if thou have borne him hence, tell me where thou hast laid him, and I will take him away. Jesus saith unto her, Mary. She turned herself, and saith unto him, Rabboni; which is to say, Master. Jesus saith unto her, Touch me not; for I am not yet ascended to my Father: but go to my brethren, and say unto them, I ascend unto my Father, and your Father; and to my God, and your God. Mary Magdalene came and told the disciples that she had seen the Lord, and that he had spoken these things unto her.

\offertory{In the day of your solemnity, saith the Lord, I will bring you into a land flowing with milk and honey, alleluia.}

\secret
\lett{W}{e} beseech thee, O Lord, mercifully to accept the oblations of thy people: that being made regenerate by the confession of thy name, and by baptism, they may attain to everlasting blessedness. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{Ye are a peculiar people, shew ye forth the praises of him, alleluia: who hath called you out of darkness into his marvellous light, alleluia.}

\postcommunion
\lett{G}{raciously} hear our prayers, O Lord: that this holy communion of our redemption may both bestow on us thy succour in this present life, and obtain for us everlasting joys. Through.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Easter Friday}
\feastday{Easter Friday}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord hath brought them out safely, alleluia: and overwhelmed their enemies with the sea, alleluia, alleluia, alleluia. \textit{Ps.} Hear my law, O my people: incline your ears unto the words of my mouth.

\collect
\lett{A}{lmighty} and everlasting God, who hast bestowed the paschal sacrament for a pledge of man's reconciliation: grant unto our hearts; that what we celebrate in outward profession we may effectually fulfil. Through.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{1 Peter 3:18}
\lett{D}{early beloved:} Christ hath once suffered for sins, the just for the unjust, that he might bring us to God, being put to death in the flesh, but quickened by the Spirit: By which also he went and preached unto the spirits in prison; Which sometime were disobedient, when once the longsuffering of God waited in the days of Noah, while the ark was a preparing, wherein few, that is, eight souls were saved by water. The like figure whereunto even baptism doth also now save us (not the putting away of the filth of the flesh, but the answer of a good conscience toward God,) by the resurrection of Jesus Christ: Who is gone into heaven, and is on the right hand of God.

\gradall{This is the day which the Lord hath made: we will rejoice and be glad in it. ℣. Blessed is he that cometh in the name of the Lord: God is the Lord who hath shewed us light.}{Alleluia, alleluia. ℣. Tell it out among the heathen: that the Lord hath reigned from the tree.}

\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{Matthew 28:16}
\lett{A}{t that time:} The eleven disciples went away into Galilee, into a mountain where Jesus had appointed them. And when they saw him, they worshipped him: but some doubted. And Jesus came and spake unto them, saying, All power is given unto me in heaven and in earth. Go ye therefore, and teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost: Teaching them to observe all things whatsoever I have commanded you: and, lo, I am with you alway, even unto the end of the world.

\offertory{This day shall be unto you for a memorial, alleluia: and ye shall keep it a feast to the Lord throughout your generations: ye shall keep it a feast by an ordinance for ever, alleluia, alleluia, alleluia.}

\secret
\lett{W}{e} beseech thee, O Lord, mercifully to accept our oblations: which we offer unto thee for the expiation of the sins of them that are born again, and for the speedy attaining of thy heavenly succour. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\communion{All power is given to me in heaven and in earth, alleluia: go ye, and teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost, alleluia, alleluia.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, look upon thy people: and as thou hast vouchsafed to renew them with thine eternal mysteries, so of thy goodness deliver them from temporal guilt. Through.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Easter Saturday}
\feastday{Easter Saturday}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord hath brought forth his people with joy, alleluia: and his chosen with gladness, alleluia, alleluia. \textit{Ps.} O give thanks unto the Lord, and call upon his name: tell the people what things he hath done.

\collect
\lett{G}{rant,} we beseech thee, almighty God: that we who have devoutly kept this paschal festival may thereby be found worthy to attain to everlasting joys. Through.
\begin{rubric}
    \nth{2} Collect is of Easter Day (p. \pageref{EasterCollect}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{1 Peter 2:1}
\lett{D}{early beloved:} Laying aside therefore all malice, and all guile, and hypocrisies, and envies, and all evil speakings, As newborn babes, desire the sincere milk of the word, that ye may grow thereby: If so be ye have tasted that the Lord is gracious. To whom coming, as unto a living stone, disallowed indeed of men, but chosen of God, and precious, Ye also, as lively stones, are built up a spiritual house, an holy priesthood, to offer up spiritual sacrifices, acceptable to God by Jesus Christ. Wherefore also it is contained in the scripture, Behold, I lay in Sion a chief corner stone, elect, precious: and he that believeth on him shall not be confounded. Unto you therefore which believe he is precious: but unto them which be disobedient, the stone which the builders disallowed, the same is made the head of the corner, And a stone of stumbling, and a rock of offence, even to them which stumble at the word, being disobedient: whereunto also they were appointed. But ye are a chosen generation, a royal priesthood, an holy nation, a peculiar people; that ye should shew forth the praises of him who hath called you out of darkness into his marvellous light: Which in time past were not a people, but are now the people of God: which had not obtained mercy, but now have obtained mercy.

\begin{rubric}
	From this day until Ember Saturday in Whitsun Week, inclusive, in all Masses, the Gradual is not said, but the Greater Alleluia. 
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. This is the day which the Lord hath made: we will rejoice and be glad in it. Alleluia. ℣. Praise the Lord, ye servants, O praise the name of the Lord.}

\begin{rubric}
    Sequence as on Easter Day (p. \pageref{PaschalSequence}).
\end{rubric}

\readingcitation{Gospel}{John 20:1}
\lett{A}{t that time:} On the first day of the week cometh Mary Magdalene early, when it was yet dark, unto the sepulchre, and seeth the stone taken away from the sepulchre. Then she runneth, and cometh to Simon Peter, and to the other disciple, whom Jesus loved, and saith unto them, They have taken away the Lord out of the sepulchre, and we know not where they have laid him. Peter therefore went forth, and that other disciple, and came to the sepulchre. So they ran both together: and the other disciple did outrun Peter, and came first to the sepulchre. And he stooping down, and looking in, saw the linen clothes lying; yet went he not in. Then cometh Simon Peter following him, and went into the sepulchre, and seeth the linen clothes lie, And the napkin, that was about his head, not lying with the linen clothes, but wrapped together in a place by itself. Then went in also that other disciple, which came first to the sepulchre, and he saw, and believed. For as yet they knew not the scripture, that he must rise again from the dead.

\offertory{Blessed is he that cometh in the name of the Lord: we have wished you good luck, ye that are of the house of the Lord: God is the Lord who hath shewed us light, alleluia, alleluia.}

\secret
\lett{G}{rant} us, we beseech thee, O Lord, alway to rejoice in these paschal mysteries: that the continual working of our redemption may be to us the cause of everlasting gladness. Through.
\begin{rubric}
    \nth{2} Secret is of Easter Day (p. \pageref{EasterSecret}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{As many of you as have been baptized into Christ have put on Christ, alleluia.}

\postcommunion
\lett{O}{Lord,} who hast quickened us with the gift of our redemption, we beseech thee: that this aid to everlasting salvation may avail for our advancement in the true faith. Through.
\begin{rubric}
    \nth{2} Postcommunion is of Easter Day (p. \pageref{EasterPostcommunion}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}



\subby{Low Sunday}
\feastday{Easter I}
\begin{inhead}
    {First Class Double, Octave Day of Easter}
\end{inhead}

\properantiphonfix

\introit
\lett{A}{s} new-born babes, alleluia: desire ye the guileless milk of the spirit, alleluia, alleluia, alleluia. \textit{Ps.} Sing we merrily unto God our helper: make a cheerful noise unto the God of Jacob.
\begin{rubric}
    \emph{Gloria in excelsis} is said on this and the following Sundays after Easter, even when on week-days the Mass of the preceding Sunday is resumed.
\end{rubric}
\collect
\lett{A}{lmighty} Father, who hast given thine only Son to die for our sins, and to rise again for our justification; Grant us so to put away the leaven of malice and wickedness, that we may always serve thee in pureness of living and truth; through the merits of the same thy Son Jesus Christ our Lord. Who with.

%1928 edition followed:
\readingcitation{Epistle}{1 John 5:4}
\lett{D}{early beloved:} Whatsoever is born of God overcometh the world: and this is the victory that overcometh the world, even our faith. Who is he that overcometh the world, but he that believeth that Jesus is the Son of God? This is he that came by water and blood, even Jesus Christ; not by water only, but by water and blood. And it is the Spirit that beareth witness, because the Spirit is truth. For there are three that bear witness, the Spirit, and the water, and the blood: and these three agree in one. If we receive the witness of men, the witness of God is greater: for this is the witness of God which he hath testified of his Son. He that believeth on the Son of God hath the witness in himself: he that believeth not God hath made him a liar; because he believeth not the record that God gave of his Son. And this is the record, that God hath given to us eternal life, and this life is in his Son. He that hath the Son hath life; and he that hath not the Son hath not life.

\alleluia{Alleluia, alleluia. ℣. In the day of my resurrection, saith the Lord, I will go before you into Galilee. Alleluia. ℣. After eight days, when the doors were shut, Jesus stood in the midst of his disciples, and said: Peace be unto you. Alleluia.}

\readingcitation{Gospel}{John 20:19}
\lett{A}{t that time:} The same day at evening, being the first day of the week, when the doors were shut where the disciples were assembled for fear of the Jews, came Jesus and stood in the midst, and saith unto them, Peace be unto you. And when he had so said, he shewed unto them his hands and his side. Then were the disciples glad, when they saw the Lord. Then said Jesus to them again, Peace be unto you: as my Father hath sent me, even so send I you. And when he had said this, he breathed on them, and saith unto them, Receive ye the Holy Ghost: Whose soever sins ye remit, they are remitted unto them; and whose soever sins ye retain, they are retained.

\offertory{The Angel of the Lord descended from heaven, and saith unto the women: He whom ye seek is risen, as he said, alleluia.}
\secret
\lett{R}{eceive,} we beseech thee, O Lord, the offerings of thy Church exultant: and grant unto her, on whom thou hast bestowed cause for so great joy, the fruit of perpetual gladness. Through.
\communion{Reach hither thy hand, and behold the print of the nails, alleluia: and be not faithless, but believing, alleluia, alleluia.}
\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that these holy mysteries which thou hast given us for the assurance of our redemption; may avail for our healing both in this life and in that which is to come. Through.

\subby{Second Sunday after Easter}\label{EasterII}
\feastday{Easter II}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} loving-kindness of the Lord filleth the whole world, alleluia: by the word of the Lord the heavens were stablished, alleluia, alleluia. \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.

\collect\label{EasterIICollect}
\lett{A}{lmighty God,} who hast given thine only Son to be unto us both a sacrifice for sin, and also an ensample of godly life; Give us grace that we may always most thankfully receive that his inestimable benefit, and also daily endeavour ourselves to follow the blessed steps of his most holy life. Through the same.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{1 Peter 2:19}
\lett{D}{early beloved:} This is thankworthy, if a man for conscience toward God endure grief, suffering wrongfully. For what glory is it, if, when ye be buffeted for your faults, ye shall take it patiently? but if, when ye do well, and suffer for it, ye take it patiently, this is acceptable with God. For even hereunto were ye called: because Christ also suffered for us, leaving us an example, that ye should follow his steps: Who did no sin, neither was guile found in his mouth: Who, when he was reviled, reviled not again; when he suffered, he threatened not; but committed himself to him that judgeth righteously: Who his own self bare our sins in his own body on the tree, that we, being dead to sins, should live unto righteousness: by whose stripes ye were healed. For ye were as sheep going astray; but are now returned unto the Shepherd and Bishop of your souls.
\alleluia{Alleluia, alleluia. ℣. The disciples knew the Lord Jesus in the breaking of bread. Alleluia. ℣. I am the good shepherd: and know my sheep, and am known of mine. Alleluia.}
\readingcitation{Gospel}{John 10:11}
\lett{A}{t that time:} Jesus said unto the Pharisees: I am the good shepherd: the good shepherd giveth his life for the sheep. But he that is an hireling, and not the shepherd, whose own the sheep are not, seeth the wolf coming, and leaveth the sheep, and fleeth: and the wolf catcheth them, and scattereth the sheep. The hireling fleeth, because he is an hireling, and careth not for the sheep. I am the good shepherd, and know my sheep, and am known of mine. As the Father knoweth me, even so know I the Father: and I lay down my life for the sheep. And other sheep I have, which are not of this fold: them also I must bring, and they shall hear my voice; and there shall be one fold, and one shepherd.
\offertory{O God, thou art my God, early will I seek thee: and lift up my hands in thy name, alleluia.}
\secret\label{EasterIISecret}
\lett{M}{ay} this sacred oblation, O Lord, ever bestow upon us thy saving benediction: that what is performed in a mystery may in power be fulfilled. Through.
\begin{rubric}
    \nth{2} Secret is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{I am the good shepherd, alleluia: and know my sheep, and am known of mine, alleluia, alleluia.}
\postcommunion\label{EasterIIPostcommunion}
\lett{G}{rant} to us, we beseech thee, almighty God: that we, receiving the quickening of thy grace, may ever glory in thy gift. Through.
\begin{rubric}
    \nth{2} Postcommunion is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Patronage of St. Joseph}
\feastday{Patronage of St. Joseph}
\begin{inhead}
    {Wednesday after the Second Sunday after Easter}\par
    {First Class Double, Simple Octave}
\end{inhead}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
%From Dominican Breviary (See Divine Compassion).
\textit{Opening Sentence.} He made him lord also of his house : and ruler of all his substance.

\introit
\lett{T}{he} Lord is our help, and our shield: our heart shall rejoice in him, because we have hoped in his holy name, alleluia, alleluia. \textit{Ps.} Hear, O thou Shepherd of Israel: thou that leadest Joseph like a sheep,
\collect
\lett{O}{God,} who by thy ineffable providence didst vouchsafe to choose blessed Joseph to be the spouse of thy most holy Mother: grant, we beseech thee; that we, who venerate him as a protector on earth, may be found worthy to have him as an intercessor in heaven. Who livest.
\readingcitation{Epistle}{Genesis 49:22}
\lett{J}{oseph} is a fruitful bough, even a fruitful bough by a well; whose branches run over the wall: The archers have sorely grieved him, and shot at him, and hated him: But his bow abode in strength, and the arms of his hands were made strong by the hands of the mighty God of Jacob; (from thence is the shepherd, the stone of Israel:) Even by the God of thy father, who shall help thee; and by the Almighty, who shall bless thee with blessings of heaven above, blessings of the deep that lieth under, blessings of the breasts, and of the womb: The blessings of thy father have prevailed above the blessings of my progenitors unto the utmost bound of the everlasting hills: they shall be on the head of Joseph, and on the crown of the head of him that was separate from his brethren.
\alleluia{Alleluia, alleluia, ℣. From whatsoever tribulation they shall cry unto me, I will hear them, and I will be their defender for ever. Alleluia. ℣. Grant us, O Joseph, to lead an innocent life: and may it ever be safe beneath thy care. Alleluia.}
\readingcitation{Gospel}{Luke 3:21}
\lett{A}{t that time:} When all the people were baptized, it came to pass, that Jesus also being baptized, and praying, the heaven was opened, And the Holy Ghost descended in a bodily shape like a dove upon him, and a voice came from heaven, which said, Thou art my beloved Son; in thee I am well pleased. And Jesus himself began to be about thirty years of age, being (as was supposed) the son of Joseph.
\offertory{Praise the Lord, O Jerusalem: for he hath made fast the bars of thy gates, and hath blessed thy children within thee, alleluia, alleluia.}
\secret
\lett{O}{Lord,} forasmuch as we put our trust in the advocacy of the spouse of thy most holy Mother, we entreat thy mercy: that thou wouldest make our hearts to despise all earthly things, and to love thee the true God with perfect charity. Who livest.
\communion{And Jacob begat Joseph the husband of Mary, of whom was born Jesus, who is called Christ, alleluia, alleluia.}
\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that we whom thou hast refreshed with the fountain of the divine gift, may by thee be enabled so to rejoice in the protection of blessed Joseph: that, by his merits and intercession, we may be made partakers of thy heavenly glory. Through.


\subby{Third Sunday after Easter}
\feastday{Easter III}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{be} joyful in God, all ye lands, alleluia: sing ye praises unto the honour of his name, alleluia: make his praise to be exceeding glorious, alleluia, alleluia, alleluia. \textit{Ps.} Say unto God, O how wonderful art thou in thy works, O Lord! through the greatness of thy power shall thine enemies be found liars unto thee.
\collect
\lett{A}{lmighty} God, who showest to them that are in error the light of thy truth, to the intent that they may return into the way of righteousness; Grant unto all those who are admitted into the fellowship of Christ's Religion, that they may avoid those things that are contrary to their profession, and follow all such things as are agreeable to the same. Through.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{1 Peter 2:11}
\lett{D}{early beloved:} I beseech you as strangers and pilgrims, abstain from fleshly lusts, which war against the soul; Having your conversation honest among the Gentiles: that, whereas they speak against you as evildoers, they may by your good works, which they shall behold, glorify God in the day of visitation. Submit yourselves to every ordinance of man for the Lord's sake: whether it be to the king, as supreme; Or unto governors, as unto them that are sent by him for the punishment of evildoers, and for the praise of them that do well. For so is the will of God, that with well doing ye may put to silence the ignorance of foolish men: As free, and not using your liberty for a cloke of maliciousness, but as the servants of God. Honour all men. Love the brotherhood. Fear God. Honour the king.
\alleluia{Alleluia, alleluia. ℣. The Lord hath sent redemption unto his people. Alleluia. ℣. It behoved Christ to suffer, and to rise from the dead: and so to enter into his glory. Alleluia.}
\readingcitation{Gospel}{John 16:16}
\lett{A}{t that time:} Jesus said to his disciples: A little while, and ye shall not see me: and again, a little while, and ye shall see me, because I go to the Father. Then said some of his disciples among themselves, What is this that he saith unto us, A little while, and ye shall not see me: and again, a little while, and ye shall see me: and, Because I go to the Father? They said therefore, What is this that he saith, A little while? we cannot tell what he saith. Now Jesus knew that they were desirous to ask him, and said unto them, Do ye enquire among yourselves of that I said, A little while, and ye shall not see me: and again, a little while, and ye shall see me? Verily, verily, I say unto you, That ye shall weep and lament, but the world shall rejoice: and ye shall be sorrowful, but your sorrow shall be turned into joy. A woman when she is in travail hath sorrow, because her hour is come: but as soon as she is delivered of the child, she remembereth no more the anguish, for joy that a man is born into the world. And ye now therefore have sorrow: but I will see you again, and your heart shall rejoice, and your joy no man taketh from you.
\offertory{Praise the Lord, O my soul: while I live will I praise the Lord: yea, as long as I have any being, I will sing praises unto my God, alleluia.}
\secret
\lett{G}{rant} to us, O Lord, that through these mysteries we may so assuage our earthly desires, that we may learn to love things heavenly. Through.

\begin{rubric}
    \nth{2} Secret is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{A little while, and ye shall not see me, alleluia: and again, a little while and ye shall see me, because I go to the Father, alleluia, alleluia.}
\postcommunion
\lett{M}{ay} the sacraments which we have received, we beseech thee, O Lord: both renew us with spiritual sustenance, and defend us with bodily succour. Through.
\begin{rubric}
    \nth{2} Postcommunion is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Fourth Sunday after Easter}
\feastday{Easter IV}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{sing} unto the Lord a new song, alleluia: for the Lord hath done marvellous things, alleluia: in the sight of the nations hath he shewed his righteousness, alleluia, alleluia, alleluia. \textit{Ps.} With his own right hand, and with his holy arm: hath he gotten himself the victory.
\collect
\lett{O}{almighty} God, who alone canst order the unruly wills and affections of sinful men; Grant unto thy people, that they may love the thing which thou commandest, and desire that which thou dost promise; that so, among the sundry and manifold changes of the world, our hearts may surely there be fixed, where true joys are to be found. Through.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{James 1:17}
\lett{D}{early beloved:} Every good gift and every perfect gift is from above, and cometh down from the Father of lights, with whom is no variableness, neither shadow of turning. Of his own will begat he us with the word of truth, that we should be a kind of firstfruits of his creatures. Wherefore, my beloved brethren, let every man be swift to hear, slow to speak, slow to wrath: For the wrath of man worketh not the righteousness of God. Wherefore lay apart all filthiness and superfluity of naughtiness, and receive with meekness the engrafted word, which is able to save your souls.
\alleluia{Alleluia, alleluia, ℣. The right hand of the Lord bringeth mighty things to pass: the right hand of the Lord hath exalted me. Alleluia. ℣. Christ being raised from the dead dieth no more: death hath no more dominion over him. Alleluia.}
\readingcitation{Gospel}{John 16:5}
\lett{A}{t that time:} Jesus said unto his disciples: I go my way to him that sent me; and none of you asketh me, Whither goest thou? But because I have said these things unto you, sorrow hath filled your heart. Nevertheless I tell you the truth; It is expedient for you that I go away: for if I go not away, the Comforter will not come unto you; but if I depart, I will send him unto you. And when he is come, he will reprove the world of sin, and of righteousness, and of judgment: Of sin, because they believe not on me; Of righteousness, because I go to my Father, and ye see me no more; Of judgment, because the prince of this world is judged. I have yet many things to say unto you, but ye cannot bear them now. Howbeit when he, the Spirit of truth, is come, he will guide you into all truth: for he shall not speak of himself; but whatsoever he shall hear, that shall he speak: and he will shew you things to come. He shall glorify me: for he shall receive of mine, and shall shew it unto you. All things that the Father hath are mine: therefore said I, that he shall take of mine, and shall shew it unto you.
\offertory{O be joyful in God, all ye lands, sing praises unto the honour of his name: O come hither and hearken, all ye that fear God, and I will tell you what the Lord hath done for my soul, alleluia.}
\secret
\lett{O}{God,} who by communion in this venerable sacrifice hast made us partakers of the one supreme Godhead: grant, we beseech thee; that as we have the knowledge of thy truth, so by worthy conversation we may attain unto the same. Through.
\begin{rubric}
    \nth{2} Secret is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{When the Comforter, the Spirit of Truth, is come, he will reprove the world of sin, and of righteousness, and of judgment, alleluia, alleluia.}
\postcommunion
\lett{A}{ssist} us, O Lord our God: that by these things, which we have faithfully received, we may both be purged from sin, and delivered from all dangers. Through.
\begin{rubric}
    \nth{2} Postcommunion is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Rogation Sunday}
\feastday{Easter V}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{ith} a voice of singing declare ye this, and let it be heard, alleluia: utter it even unto the end of the earth: the Lord hath delivered his people, alleluia, alleluia. \textit{Ps.} O be joyful in God, all ye lands, sing praises unto the honour of his name: make his praise to be glorious.
\collect
\lett{O}{Lord,} from whom all good things do come; Grant to us thy humble servants, that by thy holy inspiration we may think those things that are good, and by thy merciful guiding may perform the same. Through.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{James 1:22}
\lett{D}{early beloved:} Be ye doers of the word, and not hearers only, deceiving your own selves. For if any be a hearer of the word, and not a doer, he is like unto a man beholding his natural face in a glass: For he beholdeth himself, and goeth his way, and straightway forgetteth what manner of man he was. But whoso looketh into the perfect law of liberty, and continueth therein, he being not a forgetful hearer, but a doer of the work, this man shall be blessed in his deed. If any man among you seem to be religious, and bridleth not his tongue, but deceiveth his own heart, this man's religion is vain. Pure religion and undefiled before God and the Father is this, To visit the fatherless and widows in their affliction, and to keep himself unspotted from the world.
\alleluia{Alleluia, alleluia, ℣. Christ is risen, and hath shewed light unto us, whom he hath redeemed with his blood. Alleluia. ℣. I came forth from the Father, and am come into the world: again, I leave the world, and go to the Father. Alleluia.}
\readingcitation{Gospel}{John 16:23}
\lett{A}{t that time:} Jesus said unto his disciples: Verily, verily, I say unto you, Whatsoever ye shall ask the Father in my name, he will give it you. Hitherto have ye asked nothing in my name: ask, and ye shall receive, that your joy may be full. These things have I spoken unto you in proverbs: but the time cometh, when I shall no more speak unto you in proverbs, but I shall shew you plainly of the Father. At that day ye shall ask in my name: and I say not unto you, that I will pray the Father for you: For the Father himself loveth you, because ye have loved me, and have believed that I came out from God. I came forth from the Father, and am come into the world: again, I leave the world, and go to the Father. His disciples said unto him, Lo, now speakest thou plainly, and speakest no proverb. Now are we sure that thou knowest all things, and needest not that any man should ask thee: by this we believe that thou camest forth from God. Jesus answered them, Do ye now believe? Behold, the hour cometh, yea, is now come, that ye shall be scattered, every man to his own, and shall leave me alone: and yet I am not alone, because the Father is with me. These things I have spoken unto you, that in me ye might have peace. In the world ye shall have tribulation: but be of good cheer; I have overcome the world.

\offertory{O praise the Lord our God, ye people, and make the voice of his praise to be heard: who holdeth our soul in life, and suffereth not our feet to slip: praised be the Lord who hath not cast out my prayer, nor turned his mercy from me, alleluia.}
\secret
\lett{R}{eceive,} O Lord, the prayers of thy faithful people, together with the offering of these sacrifices: that through these observances of our bounden devotion we may attain unto heavenly glory. Through.
\begin{rubric}
    \nth{2} Secret is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{O sing unto the Lord, alleluia: sing unto the Lord, and praise his name, be telling of his salvation from day to day, alleluia, alleluia.}
\postcommunion
\lett{G}{rant} unto us, O Lord: that we who have been fulfilled with the strength of thy heavenly table: may both desire those things which be right, and obtain those things which we desire. Through.
\begin{rubric}
    \nth{2} Postcommunion is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Rogation Days}
\feastday{Rogation Days}
\begin{rubric}
	Rogation Monday is a \nth{2} Feria. Rogation Tuesday is a Feria.
\end{rubric}

\introit
\lett{H}{e} hath heard my voice out of his holy temple, alleluia: and my complaint hath come before him, it hath entered even into his ears, alleluia, alleluia. \textit{Ps.} I will love thee, O Lord, my strength: the Lord is my stony rock, my fortress, and my saviour.
\begin{rubric}
    In this Mass, \emph{Gloria in excelsis} is not said, nor the Creed.
\end{rubric}

\collect
\lett{A}{lmighty} God, Lord of heaven and earth; We beseech thee to pour forth thy blessing upon this land, and to give us a fruitful season; that we, constantly receiving thy bounty, may evermore give thanks unto thee in thy holy Church. Through.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Ezekiel 34:25}
\lett{I}{will} make with them a covenant of peace, and will cause the evil beasts to cease out of the land: and they shall dwell safely in the wilderness, and sleep in the woods. And I will make them and the places round about my hill a blessing; and I will cause the shower to come down in his season; there shall be showers of blessing. And the tree of the field shall yield her fruit, and the earth shall yield her increase, and they shall be safe in their land, and shall know that I am the \divineName{Lord}, when I have broken the bands of their yoke, and delivered them out of the hand of those that served themselves of them. And they shall no more be a prey to the heathen, neither shall the beast of the land devour them; but they shall dwell safely, and none shall make them afraid. And I will raise up for them a plant of renown, and they shall be no more consumed with hunger in the land, neither bear the shame of the heathen any more. Thus shall they know that I the \divineName{Lord} their God am with them, and that they, even the house of Israel, are my people, saith the Lord \divineName{God}. And ye my flock, the flock of my pasture, are men, and I am your God, saith the Lord \divineName{God}.

\alleluia{Alleluia, ℣. O give thanks unto the Lord, for he is gracious: and his mercy endureth for ever.}

\readingcitation{Gospel}{Luke 11:5}
\lett{A}{t that time:} Jesus said unto his disciples: And he said unto them, Which of you shall have a friend, and shall go unto him at midnight, and say unto him, Friend, lend me three loaves; For a friend of mine in his journey is come to me, and I have nothing to set before him? And he from within shall answer and say, Trouble me not: the door is now shut, and my children are with me in bed; I cannot rise and give thee. I say unto you, Though he will not rise and give him, because he is his friend, yet because of his importunity he will rise and give him as many as he needeth. And I say unto you, Ask, and it shall be given you; seek, and ye shall find; knock, and it shall be opened unto you. For every one that asketh receiveth; and he that seeketh findeth; and to him that knocketh it shall be opened. If a son shall ask bread of any of you that is a father, will he give him a stone? or if he ask a fish, will he for a fish give him a serpent? Or if he shall ask an egg, will he offer him a scorpion? If ye then, being evil, know how to give good gifts unto your children: how much more shall your heavenly Father give the Holy Spirit to them that ask him?

\offertory{I will give great thanks unto the Lord with my mouth: and praise him among the multitude, for he shall stand at the right hand of the poor: to save his soul from unrighteous judges, alleluia.}

\secret
\lett{W}{e} beseech thee, O Lord, that these our oblations may both loose the bonds of our iniquity, and obtain for us the gifts of thy loving-kindness. Through.
\begin{rubric}
    \nth{2} Secret is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{Ask, and ye shall receive : seek, and ye shall find: knock, and it shall be opened unto you: for every one that asketh receiveth; and he that seeketh findeth: and to him that knocketh it shall be opened, alleluia.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, to further with thy gracious favour these our supplications: that we, receiving thy gifts in our tribulation, may increase in thy love by the consolation of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Ascension Thursday}
\feastday{Ascension Thursday}
\begin{inhead}
    {First Class Double, Third Class Octave}
\end{inhead}

\properantiphonfix

\introit
\lett{Y}{e} men of Galilee, why marvel ye gazing up into heaven? alleluia: in like manner as ye have seen him going up into heaven, so shall he come again, alleluia, alleluia, alleluia. \textit{Ps.} O clap your hands together, all ye people: O sing unto God with the voice of melody,
\collect
\lett{G}{rant,} we beseech thee, Almighty God, that like as we do believe thy only-begotten Son our Lord Jesus Christ to have ascended into the heavens; so we may also in heart and mind thither ascend, and with him continually dwell, who liveth and reigneth with thee and the Holy Ghost, one God, world without end. \textit{Amen.}

\readingcitation{Epistle}{Acts 1:1}
\lett{T}{he} former treatise have I made, O Theophilus, of all that Jesus began both to do and teach, Until the day in which he was taken up, after that he through the Holy Ghost had given commandments unto the apostles whom he had chosen: To whom also he shewed himself alive after his passion by many infallible proofs, being seen of them forty days, and speaking of the things pertaining to the kingdom of God: And, being assembled together with them, commanded them that they should not depart from Jerusalem, but wait for the promise of the Father, which, saith he, ye have heard of me. For John truly baptized with water; but ye shall be baptized with the Holy Ghost not many days hence. When they therefore were come together, they asked of him, saying, Lord, wilt thou at this time restore again the kingdom to Israel? And he said unto them, It is not for you to know the times or the seasons, which the Father hath put in his own power. But ye shall receive power, after that the Holy Ghost is come upon you: and ye shall be witnesses unto me both in Jerusalem, and in all Judaea, and in Samaria, and unto the uttermost part of the earth. And when he had spoken these things, while they beheld, he was taken up; and a cloud received him out of their sight. And while they looked stedfastly toward heaven as he went up, behold, two men stood by them in white apparel; Which also said, Ye men of Galilee, why stand ye gazing up into heaven? this same Jesus, which is taken up from you into heaven, shall so come in like manner as ye have seen him go into heaven.
\alleluia{Alleluia, alleluia. ℣. God is gone up with a merry noise, and the Lord with the sound of the trumpet. Alleluia. ℣. The Lord in the holy place of Sinai, ascending up on high, hath led captivity captive. Alleluia.}

%1928:
%\readingcitation{Gospel}{Luke 24:49}
%\lett{A}{t that time:} Jesus said unto his disciples: Behold, I send the promise of my Father upon you: but tarry ye in the city of Jerusalem, until ye be endued with power from on high. And he led them out as far as to Bethany, and he lifted up his hands, and blessed them. And it came to pass, while he blessed them, he was parted from them, and carried up into heaven. And they worshipped him, and returned to Jerusalem with great joy: and were continually in the temple, praising and blessing God.

%Sarum:
\readingcitation{Gospel}{Mark 16:14}
\lett{A}{t that time:} Jesus appeared unto the eleven as they sat at meat, and upbraided them with their unbelief and hardness of heart, because they believed not them which had seen him after he was risen. And he said unto them, Go ye into all the world, and preach the gospel to every creature. He that believeth and is baptized shall be saved; but he that believeth not shall be damned. And these signs shall follow them that believe; In my name shall they cast out devils; they shall speak with new tongues; They shall take up serpents; and if they drink any deadly thing, it shall not hurt them; they shall lay hands on the sick, and they shall recover. So then after the Lord had spoken unto them, he was received up into heaven, and sat on the right hand of God. And they went forth, and preached every where, the Lord working with them, and confirming the word with signs following.
\offertory{God is gone up with a merry noise: and the Lord with the sound of the trumpet, alleluia.}
\secret
\lett{R}{eceive,} O Lord, the gifts which we offer for the glorious Ascension of thy Son: and mercifully grant: that we may be delivered from present dangers, and attain to everlasting life. Through the same.
\communion{Sing ye to the Lord, who ascended above the heaven of heavens, to the Sunrising, alleluia.}
\postcommunion
\lett{G}{rant} to us, we beseech thee, almighty and merciful God: that those things which we have known to be received in visible mysteries we may obtain in invisible effect. Through.
\begin{rubric}
    During the Octave, the Mass is said as on the Feast, and after the Prayers of the Day are added the Seasonal Propers for Easter II (p. \pageref{EasterII}).
\end{rubric}

\subsec{Sunday within the Ascension Octave}
\feastday{Ascension Sunday}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{earken} unto my voice, O Lord, when I cry unto thee, alleluia: unto thee my heart hath said, Thy face have I sought, thy face, Lord, will I seek: O hide not thou thy face from me, alleluia, alleluia. \textit{Ps.} The Lord is my light, and my salvation: whom then shall I fear?
\collect
\lett{O}{God,} the King of glory, who hast exalted thine only Son Jesus Christ with great triumph unto thy kingdom in heaven; We beseech thee, leave us not comfortless; but send to us thine Holy Ghost to comfort us, and exalt us unto the same place whither our Saviour Christ is gone before, who liveth and reigneth with thee and the Holy Ghost, one God, world without end. \textit{Amen.}
\Needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty God: that like as we do believe thy only-begotten Son our Lord Jesus Christ to have ascended into the heavens; so we may also in heart and mind thither ascend and with him continually dwell. Who liveth and reigneth with thee.

\readingcitation{Epistle}{1 Peter 4:7}
\lett{D}{early beloved:} The end of all things is at hand: be ye therefore sober, and watch unto prayer. And above all things have fervent charity among yourselves: for charity shall cover the multitude of sins. Use hospitality one to another without grudging. As every man hath received the gift, even so minister the same one to another, as good stewards of the manifold grace of God. If any man speak, let him speak as the oracles of God; if any man minister, let him do it as of the ability which God giveth: that God in all things may be glorified through Jesus Christ, to whom be praise and dominion for ever and ever. Amen.

\alleluia{Alleluia, alleluia, ℣. The Lord reigneth over all the heathen: God sitteth upon his holy seat. Alleluia. ℣. I will not leave you comfortless: I go away and come again unto you, and your heart shall rejoice. Alleluia.}

\readingcitation{Gospel}{John 15:26}
\lett{A}{t that time:} Jesus said unto his disciples: When the Comforter is come, whom I will send unto you from the Father, even the Spirit of truth, which proceedeth from the Father, he shall testify of me: And ye also shall bear witness, because ye have been with me from the beginning. These things have I spoken unto you, that ye should not be offended. They shall put you out of the synagogues: yea, the time cometh, that whosoever killeth you will think that he doeth God service. And these things will they do unto you, because they have not known the Father, nor me. But these things have I told you, that when the time shall come, ye may remember that I told you of them.
\offertory{God is gone up with a merry noise: and the Lord with the sound of the trump, alleluia.}

\secret
\lett{M}{ay} these spotless sacrifices purify us, O Lord: and give unto our souls the strength of thy heavenly grace. Through.\\
\par\Needspace{4\baselineskip}
\lett{R}{eceive,} O Lord, the gifts which we offer for the glorious Ascension of thy Son: and mercifully grant; that we may be delivered from present dangers, and attain to everlasting life. Through the same.

\communion{Father, while I was with them, I kept those that thou gavest me, alleluia: and now I come to thee: I pray not that thou shouldest take them out of the world; but that thou shouldest keep them from the evil, alleluia, alleluia.}
\postcommunion
\lett{G}{rant}, we beseech thee, O Lord: that we, being filled with thy sacred gifts; may ever continue in thanksgiving for the same. Through.\\
\par\Needspace{4\baselineskip}
\lett{G}{rant} to us, we beseech thee, almighty and merciful God: that those things which we have known to be received in visible mysteries we may obtain in invisible effect. Through.
\begin{rubric}
    On the Octave Day of the Ascension (Greater Double), Mass is said as on the Feast.
\end{rubric}
\begin{rubric}
    On the Friday after the Octave (Semidouble), Mass of the preceding Sunday said, with \emph{Gloria in excelsis}, Seasonal Collects as on Easter II (p. \pageref{EasterII}), and Preface of the Ascension, which is to be said even in Masses of the Saints, or Votives---which have no proper preface---according to the Rubrics: but the Creed is omitted, and \emph{Joining in communion} of the Ascension is not said.
\end{rubric}
\begin{rubric}
    If on the preceding Sunday a \nth{1} or \nth{2} Double should have been celebrated, on this Feria private votive Masses and daily read Masses of the Dead are prohibited.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage

\subby{Whitsunday}
\feastday{Whitsunday}
\begin{inhead}
    {First Class Double, First Class Octave}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Spirit of the Lord hath filled the whole world, alleluia: and that which containeth all things hath knowledge of the voice, alleluia, alleluia, alleluia. \textit{Ps.} Let God arise, and let his enemies be scattered: let them also that hate him flee before him.
\collect
\lett{O}{God,} who as at this time didst teach the hearts of thy faithful people, by sending to them the light of thy Holy Spirit; Grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort. Through.
\readingcitation{Epistle}{Acts 2:1}
\lett{W}{hen} the day of Pentecost was fully come, they were all with one accord in one place. And suddenly there came a sound from heaven as of a rushing mighty wind, and it filled all the house where they were sitting. And there appeared unto them cloven tongues like as of fire, and it sat upon each of them. And they were all filled with the Holy Ghost, and began to speak with other tongues, as the Spirit gave them utterance. And there were dwelling at Jerusalem Jews, devout men, out of every nation under heaven. Now when this was noised abroad, the multitude came together, and were confounded, because that every man heard them speak in his own language. And they were all amazed and marvelled, saying one to another, Behold, are not all these which speak Galilaeans? And how hear we every man in our own tongue, wherein we were born? Parthians, and Medes, and Elamites, and the dwellers in Mesopotamia, and in Judaea, and Cappadocia, in Pontus, and Asia, Phrygia, and Pamphylia, in Egypt, and in the parts of Libya about Cyrene, and strangers of Rome, Jews and proselytes, Cretes and Arabians, we do hear them speak in our tongues the wonderful works of God.
\alleluia{Alleluia, alleluia. ℣. O send forth thy Spirit and they shall be made, and thou shalt renew the face of the earth. Alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\subsubsec{Sequence}\label{WhitSeq}
\lett{C}{ome,} thou Holy Spirit, come, and from thy celestial home send thy light and brilliancy.

\secondline{Come, thou Father of the poor; come thou source of all our store; come, the soul’s true radiancy.}
\thirdline{Thou of comforters the best, of the soul the sweetest guest, come in toil refreshingly.}
In our labour rest most sweet, grateful shadow from the heat, comfort in adversity.

O thou light most pure and blest, shine within the inmost breast of thy faithful company.

Where thou art not, man hath nought, every holy deed and thought comes from thy divinity.

What is soiled, make thou pure, what is wounded, work its cure, water what is parched and dry.

What is rigid, gently bend, what is frozen, warmly tend, strengthen what goes erringly.

Fill thy faithful, who confide in thy power to guard and guide, with thy sevenfold mystery.

Here thy grace and virtue send, grant salvation in the end, and in heaven felicity. Amen. Alleluia.
\readingcitation{Gospel}{John 14:15}
\lett{A}{t that time:} Jesus said unto his disciples: If ye love me, keep my commandments. And I will pray the Father, and he shall give you another Comforter, that he may abide with you for ever; Even the Spirit of truth; whom the world cannot receive, because it seeth him not, neither knoweth him: but ye know him; for he dwelleth with you, and shall be in you. I will not leave you comfortless: I will come to you. Yet a little while, and the world seeth me no more; but ye see me: because I live, ye shall live also. At that day ye shall know that I am in my Father, and ye in me, and I in you. He that hath my commandments, and keepeth them, he it is that loveth me: and he that loveth me shall be loved of my Father, and I will love him, and will manifest myself to him. Judas saith unto him, not Iscariot, Lord, how is it that thou wilt manifest thyself unto us, and not unto the world? Jesus answered and said unto him, If a man love me, he will keep my words: and my Father will love him, and we will come unto him, and make our abode with him. He that loveth me not keepeth not my sayings: and the word which ye hear is not mine, but the Father's which sent me. These things have I spoken unto you, being yet present with you. But the Comforter, which is the Holy Ghost, whom the Father will send in my name, he shall teach you all things, and bring all things to your remembrance, whatsoever I have said unto you. Peace I leave with you, my peace I give unto you: not as the world giveth, give I unto you. Let not your heart be troubled, neither let it be afraid. Ye have heard how I said unto you, I go away, and come again unto you. If ye loved me, ye would rejoice, because I said, I go unto the Father: for my Father is greater than I. And now I have told you before it come to pass, that, when it is come to pass, ye might believe. Hereafter I will not talk much with you: for the prince of this world cometh, and hath nothing in me. But that the world may know that I love the Father; and as the Father gave me commandment, even so I do.

\offertory{Stablish the thing, O God, that thou hast wrought in us: for thy temple's sake at Jerusalem shall kings bring presents unto thee, alleluia.}
\secret
\lett{S}{anctify,} we beseech thee, O Lord, the gifts which we offer: and cleanse our hearts by the enlightening of the Holy Spirit. Through.\\

\communion{Suddenly there came a sound from heaven as of a rushing mighty wind, where they were sitting, alleluia: and they were all filled with the Holy Ghost, speaking the wonderful works of God, alleluia, alleluia.}

\postcommunion
\lett{P}{our} thy Holy Spirit upon us, O Lord, and cleanse our hearts: that they may be made fruitful by the inward sprinkling of his dew. Through.

\subby{Whit-Monday}
\feastday{Whit-Monday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{e} fed them with the finest wheat-flour, alleluia: and with honey out of the stony rock hath he satisfied them, alleluia, alleluia. \textit{Ps.} Sing we merrily unto God our strength: make a cheerful noise unto the God of Jacob.
\collect
\lett{S}{end,} we beseech thee, Almighty God, thy Holy Spirit into our hearts, that he may direct and rule us according to thy will, comfort us in all our afflictions, defend us from all error, and lead us into all truth; through Jesus Christ our Lord, who with thee and the same Holy Spirit liveth and reigneth, one God, world without end. \textit{Amen.}
\readingcitation{Epistle}{Acts 10:34}
\lett{I}{n those days:} Peter opened his mouth, and said, Of a truth I perceive that God is no respecter of persons: But in every nation he that feareth him, and worketh righteousness, is accepted with him. The word which God sent unto the children of Israel, preaching peace by Jesus Christ: (he is Lord of all:) That word, I say, ye know, which was published throughout all Judaea, and began from Galilee, after the baptism which John preached; How God anointed Jesus of Nazareth with the Holy Ghost and with power: who went about doing good, and healing all that were oppressed of the devil; for God was with him. And we are witnesses of all things which he did both in the land of the Jews, and in Jerusalem; whom they slew and hanged on a tree: Him God raised up the third day, and shewed him openly; Not to all the people, but unto witnesses chosen before of God, even to us, who did eat and drink with him after he rose from the dead. And he commanded us to preach unto the people, and to testify that it is he which was ordained of God to be the Judge of quick and dead. To him give all the prophets witness, that through his name whosoever believeth in him shall receive remission of sins. While Peter yet spake these words, the Holy Ghost fell on all them which heard the word. And they of the circumcision which believed were astonished, as many as came with Peter, because that on the Gentiles also was poured out the gift of the Holy Ghost. For they heard them speak with tongues, and magnify God. Then answered Peter, Can any man forbid water, that these should not be baptized, which have received the Holy Ghost as well as we? And he commanded them to be baptized in the name of the Lord. Then prayed they him to tarry certain days.

\alleluia{Alleluia, alleluia. ℣. The Apostles spake with other tongues the wonderful works of God, alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}
\readingcitation{Gospel}{John 3:16}
\lett{A}{t that time:} Jesus said unto Nicodemus: God so loved the world, that he gave his only begotten Son, that whosoever believeth in him should not perish, but have everlasting life. For God sent not his Son into the world to condemn the world; but that the world through him might be saved. He that believeth on him is not condemned: but he that believeth not is condemned already, because he hath not believed in the name of the only begotten Son of God. And this is the condemnation, that light is come into the world, and men loved darkness rather than light, because their deeds were evil. For every one that doeth evil hateth the light, neither cometh to the light, lest his deeds should be reproved. But he that doeth truth cometh to the light, that his deeds may be made manifest, that they are wrought in God.
\offertory{The Lord thundered out of heaven, and the Highest gave forth his voice: and the springs of waters were seen, alleluia.}
\secret
\lett{W}{e} beseech thee, O Lord, graciously to sanctify these gifts: and by the acceptable offering of our spiritual sacrifice, render us ourselves an everlasting oblation unto thee. Through.
\communion{The Holy Ghost shall teach you, alleluia: whatsoever I have said unto you, alleluia, alleluia.}
\postcommunion
\lett{A}{ssist,} O Lord, we beseech thee, thy people: and defend from the fury of their enemies those whom thou hast fulfilled with heavenly mysteries. Through.

\subby{Whit-Tuesday}
\feastday{Whit-Tuesday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\properantiphonfix

\introit
\lett{R}{eceive} the joyfulness of your glory, alleluia: giving thanks unto God, alleluia: who hath called you to the heavenly kingdom, alleluia, alleluia, alleluia. \textit{Ps.} Hear my law, O my people: incline your ears unto the words of my mouth.
\collect
\lett{G}{rant,} we beseech thee, merciful God, that thy Church, being gathered together in unity by thy Holy Spirit, may manifest thy power among all peoples, to the glory of thy Name; through Jesus Christ our Lord, who liveth and reigneth with thee and the same Spirit, one God, world without end. \textit{Amen.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\readingcitation{Epistle}{Acts 8:14}
\lett{I}{n those days:} When the apostles which were at Jerusalem heard that Samaria had received the word of God, they sent unto them Peter and John: Who, when they were come down, prayed for them, that they might receive the Holy Ghost: (For as yet he was fallen upon none of them: only they were baptized in the name of the Lord Jesus.) Then laid they their hands on them, and they received the Holy Ghost.
\alleluia{Alleluia, alleluia, ℣. The Holy Ghost shall teach you, whatsoever I have said unto you. Alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\readingcitation{Gospel}{John 10:1}
\lett{A}{t that time:} Jesus said unto the Pharisees: Verily, verily, I say unto you, He that entereth not by the door into the sheepfold, but climbeth up some other way, the same is a thief and a robber. But he that entereth in by the door is the shepherd of the sheep. To him the porter openeth; and the sheep hear his voice: and he calleth his own sheep by name, and leadeth them out. And when he putteth forth his own sheep, he goeth before them, and the sheep follow him: for they know his voice. And a stranger will they not follow, but will flee from him: for they know not the voice of strangers. This parable spake Jesus unto them: but they understood not what things they were which he spake unto them. Then said Jesus unto them again, Verily, verily, I say unto you, I am the door of the sheep. All that ever came before me are thieves and robbers: but the sheep did not hear them. I am the door: by me if any man enter in, he shall be saved, and shall go in and out, and find pasture. The thief cometh not, but for to steal, and to kill, and to destroy: I am come that they might have life, and that they might have it more abundantly.
\offertory{The Lord opened the doors of heaven: he rained down manna also upon them for to eat: he gave them food from heaven, so man did eat Angels' food, alleluia.}
\secret
\lett{M}{ay} the offering of this gift purify us, we beseech thee, O Lord: and render us worthy to partake of holy things. Through.\\

\communion{The Spirit which proceedeth from the Father, alleluia: he shall glorify me, alleluia, alleluia.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that the Holy Ghost may renew our minds by divine sacraments: forasmuch as he is himself the remission of all sins. Through.

\subby{Ember Wednesday in Whitsuntide}
\feastday{Whitsun Emberday}
\begin{inhead}
    {Semidouble}
\end{inhead}
\fancyhead[RE,LO]{Ember Wednesday}

\introit
\lett{O}{God,} when thou wentest forth before the people, journeying and dwelling with them, alleluia: the earth shook, and the heavens dropped, alleluia, alleluia. \textit{Ps.} Let God arise, and let his enemies be scattered: let them also that hate him flee before him.

\collect
\lett{W}{e} beseech thee, O Lord, that the Paraclete, who proceedeth from thee, may enlighten our minds: and lead us, as thy Son hath promised, into all truth; Who liveth.

\readingcitation{Epistle}{Acts 2:14}
\lett{I}{n those days:} Peter, standing up with the eleven, lifted up his voice, and said unto them, Ye men of Judaea, and all ye that dwell at Jerusalem, be this known unto you, and hearken to my words: For these are not drunken, as ye suppose, seeing it is but the third hour of the day. But this is that which was spoken by the prophet Joel; And it shall come to pass in the last days, saith God, I will pour out of my Spirit upon all flesh: and your sons and your daughters shall prophesy, and your young men shall see visions, and your old men shall dream dreams: And on my servants and on my handmaidens I will pour out in those days of my Spirit; and they shall prophesy: And I will shew wonders in heaven above, and signs in the earth beneath; blood, and fire, and vapour of smoke: The sun shall be turned into darkness, and the moon into blood, before that great and notable day of the Lord come: And it shall come to pass, that whosoever shall call on the name of the Lord shall be saved.
\alleluia{Alleluia. ℣. By the word of the Lord were the heavens made, and all the hosts of them by the breath of his mouth.}
\collect
\lett{G}{rant,} we beseech thee, almighty and merciful God: that the Holy Ghost descending on us may by his gracious indwelling render us a temple of his glory. Through.
\begin{rubric}
    From this day until the Saturday following, inclusive, after the Collect of the Day is said the \nth{2} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Acts 5:12}
\lett{I}{n those days:} By the hands of the apostles were many signs and wonders wrought among the people; (and they were all with one accord in Solomon's porch. And of the rest durst no man join himself to them: but the people magnified them. And believers were the more added to the Lord, multitudes both of men and women.) Insomuch that they brought forth the sick into the streets, and laid them on beds and couches, that at the least the shadow of Peter passing by might overshadow some of them. There came also a multitude out of the cities round about unto Jerusalem, bringing sick folks, and them which were vexed with unclean spirits: and they were healed every one.
\alleluia{Alleluia, alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}
\readingcitation{Gospel}{John 6:44}
\lett{A}{t that time}: Jesus said to the multitudes of the Jews: No man can come to me, except the Father which hath sent me draw him: and I will raise him up at the last day. It is written in the prophets, And they shall be all taught of God. Every man therefore that hath heard, and hath learned of the Father, cometh unto me. Not that any man hath seen the Father, save he which is of God, he hath seen the Father. Verily, verily, I say unto you, He that believeth on me hath everlasting life. I am that bread of life. Your fathers did eat manna in the wilderness, and are dead. This is the bread which cometh down from heaven, that a man may eat thereof, and not die. I am the living bread which came down from heaven: if any man eat of this bread, he shall live for ever: and the bread that I will give is my flesh, which I will give for the life of the world.
\offertory{My delight shall be in thy commandments, which I have loved exceedingly: my hands also will I lift up unto thy commandments, which I have loved, alleluia.}
\secret
\lett{A}{ccept,} we beseech thee, O Lord, the gift which we offer: and vouchsafe so to work, that what we perform in a mystery, we may celebrate with godly devotion. Through.
\begin{rubric}
    \nth{2} Secret is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{Peace I leave with you, alleluia: my peace I give unto you, alleluia, alleluia.}
\postcommunion
\lett{W}{e,} O Lord, who receive the heavenly sacraments, beseech thy mercy: that by those things which we perform in this life we may attain unto everlasting joys. Through.
\begin{rubric}
    \nth{2} Postcommunion is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage

\subby{Whit-Thursday}
\feastday{Whit-Thursday}
\begin{inhead}
    {Semidouble}
\end{inhead}
\fancyhead[RE,LO]{}

\properantiphonfix

\introit
\lett{T}{he} Spirit of the Lord hath filled the whole world, alleluia: and that which containeth all things hath knowledge of the voice, alleluia, alleluia, alleluia. \textit{Ps.} Let God arise, and let his enemies be scattered: let them also that hate him flee before him.
\collect
\lett{G}{od,} who as at this time didst teach the hearts of thy faithful people, by sending to them the light of thy Holy Spirit; grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in his holy comfort. Through.
\begin{rubric}
    \nth{2} Collect is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Acts 8:5}
\lett{I}{n those days:} Philip went down to the city of Samaria, and preached Christ unto them. And the people with one accord gave heed unto those things which Philip spake, hearing and seeing the miracles which he did. For unclean spirits, crying with loud voice, came out of many that were possessed with them: and many taken with palsies, and that were lame, were healed. And there was great joy in that city.
\alleluia{Alleluia, alleluia, ℣. O send forth thy Spirit, and they shall be made: and thou shalt renew the face of the earth. Alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}
\readingcitation{Gospel}{Luke 9:1}
\lett{A}{t that time:} Jesus called his twelve disciples together, and gave them power and authority over all devils, and to cure diseases. And he sent them to preach the kingdom of God, and to heal the sick. And he said unto them, Take nothing for your journey, neither staves, nor scrip, neither bread, neither money; neither have two coats apiece. And whatsoever house ye enter into, there abide, and thence depart. And whosoever will not receive you, when ye go out of that city, shake off the very dust from your feet for a testimony against them. And they departed, and went through the towns, preaching the gospel, and healing every where.
\offertory{Stablish the thing, O God, that thou hast wrought in us: for thy temple's sake at Jerusalem shall kings bring presents unto thee, alleluia.}
\secret
\lett{S}{anctify,} we beseech thee, O Lord, the gifts which we offer: and cleanse our hearts by the enlightening of the Holy Spirit. Through.
\begin{rubric}
    \nth{2} Secret is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\communion{Suddenly there came a sound from heaven, as of a rushing mighty wind, where they were sitting, alleluia: and they were all filled with the Holy Ghost speaking the wonderful works of God, alleluia, alleluia.}

\postcommunion
\lett{P}{our} thy Holy Spirit upon us, O Lord, and cleanse our hearts: that they may be made fruitful by the inward sprinkling of his dew. Through.
\begin{rubric}
    \nth{2} Postcommunion is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{Ember Friday in Whitsuntide}
\feastday{Whitsun Emberday}
\begin{inhead}
    {Semidouble}
\end{inhead}
\fancyhead[RE,LO]{Ember Friday}

\introit
\lett{O}{Let} my mouth be filled with thy praise, alleluia: that I may sing, alleluia: my lips will be fain when I sing unto thee, alleluia, alleluia. \textit{Ps.} In thee, O Lord, have I put my trust, let me never be put to confusion: but rid me, and deliver me in thy righteousness.
\collect
\lett{G}{rant,} we beseech thee, O merciful God, to thy Church: that, being gathered together in the Holy Ghost, she may be disturbed by no assaults of the enemy. Through.
\begin{rubric}
    \nth{2} Collect is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\readingcitation{Epistle}{Joel 2:23}
\lett{T}{hus saith the Lord God:} Be glad then, ye children of Zion, and rejoice in the \divineName{Lord} your God: for he hath given you the former rain moderately, and he will cause to come down for you the rain, the former rain, and the latter rain in the first month. And the floors shall be full of wheat, and the fats shall overflow with wine and oil. And I will restore to you the years that the locust hath eaten, the cankerworm, and the caterpiller, and the palmerworm, my great army which I sent among you. And ye shall eat in plenty, and be satisfied, and praise the name of the \divineName{Lord} your God, that hath dealt wondrously with you: and my people shall never be ashamed. And ye shall know that I am in the midst of Israel, and that I am the \divineName{Lord} your God, and none else: and my people shall never be ashamed, saith the Lord Almighty.
\alleluia{Alleluia, alleluia. ℣. O how good and sweet, O Lord, is thy Spirit within us. Alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}
\readingcitation{Gospel}{Luke 5:17}
\lett{A}{t that time:} And it came to pass on a certain day, as Jesus was teaching, that there were Pharisees and doctors of the law sitting by, which were come out of every town of Galilee, and Judaea, and Jerusalem: and the power of the Lord was present to heal them. And, behold, men brought in a bed a man which was taken with a palsy: and they sought means to bring him in, and to lay him before him. And when they could not find by what way they might bring him in because of the multitude, they went upon the housetop, and let him down through the tiling with his couch into the midst before Jesus. And when he saw their faith, he said unto him, Man, thy sins are forgiven thee. And the scribes and the Pharisees began to reason, saying, Who is this which speaketh blasphemies? Who can forgive sins, but God alone? But when Jesus perceived their thoughts, he answering said unto them, What reason ye in your hearts? Whether is easier, to say, Thy sins be forgiven thee; or to say, Rise up and walk? But that ye may know that the Son of man hath power upon earth to forgive sins, (he said unto the sick of the palsy,) I say unto thee, Arise, and take up thy couch, and go into thine house. And immediately he rose up before them, and took up that whereon he lay, and departed to his own house, glorifying God. And they were all amazed, and they glorified God, and were filled with fear, saying, We have seen strange things to day.
\offertory{Praise the Lord, O my soul: while I live will I praise the Lord: yea, as long as I have any being, I will sing praises unto my God, alleluia.}
\secret
\lett{M}{ay} the sacrifices, which we offer in thy sight, O Lord, be consumed by that divine fire, which through the Holy Spirit enkindled the hearts of the disciples of Christ, thy Son. Through the same.
\begin{rubric}
    \nth{2} Secret is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{I will not leave you comfortless: I will come to you again, alleluia: and your heart shall rejoice, alleluia.}
\postcommunion
\lett{W}{e} have received, O Lord, the gifts of thy sacred mystery: humbly beseeching thee; that those things which thou hast commanded us to do in remembrance of thee, may be profitable to the succour of our infirmity. Who livest.
\begin{rubric}
    \nth{2} Postcommunion is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Ember Saturday in Whitsuntide}
\feastday{Whitsun Emberday}
\begin{inhead}
    {Semidouble}
\end{inhead}
\fancyhead[RE,LO]{Ember Saturday}

\introit
\lett{T}{he} love of God is shed abroad in our hearts, alleluia: by the Holy Ghost which dwelleth in us, alleluia, alleluia. \textit{Ps.} Praise the Lord, O my soul: and all that is within me, praise his holy name.
\collect
\lett{W}{e} beseech thee, O Lord, graciously pour the Holy Ghost into our hearts: by whose wisdom we were created, and by whose providence we are governed. Through.

\readingcitation{Epistle}{Joel 2:28}
\lett{T}{hus saith the Lord God:} I will pour out my spirit upon all flesh; and your sons and your daughters shall prophesy, your old men shall dream dreams, your young men shall see visions: And also upon the servants and upon the handmaids in those days will I pour out my spirit. And I will shew wonders in the heavens and in the earth, blood, and fire, and pillars of smoke. The sun shall be turned into darkness, and the moon into blood, before the great and the terrible day of the \divineName{Lord} come. And it shall come to pass, that whosoever shall call on the name of the \divineName{Lord} shall be delivered.
\alleluia{Alleluia. ℣. It is the spirit that quickeneth: the flesh profiteth nothing.}
\collect
\lett{W}{e} beseech thee, O Lord, that the Holy Ghost may inflame us with that fire: which our Lord Jesus Christ sent upon earth, and willed that it should be kindled exceedingly. Who liveth.

\readingcitation{Epistle}{Leviticus 23:9}
\lett{I}{n those days:} The \divineName{Lord} spake unto Moses, saying, Speak unto the children of Israel, and say unto them, When ye be come into the land which I give unto you, and shall reap the harvest thereof, then ye shall bring a sheaf of the firstfruits of your harvest unto the priest: And he shall wave the sheaf before the \divineName{Lord}, to be accepted for you: on the morrow after the sabbath the priest shall wave it. And ye shall offer that day when ye wave the sheaf an he lamb without blemish of the first year for a burnt offering unto the \divineName{Lord}. And the meat offering thereof shall be two tenth deals of fine flour mingled with oil, an offering made by fire unto the \divineName{Lord} for a sweet savour: and the drink offering thereof shall be of wine, the fourth part of an hin. And ye shall eat neither bread, nor parched corn, nor green ears, until the selfsame day that ye have brought an offering unto your God: it shall be a statute for ever throughout your generations in all your dwellings. And ye shall count unto you from the morrow after the sabbath, from the day that ye brought the sheaf of the wave offering; seven sabbaths shall be complete: Even unto the morrow after the seventh sabbath shall ye number fifty days; and ye shall offer a new meat offering unto the \divineName{Lord}. Ye shall bring out of your habitations two wave loaves of two tenth deals: they shall be of fine flour; they shall be baken with leaven; they are the firstfruits unto the \divineName{Lord}. And ye shall offer with the bread seven lambs without blemish of the first year, and one young bullock, and two rams: they shall be for a burnt offering unto the \divineName{Lord}, with their meat offering, and their drink offerings, even an offering made by fire, of sweet savour unto the \divineName{Lord}. Then ye shall sacrifice one kid of the goats for a sin offering, and two lambs of the first year for a sacrifice of peace offerings. And the priest shall wave them with the bread of the firstfruits for a wave offering before the \divineName{Lord}, with the two lambs: they shall be holy to the \divineName{Lord} for the priest. And ye shall proclaim on the selfsame day, that it may be an holy convocation unto you: ye shall do no servile work therein: it shall be a statute for ever in all your dwellings throughout your generations, saith the Lord almighty.
\alleluia{Alleluia. ℣. By his Spirit he hath garnished the heavens.}
\collect
\lett{O}{God,} who for the healing of our souls hast commanded us to chasten our bodies with godly fasting: mercifully grant unto us: that we may ever serve thee both in body and soul. Through.
\readingcitation{Epistle}{Deuteronomy 26:1}
\lett{I}{n those days:} Moses said to the children of Israel: Hear, O Israel, what I command thee this day. When thou art come in unto the land which the \divineName{Lord} thy God giveth thee for an inheritance, and possessest it, and dwellest therein; That thou shalt take of the first of all the fruit of the earth, which thou shalt bring of thy land that the \divineName{Lord} thy God giveth thee, and shalt put it in a basket, and shalt go unto the place which the \divineName{Lord} thy God shall choose to place his name there. And thou shalt go unto the priest that shall be in those days, and say unto him, I profess this day unto the \divineName{Lord} thy God, that I am come unto the country which the \divineName{Lord} sware unto our fathers for to give us. And the priest shall take the basket out of thine hand, and set it down before the altar of the \divineName{Lord} thy God. And thou shalt speak and say before the \divineName{Lord} thy God, A Syrian ready to perish was my father, and he went down into Egypt, and sojourned there with a few, and became there a nation, great, mighty, and populous: And the Egyptians evil entreated us, and afflicted us, and laid upon us hard bondage: And when we cried unto the \divineName{Lord} God of our fathers, the \divineName{Lord} heard our voice, and looked on our affliction, and our labour, and our oppression: And the \divineName{Lord} brought us forth out of Egypt with a mighty hand, and with an outstretched arm, and with great terribleness, and with signs, and with wonders: And he hath brought us into this place, and hath given us this land, even a land that floweth with milk and honey. And now, behold, I have brought the firstfruits of the land, which thou, O \divineName{Lord}, hast given me. And thou shalt set it before the \divineName{Lord} thy God, and worship before the \divineName{Lord} thy God: And thou shalt rejoice in every good thing which the \divineName{Lord} thy God hath given unto thee
\alleluia{Alleluia. ℣. When the day of Pentecost was fully come, they were all sitting with one accord.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, being taught by these saving fasts, and likewise abstaining from all vices, may more readily obtain thy pardon. Through.
\readingcitation{Epistle}{Leviticus 26:3}
\lett{I}{n those days:} The Lord said unto Moses: Speak unto the children of Israel, and say unto them: If ye walk in my statutes, and keep my commandments, and do them; Then I will give you rain in due season, and the land shall yield her increase, and the trees of the field shall yield their fruit. And your threshing shall reach unto the vintage, and the vintage shall reach unto the sowing time: and ye shall eat your bread to the full, and dwell in your land safely. And I will give peace in the land, and ye shall lie down, and none shall make you afraid: and I will rid evil beasts out of the land, neither shall the sword go through your land. And ye shall chase your enemies, and they shall fall before you by the sword. And five of you shall chase an hundred, and an hundred of you shall put ten thousand to flight: and your enemies shall fall before you by the sword. For I will have respect unto you, and make you fruitful, and multiply you, and establish my covenant with you. And ye shall eat old store, and bring forth the old because of the new. And I will set my tabernacle among you: and my soul shall not abhor you. And I will walk among you, and will be your God, and ye shall be my people, saith the Lord almighty.
\alleluia{Alleluia. \inrub{Here genuflect.} ℣. Come, Holy Ghost, fill the hearts of thy faithful: and kindle in them the fire of thy love.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that we may so abstain from carnal feasting; that we may likewise fast from the vices which beset us. Through.
\readingcitation{Epistle}{Song of the Three Children 26}
\lett{I}{n those days:} The Angel of the Lord came down into the oven together with Azarias and his fellows, and smote the flame of the fire out of the oven; And made the midst of the furnace as it had been a moist whistling wind, so that the fire touched them not at all, neither hurt nor troubled them. Then the three, as out of one mouth, praised, glorified, and blessed, God in the furnace, saying,
\begin{rubric}
    Here the response \emph{Thanks be to God}, is not made.
\end{rubric}
\alleluia{Alleluia. ℣. Blessed art thou, O Lord, God of our Fathers, and worthy to be praised for ever.}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\collect
\lett{O}{God,} who to the three children didst assuage the flames of fire; mercifully grant; that the flames of sin may not kindle upon us thy servants. Through.
\begin{rubric}
    \nth{2} Collect is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\readingcitation{Epistle}{Romans 5:1}
\lett{B}{rethren:} Being justified by faith, we have peace with God through our Lord Jesus Christ: By whom also we have access by faith into this grace wherein we stand, and rejoice in hope of the glory of God. And not only so, but we glory in tribulations also: knowing that tribulation worketh patience; And patience, experience; and experience, hope: And hope maketh not ashamed; because the love of God is shed abroad in our hearts by the Holy Ghost which is given unto us.
\tract{O praise the Lord, all ye heathen: praise him, all ye nations, ℣. For his merciful kindness is ever more and more towards us: and the truth of the Lord endureth for ever.}
\begin{rubric}
{The Sequence as on Whit-Sunday (p. \pageref{WhitSeq})}
\end{rubric}
\begin{rubric}
    \textsc{Note,} At the end, \emph{Alleluia} is not said.
\end{rubric}
\readingcitation{Gospel}{Luke 4:38}
\lett{A}{t that time:} Jesus arose out of the synagogue, and entered into Simon's house. And Simon's wife's mother was taken with a great fever; and they besought him for her. And he stood over her, and rebuked the fever; and it left her: and immediately she arose and ministered unto them. Now when the sun was setting, all they that had any sick with divers diseases brought them unto him; and he laid his hands on every one of them, and healed them. And devils also came out of many, crying out, and saying, Thou art Christ the Son of God. And he rebuking them suffered them not to speak: for they knew that he was Christ. And when it was day, he departed and went into a desert place: and the people sought him, and came unto him, and stayed him, that he should not depart from them. And he said unto them, I must preach the kingdom of God to other cities also: for therefore am I sent. And he preached in the synagogues of Galilee.
\offertory{O Lord God of my salvation, I have cried day and night before thee: let my prayer enter into thy presence, O Lord, alleluia.}
\secret
\lett{T}{hat} our fasts may be acceptable to thee, O Lord: grant us, we beseech thee: by the gift of this sacrament to offer unto thee a purified heart. Through.
\begin{rubric}
    \nth{2} Secret is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\communion{The Spirit bloweth where it listeth; and thou hearest the sound thereof, alleluia, alleluia: but canst not tell whence it cometh and whither it goeth, alleluia, alleluia, alleluia.}

\postcommunion
\lett{M}{ay} thy holy things, O Lord, impart to us divine fervour: that we may delight both in the performance and likewise in the fulfilment of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Trinity Sunday}
\feastday{Trinity Sunday}
\begin{inhead}
    {First Class Double}
\end{inhead}
\fancyhead[RE,LO]{}

\properantiphonfix

\introit
\lett{B}{lessed} be the holy Trinity, and the undivided Unity: we will praise and glorify him, because he hath showed his mercy upon us. \textit{Ps.} O Lord our governor: how excellent is thy name in all the world!
\collect
\lett{A}{lmighty} and everlasting God, who hast given unto us thy servants grace by the confession of a true faith to acknowledge the glory of the eternal Trinity, and in the power of the Divine Majesty to worship the Unity: We beseech thee; that thou wouldest keep us stedfast in this faith, and evermore defend us from all adversities. Who livest and reignest, one God, world without end. \textit{Amen.}
\Needspace{4\baselineskip}
\lett{O}{God,} the strength of all those who put their trust in thee, mercifully accept our prayers: and because through the weakness of our mortal nature we can do no good thing without thee, grant us the help of thy grace; that in keeping thy commandments we may please thee, both in will and deed. Through.
%We pick up here from Ian's 1933 Lay English Missal

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\readingcitation{Epistle}{Revelation 4:1}
\lett{I}{n those days:} I looked, and, behold, a door was opened in heaven: and the first voice which I heard was as it were of a trumpet talking with me; which said, Come up hither, and I will shew thee things which must be hereafter. And immediately I was in the spirit: and, behold, a throne was set in heaven, and one sat on the throne. And he that sat was to look upon like a jasper and a sardine stone: and there was a rainbow round about the throne, in sight like unto an emerald. And round about the throne were four and twenty seats: and upon the seats I saw four and twenty elders sitting, clothed in white raiment; and they had on their heads crowns of gold. And out of the throne proceeded lightnings and thunderings and voices: and there were seven lamps of fire burning before the throne, which are the seven Spirits of God. And before the throne there was a sea of glass like unto crystal: and in the midst of the throne, and round about the throne, were four beasts full of eyes before and behind. And the first beast was like a lion, and the second beast like a calf, and the third beast had a face as a man, and the fourth beast was like a flying eagle. And the four beasts had each of them six wings about him; and they were full of eyes within: and they rest not day and night, saying, Holy, holy, holy, Lord God Almighty, which was, and is, and is to come. And when those beasts give glory and honour and thanks to him that sat on the throne, who liveth for ever and ever, The four and twenty elders fall down before him that sat on the throne, and worship him that liveth for ever and ever, and cast their crowns before the throne, saying, Thou art worthy, O Lord, to receive glory and honour and power: for thou hast created all things, and for thy pleasure they are and were created.

\gradall{Blessed art thou, O Lord, that beholdest the depths, and sittest upon the Cherubim. ℣. Blessed art thou, O Lord, in the firmament of heaven, and worthy to be praised for ever.}{Alleluia, alleluia. ℣. Blessed art thou, O Lord God of our fathers, and worthy to be praised for evermore. Alleluia.}

\readingcitation{Gospel}{John 3:1}
\lett{A}{t that time:} There was a man of the Pharisees, named Nicodemus, a ruler of the Jews: The same came to Jesus by night, and said unto him, Rabbi, we know that thou art a teacher come from God: for no man can do these miracles that thou doest, except God be with him. Jesus answered and said unto him, Verily, verily, I say unto thee, Except a man be born again, he cannot see the kingdom of God. Nicodemus saith unto him, How can a man be born when he is old? can he enter the second time into his mother's womb, and be born? Jesus answered, Verily, verily, I say unto thee, Except a man be born of water and of the Spirit, he cannot enter into the kingdom of God. That which is born of the flesh is flesh; and that which is born of the Spirit is spirit. Marvel not that I said unto thee, Ye must be born again. The wind bloweth where it listeth, and thou hearest the sound thereof, but canst not tell whence it cometh, and whither it goeth: so is every one that is born of the Spirit. Nicodemus answered and said unto him, How can these things be? Jesus answered and said unto him, Art thou a master of Israel, and knowest not these things? Verily, verily, I say unto thee, We speak that we do know, and testify that we have seen; and ye receive not our witness. If I have told you earthly things, and ye believe not, how shall ye believe, if I tell you of heavenly things? And no man hath ascended up to heaven, but he that came down from heaven, even the Son of man which is in heaven. And as Moses lifted up the serpent in the wilderness, even so must the Son of man be lifted up: That whosoever believeth in him should not perish, but have eternal life.

\offertory{Blessed be God the Father, and the only-begotten Son of God, and the Holy Spirit: because he hath showed his mercy upon us.}

\secret
\lett{S}{anctify,} we beseech thee, O Lord our God, the sacrifice which we here offer by the invocation of thy holy name: that through the same we may ourselves be made a perfect gift unto thee for evermore. Through.
\Needspace{4\baselineskip}
\lett{D}{o} thou, we beseech thee, O Lord, graciously accept the offerings which have been dedicated unto thee: and grant that they may become to us an abiding source of strength. Through.
\communion{We bless the God of heaven, and will praise him in the sight of all that live: because he hath showed his mercy upon us.}
\postcommunion
\lett{G}{rant}, O Lord our God: that we who receive this sacrament and who acknowledge the holy and everlasting Trinity, and likewise the undivided Unity: may be profited thereby unto salvation of body and soul. Through.
\Needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, O Lord, that we, who have been filled with such great gifts, may both receive these gifts unto our profit, and never cease from thy praise. Through.
\readingcitation{Last Gospel}{Luke 6:36}
\lett{A}{t that time:} Jesus said unto his disciples: Be ye therefore merciful, as your Father also is merciful. Judge not, and ye shall not be judged: condemn not, and ye shall not be condemned: forgive, and ye shall be forgiven: Give, and it shall be given unto you; good measure, pressed down, and shaken together, and running over, shall men give into your bosom. For with the same measure that ye mete withal it shall be measured to you again. And he spake a parable unto them, Can the blind lead the blind? shall they not both fall into the ditch? The disciple is not above his master: but every one that is perfect shall be as his master. And why beholdest thou the mote that is in thy brother's eye, but perceivest not the beam that is in thine own eye? Either how canst thou say to thy brother, Brother, let me pull out the mote that is in thine eye, when thou thyself beholdest not the beam that is in thine own eye? Thou hypocrite, cast out first the beam out of thine own eye, and then shalt thou see clearly to pull out the mote that is in thy brother's eye.

\subby{The Most Holy Body of Christ}
\feastday{Corpus Christi}
\begin{inhead}
	{Thursday after Trinity Sunday}\par
    {First Class Double, Second Class Octave}
\end{inhead}

\textit{Opening Sentence.} Thou gavest them Bread from heaven. Containing in itself all sweetness.\vr{Wis. 16:20}

\introit
\lett{H}{e} fed them with the finest wheat-flour, alleluia: and with honey out of the stony rock hath he satisfied them, alleluia, alleluia, alleluia. \textit{Ps.} Sing we merrily unto God, our strength: make a cheerful noise unto the God of Jacob.
\collect
\lett{O}{God,} who under a wonderful sacrament hast left unto us a memorial of thy Passion: grant us, we beseech thee, so to venerate the sacred mysteries of thy Body and Blood; that we may ever perceive within ourselves the fruit of thy redemption. Who livest and reignest with God the Father.
\readingcitation{Epistle}{1 Corinthians 11:23}
\lett{B}{rethren:} I have received of the Lord that which also I delivered unto you, That the Lord Jesus the same night in which he was betrayed took bread: and when he had given thanks, he brake it, and said, Take, eat: this is my body, which is broken for you: this do in remembrance of me. After the same manner also he took the cup, when he had supped, saying, This cup is the new testament in my blood: this do ye, as oft as ye drink it, in remembrance of me. For as often as ye eat this bread, and drink this cup, ye do shew the Lord's death till he come. Wherefore whosoever shall eat this bread, and drink this cup of the Lord, unworthily, shall be guilty of the body and blood of the Lord. But let a man examine himself, and so let him eat of that bread, and drink of that cup. For he that eateth and drinketh unworthily, eateth and drinketh damnation to himself, not discerning the Lord's body.

\gradual{The eyes of all wait upon thee, O Lord: and thou givest them their meat in due season. ℣. Thou openest thine hand: and fillest all things living with plenteousness.}

\alleluia{Alleluia, alleluia. ℣. My flesh is meat indeed, and my blood is drink indeed: he that eateth my flesh and drinketh my blood, dwelleth in me, and I in him.}

\subsubsec{Sequence}
%\begin{multicols}{2}
\lett{L}{aud,} O Sion, thy salvation, laud with hymns of exultation Christ, thy King and Shepherd true.\par
\secondline{Spend thyself, his honour raising: who surpasseth all thy praising, never canst thou reach his due.}
\thirdline{Sing to-day, the mystery shewing of the living, life-bestowing bread from heaven before thee set.}
E'en the same of old provided, where the twelve, divinely guided, at the holy table met.\par
Full and clear ring out thy chanting, joy nor sweetest grace be wanting to thy heart and soul today.\par
When we gather up the measure of that supper and its treasure, keeping feast in glad array.\par
Lo, the new King's table gracing, this new Passover of blessing hath fulfilled the elder rite.\par
Now the new the old effaceth, truth revealed the shadow chaseth, day is breaking on the night.\par
What he did at supper seated, Christ ordained to be repeated, his memorial ne'er to cease.\par
And, his word for guidance taking, bread and wine we hallow, making thus our sacrifice of peace.\par
This the truth to Christians given--bread becomes his flesh from heaven, wine becomes his holy blood.\par
Doth it pass thy comprehending? yet by faith, thy sight transcending, wondrous things are understood.\par
Yea, beneath these sign are hidden glorious things to sight forbidden, signs, not things, are all we see.\par
Wine is poured and bread is broken: but in either sacred token Christ entire we know to be.\par
Whoso of this food partaketh rendeth not the Lord nor breaketh: Christ is whole to all that taste.\par
Thousands are, as one, receivers: one, as thousands of believers, takes the Food that cannot waste.\par
Good and evil men are sharing one repast: a doom preparing varied as the heart of man.\par
Doom of life or death awarded: as their days shall be recorded which from one beginning ran.\par
When the Sacrament is broken, doubt not in each severed token, hallowed by the word once spoken, resteth all the true content.\par
Nought the precious gift divideth: breaking but the sign betideth: he himself the same abideth, nothing of his fulness spent.\par
Lo! the Angels' Food is given to the pilgrim who hath striven: see the children's Bread from heaven, which to dogs may not be cast.\par
Truth the ancient types fulfilling, Isaac bound, a Victim willing: Paschal Lamb, its life-blood spilling: Manna sent in ages past.\par
Very Bread, Good Shepherd, tend us: Jesu, of thy love befriend us: thou refresh us, thou defend us: thine eternal goodness send us in the land of life to see.\par
Thou who all things canst and knowest: who on earth such food bestowest: grant us with thy Saints, though lowest, where the heavenly feast thou showest, fellow-heirs and guests to be. Amen. Alleluia.
%\end{multicols}

\readingcitation{Gospel}{John 6:55}
\lett{A}{t that time:} Jesus said to the multitudes of the Jews: My flesh is meat indeed, and my blood is drink indeed. He that eateth my flesh, and drinketh my blood, dwelleth in me, and I in him. As the living Father hath sent me, and I live by the Father, so he that eateth me, even he shall live by me. This is that bread which came down from heaven; not as your fathers did eat manna and are dead; he that eateth of this bread shall live for ever.
\offertory{The priests of the Lord offer incense and bread unto God: and therefore shall they be holy unto their God, and shall not profane his name, alleluia.}
\secret
\lett{G}{raciously} grant, we beseech thee, O Lord, unto thy Church the gifts of unity and peace: which are shewn forth in a mystery in the gifts we offer. Through.
\communion{As often as ye eat this bread, and drink this cup, ye do show the Lord's death till he come: wherefore whosoever shall eat this bread or drink this cup of the Lord, unworthily, shall be guilty of the body and blood of the Lord, alleluia.}
\postcommunion
\lett{W}{e} beseech thee, O Lord: that like as the receiving of thy precious Body and Blood in this life doth foreshadow the everlasting fruition of thy Godhead, so thou wouldest vouchsafe unto us to be fulfilled with the same. Who livest.
\begin{rubric}
    Within the Octave (Semidouble) and on the Octave Day (Greater Double), Mass is said as on the Feast.
\end{rubric}
\begin{rubric}
    Within the Octave are added, according to the Rubrics, the Seasonal Prayers, \nth{2} of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\subby{Sunday in the Octave of Corpus Christi}
\feastday{Trinity I}
\centerline{\small{(First Sunday after Trinity)}}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{Lord,} in thy loving-kindness have I trusted: and my heart is joyful in thy salvation: I will sing unto the Lord, because he hath dealt so lovingly with me. \textit{Ps.} How long wilt thou forget me, O Lord, for ever? How long wilt thou hide thy face from me?
\begin{rubric}
    \emph{Gloria in excelsis} is said on all Sundays after Trinity. But it is not said on Ferial Days, when the Mass of the preceding Sunday is resumed.
\end{rubric}
\collect
\lett{O}{God,} the strength of all those who put their trust in thee, mercifully accept our prayers: and because through the weakness of our mortal nature we can do no good thing without thee, grant us the help of thy grace; that in keeping thy commandments we may please thee, both in will and deed. Through.
\Needspace{4\baselineskip}
\lett{O}{God,} who under a wonderful sacrament hast left unto us a memorial of thy Passion: grant us, we beseech thee, so to venerate the sacred mysteries of thy Body and Blood; that we may ever perceive within ourselves the fruit of thy redemption. Who livest and reignest with God the Father.
\readingcitation{Epistle}{1 John 4:7}
\lett{D}{early beloved:} Let us love one another: for love is of God; and every one that loveth is born of God, and knoweth God. He that loveth not knoweth not God; for God is love. In this was manifested the love of God toward us, because that God sent his only begotten Son into the world, that we might live through him. Herein is love, not that we loved God, but that he loved us, and sent his Son to be the propitiation for our sins. Beloved, if God so loved us, we ought also to love one another. No man hath seen God at any time. If we love one another, God dwelleth in us, and his love is perfected in us. Hereby know we that we dwell in him, and he in us, because he hath given us of his Spirit. And we have seen and do testify that the Father sent the Son to be the Saviour of the world. Whosoever shall confess that Jesus is the Son of God, God dwelleth in him, and he in God. And we have known and believed the love that God hath to us. God is love; and he that dwelleth in love dwelleth in God, and God in him. Herein is our love made perfect, that we may have boldness in the day of judgment: because as he is, so are we in this world. There is no fear in love; but perfect love casteth out fear: because fear hath torment. He that feareth is not made perfect in love. We love him, because he first loved us. If a man say, I love God, and hateth his brother, he is a liar: for he that loveth not his brother whom he hath seen, how can he love God whom he hath not seen? And this commandment have we from him, That he who loveth God love his brother also.

\gradall{I said, Lord, be merciful unto me: heal my soul, for I have sinned against thee. ℣. Blessed is he that considereth the poor and needy: the Lord shall deliver him in the time of trouble.}{Alleluia, alleluia. ℣. Ponder my words, O Lord: consider my meditation. Alleluia.}

\readingcitation{Gospel}{Luke 16:19}
\lett{A}{t that time:} Jesus said unto his disciples: There was a certain rich man, which was clothed in purple and fine linen, and fared sumptuously every day: and there was a certain beggar named Lazarus, which was laid at his gate, full of sores, and desiring to be fed with the crumbs which fell from the rich man's table: moreover the dogs came and licked his sores. And it came to pass, that the beggar died, and was carried by the angels into Abraham's bosom: the rich man also died, and was buried; and in hell he lift up his eyes, being in torments, and seeth Abraham afar off, and Lazarus in his bosom. And he cried and said, Father Abraham, have mercy on me, and send Lazarus, that he may dip the tip of his finger in water, and cool my tongue; for I am tormented in this flame. But Abraham said, Son, remember that thou in thy lifetime receivedst thy good things, and likewise Lazarus evil things: but now he is comforted, and thou art tormented. And beside all this, between us and you there is a great gulf fixed: so that they which would pass from hence to you cannot; neither can they pass to us, that would come from thence. Then he said, I pray thee therefore, father, that thou wouldest send him to my father's house: for I have five brethren; that he may testify unto them, lest they also come into this place of torment. Abraham saith unto him, They have Moses and the prophets; let them hear them. And he said, Nay, father Abraham: but if one went unto them from the dead, they will repent. And he said unto him, If they hear not Moses and the prophets, neither will they be persuaded, though one rose from the dead.
\begin{rubric}
    The Creed is said on all Sundays after Trinity. But it is not said on Ferial Days, when the Mass of the preceding Sunday is resumed.
\end{rubric}
\offertory{O hearken thou unto the voice of my calling, my King and my God: for unto thee, O Lord, will I make my prayer.}
\secret
\lett{D}{o} thou, we beseech thee, O Lord, graciously accept the offerings which have been dedicated unto thee: and grant that they may become to us an abiding source of strength. Through.
\Needspace{4\baselineskip}
\lett{G}{raciously} grant, we beseech thee, O Lord, unto thy Church the gifts of unity and peace: which are shewn forth in a mystery in the gifts we offer. Through.

\communion{I will speak of all thy marvellous works: I will be glad and rejoice in thee: yea, my songs will I make of thy name, O Most Highest.}

\postcommunion
\lett{O}{Lord}, who hast filled us with such great gifts: grant, we beseech thee; that we may both receive these saving gifts, and never cease from thy praise. Through.
\Needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord: that like as the receiving of thy precious Body and Blood in this life doth foreshadow the everlasting fruition of thy Godhead, so thou wouldest vouchsafe unto us to be fulfilled with the same. Who livest.


\subby{Compassion of Our Lord Jesus Christ}
\feastday{Divine Compassion}
\begin{inhead}
	{Friday after Trinity I}\par
    {Second Class Double, Simple Octave}
\end{inhead}
\par\noindent
%Taken from the 1909 Dominican Breviary's Sacerdotal Versicle for Sacred Heart (https://archive.org/details/BrevariumIuxtaRitumSacriOrdinisPraedicatorum/black-and-white%20version/Breviarium_Cormier_1909_posterior_cropped%20%285.18x6.5%29%20pg%201-1374%20%28vol.%202%29/page/n465/mode/2up). Coverdale used as translation.
\textit{Opening Sentence.} The Lord is full of compassion and mercy: long-suffering, and of great goodness.

\introit
\lett{T}{he} thoughts of his Heart are from generation to generation: to deliver their soul from death, and to feed them in time of dearth (Alleluia, alleluia). \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.
\collect
\lett{O}{God,} who in the Heart of thy Son, wounded by our sins, dost vouchsafe mercifully to bestow upon us the infinite treasures of love: grant, we beseech thee; that we, giving him the homage of our devotion and piety, may likewise perform the duty of worthy satisfaction. Through the same.
\readingcitation{Epistle}{Ephesians 3:8}
\lett{B}{rethren:} Unto me, who am less than the least of all saints, is this grace given, that I should preach among the Gentiles the unsearchable riches of Christ, and to make all men see what is the fellowship of the mystery, which from the beginning of the world hath been hid in God, who created all things by Jesus Christ: to the intent that now unto the principalities and powers in heavenly places might be known by the church the manifold wisdom of God, according to the eternal purpose which he purposed in Christ Jesus our Lord, in whom we have boldness and access with confidence by the faith of him. For this cause I bow my knees unto the Father of our Lord Jesus Christ, of whom the whole family in heaven and earth is named, that he would grant you, according to the riches of his glory, to be strengthened with might by his Spirit in the inner man; that Christ may dwell in your hearts by faith; that ye, being rooted and grounded in love, may be able to comprehend with all saints what is the breadth, and length, and depth, and height; and to know the love of Christ, which passeth knowledge, that ye might be filled with all the fulness of God.

%MANUAL ADJUSTMENT:
\gradual{Gracious and righteous is the Lord: therefore will he teach sinners in the way. ℣. Them that are meek shall he guide in judgment, and such as are gentle, them shall he learn his way.}

\alleluia{Alleluia, alleluia. ℣. Take my yoke upon you and learn of me, for I am meet and lowly in heart, and ye shall find rest unto your souls. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\textit{Tract.} The Lord is full of compassion and mercy, long suffering and of great goodness. ℣. He will not always be chiding, neither keepeth he his anger for ever. ℣. He hath not dealt with us after our sins, nor rewarded us according to our wickedness.

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\textit{Alleluia.} Alleluia, alleluia. ℣. Take my yoke upon you and learn of me, for I am meet and lowly in heart, and ye shall find rest unto your souls. Alleluia. ℣. Come unto me all ye that labor and are heavy laden, and I will refresh you. Alleluia.
\readingcitation{Gospel}{John 19:31}
\lett{A}{t that time:} The Jews, because it was the Preparation, that the bodies should not remain upon the cross on the sabbath day, (for that sabbath day was an high day,) besought Pilate that their legs might be broken, and that they might be taken away. Then came the soldiers, and brake the legs of the first, and of the other which was crucified with him. But when they came to Jesus, and saw that he was dead already, they brake not his legs, but one of the soldiers with a spear pierced his side, and forthwith came there out blood and water. And he that saw it bare record, and his record is true: and he knoweth that he saith true, that ye might believe. For these things were done, that the scripture should be fulfilled, A bone of him shall not be broken. And again another scripture saith, They shall look on him whom they pierced.
\offertory{Thy rebuke hath broken my heart, I am full of heaviness: I looked for some to have pity upon me, but there was no man: neither found I any to comfort me.  Burnt offerings and sacrifice for sin hast thou not required: then said I: Lo I come. In the volume of the book it is written of me that I should fulfill thy will, O my God: I am content to do it, yea thy law is within my heart, alleluia.}
\secret
\lett{R}{egard,} we beseech thee, O Lord, the ineffable charity of the Heart of thy beloved Son: that the gift which we offer may be acceptable unto thee, and avail for the expiation of our offences. Through the same.
\communion{One of the soldiers with a spear pierced his side, and forthwith came there out blood and water. If any man thirst, let him come to me and drink, alleluia, alleluia.}
\postcommunion
\lett{M}{ay} thy holy gifts, O Lord Jesu, endue us with divine fervour: that we, having tasted thereby the delights of thy most sweet Heart, may learn to despise things earthly, and to love things of heaven. Who livest.


\subby{Second Sunday after Trinity}
\feastday{Trinity II}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord was my upholder, he brought me forth also into a place of liberty: he delivered me because he delighted in me. \textit{Ps.} I will love thee, O Lord, my strength: the Lord is my rock, my defence, and my saviour.

\collect
\lett{O}{Lord,} who never failest to help and govern those whom thou dost bring up in thy stedfast fear and love; Keep us, we beseech thee, under the protection of thy good providence, and make us to have a perpetual fear and love of thy holy Name. Through.
\lett{O}{God,} who in the Heart of thy Son, wounded by our sins, dost vouchsafe mercifully to bestow upon us the infinite treasures of love: grant, we beseech thee; that we, giving him the homage of our devotion and piety, may likewise perform the duty of worthy satisfaction. Through the same.

\readingcitation{Epistle}{1 John 3:13}
\lett{D}{early beloved:} Marvel not, my brethren, if the world hate you. We know that we have passed from death unto life, because we love the brethren. He that loveth not his brother abideth in death. Whosoever hateth his brother is a murderer: and ye know that no murderer hath eternal life abiding in him. Hereby perceive we the love of God, because he laid down his life for us: and we ought to lay down our lives for the brethren. But whoso hath this world's good, and seeth his brother have need, and shutteth up his bowels of compassion from him, how dwelleth the love of God in him? My little children, let us not love in word, neither in tongue; but in deed and in truth. And hereby we know that we are of the truth, and shall assure our hearts before him. For if our heart condemn us, God is greater than our heart, and knoweth all things. Beloved, if our heart condemn us not, then have we confidence toward God. And whatsoever we ask, we receive of him, because we keep his commandments, and do those things that are pleasing in his sight. And this is his commandment, That we should believe on the name of his Son Jesus Christ, and love one another, as he gave us commandment. And he that keepeth his commandments dwelleth in him, and he in him. And hereby we know that he abideth in us, by the Spirit which he hath given us.

\gradall{When I was in trouble I called upon the Lord, and he heard me. ℣. Deliver my soul, O Lord, from lying lips, and from a deceitful tongue.}{Alleluia, alleluia. ℣. O Lord my God, in thee have I put my trust: save me from them that persecute me, and deliver me. Alleluia.}

\readingcitation{Gospel}{Luke 14:16}
\lett{A}{t that time:} Jesus spake this parable unto the Pharisees: A certain man made a great supper, and bade many: And sent his servant at supper time to say to them that were bidden, Come; for all things are now ready. And they all with one consent began to make excuse. The first said unto him, I have bought a piece of ground, and I must needs go and see it: I pray thee have me excused. And another said, I have bought five yoke of oxen, and I go to prove them: I pray thee have me excused. And another said, I have married a wife, and therefore I cannot come. So that servant came, and shewed his lord these things. Then the master of the house being angry said to his servant, Go out quickly into the streets and lanes of the city, and bring in hither the poor, and the maimed, and the halt, and the blind. And the servant said, Lord, it is done as thou hast commanded, and yet there is room. And the lord said unto the servant, Go out into the highways and hedges, and compel them to come in, that my house may be filled. For I say unto you, That none of those men which were bidden shall taste of my supper.

\offertory{Turn thee, O Lord, and deliver my soul: O save me for thy mercy's sake.}

\secret
\lett{M}{ay} the oblation to be offered to thy Name purify us, O Lord: and day by day renew us to the attainment of heavenly life. Through.\\

\lett{R}{egard,} we beseech thee, O Lord, the ineffable charity of the Heart of thy beloved Son: that the gift which we offer may be acceptable unto thee, and avail for the expiation of our offences. Through the same.

\communion{I will sing of the Lord, because he hath dealt so lovingly with me: yea, I will praise the name of the Lord Most Highest.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received thy sacred gifts: may by the frequenting of this mystery set forward the work of our salvation. Through.
\lett{M}{ay} thy holy gifts, O Lord Jesu, endue us with divine fervour: that we, having tasted thereby the delights of thy most sweet Heart, may learn to despise things earthly, and to love things of heaven. Who livest.
\begin{rubric}
    On the Octave Day of the Compassion of Our Lord (Greater Double), Mass is said as on the Feast.
\end{rubric}

\subby{Third Sunday after Trinity}\label{TrinityIII}
\feastday{Trinity III}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{urn} thee unto me, and have mercy upon me, O Lord: for I am desolate and in misery: look thou on mine affliction and my travail: and forgive me all mine iniquities, O my God. \textit{Ps.} Unto thee, O Lord, do I lift up my soul: my God, in thee have I trusted, let me not be confounded.

\collect\label{TrinityIIICollect}
\lett{O}{Lord,} we beseech thee mercifully to hear us; and grant that we, to whom thou hast given an hearty desire to pray, may, by thy mighty aid, be defended and comforted in all dangers and adversities. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 Peter 5:5}
\lett{D}{early beloved:} All of you be subject one to another, and be clothed with humility: for God resisteth the proud, and giveth grace to the humble. Humble yourselves therefore under the mighty hand of God, that he may exalt you in due time: casting all your care upon him; for he careth for you. Be sober, be vigilant; because your adversary the devil, as a roaring lion, walketh about, seeking whom he may devour: whom resist stedfast in the faith, knowing that the same afflictions are accomplished in your brethren that are in the world. But the God of all grace, who hath called us unto his eternal glory by Christ Jesus, after that ye have suffered a while, make you perfect, stablish, strengthen, settle you. To him be glory and dominion for ever and ever. Amen.

\gradall{O cast thy burden upon the Lord: and he shall nourish thee. ℣. When I called upon the Lord, he heard my voice from the battle that was against me.}{Alleluia, alleluia. ℣. God is a righteous Judge, strong and patient, and God is provoked every day. Alleluia.}

\readingcitation{Gospel}{Luke 15:1}
\lett{A}{t that time:} There drew near unto Jesus all the publicans and sinners for to hear him. And the Pharisees and scribes murmured, saying, This man receiveth sinners, and eateth with them. And he spake this parable unto them saying, What man of you, having an hundred sheep, if he lose one of them, doth not leave the ninety and nine in the wilderness, and go after that which is lost, until he find it? And when he hath found it, he layeth it on his shoulders, rejoicing. And when he cometh home, he calleth together his friends and neighbours, saying unto them, Rejoice with me; for I have found my sheep which was lost. I say unto you, that likewise joy shall be in heaven over one sinner that repenteth, more than over ninety and nine just persons, which need no repentance. Either what woman having ten pieces of silver, if she lose one piece, doth not light a candle, and sweep the house, and seek diligently till she find it? And when she hath found it, she calleth her friends and her neighbours together, saying, Rejoice with me; for I have found the piece which I had lost. Likewise, I say unto you, there is joy in the presence of the angels of God over one sinner that repenteth.

\offertory{They that know thy name will put their trust in thee: for thou, Lord, hast never failed them that seek thee: O praise the Lord which dwelleth in Sion: for he forgetteth not the complaint of the poor.}

\secret\label{TrinityIIISecret}
\lett{S}{anctify,} O Lord, we beseech thee, the gifts which we offer unto thee: that they may be made unto us the Body and Blood of thine only-begotten Son. Who liveth.

\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{I say unto you: There is joy in the presence of the Angels of God over one sinner that repenteth.}

\postcommunion\label{TrinityIIIPostcommunion}
\lett{O}{Lord,} through whom we have received these sacred gifts: we beseech thee; that by the power of the same, thou wouldest both cleanse us from all vices, and evermore fulfil us with the gifts of thy grace. Through.

\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Fourth Sunday after Trinity}
\feastday{Trinity IV}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord is my light, and my salvation, whom then shall I fear? The Lord is the stronghold of my life, of whom shall I be afraid? When mine enemies pressed sore against me, they stumbled and fell. \textit{Ps.} Though an host of men were laid against me, yet shall not my heart be afraid.

\collect
\lett{O}{God,} the protector of all that trust in thee, without whom nothing is strong, nothing is holy; Increase and multiply upon us thy mercy; that, thou being our ruler and guide, we may so pass through things temporal, that we finally lose not the things eternal. Grant this, O heavenly Father, for the sake of Jesus Christ our Lord. Who liveth.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Romans 8:18}
\lett{B}{rethren:} I reckon that the sufferings of this present time are not worthy to be compared with the glory which shall be revealed in us. For the earnest expectation of the creature waiteth for the manifestation of the sons of God. For the creature was made subject to vanity, not willingly, but by reason of him who hath subjected the same in hope, because the creature itself also shall be delivered from the bondage of corruption into the glorious liberty of the children of God. For we know that the whole creation groaneth and travaileth in pain together until now. And not only they, but ourselves also, which have the firstfruits of the Spirit, even we ourselves groan within ourselves, waiting for the adoption, to wit, the redemption of our body: in Christ Jesus our Lord.

\gradall{Be merciful, O Lord, unto our sins: wherefore do the heathen say: Where is now their God? ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us.}{Alleluia, alleluia. ℣. O God, who art set in the throne and judgest right: be thou the refuge of the oppressed in time of trouble. Alleluia.}

\readingcitation{Gospel}{Luke 6:36}
\lett{A}{t that time:} Jesus said unto his disciples: Be ye therefore merciful, as your Father also is merciful. Judge not, and ye shall not be judged: condemn not, and ye shall not be condemned: forgive, and ye shall be forgiven: give, and it shall be given unto you; good measure, pressed down, and shaken together, and running over, shall men give into your bosom. For with the same measure that ye mete withal it shall be measured to you again. And he spake a parable unto them, Can the blind lead the blind? shall they not both fall into the ditch? The disciple is not above his master: but every one that is perfect shall be as his master. And why beholdest thou the mote that is in thy brother's eye, but perceivest not the beam that is in thine own eye? Either how canst thou say to thy brother, Brother, let me pull out the mote that is in thine eye, when thou thyself beholdest not the beam that is in thine own eye? Thou hypocrite, cast out first the beam out of thine own eye, and then shalt thou see clearly to pull out the mote that is in thy brother's eye.

\offertory{Lighten mine eyes, that I sleep not in death: lest mine enemies says: I have prevailed against him.}

\secret
\lett{R}{egard,} O Lord, the gifts of thy suppliant Church: and grant that the partaking thereof may avail for the salvation and continual sanctification of them that believe. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\communion{The Lord is my strong rock and my defence, my saviour, my God, and my might.}

\postcommunion
\lett{M}{ay} the holy things, which we have received, quicken us, O Lord: that we being cleansed thereby may be made ready for thine everlasting mercy. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Fifth Sunday after Trinity}
\feastday{Trinity V}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{earken} unto my voice, O Lord, when I cry unto thee: be thou my succour, O cast me not away, neither forsake me, O God of my salvation. \textit{Ps.} The Lord is my light and my salvation, whom then shall I fear?

\collect
\lett{G}{rant,} O Lord, we beseech thee, that the course of this world may be so peaceably ordered by thy governance, that thy Church may joyfully serve thee in all godly quietness. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\readingcitation{Epistle}{1 Peter 3:8}
\lett{D}{early beloved:} Be ye all of one mind, having compassion one of another, love as brethren, be pitiful, be courteous: Not rendering evil for evil, or railing for railing: but contrariwise blessing; knowing that ye are thereunto called, that ye should inherit a blessing. For he that will love life, and see good days, let him refrain his tongue from evil, and his lips that they speak no guile: Let him eschew evil, and do good; let him seek peace, and ensue it. For the eyes of the Lord are over the righteous, and his ears are open unto their prayers: but the face of the Lord is against them that do evil. And who is he that will harm you, if ye be followers of that which is good? But and if ye suffer for righteousness' sake, happy are ye: and be not afraid of their terror, neither be troubled; But sanctify the Lord God in your hearts.

\gradall{Behold, O God, our defender: and look upon thy servants. ℣. O Lord God of hosts, hear the prayer of thy servants.}{Alleluia, alleluia. ℣. The king shall rejoice in thy strength, O Lord: exceeding glad shall he be of thy salvation. Alleluia.}

\readingcitation{Gospel}{Luke 5:1}
\lett{A}{t that time:} It came to pass, that, as the people pressed upon Jesus to hear the word of God, he stood by the lake of Gennesaret, and saw two ships standing by the lake: but the fishermen were gone out of them, and were washing their nets. And he entered into one of the ships, which was Simon's, and prayed him that he would thrust out a little from the land. And he sat down, and taught the people out of the ship. Now when he had left speaking, he said unto Simon, Launch out into the deep, and let down your nets for a draught. And Simon answering said unto him, Master, we have toiled all the night, and have taken nothing: nevertheless at thy word I will let down the net. And when they had this done, they inclosed a great multitude of fishes: and their net brake. And they beckoned unto their partners, which were in the other ship, that they should come and help them. And they came, and filled both the ships, so that they began to sink. When Simon Peter saw it, he fell down at Jesus' knees, saying, Depart from me; for I am a sinful man, O Lord. For he was astonished, and all that were with him, at the draught of the fishes which they had taken: and so was also James, and John, the sons of Zebedee, which were partners with Simon. And Jesus said unto Simon, Fear not; from henceforth thou shalt catch men. And when they had brought their ships to land, they forsook all, and followed him.

\offertory{I will bless the Lord who hath given me counsel: I have set God always before me: for he is on my right hand, therefore I shall not fall.}

\secret
\lett{O}{Lord,} we beseech thee mercifully to receive our oblations: and graciously turn our rebel wills to thee. Through.

\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{One thing have I desired of the Lord, which I will require: even that I may dwell in the house of the Lord all the days of my life.}

\postcommunion
\lett{W}{e} beseech thee, O Lord: that the mysteries which we have received may purify us: and defend us by the gift which they bestow. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Sixth Sunday after Trinity}
\feastday{Trinity VI}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{he} Lord is the strength of his people, and a stronghold of salvation to his Anointed: O Lord, save thy people, and give thy blessing unto thine inheritance: O feed them also for ever. \textit{Ps.} Unto thee will I cry, O Lord my God, be not silent unto me: lest, if thou make as though thou hearest not, I become like them that go down into the pit.

\collect
\lett{O}{God,} who hast prepared for them that love thee such good things as pass man's understanding: pour into our hearts such love toward thee: that we, loving thee above all things, may obtain thy promises, which exceed all that we can desire. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Romans 6:3}
\lett{B}{rethren:} Know ye not, that so many of us as were baptized into Jesus Christ were baptized into his death? Therefore we are buried with him by baptism into death: that like as Christ was raised up from the dead by the glory of the Father, even so we also should walk in newness of life. For if we have been planted together in the likeness of his death, we shall be also in the likeness of his resurrection: Knowing this, that our old man is crucified with him, that the body of sin might be destroyed, that henceforth we should not serve sin. For he that is dead is freed from sin. Now if we be dead with Christ, we believe that we shall also live with him: Knowing that Christ being raised from the dead dieth no more; death hath no more dominion over him. For in that he died, he died unto sin once: but in that he liveth, he liveth unto God. Likewise reckon ye also yourselves to be dead indeed unto sin, but alive unto God through Jesus Christ our Lord.

\gradall{Turn thee again, O Lord, at the last, and be gracious unto thy servants. ℣. Lord, thou hast been our refuge, from one generation to another.}{Alleluia, alleluia. ℣. In thee, O Lord, have I put my trust, let me never be put to confusion: rid me and deliver me in thy righteousness: bow down thine ear to me, make haste to deliver me. Alleluia.}

\readingcitation{Gospel}{Matthew 5:20}
\lett{A}{t that time:} Jesus said unto his disciples, Except your righteousness shall exceed the righteousness of the scribes and Pharisees, ye shall in no case enter into the kingdom of heaven. Ye have heard that it was said of them of old time, Thou shalt not kill; and whosoever shall kill shall be in danger of the judgment: But I say unto you, That whosoever is angry with his brother without a cause shall be in danger of the judgment: and whosoever shall say to his brother, Raca, shall be in danger of the council: but whosoever shall say, Thou fool, shall be in danger of hell fire. Therefore if thou bring thy gift to the altar, and there rememberest that thy brother hath ought against thee; Leave there thy gift before the altar, and go thy way; first be reconciled to thy brother, and then come and offer thy gift. Agree with thine adversary quickly, whiles thou art in the way with him; lest at any time the adversary deliver thee to the judge, and the judge deliver thee to the officer, and thou be cast into prison. Verily I say unto thee, Thou shalt by no means come out thence, till thou hast paid the uttermost farthing.

\offertory{O hold thou my goings in thy paths, that my footsteps slip not: incline thine ear unto me, and hearken unto my words: shew thy marvellous lovingkindness, thou that art the Saviour of them which put their trust in thee, O Lord.}

\secret
\lett{B}{e} merciful, O Lord, unto our supplications: and graciously receive these oblations of thy servants and handmaids; that those things which each hath offered to the honour of thy name may be profitable unto all for their salvation. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{I will offer in his dwelling an oblation with great gladness: I will sing, and speak praises unto the Lord.}

\postcommunion
\lett{O}{Lord} who hast satisfied us with thy heavenly gift: grant, we beseech thee; that we may be cleansed from our secret faults, and delivered from the snares of our enemies. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subby{Seventh Sunday after Trinity}
\feastday{Trinity VII}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{clap} your hands together, all ye people: O sing unto God with the voice of melody. \textit{Ps.} For the Lord is high, and to be feared: he is the great King upon all the earth.

\collect
\lett{L}{ord} of all power and might, who art the author and giver of all good things; Graft in our hearts the love of thy Name, increase in us true religion, nourish us with all goodness, and of thy great mercy keep us in the same. Through.

\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Romans 6:19}
\lett{B}{rethren:} I speak after the manner of men because of the infirmity of your flesh: for as ye have yielded your members servants to uncleanness and to iniquity unto iniquity; even so now yield your members servants to righteousness unto holiness. For when ye were the servants of sin, ye were free from righteousness. What fruit had ye then in those things whereof ye are now ashamed? for the end of those things is death. But now being made free from sin, and become servants to God, ye have your fruit unto holiness, and the end everlasting life. For the wages of sin is death; but the gift of God is eternal life, in Christ Jesus our Lord.

\gradall{Come ye children, and hearken unto me: I will teach you the fear of the Lord. ℣. They had an eye unto him and were enlightened: and their faces were not ashamed.}{Alleluia, alleluia. ℣. O clap your hands together, all ye people: O sing unto God with the voice of melody. Alleluia.}

\readingcitation{Gospel}{Mark 8:1}
\lett{I}{n those days:} The multitude being very great, and having nothing to eat, Jesus called his disciples unto him, and saith unto them, I have compassion on the multitude, because they have now been with me three days, and have nothing to eat: And if I send them away fasting to their own houses, they will faint by the way: for divers of them came from far. And his disciples answered him, From whence can a man satisfy these men with bread here in the wilderness? And he asked them, How many loaves have ye? And they said, Seven. And he commanded the people to sit down on the ground: and he took the seven loaves, and gave thanks, and brake, and gave to his disciples to set before them; and they did set them before the people. And they had a few small fishes: and he blessed, and commanded to set them also before them. So they did eat, and were filled: and they took up of the broken meat that was left seven baskets. And they that had eaten were about four thousand: and he sent them away.

\offertory{Like as in the burnt offerings of rams and bullocks, and like as in ten thousands of fat lambs: so let our sacrifice be in thy sight this day, that it may please thee: for they shall not be confounded that put their trust in thee, O Lord.}

\secret
\lett{B}{e} merciful, O Lord, unto our supplications, and graciously receive these oblations of thy people: and that none may pray amiss or fail in his petition, grant; that what we ask faithfully, we may obtain effectually. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Bow down thine ear, make haste to deliver me.}

\postcommunion
\lett{W}{e} have been fulfilled with thy gifts, O Lord: grant, we beseech thee; that we may both be cleansed by their operation, and defended by their succour. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Eighth Sunday after Trinity}
\feastday{Trinity VIII}
\begin{inhead}
    {Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{e} have waited, O God, thy loving kindness in the midst of thy temple: according to thy name, O God, so is thy praise unto the world's end: thy right hand is full of righteousness. \textit{Ps.} Great is the Lord, and highly to be praised: in the city of our God, even upon his holy hill.

\collect
\lett{O}{God,} whose never-failing providence ordereth all things both in heaven and earth; We humbly beseech thee to put away from us all hurtful things, and to give us those things which are profitable for us. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Romans 6:19}
\lett{B}{rethren:} We are debtors, not to the flesh, to live after the flesh. For if ye live after the flesh, ye shall die: but if ye through the Spirit do mortify the deeds of the body, ye shall live. For as many as are led by the Spirit of God, they are the sons of God. For ye have not received the spirit of bondage again to fear; but ye have received the Spirit of adoption, whereby we cry, Abba, Father. The Spirit itself beareth witness with our spirit, that we are the children of God: And if children, then heirs; heirs of God, and joint-heirs with Christ; if so be that we suffer with him, that we may be also glorified together.

\gradual{Be thou my strong rock and house of defence, that thou mayest save me. ℣. In thee, O God, have I put my trust: let me never be put to confusion, O Lord.}

\alleluia{Alleluia, alleluia. ℣. Great is the Lord, and highly to be praised, in the city of our God, even upon his holy hill. Alleluia.}

\readingcitation{Gospel}{Matthew 7:15}
\lett{A}{t that time:} Jesus said unto his disciples: Beware of false prophets, which come to you in sheep's clothing, but inwardly they are ravening wolves. Ye shall know them by their fruits. Do men gather grapes of thorns, or figs of thistles? Even so every good tree bringeth forth good fruit; but a corrupt tree bringeth forth evil fruit. A good tree cannot bring forth evil fruit, neither can a corrupt tree bring forth good fruit. Every tree that bringeth not forth good fruit is hewn down, and cast into the fire. Wherefore by their fruits ye shall know them. Not every one that saith unto me, Lord, Lord, shall enter into the kingdom of heaven; but he that doeth the will of my Father which is in heaven.

\offertory{Thou shalt save the people that are in adversity, O Lord, and shalt bring down the high looks of the proud: for who is God, but thou, O Lord?}

\secret
\lett{O}{God} who by the one perfect sacrifice didst ratify the divers offerings of the law: accept the sacrifice of thy bounden servants, and sanctify it with that blessing wherewith thou didst bless the gifts of Abel: that what each hath offered to the honour of thy majesty may be profitable unto all for their salvation. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{O taste, and see, how gracious the Lord is: blessed is the man that trusteth in him.}

\postcommunion
\lett{L}{et} the operation of thy healing, O Lord, both mercifully set us free from our perversities; and lead us to the things that be right. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Ninth Sunday after Trinity}
\feastday{Trinity IX}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{ehold,} God is my helper: the Lord is he that upholdeth my soul: he shall reward evil unto mine enemies: destroy thou them in thy truth, O Lord, my defender. \textit{Ps.} Save me, O God, for thy name's sake: and avenge me in thy strength.
\collect
\lett{G}{rant} to us, Lord, we beseech thee, the spirit to think and do always such things as be rightful: that we, who cannot exist without thee; may by thee be enabled to live according to thy will. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
\readingcitation{Epistle}{1 Corinthians 10:1}
\lett{B}{rethren:} I would not that ye should be ignorant, how that all our fathers were under the cloud, and all passed through the sea; And were all baptized unto Moses in the cloud and in the sea; And did all eat the same spiritual meat; And did all drink the same spiritual drink: for they drank of that spiritual Rock that followed them: and that Rock was Christ. But with many of them God was not well pleased: for they were overthrown in the wilderness. Now these things were our examples, to the intent we should not lust after evil things, as they also lusted. Neither be ye idolaters, as were some of them; as it is written, The people sat down to eat and drink, and rose up to play. Neither let us commit fornication, as some of them committed, and fell in one day three and twenty thousand. Neither let us tempt Christ, as some of them also tempted, and were destroyed of serpents. Neither murmur ye, as some of them also murmured, and were destroyed of the destroyer. Now all these things happened unto them for ensamples: and they are written for our admonition, upon whom the ends of the world are come. Wherefore let him that thinketh he standeth take heed lest he fall. There hath no temptation taken you but such as is common to man: but God is faithful, who will not suffer you to be tempted above that ye are able; but will with the temptation also make a way to escape, that ye may be able to bear it.

\gradall{O Lord our governor: how excellent is thy name in all the world! ℣. Thou that hast set thy glory above the heavens.}{Alleluia, alleluia. ℣. Deliver me from mine enemies, O God: defend me from them that rise up against me. Alleluia.}

%REVIEW:
%1928 BCP Gospel is thoroughly innovative and departs from the Sarum \& Prayer Book Tradition, so the 1892 is preferred.
\readingcitation{Gospel}{Luke 16:1}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: There was a certain rich man, which had a steward; and the same was accused unto him that he had wasted his goods. And he called him, and said unto him, How is it that I hear this of thee? give an account of thy stewardship; for thou mayest be no longer steward. Then the steward said within himself, What shall I do? for my lord taketh away from me the stewardship: I cannot dig; to beg I am ashamed. I am resolved what to do, that, when I am put out of the stewardship, they may receive me into their houses. So he called every one of his lord's debtors unto him, and said unto the first, How much owest thou unto my lord? And he said, An hundred measures of oil. And he said unto him, Take thy bill, and sit down quickly, and write fifty. Then said he to another, And how much owest thou? And he said, An hundred measures of wheat. And he said unto him, Take thy bill, and write fourscore. And the lord commended the unjust steward, because he had done wisely: for the children of this world are in their generation wiser than the children of light. And I say unto you, Make to yourselves friends of the mammon of unrighteousness; that, when ye fail, they may receive you into everlasting habitations.

\offertory{The statutes of the Lord are right, and rejoice the heart, his judgments are sweeter than honey, and the honey-comb: moreover by them is thy servant taught.}

\secret
\lett{A}{ccept,} we beseech thee, O Lord, the gifts which of thine own bounty we offer unto thee: that, by the mighty working of thy grace, these most sacred mysteries may both sanctify our conversation in this life, and bring us to everlasting joys. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{He that eateth my flesh and drinketh my blood, dwelleth in me, and I in him, saith the Lord.}

\postcommunion
\lett{M}{ay} this heavenly mystery be unto us, O Lord, the renewing both of mind and of body: that like as we do offer unto thee our worship, so we may perceive the benefit of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subsec{Tenth Sunday after Trinity}
\feastday{Trinity X}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{W}{hen} I called upon the Lord, he heard my voice from the battle that was against me: and he hath brought them down, even he that is of old, and endureth for ever: O cast thy burden upon the Lord, and he shall nourish thee. \textit{Ps.} Hear my prayer, O Lord, and hide not thyself from my petition: take heed unto me, and hear me.

\collect
\lett{L}{et} thy merciful ears, O Lord, be open to the prayers of thy humble servants; and that they may obtain their petitions make them to ask such things as shall please thee. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 12:1}
\lett{B}{rethren:} Concerning spiritual gifts, I would not have you ignorant. Ye know that ye were Gentiles, carried away unto these dumb idols, even as ye were led. Wherefore I give you to understand, that no man speaking by the Spirit of God calleth Jesus accursed: and that no man can say that Jesus is the Lord, but by the Holy Ghost. Now there are diversities of gifts, but the same Spirit. And there are differences of administrations, but the same Lord. And there are diversities of operations, but it is the same God which worketh all in all. But the manifestation of the Spirit is given to every man to profit withal. For to one is given by the Spirit the word of wisdom; to another the word of knowledge by the same Spirit; To another faith by the same Spirit; to another the gifts of healing by the same Spirit; To another the working of miracles; to another prophecy; to another discerning of spirits; to another divers kinds of tongues; to another the interpretation of tongues: But all these worketh that one and the selfsame Spirit, dividing to every man severally as he will.

\gradall{Keep me O Lord, as the apple of an eye: hide me under the shadow of thy wings. ℣. Let my sentence come forth from thy presence: and let thine eyes look upon the thing that is equal.}{Alleluia, alleluia. ℣. Thou, O God, art praised in Sion: and unto thee shall the vow be performed in Jerusalem. Alleluia.}

\readingcitation{Gospel}{Luke 19:41}
\lett{A}{t that time:} When Jesus was come near to Jerusalem, he beheld the city, and wept over it, Saying, If thou hadst known, even thou, at least in this thy day, the things which belong unto thy peace! but now they are hid from thine eyes. For the days shall come upon thee, that thine enemies shall cast a trench about thee, and compass thee round, and keep thee in on every side, And shall lay thee even with the ground, and thy children within thee; and they shall not leave in thee one stone upon another; because thou knewest not the time of thy visitation. And he went into the temple, and began to cast out them that sold therein, and them that bought; Saying unto them, It is written, My house is the house of prayer: but ye have made it a den of thieves. And he taught daily in the temple.

\offertory{Unto thee, O Lord, lift I up my soul: O my God, in thee have I trusted, let me not be confounded: neither let mine enemies triumph over me: for all they that hope for thee shall not be ashamed.}

\secret
\lett{G}{rant} to us, we beseech thee, O Lord, worthily to frequent these mysteries: for as oft as the commemoration of this victim is celebrated, the work of our redemption is performed. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Thou shalt be pleased with the sacrifice of righteousness, with the burnt offerings and oblations, upon thine altar, O Lord.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that the communion of thy sacrament may both bestow upon us purification, and grant us purity. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}



\subby{Eleventh Sunday after Trinity}
\feastday{Trinity XI}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{G}{od} is in his holy habitation: he is the God that maketh men to be of one mind in an house: he will give strength and power unto his people. \textit{Ps.} Let God arise, and let his enemies be scattered: let them also that hate him flee before him.

\collect
\lett{O}{God,} who declarest thy almighty power chiefly in showing mercy and pity; Mercifully grant unto us such a measure of thy grace, that we, running the way of thy commandments, may obtain thy gracious promises, and be made partakers of thy heavenly treasure. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 15:1}
\lett{B}{rethren:} I declare unto you the gospel which I preached unto you, which also ye have received, and wherein ye stand; By which also ye are saved, if ye keep in memory what I preached unto you, unless ye have believed in vain. For I delivered unto you first of all that which I also received, how that Christ died for our sins according to the scriptures; And that he was buried, and that he rose again the third day according to the scriptures: And that he was seen of Cephas, then of the twelve: After that, he was seen of above five hundred brethren at once; of whom the greater part remain unto this present, but some are fallen asleep. After that, he was seen of James; then of all the apostles. And last of all he was seen of me also, as of one born out of due time. For I am the least of the apostles, that am not meet to be called an apostle, because I persecuted the church of God. But by the grace of God I am what I am: and his grace which was bestowed upon me was not in vain; but I laboured more abundantly than they all: yet not I, but the grace of God which was with me. Therefore whether it were I or they, so we preach, and so ye believed.

\gradall{My heart hath trusted in God, and I am helped: therefore my heart danceth for joy and in my song will I praise him. ℣. Unto thee will I cry, O Lord: be not silent, O my God, nor depart from me.}{Alleluia, alleluia. ℣. Sing we merrily unto God our strength, make a cheerful noise unto the God of Jacob: take the merry psalm, with the lute. Alleluia.}

\readingcitation{Gospel}{Luke 18:9}
\lett{A}{t that time:} Jesus spake this parable unto certain which trusted in themselves that they were righteous, and despised others: Two men went up into the temple to pray; the one a Pharisee, and the other a publican. The Pharisee stood and prayed thus with himself, God, I thank thee, that I am not as other men are, extortioners, unjust, adulterers, or even as this publican. I fast twice in the week, I give tithes of all that I possess. And the publican, standing afar off, would not lift up so much as his eyes unto heaven, but smote upon his breast, saying, God be merciful to me a sinner. I tell you, this man went down to his house justified rather than the other: for every one that exalteth himself shall be abased; and he that humbleth himself shall be exalted.

\offertory{I will magnify thee, O Lord, for thou hast set me up, and not made my foes to triumph over me: O Lord, I cried unto thee, and thou hast healed me.}

\secret
\lett{L}{et} the sacrifices which we dedicate be offered unto thee, O Lord: which thou hast granted us so to present for the honour of thy name, that they may also be for our healing. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Honour the Lord with thy substance, and with the first-fruits of all thine increase: so shall thy barns be filled with plenty, and thy presses shall burst out with new wine.}

\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that as thou failest not to renew us with thy divine sacraments, so thou wouldest not leave us destitute of thy gracious help. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Twelfth Sunday after Trinity}
\feastday{Trinity XII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{aste} thee, O God, unto my rescue: O Lord make haste to my deliverance: let mine enemies be ashamed and confounded that seek after my soul. \textit{Ps.} Let them be turned backward and put to confusion: that wish me evil.

\collect
\lett{A}{lmighty} and everlasting God, who art always more ready to hear than we to pray, and art wont to give more than either we desire or deserve; Pour down upon us the abundance of thy mercy; forgiving us those things whereof our conscience is afraid, and giving us those good things which we are not worthy to ask, but through the merits and mediation of Jesus Christ, thy Son, our Lord. Who liveth.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{2 Corinthians 3:4}
\lett{B}{rethren:} Such trust have we through Christ to God-ward: Not that we are sufficient of ourselves to think any thing as of ourselves; but our sufficiency is of God; Who also hath made us able ministers of the new testament; not of the letter, but of the spirit: for the letter killeth, but the spirit giveth life. But if the ministration of death, written and engraven in stones, was glorious, so that the children of Israel could not stedfastly behold the face of Moses for the glory of his countenance; which glory was to be done away: How shall not the ministration of the spirit be rather glorious? For if the ministration of condemnation be glory, much more doth the ministration of righteousness exceed in glory.

\gradall{I will alway give thanks unto the Lord: his praise shall ever be in my mouth. ℣. My soul shall make her boast in the Lord: the humble shall hear thereof, and be glad.}{Alleluia, alleluia. ℣. O Lord God of my salvation, I have cried day and night before thee. Alleluia.}

\readingcitation{Gospel}{Mark 7:31}
\lett{A}{t that time:} Jesus, departing from the coasts of Tyre and Sidon, came unto the sea of Galilee, through the midst of the coasts of Decapolis. And they bring unto him one that was deaf, and had an impediment in his speech; and they beseech him to put his hand upon him. And he took him aside from the multitude, and put his fingers into his ears, and he spit, and touched his tongue; And looking up to heaven, he sighed, and saith unto him, Ephphatha, that is, Be opened. And straightway his ears were opened, and the string of his tongue was loosed, and he spake plain. And he charged them that they should tell no man: but the more he charged them, so much the more a great deal they published it; And were beyond measure astonished, saying, He hath done all things well: he maketh both the deaf to hear, and the dumb to speak.

\offertory{Moses besought the Lord his God, and said: Why, O Lord, doth thy wrath wax hot against thy people? Turn from thy fierce wrath: remember Abraham, Isaac, and Jacob, to whom thou swarest to give a land flowing with milk and honey. And the Lord repented of the evil which he thought to do unto his people.}

\secret
\lett{L}{ook} favourably, O Lord, we beseech thee, on this our service: that the gift which we offer may be acceptable to thee, and avail for the succour of our frailty. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{The earth, O Lord, is filled with the fruit of thy works: that thou mayest bring food out of the earth, and wine that maketh glad the heart of man: and oil to make him a cheerful countenance, and bread to strengthen man's heart.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received thy sacrament may in such wise feel the succour of the same: that being preserved both in body and soul, we may glory in the fulness of thy heavenly healing. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Thirteenth Sunday after Trinity}
\feastday{Trinity XIII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{L}{ook,} O Lord, upon thy covenant, and forsake not the congregation of the poor for ever: arise, O Lord, maintain thine own cause, and be not unmindful of the voices of them that seek thee. \textit{Ps.} O God, wherefore art thou absent from us so long: why is thy wrath so hot against the sheep of thy pasture?

\collect
\lett{A}{lmighty} and merciful God, of whose only gift it cometh that thy faithful people do unto thee true and laudable service; Grant, we beseech thee, that we may so faithfully serve thee in this life, that we fail not finally to attain thy heavenly promises. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Galatians 3:16}
\lett{B}{rethren:} To Abraham and his seed were the promises made. He saith not, And to seeds, as of many; but as of one, And to thy seed, which is Christ. And this I say, that the covenant, that was confirmed before of God in Christ, the law, which was four hundred and thirty years after, cannot disannul, that it should make the promise of none effect. For if the inheritance be of the law, it is no more of promise: but God gave it to Abraham by promise. Wherefore then serveth the law? It was added because of transgressions, till the seed should come to whom the promise was made; and it was ordained by angels in the hand of a mediator. Now a mediator is not a mediator of one, but God is one. Is the law then against the promises of God? God forbid: for if there had been a law given which could have given life, verily righteousness should have been by the law. But the scripture hath concluded all under sin, that the promise by faith of Jesus Christ might be given to them that believe.

\gradall{Look upon thy covenant, O Lord, and forget not the congregation of the poor for ever. ℣. Arise, O Lord, maintain thine own cause: remember the rebuke that thy servants have.}{Alleluia, alleluia. ℣. Lord, thou hast been our refuge: from one generation to another. Alleluia.}

\readingcitation{Gospel}{Luke 10:23}
\lett{A}{t that time:} Jesus said unto his disciples: Blessed are the eyes which see the things that ye see: For I tell you, that many prophets and kings have desired to see those things which ye see, and have not seen them; and to hear those things which ye hear, and have not heard them. And, behold, a certain lawyer stood up, and tempted him, saying, Master, what shall I do to inherit eternal life? He said unto him, What is written in the law? how readest thou? And he answering said, Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy strength, and with all thy mind; and thy neighbour as thyself. And he said unto him, Thou hast answered right: this do, and thou shalt live. But he, willing to justify himself, said unto Jesus, And who is my neighbour? And Jesus answering said, A certain man went down from Jerusalem to Jericho, and fell among thieves, which stripped him of his raiment, and wounded him, and departed, leaving him half dead. And by chance there came down a certain priest that way: and when he saw him, he passed by on the other side. And likewise a Levite, when he was at the place, came and looked on him, and passed by on the other side. But a certain Samaritan, as he journeyed, came where he was: and when he saw him, he had compassion on him, And went to him, and bound up his wounds, pouring in oil and wine, and set him on his own beast, and brought him to an inn, and took care of him. And on the morrow when he departed, he took out two pence, and gave them to the host, and said unto him, Take care of him; and whatsoever thou spendest more, when I come again, I will repay thee. Which now of these three, thinkest thou, was neighbour unto him that fell among the thieves? And he said, He that shewed mercy on him. Then said Jesus unto him, Go, and do thou likewise.

\offertory{My hope hath been in thee, O Lord; I have said: Thou art my God, my time is in thy hand.}

\secret
\lett{M}{ercifully} look, we beseech thee, O Lord, upon the sacrifices which we present on thy sacred altars: that while they do bring us pardon, they may minister unto the honour of thy name. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Thou hast given us bread from heaven, O Lord, having every delight, and every taste of sweetness.}

\postcommunion
\lett{Q}{uicken} us, O Lord, we pray thee, who have received these holy mysteries: that they may obtain for us both pardon and protection. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Fourteenth Sunday after Trinity}
\feastday{Trinity XIV}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{ehold,} O God, our defender, and look upon the face of thine Anointed: for one day in thy courts is better than a thousand. \textit{Ps.} O how amiable are thy dwellings, thou Lord of hosts! My soul hath a desire and longing to enter into the courts of the Lord.

\collect
\lett{A}{lmighty} and everlasting God, give unto us the increase of faith, hope, and charity; and, that we may obtain that which thou dost promise, make us to love that which thou dost command. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Galatians 5:16}
\lett{B}{rethren:} Walk in the Spirit, and ye shall not fulfil the lust of the flesh. For the flesh lusteth against the Spirit, and the Spirit against the flesh: and these are contrary the one to the other: so that ye cannot do the things that ye would. But if ye be led of the Spirit, ye are not under the law. Now the works of the flesh are manifest, which are these; Adultery, fornication, uncleanness, lasciviousness, Idolatry, witchcraft, hatred, variance, emulations, wrath, strife, seditions, heresies, Envyings, murders, drunkenness, revellings, and such like: of the which I tell you before, as I have also told you in time past, that they which do such things shall not inherit the kingdom of God. But the fruit of the Spirit is love, joy, peace, longsuffering, gentleness, goodness, faith, Meekness, temperance: against such there is no law. And they that are Christ's have crucified the flesh with the affections and lusts.

\gradall{It is better to trust in the Lord, than to put any confidence in man. ℣. It is better to trust in the Lord, than to put any confidence in princes.}{Alleluia, alleluia. ℣. O come let us sing unto the Lord: let us heartily rejoice in the strength of our salvation. Alleluia.}

\readingcitation{Gospel}{Luke 17:11}
\lett{A}{t that time:} As Jesus went to Jerusalem, he passed through the midst of Samaria and Galilee. And as he entered into a certain village, there met him ten men that were lepers, which stood afar off: And they lifted up their voices, and said, Jesus, Master, have mercy on us. And when he saw them, he said unto them, Go shew yourselves unto the priests. And it came to pass, that, as they went, they were cleansed. And one of them, when he saw that he was healed, turned back, and with a loud voice glorified God, And fell down on his face at his feet, giving him thanks: and he was a Samaritan. And Jesus answering said, Were there not ten cleansed? but where are the nine? There are not found that returned to give glory to God, save this stranger. And he said unto him, Arise, go thy way: thy faith hath made thee whole.

\offertory{The Angel of the Lord tarrieth round about them that fear him, and delivereth them: O taste, and see, how gracious the Lord is.}

\secret
\lett{B}{e} favourable, O Lord, to thy people, be favourable unto their offerings: that, being well-pleased with this oblation, thou mayest both bestow upon us pardon and grant us our petitions. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Seek ye first the kingdom of God, and all things shall be added unto you, saith the Lord.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received thy heavenly sacraments, may thereby grow in grace to the attainment of eternal redemption. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage

\subby{Fifteenth Sunday after Trinity}
\feastday{Trinity XV}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{B}{ow} down, O Lord, thine ear to me, and hear me: O my God, save thy servant, that trusteth in thee: have mercy upon me, O Lord, for I have called daily upon thee. \textit{Ps.} Comfort the soul of thy servant: for unto thee, O Lord, do I lift up my soul.

\collect
\lett{K}{eep,} we beseech thee, O Lord, thy Church with thy perpetual mercy; and, because the frailty of man without thee cannot but fall, keep us ever by thy help from all things hurtful, and lead us to all things profitable to our salvation. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Galatians 6:11}
\lett{B}{rethren:} Ye see how large a letter I have written unto you with mine own hand. As many as desire to make a fair shew in the flesh, they constrain you to be circumcised; only lest they should suffer persecution for the cross of Christ. For neither they themselves who are circumcised keep the law; but desire to have you circumcised, that they may glory in your flesh. But God forbid that I should glory, save in the cross of our Lord Jesus Christ, by whom the world is crucified unto me, and I unto the world. For in Christ Jesus neither circumcision availeth any thing, nor uncircumcision, but a new creature. And as many as walk according to this rule, peace be on them, and mercy, and upon the Israel of God. From henceforth let no man trouble me: for I bear in my body the marks of the Lord Jesus. Brethren, the grace of our Lord Jesus Christ be with your spirit. Amen.

\gradall{It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest. ℣. To tell of thy loving-kindness early in the morning, and of thy truth in the night-season.}{Alleluia, alleluia. ℣. For the Lord is a great God, and the great King upon all the earth. Alleluia.}

\readingcitation{Gospel}{Matthew 6:24}
\lett{A}{t that time:} Jesus said unto his disciples: No man can serve two masters: for either he will hate the one, and love the other; or else he will hold to the one, and despise the other. Ye cannot serve God and mammon. Therefore I say unto you, Take no thought for your life, what ye shall eat, or what ye shall drink; nor yet for your body, what ye shall put on. Is not the life more than meat, and the body than raiment? Behold the fowls of the air: for they sow not, neither do they reap, nor gather into barns; yet your heavenly Father feedeth them. Are ye not much better than they? Which of you by taking thought can add one cubit unto his stature? And why take ye thought for raiment? Consider the lilies of the field, how they grow; they toil not, neither do they spin: And yet I say unto you, That even Solomon in all his glory was not arrayed like one of these. Wherefore, if God so clothe the grass of the field, which to day is, and to morrow is cast into the oven, shall he not much more clothe you, O ye of little faith? Therefore take no thought, saying, What shall we eat? or, What shall we drink? or, Wherewithal shall we be clothed? (For after all these things do the Gentiles seek:) for your heavenly Father knoweth that ye have need of all these things. But seek ye first the kingdom of God, and his righteousness; and all these things shall be added unto you. Take therefore no thought for the morrow: for the morrow shall take thought for the things of itself. Sufficient unto the day is the evil thereof.

\offertory{I waited patiently for the Lord, and he inclined unto me: he heard my calling: and hath put a new song in my mouth, even a thanksgiving unto our God.}

%MANUAL ADJUSTMENT:
\vspace{-2ex}

\secret
\lett{G}{rant} to us, O Lord, we beseech thee, that this saving victim may avail, both for the cleansing of our offences, and for the propitiation of thy power. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{The bread that I will give is my flesh, which I will give for the life of the world.}

%MANUAL ADJUSTMENT:
\vspace{-2ex}

%MANUAL ADJUSTMENT:
%\postcommunion
\subsubsection{Postcommunion}
\lett{M}{ay} thy sacraments, O God, evermore cleanse and defend us: and bring us to the attainment of eternal salvation. Through.

%MANUAL ADJUSTMENT:
\vspace{0.25ex}

\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Sixteenth Sunday after Trinity}
\feastday{Trinity XVI}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{H}{ave} mercy upon me, O Lord, for I have called daily upon thee: for thou, O Lord, art gracious and merciful, and plenteous in thy loving kindness toward all them that call upon thee. \textit{Ps.} Bow down thine ear to me, O Lord, and hear me: for I am poor and in misery.
\collect
\lett{O}{Lord,} we beseech thee, let thy continual pity cleanse and defend thy Church; and, because it cannot continue in safety without thy succour, preserve it evermore by thy help and goodness. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Ephesians 3:13}
\lett{B}{rethren:} I desire that ye faint not at my tribulations for you, which is your glory. For this cause I bow my knees unto the Father of our Lord Jesus Christ, Of whom the whole family in heaven and earth is named, That he would grant you, according to the riches of his glory, to be strengthened with might by his Spirit in the inner man; That Christ may dwell in your hearts by faith; that ye, being rooted and grounded in love, May be able to comprehend with all saints what is the breadth, and length, and depth, and height; And to know the love of Christ, which passeth knowledge, that ye might be filled with all the fulness of God. Now unto him that is able to do exceeding abundantly above all that we ask or think, according to the power that worketh in us, Unto him be glory in the church by Christ Jesus throughout all ages, world without end. Amen.

\gradall{The heathen shall fear thy name, O Lord, and all the kings of the earth thy majesty. ℣. When the Lord shall build up Sion, and when his glory shall appear.}{Alleluia, alleluia. ℣. O sing unto the Lord a new song: for the Lord hath done marvellous things. Alleluia.}

\readingcitation{Gospel}{Luke 7:11}
\lett{A}{t that time:} Jesus went into a city called Nain; and many of his disciples went with him, and much people. Now when he came nigh to the gate of the city, behold, there was a dead man carried out, the only son of his mother, and she was a widow: and much people of the city was with her. And when the Lord saw her, he had compassion on her, and said unto her, Weep not. And he came and touched the bier: and they that bare him stood still. And he said, Young man, I say unto thee, Arise. And he that was dead sat up, and began to speak. And he delivered him to his mother. And there came a fear on all: and they glorified God, saying, That a great prophet is risen up among us; and, That God hath visited his people. And this rumour of him went forth throughout all Judaea, and throughout all the region round about.

\offertory{Look down, O Lord, to help me: let them be ashamed, and confounded that seek after my soul to destroy it: look down, O Lord, to help me.}

\secret
\lett{M}{ay} thy sacraments guard us, O Lord: and ever protect us against the assaults of the devil. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{O Lord, I will make mention of thy righteousness only: thou, O God, hast taught me from my youth up until now: forsake me not, O God, in mine old age, when I am gray-headed.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that the operation of thy heavenly gift may so possess our souls and bodies: that not our own desires, but the effectual working of the same may ever prevail within us. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage

\subby{Seventeenth Sunday after Trinity}
\feastday{Trinity XVII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{R}{ighteous} art thou, O Lord, and true is thy judgement: O deal with thy servant according unto thy merciful kindness. \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

\collect
\lett{L}{ord,} we pray thee that thy grace may always prevent and follow us, and make us continually to be given to all good works. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Ephesians 4:1}
\lett{B}{rethren:} I, the prisoner of the Lord, beseech you that ye walk worthy of the vocation wherewith ye are called, With all lowliness and meekness, with longsuffering, forbearing one another in love; Endeavouring to keep the unity of the Spirit in the bond of peace. There is one body, and one Spirit, even as ye are called in one hope of your calling; One Lord, one faith, one baptism, One God and Father of all, who is above all, and through all, and in you all. Who is blessed for ever and ever. Amen.

\gradall{Blessed are the people whose God is the Lord: and blessed are the folk that he hath chosen to him to be his inheritance. ℣. By the word of the Lord were the heavens made: and all the hosts of them by the breath of his mouth.}{Alleluia, alleluia. ℣. Hear my prayer, O Lord, and let my cry come unto thee. Alleluia.}

\readingcitation{Gospel}{Luke 14:1}
\lett{A}{t that time:} When Jesus went into the house of one of the chief Pharisees to eat bread on the sabbath day, that they watched him. And, behold, there was a certain man before him which had the dropsy. And Jesus answering spake unto the lawyers and Pharisees, saying, Is it lawful to heal on the sabbath day? And they held their peace. And he took him, and healed him, and let him go; And answered them, saying, Which of you shall have an ass or an ox fallen into a pit, and will not straightway pull him out on the sabbath day? And they could not answer him again to these things. And he put forth a parable to those which were bidden, when he marked how they chose out the chief rooms; saying unto them, When thou art bidden of any man to a wedding, sit not down in the highest room; lest a more honourable man than thou be bidden of him; And he that bade thee and him come and say to thee, Give this man place; and thou begin with shame to take the lowest room. But when thou art bidden, go and sit down in the lowest room; that when he that bade thee cometh, he may say unto thee, Friend, go up higher: then shalt thou have worship in the presence of them that sit at meat with thee. For whosoever exalteth himself shall be abased; and he that humbleth himself shall be exalted.

\offertory{I, Daniel, prayed unto my God, and said: Hear, O Lord, the prayers of thy servant: cause thy face to shine upon thy sanctuary: and behold, O God, this thy people, who are called by thy name.}

\secret
\lett{C}{leanse} us, O Lord, we pray thee, by the effectual working of this our sacrifice: and so accomplish in us the work of thy mercy; that we may be found worthy to be made partakers of the same. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Promise unto the Lord your God, and keep it, all ye that are round about him, bring presents unto him that ought to be feared: he shall refrain the spirit of princes: and is wonderful among all the kings of the earth.}
\postcommunion
\lett{O}{Lord,} we beseech thee, mercifully to purify our minds and renew them with thy heavenly sacraments: that we may thereby have help also for our bodies both now and for evermore. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subby{Eighteenth Sunday after Trinity}
\feastday{Trinity XVIII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{G}{ive} peace, O Lord, to them that wait for thee, and let thy prophets be found faithful: hear the prayers of thy servant, and of thy people Israel. \textit{Ps.} I was glad when they said unto me: We will go into the house of the Lord.
\collect
\lett{L}{ord,} we beseech thee, grant thy people grace to withstand the temptations of the world, the flesh, and the devil; and with pure hearts and minds to follow thee, the only God. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 1:4}
\lett{B}{rethren:} I thank my God always on your behalf, for the grace of God which is given you by Jesus Christ; That in every thing ye are enriched by him, in all utterance, and in all knowledge; Even as the testimony of Christ was confirmed in you: So that ye come behind in no gift; waiting for the coming of our Lord Jesus Christ: Who shall also confirm you unto the end, that ye may be blameless in the day of our Lord Jesus Christ.

\gradall{I was glad when they said unto me: We will go into the house of the Lord. ℣. Peace be within thy walls: and plenteousness within thy palaces.}{Alleluia, alleluia. ℣. The heathen shall fear thy name, O Lord, and all the kings of the earth thy majesty. Alleluia.}

\readingcitation{Gospel}{Matthew 22:34}
\lett{A}{t that time:} When the Pharisees had heard that Jesus had put the Sadduces to silence, they were gathered together. Then one of them, which was a lawyer, asked him a question, tempting him, and saying, Master, which is the great commandment in the law? Jesus said unto him, Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind. This is the first and great commandment. And the second is like unto it, Thou shalt love thy neighbour as thyself. On these two commandments hang all the law and the prophets. While the Pharisees were gathered together, Jesus asked them, Saying, What think ye of Christ? whose son is he? They say unto him, The Son of David. He saith unto them, How then doth David in spirit call him Lord, saying, The \divineName{Lord} said unto my Lord, Sit thou on my right hand, till I make thine enemies thy footstool? If David then call him Lord, how is he his son? And no man was able to answer him a word, neither durst any man from that day forth ask him any more questions.

\offertory{Moses consecrated an altar unto the Lord, offering burnt offerings upon it, and sacrificing peace offerings: and he made an evening sacrifice for a sweet-smelling savour unto the Lord God, in the sight of the children of Israel.}

\secret
\lett{W}{e} humbly entreat thy majesty, O Lord: that these holy mysteries which we perform may deliver us both from past and future transgressions. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Bring offerings and come into his courts: O worship the Lord in the beauty of holiness.}

\postcommunion
\lett{A}{lmighty} God, let thy holy mysteries both heal our vices, and bestow upon us everlasting remedies. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Nineteenth Sunday after Trinity}
\feastday{Trinity XIX}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{I}{am} the saving health of the people, saith the Lord: out of whatsoever tribulation they shall cry to me, I will hear them: and I will be their Lord for ever. \textit{Ps.} Hear my law, O my people: incline your ears unto the words of my mouth.

\collect
\lett{O}{God,} forasmuch as without thee we are not able to please thee; Mercifully grant that thy Holy Spirit may in all things direct and rule our hearts. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Ephesians 4:17}
\lett{B}{rethren:} This I say therefore, and testify in the Lord, that ye henceforth walk not as other Gentiles walk, in the vanity of their mind, Having the understanding darkened, being alienated from the life of God through the ignorance that is in them, because of the blindness of their heart: Who being past feeling have given themselves over unto lasciviousness, to work all uncleanness with greediness. But ye have not so learned Christ; If so be that ye have heard him, and have been taught by him, as the truth is in Jesus: That ye put off concerning the former conversation the old man, which is corrupt according to the deceitful lusts; Be renewed in the spirit of your mind; And that ye put on the new man, which after God is created in righteousness and true holiness. Wherefore putting away lying, speak every man truth with his neighbour: for we are members one of another. Be ye angry, and sin not: let not the sun go down upon your wrath: Neither give place to the devil. Let him that stole steal no more: but rather let him labour, working with his hands the thing which is good, that he may have to give to him that needeth. Let no corrupt communication proceed out of your mouth, but that which is good to the use of edifying, that it may minister grace unto the hearers. And grieve not the holy Spirit of God, whereby ye are sealed unto the day of redemption. Let all bitterness, and wrath, and anger, and clamour, and evil speaking, be put away from you, with all malice: And be ye kind one to another, tenderhearted, forgiving one another, even as God for Christ's sake hath forgiven you.

\gradall{Let my prayer be set forth in thy sight, O Lord, as the incense. ℣. Let the lifting up of my hands be an evening sacrifice.}{Alleluia, alleluia. ℣. O give thanks unto the Lord, and call upon his name: tell the people what things he hath done. Alleluia.}

\readingcitation{Gospel}{Matthew 9:1}
\lett{A}{t that time:} Jesus entered into a ship, and passed over, and came into his own city. And, behold, they brought to him a man sick of the palsy, lying on a bed: and Jesus seeing their faith said unto the sick of the palsy; Son, be of good cheer; thy sins be forgiven thee. And, behold, certain of the scribes said within themselves, This man blasphemeth. And Jesus knowing their thoughts said, Wherefore think ye evil in your hearts? For whether is easier, to say, Thy sins be forgiven thee; or to say, Arise, and walk? But that ye may know that the Son of man hath power on earth to forgive sins, (then saith he to the sick of the palsy,) Arise, take up thy bed, and go unto thine house. And he arose, and departed to his house. But when the multitudes saw it, they marvelled, and glorified God, which had given such power unto men.

\offertory{Though I walk in the midst of trouble, yet shalt thou refresh me, O Lord: thou shalt stretch forth thy hand upon the furiousness of mine enemies, and thy right hand shall save me.}

\secret
\lett{O}{God,} who, by communion in this venerable sacrifice, hast made us partakers of the one supreme Godhead: grant, we beseech thee; that as we have the knowledge of thy truth, so by worthy conversation we may attain unto the same. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Thou hast charged that we shall diligently keep thy commandments: O that my ways were made so direct, that I might keep thy statutes.}

\postcommunion
\lett{W}{e} render thanks unto thee, O Lord, for that thou hast quickened us with thy sacred gift: humbly beseeching thy mercy; that thou wouldest make us worthy partakers of the same. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Twentieth Sunday after Trinity}
\feastday{Trinity XX}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{E}{verything} that thou hast done to us, O Lord, thou hast done in true judgment: for we have sinned and have not obeyed thy commandments: but give glory to thy name, and deal with us according to the multitude of thy mercies. \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

\collect
\lett{O}{almighty} and most merciful God, of thy bountiful goodness keep us, we beseech thee, from all things that may hurt us; that we, being ready both in body and soul, may cheerfully accomplish those things which thou commandest. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Ephesians 5:15}
\lett{B}{rethren:} See that ye walk circumspectly, not as fools, but as wise, Redeeming the time, because the days are evil. Wherefore be ye not unwise, but understanding what the will of the Lord is. And be not drunk with wine, wherein is excess; but be filled with the Spirit; Speaking to yourselves in psalms and hymns and spiritual songs, singing and making melody in your heart to the Lord; Giving thanks always for all things unto God and the Father in the name of our Lord Jesus Christ; Submitting yourselves one to another in the fear of God.

\gradall{The eyes of all wait upon thee, O Lord: and thou givest them their meat in due season. ℣. Thou openest thine hand: and fillest all things living with plenteousness.}{Alleluia, alleluia. ℣. O God, my heart is ready, my heart is ready: I will sing and give praise unto thee, my glory. Alleluia.}

\readingcitation{Gospel}{Matthew 22:1}
\lett{A}{t that time:} Jesus spake unto the chief priests and Pharisees in parables, saying: The kingdom of heaven is like unto a certain king, which made a marriage for his son, And sent forth his servants to call them that were bidden to the wedding: and they would not come. Again, he sent forth other servants, saying, Tell them which are bidden, Behold, I have prepared my dinner: my oxen and my fatlings are killed, and all things are ready: come unto the marriage. But they made light of it, and went their ways, one to his farm, another to his merchandise: And the remnant took his servants, and entreated them spitefully, and slew them. But when the king heard thereof, he was wroth: and he sent forth his armies, and destroyed those murderers, and burned up their city. Then saith he to his servants, The wedding is ready, but they which were bidden were not worthy. Go ye therefore into the highways, and as many as ye shall find, bid to the marriage. So those servants went out into the highways, and gathered together all as many as they found, both bad and good: and the wedding was furnished with guests. And when the king came in to see the guests, he saw there a man which had not on a wedding garment: And he saith unto him, Friend, how camest thou in hither not having a wedding garment? And he was speechless. Then said the king to the servants, Bind him hand and foot, and take him away, and cast him into outer darkness; there shall be weeping and gnashing of teeth. For many are called, but few are chosen.

\offertory{By the waters of Babylon we sat down and wept: when we remembered thee, O Sion.}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\secret
\lett{G}{rant,} we beseech thee, O Lord, that these gifts which we offer in the sight of thy majesty, may be for our salvation. Through.

%MANUAL ADJUSTMENT:
\vspace{0.25ex}

\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\communion{Remember thy word unto thy servant, O Lord, wherein thou hast caused me to put my trust: the same is my comfort in my affliction.}
\postcommunion
\lett{M}{ay} the operation of thy healing, O Lord, both mercifully cleanse us from our perversities, and make us ever to cleave unto thy commandments. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Twenty-First Sunday after Trinity}\label{trinity}
\feastday{Trinity XXI}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{O}{Lord,} everything is in subjection to thy will, and there is no man that is able to resist thy power: for thou hast created everything, heaven and earth, and all the wonders which under heaven's vault are contained: thou art the Lord of all things. \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.
\collect
\lett{G}{rant,} we beseech thee, merciful Lord, to thy faithful people pardon and peace, that they may be cleansed from all their sins, and serve thee with a quiet mind. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Ephesians 6:10}
\lett{B}{rethren:} Be strong in the Lord, and in the power of his might. Put on the whole armour of God, that ye may be able to stand against the wiles of the devil. For we wrestle not against flesh and blood, but against principalities, against powers, against the rulers of the darkness of this world, against spiritual wickedness in high places. Wherefore take unto you the whole armour of God, that ye may be able to withstand in the evil day, and having done all, to stand. Stand therefore, having your loins girt about with truth, and having on the breastplate of righteousness; And your feet shod with the preparation of the gospel of peace; Above all, taking the shield of faith, wherewith ye shall be able to quench all the fiery darts of the wicked. And take the helmet of salvation, and the sword of the Spirit, which is the word of God: Praying always with all prayer and supplication in the Spirit, and watching thereunto with all perseverance and supplication for all saints; And for me, that utterance may be given unto me, that I may open my mouth boldly, to make known the mystery of the gospel, For which I am an ambassador in bonds: that therein I may speak boldly, as I ought to speak.

\gradall{Lord, thou hast been our refuge, from one generation to another. ℣. Before the mountains were brought forth, or ever the earth and the world were made: thou art God from everlasting, and world without end.}{Alleluia, alleluia. ℣. When Israel came out of Egypt, and the house of Jacob from among the strange people. Alleluia.}

\readingcitation{Gospel}{John 4:46}
\lett{A}{t that time:} There was a certain nobleman, whose son was sick at Capernaum. When he heard that Jesus was come out of Judaea into Galilee, he went unto him, and besought him that he would come down, and heal his son: for he was at the point of death. Then said Jesus unto him, Except ye see signs and wonders, ye will not believe. The nobleman saith unto him, Sir, come down ere my child die. Jesus saith unto him, Go thy way; thy son liveth. And the man believed the word that Jesus had spoken unto him, and he went his way. And as he was now going down, his servants met him, and told him, saying, Thy son liveth. Then enquired he of them the hour when he began to amend. And they said unto him, Yesterday at the seventh hour the fever left him. So the father knew that it was at the same hour, in the which Jesus said unto him, Thy son liveth: and himself believed, and his whole house. This is again the second miracle that Jesus did, when he was come out of Jud{\ae}a into Galilee.

\offertory{There was a man in the land of Uz, whose name was Job: perfect and upright, and one that feared God: and Satan sought to tempt him: and power was given him by the Lord over his possessions and over his flesh: and he destroyed all his substance and his sons: and he smote his flesh with sore boils.}

%MANUAL ADJUSTMENT:
\vspace{-2ex}

\secret
\lett{W}{e} beseech thee, O Lord, that these mysteries may bestow upon us thy heavenly healing: and purge away the vices of our hearts. Through.

%MANUAL ADJUSTMENT:
\vspace{0.25ex}

\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.25ex}

\communion{My soul hath longed for thy salvation, and I have a good hope in thy word: when wilt thou be avenged of them that persecute me? They persecuted me falsely, O be thou my help, O Lord my God.}

%MANUAL ADJUSTMENT:
\vspace{-0.25ex}

%MANUAL ADJUSTMENT:
%\postcommunion
\subsubsection{Postcommunion}
\lett{W}{e} beseech thee, O Lord, make us ever to obey thy commandments: that we may be rendered worthy of thy sacred gifts. Through.

%MANUAL ADJUSTMENT:
\vspace{0.25ex}

\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\subby{Twenty-Second Sunday after Trinity}
\feastday{Trinity XXII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{I}{f} thou, O Lord, wilt be extreme to mark iniquities: O Lord, who may abide it? for there is mercy with thee, O God of Israel. \textit{Ps.} Out of the deep have I called unto thee, O Lord: Lord, hear my voice.
\collect
\lett{L}{ord,} we beseech thee to keep thy household the Church in continual godliness; that through thy protection it may be free from all adversities, and devoutly given to serve thee in good works, to the glory of thy Name. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Philippians 1:3}
\lett{B}{rethren:} I thank my God upon every remembrance of you, Always in every prayer of mine for you all making request with joy, For your fellowship in the gospel from the first day until now; Being confident of this very thing, that he which hath begun a good work in you will perform it until the day of Jesus Christ: Even as it is meet for me to think this of you all, because I have you in my heart; inasmuch as both in my bonds, and in the defence and confirmation of the gospel, ye all are partakers of my grace. For God is my record, how greatly I long after you all in the bowels of Jesus Christ. And this I pray, that your love may abound yet more and more in knowledge and in all judgment; That ye may approve things that are excellent; that ye may be sincere and without offence till the day of Christ; Being filled with the fruits of righteousness, which are by Jesus Christ, unto the glory and praise of God.

\gradall{Behold, how good and joyful a thing it is, brethren, to dwell together in unity. ℣. It is like the precious ointment upon the head, that ran down unto the beard, even unto Aaron's beard.}{Alleluia, alleluia. ℣. Ye that fear the Lord, put your trust in him: he is their helper and defender. Alleluia.}

\readingcitation{Gospel}{Matthew 18:21}
\lett{A}{t that time:} Peter came to Jesus, and said, Lord, how oft shall my brother sin against me, and I forgive him? till seven times? Jesus saith unto him, I say not unto thee, Until seven times: but, Until seventy times seven. Therefore is the kingdom of heaven likened unto a certain king, which would take account of his servants. And when he had begun to reckon, one was brought unto him, which owed him ten thousand talents. But forasmuch as he had not to pay, his lord commanded him to be sold, and his wife, and children, and all that he had, and payment to be made. The servant therefore fell down, and worshipped him, saying, Lord, have patience with me, and I will pay thee all. Then the lord of that servant was moved with compassion, and loosed him, and forgave him the debt. But the same servant went out, and found one of his fellowservants, which owed him an hundred pence: and he laid hands on him, and took him by the throat, saying, Pay me that thou owest. And his fellowservant fell down at his feet, and besought him, saying, Have patience with me, and I will pay thee all. And he would not: but went and cast him into prison, till he should pay the debt. So when his fellowservants saw what was done, they were very sorry, and came and told unto their lord all that was done. Then his lord, after that he had called him, said unto him, O thou wicked servant, I forgave thee all that debt, because thou desiredst me: Shouldest not thou also have had compassion on thy fellowservant, even as I had pity on thee? And his lord was wroth, and delivered him to the tormentors, till he should pay all that was due unto him. So likewise shall my heavenly Father do also unto you, if ye from your hearts forgive not every one his brother their trespasses.

\offertory{Remember me, O Lord, King of all power: and put a well-ordered speech in my mouth, that my words may be pleasing in the sight of the prince.}

\secret
\lett{M}{ercifully} receive, O Lord, these offerings: wherewith thou hast willed that atonement should be made unto thee, and that salvation should be restored to us by the power of thy loving kindness. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{I have called upon thee, O God, for thou hast heard me: incline thine ear, and hearken unto my words.}

\postcommunion
\lett{O}{Lord,} who hast bestowed upon us the food of immortality, we beseech thee: that those things which we have taken with our lips we may seek after in purity of heart. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Twenty-Third Sunday after Trinity}
\feastday{Trinity XXIII}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{hus} saith the Lord: I think thoughts of peace, and not of affliction: ye shall call upon me, and I will hearken unto you, and I will bring again your captivity from all places. \textit{Ps.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob.

\collect
\lett{O}{God,} our refuge and strength, who art the author of all godliness; Be ready, we beseech thee, to hear the devout prayers of thy Church; and grant that those things which we ask faithfully we may obtain effectually. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Philippians 3:17}
\lett{B}{rethren:} Be followers together of me, and mark them which walk so as ye have us for an ensample. (For many walk, of whom I have told you often, and now tell you even weeping, that they are the enemies of the cross of Christ: Whose end is destruction, whose God is their belly, and whose glory is in their shame, who mind earthly things.) For our conversation is in heaven; from whence also we look for the Saviour, the Lord Jesus Christ: Who shall change our vile body, that it may be fashioned like unto his glorious body, according to the working whereby he is able even to subdue all things unto himself.

%MANUAL ADJUSTMENT:
\gradual{It is thou, O Lord, that savest us from our enemies: and puttest them to confusion that hate us. ℣. We make our boast of God all day long, and will praise thy name for ever.}

\alleluia{Alleluia, alleluia. ℣. Out of the deep have I called unto thee, O Lord: Lord, hear my prayer. Alleluia.}

\readingcitation{Gospel}{Matthew 22:15}
\lett{A}{t that time:} The Pharisees went, and took counsel how they might entangle Jesus in his talk. And they sent out unto him their disciples with the Herodians, saying, Master, we know that thou art true, and teachest the way of God in truth, neither carest thou for any man: for thou regardest not the person of men. Tell us therefore, What thinkest thou? Is it lawful to give tribute unto Caesar, or not? But Jesus perceived their wickedness, and said, Why tempt ye me, ye hypocrites? Shew me the tribute money. And they brought unto him a penny. And he saith unto them, Whose is this image and superscription? They say unto him, Caesar's. Then saith he unto them, Render therefore unto Caesar the things which are Caesar's; and unto God the things that are God's. When they had heard these words, they marvelled, and left him, and went their way.

\offertory{Out of the deep have I called unto thee, O Lord: Lord, hear my prayer: out of the deep have I called unto thee, O Lord.}

\secret
\lett{G}{rant,} O merciful God: that this saving oblation may ever free us from our sins, and shield us from all adversities. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Verily I say unto you, what things soever ye desire, when ye pray, believe that ye receive them, and it shall be done unto you.}

\postcommunion
\lett{W}{e} have received, O Lord, the gifts of thy sacred mystery, humbly beseeching thee: that those things, which thou hast commanded us to do in remembrance of thee, may be profitable to the succour of our infirmity. Who livest.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Twenty-Fourth Sunday after Trinity}
\feastday{Trinity XXIV}
\begin{inhead}
{Semidouble}
\end{inhead}
\begin{rubric}
If this Sunday be hindered by the Sunday Next before Advent, it is anticipated on Saturday with all the privileges proper to a Sunday occurring, and in it are said \emph{Gloria in excelsis}, the Nicene Creed, Preface of the Trinity, and \emph{Ite, Missa est}.
\end{rubric}

\properantiphonfix

\introit
\lett{T}{hus} saith the Lord: I think thoughts of peace, and not of affliction: ye shall call upon me, and I will hearken unto you: and will bring again your captivity from all places. \textit{Ps.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob.
\collect
\lett{O}{Lord,} we beseech thee, absolve thy people from their offences; that through thy bountiful goodness we may all be delivered from the bands of those sins, which by our frailty we have committed. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Colossians 1:3}
\lett{B}{rethren:} We give thanks to God and the Father of our Lord Jesus Christ, praying always for you, Since we heard of your faith in Christ Jesus, and of the love which ye have to all the saints, For the hope which is laid up for you in heaven, whereof ye heard before in the word of the truth of the gospel; Which is come unto you, as it is in all the world; and bringeth forth fruit, as it doth also in you, since the day ye heard of it, and knew the grace of God in truth: As ye also learned of Epaphras our dear fellowservant, who is for you a faithful minister of Christ; Who also declared unto us your love in the Spirit. For this cause we also, since the day we heard it, do not cease to pray for you, and to desire that ye might be filled with the knowledge of his will in all wisdom and spiritual understanding; That ye might walk worthy of the Lord unto all pleasing, being fruitful in every good work, and increasing in the knowledge of God; Strengthened with all might, according to his glorious power, unto all patience and longsuffering with joyfulness; Giving thanks unto the Father, which hath made us meet to be partakers of the inheritance of the saints in light.

\gradall{It is thou, O Lord, that savest us from our enemies: and puttest them to confusion that hate us. ℣. We make our boast of God all day long, and will praise thy name for ever.}{Alleluia, alleluia. ℣. Out of the deep have I called unto thee, O Lord: Lord, hear my prayer. Alleluia.}

\readingcitation{Gospel}{Matthew 9:18}
\lett{A}{t that time:} While Jesus spake these things unto John's disciples: Behold, there came a certain ruler, and worshipped him, saying, My daughter is even now dead: but come and lay thy hand upon her, and she shall live. And Jesus arose, and followed him, and so did his disciples. And, behold, a woman, which was diseased with an issue of blood twelve years, came behind him, and touched the hem of his garment: For she said within herself, If I may but touch his garment, I shall be whole. But Jesus turned him about, and when he saw her, he said, Daughter, be of good comfort; thy faith hath made thee whole. And the woman was made whole from that hour. And when Jesus came into the ruler's house, and saw the minstrels and the people making a noise, He said unto them, Give place: for the maid is not dead, but sleepeth. And they laughed him to scorn. But when the people were put forth, he went in, and took her by the hand, and the maid arose. And the fame hereof went abroad into all that land.

\offertory{Out of the deep have I called unto thee, O Lord: Lord, hear my prayer: out of the deep have I called unto thee, O Lord.}

\secret
\lett{W}{e} offer thee, O Lord, this sacrifice of praise for the increase of our dutiful service: that thou wouldest mercifully accomplish that which thou hast bestowed upon us beyond our deserving. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\communion{Verily I say unto you, What things soever ye desire, when ye pray, believe that ye receive them, and it shall be done unto you.}

\postcommunion
\lett{W}{e} beseech thee, almighty God: that we, whom thou makest to rejoice in the partaking of heavenly things, may by thee be defended against all earthly perils. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\begin{rubric}
    If there be more than 25 Sundays after Trinity, then the Sundays after the \nth{24} after Trinity are supplied from the unused Sundays after Epiphany (p. \pageref{epiphany}). Then, the final Sunday is the Sunday Next before Advent.
\end{rubric}

\subby{Sunday Next before Advent}
\feastday{Sunday before Advent}
\begin{inhead}
{Semidouble}
\end{inhead}

\properantiphonfix

\introit
\lett{T}{hus} saith the Lord: I think thoughts of peace, and not of affliction: ye shall call upon me, and I will hearken unto you: and will bring again your captivity from all places. \textit{Ps.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob.

\collect
\lett{S}{tir} up, we beseech thee, O Lord, the wills of thy faithful people; that they, plenteously bringing forth the fruit of good works, may by thee be plenteously rewarded. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\readingcitation{Epistle}{Jeremiah 23:5}
\lett{I}{n those days:} Jeremiah the Prophet spake, saying: Behold, the days come, saith the \divineName{Lord}, that I will raise unto David a righteous Branch, and a King shall reign and prosper, and shall execute judgment and justice in the earth. In his days Judah shall be saved, and Israel shall dwell safely: and this is his name whereby he shall be called, \textsc{The Lord Our Righteousness}. Therefore, behold, the days come, saith the \divineName{Lord}, that they shall no more say, The \divineName{Lord} liveth, which brought up the children of Israel out of the land of Egypt; But, The \divineName{Lord} liveth, which brought up and which led the seed of the house of Israel out of the north country, and from all countries whither I had driven them; and they shall dwell in their own land.

\gradual{It is thou, O Lord, that savest us from our enemies: and puttest them to confusion that hate us. ℣. We make our boast of God all day long, and will praise thy name for ever.}

\alleluia{Alleluia, alleluia. ℣. Out of the deep have I called unto thee, O Lord: Lord, hear my prayer. Alleluia.}

\readingcitation{Gospel}{John 6:5}
\lett{A}{t that time:} When Jesus then lifted up his eyes, and saw a great company come unto him, he saith unto Philip, Whence shall we buy bread, that these may eat? And this he said to prove him: for he himself knew what he would do. Philip answered him, Two hundred pennyworth of bread is not sufficient for them, that every one of them may take a little. One of his disciples, Andrew, Simon Peter's brother, saith unto him, There is a lad here, which hath five barley loaves, and two small fishes: but what are they among so many? And Jesus said, Make the men sit down. Now there was much grass in the place. So the men sat down, in number about five thousand. And Jesus took the loaves; and when he had given thanks, he distributed to the disciples, and the disciples to them that were set down; and likewise of the fishes as much as they would. When they were filled, he said unto his disciples, Gather up the fragments that remain, that nothing be lost. Therefore they gathered them together, and filled twelve baskets with the fragments of the five barley loaves, which remained over and above unto them that had eaten. Then those men, when they had seen the miracle that Jesus did, said, This is of a truth that prophet that should come into the world.

\offertory{Out of the deep have I called unto thee, O Lord: Lord, hear my prayer: out of the deep have I called unto thee, O Lord.}

\secret
\lett{B}{e} favourable, O Lord, to our supplications: and, accepting the offerings and prayers of us thy people, turn the hearts of all to thee; that we, being delivered from earthly lusts, may turn to heavenly desires. Through.

\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}


\communion{Verily I say unto you, what things soever ye desire, when ye pray, believe that ye receive them, and it shall be done unto you.}

\subsubsec{Postcommunion}
\lett{G}{rant} to us, we beseech thee, O Lord: that through these sacraments, which we have received, whatsoever is corrupt in our minds may be made whole by the healing power of the same. Through.

\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}