\fancyhead[C]{{\LARGE Proper of Saints}}
\fancyhead[RO,LE]{}
\fancyhead[RE,LO]{}
\phantomsection
\addcontentsline{toc}{section}{Proper of Saints}
%In the Proper of Saints, to retain the simplicity of the BCP, we are only including First and Second Class Doubles, except where it is a Prayer Book Holiday.
%Since the Proper of Saints only provides the Tikhonite Tradition in the BCP, we readily look to the Sarum for the Missal propers, beyond the BCP.
%Therefore, for Feast Days not in the 1928 BCP, we provide a score of Sarum-Roman adherence. We check both the Roman and Sarum propers. If the Sarum Missal has both Major and Minor Propers---or the Roman only has Commons---then the coherence of the Major Propers (Collect, Epistle, or Gospel) are given two points and that of Minor Propers (Introit, Gradual, Alleluia, Offertory, Secret, Communion, and Post-Communion) one point. Then, any Feast Day with a score less than 0.5 gets Sarum propers. This isn't perfect, but it sufficiently avoids the Proper of Saints becoming an ugly patchwork quilt.
\subby{St. Andrew}
\feastday{St. Andrew}
\begin{inhead}
    {Second Class Double\\
30 November}
\end{inhead}
\feastdate{30}{November}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} The Lord loved Andrew in the odour of sweetness.

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine uprising.
\collect
%1928 BCP Collect:
%\lett{A}{lmighty} God, who didst give such grace unto thy holy Apostle Saint Andrew, that he readily obeyed the calling of thy Son Jesus Christ, and followed him without delay; Grant unto us all, that we, being called by thy holy Word, may forthwith give up ourselves obediently to fulfil thy holy commandments. Through the same.
\lett{W}{e} humbly entreat thy majesty, O Lord: that as blessed Andrew the Apostle was to thy Church a preacher and governor; so he may be a perpetual intercessor for us in thy sight. Through.

\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\readingcitation{Epistle}{Romans 10:9}
\lett{B}{rethren:} If thou shalt confess with thy mouth the Lord Jesus, and shalt believe in thine heart that God hath raised him from the dead, thou shalt be saved. For with the heart man believeth unto righteousness; and with the mouth confession is made unto salvation. For the scripture saith, Whosoever believeth on him shall not be ashamed. For there is no difference between the Jew and the Greek: for the same Lord over all is rich unto all that call upon him. For whosoever shall call upon the name of the Lord shall be saved. How then shall they call on him in whom they have not believed? and how shall they believe in him of whom they have not heard? and how shall they hear without a preacher? And how shall they preach, except they be sent? as it is written, How beautiful are the feet of them that preach the gospel of peace, and bring glad tidings of good things! But they have not all obeyed the gospel. For Esaias saith, Lord, who hath believed our report? So then faith cometh by hearing, and hearing by the word of God. But I say, Have they not heard? Yes verily, their sound went into all the earth, and their words unto the ends of the world. But I say, Did not Israel know? First Moses saith, I will provoke you to jealousy by them that are no people, and by a foolish nation I will anger you. But Esaias is very bold, and saith, I was found of them that sought me not; I was made manifest unto them that asked not after me. But to Israel he saith, All day long I have stretched forth my hands unto a disobedient and gainsaying people.

%MANUAL ADJUSTMENT:
\gradual{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}

\alleluia{Alleluia, alleluia. ℣. The Lord loved Andrew as a sweet savour. Alleluia.}

\readingcitation{Gospel}{Matthew 4:18}
\lett{A}{t that time:} Jesus, walking by the sea of Galilee, saw two brethren, Simon called Peter, and Andrew his brother, casting a net into the sea: for they were fishers. And he saith unto them, Follow me, and I will make you fishers of men. And they straightway left their nets, and followed him. And going on from thence, he saw other two brethren, James the son of Zebedee, and John his brother, in a ship with Zebedee their father, mending their nets; and he called them. And they immediately left the ship and their father, and followed him.
\offertory{Right honourable are thy friends unto me, O God: right well is their princedom established.}
\secret
\lett{O}{Lord,} we beseech thee that the holy prayer of thy blessed Apostle Andrew may commend our sacrifice unto thee: that it may be rendered acceptable by the merits of him in whose honour it is solemnly shewn forth. Through.
\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\communion{Follow me: I will make you fishers of men: and they straightway left their nets, and followed the Lord.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that these divine mysteries, which we have joyfully received on the festivity (commemoration) of blessed Andrew thine Apostle: may effectually avail for the glory of thy Saints, and likewise for our pardon. Through.
\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\subby{Conception of the Blessed Virgin Mary}
\feastday{Conception of BVM}
\begin{inhead}
    {Second Class Double with a Common Octave\\
8 December}
\end{inhead}
\feastdate{8}{December}

\introit
\lett{L}{et} us all rejoice in the Lord, and celebrate this feast in honour of the Virgin Mary, at whose Conception Angels rejoice and praise the Son of God. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
\lett{O}{God,} mercifully hear the supplication of thy servants; that we who are assembled together on the Conception of the Virgin Mother of God, may at her intercession be delivered by thee from the dangers which beset us. Through.
\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\readingcitation{Epistle}{Ecclesiasticus 24:17}
%\lett{A}{s} the vine brought I forth pleasant savour, and my flowers are the fruit of honour and riches. I am the mother of fair love, and fear, and knowledge, and holy hope: I therefore, being eternal, am given to all my children which are named of him. Come unto me, all ye that be desirous of me, and fill yourselves with my fruits. For my memorial is sweeter than honey, and mine inheritance than the honeycomb. They that eat me shall yet be hungry, and they that drink me shall yet be thirsty. He that obeyeth me shall never be confounded, and they that work by me shall not do amiss.
%RV:
\lett{A}{s} the vine I put forth grace; And my flowers are the fruit of glory and riches. Come unto me, ye that are desirous of me, And be ye filled with my produce. For my memorial is sweeter than honey, And mine inheritance than the honeycomb. They that eat me shall yet be hungry; And they that drink me shall yet be thirsty. He that obeyeth me shall not be ashamed; And they that work in me shall not do amiss. 
\gradall{Hearken, O daughter, and consider: incline thine ear. So shall the king have pleasure in thy beauty. ℣. According to thy worship and renown, good luck have thou with thine honour: ride on, because of the word of truth, of meekness, and righteousness; and thy right hand shall teach thee terrible things.}{Alleluia, alleluia. ℣. The Conception of the glorious Virgin Mary who sprang from the seed of Abraham, of the tribe of Judah, of the root of David. Alleluia.}
\subbysub{Sequence}
\begin{multicols}{2}
\lett{L}{et} us celebrate this day\par
\secondline{Whereon piously we say}
\thirdline{Mary was conceived.}
Begotten is the Mother Maid,

Conceived, created, channel made

Of pardon to the world.

Adam's primeval banishment

%MANUAL CHANGE BASED ON LATIN:
Joachim's childless discontent

There find a remedy.\\
\par
This the Prophets have foreshown,

This was to the Patriarchs known:

The Virgin whence a flower should spring,

The Star which forth the Sun should bring,

On this day is conceived;

The flower which from the rod should bloom,

The Sun which of the Star should come,

Is Christ interpreted.\\
\par
O how happy, O how fair!

Sweet to us, to God how dear,

Hath this Conception been!

Misery now is at an end,

Mercy doth on earth descend,

For sorrow joy is seen.

A Mother new new offspring bears,

From a new Star new Sun appears,

New grace doth all inspire;

The Mother bears the Generator,

The Creature brings forth the Creator,

The Daughter bears the Sire:\\
\par
O unexampled novelty!

O new, unheard of dignity!

The Mother's holy chastity

The Son's Conception shows.

Rejoice, O gracious Virgin mild!

Fair rod with blossoms undefiled,

Mother ennobled by her Child,

Such grace no other knows.\\
\par
That which lay hid, in figure sealed,

By clouds mysterious concealed,

The future Mother hath revealed:

For once a Virgin pure and good

Reversed the laws of Motherhood:

Nature, surprised, beheld a flood

Of Deity outpoured.\\
\par
\selectlanguage{latin}
Triste fuit in Eva ve,

Sed ex Eva formans Ave,

Versa vice, sed non prave

Intus ferens in conclave

Verbum bonum et suave

Nobis, Mater Virgo, fave

Tua frui gratia.\\
\par
\selectlanguage{british}
Whoe'r thou art, without delay

Open thy lips, her praises pay;

Offer her homage, to her pray

At every hour, on every day.

With swelling voice, with spirit sage,

By supplicating prayer engage

A portion in her patronage.\\
\par
Thou of the sad art comfort sure,

True Mother of the orphans poor,

Of the opprest the help secure,

Thou of the sick the healing cure,

All things to all thou givest.

With one consent we ask of thee,

Whom praise awaits especially,

Conduct us wanderers o'er this sea

Unto salvation's port, where we

By grace may be at rest. Amen.
\end{multicols}
\readingcitation{Gospel}{Matthew 1:1}
\lett{T}{he} book of the generation of Jesus Christ, the son of David, the son of Abraham. Abraham begat Isaac; and Isaac begat Jacob; and Jacob begat Judas and his brethren; And Judas begat Phares and Zara of Thamar; and Phares begat Esrom; and Esrom begat Aram; And Aram begat Aminadab; and Aminadab begat Naasson; and Naasson begat Salmon; And Salmon begat Booz of Rachab; and Booz begat Obed of Ruth; and Obed begat Jesse; And Jesse begat David the king; and David the king begat Solomon of her that had been the wife of Urias; And Solomon begat Roboam; and Roboam begat Abia; and Abia begat Asa; And Asa begat Josaphat; and Josaphat begat Joram; and Joram begat Ozias; And Ozias begat Joatham; and Joatham begat Achaz; and Achaz begat Ezekias; And Ezekias begat Manasses; and Manasses begat Amon; and Amon begat Josias; And Josias begat Jechonias and his brethren, about the time they were carried away to Babylon: And after they were brought to Babylon, Jechonias begat Salathiel; and Salathiel begat Zorobabel; And Zorobabel begat Abiud; and Abiud begat Eliakim; and Eliakim begat Azor; And Azor begat Sadoc; and Sadoc begat Achim; and Achim begat Eliud; And Eliud begat Eleazar; and Eleazar begat Matthan; and Matthan begat Jacob; And Jacob begat Joseph the husband of Mary, of whom was born Jesus, who is called Christ.
\offertory{Full of grace are thy lips, because God hath blessed thee for ever.}
\secret
\lett{O}{Lord,} let the human nature of thine only-begotten succour us, that he, who being born of a Virgin diminished not but sanctified the chastity of his Mother, may put away our offences from us on the feast of her Conception, and make our oblations acceptable to himself, even Jesus Christ our Lord. Who liveth.
\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\communion{Thy Son's true faith hath cleansed the world of sin. Immaculate thy virginity abideth.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord, that the sacrament which we have received of our bounden duty on this yearly celebration, may, at the intercession of blessed Mary ever-Virgin, afford us relief in this present life and in the world to come. Through.
\begin{rubric}
    In Advent, Commemoration is made of the Feria.
\end{rubric}

\subby{St. Thomas}
\feastday{St. Thomas}
\begin{inhead}
    {Second Class Double\\
21 December}
\end{inhead}
\feastdate{21}{December}

\properantiphonfix

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine uprising.
\collect
%1928 BCP:
%\lett{A}{lmighty} and everliving God, who, for the greater confirmation of the faith, didst suffer thy holy Apostle Thomas to be doubtful in thy Son's resurrection; Grant us so perfectly, and without all doubt, to believe in thy Son Jesus Christ, that our faith in thy sight may never be reproved. Hear us, O Lord, through the same Jesus Christ, to whom, with thee and the Holy Ghost, be all honour and glory, now and for evermore. Amen.
\lett{G}{rant} us, we beseech thee, O Lord, to glory in the solemnity of thy blessed Apostle Thomas: that we may ever be succoured by his protection; and follow his faith with worthy devotion. Through.

\begin{rubric}
    Commemoration of the Feria.
\end{rubric}

%1892 \& Sarum:
\readingcitation{Epistle}{Ephesians 2:19}
\lett{B}{rethren:} Ye are no more strangers and foreigners, but fellowcitizens with the saints, and of the household of God; and are built upon the foundation of the apostles and prophets, Jesus Christ himself being the chief corner stone; in whom all the building fitly framed together groweth unto an holy temple in the Lord: in whom ye also are builded together for an habitation of God through the Spirit.

%1928 reversed to conform to Sarum:
%\readingcitation{Epistle}{Hebrews 10:35}
%\lett{B}{rethren:} Cast not away therefore your confidence, which hath great recompence of reward. For ye have need of patience, that, after ye have done the will of God, ye might receive the promise. For yet a little while, and he that shall come will come, and will not tarry. Now the just shall live by faith: but if any man draw back, my soul shall have no pleasure in him. But we are not of them who draw back unto perdition; but of them that believe to the saving of the soul. Now faith is the substance of things hoped for, the evidence of things not seen.

\gradall{Right honourable are thy friends, O God: right well is their princedom established. ℣. If I tell them, they are more in number than the sand.}{Alleluia, alleluia. ℣. Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful. Alleluia.}
\readingcitation{Gospel}{John 20:24}
\lett{A}{t that time:} Thomas, one of the twelve, called Didymus, was not with them when Jesus came. The other disciples therefore said unto him, We have seen the Lord. But he said unto them, Except I shall see in his hands the print of the nails, and put my finger into the print of the nails, and thrust my hand into his side, I will not believe. And after eight days again his disciples were within, and Thomas with them: then came Jesus, the doors being shut, and stood in the midst, and said, Peace be unto you. Then saith he to Thomas, Reach hither thy finger, and behold my hands; and reach hither thy hand, and thrust it into my side: and be not faithless, but believing. And Thomas answered and said unto him, My Lord and my God. Jesus saith unto him, Thomas, because thou hast seen me, thou hast believed: blessed are they that have not seen, and yet have believed. And many other signs truly did Jesus in the presence of his disciples, which are not written in this book: But these are written, that ye might believe that Jesus is the Christ, the Son of God; and that believing ye might have life through his name.
\offertory{Their sound is gone out into all lands: and their words into the ends of the world.}
\secret
\lett{W}{e} render unto thee, O Lord, this duty of our bounden service, humbly beseeching thee: that like as we do offer unto thee this sacrifice of praise in honour of the confession of blessed Thomas the Apostle, so by his prayers thou wouldest preserve thy gifts within us. Through.
\begin{rubric}
    Commemoration of the Feria.%, except in Conventual Masses if an Ember Day occur.
\end{rubric}
\communion{Reach hither thy hand, and behold the print of the nails: and be not faithless, but believing.}
\postcommunion
\lett{A}{ssist} us, O merciful God: and, at the intercession of blessed Thomas the Apostle on our behalf, continue towards us the gifts of thy loving-kindness. Through.
\begin{rubric}
    Commemoration of the Feria.%, except in Conventual Masses if an Ember Day occur.
\end{rubric}
\begin{rubric}
    The Last Gospel is read of the Ember Day, if one occur.%, in non-conventual Masses.
\end{rubric}


\subby{Conversion of St. Paul}
\feastday{Conversion of St. Paul}
\begin{inhead}
    {Greater Double\\
25 January}
\end{inhead}
\feastdate{25}{January}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} Thou hast given an heritage unto those that fear thy Name, O Lord.%\vr{Ps. 61:5}

\introit
\lett{I}{know} whom I have believed, and am persuaded that he is able to keep that which I have committed unto him against that day, a just judge. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest, my down-sitting, and mine up-rising.
\collect
\lett{O}{God,} who, through the preaching of the blessed Apostle Saint Paul, hast caused the light of the Gospel to shine throughout the world; Grant, we beseech thee, that we, having his wonderful conversion in remembrance, may show forth our thankfulness unto thee for the same, by following the holy doctrine which he taught. Through.
\needspace{4\baselineskip}
\lett{O}{God,} who didst bestow upon thy blessed Apostle Peter the keys of the kingdom of heaven, and didst appoint unto him the high priesthood of binding and loosing: vouchsafe; that by the help of his intercession we may be delivered from the bonds of our iniquities. Who livest.

\readingcitation{Epistle}{Acts 9:1}
\lett{I}{n those days:} Saul, yet breathing out threatenings and slaughter against the disciples of the Lord, went unto the high priest, And desired of him letters to Damascus to the synagogues, that if he found any of this way, whether they were men or women, he might bring them bound unto Jerusalem. And as he journeyed, he came near Damascus: and suddenly there shined round about him a light from heaven: And he fell to the earth, and heard a voice saying unto him, Saul, Saul, why persecutest thou me? And he said, Who art thou, Lord? And the Lord said, I am Jesus whom thou persecutest: it is hard for thee to kick against the pricks. And he trembling and astonished said, Lord, what wilt thou have me to do? And the Lord said unto him, Arise, and go into the city, and it shall be told thee what thou must do. And the men which journeyed with him stood speechless, hearing a voice, but seeing no man. And Saul arose from the earth; and when his eyes were opened, he saw no man: but they led him by the hand, and brought him into Damascus. And he was three days without sight, and neither did eat nor drink. And there was a certain disciple at Damascus, named Ananias; and to him said the Lord in a vision, Ananias. And he said, Behold, I am here, Lord. And the Lord said unto him, Arise, and go into the street which is called Straight, and enquire in the house of Judas for one called Saul, of Tarsus: for, behold, he prayeth, And hath seen in a vision a man named Ananias coming in, and putting his hand on him, that he might receive his sight. Then Ananias answered, Lord, I have heard by many of this man, how much evil he hath done to thy saints at Jerusalem: And here he hath authority from the chief priests to bind all that call on thy name. But the Lord said unto him, Go thy way: for he is a chosen vessel unto me, to bear my name before the Gentiles, and kings, and the children of Israel: For I will shew him how great things he must suffer for my name's sake. And Ananias went his way, and entered into the house; and putting his hands on him said, Brother Saul, the Lord, even Jesus, that appeared unto thee in the way as thou camest, hath sent me, that thou mightest receive thy sight, and be filled with the Holy Ghost. And immediately there fell from his eyes as it had been scales: and he received sight forthwith, and arose, and was baptized. And when he had received meat, he was strengthened. Then was Saul certain days with the disciples which were at Damascus. And straightway he preached Christ in the synagogues, that he is the Son of God. But all that heard him were amazed, and said; Is not this he that destroyed them which called on this name in Jerusalem, and came hither for that intent, that he might bring them bound unto the chief priests? But Saul increased the more in strength, and confounded the Jews which dwelt at Damascus, proving that this is very Christ.
\gradall{He that wrought effectually in Peter to the apostleship, the same was mighty in me toward the Gentiles: and they perceived the grace that was given unto me. ℣. The grace of God which was bestowed upon me was not in vain: but his grace ever abideth in me.}{Alleluia, alleluia. ℣. Great and holy is Paul, a chosen vessel, meet indeed to be glorified, who also was worthy to possess the twelfth throne. Alleluia.}
\readingcitation{Gospel}{Matthew 19:27}
\lett{A}{t that time:} Peter said unto Jesus: Behold, we have forsaken all, and followed thee; what shall we have therefore? And Jesus said unto them, Verily I say unto you, That ye which have followed me, in the regeneration when the Son of man shall sit in the throne of his glory, ye also shall sit upon twelve thrones, judging the twelve tribes of Israel. And every one that hath forsaken houses, or brethren, or sisters, or father, or mother, or wife, or children, or lands, for my name's sake, shall receive an hundredfold, and shall inherit everlasting life. But many that are first shall be last; and the last shall be first.
\offertory{Right honourable are thy friends unto me, O God: right well is their princedom established.}
\secret
\lett{S}{anctify,} O Lord, through the prayers of thine Apostle Paul, the gifts of thy people: that those things, which by thine institution are pleasing unto thee, may be made more pleasing by his prayer and advocacy. Through.
\lett{W}{e} beseech thee, O Lord, that the intercession of blessed Peter the Apostle may commend unto thee the prayers and sacrifices of thy Church: that those things which we celebrate for his glory may avail for our pardon. Through.
\communion{Verily, I say unto you: that ye which have forsaken all and followed me, shall receive an hundredfold, and shall inherit everlasting life.}
\postcommunion
\lett{O}{Lord,} who hast sanctified us with this saving mystery: we beseech thee; that he, whom thou hast given to be our advocate and guide, may never fail in prayer for us. Through.
\lett{M}{ay} the gift, O Lord, which we have offered make us to rejoice: that as we proclaim thy wonders in thine Apostle Peter; so through him we may receive the abundance of thy loving-kindness. Through.

\subby{Purification of the Blessed Virgin Mary}
\feastday{Candlemas}
\begin{inhead}
    {Second Class Double\\
2 February}
\end{inhead}
\feastdate{2}{February}
\begin{rubric}
    If it happen that this Feast be transferred, only the Blessing and Distribution of the Candles and the Procession take place today. Mass is said of the Office occurring, and in it lighted Candles are not held in the hands.
\end{rubric}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} It was revealed unto Simeon by the Holy Ghost, that he should not see death, until he had seen the Lord's Christ.

\begin{rubric}
    Terce (or, Mattins) being ended, the Priest vested in a violet Cope, or without a Chasuble, with the Ministers vested in like manner, proceeds to bless the Candles, which are placed in the midst before the Altar or at the Epistle corner, and he himself, standing in the same place facing the Altar, says with joined hands in the tone of the Collect of the Ferial Mass:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{O}{Lord} holy Father almighty, everlasting God, who hast created all things out of nothing, and by thy command hast caused this liquid through the labour of bees to come to the perfection of wax: and who on this day didst fulfil the petition of Simeon the just: we humbly entreat thee; that through the invocation of thy most holy name and the intercession of blessed Mary ever Virgin, whose festival we this day devoutly celebrate, and through the prayers of all thy Saints, thou wouldest vouchsafe to bl {\ding{64}} ess and sanc {\ding{64}} tify these candles for the use of men, and the health of their bodies and souls, whether on land or on water: and that from thy holy heaven, and from the seat of thy majesty, thou wouldest hearken to the voices of this thy people, who desire to bear them in their hands to thine honour and to sing thy praises: and that thou wouldest be gracious to all that cry unto thee, whom thou hast redeemed with the precious Blood of thy Son: Who liveth and reigneth with thee. \textit{Amen.}
\letuspray
\lett{A}{lmighty} and everlasting God, who on this day didst present thine only-begotten Son in thy holy temple, to be received in the arms of Saint Simeon: we humbly entreat thy mercy; that thou wouldest vouchsafe to bl {\ding{64}} ess, sanc {\ding{64}} tify, and kindle with the light of thy heavenly benediction these candles, which we thy servants, receiving to the glory of thy name, desire to kindle and bear forth: to the intent that we, being made worthy by the offering of them unto thee, the Lord our God, and being enkindled with the holy fire of thy most sweet charity, may be found meet to be presented in the holy temple of thy glory. Through the same. \textit{Amen.}
\letuspray
\lett{O}{Lord} Jesu Christ, the true light, that lightenest every man that cometh into this world: pour forth thy bles {\ding{64}} sing upon these candles, and sanc {\ding{64}} tify them with the light of thy grace, and mercifully grant; that even as these lights, kindled with visible fire, do scatter the darkness of night; so our hearts, illumined by invisible fire, that is, by the brightness of the Holy Spirit, may be free from the blindness of every iniquity: that, the eyes of our minds being purified, we may be enabled to discern such things as are pleasing unto thee and profitable for our salvation; to the end that after the darkness and dangers of this world, we may be found meet to attain unto the light unfailing. Through thee, Christ Jesu, Saviour of the world who in the perfect Trinity livest and reignest God. \textit{Amen.}
\letuspray
\lett{A}{lmighty} and everlasting God, who through thy servant Moses didst command the purest liquid of oil to be prepared for the lamps to burn continually before thy face: mercifully pour upon these candles the grace of thy bles {\ding{64}} sing; that they may so outwardly shed forth light, that, by thy gift, the light of thy Spirit may not inwardly be wanting to our souls. Through . . . in the unity of the same. \textit{Amen.}
\letuspray
\lett{O}{Lord} Jesu Christ, who on this day appearing among men in the substance of our flesh wast presented by thy parents in the temple: whom the old man, the venerable Simeon, illumined by the light of thy Spirit, acknowledged, received, and blessed: mercifully grant; that we, being enlightened and instructed by the grace of the same Holy Spirit, may truly acknowledge and faithfully love thee: Who livest and reignest with God the Father in the unity of the same Holy Spirit, God. \textit{Amen.}

\begin{rubric}
    The Prayers being ended, the Celebrant sets incense in the thurible; and thrice sprinkles the Candles with holy water, saying the Antipbon, \emph{Thou shalt purge me}, without chant and without the Psalm: and then thrice incenses them.
\end{rubric}
\begin{rubric}
    Then the senior of the Clergy approaches the Altar, and from him the Celebrant receives a Candle without genuflecting or kissing his hand. Then the Celebrant, standing in the midst before the Altar and turning to the people, distributes candles first to the senior, from whom he himself received; then to the Deacon and Subdeacon vested, and to the rest of the Clergy, one by one in order, lastly to the laity: all kneeling and kissing the Candle and the Celebrant's hand, with the exception of Prelates, if they be present. And when he begins to distribute the Candles the Choir sings:
\end{rubric}

\antiphon{}{A light to lighten the Gentiles and the glory of thy people Israel.}

\antiphon{}{Lord, now lettest thou thy servant depart in peace, according to thy word.}

\begin{rubric}
    Then is repeated the whole Antiphon, \emph{A light to lighten the Gentiles}, which is likewise repeated after each Verse.
\end{rubric}

\antiphon{}{For mine eyes have seen thy salvation. \inrub{Antiphon.}}

\antiphon{}{Which thou hast prepared before the face of all people. \inrub{Antiphon.}}

\antiphon{}{Glory be to the Father, and to the Son, and to the Holy Ghost. \inrub{Antiphon.}}

\antiphon{}{As it was in the beginning, is now, and ever shall be, world without end. Amen. \inrub{Antiphon.}}

\begin{rubric}
    These being ended, shall be sung:
\end{rubric}

\antiphon{}{O Lord, arise, help us: and deliver us for thy name's sake. \textit{Ps.} O God, we have heard with our ears: our fathers have declared unto us. ℣. Glory be . . .}

\antiphon{}{O Lord, arise, help us: and deliver us for thy name's sake.}

\elcol{℣. Let us pray.}{℣. Orémus.}

\begin{rubric}
    If it be after Septuagesima, and not a Sunday, the Deacon says the following versicle.
\end{rubric}

\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord, hearken unto thy people: and grant, that by the light of thy grace we may inwardly attain to those things, which in this yearly devotion thou dost suffer us outwardly to venerate. Through Christ, our Lord. \textit{Amen.}

\begin{rubric}
    Then follows the Procession. And first the Celebrant sets incense in the censer: then the Deacon, turning himself to the people, says: \emph{Let us proceed in peace.} And the Choir answers: \emph{In the name of Christ. Amen.}
\end{rubric}

\begin{rubric}
    The Thurifer goes first, with smoking thurible: then the Subdeacon vested, carrying the Cross, in the midst between two Acolytes with lighted candles: the Clergy follow in order, last of all the Celebrant with the Deacon on his left, all with lighted Candles in their hands: and the following Antiphons are sung:
\end{rubric}

\antiphon{}{O Sion, adorn thy bride-chamber, and receive Christ the King: greet Mary, who is the gate of heaven: for she beareth the King of the glory of the new light: she remaineth a Virgin, yet beareth in her hands a Son begotten before the morning star: whom Simeon took into his arms, declaring to the nations that he is the Lord of life and death, and Saviour of the world.}

\antiphon{Another}{It was revealed unto Simeon by the Holy Ghost, that he should not see death, before he had seen the Lord's Christ: and when they brought the Child into the temple, then took he him up in his arms, and blessed God, and said: Lord, now lettest thou thy servant depart in peace. ℣. When his parents brought in the Child Jesus, to do for him according to the custom of the law, then took he him up in his arms.}

\begin{rubric}
    As they enter the Church, the following is sung:
\end{rubric}

℣. They offered for him unto the Lord a pair of turtle doves or two young pigeons: As it is written in the law of the Lord.

℣. When the days of Mary's purification according to the law of Moses were accomplished, they brought Jesus to Jerusalem, to present him to the Lord. As it is written in the law of the Lord.

℣. Glory be to the Father, and to the Son, and to the Holy Ghost. As it is written in the law of the Lord.

\begin{rubric}
    The Procession being ended, the Celebrant and Ministers lay aside their violet vestments, and put on white for the Mass. And Candles are held in the hands lighted, while the Gospel is read, and again from the beginning of the Canon to the end of the Communion.
\end{rubric}

\begin{rubric}
    Today any Votive Mass, even solemn, of Christ the Lord, is forbidden.
\end{rubric}
\begin{rubric}
    The Mass is then said.
\end{rubric}
\introit
\lett{W}{e} have waited, O God, for thy loving-kindness in the midst of thy temple: according to thy name, O God, so is thy praise unto the world's end: thy right hand is full of righteousness. \textit{Ps.} Great is the Lord, and highly to be praised: in the city of our God, even upon his holy hill.
\collect
\lett{A}{lmighty} and everliving God, we humbly beseech thy Majesty, that, as thy only-begotten Son was this day presented in the temple in substance of our flesh, so we may be presented unto thee with pure and clean hearts. By the same thy Son Jesus Christ our Lord. Who liveth.
\readingcitation{Epistle}{Malachi 3:1}
\lett{B}{ehold,} I will send my messenger, and he shall prepare the way before me: and the Lord, whom ye seek, shall suddenly come to his temple, even the messenger of the covenant, whom ye delight in: behold, he shall come, saith the \textsc{Lord} of hosts. But who may abide the day of his coming? and who shall stand when he appeareth? for he is like a refiner's fire, and like fullers' soap: And he shall sit as a refiner and purifier of silver: and he shall purify the sons of Levi, and purge them as gold and silver, that they may offer unto the \textsc{Lord} an offering in righteousness. Then shall the offering of Judah and Jerusalem be pleasant unto the \textsc{Lord}, as in the days of old, and as in former years. And I will come near to you to judgment; and I will be a swift witness against the sorcerers, and against the adulterers, and against false swearers, and against those that oppress the hireling in his wages, the widow, and the fatherless, and that turn aside the stranger from his right, and fear not me, saith the \textsc{Lord} of hosts.

\gradual{We have waited, O God, for thy loving-kindness in the midst of thy temple: according to thy name, O God, so is thy praise unto the world's end. ℣. Like as we have heard, so have we seen, in the city of our God, even upon his holy hill.}

%MANUAL ADJUSTMENT:

\alleluia{Alleluia, alleluia. ℣. The old man carried the Child: but the Child was the old man's King. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent

\tract{Lord, now lettest thou thy servant depart in peace, according to thy word. ℣. For mine eyes have seen thy salvation. ℣. Which thou hast prepared before the face of all people. ℣. A light to lighten the Gentiles and the glory of thy people Israel.}

\readingcitation{Gospel}{Luke 2:22}
\lett{A}{t that time:} When the days of her purification according to the law of Moses were accomplished, they brought him to Jerusalem, to present him to the Lord; (As it is written in the law of the Lord, Every male that openeth the womb shall be called holy to the Lord;) And to offer a sacrifice according to that which is said in the law of the Lord, A pair of turtledoves, or two young pigeons. And, behold, there was a man in Jerusalem, whose name was Simeon; and the same man was just and devout, waiting for the consolation of Israel: and the Holy Ghost was upon him. And it was revealed unto him by the Holy Ghost, that he should not see death, before he had seen the Lord's Christ. And he came by the Spirit into the temple: and when the parents brought in the child Jesus, to do for him after the custom of the law, Then took he him up in his arms, and blessed God, and said, Lord, now lettest thou thy servant depart in peace, according to thy word: For mine eyes have seen thy salvation, Which thou hast prepared before the face of all people; A light to lighten the Gentiles, and the glory of thy people Israel. And Joseph and his mother marvelled at those things which were spoken of him. And Simeon blessed them, and said unto Mary his mother, Behold, this child is set for the fall and rising again of many in Israel; and for a sign which shall be spoken against; (Yea, a sword shall pierce through thy own soul also,) that the thoughts of many hearts may be revealed. And there was one Anna, a prophetess, the daughter of Phanuel, of the tribe of Aser: she was of a great age, and had lived with an husband seven years from her virginity; And she was a widow of about fourscore and four years, which departed not from the temple, but served God with fastings and prayers night and day. And she coming in that instant gave thanks likewise unto the Lord, and spake of him to all them that looked for redemption in Jerusalem. And when they had performed all things according to the law of the Lord, they returned into Galilee, to their own city Nazareth. And the child grew, and waxed strong in spirit, filled with wisdom: and the grace of God was upon him.
\offertory{Full of grace are thy lips: because God hath blessed thee for ever and ever.}

%MANUAL ADJUSTMENT:
\vspace{-3ex}

\secret
\lett{G}{raciously} hear our prayers, O Lord: and, that the gifts which we offer before the eyes of thy majesty may be worthy, do thou bestow upon us the succour of thy loving-kindness. Through.
\communion{It was revealed unto Simeon by the Holy Ghost, that he should not see death, before he had seen the Lord's Christ.}

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that these holy mysteries, which thou hast given to us for the assurance of our redemption, may, by the intercession of blessed Mary ever Virgin, be for our healing both in this life and that which is to come. Through.


\subby{Chair of St. Peter at Antioch}
\feastday{Chair of St. Peter}
\begin{inhead}
    {Second Class Double\\
22 February}
\end{inhead}
\feastdate{22}{February}
%S-R: 1
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
%Due to lacking in Sarum, Opening Sentence is taken from Dominican Breviary.
\textit{Opening Sentence.} That they would exalt him also in the congregation of the people : and praise him in the seat of the elders!%\vr{Ps. 107:32}

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\introit
\lett{T}{he} Lord hath established a covenant of peace with him, and made him a prince: that he should have the dignity of the priesthood for ever. \textit{Ps.} Lord, remember David: and all his trouble.

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\collect
\lett{O}{God,} who didst bestow upon thy blessed Apostle Peter the keys of the kingdom of heaven, and didst appoint unto him the high priesthood of binding and loosing: vouchsafe; that by the help of his intercession we may be delivered from the bonds of our iniquities. Who livest.

%MANUAL ADJUSTMENT:
%\needspace{4\baselineskip}
\lett{O}{God,} who by the preaching of the blessed Apostle Paul didst teach the multitude of the Gentiles: grant to us, we beseech thee; that we who celebrate his commemoration may know him to be our advocate with thee. (Through.)
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\begin{rubric}
    If this Feast fall on a Saturday---not in a Leap Year---the following Commemoration is made of the Vigil of St. Matthias.
\end{rubric}
\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty God: that the venerable solemnity of blessed Matthias thine Apostle, which we here prevent, may increase our devotion and set forward our salvation. Through.
\readingcitation{Epistle}{1 Peter 1:1}
\lett{P}{eter,} an apostle of Jesus Christ, to the strangers scattered throughout Pontus, Galatia, Cappadocia, Asia, and Bithynia, Elect according to the foreknowledge of God the Father, through sanctification of the Spirit, unto obedience and sprinkling of the blood of Jesus Christ: Grace unto you, and peace, be multiplied. Blessed be the God and Father of our Lord Jesus Christ, which according to his abundant mercy hath begotten us again unto a lively hope by the resurrection of Jesus Christ from the dead, To an inheritance incorruptible, and undefiled, and that fadeth not away, reserved in heaven for you, Who are kept by the power of God through faith unto salvation ready to be revealed in the last time. Wherein ye greatly rejoice, though now for a season, if need be, ye are in heaviness through manifold temptations: That the trial of your faith, being much more precious than of gold that perisheth, though it be tried with fire, might be found unto praise and honour and glory at the appearing of Jesus Christ our Lord.

\gradall{Let them exalt him in the congregation of the people: and praise him in the seat of the elders. ℣. O that men would praise the Lord for his goodness, and declare the wonder that he doeth for the children of men.}{Alleluia, alleluia. ℣. Thou art Peter, and upon this rock I will build my Church. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent

\tract{Thou art Peter, and upon this rock I will build my Church. ℣. And the gates of hell shall not prevail against it: and I will give unto thee the keys of the kingdom of heaven. ℣. Whatsoever thou shalt bind on earth shall be bound in heaven. ℣. And whatsoever thou shalt loose on earth shall be loosed in heaven.}

\readingcitation{Gospel}{Matthew 16:13}
\lett{A}{t that time:} When Jesus came into the coasts of Caesarea Philippi, he asked his disciples, saying, Whom do men say that I the Son of man am? And they said, Some say that thou art John the Baptist: some, Elias; and others, Jeremias, or one of the prophets. He saith unto them, But whom say ye that I am? And Simon Peter answered and said, Thou art the Christ, the Son of the living God. And Jesus answered and said unto him, Blessed art thou, Simon Barjona: for flesh and blood hath not revealed it unto thee, but my Father which is in heaven. And I say also unto thee, That thou art Peter, and upon this rock I will build my church; and the gates of hell shall not prevail against it. And I will give unto thee the keys of the kingdom of heaven: and whatsoever thou shalt bind on earth shall be bound in heaven: and whatsoever thou shalt loose on earth shall be loosed in heaven.
\offertory{Thou art Peter, and upon this rock I will build my Church: and the gates of hell shall not prevail against it: and I will give unto thee the keys of the kingdom of heaven.}
\secret
\lett{W}{e} beseech thee, O Lord, that the intercession of blessed Peter the Apostle may commend unto thee the prayers and sacrifices of thy Church: that those things which we celebrate for his glory may avail for our pardon. Through.
\needspace{4\baselineskip}
\lett{S}{anctify,} O Lord, through the prayers of thine Apostle Paul, the gifts of thy people: that those things, which by thine institution are pleasing unto thee, may be made more pleasing by his prayer and advocacy. (Through.)
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\begin{rubric}
    If this Feast fall on a Saturday---not in a Leap Year---the following Commemoration is made of the Vigil of St. Matthias.
\end{rubric}
\needspace{4\baselineskip}
\lett{O}{Lord,} who didst exalt blessed Matthias to be numbered among thine Apostles: grant that we thy people, who on this day, preventing his heavenly birth, do offer unto thee these holy mysteries, may be assisted by his intercession both in the making of our supplications before thee, and in the obtaining of all that we desire. Through.

\communion{Thou art Peter, and upon this rock I will build my Church.}

\postcommunion
\lett{M}{ay} the gift, O Lord, which we have offered, make us to rejoice: that as we proclaim thy wonders in thine Apostle Peter; so through him we may receive the abundance of thy loving-kindness. Through.
%\needspace{4\baselineskip}

\lett{O}{Lord,} who hast sanctified us with this saving mystery: we beseech thee; that he whom thou hast given to be our advocate and guide may never fail in prayer for us. (Through.)
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\begin{rubric}
    If this Feast fall on a Saturday---not in a Leap Year---the following Commemoration is made of the Vigil of St. Matthias.
\end{rubric}
\needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord, mercifully to hear the supplication of thy holy Apostle Matthias: that thou wouldest vouchsafe unto us thy pardon, and grant us thy everlasting healing. Through.
\begin{rubric}
    In Lent, the Last Gospel is of the Feria.
\end{rubric}


\subby{St. Scholastica}
\feastday{{St. Scholastica}}
\fancyhead[RE,LO]{10 February}
\begin{inhead}
    {Second Class Double\\
10 February}
\end{inhead}

%From the 1781 Missale Romano-Monasticum:

%Translation from D-R
\introit
\lett{O}{how} beautiful is the chaste generation with glory: for the memory thereof is immortal: because it is known both with God and with men. \textit{Ps.} My heart and my flesh rejoice in the living God.

\collect
\lett{O}{God,} who didst reveal in a vision the soul of blessed Scholastica thy Virgin entering heaven in the likeness of a dove, that thou mightest shew the way of the undefiled: grant us by the aid of her merits and prayers so innocently to live, that we may worthily attain unto joys eternal. Through.

\readingcitation{Epistle}{Song of Songs 2:10}
\lett{M}{y} beloved spake, and said unto me, Rise up, my love, my fair one, and come away. For, lo, the winter is past, the rain is over and gone; The flowers appear on the earth; the time of the singing of birds is come, and the voice of the turtle is heard in our land; The fig tree putteth forth her green figs, and the vines with the tender grape give a good smell. Arise, my love, my fair one, and come away. O my dove, that art in the clefts of the rock, in the secret places of the stairs, let me see thy countenance, let me hear thy voice; for sweet is thy voice, and thy countenance is comely.

\gradall{Whom have I in heaven but thee: and there is none upon earth that I desire in comparison of thee. ℣. My flesh and my heart faileth: but God is the strength of my heart, and my portion for ever.}{Alleluia, alleluia. ℣. They are virgins. These are they which follow the Lamb whithersoever he goeth. And in their mouth was found no guile: for they are without fault before the throne of God. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent

\tract{Blessed are those that are undefiled in the way: and walk in the law of the Lord. ℣. Blessed are they that keep his testimonies: and seek him with their whole heart. ℣. O that my ways were made so direct: that I might keep thy statutes!}

\readingcitation{Gospel}{Matthew 19:3}
\lett{A}{t that time:} The Pharisees came unto Jesus, tempting him, and saying unto him: Is it lawful for a man to put away his wife for every cause? And he answered and said unto them, Have ye not read, that he which made them at the beginning made them male and female, And said, For this cause shall a man leave father and mother, and shall cleave to his wife: and they twain shall be one flesh? Wherefore they are no more twain, but one flesh. What therefore God hath joined together, let not man put asunder. They say unto him, Why did Moses then command to give a writing of divorcement, and to put her away? He saith unto them, Moses because of the hardness of your hearts suffered you to put away your wives: but from the beginning it was not so. And I say unto you, Whosoever shall put away his wife, except it be for fornication, and shall marry another, committeth adultery: and whoso marrieth her which is put away doth commit adultery. His disciples say unto him, If the case of the man be so with his wife, it is not good to marry. But he said unto them, All men cannot receive this saying, save they to whom it is given. For there are some eunuchs, which were so born from their mother's womb: and there are some eunuchs, which were made eunuchs of men: and there be eunuchs, which have made themselves eunuchs for the kingdom of heaven's sake. He that is able to receive it, let him receive it.

\offertory{The virgins that be her fellows shall bear her company, and shall be brought unto thee. With joy and gladness shall they be brought, and shall enter into the King's palace.}

\secret
\lett{R}{eceive,} O Lord, through the intercession of blessed Scholastica thy Virgin, our humble obedience; and, through the immaculate Victim, cleanse us from every stain of the passions, that we may ever burn with pure and holy love in thy sight. Through.

\communion{For how great is his goodness, and how great is his beauty! corn shall make the young men cheerful, and new wine the virgins.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that, refreshed by the sweetness of thine ineffable Sacraments, and assisted by the suffrages of the blessed Virgin Scholastica, we may never be hindered by any earthly things from following thee, the Maker of all good things. Through.


\subby{St. Matthias}
\feastday{St. Matthias}
\begin{inhead}
    {Second Class Double\\
24 February (25 February in a Leap Year)}
\end{inhead}
\feastdate{24}{February}

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine up-rising.
\collect
%1928 BCP:
%\lett{O}{almighty} God, who into the place of the traitor Judas didst choose thy faithful servant Matthias to be of the number of the twelve Apostles; Grant that thy Church, being alway preserved from false Apostles, may be ordered and guided by faithful and true pastors. Through.
%Sarum:
\lett{O}{God,} who didst join blessed Matthias to the college of thine Apostles: grant, we beseech thee; that through his mediation we may ever perceive thy tender mercy towards us. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\readingcitation{Epistle}{Acts 1:15}
\lett{I}{n those days:} Peter stood up in the midst of the disciples, and said, (the number of names together were about an hundred and twenty,) Men and brethren, this scripture must needs have been fulfilled, which the Holy Ghost by the mouth of David spake before concerning Judas, which was guide to them that took Jesus. For he was numbered with us, and had obtained part of this ministry. Now this man purchased a field with the reward of iniquity; and falling headlong, he burst asunder in the midst, and all his bowels gushed out. And it was known unto all the dwellers at Jerusalem; insomuch as that field is called in their proper tongue, Aceldama, that is to say, The field of blood. For it is written in the book of Psalms, Let his habitation be desolate, and let no man dwell therein: and his bishoprick let another take. Wherefore of these men which have companied with us all the time that the Lord Jesus went in and out among us, Beginning from the baptism of John, unto that same day that he was taken up from us, must one be ordained to be a witness with us of his resurrection. And they appointed two, Joseph called Barsabas, who was surnamed Justus, and Matthias. And they prayed, and said, Thou, Lord, which knowest the hearts of all men, shew whether of these two thou hast chosen, That he may take part of this ministry and apostleship, from which Judas by transgression fell, that he might go to his own place. And they gave forth their lots; and the lot fell upon Matthias; and he was numbered with the eleven apostles.

\gradall{Right honourable are thy friends, O God: right well is their princedom established. ℣. If I tell them, they are more in number than the sand.}{%Taken from the Votive Mass of the Apostles:
Alleluia, alleluia. ℣. Right honourable are thy friends unto me, O God: right well is their princedom established. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent

\tract{Thou hast given him his heart's desire: and hast not denied him the request of his lips. ℣. For thou hast prevented him with the blessings of goodness: ℣. Thou hast set a crown of pure gold upon his head.}

\readingcitation{Gospel}{Matthew 11:25}
\lett{A}{t that time:} Jesus answered and said, I thank thee, O Father, Lord of heaven and earth, because thou hast hid these things from the wise and prudent, and hast revealed them unto babes. Even so, Father: for so it seemed good in thy sight. All things are delivered unto me of my Father: and no man knoweth the Son, but the Father; neither knoweth any man the Father, save the Son, and he to whomsoever the Son will reveal him. Come unto me, all ye that labour and are heavy laden, and I will give you rest. Take my yoke upon you, and learn of me; for I am meek and lowly in heart: and ye shall find rest unto your souls. For my yoke is easy, and my burden is light.
\offertory{Thou shalt make them princes in all lands: they shall remember thy name, O Lord, from one generation to another.}
\secret
\lett{M}{ay} the prayer of thy Apostle, Saint Matthias, O Lord, accompany the oblations which we offer for consecration to thy name: and grant that we may thereby be cleansed and defended. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\communion{Ye which have followed me, shall sit upon thrones, judging the twelve tribes of Israel.}
\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that through these holy things which we have received; we may, at the intercession of blessed Matthias, thine Apostle, obtain pardon and peace. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}

%MANUAL ADJUSTMENT
\clearpage
\subby{Pope St. Gregory}
\feastday{Pope St. Gregory}
\begin{inhead}
    {Second Class Double\\
12 March}
\end{inhead}
\feastdate{12}{March}

\introit
\lett{O}{ye} priests of the Lord, bless ye the Lord: O ye holy and humble
men of heart, bless ye the Lord. \textit{Cant.} O all ye works of the Lord, bless ye the Lord: praise him, and magnify him for ever.
\collect
\lett{O}{God,} who on the soul of thy servant Gregory, hast bestowed the rewards of everlasting bliss: mercifully grant; that we, who are oppressed by the burden of our sins, may by his prayers before thee be relieved. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\readingcitation{Epistle}{2 Timothy 4:1}
\lett{D}{early beloved:} I charge thee therefore before God, and the Lord Jesus Christ, who shall judge the quick and the dead at his appearing and his kingdom; Preach the word; be instant in season, out of season; reprove, rebuke, exhort with all longsuffering and doctrine. For the time will come when they will not endure sound doctrine; but after their own lusts shall they heap to themselves teachers, having itching ears; And they shall turn away their ears from the truth, and shall be turned unto fables. But watch thou in all things, endure afflictions, do the work of an evangelist, make full proof of thy ministry. For I am now ready to be offered, and the time of my departure is at hand. I have fought a good fight, I have finished my course, I have kept the faith: Henceforth there is laid up for me a crown of righteousness, which the Lord, the righteous judge, shall give me at that day: and not to me only, but unto all them also that love his appearing.

\gradtr{The Lord sware, and will not repent: Thou art a priest for ever after the order of Melchisedech. ℣. The Lord said unto my Lord: Sit thou on my right hand.}{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\readingcitation{Gospel}{Matthew 5:13}
\lett{A}{t that time:} Jesus said unto his disciples: Ye are the salt of the earth: but if the salt have lost his savour, wherewith shall it be salted? it is thenceforth good for nothing, but to be cast out, and to be trodden under foot of men. Ye are the light of the world. A city that is set on an hill cannot be hid. Neither do men light a candle, and put it under a bushel, but on a candlestick; and it giveth light unto all that are in the house. Let your light so shine before men, that they may see your good works, and glorify your Father which is in heaven. Think not that I am come to destroy the law, or the prophets: I am not come to destroy, but to fulfil. For verily I say unto you, Till heaven and earth pass, one jot or one tittle shall in no wise pass from the law, till all be fulfilled. Whosoever therefore shall break one of these least commandments, and shall teach men so, he shall be called the least in the kingdom of heaven: but whosoever shall do and teach them, the same shall be called great in the kingdom of heaven.
\offertory{My truth and my mercy shall be with him: and in my name shall his horn be exalted.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\secret
\lett{G}{rant} to us, we beseech thee, O Lord: that by the intercession of blessed Gregory this oblation, by the offering of which thou hast bestowed remission of sins on the whole world, may be profitable unto us. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\communion{A faithful and wise servant, whom the lord hath made ruler over his household: to give them their portion of meat in due season.}

%MANUAL ADJUSTMENT:
\vspace{-0.5ex}

\postcommunion
\lett{O}{God,} who didst make the blessed Bishop Gregory by his merits the equal of thy Saints: mercifully grant; that we, who celebrate the festival of his commemoration, may likewise follow the example of his life. Through.

\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}


\subby{St. Joseph, Spouse of the Blessed Virgin Mary}
\feastday{St. Joseph}
\begin{inhead}
    {First Class Double\\
19 March}
\end{inhead}
\feastdate{19}{March}

\introit
\lett{T}{he} just shall flourish like a palm-tree, and shall spread abroad like a cedar in Libanus: planted in the house of the Lord: in the courts of the house of our God. \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.
\collect
\lett{W}{e} beseech thee, O Lord, that we may be aided through the merits of the Spouse of thy most holy Mother: that those things, which by our own power we cannot obtain, may through his intercession be granted unto us. Who livest.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 45:1}
%\lett{B}{eloved} of God and men, whose memorial is blessed. He made him like to the glorious saints, and magnified him, so that his enemies stood in fear of him. By his words he caused the wonders to cease, and he made him glorious in the sight of kings, and gave him a commandment for his people, and shewed him part of his glory. He sanctified him in his faithfulness and meekness, and chose him out of all men. He made him to hear his voice, and brought him into the dark cloud, and gave him commandments before his face, even the law of life and knowledge.
%RV:
\lett{B}{eloved} of God and men, even Moses, Whose memorial is blessed. He made him like to the glory of the saints, And magnified him in the fears of his enemies. By his words he caused the wonders to cease; He glorified him in the sight of kings; He gave him commandment for his people, And shewed him part of his glory. He sanctified him in his faithfulness and meekness; He chose him out of all flesh. He made him to hear his voice, And led him into the thick darkness, And gave him commandments face to face, Even the law of life and knowledge.

\gradtr{Thou hast prevented him, O Lord, with the blessings of goodness: thou hast set a crown of pure gold upon his head. ℣. He asked life of thee, and thou gavest him a long life, even for ever and ever.}{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\readingcitation{Gospel}{Matthew 1:18}
\lett{W}{hen} Mary the Mother of Jesus was espoused to Joseph, before they came together, she was found with child of the Holy Ghost. Then Joseph her husband, being a just man, and not willing to make her a publick example, was minded to put her away privily. But while he thought on these things, behold, the angel of the Lord appeared unto him in a dream, saying, Joseph, thou son of David, fear not to take unto thee Mary thy wife: for that which is conceived in her is of the Holy Ghost. And she shall bring forth a son, and thou shalt call his name \divineName{Jesus}: for he shall save his people from their sins.
\offertory{My truth and my mercy shall be with him: and in my name shall his horn be exalted.}
\secret
\lett{W}{e} render unto thee, O Lord, this duty of our bounden service, humbly beseeching thee: that like as we do offer unto thee this sacrifice of praise in honour of the venerable festivity of blessed Joseph, the Spouse of the Mother of thy Son Jesus Christ our Lord, so by his prayers thou wouldest preserve thy gifts within us. Through the same.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\communion{Joseph, thou son of David, fear not to take unto thee Mary thy wife: for that which is conceived in her is of the Holy Ghost.}
\postcommunion
\lett{A}{ssist} us, we beseech thee, O merciful God: and at the intercession of blessed Joseph, thy Confessor on our behalf, continue towards us the gifts of thy loving-kindness. Through.
%RUBRIC REWORDED/COMBINED FOR SPACE:
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
\subby{St. Benedict}
\feastday{St. Benedict}
\begin{inhead}
    {Second Class Double\\
21 March}
\end{inhead}
\feastdate{21}{March}
%Sarum-Roman: 0.3

\introit
%Sarum \& Roman:
\lett{T}{he} mouth of the righteous is exercised in wisdom: and his tongue will be talking of judgment. The law of his God is in his heart. \textit{Ps.} Fret not thyself because of the ungodly: neither be thou envious against the evil doers.
\collect
\lett{A}{lmighty} and everlasting God, who on this day didst release thy most blessed Confessor Benedict from the bondage of the flesh and take him up to heaven, grant, we beseech thee, to us thy servants who celebrate this feast pardon of all our sins, that we who with joyful hearts take pleasure in his renown may for his sake, and at his intercession, have fellowship with him. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 39:5}
%\lett{H}{e} will give his heart to resort early to the Lord that made him, and will pray before the most High, and will open his mouth in prayer, and make supplication for his sins. When the great Lord will, he shall be filled with the spirit of understanding: he shall pour out wise sentences, and give thanks unto the Lord in his prayer. He shall direct his counsel and knowledge, and in his secrets shall he meditate. He shall shew forth that which he hath learned, and shall glory in the law of the covenant of the Lord. Many shall commend his understanding; and so long as the world endureth, it shall not be blotted out; his memorial shall not depart away, and his name shall live from generation to generation.
%RV:
\lett{H}{e} will apply his heart to resort early to the Lord that made him, And will make supplication before the Most High, And will open his mouth in prayer, And will make supplication for his sins. If the great Lord will, He shall be filled with the spirit of understanding: He shall pour forth the words of his wisdom, And in prayer give thanks unto the Lord. He shall direct his counsel and knowledge, And in his secrets shall he meditate. He shall shew forth the instruction which he hath been taught, And shall glory in the law of the covenant of the Lord. Many shall commend his understanding; And so long as the world endureth, it shall not be blotted out: His memorial shall not depart, And his name shall live from generation to generation. 
%TractL Sarum (LXX from Comm. Martyrs):
\gradtr{The mouth of the righteous is exercised in wisdom: and his tongue will be talking of judgment. ℣. The law of his God is in his heart, and his goings shall not slide.}{They that sow in tears: shall reap in joy. ℣. He that now goeth on his way weeping, and beareth forth good seed, ℣. Shall doubtless come again with joy, and bring his sheaves with him.}

%Sarum Sequence:
\subsubsec{Sequence}
%Sequence I:
\begin{multicols}{2}
\lett{T}{he} festive day is come\par
\secondline{Which brings to us great joy,\\}

When Holy Church gives thanks acceptable to God.

This day the Heavenly Host sings, Glory in the highest,

With voice and symphony melodious.

This day the holy armies of the heavens

Unite with us in praise of God our King,\\

Christ, of the holy Virgin Mary born.\\

Hail, great Confessor Benedict, thou flower of saints

With stedfast faith, Christ's footsteps following,

Thou hast attained the everlasting Kingdom;

Already to their courts the blessed welcome thee.

To thee unitedly our suit we make,

Our frail life by thy blest protection aid.\\

Praise, glory to thee, blessed, holy Benedict

Entreat for us, O blessed, holy Benedict,\\

That in the heavenly Temple we may stand,

Fulness of joy beholding in thy presence.

Praise, honour, and thanksgiving pure

Be unto God, who doth for evermore endure.\\

Let all things say, Amen.
\end{multicols}
\readingcitation{Gospel}{Luke 11:33}
\lett{A}{t that time:} Jesus said to his disciples: No man, when he hath lighted a candle, putteth it in a secret place, neither under a bushel, but on a candlestick, that they which come in may see the light. The light of the body is the eye: therefore when thine eye is single, thy whole body also is full of light; but when thine eye is evil, thy body also is full of darkness. Take heed therefore that the light which is in thee be not darkness. If thy whole body therefore be full of light, having no part dark, the whole shall be full of light, as when the bright shining of a candle doth give thee light.
%Roman \& Sarum:
\offertory{Thou hast given him his heart's desire, O Lord, and hast not denied him the request of his lips: Thou hast set a crown of pure gold upon his head.}
\secret
\lett{O}{Lord,} let the gifts offered in honour of Saint Benedict, thy Confessor and Abbot, be pleasing unto thee; and, at his intercession, grant us pardon of all our sins. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\communion{Blessed is that servant whom his lord, when he cometh, shall find watching.}
\postcommunion
\lett{L}{et} the receiving of thy Sacrament, O Lord, at the intercession of Saint Benedict, thy Confessor and Abbot, lead us to follow the example of his conversation, and also to attain unto his reward. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}


%MANUAL ADJUSTMENT:
%\subby{St. Gabriel}
%\feastday{St. Gabriel}
%\begin{inhead}
%    {Greater Double\\
%24 March}
%\end{inhead}
%\feastdate{24}{March}
%\par\noindent
%AOB:
%\textit{Opening Sentence.} I saw another angel come down from heaven, having great power; and the earth was lightened with his glory.

%\introit
%\lett{O}{praise} the Lord, ye Angels of his, ye that excel in strength: ye that fulfil his commandment, and hearken unto the voice of his words. \textit{Ps.} Praise the Lord, O my soul: and all that is within me praise his holy name.

%\collect
%\lett{O}{God,} who from the company of Angels didst choose the Archangel Gabriel to proclaim the mystery of thine Incarnation: mercifully grant; that we who celebrate his festival (commemoration) on earth, may perceive his advocacy in heaven. Who livest.

%\begin{rubric}
%	Commemoration is made of the Feria.
%\end{rubric}

%\readingcitation{Epistle}{Daniel 9:21}
%\lett{I}{n those days:} Behold the man Gabriel, whom I had seen in the vision at the beginning, being caused to fly swiftly, touched me about the time of the evening oblation. And he informed me, and talked with me, and said, O Daniel, I am now come forth to give thee skill and understanding. At the beginning of thy supplications the commandment came forth, and I am come to shew thee; for thou art greatly beloved: therefore understand the matter, and consider the vision. Seventy weeks are determined upon thy people and upon thy holy city, to finish the transgression, and to make an end of sins, and to make reconciliation for iniquity, and to bring in everlasting righteousness, and to seal up the vision and prophecy, and to anoint the most Holy. Know therefore and understand, that from the going forth of the commandment to restore and to build Jerusalem unto the Messiah the Prince shall be seven weeks, and threescore and two weeks: the street shall be built again, and the wall, even in troublous times. And after threescore and two weeks shall Messiah be cut off, but not for himself: and the people of the prince that shall come shall destroy the city and the sanctuary; and the end thereof shall be with a flood, and unto the end of the war desolations are determined.

%\gradtr{O praise the Lord, ye Angels of his, ye that excel in strength, ye that fulfil his commandment. ℣. Praise the Lord, O my soul, and all that is within me praise his holy name.}{Hail, Mary, full of grace; the Lord is with thee. ℣. Blessed art thou among women: and blessed is the fruit of thy womb. ℣. Behold, thou shalt conceive, and bring forth a Son, and shalt call his name Emmanuel. ℣. The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee. ℣. Therefore also that Holy Thing which shall be born of thee, shall be called the Son of God.}

%\begin{rubric}
%	In Votive Masses before Septuagesima or after Pentecost, the Gradual as above, but in place of the Tract is said:
%\end{rubric}
%\alleluia{Alleluia, alleluia. ℣. O praise the Lord, all ye his hosts: ye servants of his that do his pleasure. Alleluia.}

%\begin{rubric}
%	In Eastertide, replacing the Gradual \& Tract:
%\end{rubric}

%\alleluia{Alleluia, alleluia. ℣. Who maketh his Angels spirits: and his ministers a flaming fire. Alleluia. ℣. Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women. Alleluia.}

%\readingcitation{Gospel}{Luke 1:26}
%\lett{A}{t that time:} The Angel Gabriel was sent from God unto a city of Galilee named Nazareth, to a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin's name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name \divineName{Jesus}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: and he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word.

%\offertory{An Angel stood at the altar of the temple, having a golden censer in his hand; and there was given unto him much incense: and the smoke of the incense ascended up before God.}

%\secret
%\lett{M}{ay} the offering of our service, and the prayer of blessed Gabriel the Archangel, be accepted in thy sight, O Lord: that he, who is venerated by us on earth, may be an advocate for us with thee in heaven. Through.
%\begin{rubric}
%	Commemoration of the Feria.
%\end{rubric}

%\communion{O all ye Angels of the Lord, bless ye the Lord: sing ye praises, and magnify him above all for ever.}

%\postcommunion
%\lett{H}{aving} received the mysteries of thy Body and Blood, we entreat thy loving kindness, O Lord our God: that, as we have known thine Incarnation by the message of Gabriel, so we may through his help obtain the benefits of that Incarnation. Who livest.

%\begin{rubric}
%	Commemoration \& Last Gospel of the Feria.
%\end{rubric}


\subby{Annunciation of the Blessed Virgin Mary}
\feastday{Annunciation}
\begin{inhead}
    {First Class Double\\
25 March}
\end{inhead}
\feastdate{25}{March}
%S-R: N/A
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} Send, O Lord, the Lamb, the Ruler of the land. From the rock of the wilderness unto the mountain of the daughter of Sion.%\vr{Is. 16:1}

%\begin{rubric}
%    Today in Choir after Terce is said a Conventual Mass of the Feast, and outside Choir after None is read a Mass of the Feria.
%\end{rubric}

\introit
\lett{A}{ll} the rich among the people shall make their supplication before thee: the Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy and gladness. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
%1928 BCP:
%\lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts; that, as we have known the incarnation of thy Son Jesus Christ by the message of an Angel, so by his cross and passion we may be brought unto the glory of his resurrection. Through the same.
%Sarum:
\lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the Blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we who believe her to be truly the Mother of God may be aided by her intercession in thy sight. Through the same.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\readingcitation{Epistle}{Isaiah 7:10}
\lett{I}{n those days:} The \divineName{Lord} spake again unto Ahaz, saying, Ask thee a sign of the \divineName{Lord} thy God; ask it either in the depth, or in the height above. But Ahaz said, I will not ask, neither will I tempt the \divineName{Lord}. And he said, Hear ye now, O house of David; Is it a small thing for you to weary men, but will ye weary my God also? Therefore the Lord himself shall give you a sign; Behold, a virgin shall conceive, and bear a son, and shall call his name Immanuel. Butter and honey shall he eat, that he may know to refuse the evil, and choose the good.
\gradtr{Full of grace are thy lips: because God hath blessed thee for ever. ℣. Because of the word of truth, of meekness, and righteousness: and thy right hand shall teach thee terrible things.}{Hearken, O daughter, and consider, incline thine ear: so shall the King have pleasure in thy beauty. ℣. The rich among the people shall make their supplication before thee: kings' daughters were among thy honourable women. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee. ℣. With joy and gladness shall they be brought: and shall enter into the King's palace.}
\readingcitation{Gospel}{Luke 1:26}
\lett{A}{t that time:} In the sixth month the angel Gabriel was sent from God unto a city of Galilee, named Nazareth, To a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin's name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name \divineName{Jesus}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: And he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word. And the angel departed from her.
\offertory{Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women, and blessed is the fruit of thy womb.}
\secret
\lett{S}{tablish} in our hearts, we beseech thee, O Lord, the mysteries of the true faith: that like as we acknowledge thy Son conceived of a Virgin to be very God and very man; so by the power of his life giving Resurrection, we may be found worthy to attain unto everlasting gladness. Through the same.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\communion{Behold, a Virgin shall conceive, and bear a son: and shall call his name Immanuel.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts; that, as we have known the incarnation of thy Son Jesus Christ by the message of an Angel, so by his cross and passion we may be brought unto the glory of his resurrection. Through the same.
%\lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the Blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we who believe her to be truly the Mother of God may be aided by her intercession in thy sight. Through the same.
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}

\begin{rubric}
    If any Feast fall during Holy Week or the Easter Octave, it is transferred to the next available day.
\end{rubric}


\subby{St. Tikhon of Moscow}
\feastday{St. Tikhon Moscow}
\begin{inhead}
    {First Class Double with a Common Octave\\
7 April}
\end{inhead}
\feastdate{7}{April}
%S-R: N/A

\introit
\lett{T}{he} Lord hath established a covenant of peace with him, and made him a prince: that he should have the dignity of the priesthood for ever (alleluia, alleluia). \textit{Ps.} Lord, remember David: and all his trouble.
\collect
%From Common of a Confessor Bishop I
\lett{G}{rant,} we beseech thee, almighty God: that the venerable solemnity of blessed Tikhon, thy Confessor and Bishop may increase our devotion and set forward our salvation. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 44:16}
%This proper presents a unique problem. This Scripture is not just different between the LXX and the Vulgate, since both differ from this text. Instead, a sort of summarised or traditional text inspired by the original has been placed here, with the citation of the Scripture from which it was inspired. Therefore, since a unique traditional text is provided in the missal (such as when the Comma Johanneum or the Pericope Adulterae is provided in the missal), it is simply placed here as-is, as translated by Knott.
\lett{B}{ehold,} a great priest, who in his days pleased God, and was found just; in the time of wrath he was made a means of reconciliation. There was none like unto him who kept the law of the Most High. Therefore he assured him by an oath that he would increase him among his people. He established him with the blessing of all men and the covenant and made it rest upon his head. He acknowledged him in his blessing: he preserved for him his mercy: so that he found favour in the sight of the Lord. He made him glorious in the sight of kings, and gave him a crown of glory. An everlasting covenant he made with him, and gave him a great priesthood, and clothed him with a robe of glory. To execute the office of the priesthood, and bless the people in his name, and to offer him a worthy incense and a sweet savour.

\gradtr{Behold, a great priest, who in his days pleased God. ℣. There was none found like unto him who kept the law of the Most High.}{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Gradual \& Tract:}
\end{rubric}\par\noindent

\alleluia{Alleluia, alleluia. ℣. Thou art a priest for ever after the order of Melchisedech. Alleluia. ℣. This is a priest whom the Lord hath crowned. Alleluia.}

\readingcitation{Gospel}{Matthew 25:14}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: A man travelling into a far country, called his own servants, and delivered unto them his goods. And unto one he gave five talents, to another two, and to another one; to every man according to his several ability; and straightway took his journey. Then he that had received the five talents went and traded with the same, and made them other five talents. And likewise he that had received two, he also gained other two. But he that had received one went and digged in the earth, and hid his lord's money. After a long time the lord of those servants cometh, and reckoneth with them. And so he that had received five talents came and brought other five talents, saying, Lord, thou deliveredst unto me five talents: behold, I have gained beside them five talents more. His lord said unto him, Well done, thou good and faithful servant: thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord. He also that had received two talents came and said, Lord, thou deliveredst unto me two talents: behold, I have gained two other talents beside them. His lord said unto him, Well done, good and faithful servant; thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord.
\offertory{I have found David my servant, with my holy oil have I anointed them; my hand shall hold him fast, and my arm shall strengthen him (alleluia).}
\secret
\lett{W}{e} beseech thee, O Lord, that we, remembering with gladness the merits of thy Saints, may in all places feel the succour of their intercession. Through.

\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}

\communion{A faithful and wise steward, whom the lord hath made ruler over his household: to give them their portion of meat in due season (alleluia).}
\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that we, shewing forth our thankfulness for the gifts which we have received, may, at the intercession of blessed Tikhon, thy Confessor and Bishop, obtain yet more abundant mercies. Through.

\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}


\subby{St. George}
\feastday{St. George}
\begin{inhead}
    {First Class Double with a Common Octave\\
23 April}
\end{inhead}
\feastdate{23}{April}
%Sarum-Roman: 0.2

\properantiphonfix

\introit
%The proper introit is identical to the Common of One Martyr in Eastertide, probably because this feast would always be in Eastertide in the Gregorian Kalendar, or would fall in the Triduum, so would be transferred into Easter.
%Because of our kalendar situation in the AWRV, I have provided Lenten propers from the Common (III Martyr not a Bishop) where the Proper of Saints is impossible (such as when Alleluia is used).
\ttyie{\centerline{\textit{Lent}}}{\centerline{\textit{Eastertide}}}
\ttyie{\lett{T}{he} righteous shall rejoice in thy strength, O Lord: exceeding glad shall he be of thy salvation: thou hast given him his heart's desire. \textit{Ps.} For thou hast prevented him with the blessings of goodness: and hast set a crown of pure gold upon his head.}{\lett{T}{hou} hast hidden me, O God, from the gathering together of the froward, alleluia: and from the insurrection of wicked doers, alleluia, alleluia. \textit{Ps.} Hear my voice, O God, in my prayer: preserve my life from fear of the enemy.}
\collect
\lett{O}{God,} who causest us to rejoice in the good deeds and intercession of Saint George, thy Martyr, mercifully grant that, by the gift of thy grace, we may obtain the benefits we ask of him. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\readingcitation{Epistle}{James 1:2}
\lett{B}{eloved:} Count it all joy when ye fall into divers temptations; Knowing this, that the trying of your faith worketh patience. But let patience have her perfect work, that ye may be perfect and entire, wanting nothing. If any of you lack wisdom, let him ask of God, that giveth to all men liberally, and upbraideth not; and it shall be given him. But let him ask in faith, nothing wavering. For he that wavereth is like a wave of the sea driven with the wind and tossed. For let not that man think that he shall receive any thing of the Lord. A double minded man is unstable in all his ways. Let the brother of low degree rejoice in that he is exalted: But the rich, in that he is made low: because as the flower of the grass he shall pass away. For the sun is no sooner risen with a burning heat, but it withereth the grass, and the flower thereof falleth, and the grace of the fashion of it perisheth: so also shall the rich man fade away in his ways. Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life, which the Lord hath promised to them that love him.

\gradtr{The righteous shall not be cast away: for the Lord upholdeth him with his hand. ℣. He is ever merciful, and lendeth: and his seed is blessed.}{Thou hast given him his heart's desire, and hast not denied him the request of his lips. ℣. For thou shalt prevent him with the blessing of goodness. ℣. Thou shalt set a crown of pure gold upon his head.}

\begin{rubric}
{In Eastertide, replacing the Gradual \& Tract:}
\end{rubric}\par\noindent

%Check Sarum Alleluia
\alleluia{Alleluia, alleluia. ℣. Blessed is the man that feareth the Lord: he hath great delight in his commandments. Alleluia.}

\subsubsec{Sequence}
\begin{multicols}{2}
\lett{N}{ow} let us sing, with instruments well tuned,\par
\secondline{The feast of George, in divers kinds of voice,}
\thirdline{Praying our due thanksgivings to the Lord,}
Much to be had in reverence of His saints,

Whom, with high gifts and virtues manifold,

He doth adorn and richly beautify.

In them, as though in instruments of music,

Faith doth with her own finger touch the strings,

Discoursing high of virtues excellent---

As on each single string she lays her hand

She blends it in the fourfold melody

Which she, that mother of all grace, evokes;

Composing thus harmonious symphony;

Without whom all is dissonant,

Yea, trifling, poor, and vain;

With whom all is unisonant,

Yea fraught with endless gain;

Aided by whom the just, in holy lives

Seeking to climb the heights of starry heaven,

Sing forth new songs, and tune their harps in gladness.

May we who keep their feast be counted meet

To hold with them in heaven communion sweet.
\end{multicols}

\readingcitation{Gospel}{John 15:1}
\lett{A}{t that time:} Jesus said to his disciples: I am the true vine, and my Father is the husbandman. Every branch in me that beareth not fruit he taketh away: and every branch that beareth fruit, he purgeth it, that it may bring forth more fruit. Now ye are clean through the word which I have spoken unto you. Abide in me, and I in you. As the branch cannot bear fruit of itself, except it abide in the vine; no more can ye, except ye abide in me. I am the vine, ye are the branches: He that abideth in me, and I in him, the same bringeth forth much fruit: for without me ye can do nothing. If a man abide not in me, he is cast forth as a branch, and is withered; and men gather them, and cast them into the fire, and they are burned. If ye abide in me, and my words abide in you, ye shall ask what ye will, and it shall be done unto you.
\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, (alleluia, alleluia.)}

\secret
\lett{W}{e} offer unto thee, O Lord, the wonted Sacrifice of the death of thy Martyr, Saint George, entreating of thy mercy that through these holy Mysteries we may, in thy victory, overcome the temptations of the old enemy, and of thy bounty obtain an everlasting recompense of reward. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}

\communion{The righteous shall rejoice in the Lord, and put his trust in him : and all they that are true of heart shall be glad, (alleluia, alleluia.)}

\postcommunion
\lett{W}{e} humbly pray thee, almighty Father, that we who are satisfied with the sweetness of the heavenly Table, may at the intercession of thy Martyr, Saint George, also be partakers of his resurrection by whose death we are redeemed. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
\subby{St. Mark}
\feastday{St. Mark}
\begin{inhead}
    {Second Class Double\\
25 April}
\end{inhead}
\feastdate{25}{April}
%S-R: N/A
%Latest Date of Gregorian Easter

\properantiphonfix

\introit
\lett{T}{hou} hast hidden me, O God, from the gathering together of the froward, (alleluia): and from the insurrection of wicked doers, (alleluia, alleluia). \textit{Ps.} Hear my voice, O God, in my prayer: preserve my life from fear of the enemy.
\collect
%1928 BCP:
%\lett{O}{almighty} God, who hast instructed thy holy Church with the heavenly doctrine of thy Evangelist Saint Mark; Give us grace that, being not like children carried away with every blast of vain doctrine, we may be established in the truth of thy holy Gospel. Through.
%Sarum:
\lett{O}{God,} who hast exalted blessed Mark thine Evangelist by the grace of the preaching of the Gospel: grant, we beseech thee; that we may ever both profit by his learning, and be defended by his prayer. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}

\readingcitation{Epistle}{Ephesians 4:7}
\lett{B}{rethren:} Unto every one of us is given grace according to the measure of the gift of Christ. Wherefore he saith, When he ascended up on high, he led captivity captive, and gave gifts unto men. (Now that he ascended, what is it but that he also descended first into the lower parts of the earth? He that descended is the same also that ascended up far above all heavens, that he might fill all things.) And he gave some, apostles; and some, prophets; and some, evangelists; and some, pastors and teachers; For the perfecting of the saints, for the work of the ministry, for the edifying of the body of Christ: Till we all come in the unity of the faith, and of the knowledge of the Son of God, unto a perfect man, unto the measure of the stature of the fulness of Christ: That we henceforth be no more children, tossed to and fro, and carried about with every wind of doctrine, by the sleight of men, and cunning craftiness, whereby they lie in wait to deceive; But speaking the truth in love, may grow up into him in all things, which is the head, even Christ: From whom the whole body fitly joined together and compacted by that which every joint supplieth, according to the effectual working in the measure of every part, maketh increase of the body unto the edifying of itself in love.

%MANUAL ADJUSTMENT:
\gradual{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}

\tract{They that sow in tears, shall reap in joy. ℣. He that now goeth on his way weeping, and beareth forth good seed. ℣. Shall doubtless come again with joy, and bring his sheaves with him.}

\begin{rubric}
{In Eastertide, replacing the Gradual \& Tract:}
\end{rubric}\par\noindent

\alleluia{Alleluia, alleluia. ℣. O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints. Alleluia. ℣. Thou hast set, O Lord, a crown of pure gold upon his head. Alleluia.}

\readingcitation{Gospel}{John 15:1}
\lett{A}{t that time:} Jesus said unto his disciples: I am the true vine, and my Father is the husbandman. Every branch in me that beareth not fruit he taketh away: and every branch that beareth fruit, he purgeth it, that it may bring forth more fruit. Now ye are clean through the word which I have spoken unto you. Abide in me, and I in you. As the branch cannot bear fruit of itself, except it abide in the vine; no more can ye, except ye abide in me. I am the vine, ye are the branches: He that abideth in me, and I in him, the same bringeth forth much fruit: for without me ye can do nothing. If a man abide not in me, he is cast forth as a branch, and is withered; and men gather them, and cast them into the fire, and they are burned. If ye abide in me, and my words abide in you, ye shall ask what ye will, and it shall be done unto you. Herein is my Father glorified, that ye bear much fruit; so shall ye be my disciples. As the Father hath loved me, so have I loved you: continue ye in my love. If ye keep my commandments, ye shall abide in my love; even as I have kept my Father's commandments, and abide in his love. These things have I spoken unto you, that my joy might remain in you, and that your joy might be full.

\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, (alleluia, alleluia.)}
\secret
\lett{W}{e} offer unto thee our gifts on the solemnity of blessed Mark, thine Evangelist, beseeching thee, O Lord: that as the preaching of the Gospel made him glorious; so his intercession may render us both in word and deed acceptable unto thee. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.% in non-conventual Masses.
\end{rubric}
\communion{The righteous shall rejoice in the Lord, and put his trust in him: and all they that are true of heart shall be glad, (alleluia, alleluia.)}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that thy holy mysteries may give unto us continual protection: and through the prayers of blessed Mark thine Evangelist ever defend us from all adversities. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria \& Last Gospel is of the Feria.
\end{rubric}


\subby{Sts. Philip \& James}
\feastday{Sts. Philip \& James}
\begin{inhead}
    {Second Class Double\\
1 May}
\end{inhead}
\feastdate{1}{May}
%This Feast will *always* occur during Easter. At the latest Julian Easter, it would fall on Palm Sunday and be transferred to after the Easter Octave.
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} He was known of them in breaking of bread. Alleluia.\par

\introit
\lett{T}{hey} cried unto thee, O Lord, in the time of their trouble: and thou didst hear them from heaven, alleluia, alleluia. \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.
\collect
\lett{O}{almighty} God, whom truly to know is everlasting life; Grant us perfectly to know thy Son Jesus Christ to be the way, the truth, and the life; that, following the steps of thy holy Apostles, Saint Philip and Saint James, we may stedfastly walk in the way that leadeth to eternal life through the same thy Son Jesus Christ our Lord. Who liveth.

\readingcitation{Epistle}{James 1:1}
\lett{J}{ames,} a servant of God and of the Lord Jesus Christ, to the twelve tribes which are scattered abroad, greeting. My brethren, count it all joy when ye fall into divers temptations; Knowing this, that the trying of your faith worketh patience. But let patience have her perfect work, that ye may be perfect and entire, wanting nothing. If any of you lack wisdom, let him ask of God, that giveth to all men liberally, and upbraideth not; and it shall be given him. But let him ask in faith, nothing wavering. For he that wavereth is like a wave of the sea driven with the wind and tossed. For let not that man think that he shall receive any thing of the Lord. A double minded man is unstable in all his ways. Let the brother of low degree rejoice in that he is exalted: But the rich, in that he is made low: because as the flower of the grass he shall pass away. For the sun is no sooner risen with a burning heat, but it withereth the grass, and the flower thereof falleth, and the grace of the fashion of it perisheth: so also shall the rich man fade away in his ways. Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life, which the Lord hath promised to them that love him.

\alleluia{Alleluia, alleluia. ℣. O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints. Alleluia. ℣. Have I been so long time with you, and yet hast thou not known me, Philip? He that hath seen me hath seen my Father also. Alleluia.}

\readingcitation{Gospel}{John 14:1}
\lett{A}{t that time:} Jesus said unto his disciples: Let not your heart be troubled: ye believe in God, believe also in me. In my Father's house are many mansions: if it were not so, I would have told you. I go to prepare a place for you. And if I go and prepare a place for you, I will come again, and receive you unto myself; that where I am, there ye may be also. And whither I go ye know, and the way ye know. Thomas saith unto him, Lord, we know not whither thou goest; and how can we know the way? Jesus saith unto him, I am the way, the truth, and the life: no man cometh unto the Father, but by me. If ye had known me, ye should have known my Father also: and from henceforth ye know him, and have seen him. Philip saith unto him, Lord, shew us the Father, and it sufficeth us. Jesus saith unto him, Have I been so long time with you, and yet hast thou not known me, Philip? he that hath seen me hath seen the Father; and how sayest thou then, Shew us the Father? Believest thou not that I am in the Father, and the Father in me? the words that I speak unto you I speak not of myself: but the Father that dwelleth in me, he doeth the works. Believe me that I am in the Father, and the Father in me: or else believe me for the very works' sake. Verily, verily, I say unto you, He that believeth on me, the works that I do shall he do also; and greater works than these shall he do; because I go unto my Father. And whatsoever ye shall ask in my name, that will I do, that the Father may be glorified in the Son. If ye shall ask any thing in my name, I will do it.
\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, alleluia, alleluia.}
\secret
\lett{G}{raciously} receive, O Lord, the gifts which we bring for the solemnity of thine Apostles Philip and James: and turn aside all the evils which we deserve. Through.
\communion{Have I been so long time with you, and yet hast thou not known me, Philip? He that hath seen me hath seen my Father also, alleluia: believest thou not that I am in the Father, and the Father in me? Alleluia, alleluia.}
\postcommunion
\lett{O}{Lord,} who hast fulfilled us with saving mysteries: we beseech thee that we may be aided by the prayers of those whose festival we celebrate. Through.


\subby{Invention of the Holy Cross}
\feastday{Invention of the Cross}
\begin{inhead}
    {Second Class Double\\
3 May}
\end{inhead}
\feastdate{3}{May}
%This Feast will *always* occur during Easter. At the latest Julian Easter, it would fall on Holy Tuesday and be transferred to after the Easter Octave.
%Sarum-Roman: ~0.7
\par\noindent
\textit{Opening Sentence.} Tell it out among the nations that the Lord reigneth from the tree.\par

\introit
\lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life and resurrection: by whom we are saved and set free, alleluia, alleluia. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.
\collect
\lett{O}{God,} who in the wondrous Finding of the Cross of salvation didst renew the miracles of thy Passion: vouchsafe; that by the ransom of the tree of life we may attain thy succour unto life eternal. Who livest.

\readingcitation{Epistle}{Philippians 2:5}
\lett{B}{rethren:} Let this mind be in you, which was also in Christ Jesus: Who, being in the form of God, thought it not robbery to be equal with God: But made himself of no reputation, and took upon him the form of a servant, and was made in the likeness of men: And being found in fashion as a man, he humbled himself, and became obedient unto death, even the death of the cross. Wherefore God also hath highly exalted him, and given him a name which is above every name: That at the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth; And that every tongue should confess that Jesus Christ is Lord, to the glory of God the Father.

\alleluia{Alleluia, alleluia. ℣. Tell it out among the heathen that the Lord hath reigned from the tree. Alleluia. ℣. Sweetest wood, and sweetest iron, sweetest weight is hung on thee: thou alone wast worthy to bear the King of heaven, and its Lord. Alleluia.}

\readingcitation{Gospel}{John 3:1}
\lett{A}{t that time:} There was a man of the Pharisees, named Nicodemus, a ruler of the Jews: The same came to Jesus by night, and said unto him, Rabbi, we know that thou art a teacher come from God: for no man can do these miracles that thou doest, except God be with him. Jesus answered and said unto him, Verily, verily, I say unto thee, Except a man be born again, he cannot see the kingdom of God. Nicodemus saith unto him, How can a man be born when he is old? can he enter the second time into his mother's womb, and be born? Jesus answered, Verily, verily, I say unto thee, Except a man be born of water and of the Spirit, he cannot enter into the kingdom of God. That which is born of the flesh is flesh; and that which is born of the Spirit is spirit. Marvel not that I said unto thee, Ye must be born again. The wind bloweth where it listeth, and thou hearest the sound thereof, but canst not tell whence it cometh, and whither it goeth: so is every one that is born of the Spirit. Nicodemus answered and said unto him, How can these things be? Jesus answered and said unto him, Art thou a master of Israel, and knowest not these things? Verily, verily, I say unto thee, We speak that we do know, and testify that we have seen; and ye receive not our witness. If I have told you earthly things, and ye believe not, how shall ye believe, if I tell you of heavenly things? And no man hath ascended up to heaven, but he that came down from heaven, even the Son of man which is in heaven. And as Moses lifted up the serpent in the wilderness, even so must the Son of man be lifted up: That whosoever believeth in him should not perish, but have eternal life.
\offertory{The right hand of the Lord bringeth mighty things to pass, the right hand of the Lord hath exalted me: I shall not die, but live and declare the works of the Lord, alleluia.}
\secret
\lett{M}{ercifully} give heed to the sacrifice which we offer unto thee, O Lord: that we, being preserved from all the wickedness of war, may through the standard of the holy Cross of thy Son ever be stablished with the protection of thy sure defence, that we may destroy the snares of the power of the enemy. Through the same.
\communion{By the sign of the Cross deliver us from our enemies, O our God, alleluia.}
\postcommunion
\lett{A}{lmighty} God, who hast fulfilled us with heavenly meat, and refreshed us with spiritual drink, we beseech thee: that as thou hast bidden us to triumph by the wood of the holy Cross of thy Son, the armour of righteousness for the salvation of the world, so thou wouldest defend us from the malicious enemy. Through the same.


\subby{St. John before the Latin Gate}
\feastday{St. John Latin Gate}
\begin{inhead}
    {Greater Double\\
6 May}
\end{inhead}
\feastdate{6}{May}

\introit
\lett{T}{hou} hast hidden me, O God, from the gathering together of the froward, alleluia: and from the insurrection of wicked doers, alleluia, alleluia. \textit{Ps.} Hear my voice, O God, in my prayer: preserve my life from fear of the enemy.

\collect
\lett{O}{God,} who seest that we are beset by evils on every side: grant, we beseech thee; that the glorious intercession of blessed John, thine Apostle and Evangelist, may protect us. Through.

\readingcitation{Epistle}{Wisdom 5:1}
\lett{T}{he} righteous man shall stand in great boldness before the face of such as have afflicted him, and made no account of his labours. When they see it, they shall be troubled with terrible fear, and shall be amazed at the strangeness of his salvation, so far beyond all that they looked for. And they repenting and groaning for anguish of spirit shall say within themselves, This was he, whom we had sometimes in derision, and a proverb of reproach: We fools accounted his life madness, and his end to be without honour: How is he numbered among the children of God, and his lot is among the saints!

\alleluia{Alleluia, alleluia. ℣. The righteous shall flourish like a palm tree: and shall spread abroad like a cedar in Libanus. Alleluia. ℣. The righteous shall grow as the lily: and shall flourish for ever before the Lord. Alleluia.}

\readingcitation{Gospel}{Matthew 20:20}
\lett{A}{t that time:} There came unto Jesus the mother of Zebedee's children with her sons, worshipping him, and desiring a certain thing of him. And he said unto her, What wilt thou? She saith unto him, Grant that these my two sons may sit, the one on thy right hand, and the other on the left, in thy kingdom. But Jesus answered and said, Ye know not what ye ask. Are ye able to drink of the cup that I shall drink of, and to be baptized with the baptism that I am baptized with? They say unto him, We are able. And he saith unto them, Ye shall drink indeed of my cup, and be baptized with the baptism that I am baptized with: but to sit on my right hand, and on my left, is not mine to give, but it shall be given to them for whom it is prepared of my Father.

\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, alleluia, alleluia.}

\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us whom thou dost cleanse by thy heavenly mysteries. Through.

\communion{The just shall rejoice in the Lord, and put his trust in him: and all they that are true of heart shall be glad, alleluia, alleluia.}

\postcommunion
\lett{W}{e} beseech thee, O Lord: that we, who have been refreshed with heavenly bread, may be nourished unto life eternal. Through.

%8 May is the latest day of Easter

%MANUAL ADJUSTMENT:
\clearpage

\subby{St. Barnabas}
\feastday{St. Barnabas}
\begin{inhead}
    {Greater Double\\
11 June}
\end{inhead}
\feastdate{11}{June}
%S-R: N/A

\properantiphonfix

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine uprising.

\collect
%1928 BCP:
%\lett{O}{Lord} God Almighty, who didst endue thy holy Apostle Barnabas with singular gifts of the Holy Ghost; Leave us not, we beseech thee, destitute of thy manifold gifts, nor yet of grace to use them alway to thy honour and glory. Through.
%Sarum:
\lett{W}{e} beseech thee, O Lord, let the prayers of thy Apostle, Saint Barnabas, commend thy Church to thee, and let him continue to intercede for her whom by his doctrine and death he doth glorify. Through.

\readingcitation{Epistle}{Acts 11:21}
\lett{I}{n those days:} Tidings of these things came unto the ears of the church which was in Jerusalem: and they sent forth Barnabas, that he should go as far as Antioch. Who, when he came, and had seen the grace of God, was glad, and exhorted them all, that with purpose of heart they would cleave unto the Lord. For he was a good man, and full of the Holy Ghost and of faith: and much people was added unto the Lord. Then departed Barnabas to Tarsus, for to seek Saul. And when he had found him, he brought him unto Antioch. And it came to pass that a whole year they assembled themselves with the church, and taught much people. And the disciples were called Christians first in Antioch. And in these days came prophets from Jerusalem unto Antioch. And there stood up one of them named Agabus, and signified by the spirit that there should be great dearth throughout all the world: which came to pass in the days of Claudius C{\ae}sar. Then the disciples, every man according to his ability, determined to send relief unto the brethren which dwelt in Jud{\ae}a: which also they did, and sent it to the elders by the hands of Barnabas and Saul.

\gradall{Their sound is gone out into all lands: and their words into the ends of the world. ℣. The heavens declare the glory of God: and the firmament sheweth his handy-work.}{Alleluia, alleluia. ℣. I have chosen you out of the world, that ye should go and bring forth fruit: and that your fruit should remain. Alleluia.}

\begin{rubric}
{In Eastertide, replacing the Gradual \& Alleluia:}
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. O Lord, the very heavens shall praise thy wondrous work: and thy truth in the congregation of the saints. Alleluia. ℣. Thou hast set, O Lord, a crown of pure gold upon his head. Alleluia.}

\readingcitation{Gospel}{John 15:12}
\lett{A}{t that time:} Jesus said unto his disciples: This is my commandment, That ye love one another, as I have loved you. Greater love hath no man than this, that a man lay down his life for his friends. Ye are my friends, if ye do whatsoever I command you. Henceforth I call you not servants; for the servant knoweth not what his lord doeth: but I have called you friends; for all things that I have heard of my Father I have made known unto you. Ye have not chosen me, but I have chosen you, and ordained you, that ye should go and bring forth fruit, and that your fruit should remain: that whatsoever ye shall ask of the Father in my name, he may give it you.

\offertory{Thou shalt make them princes in all lands: they shall remember thy name, O Lord, from one generation to another.}
\secret
\lett{S}{anctify,} O Lord, the gifts which we offer: and, at the intercession of blessed Barnabas, thine Apostle, cleanse us thereby from the defilements of our iniquities. Through.
\communion{Ye which have followed me shall sit upon thrones, judging the twelve tribes of Israel.}
\postcommunion
\lett{W}{e} humbly beseech thee, almighty God: that as thou dost refresh us with thy sacraments, so, at the intercession of blessed Barnabas, thine Apostle, thou wouldest vouchsafe unto us to do thee worthy and acceptable service. Through.


\subby{St. John Baptist}\label{StJohnBaptist}
\feastday{St. John Baptist}
\begin{inhead}
    {First Class Double with a Common Octave\\
24 June}
%(25 June is the last possible day of Eastertide)
\end{inhead}
\feastdate{24}{June}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} There was a man sent from God whose name was John.

\introit
\lett{F}{rom} the womb of my mother the Lord hath called me by my name: and hath made my mouth as it were a sharp sword: beneath the shadow of his hand hath he hid me, and hath made me like to a polished arrow. \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.
\collect
\label{JohnCollect}
\lett{A}{lmighty} God, by whose providence thy servant John Baptist was wonderfully born, and sent to prepare the way of thy Son our Saviour by preaching repentance; Make us so to follow his doctrine and holy life, that we may truly repent according to his preaching; and after his example constantly speak the truth, boldly rebuke vice, and patiently suffer for the truth's sake. Through the same.

\readingcitation{Epistle}{Isaiah 40:1}
\lett{C}{omfort} ye, comfort ye my people, saith your God. Speak ye comfortably to Jerusalem, and cry unto her, that her warfare is accomplished, that her iniquity is pardoned: for she hath received of the {\textsc{Lord's}} hand double for all her sins. The voice of him that crieth in the wilderness, Prepare ye the way of the \divineName{Lord}, make straight in the desert a highway for our God. Every valley shall be exalted, and every mountain and hill shall be made low: and the crooked shall be made straight, and the rough places plain: And the glory of the \divineName{Lord} shall be revealed, and all flesh shall see it together: for the mouth of the \divineName{Lord} hath spoken it. The voice said, Cry. And he said, What shall I cry? All flesh is grass, and all the goodliness thereof is as the flower of the field: The grass withereth, the flower fadeth: because the spirit of the \divineName{Lord} bloweth upon it: surely the people is grass. The grass withereth, the flower fadeth: but the word of our God shall stand for ever. O Zion, that bringest good tidings, get thee up into the high mountain; O Jerusalem, that bringest good tidings, lift up thy voice with strength; lift it up, be not afraid; say unto the cities of Judah, Behold your God! Behold, the Lord \divineName{God} will come with strong hand, and his arm shall rule for him: behold, his reward is with him, and his work before him. He shall feed his flock like a shepherd: he shall gather the lambs with his arm, and carry them in his bosom, and shall gently lead those that are with young.

\gradall{Before I formed thee in the belly I knew thee: and before thou camest forth out of the womb I sanctified thee. ℣. The Lord put forth his hand, and touched my mouth, and said unto me.}{Alleluia, alleluia. ℣. Thou, child, shalt be called the Prophet of the Highest: thou shalt go before the face of the Lord, to prepare his ways. Alleluia.}

\readingcitation{Gospel}{Luke 1:57}
\lett{E}{\smash{lisabeth's}} full time came that she should be delivered; and she brought forth a son. And her neighbours and her cousins heard how the Lord had shewed great mercy upon her; and they rejoiced with her. And it came to pass, that on the eighth day they came to circumcise the child; and they called him Zacharias, after the name of his father. And his mother answered and said, Not so; but he shall be called John. And they said unto her, There is none of thy kindred that is called by this name. And they made signs to his father, how he would have him called. And he asked for a writing table, and wrote, saying, His name is John. And they marvelled all. And his mouth was opened immediately, and his tongue loosed, and he spake, and praised God. And fear came on all that dwelt round about them: and all these sayings were noised abroad throughout all the hill country of Judaea. And all they that heard them laid them up in their hearts, saying, What manner of child shall this be! And the hand of the Lord was with him. And his father Zacharias was filled with the Holy Ghost, and prophesied, saying, Blessed be the Lord God of Israel; for he hath visited and redeemed his people, And hath raised up an horn of salvation for us in the house of his servant David; As he spake by the mouth of his holy prophets, which have been since the world began: That we should be saved from our enemies, and from the hand of all that hate us; To perform the mercy promised to our fathers, and to remember his holy covenant; The oath which he sware to our father Abraham, That he would grant unto us, that we being delivered out of the hand of our enemies might serve him without fear, In holiness and righteousness before him, all the days of our life. And thou, child, shalt be called the prophet of the Highest: for thou shalt go before the face of the Lord to prepare his ways; To give knowledge of salvation unto his people by the remission of their sins, Through the tender mercy of our God; whereby the dayspring from on high hath visited us, To give light to them that sit in darkness and in the shadow of death, to guide our feet into the way of peace. And the child grew, and waxed strong in spirit, and was in the deserts till the day of his shewing unto Israel.
\offertory{The righteous shall flourish like a palm tree: and shall spread abroad like a cedar in Libanus.}

\secret\label{JohnSecret}
\lett{W}{e} set upon thine altars, O Lord, these gifts: celebrating with due honour the nativity of him, who sang of the coming and proclaimed the presence of the Saviour of the world, Jesus Christ thy Son our Lord. Who liveth.
\communion{Thou, child, shalt be called the Prophet of the Highest: for thou shalt go before the face of the lord to prepare his ways.}
\postcommunion
\label{JohnPC}
\lett{L}{et} thy Church, O God, rejoice at the birth of blessed John Baptist: through whom she hath known the author of her new birth, Jesus Christ thy Son our Lord. Who liveth.
\begin{rubric}
    Within the Octave, Mass is said as on the Feast, with \nth{2} Collect of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{St. Peter}

\centerline{\small{(Sts. Peter \& Paul)}}

\feastday{St. Peter}
\begin{inhead}
    {First Class Double with a Common Octave\\
29 June}
\end{inhead}
\feastdate{29}{June}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} Thou art Peter and upon this rock I will build my Church.%\vr{Mt. 16:18}

\introit
\lett{N}{ow} I know of a surety that the Lord hath sent his Angel: and hath delivered me out of the hand of Herod, and from all the expectation of the people of the Jews. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine uprising.
\collect
\lett{O}{almighty} God, who by thy Son Jesus Christ didst give to thy Apostle Saint Peter many excellent gifts, and commandedst him earnestly to feed thy flock; Make, we beseech thee, all Bishops and Pastors diligently to preach thy holy Word, and the people obediently to follow the same, that they may receive the crown of everlasting glory. Through the same.

\readingcitation{Epistle}{Acts 12:1}
\lett{I}{n those days:} Herod the king stretched forth his hands to vex certain of the church. And he killed James the brother of John with the sword. And because he saw it pleased the Jews, he proceeded further to take Peter also. (Then were the days of unleavened bread.) And when he had apprehended him, he put him in prison, and delivered him to four quaternions of soldiers to keep him; intending after Easter to bring him forth to the people. Peter therefore was kept in prison: but prayer was made without ceasing of the church unto God for him. And when Herod would have brought him forth, the same night Peter was sleeping between two soldiers, bound with two chains: and the keepers before the door kept the prison. And, behold, the angel of the Lord came upon him, and a light shined in the prison: and he smote Peter on the side, and raised him up, saying, Arise up quickly. And his chains fell off from his hands. And the angel said unto him, Gird thyself, and bind on thy sandals. And so he did. And he saith unto him, Cast thy garment about thee, and follow me. And he went out, and followed him; and wist not that it was true which was done by the angel; but thought he saw a vision. When they were past the first and the second ward, they came unto the iron gate that leadeth unto the city; which opened to them of his own accord: and they went out, and passed on through one street; and forthwith the angel departed from him. And when Peter was come to himself, he said, Now I know of a surety, that the Lord hath sent his angel, and hath delivered me out of the hand of Herod, and from all the expectation of the people of the Jews.
\gradall{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}{Alleluia, alleluia. ℣. Thou art Peter, and upon this rock I will build my Church. Alleluia.}

\readingcitation{Gospel}{Matthew 16:13}
\lett{A}{t that time:} When Jesus came into the coasts of Caesarea Philippi, he asked his disciples, saying, Whom do men say that I the Son of man am? And they said, Some say that thou art John the Baptist: some, Elias; and others, Jeremias, or one of the prophets. He saith unto them, But whom say ye that I am? And Simon Peter answered and said, Thou art the Christ, the Son of the living God. And Jesus answered and said unto him, Blessed art thou, Simon Barjona: for flesh and blood hath not revealed it unto thee, but my Father which is in heaven. And I say also unto thee, That thou art Peter, and upon this rock I will build my church; and the gates of hell shall not prevail against it. And I will give unto thee the keys of the kingdom of heaven: and whatsoever thou shalt bind on earth shall be bound in heaven: and whatsoever thou shalt loose on earth shall be loosed in heaven.

\offertory{Thou shalt make them princes in all lands: they shall remember thy name, O Lord, from one generation to another.}

\secret
\lett{M}{ay} the prayers of thine Apostles, O Lord, accompany the oblations which we offer for consecration to thy name: and grant that we may thereby be cleansed and defended. Through.

\communion{Thou art Peter, and upon this rock I will build my Church.}

\postcommunion
\lett{O}{Lord,} who hast satisfied us with heavenly food: defend us by the intercession of thine Apostles against all adversity. Through.


\subby{Commemoration of St. Paul the Apostle}
\feastday{St. Paul}
\begin{inhead}
    {Greater Double\\
30 June}
\end{inhead}
\feastdate{30}{June}

\introit
\lett{I}{know} whom I have believed, and am persuaded that he is able to keep that which I have committed unto him against that day, a just judge. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine up-rising.

\collect
\lett{O}{God,} who by the preaching of the blessed Apostle Paul didst teach the multitude of the Gentiles: grant to us, we beseech thee; that we who celebrate his birthday (commemoration) may know him to be our advocate with thee. Through.

\begin{rubric}
	Commemoration of St. Peter is made before all other Commemorations.
\end{rubric}

\lett{O}{God,} who didst bestow upon thy blessed Apostle Peter the keys of the kingdom of heaven, and didst  appoint unto him the high priesthood of binding and loosing: vouchsafe; that by the help of his intercession we may be delivered from the bonds of our iniquities. (Who livest.)
\begin{rubric}
	Commemoration is also made of the Octave of St. John (p. \pageref{JohnCollect}).
\end{rubric}

\readingcitation{Epistle}{Galatians 1:11}
\lett{B}{rethren:} I certify you that the gospel which was preached of me is not after man. For I neither received it of man, neither was I taught it, but by the revelation of Jesus Christ. For ye have heard of my conversation in time past in the Jews' religion, how that beyond measure I persecuted the church of God, and wasted it: and profited in the Jews' religion above many my equals in mine own nation, being more exceedingly zealous of the traditions of my fathers. But when it pleased God, who separated me from my mother’s womb, and called me by his grace, to reveal his Son in me, that I might preach him among the heathen; immediately I conferred not with flesh and blood: neither went I up to Jerusalem to them which were apostles before me; but I went into Arabia, and returned again unto Damascus. Then after three years I went up to Jerusalem to see Peter, and abode with him fifteen days. But other of the apostles saw I none, save James the Lord’s brother. Now the things which I write unto you, behold, before God, I lie not.

\gradall{He that wrought effectually in Peter to the apostleship, the same was mighty in me toward the Gentiles: and they perceived the grace of God that was given unto me. ℣. The grace of God which was bestowed upon me was not in vain: but his grace ever abideth in me.}{Alleluia, alleluia. ℣. Holy Paul Apostle, preacher of truth and teacher of the Gentiles, intercede for us. Alleluia.}

\readingcitation{Gospel}{Matthew 10:16}
\lett{A}{t that time:} Jesus said unto his disciples: Behold, I send you forth as sheep in the midst of wolves: be ye therefore wise as serpents, and harmless as doves. But beware of men: for they will deliver you up to the councils, and they will scourge you in their synagogues; and ye shall be brought before governors and kings for my sake, for a testimony against them and the Gentiles. But when they deliver you up, take no thought how or what ye shall speak: for it shall be given you in that same hour what ye shall speak. For it is not ye that speak, but the Spirit of your Father which speaketh in you. And the brother shall deliver up the brother to death, and the father the child: and the children shall rise up against their parents, and cause them to be put to death. And ye shall be hated of all men for my name's sake: but he that endureth to the end shall be saved. 

\offertory{Right honourable are thy friends unto me, O God: right well is their princedom established.}

%MANUAL ADJUSTMENT:
\vspace{-2.5ex}

\secret
\lett{S}{anctify,} O Lord, through the prayers of thine Apostle Paul the gifts of thy people: that those things, which by thine institution are pleasing unto thee, may be made more pleasing by his prayer and advocacy. Through.

\lett{W}{e} beseech thee, O Lord, that the intercession of blessed Peter the Apostle may commend unto thee the prayers and sacrifices of thy Church: that those things which we celebrate for his glory may avail for our pardon. (Through.)

\begin{rubric}
	Commemoration is also made of the Octave of St. John (p. \pageref{JohnSecret}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-2.5ex}

\communion{Verily I say unto you: That ye which have forsaken all and followed me, shall receive an hundredfold, and shall inherit everlasting life.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\postcommunion
\lett{O}{Lord,} who hast made us partakers of thy sacrament: we entreat thee; that at the intercession of thy Blessed Apostle Paul, those things which we have celebrated for his glory may be profitable for our healing. Through.

\lett{M}{ay} the gift, O Lord, which we have offered make us to rejoice: that as we proclaim thy wonders in thine Apostle Peter; so through him we may receive the abundance of thy loving-kindness. (Through).

\begin{rubric}
	Commemoration is also made of the Octave of St. John (p. \pageref{JohnPC}).
\end{rubric}


\subby{The Most Precious Blood of Our Lord Jesus Christ}
\feastday{Precious Blood}
\begin{inhead}
    {Second Class Double\\
1 July}
\end{inhead}
\feastdate{1}{July}
%No Sarum
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
%Dominican (https://archive.org/details/BrevariumIuxtaRitumSacriOrdinisPraedicatorum/black-and-white%20version/Breviarium_Cormier_1909_prior_cropped%20%285.18x6.5%29%20pg%201-1577%20%28vol.%201%29/page/n743/mode/1up)
\textit{Opening Sentence.} We therefore pray thee, help thy servants, whom thou hast redeemed with thy precious blood.

\introit
\lett{T}{hou} hast redeemed us, O Lord, by thy blood, out of every kindred, and tongue, and people, and nation: and hast made us a kingdom for our God. \textit{Ps.} My song shall be alway of the loving-kindness of the Lord: with my mouth will I ever be shewing thy truth from one generation to another.
\collect
\lett{A}{lmighty} and everlasting God, who didst appoint thine only-begotten Son to be the Redeemer of the world, and hast vouchsafed to be reconciled to us by his Blood: grant us, we beseech thee, so to venerate with solemn worship the price of our salvation, that by its power we may be defended from the evils of this present life on earth; and may rejoice in the everlasting fruit thereof in heaven. Through.
\begin{rubric}
    Commemoration is made of the Octave Day of St. John Baptist (p. \pageref{JohnCollect}).
\end{rubric}
\readingcitation{Epistle}{Hebrews 9:11}
\lett{B}{rethren:} Christ being come an high priest of good things to come, by a greater and more perfect tabernacle, not made with hands, that is to say, not of this building; Neither by the blood of goats and calves, but by his own blood he entered in once into the holy place, having obtained eternal redemption for us. For if the blood of bulls and of goats, and the ashes of an heifer sprinkling the unclean, sanctifieth to the purifying of the flesh: How much more shall the blood of Christ, who through the eternal Spirit offered himself without spot to God, purge your conscience from dead works to serve the living God? And for this cause he is the mediator of the new testament, that by means of death, for the redemption of the transgressions that were under the first testament, they which are called might receive the promise of eternal inheritance, %Added by Missal
in Christ Jesus our Lord.

%MANUAL ADJUSTMENT:

\gradual{%The inauthenticity of this text in the original text of St. John does not exclude its liturgical use, as we occasionally use elements from the tradition or deuterocanon in the liturgy. It is certainly evidence of its later date.
This is he that came by water and blood, even Jesus Christ: not by water only, but by water and blood. ℣. There are three that bear record in heaven: the Father, the Word, and the Holy Ghost; and these three are one. And there are three that bear witness in earth: the spirit, and the water, and the blood: and these three agree in one.}

\alleluia{Alleluia, alleluia. ℣. If we receive the witness of men, the witness of God is greater. Alleluia.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\readingcitation{Gospel}{John 19:30}
\lett{A}{t that time:} When Jesus had received the vinegar, he said, It is finished: and he bowed his head, and gave up the ghost. The Jews therefore, because it was the preparation, that the bodies should not remain upon the cross on the sabbath day, (for that sabbath day was an high day,) besought Pilate that their legs might be broken, and that they might be taken away. Then came the soldiers, and brake the legs of the first, and of the other which was crucified with him. But when they came to Jesus, and saw that he was dead already, they brake not his legs: But one of the soldiers with a spear pierced his side, and forthwith came there out blood and water. And he that saw it bare record, and his record is true.

\offertory{The cup of blessing which we bless, is it not the communion of the blood of Christ? the bread which we break, is it not the communion of the body of Christ?}

%MANUAL ADJUSTMENT:
\vspace{-2.5ex}

\secret
\lett{W}{e} beseech thee, that through these divine mysteries we may draw near unto Jesus, the mediator of the new Testament: and renew upon thine altars, O Lord of Hosts, the sprinkling of the blood that speaketh better things than that of Abel. Through the same.

\begin{rubric}
    Commemoration is made of the Octave Day of St. John Baptist (p. \pageref{JohnSecret}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-2.5ex}

\communion{Christ was once offered to bear the sins of many: and unto them that look for him shall he appear the second time without sin unto salvation.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\postcommunion
\lett{O}{Lord,} who hast suffered us to approach thy holy table, and with joy to draw water out of the wells of the Saviour: we beseech thee, that his blood may be to us a well of water springing up unto life eternal. Who liveth.
\begin{rubric}
    Commemoration is made of the Octave Day of St. John Baptist (p. \pageref{JohnPC}).
\end{rubric}


\subby{Visitation of the Blessed Virgin Mary}
\feastday{Visitation}
\begin{inhead}
    {Second Class Double\\
2 July}
\end{inhead}
\feastdate{2}{July}

\introit
\lett{H}{ail,} O Mother most holy, who in child-birth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
\lett{W}{e} beseech thee, O Lord, to grant unto us thy servants the gift of heavenly grace: that as the child-bearing of the blessed Virgin was unto us the beginning of salvation; so the devout observance of her Visitation may bestow on us an increase of peace. Through.
\readingcitation{Epistle}{Song of Songs 2:8}
\lett{B}{ehold,} he cometh leaping upon the mountains, skipping upon the hills. My beloved is like a roe or a young hart: behold, he standeth behind our wall, he looketh forth at the windows, shewing himself through the lattice. My beloved spake, and said unto me, Rise up, my love, my fair one, and come away. For, lo, the winter is past, the rain is over and gone; The flowers appear on the earth; the time of the singing of birds is come, and the voice of the turtle is heard in our land; The fig tree putteth forth her green figs, and the vines with the tender grape give a good smell. Arise, my love, my fair one, and come away. O my dove, that art in the clefts of the rock, in the secret places of the stairs, let me see thy countenance, let me hear thy voice; for sweet is thy voice, and thy countenance is comely.
\gradall{Blessed and venerable art thou, O Virgin Mary: who without spot wast found the Mother of the Saviour. ℣. Virgin, Mother of God, he whom the whole world containeth not, being made man hid in thy womb.}{Alleluia, alleluia. ℣. Happy art thou, O sacred Virgin Mary, and most worthy of all praise: for our of thee hath arisen the sun of righteousness, Christ our God. Alleluia.}
\readingcitation{Gospel}{Luke 1:39}
\lett{A}{t that time:} Mary arose and went into the hill country with haste, into a city of Juda; And entered into the house of Zacharias, and saluted Elisabeth. And it came to pass, that, when Elisabeth heard the salutation of Mary, the babe leaped in her womb; and Elisabeth was filled with the Holy Ghost: And she spake out with a loud voice, and said, Blessed art thou among women, and blessed is the fruit of thy womb. And whence is this to me, that the mother of my Lord should come to me? For, lo, as soon as the voice of thy salutation sounded in mine ears, the babe leaped in my womb for joy. And blessed is she that believed: for there shall be a performance of those things which were told her from the Lord. And Mary said, My soul doth magnify the Lord, And my spirit hath rejoiced in God my Saviour.
\offertory{Blessed art thou, O Virgin Mary, who didst bear the Creator of all things: thou broughtest forth him who made thee, and for ever remainest a Virgin.}
\secret
\lett{M}{ay} the manhood of thine only-begotten Son, O Lord, avail for our succour: that even as he, being born of a Virgin, destroyed not but hallowed the innocence of his Mother; so on this feast of her Visitation, he may deliver us from our offences, and render us an oblation acceptable unto thee, even Jesus Christ our Lord. Who liveth.
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord: that the sacrament which we have received in the observance of this yearly festival; may both in this life and in that which is to come be profitable for our healing. Through.


%MANUAL ADJUSTMENT:
\clearpage
%The Feast of St. Mary Magdalene is very dear to the English tradition, so much so that the 1549 BCP included it. Therefore, it is supplied here.
\subby{St. Mary Magdalene, Penitent}
\feastday{St. Mary Magdalene}
\begin{inhead}
    {Greater Double\\
22 July}
\end{inhead}
\feastdate{22}{July}
\par\noindent
\textit{Opening Sentence.} Mary hath chosen that good part, which shall not be taken away from her.%\vr{Lk. 10:42}

\introit
\lett{T}{he} ungodly laid wait for me to destroy me: O Lord, I will consider thy testimonies: I see that all things come to an end: but thy commandment is exceeding broad. \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.
\collect
%1549 BCP:
%\lett{M}{erciful} Father, give us grace, that we never presume to sin through the example of any creature; but if it shall chance us at any time to offend thy divine majesty; that then we may truly repent, and lament the same, after the example of Mary Magdalene, and by lively faith obtain remission of all our sins; through the only merits of thy Son our Saviour Christ. Who liveth.
%Sarum:
\lett{G}{rant} us, most merciful Father, like as St. Mary Magdalene, by loving thy Only Begotten Son above all things, obtained forgiveness of her sins, so she may procure for us in thy compassionate presence everlasting blessedness. Through.

\readingcitation{Epistle}{Proverbs 31:10}
\lett{W}{ho} can find a virtuous woman? for her price is far above rubies. The heart of her husband doth safely trust in her, so that he shall have no need of spoil. She will do him good and not evil all the days of her life. She seeketh wool, and flax, and worketh willingly with her hands. She is like the merchants' ships; she bringeth her food from afar. She riseth also while it is yet night, and giveth meat to her household, and a portion to her maidens. She considereth a field, and buyeth it: with the fruit of her hands she planteth a vineyard. She girdeth her loins with strength, and strengtheneth her arms. She perceiveth that her merchandise is good: her candle goeth not out by night. She layeth her hands to the spindle, and her hands hold the distaff. She stretcheth out her hand to the poor; yea, she reacheth forth her hands to the needy. She is not afraid of the snow for her household: for all her household are clothed with scarlet. She maketh herself coverings of tapestry; her clothing is silk and purple. Her husband is known in the gates, when he sitteth among the elders of the land. She maketh fine linen, and selleth it; and delivereth girdles unto the merchant. Strength and honour are her clothing; and she shall rejoice in time to come. She openeth her mouth with wisdom; and in her tongue is the law of kindness. She looketh well to the ways of her household, and eateth not the bread of idleness. Her children arise up, and call her blessed; her husband also, and he praiseth her. Many daughters have done virtuously, but thou excellest them all. Favour is deceitful, and beauty is vain: but a woman that feareth the \divineName{Lord}, she shall be praised. Give her of the fruit of her hands; and let her own works praise her in the gates.

\gradall{Thou hast loved righteousness and hated iniquity. ℣. Wherefore God, even thy God, hath anointed thee with the oil of gladness.}{Alleluia, alleluia. ℣. Full of grace are thy lips: because God hath blessed thee for ever. Alleluia.}

\readingcitation{Gospel}{Luke 7:36}
\lett{A}{t that time:} One of the Pharisees desired Jesus that he would eat with him. And he went into the Pharisee's house, and sat down to meat. And, behold, a woman in the city, which was a sinner, when she knew that Jesus sat at meat in the Pharisee's house, brought an alabaster box of ointment, And stood at his feet behind him weeping, and began to wash his feet with tears, and did wipe them with the hairs of her head, and kissed his feet, and anointed them with the ointment. Now when the Pharisee which had bidden him saw it, he spake within himself, saying, This man, if he were a prophet, would have known who and what manner of woman this is that toucheth him: for she is a sinner. And Jesus answering said unto him, Simon, I have somewhat to say unto thee. And he saith, Master, say on. There was a certain creditor which had two debtors: the one owed five hundred pence, and the other fifty. And when they had nothing to pay, he frankly forgave them both. Tell me therefore, which of them will love him most? Simon answered and said, I suppose that he, to whom he forgave most. And he said unto him, Thou hast rightly judged. And he turned to the woman, and said unto Simon, Seest thou this woman? I entered into thine house, thou gavest me no water for my feet: but she hath washed my feet with tears, and wiped them with the hairs of her head. Thou gavest me no kiss: but this woman since the time I came in hath not ceased to kiss my feet. My head with oil thou didst not anoint: but this woman hath anointed my feet with ointment. Wherefore I say unto thee, Her sins, which are many, are forgiven; for she loved much: but to whom little is forgiven, the same loveth little. And he said unto her, Thy sins are forgiven. And they that sat at meat with him began to say within themselves, Who is this that forgiveth sins also? And he said to the woman, Thy faith hath saved thee; go in peace.
\offertory{Kings' daughters were among thy honourable women, upon thy right hand did stand the queen in a vesture of gold, wrought about with divers colours.}

\secret
\lett{W}{e} beseech thee, O Lord, that the glorious merits of blessed Mary Magdalene may render our gifts pleasing unto thee: even as the oblation of her devout obedience was mercifully accepted by thine only-begotten Son. Who liveth.
\communion{I deal with the thing that is lawful and right, O Lord, let the proud do me no wrong: I hold straight all thy commandments, and all false ways I utterly abhor.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received the only and saving remedy, thy precious Body and Blood: may by the protection of Saint Mary Magdalene be delivered from all evils. Who livest.


\subby{St. James}
\feastday{St. James}
\begin{inhead}
    {Second Class Double\\
25 July}
\end{inhead}
\feastdate{25}{July}

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting and mine up-rising.

\collect
%1928 BCP:
%\lett{G}{rant,} O merciful God, that, as thine holy Apostle Saint James, leaving his father and all that he had, without delay was obedient unto the calling of thy Son Jesus Christ, and followed him; so we, forsaking all worldly and carnal affections, may be evermore ready to follow thy holy commandments. Through.
%Sarum:
\lett{B}{e} thou, O Lord, the sanctifier and guardian of thy people, that under the protection of thy Apostle, James, they may please thee in their conversation, and serve thee in all quietness. Through.

\readingcitation{Epistle}{Acts 11:27}
\lett{I}{n those days:} Came prophets from Jerusalem unto Antioch. And there stood up one of them named Agabus, and signified by the Spirit that there should be great dearth throughout all the world: which came to pass in the days of Claudius Caesar. Then the disciples, every man according to his ability, determined to send relief unto the brethren which dwelt in Judaea: Which also they did, and sent it to the elders by the hands of Barnabas and Saul. Now about that time Herod the king stretched forth his hands to vex certain of the church. And he killed James the brother of John with the sword. And because he saw it pleased the Jews, he proceeded further to take Peter also.

\gradual{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}

\alleluia{Allelluia, alleluia. ℣. I have chosen you out of the world, that ye should go and bring forth fruit: and that your fruit should remain. Alleluia.}

\readingcitation{Gospel}{Matthew 20:20}
\lett{A}{t that time:} There came unto Jesus the mother of Zebedee's children with her sons, worshipping him, and desiring a certain thing of him. And he said unto her, What wilt thou? She saith unto him, Grant that these my two sons may sit, the one on thy right hand, and the other on the left, in thy kingdom. But Jesus answered and said, Ye know not what ye ask. Are ye able to drink of the cup that I shall drink of, and to be baptized with the baptism that I am baptized with? They say unto him, We are able. And he saith unto them, Ye shall drink indeed of my cup, and be baptized with the baptism that I am baptized with: but to sit on my right hand, and on my left, is not mine to give, but it shall be given to them for whom it is prepared of my Father. And when the ten heard it, they were moved with indignation against the two brethren. But Jesus called them unto him, and said, Ye know that the princes of the Gentiles exercise dominion over them, and they that are great exercise authority upon them. But it shall not be so among you: but whosoever will be great among you, let him be your minister; And whosoever will be chief among you, let him be your servant: Even as the Son of man came not to be ministered unto, but to minister, and to give his life a ransom for many.
\offertory{Their sound is gone out into all lands: and their words into the ends of the world.}
\secret
\lett{W}{e} beseech thee, O Lord, that the blessed passion of blessed James thy Apostle may commend unto thee the oblations of thy people: that whereas they are not worthy through our merits, they may by his prayers be made acceptable unto thee. Through.
\communion{Ye, which have followed me, shall sit upon thrones, judging the twelve tribes of Israel.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who on this feast of thy blessed Apostle James, have joyfully received thy holy mysteries: may by his intercession obtain thy succour. Through.

\subby{St. Anne}
\feastday{St. Anne}
\begin{inhead}
    {Second Class Double\\
26 July}
\end{inhead}
\feastdate{26}{July}

\introit
\lett{R}{ejoice} we all in the Lord, keeping feast day in honour of blessed Anne: in whose solemnity the Angels rejoice and glorify the Son of God. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
\lett{O}{God,} who on blessed Anne didst vouchsafe to bestow grace, that she might be made worthy to become the mother of her who bore thine only-begotten Son: mercifully grant; that we, who celebrate her festival (commemoration) may be aided by her intercession with thee. Through the same.
\readingcitation{Epistle}{Proverbs 31:10}
\lett{W}{ho} can find a virtuous woman? for her price is far above rubies. The heart of her husband doth safely trust in her, so that he shall have no need of spoil. She will do him good and not evil all the days of her life. She seeketh wool, and flax, and worketh willingly with her hands. She is like the merchants’ ships; she bringeth her food from afar. She riseth also while it is yet night, and giveth meat to her household, and a portion to her maidens. She considereth a field, and buyeth it: with the fruit of her hands she planteth a vineyard. She girdeth her loins with strength, and strengtheneth her arms. She perceiveth that her merchandise is good: her candle goeth not out by night. She layeth her hands to the spindle, and her hands hold the distaff. She stretcheth out her hand to the poor; yea, she reacheth forth her hands to the needy. She is not afraid of the snow for her household: for all her household are clothed with scarlet. She maketh herself coverings of tapestry; her clothing is silk and purple. Her husband is known in the gates, when he sitteth among the elders of the land. She maketh fine linen, and selleth it; and delivereth girdles unto the merchant. Strength and honour are her clothing; and she shall rejoice in time to come. She openeth her mouth with wisdom; and in her tongue is the law of kindness. She looketh well to the ways of her household, and eateth not the bread of idleness. Her children arise up, and call her blessed; her husband also, and he praiseth her. Many daughters have done virtuously, but thou excellest them all. Favour is deceitful, and beauty is vain: but a woman that feareth the Lord, she shall be praised. Give her of the fruit of her hands; and let her own works praise her in the gates.
\gradall{Thou hast loved righteousness and hated iniquity. ℣. Wherefore God, even thy God, hath anointed thee with the oil of gladness.}{Alleluia, alleluia. ℣. Full of grace are thy lips: because God hath blessed thee for ever. Alleluia.}
\readingcitation{Gospel}{Matthew 13:44}
\lett{A}{t that time:} Jesus spoke this parable unto his disciples: The kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field. Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: who, when he had found one pearl of great price, went and sold all that he had, and bought it. Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, and shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth. Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord. Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.
\offertory{Kings' daughters were among thy honourable women, upon thy right hand did stand the queen in a vesture of gold, wrought about with divers colours.}
\secret
\lett{W}{e} beseech thee, O Lord, mercifully to have respect unto these our sacrifices: that through the intercession of blessed Anne, who brought forth the Mother of thy Son, our Lord Jesus Christ, they may increase our devotion and set forward our salvation. Through the same.
\communion{Full of grace are thy lips: because God hath blessed thee for ever, and world without end.}
\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that by the intercession of blessed Anne, whom thou didst choose to bring forth the Mother of thy Son, we, being quickened with these heavenly sacraments may be found worthy to attain to everlasting salvation. Through the same.

\subby{St. Peter in Chains}
\feastday{Lammas}
\begin{inhead}
    {Greater Double\\
1 August}
\end{inhead}
\feastdate{1}{August}
\par\noindent
%%From Sarum:
\textit{Opening Sentence.} I will call upon the Lord, which is worthy to be praised. So shall I be saved from mine enemies.%\vr{2 Sam. 22:4}

\introit
\lett{N}{ow} I know of a surety, that the Lord hath sent his Angel, and hath delivered me out of the hand of Herod, and from all the expectation of the people of the Jews. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting, and mine up-rising.

\collect
\lett{O}{God,} who didst deliver blessed Peter the Apostle from his chains, and didst cause him to depart unhurt: loose, we beseech thee, the chains of our sins; and graciously keep from us all evils. Through.
\needspace{4\baselineskip}
\lett{O}{God,} who by the preaching of the blessed Apostle Paul didst teach the multitude of the Gentiles: grant to us, we beseech thee; that we who celebrate his commemoration may know him to be our advocate with thee. (Through.)
\needspace{4\baselineskip}
\lett{O}{Lord,} let the crown of the brethren, thy Martyrs, cause us to rejoice: that we may thereby be strengthened and increased in our faith; and comforted by their manifold intercession. Through.

\readingcitation{Epistle}{Acts 12:1}
\lett{I}{n those days:} Herod the king stretched forth his hands to vex certain of the church. And he killed James the brother of John with the sword. And because he saw it pleased the Jews, he proceeded further to take Peter also. (Then were the days of unleavened bread.) And when he had apprehended him, he put him in prison, and delivered him to four quaternions of soldiers to keep him; intending after Easter to bring him forth to the people. Peter therefore was kept in prison: but prayer was made without ceasing of the church unto God for him.

And when Herod would have brought him forth, the same night Peter was sleeping between two soldiers, bound with two chains: and the keepers before the door kept the prison. And, behold, the angel of the Lord came upon him, and a light shined in the prison: and he smote Peter on the side, and raised him up, saying, Arise up quickly. And his chains fell off from his hands. And the angel said unto him, Gird thyself, and bind on thy sandals. And so he did. And he saith unto him, Cast thy garment about thee, and follow me. And he went out, and followed him; and wist not that it was true which was done by the angel; but thought he saw a vision. When they were past the first and the second ward, they came unto the iron gate that leadeth unto the city; which opened to them of his own accord: and they went out, and passed on through one street; and forthwith the angel departed from him. And when Peter was come to himself, he said, Now I know of a surety, that the Lord hath sent his angel, and hath delivered me out of the hand of Herod, and from all the expectation of the people of the Jews.

\gradall{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}{Alleluia, alleluia. ℣. Loosen at God's command, O Peter, the chains of earth's bondage: thou who dost make to lie open the heavenly realms to the blessed. Alleluia.}

\readingcitation{Gospel}{Matthew 16:13}
\lett{A}{t that time:} When Jesus came into the coasts of C{\ae}sarea Philippi, he asked his disciples, saying, Whom do men say that I the Son of man am? And they said, Some say that thou art John the Baptist: some, Elias; and others, Jeremias, or one of the prophets. He saith unto them, But whom say ye that I am? And Simon Peter answered and said, Thou art the Christ, the Son of the living God. And Jesus answered and said unto him, Blessed art thou, Simon Bar-jona: for flesh and blood hath not revealed it unto thee, but my Father which is in heaven. And I say also unto thee, That thou art Peter, and upon this rock I will build my church; and the gates of hell shall not prevail against it. And I will give unto thee the keys of the kingdom of heaven: and whatsoever thou shalt bind on earth shall be bound in heaven: and whatsoever thou shalt loose on earth shall be loosed in heaven.

\offertory{Thou shalt make them princes in all lands: they shall remember thy name, O Lord, from one generation to another.}

\secret
\lett{M}{ay} the sacrifice which we offer unto thee, O Lord, through the intercession of blessed Peter thine Apostle, evermore quicken and defend us. Through.
\needspace{4\baselineskip}
\lett{S}{anctify,} O Lord, through the prayers of thine Apostle Paul, the gifts of thy people: that those things which, by thine institution, are pleasing unto thee, may be made more pleasing by his prayer and advocacy. (Through.)
\needspace{4\baselineskip}
\lett{G}{rant,} O Lord, that we may with devout hearts celebrate thy mysteries in honour of thy holy Martyrs: and thereby obtain an increase both of protection and joy. Through.

\communion{Thou art Peter, and upon this rock I will build my Church.}

\postcommunion
\lett{O}{Lord} our God, who hast fulfilled us with the partaking of the sacred Body and the precious Blood, we beseech thee: that, those things which we perform with godly devotion, we may attain in the assurance of our redemption. Through the same.
\needspace{4\baselineskip}
\lett{O}{Lord,} who hast made us partakers of thy sacrament: we entreat thee; that at the intercession of thy holy Apostle Paul, those things which we have celebrated for his glory may be profitable for our healing. (Through.)
\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, almighty God: that growing in virtue we may follow the faith of them whose memory we recall by the partaking of this sacrament. Through.


\subby{Transfiguration of Our Lord Jesus Christ}
\feastday{Transfiguration}
\begin{inhead}
    {Second Class Double\\
6 August}
\end{inhead}
\feastdate{6}{August}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} Let us worship the Father, the Son, and the Holy Ghost, reigning in his Majesty.%\vr{Ps. 93:1}

\introit
\lett{C}{ome} and shew the light of thy countenance, O Lord, thou that sittest upon the cherubims, and we shall be whole. \textit{Ps.} Hear, O thou Shepherd of Israel, thou that leadest Joseph like a sheep.
\collect
%Sarum:
\lett{O}{God,} who on this day didst reveal from heaven thine only-begotten Son, transfigured in a wonderful manner, to the fathers of both testaments, grant unto us, we beseech thee, that by actions acceptable unto thee we may attain unto the perpetual contemplation of his glory in whom thou hast testified that thou, the Father, wast well pleased. Through the same.
%1928 BCP:
%The issue here is that the triple directed prayer doesn't follow the traditional style of Collect and comes off as jarring.
%\lett{O}{God,} who on the mount didst reveal to chosen witnesses thine only-begotten Son wonderfully transfigured, in raiment white and glistering; Mercifully grant that we, being delivered from the disquietude of this world, may be permitted to behold the King in his beauty, who with thee, O Father, and thee, O Holy Ghost, liveth and reigneth, one God, world without end. \textit{Amen.}

\readingcitation{Epistle}{2 Peter 1:13}
\lett{D}{early beloved:} I think it meet, as long as I am in this tabernacle, to stir you up by putting you in remembrance; knowing that shortly I must put off this my tabernacle, even as our Lord Jesus Christ hath shewed me. Moreover I will endeavour that ye may be able after my decease to have these things always in remembrance. We have not followed cunningly devised fables, when we made known unto you the power and coming of our Lord Jesus Christ, but were eyewitnesses of his majesty. For he received from God the Father honour and glory, when there came such a voice to him from the excellent glory, This is my beloved Son, in whom I am well pleased. And this voice which came from heaven we heard, when we were with him in the holy mount.

\gradall{In the day of thy power shall the people offer thee free-will offerings with an holy worship. The dew of thy birth is of the womb of the morning. ℣. The Lord said unto my Lord, Sit thou on my right hand until I make thy enemies thy footstool.}{Alleluia, alleluia. ℣. The hallowed day hath lightened upon us; come ye nations and adore the Lord, for a great light hath this day descended upon the earth.}

%1928:
\readingcitation{Gospel}{Luke 9:28}
\lett{A}{t that time:} it came to pass about an eight days after these sayings, he took Peter and John and James, and went up into a mountain to pray. And as he prayed, the fashion of his countenance war altered, and his raiment was white and glistering. And, behold, there talked with him two men, which were Moses and Elias: who appeared in glory, and spake of his decease, which he should accomplish at Jerusalem. But Peter and they that were with him were heavy with sleep: and when they were awake, they saw his glory, and the two men that stood with him. And it came to pass, as they departed from him, Peter said unto Jesus, Master, it is good for us to be here: and let us make three tabernacles; one for thee, and one for Moses, and one for Elias: not knowing what he said. While he thus spake, there came a cloud, and overshadowed them: and they feared as they entered into the cloud. And there came a voice out of the cloud, saying, This is my beloved Son: hear him. And when the voice was past, Jesus was found alone. And they kept it close, and told no man in those days any of those things which they had seen.
%Sarum:
%\lett{A}{t that time:} Jesus taketh Peter, James, and John his brother, and bringeth them up into an high mountain apart, And was transfigured before them: and his face did shine as the sun, and his raiment was white as the light. And, behold, there appeared unto them Moses and Elias talking with him. Then answered Peter, and said unto Jesus, Lord, it is good for us to be here: if thou wilt, let us make here three tabernacles; one for thee, and one for Moses, and one for Elias. While he yet spake, behold, a bright cloud overshadowed them: and behold a voice out of the cloud, which said, This is my beloved Son, in whom I am well pleased; hear ye him. And when the disciples heard it, they fell on their face, and were sore afraid. And Jesus came and touched them, and said, Arise, and be not afraid. And when they had lifted up their eyes, they saw no man, save Jesus only. And as they came down from the mountain, Jesus charged them, saying, Tell the vision to no man, until the Son of man be risen again from the dead.

\offertory{God hath made the round world so sure that it cannot be moved. Ever since the world began hath thy seat been prepared, O God; thou art from everlasting.}
\secret
\lett{O}{Lord,} heavenly Father, almighty, everlasting God, receive, we beseech thee, the gifts which we present on the glorious Transfiguration of thy Son; and mercifully grant that by them we may be delivered from the disquietude of this world, and may be knit together in happiness eternal. Through the same.

\communion{The dew of thy birth is of the womb of the morning.}

\postcommunion
\lett{O}{God,} who hast hallowed this day by the Transfiguration of thine Incarnate Word, and by the voice of thee, the Father, sent down to him, grant, we beseech thee, that we who have been fed with the Bread of Heaven, may be found meet to become members of him who gave commandment that this should be done in remembrance of him, Jesus Christ, thy Son, our Lord. Who.

\subby{Most Holy Name of Jesus Christ}
\feastday{Holy Name}
\begin{inhead}
    {Second Class Double\\
7 August}
\end{inhead}
\feastdate{7}{August}
\begin{rubric}
    The propers are the same as on the Christmastide Feast Day of the same name (p. \pageref{MostHolyName}).
\end{rubric}

\subby{St. Lawrence}
\feastday{St. Lawrence}
\begin{inhead}
    {Second Class Double with Common Octave\\
10 August}
\end{inhead}
\feastdate{10}{August}
%S-R: No Sarum
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} He hath dispersed abroad, and given to the poor : and his righteousness remaineth for ever.%\vr{Ps. 112:9}

\introit
\lett{G}{lory} and worship are before him: power and honour are in his sanctuary. \textit{Ps.} O sing unto the Lord a new song: sing unto the lord, all the whole earth.

\collect
\lett{G}{rant} to us, we beseech thee, almighty God: that we may quench the flames of our sins; as thou didst enable blessed Lawrence to overcome the fires of his torments. Through.

\readingcitation{Epistle}{2 Corinthians 9:6}
\lett{B}{rethren:} He which soweth sparingly shall reap also sparingly; and he which soweth bountifully shall reap also bountifully. Every man according as he purposeth in his heart, so let him give; not grudgingly, or of necessity: for God loveth a cheerful giver. And God is able to make all grace abound toward you; that ye, always having all sufficiency in all things, may abound to every good work: (As it is written, He hath dispersed abroad; he hath given to the poor: his righteousness remaineth for ever. Now he that ministereth seed to the sower both minister bread for your food, and multiply your seed sown, and increase the fruits of your righteousness;)

\gradual{Thou hast proved and visited mine heart, O Lord, in the night-season. ℣. Thou hast tried me with fire, and hast found no wickedness in me.}

\alleluia{Alleluia, alleluia. ℣. The Levite Lawrence wrought a good work: who by the sign of the cross gave light to the blind. Alleluia.}

\readingcitation{Gospel}{John 12:24}
\lett{A}{t that time:} Jesus said unto his disciples: Verily, verily, I say unto you, Except a corn of wheat fall into the ground and die, it abideth alone: but if it die, it bringeth forth much fruit. He that loveth his life shall lose it; and he that hateth his life in this world shall keep it unto life eternal. If any man serve me, let him follow me; and where I am, there shall also my servant be: if any man serve me, him will my Father honour.

\offertory{Glory and worship are before him; power and honour are in his sanctuary.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\secret
\lett{A}{ccept,} we beseech thee, O Lord, the gifts which we duly offer: that by the merits and intercession of blessed Lawrence, thy Martyr, they may be profitable to the advancement of our salvation. Through.

\communion{He that serveth me, let him follow me: and where I am there shall also my servant be.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\postcommunion
\lett{O}{Lord,} who hast satisfied us with this sacred gift, we humbly beseech thee: that, we who celebrate the service of our bounden duty may, at the intercession of blessed Lawrence thy Martyr perceive the increase of thy saving grace. Through.


\subby{Assumption of the Blessed Virgin Mary}
\feastday{Assumption}
\begin{inhead}
    {First Class Double with Common Octave\\
15 August}
\end{inhead}
\feastdate{15}{August}

\introit
\lett{R}{ejoice} we all in the Lord, keeping feast day in honour of the blessed Virgin Mary: in whose Assumption the Angels rejoice and glorify the Son of God. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
\lett{W}{e} beseech thee, O Lord, let us be continually aided by the sacred feast of this day, whereon the holy Mother of God underwent death in this world, and yet could not be holden by the chains of death: who did bring forth thy Son our Lord. Who liveth.

\readingcitation{Epistle}{Ecclesiasticus 24:7}
%\lett{W}{ith} all these I sought rest: and in whose inheritance shall I abide? So the Creator of all things gave me a commandment, and he that made me caused my tabernacle to rest, and said, Let thy dwelling be in Jacob, and thine inheritance in Israel. He created me from the beginning before the world, and I shall never fail. In the holy tabernacle I served before him; and so was I established in Sion. Likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance. I was exalted like a cedar in Libanus, and as a cypress tree upon the mountains of Hermon. I was exalted like a palm tree in En-gaddi, and as a rose plant in Jericho, as a fair olive tree in a pleasant field, and grew up as a plane tree by the water. I gave a sweet smell like cinnamon and aspalathus, and I yielded a pleasant odour like the best myrrh.
%RV:
\lett{W}{ith} all these I sought rest; And in whose inheritance shall I lodge? Then the Creator of all things gave me a commandment; And he that created me made my tabernacle to rest, And said, Let thy tabernacle be in Jacob, And thine inheritance in Israel. He created me from the beginning before the world; And to the end I shall not fail. In the holy tabernacle I ministered before him; And so was I established in Sion. In the beloved city likewise he gave me rest; And in Jerusalem was my authority. And I took root in a people that was glorified, Even in the portion of the Lord’s own inheritance. I was exalted like a cedar in Libanus, And as a cypress tree on the mountains of Hermon. I was exalted like a palm tree on the sea shore, And as rose plants in Jericho, And as a fair olive tree in the plain; And I was exalted as a plane tree. As cinnamon and aspalathus, I have given a scent of perfumes; And as choice myrrh, I spread abroad a pleasant odour.

\gradall{Because of the word of truth, of meekness, and righteousness, and thy right hand shall teach thee terrible things. ℣. Hearken, O daughter, and consider, incline thine ear: so shall the King have pleasure in thy beauty.}{Alleluia, alleluia. ℣. This day the Virgin Mary entered heaven: rejoice, for she reigneth with Christ for ever. Alleluia.}

\subbysub{Sequence}
\begin{multicols}{2}
From our first Mother Eva's sickly branch

Mary the blooming rose proceeded forth;

Bright as amidst the stars the Morning Star,

And fair in beauty as the moon she came;\\

Sweet beyond balsam, ointments, frankincense;

As violet glowing, dewy as the rose,

White as the lily, she who was preferred

To bear the highest Father's holy Child,\\

%\columnbreak

That of a Virgin's flesh immaculate

He might upon him take flesh hallowed.

Great Gabriel brings the message of new joy,

Th' arising of the eternal King on earth,\\

And to His Mother thus gives salutation:

Blessed art thou, Queen of the universe!

Blessed art thou, Queen of the universe!

Blessed art thou, Queen of the universe!

Thou shalt bring forth the Everlasting King.

She answered, How can I fruitful be,\\

Seeing a man I know not, from my birth

Ever a Virgin chaste continuing?

Fear not (the Angel answered), upon thee,

Chaste as thou art, the Holy Ghost shall come,\\

Whereby thou shalt bear God and Man in one.

O truly holy, truly to be loved!

Of whom redemption hath for us arisen,

Salvation of the world, and our true life.\\

Mother of God, accept our prayers this day,

Whereon to heaven's portals thou wast borne.

Dear to the Father, Jesus' Mother pure,

The Holy Spirit's temple thou wast made.\\

Fair Spouse of God, thou Christ the King hast borne:

Lady thou art in heaven and in earth.

This day hosts met thee from the court of heaven,

And to the starry palace led thee up.\\

Jesus himself, to welcome thee his Mother,

Came with the angels forth, and set thee up

With him for ever in his Father's seat.

With God now reigning, mercifully pardon\\

Our evil deeds, and ask for us all good.

O gracious Mediatrix, next to God

Our only hope, commend us to thy Son,

That we in highest heaven may Alleluias sing.
\end{multicols}

%MANUAL ADJUSTMENT:
\vspace{-2.5ex}

\readingcitation{Gospel}{Luke 10:38}
\lett{A}{t that time:} It came to pass, as they went, that Jesus entered into a certain village: and a certain woman named Martha received him into her house. And she had a sister called Mary, which also sat at Jesus' feet, and heard his word. But Martha was cumbered about much serving, and came to him, and said, Lord, dost thou not care that my sister hath left me to serve alone? bid her therefore that she help me. And Jesus answered and said unto her, Martha, Martha, thou art careful and troubled about many things: but one thing is needful: and Mary hath chosen that good part, which shall not be taken away from her.

\offertory{Full of grace are thy lips, because God hath blessed thee for ever.}

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\secret
\lett{W}{e} beseech thee, O Lord, that the prayers of the Blessed Mother of God may make our gifts acceptable to thee, that albeit we know her to have departed this life in respect of this flesh we may perceive her prayers for us continually in the glory of heaven. Through.

\communion{Blessed is the womb of the Virgin Mary, which bare the Son of the everlasting Father.}

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\postcommunion
\lett{W}{e} that have partaken of thy heavenly table implore thy mercy, O Lord our God, that we who keep the feast of the Mother of God may at her intercession be delivered from all evils which beset us. Through.

\begin{rubric}
    Within the Octave, Mass is said as on the Feast. And the \nth{2} Collect is of the Holy Ghost (p. \pageref{SPHolyGhost}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}
\begin{rubric}
    The following Alleluia is said daily, except on the Sunday in the Octave and the Octave Day.
\end{rubric}
\alleluia{Alleluia, alleluia. ℣. Mary is taken up into heaven: angels rejoice, praise, and bless the Lord.}

\subby{St. Joachim, Father of the Blessed Virgin Mary}
\feastday{St. Joachim}
\begin{inhead}
    {Second Class Double\\
16 August}
\end{inhead}
\feastdate{16}{August}

\introit
\lett{H}{e} hath dispersed abroad, and given to the poor his righteousness remaineth for ever: his horn shall be exalted with honour. \textit{Ps.} Blessed is the man that feareth the Lord he hath great delight in his commandments.

\collect
\lett{O}{God,} who from amongst all thy Saints didst choose blessed Joachim to be the father of the Mother of thy Son: grant, we beseech thee; that we who venerate his festival may also continually perceive his advocacy. Through the same.
\readingcitation{Epistle}{Ecclesiasticus 31:8}
%\lett{B}{lessed} is the rich that is found without blemish, and hath not gone after gold. Who is he? and we will call him blessed: for wonderful things hath he done among his people. Who hath been tried thereby, and found perfect? then let him glory. Who might offend, and hath not offended? or done evil, and hath not done it? His goods shall be established, and the congregation shall declare his alms.
%RV:
\lett{B}{lessed} is the rich that is found without blemish, And that goeth not after gold. Who is he? and we will call him blessed: For wonderful things hath he done among his people. Who hath been tried thereby, and found perfect? Then let him glory. Who hath had the power to transgress, and hath not transgressed? And to do evil, and hath not done it? His goods shall be made sure, And the congregation shall declare his alms.

\gradall{He hath dispersed abroad, and given to the poor: his righteousness remaineth for ever. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed.}{Alleluia, alleluia. ℣. Joachim, spouse of Saint Anne, of the gracious Virgin the father, here to thy servants bring safety and aid from on high. Alleluia.}
\readingcitation{Gospel}{Matthew 1:1}
\lett{T}{he} book of the generation of Jesus Christ, the son of David, the son of Abraham. Abraham begat Isaac; and Isaac begat Jacob; and Jacob begat Judas and his brethren; And Judas begat Phares and Zara of Thamar; and Phares begat Esrom; and Esrom begat Aram; And Aram begat Aminadab; and Aminadab begat Naasson; and Naasson begat Salmon; And Salmon begat Booz of Rachab; and Booz begat Obed of Ruth; and Obed begat Jesse; And Jesse begat David the king; and David the king begat Solomon of her that had been the wife of Urias; And Solomon begat Roboam; and Roboam begat Abia; and Abia begat Asa; And Asa begat Josaphat; and Josaphat begat Joram; and Joram begat Ozias; And Ozias begat Joatham; and Joatham begat Achaz; and Achaz begat Ezekias; And Ezekias begat Manasses; and Manasses begat Amon; and Amon begat Josias; And Josias begat Jechonias and his brethren, about the time they were carried away to Babylon: And after they were brought to Babylon, Jechonias begat Salathiel; and Salathiel begat Zorobabel; And Zorobabel begat Abiud; and Abiud begat Eliakim; and Eliakim begat Azor; And Azor begat Sadoc; and Sadoc begat Achim; and Achim begat Eliud; And Eliud begat Eleazar; and Eleazar begat Matthan; and Matthan begat Jacob; And Jacob begat Joseph the husband of Mary, of whom was born Jesus, who is called Christ.
\begin{rubric}
    The Creed is said, on account of the Octave of the Assumption.
\end{rubric}
\offertory{Thou hast crowned him with glory and worship: and hast made him to have dominion of the works of thy hands, O Lord.}
\secret
\lett{R}{eceive,} most merciful God, the sacrifice which we offer to thy majesty, in honour of the holy Patriarch Joachim, father of the Virgin Mary: and grant that, he with his wife and their most blessed offspring interceding for us, we may be counted worthy to receive perfect remission of sins and everlasting glory. Through the same.
\begin{rubric}
    The Preface of the B.V.M. is said.
\end{rubric}
\communion{A faithful and wise servant, whom the lord hath made ruler over his household: to give them their portion of meat in due season.}
\postcommunion
\lett{W}{e} beseech thee, almighty God: that through these sacraments which we have received, with the pleading of the merits and prayers of blessed Joachim, father of the Mother of thy beloved Son Jesus Christ our Lord, we may be found worthy to be made partakers of thy grace in this life and of eternal glory in the life to come. Through the same.

\subby{St. Bartholomew}
\feastday{St. Bartholomew}
\begin{inhead}
    {Second Class Double\\
24 August}
\end{inhead}
\feastdate{24}{August}

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting and mine up-rising.

\collect
\lett{O}{almighty} and everlasting God, who didst give to thine Apostle Bartholomew grace truly to believe and to preach thy Word; Grant, we beseech thee, unto thy Church to love that Word which he believed, and both to preach and receive the same. Through.

\readingcitation{Epistle}{Acts 5:12}
\lett{I}{n those days:} By the hands of the apostles were many signs and wonders wrought among the people; (and they were all with one accord in Solomon's porch. And of the rest durst no man join himself to them: but the people magnified them. And believers were the more added to the Lord, multitudes both of men and women.) Insomuch that they brought forth the sick into the streets, and laid them on beds and couches, that at the least the shadow of Peter passing by might overshadow some of them. There came also a multitude out of the cities round about unto Jerusalem, bringing sick folks, and them which were vexed with unclean spirits: and they were healed every one.

\gradall{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}{Alleluia, alleluia. ℣. The glorious company of the Apostles praise thee, O Lord. Alleluia.}

\readingcitation{Gospel}{Luke 22:24}
\lett{A}{t that time:} There was a strife among the disciples, which of them should be accounted the greatest. And he said unto them, The kings of the Gentiles exercise lordship over them; and they that exercise authority upon them are called benefactors. But ye shall not be so: but he that is greatest among you, let him be as the younger; and he that is chief, as he that doth serve. For whether is greater, he that sitteth at meat, or he that serveth? is not he that sitteth at meat? but I am among you as he that serveth. Ye are they which have continued with me in my temptations. And I appoint unto you a kingdom, as my Father hath appointed unto me; That ye may eat and drink at my table in my kingdom, and sit on thrones judging the twelve tribes of Israel.

\offertory{Right honourable are thy friends unto me, O God: right well is their princedom established.}
\secret
\lett{W}{e} beseech thee, O Lord, that we, who celebrate the festival of thy blessed Apostle Bartholomew: may, by the succour of him in whose honour we offer thee these sacrifices of praise, receive thy benefits. Through.
\communion{Ye which have followed me shall sit upon thrones, judging the twelve tribes of Israel, saith the Lord.}
\postcommunion
\lett{O}{Lord,} through whom we have received the pledge of eternal redemption: we beseech thee; that at the intercession of blessed Bartholomew, thine Apostle, it may avail for our succour both in this life and that which is to come. Through.

\subby{The Beheading of St. John Baptist}
\feastday{Decollation}
\begin{inhead}
    {Greater Double\\
29 August}
\end{inhead}
\feastdate{29}{August}

\introit
\lett{I}{will} speak of thy testimonies, even before kings, and will not be ashamed: and my delight shall be in thy commandments, which I have loved exceedingly. \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.

\collect
\lett{W}{e} beseech thee, O Lord: that the venerable festival of thy Forerunner and Martyr, Saint John Baptist, may effectually bestow upon us thy succour unto our salvation. Who livest.
\needspace{4\baselineskip}
\lett{O}{God,} who among the manifold works of thy power hast bestowed even upon the weakness of women the victory of martyrdom: mercifully grant; that we, who celebrate the birthday of blessed Sabina, thy Virgin and Martyr, may by her example be drawn nearer unto thee. Through.

\readingcitation{Epistle}{Jeremiah 1:17}
\lett{I}{n those days:} The word of the Lord came unto me, saying: Gird up thy loins, and arise, and speak unto Judah all that I command thee: be not dismayed at their faces, lest I confound thee before them. For, behold, I have made thee this day a defenced city, and an iron pillar, and brasen walls against the whole land, against the kings of Judah, against the princes thereof, against the priests thereof, and against the people of the land. And they shall fight against thee; but they shall not prevail against thee; for I am with thee, saith the Lord, to deliver thee.

\gradall{The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus in the house of the Lord. ℣. To tell of thy loving-kindness early in the morning, and of thy truth in the night-season.}{Alleluia, alleluia. ℣. The righteous shall grow as the lily: and shall flourish for ever before the Lord. Alleluia.}

\readingcitation{Gospel}{Mark 6:17}
\lett{A}{t that time:} Herod sent forth and laid hold upon John, and bound him in prison for Herodias' sake, his brother Philip’s wife: for he had married her. For John had said unto Herod, It is not lawful for thee to have thy brother’s wife. Therefore Herodias had a quarrel against him, and would have killed him; but she could not: for Herod feared John, knowing that he was a just man and an holy, and observed him; and when he heard him, he did many things, and heard him gladly. And when a convenient day was come, that Herod on his birthday made a supper to his lords, high captains, and chief estates of Galilee; and when the daughter of the said Herodias came in, and danced, and pleased Herod and them that sat with him, the king said unto the damsel, Ask of me whatsoever thou wilt, and I will give it thee. And he sware unto her, Whatsoever thou shalt ask of me, I will give it thee, unto the half of my kingdom. And she went forth, and said unto her mother, What shall I ask? And she said, The head of John the Baptist. And she came in straightway with haste unto the king, and asked, saying, I will that thou give me by and by in a charger the head of John the Baptist. And the king was exceeding sorry; yet for his oath’s sake, and for their sakes which sat with him, he would not reject her. And immediately the king sent an executioner, and commanded his head to be brought: and he went and beheaded him in the prison, and brought his head in a charger, and gave it to the damsel: and the damsel gave it to her mother. And when his disciples heard of it, they came and took up his corpse, and laid it in a tomb.

\offertory{The just shall rejoice in thy strength, O Lord, exceeding glad shall he be of thy salvation: thou hast given him his heart's desire.}

\secret
\lett{W}{e} beseech thee, O Lord, that the gifts which we offer unto thee for the passion of thy holy Martyr John Baptist: may through his intercession be profitable for our salvation. Through.
\needspace{4\baselineskip}
\lett{G}{raciously} receive, O Lord, through the merits of blessed Sabina, thy Virgin and Martyr, the sacrifices which we offer unto thee: and grant that they may avail for our continual help. Through.

\communion{Thou hast set, O Lord, a crown of pure gold upon his head.}

\postcommunion
\lett{L}{et} the solemnity, of Saint John Baptist enable us, O Lord: both to reverence those things which are signified in the wondrous sacraments which we have received, and to rejoice abundantly in the fruits that they bring forth in us. Through.
\needspace{4\baselineskip}
\lett{O}{Lord} our God, who hast fulfilled us with the bounty of thy heavenly gift; we beseech thee, that, at the intercession of blessed Sabina, thy Virgin and Martyr, we may ever live by the partaking of the same. Through.


\subby{Nativity of the Blessed Virgin Mary}
\feastday{Nativity of the BVM}
\begin{inhead}
    {Second Class Double with a Common Octave\\
8 September}
\end{inhead}
\feastdate{8}{September}
%S-R:
\par\noindent
\textit{Opening Sentence.} The Lord shall come down like the rain into a fleece of wool. Even as the drops that water the earth.%\vr{Judg. 6:37}

\introit
\lett{H}{ail,} O Mother most holy, who in childbirth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.
\collect
\lett{W}{e} beseech thee, O Lord, to grant unto us thy servants the gift of heavenly grace: that as the childbearing of the blessed Virgin was unto us the beginning of salvation; so the devout observance of her Nativity may bestow an increase of peace. Through.
\readingcitation{Epistle}{Proverbs 8:22}
\lett{T}{he} \textsc{Lord} possessed me in the beginning of his way, before his works of old. I was set up from everlasting, from the beginning, or ever the earth was. When there were no depths, I was brought forth; when there were no fountains abounding with water. Before the mountains were settled, before the hills was I brought forth: While as yet he had not made the earth, nor the fields, nor the highest part of the dust of the world. When he prepared the heavens, I was there: when he set a compass upon the face of the depth: When he established the clouds above: when he strengthened the fountains of the deep: When he gave to the sea his decree, that the waters should not pass his commandment: when he appointed the foundations of the earth: Then I was by him, as one brought up with him: and I was daily his delight, rejoicing always before him; Rejoicing in the habitable part of his earth; and my delights were with the sons of men. Now therefore hearken unto me, O ye children: for blessed are they that keep my ways. Hear instruction, and be wise, and refuse it not. Blessed is the man that heareth me, watching daily at my gates, waiting at the posts of my doors. For whoso findeth me findeth life, and shall obtain favour of the \textsc{Lord}.
\gradall{Blessed and venerable art thou, O Virgin Mary: who without spot wast found Mother of the Saviour. ℣. Virgin, Mother of God, he, whom the whole world containeth not, being made man lay hid in thy womb.}{Alleluia, alleluia, ℣. Happy art thou, O sacred Virgin Mary, and most worthy of all praise: for out of thee hath arisen the son of righteousness, Christ our God. Alleluia.}
\readingcitation{Gospel}{Matthew 1:1}
\lett{T}{he} book of the generation of Jesus Christ, the son of David, the son of Abraham. Abraham begat Isaac; and Isaac begat Jacob; and Jacob begat Judas and his brethren; And Judas begat Phares and Zara of Thamar; and Phares begat Esrom; and Esrom begat Aram; And Aram begat Aminadab; and Aminadab begat Naasson; and Naasson begat Salmon; And Salmon begat Booz of Rachab; and Booz begat Obed of Ruth; and Obed begat Jesse; And Jesse begat David the king; and David the king begat Solomon of her that had been the wife of Urias; And Solomon begat Roboam; and Roboam begat Abia; and Abia begat Asa; And Asa begat Josaphat; and Josaphat begat Joram; and Joram begat Ozias; And Ozias begat Joatham; and Joatham begat Achaz; and Achaz begat Ezekias; And Ezekias begat Manasses; and Manasses begat Amon; and Amon begat Josias; And Josias begat Jechonias and his brethren, about the time they were carried away to Babylon: And after they were brought to Babylon, Jechonias begat Salathiel; and Salathiel begat Zorobabel; And Zorobabel begat Abiud; and Abiud begat Eliakim; and Eliakim begat Azor; And Azor begat Sadoc; and Sadoc begat Achim; and Achim begat Eliud; And Eliud begat Eleazar; and Eleazar begat Matthan; and Matthan begat Jacob; And Jacob begat Joseph the husband of Mary, of whom was born Jesus, who is called Christ.
\offertory{Blessed art thou, O Virgin Mary, who didst bear the Creator of all things: thou broughtest forth him who made thee, and for ever remainest a Virgin.}
\secret
\lett{M}{ay} the manhood of thine only-begotten Son, O Lord, avail for our succour: that, even as he, being born of a Virgin, destroyed not but hallowed the innocence of his Mother; so on this feast of her Nativity, he may deliver us from out offences, and render us an oblation acceptable unto thee, even Jesus Christ our Lord. Who liveth.
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord: that the sacraments which we have received in the observance of this yearly festival; may both in this life and that which is to come be profitable for our healing. Through.

\begin{rubric}
    Within the Octave, nothing is said of it; but Votive Masses of the Blessed Virgin Mary and her Mass on Saturday are said as on the Feast, with \emph{Gloria in excelsis}, \nth{2} Collect of the Holy Ghost (p. \pageref{SPHolyGhost}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}), and the Creed is omitted.
\end{rubric}
\begin{rubric}
    Also, if by reason of another Octave celebrated in any place the \nth{2} or \nth{3} Collect is to be said of St. Mary, it is said of the Nativity, as above.
\end{rubric}


\subby{Exaltation of the Holy Cross}
\feastday{Roodmas}
\begin{inhead}
	{Greater Double\\
		14 September}
\end{inhead}
\feastdate{14}{September}

\properantiphonfix

\introit
\lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life, and resurrection: by whom we are saved and set free. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.

\collect
\lett{O}{God,} who on this day dost gladden us with the yearly solemnity of the exaltation of the holy Cross: grant, we beseech thee; that as we have known the mystery of thy Son on earth, so we may attain unto the rewards of his redemption in heaven. Through the same

\readingcitation{Epistle}{Philippians 2:5}
\lett{B}{rethren:} Let this mind be in you, which was also in Christ Jesus: who, being in the form of God, thought it not robbery to be equal with God: but made himself of no reputation, and took upon him the form of a servant, and was made in the likeness of men: and being found in fashion as a man, he humbled himself, and became obedient unto death, even the death of the cross. Wherefore God also hath highly exalted him, and given him a name which is above every name: \inrub{Here genuflect.} that at the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth; and that every tongue should confess that Jesus Christ is Lord, to the glory of God the Father.

%MANUAL ADJUSTMENT:
\gradual{Christ for us became obedient unto death even the death of the cross. ℣. Wherefore God also hath highly exalted him, and given him a name which is above every name.}

\alleluia{Alleluia, alleluia. ℣. Sweetest wood, sweetest iron, sweetest weight is hung on thee: thou alone wast counted worthy to bear the King of heaven and its Lord. Alleluia.}

\readingcitation{Gospel}{John 12:31}
\lett{A}{t that time:} Jesus said unto the multitudes of the Jews: Now is the judgment of this world: now shall the prince of this world be cast out. And I, if I be lifted up from the earth, will draw all men unto me. This he said, signifying what death he should die. The people answered him, We have heard out of the law that Christ abideth for ever: and how sayest thou, The Son of man must be lifted up? who is this Son of man? Then Jesus said unto them, Yet a little while is the light with you. Walk while ye have the light, lest darkness come upon you: for he that walketh in darkness knoweth not whither he goeth. While ye have light, believe in the light, that ye may be the children of light.

\offertory{Protect, O Lord, thy people, by the sign of the holy Cross, from all the snares of every enemy: that we may render thee pleasing service, and that our sacrifice may be acceptable unto thee, alleluia.}

\secret
\lett{O}{Lord,} our God, who dost fulfil us with the Body and Blood of Jesus Christ our Lord, through whom thou hast sanctified the standard of the Cross: we beseech thee; that as we have been counted worthy to venerate that Cross, so we may effectually be made partakers of its glory unto everlasting salvation. Through the same.

\communion{By the sign of the cross deliver us from our enemies, O our God.}

\postcommunion
\lett{A}{ssist} us, O Lord our God: and as thou dost vouchsafe unto us to rejoice in honouring the holy Cross, so defend us by the perpetual succour of the same. Through.

%NEED TO ADD PAGE OR REFERENCE:
%\begin{rubric}
%	For Votive Masses, the Mass of the Holy Cross is said.
%\end{rubric}


\subby{Ember Wednesday in Autumn}
\feastday{Autumnal Emberday}
\begin{inhead}
    {Second Class Feria\\
Wednesday after the Exaltation of the Holy Cross (14 September)\\
On this day and the following Friday and Saturday, no Feast is kept unless it be a First or Second Class Double.}
\end{inhead}
\fancyhead[RE,LO]{Ember Wednesday}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\properantiphon{Ben.}{This kind {\dag} of evil spirit can come forth by nothing but by prayer and fasting.}
\introit
\lett{S}{ing} we merrily unto God our strength: make a cheerful noise unto the God of Jacob: take the merry psalm with the lute: sound the trumpet in the beginning of the month, for this was made a statute for Israel, and a law of the God of Jacob. \textit{Ps.} This he ordained in Joseph for a testimony, when he came out of the land of Egypt: and had heard a strange language.
\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{W}{e} beseech thee, O Lord, that our frailty may be upheld by the healing of thy loving kindness: that what by its own nature is ready to decay, may by thy mercy be renewed. Through.
\readingcitation{Epistle}{Amos 9:13}
\lett{T}{hus} saith the Lord God: Behold, the days come: that the plowman shall overtake the reaper, and the treader of grapes him that soweth seed; and the mountains shall drop sweet wine, and all the hills shall melt. And I will bring again the captivity of my people of Israel, and they shall build the waste cities, and inhabit them; and they shall plant vineyards, and drink the wine thereof; they shall also make gardens, and eat the fruit of them. And I will plant them upon their land, and they shall no more be pulled up out of their land which I have given them, saith the \divineName{Lord} thy God.
\gradual{Who is like unto the Lord our God, that hath his dwelling so high, and yet humbleth himself to behold the things that are in heaven and earth? ℣. He taketh up the simple out of the dust, and lifteth the poor out of the mire.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\collect
\lett{G}{rant,} we beseech thee, O Lord, unto thy family who pray unto thee: that, as they do abstain from bodily food, so they may also spiritually fast from sin. Through.
\begin{rubric}
    Additional Collects are as on Trinity III (p. \pageref{TrinityIIICollect}).
\end{rubric}

\readingcitation{Epistle}{Nehemiah 8:1}
\lett{I}{n those days:} All the people gathered themselves together as one man into the street that was before the water gate; and they spake unto Ezra the scribe to bring the book of the law of Moses, which the \divineName{Lord} had commanded to Israel. And Ezra the priest brought the law before the congregation both of men and women, and all that could hear with understanding, upon the first day of the seventh month. And he read therein before the street that was before the water gate from the morning until midday, before the men and the women, and those that could understand; and the ears of all the people were attentive unto the book of the law. And Ezra the scribe stood upon a pulpit of wood, which they had made for the purpose. And Ezra opened the book in the sight of all the people; (for he was above all the people;) and when he opened it, all the people stood up: And Ezra blessed the \divineName{Lord}, the great God. And all the people answered, Amen, Amen, with lifting up their hands: and they bowed their heads, and worshipped the \divineName{Lord} with their faces to the ground. And the Levites, caused the people to understand the law: and the people stood in their place. So they read in the book in the law of God distinctly, and gave the sense, and caused them to understand the reading. And Nehemiah, and Ezra the priest the scribe, and the Levites that taught the people, said unto all the people, This day is holy unto the \divineName{Lord} your God; mourn not, nor weep. For all the people wept, when they heard the words of the law. Then he said unto them, Go your way, eat the fat, and drink the sweet, and send portions unto them for whom nothing is prepared: for this day is holy unto our Lord: neither be ye sorry; for the joy of the \divineName{Lord} is your strength.

\gradual{Blessed are the people, whose God is the Lord: and blessed are the folk, that he hath chosen to him to be his inheritance. ℣. By the word of the Lord were the heavens made: and all the hosts of them by the breath of his mouth.}

\readingcitation{Gospel}{Mark 9:17}
\lett{A}{t that time:} One of the multitude answered and said unto Jesus: Master, I have brought unto thee my son, which hath a dumb spirit; And wheresoever he taketh him, he teareth him: and he foameth, and gnasheth with his teeth, and pineth away: and I spake to thy disciples that they should cast him out; and they could not. He answereth him, and saith, O faithless generation, how long shall I be with you? how long shall I suffer you? bring him unto me. And they brought him unto him: and when he saw him, straightway the spirit tare him; and he fell on the ground, and wallowed foaming. And he asked his father, How long is it ago since this came unto him? And he said, Of a child. And ofttimes it hath cast him into the fire, and into the waters, to destroy him: but if thou canst do any thing, have compassion on us, and help us. Jesus said unto him, If thou canst believe, all things are possible to him that believeth. And straightway the father of the child cried out, and said with tears, Lord, I believe; help thou mine unbelief. When Jesus saw that the people came running together, he rebuked the foul spirit, saying unto him, Thou dumb and deaf spirit, I charge thee, come out of him, and enter no more into him. And the spirit cried, and rent him sore, and came out of him: and he was as one dead; insomuch that many said, He is dead. But Jesus took him by the hand, and lifted him up; and he arose. And when he was come into the house, his disciples asked him privately, Why could not we cast him out? And he said unto them, This kind can come forth by nothing, but by prayer and fasting.

%MANUAL ADJUSTMENT:
\clearpage

\offertory{My delight shall be in thy commandments, which I have loved exceedingly: my hands also will I lift up unto thy commandments, which I have loved.}

\secret
\lett{W}{e} beseech thee, O Lord, that this oblation may cleanse us from our sins: and sanctify thy servants both in body and mind for the celebration of this sacrifice. Through.
\begin{rubric}
    Additional Secrets are as on Trinity III (p. \pageref{TrinityIIISecret}).
\end{rubric}

\communion{Eat the fat, and drink the sweet, and send portions unto them for whom nothing is prepared: for this day is holy unto our Lord, neither be ye sorry: for the joy of the Lord is your strength.}

\postcommunion
\lett{W}{e} humbly beseech thee, O Lord: that like as of thy bounty we, receiving thy heavenly gifts, do with continual devotion offer the same, so by thy goodness we may worthily receive them in our souls. Through.
\begin{rubric}
    Additional Postcommunions are as on Trinity III (p. \pageref{TrinityIIIPostcommunion}).
\end{rubric}

\subby{Ember Friday in Autumn}
\feastday{Autumnal Emberday}
\begin{inhead}
    {Second Class Feria\\
Friday after Ember Wednesday}
\end{inhead}
\fancyhead[RE,LO]{Ember Friday}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\properantiphon{Ben.}{A woman {\dag} in the city, which was a sinner, stood at the Lord's feet behind him, and began to wash his feet with tears, and did wipe them with the hairs of her head, and kissed his feet, and anointed them with the ointment.}
\introit
\lett{L}{et} the heart of them rejoice that seek the Lord: seek the Lord and his strength: seek his face evermore. \textit{Ps.} O give thanks unto the Lord, and call upon his name: tell the people what things he hath done.
\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, who year by year devoutly keep this holy observance, may be acceptable unto thee both in body and in soul. Through.
\begin{rubric}
    Additional Collects are as on Trinity III (p. \pageref{TrinityIIICollect}).
\end{rubric}

\readingcitation{Epistle}{Hosea 14:1}
\lett{T}{hus} saith the Lord God: O Israel, return unto the \divineName{Lord} thy God; for thou hast fallen by thine iniquity. Take with you words, and turn to the \divineName{Lord}: say unto him, Take away all iniquity, and receive us graciously: so will we render the calves of our lips. Asshur shall not save us; we will not ride upon horses: neither will we say any more to the work of our hands, Ye are our gods: for in thee the fatherless findeth mercy. I will heal their backsliding, I will love them freely: for mine anger is turned away from him. I will be as the dew unto Israel: he shall grow as the lily, and cast forth his roots as Lebanon. His branches shall spread, and his beauty shall be as the olive tree, and his smell as Lebanon. They that dwell under his shadow shall return; they shall revive as the corn, and grow as the vine: the scent thereof shall be as the wine of Lebanon. Ephraim shall say, What have I to do any more with idols? I have heard him, and observed him: I am like a green fir tree. From me is thy fruit found. Who is wise, and he shall understand these things? prudent, and he shall know them? for the ways of the \divineName{Lord} are right, and the just shall walk in them: but the transgressors shall fall therein.
\gradual{Turn thee again, O Lord, at the last, and be gracious unto thy servants. ℣. Lord, thou hast been our refuge, from one generation to another.}
\readingcitation{Gospel}{Luke 7:36}
\lett{A}{t that time:} One of the Pharisees desired Jesus that he would eat with him. And he went into the Pharisee's house, and sat down to meat. And, behold, a woman in the city, which was a sinner, when she knew that Jesus sat at meat in the Pharisee's house, brought an alabaster box of ointment, And stood at his feet behind him weeping, and began to wash his feet with tears, and did wipe them with the hairs of her head, and kissed his feet, and anointed them with the ointment. Now when the Pharisee which had bidden him saw it, he spake within himself, saying, This man, if he were a prophet, would have known who and what manner of woman this is that toucheth him: for she is a sinner. And Jesus answering said unto him, Simon, I have somewhat to say unto thee. And he saith, Master, say on. There was a certain creditor which had two debtors: the one owed five hundred pence, and the other fifty. And when they had nothing to pay, he frankly forgave them both. Tell me therefore, which of them will love him most? Simon answered and said, I suppose that he, to whom he forgave most. And he said unto him, Thou hast rightly judged. And he turned to the woman, and said unto Simon, Seest thou this woman? I entered into thine house, thou gavest me no water for my feet: but she hath washed my feet with tears, and wiped them with the hairs of her head. Thou gavest me no kiss: but this woman since the time I came in hath not ceased to kiss my feet. My head with oil thou didst not anoint: but this woman hath anointed my feet with ointment. Wherefore I say unto thee, Her sins, which are many, are forgiven; for she loved much: but to whom little is forgiven, the same loveth little. And he said unto her, Thy sins are forgiven. And they that sat at meat with him began to say within themselves, Who is this that forgiveth sins also? And he said to the woman, Thy faith hath saved thee; go in peace.
\offertory{Praise the Lord, O my soul, and forget not all his benefits: who maketh thee young and lusty as an eagle.}
\secret
\lett{W}{e} beseech thee, O Lord, that the offerings of our fast may be acceptable in thy sight: that we, being cleansed thereby and made worthy of thy grace, may be brought unto thine everlasting promises. Through.
\begin{rubric}
    Additional Secrets are as on Trinity III (p. \pageref{TrinityIIISecret}).
\end{rubric}

\communion{O turn from me shame and rebuke, for I have kept thy commandments, O Lord: for thy testimonies are my delight.}
\postcommunion
\lett{W}{e} beseech thee, almighty God: that shewing forth our thankfulness for the gifts which we have received, we may obtain yet more abundant benefits. Through.
\begin{rubric}
    Additional Postcommunions are as on Trinity III (p. \pageref{TrinityIIIPostcommunion}).
\end{rubric}


\subby{Ember Saturday in Autumn}
\feastday{Autumnal Emberday}
\begin{inhead}
    {Second Class Feria\\
Saturday after Ember Friday}
\end{inhead}
\fancyhead[RE,LO]{Ember Saturday}
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\properantiphon{Ben.}{Give light, O Lord, {\dag} to them that sit in darkness: and gudie our feet into the way of peace, thou God of Israel.}
\introit
\lett{O}{come,} let us worship and fall down, and kneel before the Lord our maker: for he is the Lord our God. \textit{Ps.} O come, let us sing unto the Lord: let us heartily rejoice in the strength of our salvation.

\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{A}{lmighty} and everlasting God, who by salutary continence bestowest healing in body and soul: we humbly entreat thy majesty; that, thou wouldest mercifully look upon the devout prayers and fasting of thy people, and grant us help both in this life and that which is to come. Through.
\readingcitation{Epistle}{Leviticus 23:26}
\lett{I}{n those days:} The \divineName{Lord} spake unto Moses, saying, Also on the tenth day of this seventh month there shall be a day of atonement: it shall be an holy convocation unto you; and ye shall afflict your souls, and offer an offering made by fire unto the \divineName{Lord}. And ye shall do no work in that same day: for it is a day of atonement, to make an atonement for you before the \divineName{Lord} your God. For whatsoever soul it be that shall not be afflicted in that same day, he shall be cut off from among his people. And whatsoever soul it be that doeth any work in that same day, the same soul will I destroy from among his people. Ye shall do no manner of work: it shall be a statute for ever throughout your generations in all your dwellings. It shall be unto you a sabbath of rest, and ye shall afflict your souls: in the ninth day of the month at even, from even unto even, shall ye celebrate your sabbath.
\gradual{Be merciful unto our sins, O Lord: wherefore do the heathen say: Where is now their God? ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{G}{rant} to us, we beseech thee, almighty God: that we may through fasting be satisfied with thy grace; and through abstinence strengthened to overcome all our enemies. Through.
\readingcitation{Epistle}{Leviticus 23:39}
\lett{I}{n those days:} The Lord spake unto Moses, saying: In the fifteenth day of the seventh month, when ye have gathered in the fruit of the land, ye shall keep a feast unto the \divineName{Lord} seven days: on the first day shall be a sabbath, and on the eighth day shall be a sabbath. And ye shall take you on the first day the boughs of goodly trees, branches of palm trees, and the boughs of thick trees, and willows of the brook; and ye shall rejoice before the \divineName{Lord} your God seven days. And ye shall keep it a feast unto the \divineName{Lord} seven days in the year. It shall be a statute for ever in your generations: ye shall celebrate it in the seventh month. Ye shall dwell in booths seven days; all that are Israelites born shall dwell in booths: That your generations may know that I made the children of Israel to dwell in booths, when I brought them out of the land of Egypt: I am the \divineName{Lord} your God.
\gradual{Behold, O God our defender: and look upon thy servants. ℣. O Lord God of hosts, hear the prayers of thy servants.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}

\collect
\lett{K}{eep,} we beseech thee, O Lord, this thy family: that as by thine inspiration we do seek thy healing unto everlasting salvation, so by thy bounty we may obtain the same. Through.

\readingcitation{Epistle}{Micah 7:14}
\lett{O}{Lord} our God, Feed thy people with thy rod, the flock of thine heritage, which dwell solitarily in the wood, in the midst of Carmel: let them feed in Bashan and Gilead, as in the days of old. According to the days of thy coming out of the land of Egypt will I shew unto him marvellous things. The nations shall see and be confounded at all their might: they shall lay their hand upon their mouth, their ears shall be deaf. They shall lick the dust like a serpent, they shall move out of their holes like worms of the earth: they shall be afraid of the \divineName{Lord} our God, and shall fear because of thee. Who is a God like unto thee, that pardoneth iniquity, and passeth by the transgression of the remnant of his heritage? he retaineth not his anger for ever, because he delighteth in mercy. He will turn again, he will have compassion upon us; he will subdue our iniquities; and thou wilt cast all their sins into the depths of the sea. Thou wilt perform the truth to Jacob, and the mercy to Abraham, which thou hast sworn unto our fathers from the days of old: O Lord our God.
\gradual{Turn thee again, O Lord, at the last, and be gracious unto thy servants. ℣. Lord, thou hast been our refuge, from one generation to another.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\collect
\lett{G}{rant,} we beseech thee, almighty God: that we may so abstain from carnal feasting: that we may likewise fast from the vices which beset us. Through.

\readingcitation{Epistle}{Zechariah 8:14}
\lett{I}{n those days:} The word of the \divineName{Lord} came unto me, saying: Thus saith the \divineName{Lord} of hosts: As I thought to punish you, when your fathers provoked me to wrath, saith the \divineName{Lord} of hosts, and I repented not: So again have I thought in these days to do well unto Jerusalem and to the house of Judah: fear ye not. These are the things that ye shall do; Speak ye every man the truth to his neighbour; execute the judgment of truth and peace in your gates: And let none of you imagine evil in your hearts against his neighbour; and love no false oath: for all these are things that I hate, saith the \divineName{Lord}. And the word of the \divineName{Lord} of hosts came unto me, saying, Thus saith the \divineName{Lord} of hosts; The fast of the fourth month, and the fast of the fifth, and the fast of the seventh, and the fast of the tenth, shall be to the house of Judah joy and gladness, and cheerful feasts; therefore love the truth and peace: saith the \divineName{Lord} of hosts.
\gradual{Let my prayer be set forth in thy sight, O Lord, as the incense. ℣. Let the lifting up of my hands be an evening sacrifice.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}

\collect
\lett{O}{Lord,} who sufferest us to offer unto thee this solemn fast: we beseech thee that thou wouldest likewise bestow upon us the succour of thy pardon. Through.

\readingcitation{Epistle}{Song of the Three Children 26}
\lett{I}{n those days:} The Angel of the Lord came down into the oven together with Azarias and his fellows, and smote the flame of the fire out of the oven; And made the midst of the furnace as it had been a moist whistling wind, so that the fire touched them not at all, neither hurt nor troubled them. Then the three, as out of one mouth, praised, glorified, and blessed, God in the furnace, saying:
\begin{rubric}
    `Thanks be to God' is not said here.
\end{rubric}
\subbysub{Hymn}
\lett{B}{lessed} art thou, O Lord God of our fathers. And worthy to be praised and glorious for ever.

And blessed is the name of thy glory, which is holy. And worthy to be praised and glorious for ever.

Blessed art thou in the holy temple of thy glory. And worthy to be praised and glorious for ever.

Blessed art thou on the holy throne of thy kingdom. And worthy to be praised and glorious for ever.

Blessed art thou in the sceptre of thy godhead. And worthy to be praised and glorious for ever.

Blessed art thou that beholdest the depths, and sittest upon the Cherubim. And worthy to be praised and glorious for ever.

Blessed art thou that walkest on the wings of the winds, and on the waves of the sea. And worthy to be praised and glorious for ever.

Let all thine Angels and Saints bless thee. And let them praise thee and glorify thee for ever.

Let the heavens, the earth, the sea, and all that in them is, bless thee. And let them praise thee and glorify thee for ever.

Glory be to the Father, and to the Son, and to the Holy Ghost. And worthy to be praised and glorious for ever.

As it was in the beginning, is now, and ever shall be: world without end. Amen. And worthy to be praised and glorious for ever.

Blessed art thou, O Lord God of our fathers. And worthy to be praised and glorious for ever.
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\collect
\lett{O}{God,} who to the three children didst assuage the flames of fire: mercifully grant; that the flames of sin may not kindle upon us thy servants. Through.
\begin{rubric}
    Additional Collects are as on Trinity III (p. \pageref{TrinityIIICollect}).
\end{rubric}

\readingcitation{Epistle}{Hebrews 9:2}
\lett{B}{rethren:} There was a tabernacle made; the first, wherein was the candlestick, and the table, and the shewbread; which is called the sanctuary. And after the second veil, the tabernacle which is called the Holiest of all; Which had the golden censer, and the ark of the covenant overlaid round about with gold, wherein was the golden pot that had manna, and Aaron's rod that budded, and the tables of the covenant; And over it the cherubims of glory shadowing the mercyseat; of which we cannot now speak particularly. Now when these things were thus ordained, the priests went always into the first tabernacle, accomplishing the service of God. But into the second went the high priest alone once every year, not without blood, which he offered for himself, and for the errors of the people: The Holy Ghost this signifying, that the way into the holiest of all was not yet made manifest, while as the first tabernacle was yet standing: Which was a figure for the time then present, in which were offered both gifts and sacrifices, that could not make him that did the service perfect, as pertaining to the conscience; Which stood only in meats and drinks, and divers washings, and carnal ordinances, imposed on them until the time of reformation. But Christ being come an high priest of good things to come, by a greater and more perfect tabernacle, not made with hands, that is to say, not of this building; Neither by the blood of goats and calves, but by his own blood he entered in once into the holy place, having obtained eternal redemption for us.
\tract{O praise the Lord, all ye heathen: praise him, all ye nations. ℣. For his merciful kindness is ever more and more towards us: and the truth of the Lord endureth for ever.}
\readingcitation{Gospel}{Luke 13:6}
\lett{A}{t that time:} Jesus spake this parable unto the multitudes: A certain man had a fig tree planted in his vineyard; and he came and sought fruit thereon, and found none. Then said he unto the dresser of his vineyard, Behold, these three years I come seeking fruit on this fig tree, and find none: cut it down; why cumbereth it the ground? And he answering said unto him, Lord, let it alone this year also, till I shall dig about it, and dung it: And if it bear fruit, well: and if not, then after that thou shalt cut it down. And he was teaching in one of the synagogues on the sabbath. And, behold, there was a woman which had a spirit of infirmity eighteen years, and was bowed together, and could in no wise lift up herself. And when Jesus saw her, he called her to him, and said unto her, Woman, thou art loosed from thine infirmity. And he laid his hands on her: and immediately she was made straight, and glorified God. And the ruler of the synagogue answered with indignation, because that Jesus had healed on the sabbath day, and said unto the people, There are six days in which men ought to work: in them therefore come and be healed, and not on the sabbath day. The Lord then answered him, and said, Thou hypocrite, doth not each one of you on the sabbath loose his ox or his ass from the stall, and lead him away to watering? And ought not this woman, being a daughter of Abraham, whom Satan hath bound, lo, these eighteen years, be loosed from this bond on the sabbath day? And when he had said these things, all his adversaries were ashamed: and all the people rejoiced for all the glorious things that were done by him.
\offertory{O Lord God of my salvation, I have cried day and night before thee: let my prayer enter into thy presence, O Lord.}

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\secret
\lett{G}{rant,} we beseech thee, almighty God: that the gift which we offer in the sight of thy majesty may both obtain for us the grace of devotion, and effectually bring us to everlasting happiness. Through.
\begin{rubric}
    Additional Secrets are as on Trinity III (p. \pageref{TrinityIIISecret}).
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\communion{In the seventh month ye shall celebrate a feast, when I made the children of Israel to dwell in booths, when I brought them out of the land of Egypt, I the Lord your God.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that thy sacraments may accomplish in us that which they contain: that those things which we now offer in outward fashion, we may receive in verity and truth. Through.
\begin{rubric}
    Additional Postcommunions are as on Trinity III (p. \pageref{TrinityIIIPostcommunion}).
\end{rubric}


\subby{Seven Sorrows of the Blessed Virgin Mary}
\feastday{Seven Sorrows}
\begin{inhead}
    {Second Class Double with a Common Octave\\
15 September}
\end{inhead}
\feastdate{15}{September}
%S-R: N/A (Sarum has very few propers)

\introit
\lett{T}{here} stood by the Cross of Jesus his Mother, and his Mother's sister, Mary the wife of Cleophas, and Salome, and Mary Magdalene. ℣. Woman, behold thy son: said Jesus; and to the disciple: Behold thy Mother.
\collect
\lett{O}{God,} in whose passion according to the prophecy of Simeon, a sword of sorrow did pierce the most sweet soul of the glorious Virgin Mother Mary: mercifully grant; that we, who devoutly call to mind her sorrows, may obtain the blessed effects of thy passion. Who livest.
\readingcitation{Epistle}{Judith 13:17}
%The KJV is used, for it is very different from the Vulgate in regard to its letters but the same in regard to content and is faithful to the letters and meaning of the LXX.
\lett{B}{lessed} be thou, O our God, which hast this day brought to nought the enemies of thy people. O daughter, blessed art thou of the most high God above all the women upon the earth; and blessed be the Lord God, which hath created the heavens and the earth, which hath directed thee to the cutting off of the head of the chief of our enemies. For this thy confidence shall not depart from the heart of men, which remember the power of God for ever. And God turn these things to thee for a perpetual praise, to visit thee in good things because thou hast not spared thy life for the affliction of our nation, but hast revenged our ruin, walking a straight way before our God.
\gradall{Mournful and weeping art thou, O Virgin Mary, standing by the Cross of the Lord Jesus, thy Son, the Redeemer. ℣. Virgin, Mother of God, he whom the whole world containeth not, endureth this torment of the cross, the author of life made man.}{Alleluia, alleluia, ℣. There stood mournful by the Cross of our Lord Jesus Christ, holy Mary, Queen of heaven and Lady of the world.}
\subbysub{Sequence}\label{dolorosa}
\begin{multicols}{2}
\lett{A}{t} the Cross her station keeping.

\secondline{Stood the mournful Mother weeping.}

\thirdline{Where he hung, her dying Son,\\}

Through her soul, of joy bereaved

Torn with anguish, deeply grieved,

Lo! the piercing sword hath run.\\

O, how sad and sore distressed,

Then was she, that Mother blessed

Of the sole-begotten One!\\

Torn with grief and desolation,

Mother meek, the bitter Passion

Saw she of her glorious Son.\\

%\columnbreak

Who, on Christ's dear Mother gazing.

Bowed with sorrow so amazing,

Born of woman, would not weep?\\

Who, on Christ's dear Mother thinking.

With her Son in sorrow sinking.

Would not share her sadness deep?\\

For his people's sins chastised.

She her Jesus saw despised.

Saw him by the scourges rent.\\

Saw her own sweet Offspring taken,

And in death by all forsaken.

While his spirit forth he sent.\\

Mother, fount of love o'erflowing.

Ah, that I, thy sorrow knowing.

In thy grief may mourn with thee.\\

That my heart, fresh ardour gaining,

Love of Christ my God attaining,

Unto him may pleasing be.\\

Holy Mother, be there written

Every wound of Jesus smitten

In my heart, and there remain.\\

As thy Son through tribulation

Deigned to purchase my salvation.

Let me share with thee the pain.\\

Let me weep with thee beside him

For the sins which crucified him.

While my life remains in me.\\

Take beneath the Cross my station.

Share with thee thy desolation,

Humbly this I ask of thee.\\

Virgin, virgins all excelling,

Spurn me not, my prayer repelling;

Make me weep and mourn with thee.\\

So Christ's death within me bearing,

Let me, in his Passion sharing.

Keep his wounds in memory.\\

Let thy Son's wounds penetrate me:

Let the Cross inebriate me

And his own most precious blood.\\

Lest in flames I burn and perish.

On the judgment day O cherish

And defend me, Virgin good.\\

Christ, whene'er the world shall leave me,

Through thy Mother then receive me

To the palm of victory.\\

When the bonds of flesh are riven.

Glory to my soul be given

In thy Paradise with thee.

Amen. Alleluia.
\end{multicols}
\readingcitation{Gospel}{John 19:25}
\lett{A}{t that time:} There stood by the cross of Jesus his mother, and his mother's sister, Mary the wife of Cleophas, and Mary Magdalene. When Jesus therefore saw his mother, and the disciple standing by, whom he loved, he saith unto his mother, Woman, behold thy son! Then saith he to the disciple, Behold thy mother! And from that hour that disciple took her unto his own home.

\offertory{Remember, O Virgin, Mother of God, when thou standest in the sight of the Lord, that thou speak good things for us, and that he may turn away his indignation from us.}
\secret
\lett{W}{e} offer unto thee prayers and sacrifices, O Lord Jesu Christ, humbly beseeching thee: that we, who in our prayers recall the Piercing of the most sweet spirit of blessed Mary, thy Mother; may by the manifold and most loving intercession of her and of the Saints her companions beneath the Cross, through the merits of thy death, be made worthy of a portion with the blessed. Who livest.
\communion{Happy the heart of the blessed Virgin Mary, which without death gained the palm of martyrdom beneath the Cross of the Lord.}
\postcommunion
\lett{O}{Lord} Jesu Christ, let the sacrifices which we have received in devout remembrance of the Piercing of thy Virgin Mother: obtain for us of thy mercy the effect of every good and saving gift. Who livest.


\subby{St. Matthew}
\feastday{St. Matthew}
\begin{inhead}
    {Second Class Double\\
21 September}
\end{inhead}
\feastdate{21}{September}
%S-R: N/A
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} Thou hast given an heritage unto those that fear thy Name, O Lord.%\vr{Ps. 61:5}

\introit
\lett{T}{he} mouth of the just is exercised in wisdom, and his tongue will be talking of judgment: the law of his God is in his heart. \textit{Ps.} Fret not thyself because of the ungodly: neither be thou envious against the evil doers.
\collect
%1928 BCP:
%\lett{O}{almighty} God, who by thy blessed Son didst call Matthew from the receipt of custom to be an Apostle and Evangelist; Grant us grace to forsake all covetous desires, and inordinate love of riches, and to follow the same thy Son Jesus Christ, who liveth and reigneth with thee and the Holy Ghost, one God, world without end. \textit{Amen.}
\lett{M}{ay} we be assisted, O Lord, by the prayers of blessed Matthew, the Apostle and Evangelist: that those things, which of ourselves we cannot obtain, may be vouchsafed unto us by his intercession. Through.

\readingcitation{Epistle}{2 Corinthians 4:1}
\lett{B}{rethren:} Seeing we have this ministry, as we have received mercy, we faint not; But have renounced the hidden things of dishonesty, not walking in craftiness, nor handling the word of God deceitfully; but by manifestation of the truth commending ourselves to every man's conscience in the sight of God. But if our gospel be hid, it is hid to them that are lost: In whom the god of this world hath blinded the minds of them which believe not, lest the light of the glorious gospel of Christ, who is the image of God, should shine unto them. For we preach not ourselves, but Christ Jesus the Lord; and ourselves your servants for Jesus' sake. For God, who commanded the light to shine out of darkness, hath shined in our hearts, to give the light of the knowledge of the glory of God in the face of Jesus Christ.

\gradall{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed.}{Alleluia, alleluia. ℣. The glorious company of the Apostles praise thee, O Lord. Alleluia.}

\readingcitation{Gospel}{Matthew 9:9}
\lett{A}{t that time:} As Jesus passed forth from thence, he saw a man, named Matthew, sitting at the receipt of custom: and he saith unto him, Follow me. And he arose, and followed him. And it came to pass, as Jesus sat at meat in the house, behold, many publicans and sinners came and sat down with him and his disciples. And when the Pharisees saw it, they said unto his disciples, Why eateth your Master with publicans and sinners? But when Jesus heard that, he said unto them, They that be whole need not a physician, but they that are sick. But go ye and learn what that meaneth, I will have mercy, and not sacrifice: for I am not come to call the righteous, but sinners to repentance.
\offertory{Thou hast set, O Lord, a crown of pure gold upon his head: he asked life of thee, and thou gavest it him, alleluia.}
\secret
\lett{O}{Lord,} who dost instruct thy Church with the wondrous preaching of the blessed Apostle and Evangelist Matthew: we beseech thee, that by his supplication her oblation may be commended unto thee. Through.
\communion{His honour is great in thy salvation: glory and great worship shalt thou lay upon him, O Lord.}
\postcommunion
\lett{O}{Lord,} who hast bestowed on us thy sacraments, we beseech thee: that through the mediation of blessed Matthew, thine Apostle and Evangelist, the mysteries which we have celebrated may be profitable for our healing. Through.


%MANUAL ADJUSTMENT:
\clearpage
\subby{St. Michael}
\feastday{Michaelmas}
\begin{inhead}
    {First Class Double\\
29 September}
\end{inhead}
\feastdate{29}{September}
%S-R: N/A
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} The smoke of the incense ascended up before God out of the Angel's hand.

\introit
\lett{O}{praise} the Lord, all ye Angels of his: ye that excel in strength ye that fulfil his commandment, and hearken unto the voice of his words. \textit{Ps.} Praise the Lord, O my soul; and all that is within me praise his holy name

\collect
\lett{O}{everlasting} God, who hast ordained and constituted the services of Angels and men in a wonderful order; Mercifully grant that, as thy holy Angels always do thee service in heaven, so, by thy appointment, they may succour and defend us on earth. Through.

\readingcitation{Epistle}{Revelation 12:7}
\lett{T}{here} was war in heaven: Michael and his angels fought against the dragon; and the dragon fought and his angels, And prevailed not; neither was their place found any more in heaven. And the great dragon was cast out, that old serpent, called the Devil, and Satan, which deceiveth the whole world: he was cast out into the earth, and his angels were cast out with him. And I heard a loud voice saying in heaven, Now is come salvation, and strength, and the kingdom of our God, and the power of his Christ: for the accuser of our brethren is cast down, which accused them before our God day and night. And they overcame him by the blood of the Lamb, and by the word of their testimony; and they loved not their lives unto the death. Therefore rejoice, ye heavens, and ye that dwell in them. Woe to the inhabiters of the earth and of the sea! for the devil is come down unto you, having great wrath, because he knoweth that he hath but a short time.

\gradall{O praise the Lord, all ye Angels of his; ye that excel in strength, ye that fulfil his commandment. ℣. Praise the Lord, O my soul; and all that is within me praise his holy name.}{Alleluia, alleluia. ℣. Holy Michael, Archangel, defend us in the battle: that we perish not in the dreadful judgment. Alleluia.}

\readingcitation{Gospel}{Matthew 18:1}
\lett{A}{t that time:} The disciples came unto Jesus, saying, Who is the greatest in the kingdom of heaven? And Jesus called a little child unto him, and set him in the midst of them, And said, Verily I say unto you, Except ye be converted, and become as little children, ye shall not enter into the kingdom of heaven. Whosoever therefore shall humble himself as this little child, the same is greatest in the kingdom of heaven. And whoso shall receive one such little child in my name receiveth me. But whoso shall offend one of these little ones which believe in me, it were better for him that a millstone were hanged about his neck, and that he were drowned in the depth of the sea. Woe unto the world because of offences! for it must needs be that offences come; but woe to that man by whom the offence cometh! Wherefore if thy hand or thy foot offend thee, cut them off, and cast them from thee: it is better for thee to enter into life halt or maimed, rather than having two hands or two feet to be cast into everlasting fire. And if thine eye offend thee, pluck it out, and cast it from thee: it is better for thee to enter into life with one eye, rather than having two eyes to be cast into hell fire. Take heed that ye despise not one of these little ones; for I say unto you, That in heaven their angels do always behold the face of my Father which is in heaven.

\offertory{An Angel stood by the altar of the temple, having a golden censer in his hand, and there was given unto him much incense; and the smoke of the incense ascended up before God, alleluia.}

\secret
\lett{W}{e} offer thee, O Lord, sacrifices of praise, humbly beseeching thee: that by the prayers of the Angels interceding for us, thou wouldest both graciously accept the same, and grant that they may avail to our salvation. Through.

\communion{O all ye Angels of the Lord, bless ye the Lord: sing ye praises, and magnify him above all for ever.}

\postcommunion
\lett{O}{Lord,} forasmuch as we put our trust in the intercession of thy blessed Archangel Michael; we humbly beseech thee; that those things which we touch with our lips we may likewise receive into our hearts. Through.


\subby{Holy Guardian Angels}
\feastday{Guardian Angels}
\begin{inhead}
	{Greater Double\\
		2 October}
\end{inhead}
\feastdate{2}{October}
\par\noindent
\textit{Opening Sentence.} O praise the Lord, ye Angels of his. Ye that excel in strength, and hearken unto the voice of his words.

\introit
\lett{O}{praise} the Lord, all ye Angels of his: ye that excel in strength, ye that fulfil his commandment, and hearken unto the voice of his words. \textit{Ps.} Praise the Lord, O my soul: and that is within me praise his holy name.

\collect
\lett{O}{God,} who of thy ineffable providence dost vouchsafe to send thy holy Angels to be our guardians: grant unto us thy humble servants; that we may ever both be defended by their protection, and rejoice in their everlasting fellowship. Through.

\readingcitation{Epistle}{Exodus 23:20}
\lett{T}{hus} saith the Lord God: Behold, I send an Angel before thee, to keep thee in the way, and to bring thee into the place which I have prepared. Beware of him, and obey his voice, provoke him not; for he will not pardon your transgressions: for my name is in him. But if thou shalt indeed obey his voice, and do all that I speak; then I will be an enemy unto thine enemies, and an adversary unto thine adversaries. For mine Angel shall go before thee.

\gradual{God shall give his Angels charge over thee, to keep thee in all thy ways. ℣. They shall bear thee in their hands, that thou hurt not thy foot against a stone.}

\alleluia{Alleluia, alleluia. ℣. O praise the Lord, all ye his hosts: ye servants of his that do his pleasure. Alleluia.}

\begin{rubric}
	In Votive Masses after Septuagesima, the Tract---and in Eastertide the Greater Alleluia---are as in Votives of the Holy Angels (p. \pageref{MassOfTheAngels}).
\end{rubric}

\readingcitation{Gospel}{Matthew 18:1}
\lett{A}{t that time:} The disciples came unto Jesus, saying, Who is the greatest in the kingdom of heaven? And Jesus called a little child unto him, and set him in the midst of them, and said, Verily I say unto you, Except ye be converted, and become as little children, ye shall not enter into the kingdom of heaven. Whosoever therefore shall humble himself as this little child, the same is greatest in the kingdom of heaven. And whoso shall receive one such little child in my name receiveth me. But whoso shall offend one of these little ones which believe in me, it were better for him that a millstone were hanged about his neck, and that he were drowned in the depth of the sea. Woe unto the world because of offences! for it must needs be that offences come; but woe to that man by whom the offence cometh! Wherefore if thy hand or thy foot offend thee, cut them off, and cast them from thee: it is better for thee to enter into life halt or maimed, rather than having two hands or two feet to be cast into everlasting fire. And if thine eye offend thee, pluck it out, and cast it from thee: it is better for thee to enter into life with one eye, rather than having two eyes to be cast into hell fire. Take heed that ye despise not one of these little ones; for I say unto you, That in heaven their angels do always behold the face of my Father which is in heaven.

\offertory{O praise the Lord, all ye Angels of his: ye servants of his, that fulfil his commandment, and hearken unto the voice of his words.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\secret
\lett{R}{eceive,} O Lord, the gifts which we present in veneration of thy holy Angels: and mercifully grant; that by their continual protection we may be delivered from present dangers and attain unto everlasting life. Through.

\communion{O all ye Angels of the Lord, bless ye the Lord: sing ye praises, and magnify him above all for ever.}

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\postcommunion
\lett{O}{Lord,} who on this festival of thy holy Angels, hast suffered us to receive in gladness thy heavenly mysteries: we beseech thee; that by their protection we may ever be delivered from the snares of our enemies, and defended against all adversities. Through.

\subby{Holy Rosary of the Blessed Virgin Mary}
\feastday{Holy Rosary}
\begin{inhead}
	{Greater Double\\
		7 October}
\end{inhead}
\feastdate{7}{October}
\begin{rubric}
	In Low Mass, Commemoration of St. Mark of Rome and Sts. Sergius, Bacchus, Marcellus, \& Apuleius is made, with the Prayers as below.
	\textsc{Note,} The Creed is said.
\end{rubric}

\introit
\lett{R}{ejoice} we all in the Lord, keeping feast day in honour of the blessed Virgin Mary: in whose solemnity the Angels rejoice and glorify the Son of God. \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.

\collect
\lett{O}{God,} whose only-begotten Son by his life, death, and resurrection hath purchased for us the rewards of everlasting salvation: grant, we beseech thee; that we, who meditate upon these mysteries in the most sacred Rosary of the blessed Virgin Mary, may both imitate those things which they set forth, and attain unto those things which they promise. Through.
\needspace{4\baselineskip}
\lett{G}{raciously} hear our prayers, O Lord: and at the intercession of blessed Mark, thy Confessor and Bishop, mercifully grant us pardon and peace. Through.
\needspace{4\baselineskip}
\lett{M}{ay} the blessed merits of thy holy Martyrs Sergius, Bacchus, Marcellus, and Apuleius uphold us, O Lord: and ever make us fervent in thy love. Through.


\readingcitation{Epistle}{Proverbs 8:22}
\lett{T}{he} Lord possessed me in the beginning of his way, before his works of old. I was set up from everlasting, from the beginning, or ever the earth was. When there were no depths, I was brought forth. Now therefore hearken unto me, O ye children: for blessed are they that keep my ways. Hear instruction, and be wise, and refuse it not. Blessed is the man that heareth me, watching daily at my gates, waiting at the posts of my doors. For whoso findeth me findeth life, and shall obtain favour of the Lord.

\gradall{Because of the word of truth, of meekness, and righteousness, and thy right hand shall teach thee terrible things. ℣. Hearken, O daughter, and consider, incline thine ear: so shall the King have pleasure in thy beauty.}{Alleluia, alleluia. ℣. The solemnity of the glorious Virgin Mary of the seed of Abraham, sprung from the tribe of Juda, of David's noble stock. Alleluia.}

\readingcitation{Gospel}{Luke 1:26}
\lett{A}{t that time:} The angel Gabriel was sent from God unto a city of Galilee, named Nazareth, to a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin’s name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name \divineName{Jesus}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: and he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word.

\offertory{In me is all grace of the way and of truth: in me is all hope of life and of virtue: I have budded forth as a rose growing by the brooks of water.}

\secret
\lett{G}{rant,} we beseech thee, O Lord, that we may be enabled worthily to offer these our gifts: and through the mysteries of the most sacred Rosary so to recall the life, the passion and the glory of thy only-begotten Son; that we may be made worthy of his promises. Who liveth and reigneth with thee.
\needspace{4\baselineskip}
\lett{G}{rant,} O Lord, that like as thy dedicated people do acknowledge that in tribulation they have been succoured by the merits of thy Saints: so this oblation, which they offer unto thee in honour of the same, may be acceptable in thy sight. Through.
\needspace{4\baselineskip}
\lett{W}{e} beseech thee, O Lord, that, through the meritorious supplication of thy Saints: this sacrifice, which we offer, may obtain for us the favour of thy majesty. Through.

\communion{Put forth flowers as a lily, spread abroad a sweet smell, and bring forth leaves in grace, sing a song of praise, and bless the Lord in all his works.}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who celebrate the Rosary of thy most holy Mother, may in such wise be succoured by her prayers: that we may be partakers of the power of the mysteries which we revere; and obtain the effect of the sacraments which we have received. Who livest.
\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, O Lord, that thy faithful people may ever rejoice in the veneration of thy Saints: and be defended by their perpetual supplication. Through.
%\needspace{4\baselineskip}
\lett{G}{rant,} O Lord, that, being strengthened by the sacraments which we have received: we may, at the intercession of thy holy Martyrs, Sergius, Bacchus, Marcellus, and Apuleius, fight against all iniquity, and be defended by thy heavenly armour. Through.


\subby{Motherhood of the Blessed Virgin Mary}
\feastday{Motherhood}
\begin{inhead}
    {Second Class Double\\
11 October}
\end{inhead}
\feastdate{11}{October}

\introit
\lett{B}{ehold,} a Virgin shall conceive, and bear a Son, and shall call his name Emmanuel. \textit{Ps.} O sing unto the Lord a new song, for he hath done marvellous things.
\collect
\lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the Blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we, who believe her to be truly the Mother of God, may be aided by her intercession in thy sight. Through the same.
\readingcitation{Epistle}{Ecclesiasticus 24:17}
%\lett{A}{s} the vine brought I forth pleasant savour, and my flowers are the fruit of honour and riches. I am the mother of fair love, and fear, and knowledge, and holy hope: I therefore, being eternal, am given to all my children which are named of him. Come unto me, all ye that be desirous of me, and fill yourselves with my fruits. For my memorial is sweeter than honey, and mine inheritance than the honeycomb. They that eat me shall yet be hungry, and they that drink me shall yet be thirsty. He that obeyeth me shall never be confounded, and they that work by me shall not do amiss.
%The interpolation from the Vulgate is omitted in the KJV in favour of the LXX, and is thus followed.
%RV:
\lett{A}{s} the vine I put forth grace; And my flowers are the fruit of glory and riches. Come unto me, ye that are desirous of me, And be ye filled with my produce. For my memorial is sweeter than honey, And mine inheritance than the honeycomb. They that eat me shall yet be hungry; And they that drink me shall yet be thirsty. He that obeyeth me shall not be ashamed; And they that work in me shall not do amiss. 
\gradall{There shall come forth a rod out of the stem of Jesse: and a branch shall rise up out of his roots. ℣. And the Spirit of the Lord shall rest upon him.}{Alleluia, alleluia. ℣. Virgin, Mother of God, he whom the whole world containeth not, being made man lay hid in thy womb. Alleluia.}
\readingcitation{Gospel}{Luke 2:43}
\lett{A}{t that time:} As they returned, the child Jesus tarried behind in Jerusalem; and Joseph and his mother knew not of it. But they, supposing him to have been in the company, went a day's journey; and they sought him among their kinsfolk and acquaintance. And when they found him not, they turned back again to Jerusalem, seeking him. And it came to pass, that after three days they found him in the temple, sitting in the midst of the doctors, both hearing them, and asking them questions. And all that heard him were astonished at his understanding and answers. And when they saw him, they were amazed: and his mother said unto him, Son, why hast thou thus dealt with us? behold, thy father and I have sought thee sorrowing. And he said unto them, How is it that ye sought me? wist ye not that I must be about my Father's business? And they understood not the saying which he spake unto them. And he went down with them, and came to Nazareth, and was subject unto them.
\offertory{When Mary his Mother was espoused to Joseph, she was found with child of the Holy Ghost.}
\secret
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, Mother of thine only-begotten Son, may this oblation avail for our prosperity and peace, both now and for ever. Through the same.
\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father.}
\postcommunion
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary, the Virgin Mother of God, make us partakers of thy heavenly healing. Through the same.


\subby{Our Lady of Walsingham}
\feastday{Walsingham}
\begin{inhead}
	{Double\\
		15 October}
\end{inhead}
\feastdate{15}{October}
\par\noindent
%Daily Office from the AOB:
\textit{Opening Sentence.} How dreadful is this place! this is none other but the house of God, and this is the gate of heaven.%\vr{Gn 28:17}

%From the LAP BCP:
\introit
\lett{O}{how} dreadful is this place! this is none other but the house of God, and this is the gate of heaven. \textit{Ps.} O how amiable are thy dwellings, thou Lord of hosts! Blessed are they that dwell in thy house, they will be alway praising thee.

\collect
\lett{O}{God,} who in the blessed Virgin Mary didst make a fit dwelling-place for thy Son: grant, we beseech thee, that we who honour her shrine at Walsingham may also become temples of thy Holy Ghost. Through the same.

\readingcitation{Epistle}{Ecclesiasticus 24:8}
%KJV:
%\lett{T}{he} Creator of all things gave me a commandment, and he that made me caused my tabernacle to rest, and said, Let thy dwelling be in Jacob, and thine inheritance in Israel. He created me from the beginning before the world, and I shall never fail. In the holy tabernacle I served before him; and so was I established in Sion. Likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance. I was exalted like a cedar in Libanus, and as a cypress tree upon the mountains of Hermon. I was exalted like a palm tree in En-gaddi, and as a rose plant in Jericho, as a fair olive tree in a pleasant field, and grew up as a plane tree by the water. I am the mother of fair love, and fear, and knowledge, and holy hope: I therefore, being eternal, am given to all my children which are named of him. Come unto me, all ye that be desirous of me, and fill yourselves with my fruits. For my memorial is sweeter than honey, and mine inheritance than the honeycomb.
%RV:
\lett{T}{he} Creator of all things gave me a commandment; And he that created me made my tabernacle to rest, And said, Let thy tabernacle be in Jacob, And thine inheritance in Israel. He created me from the beginning before the world; And to the end I shall not fail. In the holy tabernacle I ministered before him; And so was I established in Sion. In the beloved city likewise he gave me rest; And in Jerusalem was my authority.  And I took root in a people that was glorified, Even in the portion of the Lord's own inheritance. I was exalted like a cedar in Libanus, And as a cypress tree on the mountains of Hermon. I was exalted like a palm tree on the sea shore, And as rose plants in Jericho, And as a fair olive tree in the plain; And I was exalted as a plane tree. And my flowers are the fruit of glory and riches. Come unto me, ye that are desirous of me, And be ye filled with my produce. For my memorial is sweeter than honey, And mine inheritance than the honeycomb.

\gradall{My soul hath a desire and longing to enter into the courts of the Lord. ℣. Yea, the sparrow hath found her an house, and the swallow a nest, where she may lay her young: even thy altars, O Lord of hosts, my King and my God.}{Alleluia, alleluia. ℣. I had rather be a door-keeper in the house of my God: than to dwell in the tents of ungodliness. Alleluia.}

\readingcitation{Gospel}{Luke 2:25}
\lett{A}{t that time:} There was a man in Jerusalem, whose name was Simeon; and the same man was just and devout, waiting for the consolation of Israel: and the Holy Ghost was upon him. And it was revealed unto him by the Holy Ghost, that he should not see death, before he had seen the Lord’s Christ. And he came by the Spirit into the temple: and when the parents brought in the child Jesus, to do for him after the custom of the law, then took he him up in his arms, and blessed God, and said, Lord, now lettest thou thy servant depart in peace, according to thy word: for mine eyes have seen thy salvation, which thou hast prepared before the face of all people; a light to lighten the Gentiles, and the glory of thy people Israel. And Joseph and his mother marvelled at those things which were spoken of him. And Simeon blessed them, and said unto Mary his mother, Behold, this child is set for the fall and rising again of many in Israel; and for a sign which shall be spoken against; (yea, a sword shall pierce through thy own soul also,) that the thoughts of many hearts may be revealed. And there was one Anna, a prophetess, the daughter of Phanuel, of the tribe of Aser: she was of a great age, and had lived with an husband seven years from her virginity; and she was a widow of about fourscore and four years, which departed not from the temple, but served God with fastings and prayers night and day. And she coming in that instant gave thanks likewise unto the Lord, and spake of him to all them that looked for redemption in Jerusalem. And when they had performed all things according to the law of the Lord, they returned into Galilee, to their own city Nazareth. And the child grew, and waxed strong in spirit, filled with wisdom: and the grace of God was upon him.

\offertory{And Jesus went down with Mary and Joseph, and came to Nazareth, and was subject unto them. But his Mother kept all these sayings in her heart. And Jesus increased in wisdom and stature, and in favour with God and man.}

\secret
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever-Virgin, Mother of thine only-begotten Son, may this oblation avail for our prosperity and peace, both now and for ever. Through the same.

\communion{Blessed is the womb of the Virgin Mary, that bare the Son of the everlasting Father.}

\postcommunion
\lett{O}{Lord,} let this holy Communion cleanse us from every guilty stain: that at the intercession of the blessed Virgin Mary, Mother of God, we may be made partakers thereby of thy healing unto life eternal. Through the same.


\subby{St. Luke}
\feastday{St. Luke}
\begin{inhead}
    {Second Class Double\\
18 October}
\end{inhead}
\feastdate{18}{October}
%S-R: N/A

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting and mine up-rising.
\collect
%1928:
%\lett{A}{lmighty} God, who didst inspire thy servant Saint Luke the Physician, to set forth in the Gospel the love and healing power of thy Son; Manifest in thy Church the like power and love, to the healing of our bodies and our souls. Through the same.
%1892:
%\lett{A}{lmighty} God, who calledst Luke the Physician, whose praise is in the Gospel, to an Evangelist, and Physician of the soul: May it please thee that, by the wholesome medicines of the doctrine delivered by him, all the diseases of our souls may be healed; through the merits of thy Son Jesus Christ our Lord. Who liveth.
%Sarum/Roman:
\lett{W}{e} beseech thee, O Lord: that thy holy Evangelist Luke may intercede for us: who for the honour of thy name continually bare in his own body the mortification of the Cross. Through.

\readingcitation{Epistle}{2 Timothy 4:5}
\lett{D}{early beloved:} Watch thou in all things, endure afflictions, do the work of an evangelist, make full proof of thy ministry. For I am now ready to be offered, and the time of my departure is at hand. I have fought a good fight, I have finished my course, I have kept the faith: Henceforth there is laid up for me a crown of righteousness, which the Lord, the righteous judge, shall give me at that day: and not to me only, but unto all them also that love his appearing. Do thy diligence to come shortly unto me: For Demas hath forsaken me, having loved this present world, and is departed unto Thessalonica; Crescens to Galatia, Titus unto Dalmatia. Only Luke is with me. Take Mark, and bring him with thee: for he is profitable to me for the ministry. And Tychicus have I sent to Ephesus. The cloke that I left at Troas with Carpus, when thou comest, bring with thee, and the books, but especially the parchments. Alexander the coppersmith did me much evil: the Lord reward him according to his works: Of whom be thou ware also; for he hath greatly withstood our words.

\gradall{Their sound is gone out into all lands: and their words into the ends of the world. ℣. The heavens declare the glory of God: and the firmament sheweth his handy-work.}{Alleluia, alleluia. ℣. I have chosen you out of the world, that ye should go and bring forth fruit: and that your fruit should remain. Alleluia.}

\readingcitation{Gospel}{Luke 10:1}
\lett{A}{t that time:} The Lord appointed other seventy also, and sent them two and two before his face into every city and place, whither he himself would come. Therefore said he unto them, The harvest truly is great, but the labourers are few: pray ye therefore the Lord of the harvest, that he would send forth labourers into his harvest. Go your ways: behold, I send you forth as lambs among wolves. Carry neither purse, nor scrip, nor shoes: and salute no man by the way. And into whatsoever house ye enter, first say, Peace be to this house. And if the son of peace be there, your peace shall rest upon it: if not, it shall turn to you again. And in the same house remain, eating and drinking such things as they give: for the labourer is worthy of his hire.

\offertory{Right honourable are thy friends unto me, O God: right well is their princedom established.}
\secret
\lett{G}{rant} us, O Lord, we beseech thee, by thy heavenly gifts to serve thee in freedom of spirit: that, through the intercession of thy blessed Evangelist Luke, the oblations which we present may work in us for our healing and glory. Through.
\communion{Ye which have followed me shall sit upon thrones, judging the twelve tribes of Israel.}
\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that those things which we have receive from thy holy altar may, through the prayers of thy blessed Evangelist Luke, sanctify our souls, and avail for our protection. Through.

\subby{St. Raphael}
\feastday{St. Raphael}
\begin{inhead}
	{Greater Double\\
		24 October}
\end{inhead}
\feastdate{24}{October}
%AOB:
\par\noindent
\textit{Opening Sentence.} And there appeared an angel unto him from heaven strengthening him.

\introit
\lett{O}{praise} the Lord, ye Angels of his, ye that excel in strength: ye that fulfil his commandment, and hearken unto the voice of his words. \textit{Ps.} Praise the Lord, O my soul: and all that is within me praise his holy name.

\collect
\lett{O}{God,} who didst give blessed Raphael the Archangel unto thy servant Tobias for a companion on his way: grant to us thy servants; that we may ever be guarded by his protection and strengthened by his help. Through.

\readingcitation{Epistle}{Tobit 12:7}
%\lett{I}{n those days:} The Angel Raphael said to Tobias: It is good to keep close the secret of a king, but it is honourable to reveal the works of God. Do that which is good, and no evil shall touch you. Prayer is good with fasting and alms and righteousness. A little with righteousness is better than much with unrighteousness. It is better to give alms than to lay up gold: For alms doth deliver from death, and shall purge away all sin. Those that exercise alms and righteousness shall be filled with life: But they that sin are enemies to their own life. Surely I will keep close nothing from you. For I said, It was good to keep close the secret of a king, but that it was honourable to reveal the works of God. Now therefore, when thou didst pray, and Sara thy daughter in law, I did bring the remembrance of your prayers before the Holy One: and when thou didst bury the dead, I was with thee likewise. And when thou didst not delay to rise up, and leave thy dinner, to go and cover the dead, thy good deed was not hid from me: but I was with thee. And now God hath sent me to heal thee and Sara thy daughter in law. I am Raphael, one of the seven holy angels, which present the prayers of the saints, and which go in and out before the glory of the Holy One.
%RV:
\lett{I}{n those days:} The Angel Raphael said to Tobias: It is good to keep close the secret of a king, but to reveal gloriously the works of God. Do good, and evil shall not find you. Good is prayer with fasting and alms and righteousness. A little with righteousness is better than much with unrighteousness. It is better to give alms than to lay up gold: alms doth deliver from death, and it shall purge away all sin. They that do alms and righteousness shall be filled with life; but they that sin are enemies to their own life. Surely I will keep close nothing from you. I have said, It is good to keep close the secret of a king, but to reveal gloriously the works of God. And now, when thou didst pray, and Sarah thy daughter in law, I did bring the memorial of your prayer before the Holy One: and when thou didst bury the dead, I was with thee likewise.  And when thou didst not delay to rise up, and leave thy dinner, that thou mightest go and cover the dead, thy good deed was not hid from me: but I was with thee. And now God did send me to heal thee and Sarah thy daughter in law. I am Raphael, one of the seven holy angels, which present the prayers of the saints, and go in before the glory of the Holy One. 

\gradall{The Angel of the Lord, Raphael, took the devil and bound him. ℣. Great is our Lord, and great is his power.}{Alleluia, alleluia. ℣. In the sight of the Angels will I sing praise unto thee: I will worship toward thy holy temple and praise thy name, O Lord. Alleluia.}

\readingcitation{Gospel}{John 5:1}
\lett{A}{t that time:} There was a feast of the Jews: and Jesus went up to Jerusalem. Now there is at Jerusalem by the sheep market a pool, which is called in the Hebrew tongue Bethesda, having five porches. In these lay a great multitude of impotent folk, of blind, halt, withered, waiting for the moving of the water. For an angel went down at a certain season into the pool, and troubled the water: whosoever then first after the troubling of the water stepped in was made whole of whatsoever disease he had.

\offertory{An Angel stood by the altar of the temple, having a golden censer in his hand, and there was given unto him much incense: and the smoke of the incense ascended up before God.}

\secret
\lett{W}{e} offer thee, O Lord, sacrifices of praise, humbly beseeching thee: that, by the prayers of the Angels interceding for us, thou wouldest both graciously accept the same, and grant that they may avail to our salvation. Through.

\communion{O all ye Angels of the Lord, bless ye the Lord: sing ye praises, and magnify him above all for ever.}

\postcommunion
\lett{V}{ouchsafe,} O Lord God, to direct thy holy Archangel Raphael to our succour: that, as we believe him ever to stand before thy Majesty, so he may present our unworthy supplications for thy blessing. Through.


\subby{Sts. Simon and Jude}
\feastday{Sts. Simon \& Jude}
\begin{inhead}
    {Second Class Double\\
28 October}
\end{inhead}
\feastdate{28}{October}
%S-R: N/A

\introit
\lett{R}{ight} honourable are thy friends unto me, O God: right well is their princedom established. \textit{Ps.} O Lord, thou hast searched me out, and known me: thou knowest my down-sitting and mine up-rising.
\collect
\lett{A}{lmighty} God, who hast built thy Church upon the foundation of the Apostles and Prophets, Jesus Christ himself being the head corner-stone; Grant us so to be joined together in unity of spirit by their doctrine, that we may be made an holy temple acceptable unto thee. Through.

%CHECK:
%1928 is reverted for innovation:
\readingcitation{Epistle}{Jude 1}
\lett{J}{ude,} the servant of Jesus Christ, and brother of James, to them that are sanctified by God the Father, and preserved in Jesus Christ, and called: Mercy unto you, and peace, and love, be multiplied. Beloved, when I gave all diligence to write unto you of the common salvation, it was needful for me to write unto you, and exhort you that ye should earnestly contend for the faith which was once delivered unto the saints. For there are certain men crept in unawares, who were before of old ordained to this condemnation, ungodly men, turning the grace of our God into lasciviousness, and denying the only Lord God, and our Lord Jesus Christ. I will therefore put you in remembrance, though ye once knew this, how that the Lord, having saved the people out of the land of Egypt, afterward destroyed them that believed not. And the angels which kept not their first estate, but left their own habitation, he hath reserved in everlasting chains under darkness unto the judgment of the great day. Even as Sodom and Gomorrha, and the cities about them in like manner, giving themselves over to fornication, and going after strange flesh, are set forth for an example, suffering the vengeance of eternal fire. Likewise also these filthy dreamers defile the flesh, despise dominion, and speak evil of dignities.

\gradall{Thou shalt make them princes in all lands: they shall remember thy name, O Lord. ℣. Instead of thy fathers thou shalt have children: therefore shall the people give thanks unto thee.}{Alleluia, alleluia. ℣. Right honourable are thy friends, O God: right well is their princedom established. Alleluia.}

\readingcitation{Gospel}{John 15:17}
\lett{A}{t that time:} Jesus said unto his disciples: 17 These things I command you, that ye love one another. If the world hate you, ye know that it hated me before it hated you. If ye were of the world, the world would love his own: but because ye are not of the world, but I have chosen you out of the world, therefore the world hateth you. Remember the word that I said unto you, The servant is not greater than his lord. If they have persecuted me, they will also persecute you; if they have kept my saying, they will keep yours also. But all these things will they do unto you for my name's sake, because they know not him that sent me. If I had not come and spoken unto them, they had not had sin: but now they have no cloke for their sin. He that hateth me hateth my Father also. If I had not done among them the works which none other man did, they had not had sin: but now have they both seen and hated both me and my Father. But this cometh to pass, that the word might be fulfilled that is written in their law, They hated me without a cause. But when the Comforter is come, whom I will send unto you from the Father, even the Spirit of truth, which proceedeth from the Father, he shall testify of me: And ye also shall bear witness, because ye have been with me from the beginning.

\offertory{Their sound is gone out into all lands: and their words into the ends of the world.}
\secret
\lett{W}{e} beseech thee, O Lord: that as we do venerate the everlasting glory of thy holy Apostles Simon and Jude; so, being cleansed by these sacred mysteries, we may more worthily celebrate the same. Through.
\communion{Ye which have followed me shall sit upon thrones, judging the twelve tribes of Israel.}
\postcommunion
\lett{O}{Lord,} through whom we have received these sacraments, we humbly entreat thee: that by the intercession of thy blessed Apostles Simon and Jude, the mysteries which we celebrate for their venerable passion may be profitable for our healing. Through.

\subby{Our Lord Jesus Christ the King}
\feastday{Christ the King}
\begin{inhead}
    {First Class Double\\
Last Sunday in October}
\end{inhead}
\fancyhead[RE,LO]{}
%S-R: N/A
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} All power is given unto me in heaven and in earth.%\vr{Mt 28:18}

\introit
\lett{W}{orthy} is the Lamb that was slain to receive power, and riches, and wisdom, and strength, and honour. To him be glory and dominion for ever and ever. \textit{Ps.} Give the King thy judgments, O God: and thy righteousness unto the King's Son.

\collect
\lett{A}{lmighty} and everlasting God, who in thy beloved Son, the King of all, hast been pleased to make all things new: mercifully grant; that all the families of the Gentiles, dispersed by the wounds of sin, may be made subject to his most gracious governance. Who liveth.
\begin{rubric}
    Commemoration is made of the Sunday occurring.
\end{rubric}

\readingcitation{Epistle}{Colossians 1:12}
\lett{B}{rethren:} We give thanks unto the Father, which hath made us meet to be partakers of the inheritance of the saints in light: Who hath delivered us from the power of darkness, and hath translated us into the kingdom of his dear Son: In whom we have redemption through his blood, even the forgiveness of sins: Who is the image of the invisible God, the firstborn of every creature: For by him were all things created, that are in heaven, and that are in earth, visible and invisible, whether they be thrones, or dominions, or principalities, or powers: all things were created by him, and for him: And he is before all things, and by him all things consist. And he is the head of the body, the church: who is the beginning, the firstborn from the dead; that in all things he might have the preeminence. For it pleased the Father that in him should all fulness dwell; And, having made peace through the blood of his cross, by him to reconcile all things unto himself; by him, I say, whether they be things in earth, or things in heaven, %Added from the 1933 English Missal:
in Christ Jesus our Lord.

\gradual{His dominion shall be from the one sea to the other, and from the flood unto the world's end. ℣. All kings shall fall down before him: all nations shall do him service.}

\alleluia{Alleluia, alleluia. ℣. His dominion is an everlasting dominion, which shall not pass away: and his kingdom that which shall not be destroyed. Alleluia.}

\readingcitation{Gospel}{John 18:33}
\lett{A}{t that time:} Pilate said unto Jesus: Art thou the King of the Jews? Jesus answered him, Sayest thou this thing of thyself, or did others tell it thee of me? Pilate answered, Am I a Jew? Thine own nation and the chief priests have delivered thee unto me: what hast thou done? Jesus answered, My kingdom is not of this world: if my kingdom were of this world, then would my servants fight, that I should not be delivered to the Jews: but now is my kingdom not from hence. Pilate therefore said unto him, Art thou a king then? Jesus answered, Thou sayest that I am a king. To this end was I born, and for this cause came I into the world, that I should bear witness unto the truth. Every one that is of the truth heareth my voice.
\offertory{Desire of me, and I shall give thee the heathen for thine inheritance, and the utmost parts of the earth for thy possession.}
\secret
\lett{W}{e} offer unto thee, O Lord, the victim of man's reconciliation: grant, we beseech thee; that he, whom we offer in these present sacrifices, may himself grant unto all nations the gifts of unity and peace, Jesus Christ thy Son our Lord. Who liveth.
\begin{rubric}
    Commemoration is made of the Sunday occurring.
\end{rubric}
\communion{The Lord remaineth a King for ever: the Lord shall give his people the blessing of peace.}
\postcommunion
\lett{O}{Lord,} who hast bestowed on us the food of immortality, we beseech thee: that we, who glory to fight beneath the banner of Christ the King, may with him be enabled to reign for ever on the heavenly throne. Who liveth.
\begin{rubric}
    Commemoration is made of the Sunday occurring.
\end{rubric}
\begin{rubric}
    The Last Gospel is of the Sunday occurring.
\end{rubric}

\subby{All Hallows}
\feastday{All Hallows}
\begin{inhead}
    {Second Class Double\\
1 November}
\end{inhead}
\feastdate{1}{November}
%S-R: N/A
%\subsubsec{Daily Office Propers}\par\noindent
\par\noindent
\textit{Opening Sentence.} The righteous live for evermore, but their reward is with the Lord.%\vr{Wis. 5:15}

\introit
\lett{R}{ejoice} we all in the Lord, keeping feast day in honour of all the saints: in whose solemnity the Angels rejoice, and glorify the Son of God. \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.
\collect
%1928 BCP:
%\lett{A}{lmighty} God, who hast knit together thine elect in one communion and fellowship, in the mystical body of thy Son Christ our Lord; Grant us grace so to follow thy blessed Saints in all virtuous and godly living, that we may come to those unspeakable joys which thou hast prepared for those who unfeignedly love thee. Through the same.
\lett{A}{lmighty} and everlasting God, who in one solemnity hast vouchsafed unto us to venerate the merits of all thy Saints: we beseech thee; that, at the intercession of so great a multitude, thou wouldest bestow upon us, who entreat thee, the abundance of thy mercy. Through.

\readingcitation{Epistle}{Revelation 7:2}
\lett{I}{n those days:} Behold, I, John, saw another angel ascending from the east, having the seal of the living God: and he cried with a loud voice to the four angels, to whom it was given to hurt the earth and the sea, Saying, Hurt not the earth, neither the sea, nor the trees, till we have sealed the servants of our God in their foreheads. And I heard the number of them which were sealed: and there were sealed an hundred and forty and four thousand of all the tribes of the children of Israel. After this I beheld, and, lo, a great multitude, which no man could number, of all nations, and kindreds, and people, and tongues, stood before the throne, and before the Lamb, clothed with white robes, and palms in their hands; And cried with a loud voice, saying, Salvation to our God which sitteth upon the throne, and unto the Lamb. And all the angels stood round about the throne, and about the elders and the four beasts, and fell before the throne on their faces, and worshipped God, Saying, Amen: Blessing, and glory, and wisdom, and thanksgiving, and honour, and power, and might, be unto our God for ever and ever. Amen. And one of the elders answered, saying unto me, What are these which are arrayed in white robes? and whence came they? And I said unto him, Sir, thou knowest. And he said to me, These are they which came out of great tribulation, and have washed their robes, and made them white in the blood of the Lamb. Therefore are they before the throne of God, and serve him day and night in his temple: and he that sitteth on the throne shall dwell among them. They shall hunger no more, neither thirst any more; neither shall the sun light on them, nor any heat. For the Lamb which is in the midst of the throne shall feed them, and shall lead them unto living fountains of waters: and God shall wipe away all tears from their eyes.

%MANUAL ADJUSTMENT:
\gradual{O fear the Lord, all ye Saints of his: for they that fear him lack nothing. ℣. But they that seek the Lord, shall want no manner of thing that is good.}

\alleluia{Alleluia, alleluia. ℣. Come unto me all ye that travail and are heavy laden: and I will refresh you. Alleluia.}

\readingcitation{Gospel}{Matthew 5:1}
\lett{A}{t that time:} Jesus, seeing the multitudes, went up into a mountain: and when he was set, his disciples came unto him: And he opened his mouth, and taught them, saying, Blessed are the poor in spirit: for theirs is the kingdom of heaven. Blessed are they that mourn: for they shall be comforted. Blessed are the meek: for they shall inherit the earth. Blessed are they which do hunger and thirst after righteousness: for they shall be filled. Blessed are the merciful: for they shall obtain mercy. Blessed are the pure in heart: for they shall see God. Blessed are the peacemakers: for they shall be called the children of God. Blessed are they which are persecuted for righteousness' sake: for theirs is the kingdom of heaven. Blessed are ye, when men shall revile you, and persecute you, and shall say all manner of evil against you falsely, for my sake. Rejoice, and be exceeding glad: for great is your reward in heaven: for so persecuted they the prophets which were before you.
\offertory{The souls of the righteous are in the hand of God, and there shall no torment of malice touch them: in the sight of the unwise they seemed to die: but they are in peace, alleluia.}
\secret
\lett{W}{e} offer unto thee, O Lord, the gifts of our devotion: that they may both be pleasing unto thee for the honour of all the Just, and through thy mercy may be rendered profitable for our salvation. Through.
\communion{Blessed are the pure in heart, for they shall see God; blessed are the peacemakers, for they shall be called the children of God; blessed are they which are persecuted for righteousness' sake, for theirs is the kingdom of heaven.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord, unto thy faithful people ever to rejoice in the veneration of all thy Saints: and to be defended by their perpetual supplication. Through.

\subby{All Souls}\label{AllSouls}
\feastday{All Souls}
\begin{inhead}
    {Double\\
2 November}
\end{inhead}\feastdate{2}{November}

\centerline{\small{(The First Mass)}}

\introit
\lett{R}{est} eternal grant unto them, O Lord: and let light perpetual shine upon them. \textit{Ps.} Thou, O God, art praised in Sion, and unto thee shall the vow be performed in Jerusalem: thou that hearest the prayer, unto thee shall all flesh come.

\collect
\lett{O}{God,} the Creator and Redeemer of all the faithful: grant unto the souls of thy servants and handmaids the remission of all their sins: that through devout supplications they may obtain the pardon which they have alway desired. Who livest.
\readingcitation{Epistle}{1 Corinthians 15:51}
\lett{B}{rethren:} Behold, I shew you a mystery; We shall not all sleep, but we shall all be changed, In a moment, in the twinkling of an eye, at the last trump: for the trumpet shall sound, and the dead shall be raised incorruptible, and we shall be changed. For this corruptible must put on incorruption, and this mortal must put on immortality. So when this corruptible shall have put on incorruption, and this mortal shall have put on immortality, then shall be brought to pass the saying that is written, Death is swallowed up in victory. O death, where is thy sting? O grave, where is thy victory? The sting of death is sin; and the strength of sin is the law. But thanks be to God, which giveth us the victory through our Lord Jesus Christ.

%MANUAL ADJUSTMENT:
\gradual{Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. ℣. The righteous shall be had in everlasting remembrance: he will not be afraid of any evil tidings.}

\tract{Absolve, O Lord, the souls of all the faithful departed from every bond of sin. ℣. And by the help of thy grace may they be worthy to escape the avenging judgment. ℣. And enjoy the bliss of everlasting light.}

\subsubsec{Sequence}\label{diesirae}
\fakecol{
\begin{multicols}{2}
\begin{hangparas}{1.25em}{1}
\lett{D}{ay} of wrath and doom impending,\par
\secondline{David's word with Sibyl's blending:}
\thirdline{\par\noindent%
Heaven and earth in ashes ending.\\}
\par\noindent
Oh, what fear man's bosom rendeth\par\noindent
When from heaven the judge descendeth,\par\noindent
On whose sentence all dependeth!\\
\par\noindent
Wondrous sound the trumpet flingeth,\par\noindent
Through earth's sepulchres it ringeth,\par\noindent
All before the throne it bringeth.\\
\par\noindent
Death is struck, and nature quaking,\par\noindent
All creation is awaking,\par\noindent
To its judge an answer making.\\
\par\noindent
Lo! the book, exactly worded,\par\noindent
Wherein all hath been recorded,\par\noindent
Thence shall judgement be awarded.\\
\par\noindent
When the judge his seat attaineth,\par\noindent
And each hidden deed arraigneth,\par\noindent
Nothing unavenged remaineth.\\
\par\noindent
What shall I, frail man, be pleading?\par\noindent
Who for me be interceding,\par\noindent
When the just are mercy needing?%MANUAL ADJUSTMENT:\\\par\noindent

\columnbreak
\par\noindent
King of majesty tremendous,\par\noindent
Who dost free salvation send us,\par\noindent
Fount of pity, then befriend us!\\
\par\noindent
Think, kind Jesu, my salvation\par\noindent
Caused thy wondrous Incarnation:\par\noindent
Leave me not to reprobation.\\
\par\noindent
Faint and weary thou hast sought me:\par\noindent
On the Cross of suffering bought me:\par\noindent
Shall such grace be vainly brought me?\\
\par\noindent
Righteous judge, for sin's pollution\par\noindent
Grant thy gift of absolution\par\noindent
Ere the day of retribution.\\
\par\noindent
Guilty, now I pour my moaning:\par\noindent
All my shame with anguish owning:\par\noindent
Spare, O God, thy suppliant groaning.\\
\par\noindent
%Faithfulness to the Latin is kept by translating `Mary' not `woman'
Through the sinful Mary shriven,\par\noindent
Through the dying thief forgiven,\par\noindent
Thou to me a hope hast given.\\
\par\noindent
Worthless are my prayers and sighing:\par\noindent
Yet, good Lord, in grace complying,\par\noindent
Rescue me from fires undying.\\
\par\noindent
With thy sheep a place provide me,\par\noindent
From the goats afar divide me,\par\noindent
To thy right hand do thou guide me.\\
\par\noindent
When the wicked are confounded,\par\noindent
Doomed to flames of woe unbounded:\par\noindent
Call me, with thy Saints surrounded.\\
\par\noindent
Low I kneel, with heart submission,\par\noindent
See, like ashes, my contrition:\par\noindent
Help me in my last condition.%MANUAL ADJUSTMENT:\\\par\noindent

\columnbreak
\par\noindent
Ah! that day of tears and mourning,\par\noindent
From the dust of earth returning,\par\noindent
Man for judgment must prepare him.\\
\par\noindent
Spare, O God, in mercy spare him:\par\noindent
Lord, all-pitying, Jesu blest,\par\noindent
Grant them thine eternal rest. Amen.
\end{hangparas}
\end{multicols}
}
{\lett{D}{ies} ir{\ae}, dies illa\par
\secondline{Solvet s{\ae}clum in favílla:}
\thirdline{Teste David cum Sibýlla.\\}
Quantus tremor est futúrus,

Quando judex est ventúrus,

Cuncta stricte discussúrus!\\

Tuba mirum spargens sonum

Per sepúlcra regiónum,

Coget omnes ante thronum.\\

Mors stupébit et natúra,

Cum resúrget creatúra,

Judicánti responsúra.\\

Liber scriptus proferétur,

In quo totum continétur,

Unde mundus judicétur.\\

Judex ergo cum sedébit,

Quidquid latet, apparébit:

Nil inúltum remanébit.\\

Quid sum miser tunc dictúrus?

Quem patrónum rogatúrus,

Cum vix justus sit secúrus?\\

Rex treménd{\ae} majestátis,

Qui salvándos salvas gratis,

Salva me, fons pietátis.\\

Recordáre, Jesu pie,

Quod sum causa tu{\ae} vi{\ae}:

Ne me perdas illa die.\\

Qu{\ae}rens me, sedísti lassus:

Redemísti Crucem passus:

Tantus labor non sit cassus.\\

Juste judex ultiónis,

Donum fac remissiónis

Ante diem ratiónis.\\

Ingemísco, tamquam reus:

Culpa rubet vultus meus:

Supplicánti parce, Deus.\\

Qui Maríam absolvísti,

Et latrónem exaudísti,

Mihi quoque spem dedísti.\\

Preces me{\ae} non sunt dign{\ae}:

Sed tu bonus fac benígne,

Ne perénni cremer igne.\\

Inter oves locum pr{\ae}sta,

Et ab h{\oe}dis me sequéstra,

Státuens in parte dextra.\\

Confutátis maledíctis,

Flammis ácribus addíctis:

Voca me cum benedíctis.\\

Oro supplex et acclínis,

Cor contrítum quasi cinis:

Gere curam mei finis.\\

Lacrimósa dies illa,

Qua resúrget ex favílla

Judicándus homo reus.\\

Huic ergo parce, Deus:

Pie Jesu Dómine,

Dona eis réquiem. Amen.}
\readingcitation{Gospel}{John 5:25}
\lett{A}{t that time:} Jesus said to the multitude of the Jews: Verily, verily, I say unto you, The hour is coming, and now is, when the dead shall hear the voice of the Son of God: and they that hear shall live. For as the Father hath life in himself; so hath he given to the Son to have life in himself; And hath given him authority to execute judgment also, because he is the Son of man. Marvel not at this: for the hour is coming, in the which all that are in the graves shall hear his voice, And shall come forth; they that have done good, unto the resurrection of life; and they that have done evil, unto the resurrection of damnation.
\offertory{O Lord Jesu Christ, King of glory, deliver the souls of all the faithful departed from the pains of hell and from the depths of the pit: deliver them from the lion's mouth, that hell devour them not, that they fall not into darkness: but let the standard-bearer, Saint Michael, bring them into the holy light: Which of old thou didst promise unto Abraham and his seed. ℣. We offer unto thee, O Lord, sacrifices of prayer and praise: do thou receive them for the souls of those whose memory we this day recall: make them, O Lord, to pass from death unto life. Which of old thou didst promise unto Abraham and his seed.}
\secret
\lett{W}{e} beseech thee, O Lord, look favourably upon the sacrifices which we offer unto thee for the souls of thy servants and handmaids: that, as thou hast bestowed on them the merit of the Christian faith, thou wouldest grant them the reward of the same. Through.
\communion{Let light eternal shine upon them, O Lord: With thy Saints for evermore: for thou art gracious. ℣. Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. With thy Saints for evermore: for thou art gracious.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, that the prayer of thy suppliant people may be profitable for the souls of thy servants and handmaids: that thou wouldest deliver them from all their sins, and make them to be partakers of thy redemption. Who livest.

%MANUAL ADJUSTMENT:
%\clearpage
\subby{All Souls}
\centerline{\small{(The Second Mass)}}\label{AllSoulsII}
\par
\introit
\lett{R}{est} eternal grant unto them, O Lord: and let light perpetual shine upon them. \textit{Ps.} Thou, O God, art praised in Sion, and unto thee shall the vow be performed in Jerusalem: thou that hearest the prayer, unto thee shall all flesh come.
\collect
\lett{O}{God,} the Lord of mercies: grant unto the souls of thy servants and handmaids a place of refreshment, the blessedness of rest, and the brightness of thy light. Through.
\readingcitation{Epistle}{2 Maccabees 12:43}
\lett{I}{n those days:} That most valiant man Judas, when he had made a gathering throughout the company to the sum of two thousand drachms of silver, he sent it to Jerusalem to offer a sin offering, doing therein very well and honestly, in that he was mindful of the resurrection: For if he had not hoped that they that were slain should have risen again, it had been superfluous and vain to pray for the dead. And also in that he perceived that there was great favour laid up for those that died godly, it was an holy and good thought. Whereupon he made a reconciliation for the dead, that they might be delivered from sin.
\gradtr{Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. ℣. The righteous shall be had in everlasting remembrance: he will not be afraid of any evil tidings.}{Absolve, O Lord, the souls of all the faithful departed from every bond of sin. ℣. And by the help of thy grace may they be worthy to escape the avenging judgment. ℣. And enjoy the bliss of everlasting light.}
\begin{rubric}
    The Sequence \emph{Dies irae} is here said or sung (p. \pageref{diesirae}).
\end{rubric}
\readingcitation{Gospel}{John 6:37}
\lett{A}{t that time:} Jesus said to the multitudes of the Jews: All that the Father giveth me shall come to me; and him that cometh to me I will in no wise cast out. For I came down from heaven, not to do mine own will, but the will of him that sent me. And this is the Father's will which hath sent me, that of all which he hath given me I should lose nothing, but should raise it up again at the last day. And this is the will of him that sent me, that every one which seeth the Son, and believeth on him, may have everlasting life: and I will raise him up at the last day.
\offertory{O Lord Jesu Christ, King of glory, deliver the souls of all the faithful departed from the pains of hell and from the depths of the pit: deliver them from the lion's mouth, that hell devour them not, that they fall not into darkness: but let the standard-bearer, Saint Michael, bring them into the holy light: Which of old thou didst promise unto Abraham and his seed. ℣. We offer unto thee, O Lord, sacrifices of prayer and praise: do thou receive them for the souls of those whose memory we this day recall: make them, O Lord, to pass from death unto life. Which of old thou didst promise unto Abraham and his seed.}
\secret
\lett{B}{e} favourable, O Lord, to our supplications for the souls of thy servants and handmaids, for whom we offer unto thee this sacrifice of praise: that thou wouldest vouchsafe to join them to the fellowship of thy Saints. Through.
\communion{Let light eternal shine upon them, O Lord: With thy Saints for evermore: for thou art gracious. ℣. Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. With thy Saints for evermore: for thou art gracious.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord: that the souls of thy servants and handmaids, being purged by these sacrifices, may obtain both pardon and everlasting rest. Through.

\subby{All Souls}
\centerline{\small{(The Third Mass)}}\label{AllSoulsIII}

\introit
\lett{R}{est} eternal grant unto them, O Lord: and let light perpetual shine upon them. \textit{Ps.} Thou, O God, art praised in Sion, and unto thee shall the vow be performed in Jerusalem: thou that hearest the prayer, unto thee shall all flesh come.
\collect
\lett{O}{God,} the giver of pardon and lover of man's salvation: we beseech thee of thy mercy to grant; that the souls of thy servants and handmaids, who have passed out of this world, may, at the intercession of the blessed Mary ever Virgin and of all thy Saints, attain unto the fellowship of eternal blessedness. Through.
\readingcitation{Epistle}{Revelation 14:13}
\lett{I}{n those days:} I heard a voice from heaven saying unto me, Write, Blessed are the dead which die in the Lord from henceforth: Yea, saith the Spirit, that they may rest from their labours; and their works do follow them.
\gradtr{Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. ℣. The righteous shall be had in everlasting remembrance: he will not be afraid of any evil tidings.}{Absolve, O Lord, the souls of all the faithful departed from every bond of sin. ℣. And by the help of thy grace may they be worthy to escape the avenging judgment. ℣. And enjoy the bliss of everlasting light.}
\begin{rubric}
    The Sequence \emph{Dies irae} is here said or sung (p. \pageref{diesirae}).
\end{rubric}
\readingcitation{Gospel}{John 6:51}
\lett{A}{t that time:} Jesus said to the multitude of the Jews: I am the living bread which came down from heaven: if any man eat of this bread, he shall live for ever: and the bread that I will give is my flesh, which I will give for the life of the world. The Jews therefore strove among themselves, saying, How can this man give us his flesh to eat? Then Jesus said unto them, Verily, verily, I say unto you, Except ye eat the flesh of the Son of man, and drink his blood, ye have no life in you. Whoso eateth my flesh, and drinketh my blood, hath eternal life; and I will raise him up at the last day.
\offertory{O Lord Jesu Christ, King of glory, deliver the souls of all the faithful departed from the pains of hell and from the depths of the pit: deliver them from the lion's mouth, that hell devour them not, that they fall not into darkness: but let the standard-bearer, Saint Michael, bring them into the holy light: Which of old thou didst promise unto Abraham and his seed. ℣. We offer unto thee, O Lord, sacrifices of prayer and praise: do thou receive them for the souls of those whose memory we this day recall: make them, O Lord, to pass from death unto life. Which of old thou didst promise unto Abraham and his seed.}
\secret
\lett{O}{God,} whose mercies are without number, graciously receive our humble prayers: and through these sacraments of our salvation grant unto the souls of all the faithful departed, to whom thou hast given the confession of thy name, the remission of all their sins. Through.
\communion{Let light eternal shine upon them, O Lord: With thy Saints for evermore: for thou art gracious. ℣. Rest eternal grant unto them, O Lord: and let light perpetual shine upon them. With thy Saints for evermore: for thou art gracious.}
\postcommunion
\lett{G}{rant,} we beseech thee, almighty and merciful God: that the souls of thy servants and handmaids, for whom we have offered unto thy majesty this sacrifice of praise; may by the power of this sacrament be cleansed from all their sins, and by thy mercy obtain the blessedness of everlasting light. Through.

