   %\section{Table of Prayers}
   \fancyhead[C]{{\LARGE Table of Prayers}}
   \fancyhead[RO,LE]{}
   \fancyhead[RE,LO]{}
   \phantomsection
	\addcontentsline{toc}{section}{Table of Prayers}
\begin{secrubric}
	These are provided for the additional Prayers during Mass.
\end{secrubric}
   \begin{multicols}{2}
   \subby{Of St. Mary in Advent}\label{SPMaryInAdvent}
   \subsubsec{Collect}
   \lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we, who believe her to be truly the Mother of God, may be aided by her intercession in thy sight. Through the same.
   \subsubsec{Secret}
   \lett{W}{e} beseech thee, O Lord, stablish in our hearts the mysteries of thy true religion: that, as we believe thy Son, conceived of a Virgin, to be very God and very man; so by the power of his life-giving resurrection we may be found worthy to attain unto everlasting felicity. Through the same.
   \subsubsec{Postcommunion}
   \lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts: that, as we have known the incarnation of thy Son Christ by the message of an Angel; so by his cross and passion we may be brought unto the glory of his resurrection. (Through the same.)

    \subby{Of St. Mary after Christmas}\label{SPMaryPostChristmas}
    \subsubsec{Collect}
    \lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son, our Lord. (Who liveth.)
    \subsubsec{Secret}
    \lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace both now and for ever. (Through.)
    \subsubsec{Postcommunion}
    \lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary the Virgin Mother of God, make us partakers of thy heavenly healing. (Through the same.)
    
    %MANUAL ADJUSTMENT:
    \subsec{Of St. Mary in Eastertide}\label{SPMaryInEaster}
    \subsubsec{Collect}
    \lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. (Through.)
    \subsubsec{Secret}
    \lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and ever. (Through.)
    \subsubsec{Postcommunion}
    \lett{G}{rant,} we beseech thee, O Lord: that we who have received these aids to our salvation may at all times and in all places be protected by the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. (Through.)\\

    \subby{Of Saints}\label{SPSaints}
    \subsubsec{Collect}
    \lett{D}{efend} us, O Lord, we beseech thee, from all dangers of mind and of body: and at the intercession of the blessed and glorious Mother of God, Mary ever Virgin, with blessed Joseph, thy blessed Apostles Peter and Paul, blessed \emph{N.}, and all the Saints, graciously grant us health and peace; that all adversities and errors being done away, thy Church may serve thee in freedom and quietness. (Through the same.)
    \subsubsec{Secret}
    \lett{G}{raciously} hear us, O God our Saviour: and by the power of this sacrament defend us from all enemies of mind and of body; granting us grace in this world, and glory in that which is to come. (Through.)
    \subsubsec{Postcommunion}
    \lett{W}{e} beseech thee, O Lord, that the gift now offered in this divine sacrament may cleanse and defend us: that at the intercession of blessed Mary the Virgin Mother of God, of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \emph{N.} and of all the Saints; it may set us free from every perverse way, and deliver us from all adversities. (Through the same.)
    
   \subby{Against the Persecutors of the Church}\label{SPAgainst}
   \subsubsec{Collect}
   \lett{W}{e} beseech thee, O Lord, mercifully to hear the prayers of thy Church: that all adversities and errors being done away, she may serve thee in freedom and quietness. Through.
   \subsubsec{Secret}
   \lett{D}{efend} us, O Lord, who wait upon thy mysteries: that we, cleaving fast to things heavenly, may serve thee both in body and soul. Through.
   \subsubsec{Postcommunion}
   \lett{W}{e} beseech thee, O Lord our God: that we, whom thou makest to rejoice in the partaking of heavenly things, may by thee be defended against all earthly perils. Through.
   
   \subby{For the Chief Bishop}\label{SPChiefBishop}
   \subsubsec{Collect}
   \lett{O}{God,} the pastor and ruler of all the faithful, mercifully look upon thy servant \emph{N.}, whom thou hast been pleased to set as pastor over thy Church; grant him, we beseech thee, to be in word and conversation a wholesome example to the people committed to his charge; that he with them may attain unto everlasting life. Through.
   \subsubsec{Secret}
   \lett{L}{ook} favourably, we beseech thee, O Lord, upon the gifts which we offer: and guide with thy continual protection thy servant \textit{N.}, whom thou hast been pleased to set as pastor over thy Church. Through.
   \subsubsec{Postcommunion}
   \lett{W}{e} beseech thee, O Lord, that the divine sacrament which we have here received may be our protection: and together with the flock entrusted to his charge ever save and defend thy servant \textit{N.}; whom thou hast been pleased to set as pastor over thy Church. Through.
   
   \subby{Of the Holy Ghost}\label{SPHolyGhost}
   \subsubsec{Collect}
   \lett{G}{od,} who didst teach the hearts of thy faithful people, by the sending to them the light of thy Holy Spirit: grant us by the same Spirit to have a right judgment in all things; and evermore to rejoice in his holy comfort. (Through . . . in the unity of the same Holy Spirit.)
   \subsubsec{Secret}
   \lett{S}{anctify,} we beseech thee, O Lord, the gifts which we offer: and cleanse our hearts by the enlightening of the Holy Spirit. (Through . . . in the unity of the same Holy Spirit.)
   \subsubsec{Postcommunion}
   \lett{P}{our} thy Holy Spirit upon us, O Lord, and cleanse our hearts: that they may be made fruitful by the inward sprinkling of his dew. (Through . . . in the unity of the same Holy Spirit.)

    \subby{For the Living and Departed}\label{SPLivingDeparted}
   \subsubsec{Collect}
   \lett{A}{lmighty} and everlasting God, who hast dominion both of the living and of the dead, and hast mercy upon all who thou foreknowest will be thine in faith and works: we humbly beseech thee; that all those for whom we are minded to pour forth our prayers, whether in this present world they still be held in the flesh, or being delivered from the body have passed into that which is to come, may at the intercession of all thy Saints obtain of thy bountiful goodness the remission of all their sins. Through.
   \subsubsec{Secret}
   \lett{O}{God,} to whom alone is known the number of the elect, whom thou hast appointed unto heavenly felicity: grant, we beseech thee: that at the intercession of all thy Saints, the names of all those who have been commended to our prayers, and of all thy faithful people, may ever remain written in the book of those that are predestined to everlasting blessedness. Through.
   \subsubsec{Postcommunion}
   \lett{A}{lmighty} and merciful God, cleanse us, we pray thee, by the sacraments which we have received: and, at the intercession of all thy Saints, grant; that this thy sacrament may not bring upon us guilt to our punishment but pardon to our salvation: that it may cleanse away sins, that it may confirm the feeble, that it may defend us against every danger in this world: that it may obtain for thy faithful people both living and dead the remission of all their sins. Through.
 \end{multicols}
\vspace{-1ex}
\begin{rubric}
    The following propers may also be used when \emph{ad libitum} prayers are permitted.
\end{rubric}
\vspace{-1ex}
\begin{multicols}{2}
\subby{To Ask the Prayers of the Saints}
\subsubsec{Collect}
\lett{G}{rant,} we beseech thee, almighty God: that the intercession of holy Mary, Mother of God, and of all the holy Apostles, Martyrs, Confessors and Virgins, and of all thy elect, may everywhere cause us to rejoice; that while we call to mind their merits, we may perceive their advocacy. Through the same.
\subsubsec{Secret}
\lett{A}{ccept,} O Lord, the gifts we offer: and, at the intercession of blessed Mary, ever Virgin, with all thy Saints, defend us from all perils. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, O Lord: that we who have received these heavenly sacraments, in veneration of the memory of blessed Mary, ever Virgin, and of all thy Saints; may attain in everlasting joys the fulfilment of our service in this life. Through.

%\begin{inhead}
%    or
%\end{inhead}
%
%\subsubsec{Collect}
%\lett{D}{efend} us, O Lord, we beseech thee, from all dangers of mind and of body: and at the intercession of the blessed and glorious Mother of God, Mary, ever Virgin, with blessed Joseph, thy blessed Apostles, Peter and Paul, blessed \textit{N.}, and all the Saints, graciously grant us health and peace; that, all adversities and errors being done away, thy Church may serve thee in freedom and quietness. Through the same.
%\subsubsec{Secret}
%\lett{G}{raciously} hear us, O God our Saviour: and by the power of this sacrament defend us from all enemies of mind and of body; granting us grace in this world, and glory in that which is to come. Through.
%\subsubsec{Postcommunion}
%\lett{W}{e} beseech thee, O Lord, that the gift now offered in this divine sacrament may cleanse and defend us: that at the intercession of blessed Mary the Virgin Mother of God, of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.}, and of all the Saints; it may set us free from every perverse way and deliver us from all adversities. Through the same.

\subby{For all Estates of Men in the Church}
\subsubsec{Collect}
\lett{A}{lmighty} and everlasting God, by whose Spirit the whole body of the Church is governed and sanctified: receive our supplications and prayers, which we offer before thee for all estates of men thy holy Church; that every member of the same, in his vocation and ministry, may truly and godly serve thee. Through . . . in the unity of the same Holy Spirit.
\subsubsec{Secret}
\lett{G}{rant} unto thy servants, O Lord, the pardon of their sins, comfort in their life, and thy continual governance: so that they, serving thee aright, may continually obtain thy mercy. Through.
%MANUAL ADJUSTMENT:
\postcommunion
\lett{D}{eliver,} O Lord, we beseech thee, thy servants who call upon thee from all sin and from every foe: that they, walking in godly conversation, may be hurt by no adversity. Through.

%For the Chief Bishop

\subby{For the Emperor}
\subsubsec{Collect}
\lett{O}{God,} the protector of all kingdoms, and specially of the Christian empire: grant unto thy servant, our emperor, \textit{N.}, to labour with wisdom for the triumph of virtue, that he who by thy institution is a prince, may ever by thy gift be powerful. Through.
\begin{rubric}
    \textsc{Note,} If he be not crowned, say \emph{Emperor-elect}.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-3ex}

\subsubsec{Secret}
\lett{R}{eceive,} O Lord, the prayers and offerings of thy Church who calleth upon thee for the safety of thy servant: and for the protection of the faithful peoples perform with thy mighty arm thy wondrous works of old, that they may overcome the enemies of peace, and serve thee in freedom and safety. Through.
\subsubsec{Postcommunion}
\lett{O}{God,} who didst prepare the Roman empire for the preaching of the Gospel of the eternal King: defend thy servant, our emperor, \textit{N.}, with thy heavenly armour, that the peace of the churches may be troubled by no war nor tumult. Through.

\subby{For the King}
\subsubsec{Collect}
\lett{W}{e} beseech thee, almighty God: that thy servant \textit{N.}, our King, who by thy mercy hath received the government of this realm, may also obtain an increase of every virtue; that, being meetly adorned therewith, he may flee from sin and iniquity, and by thy grace may attain unto thee who art the way, the truth, and the life. Through.
\subsubsec{Secret}
\lett{S}{anctify,} we beseech thee, O Lord, the gifts which we offer: that they may be made unto us the Body and Blood of thine only-begotten Son; and of thy great goodness grant that they may at all times be profitable unto our King, both for the obtaining of health of body and soul, and for the fulfilling of the charge committed unto him. Through.
\subsubsec{Postcommunion}
\lett{M}{ay} this saving oblation, O Lord, keep thy servant, \textit{N.}, our King, from all adversity: that he may preserve peace and tranquillity for thy Church; and after the course of this life may attain unto his eternal inheritance. Through.

\subby{For Prelates \& Congregations Committed to their Charge}
\subsubsec{Collect}
\lett{A}{lmighty} and everlasting God, who alone workest great marvels: send down upon our Bishops, and all Congregations committed to their charge, the healthful spirit of thy grace; and that they may truly please thee, pour upon them the continual dew of thy blessing. Through.
\subsubsec{Secret}
\lett{L}{ook} down, O Lord, in mercy on the sacrifices of thy servants, which we devoutly celebrate for them unto the honour of thy name: that they may perceive the same to be profitable for the healing of their souls. Through.
\subsubsec{Postcommunion}
\lett{A}{ssist,} O Lord, with thy continual protection those whom thou dost refresh with this heavenly gift: and grant that, as thou never failest to comfort them, so thou wouldest make them worthy of everlasting redemption. Through.

\subby{For the Congregation and Family}
\subsubsec{Collect}
\lett{W}{e} beseech thee, O Lord, at the intercession of blessed Mary, ever Virgin, to defend this thy family from all adversities: that they, devoutly serving thee with all their hearts, may by thy mercy be protected from the assaults of their enemies. Through.
\subsubsec{Secret}
\lett{A}{lmighty} God, we beseech thee to accept this offering of our devotion: that by the power of this sacrament we thy servants may be protected against all adversities. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, O merciful God: that we who have here received the gifts of our redemption; may through the celebration of the same obtain the help of thy defence against all adversities. Through.

%MANUAL ADJUSTMENT:
%\columnbreak
\subby{For the Preservation of Concord in the Congregation}
\subsubsec{Collect}
\lett{O}{God,} who art the author of peace and lover of charity: grant unto thy servants that true unity which is according to thy will; that we may be delivered from all the temptations which beset us. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, mercifully to accept these sacrifices: and grant that we who pray to be absolved from our own offences may not be burdened with the sins of other men. Through.
\subsubsec{Postcommunion}
\lett{P}{our} forth upon us, O Lord, the Spirit of thy charity: that as thou hast fulfilled us with one heavenly bread, so of thy goodness thou wouldest make us to be of one heart and mind. Through . . . in the unity of the same.


%MANUAL ADJUSTMENT:
%\columnbreak
%Against the Persecutors of the Church

\subby{Against Persecutors \& Malefactors}
\subsubsec{Collect}
\lett{W}{e} beseech thee, O Lord, to bring down the pride of our enemies: and by the right hand of thy power destroy their obstinacy. Through.
\subsubsec{Secret}
\lett{G}{rant,} O Lord, that by the power of this mystery, we may be cleansed from our secret faults, and delivered from the snares of our enemies. Through.
\subsubsec{Postcommunion}
\lett{B}{ehold,} O God our defender, and protect us against the dangers of our enemies: that we may be delivered from all distress, and serve thee in freedom of spirit. Through.

\subby{For any Necessity}
\subsubsec{Collect}
\lett{O}{God,} our refuge and strength, who art the author of all godliness: be ready, we beseech thee, to hear the devout prayers of thy Church; and grant that those things which we ask faithfully we may obtain effectually. Through.
\subsubsec{Secret}
\lett{G}{rant,} O merciful God: that this saving oblation may continually deliver us from our sins, and shield us from all adversities. Through.
\subsubsec{Postcommunion}
\lett{W}{e} who have partaken, O Lord, of the gifts of this sacred mystery, humbly beseech thee: that those things which thou hast commanded us to do in remembrance of thee may be profitable for the assistance of our infirmity. Who livest.

\subby{In any Calamity}
\subsubsec{Collect}
\lett{A}{lmighty} God, despise not thy people who cry to thee in their affliction: but for the glory of thy name mercifully assist them in their tribulation. Through.
\subsubsec{Secret}
\lett{M}{ercifully} accept, O Lord, these sacrifices which thou hast ordained to be the propitiation of our sins, and the means whereby of thy loving mercy we are restored to salvation. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, mercifully look upon our tribulation: and turn from us the fury of thine anger, which we most righteously deserve. Through.

\subby{In time of Famine}
\subsubsec{Collect}
\lett{G}{rant} to us, we beseech thee, O Lord, the fulfilment of our devout supplications: and graciously turn away this famine; that the hearts of men may know that in thine anger thou sendest forth such scourges, and in thy mercy makest them to cease. Through.
\subsubsec{Secret}
\lett{O}{God,} who dost sustain the twofold nature of man by the food of the gifts we offer, and by this sacrament dost renew the same: grant, we beseech thee; that these thy gifts may never be lacking for our assistance both in body and soul. Through.
%MANUAL ADJUSTMENT:
\subsubsection{Postcommunion}
\lett{O}{Lord,} support us, we beseech thee, with temporal food: even as thou dost vouchsafe to dwell in us in everlasting mysteries. Through.

\subby{In time of Earthquake}
\subsubsec{Collect}
\lett{A}{lmighty} and everlasting God, who lookest upon the earth, and makest it to tremble: spare them that fear thee, be merciful to them that call upon thee; that whereas we are sore afraid for thy wrath that shaketh the foundations of the earth, we may likewise feel thy mercy when thou healest the sores thereof. Through.
\subsubsec{Secret}
\lett{O}{God,} who hast laid the foundations of the earth that they should never be moved, receive the prayers and oblations of thy people: put far from us the present perils of earthquake, and turn the terrors of thy divine anger into a wholesome medicine for the safety of mankind; that they who are of the earth, and shall return to earth, may rejoice to be made citizens of heaven by holy conversation. Through.
\subsubsec{Postcommunion}
\lett{P}{rotect} us, O Lord, we beseech thee, who receive thy holy things: and by thy heavenly gift stablish the earth which we see trembling for our sins; that the hearts of men may know that in thine anger thou sendest forth thy scourges, and in thy mercy makest them to cease. Through.

\subby{For Rain}
\subsubsec{Collect}
\lett{O}{God,} in whom we live, and move, and have our being: grant us such a seasonable rain; that we, receiving those things that be requisite for our necessities in this life, may the more confidently seek the things that are eternal. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, mercifully to receive the gifts which we offer: and send unto us a seasonable rain that may suffice for our necessities. Through.
\subsubsec{Postcommunion}
\lett{G}{rant} us, O Lord, we beseech thee, a refreshing rain: and graciously water the parched face of the earth by showers from heaven. Through.

\subby{For Fine Weather}
\subsubsec{Collect}
\lett{G}{raciously} hear us, O Lord, who call upon thee: and grant fair weather in answer to our prayers; that we, who are justly afflicted for our sins, may by thy preventing mercy feel thy pity. Through.
\subsubsec{Secret}
\lett{L}{ord,} we pray thee, that thy grace may alway prevent and follow us: and graciously accept these oblations which for our sins we bring for consecration unto thy name; that, through the intercession of thy Saints, they may be profitable for the safety of us all. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} God, we humbly beseech thy mercy: that thou wouldest assuage this plague of immoderate rains, and vouchsafe to shew forth on us the gladness of thy countenance. Through.

\subby{Against Storms}
\subsubsec{Collect}
\lett{W}{e} beseech thee, O Lord, that all spiritual wickedness may be driven far from thy house: and that the malignity of the tempests of the air be likewise dispelled. Through.
\subsubsec{Secret}
\lett{W}{e} offer unto thee, O Lord, our praises and gifts, rendering thanks for the blessings bestowed upon us, and ever humbly praying for the continuance of the same. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} and everlasting God, who dost heal us by thy chastisement, and preserve us by thy pardon: grant unto thy suppliant people; that in tranquillity of weather we may rejoice in the comfort we have ever desired, and may alway make use of thy bountiful goodness. Through.

%MANUAL ADJUSTMENT:
\subsec{Against Cattle Disease}
\subsubsec{Collect}
\lett{O}{God,} who even by the dumb beasts hast brought comfort to man's labour: we humbly beseech thee; that forasmuch as without them the sustenance of man cannot but fail, so thou wouldest preserve them for our enjoyment. Through.
\subsubsec{Secret}
\lett{L}{ook} mercifully, O Lord, upon the sacrifices which we offer: and of thy mercy grant us thy help in our time. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} O Lord, that thy faithful people, being filled with thy blessing, may be preserved both in body and soul: that they may render unto thee worthy service, and evermore receive the benefits of thy mercy. Through.

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\subby{For the Priest himself}
\subsubsec{Collect}
\lett{A}{lmighty} and merciful God, graciously give heed to my humble prayers: and make me thy servant, to whom for no merits of my own but by the boundless compassion of thy bounty thou hast vouchsafed to serve thee in these heavenly mysteries, a worthy minister at the sacred altars; that the words which I utter with my lips may by thy sanctifying power be brought to good effect. Through.

%MANUAL ADJUSTMENT:
\vspace{-1.5ex}

\subsubsec{Secret}
\lett{C}{leanse} me, O Lord, by the power of this sacrament, from all the stains of my sins: and grant; that with the help of thy grace I may be made worthy thereby to fulfil the ministry committed to my charge. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} and everlasting God, who hast vouchsafed to me, a sinner, to stand before thy sacred altars, and to praise the power of thy holy name: mercifully grant to me by the mystery of this sacrament the pardon of my sins; that I may offer unto thy majesty true and worthy service. Through.

\subby{To ask for Tears}
\subsubsec{Collect}
\lett{A}{lmighty} and most merciful God, who didst bring forth living water from the rock for thy thirsting people: draw out from our stony hearts the tears of repentance; that we, worthily lamenting our sins, may of thy mercy obtain forgiveness of the same. Through.
\subsubsec{Secret}
\lett{O}{Lord} God, we beseech thee, look down in mercy on this oblation, which we offer to thy majesty for our sins: and draw forth from our eyes such streams of tears, as shall quench the burning flames which we most righteously deserve. Through.
\subsubsec{Postcommunion}
\lett{O}{Lord} God, mercifully pour the grace of thy Holy Spirit into our heart: that we may be enabled thereby with tears of sorrow to wash away the defilements of our sins; that like as we do desire thy merciful pardon, so of thy bountiful goodness we may effectually receive the same. Through. . . in the unity of the same.

\subby{For the Remission of Sins}
\subsubsec{Collect}
\lett{O}{God,} who dost not cast out sinners, and of thy loving mercy sparest them that are penitent, howsoever they sin against thee: mercifully incline to the prayers of thy humble servants, and enlighten our hearts; that we may be enabled to fulfil all that thou dost command. Through.
\subsubsec{Secret}
\lett{L}{et} this present sacrifice, O Lord, which we offer to thee for our offences, be an acceptable offering unto thee: and avail for salvation both to the living and the dead. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} God, graciously hear the prayers of this thy family: and grant; that, like as we have received these holy things from thee, so of thy bounty they may be preserved pure and undefiled within us. Through.

\subby{For Public Penitents}
\subsubsec{Collect}
\lett{A}{lmighty} and everlasting God, of thy great goodness forgive the sins of thy servants who confess the same to thee: that they whose consciences by sin are accused may not be delivered unto punishment, but may obtain of thy loving mercy pardon and forgiveness. Through.
\subsubsec{Secret}
\lett{G}{rant,} we beseech thee, almighty and merciful God: that this saving oblation may continually deliver thy servants from their offences, and defend them against all adversities. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} and merciful God, who wouldest not the death of sinners, but rather that they should repent, and confess to thee, and amend their lives: look down upon these thy servants; and by these holy sacraments which we have received turn from them thy wrathful indignation, and grant them forgiveness of all their sins. Through.

\subby{For the Tempted and Troubled}
\subsubsec{Collect}
\lett{O}{God,} who dost justify the wicked, and wouldest not the death of a sinner, we humbly beseech thy majesty: that thy servants who put their trust and confidence in thy mercy may be defended by the loving-kindness of thy heavenly succour and preserved by thy continual help; that in all temptations they may steadfastly cleave unto thy service. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, that by the power of this mystery thou wouldest cleanse us from our faults, and absolve thy servants from all their offences. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, O Lord, that the sacraments which we have received may purify our hearts: and that thy servants may be delivered from all transgression; that whereas their consciences are bound by reason of their guilt, they may receive with gladness the plenteous healing of thy heavenly power. Through.

\subby{To repel Evil Thoughts}
\subsubsec{Collect}
\lett{A}{lmighty} and most merciful God, graciously regard our prayers: that, our hearts being delivered from the temptations of evil thoughts; we may be found worthy to be made a dwelling-place acceptable to thy Holy Spirit. Through . . . in the unity of the same Holy Spirit.
\subsubsec{Secret}
\lett{W}{e} here present unto thee, O Lord, these offerings for our salvation: that the thoughts of our hearts being purged from all uncleanness and preserved from harm, we may be enlightened by the grace of thy Holy Spirit. Through . . . in the unity of the same Holy Spirit.
\subsubsec{Postcommunion}
\lett{O}{God,} who lightenest every man that cometh into this world: lighten, we beseech thee, our hearts with the brightness of thy grace; that we may ever be enabled to think such things as be good and pleasing to thy majesty, and that we may love thee with an unfeigned heart. Through.

\subby{To obtain the Grace of Continence}
\subsubsec{Collect}
\lett{K}{indle,} O Lord, with the fire of the Holy Spirit our reins and our hearts: that we, serving thee in pureness both of body and soul, may be found pleasing in thy sight. Through . . . in the unity of the same.
\subsubsec{Secret}
\lett{S}{mite} asunder, O Lord, the chains of our sins: that we who seek to offer unto thee the sacrifice of praise in perfect freedom and in innocency of heart, may be restored by thee unto our first estate; that like as thou hast saved us by thy grace, so by thy pardon thou wouldest restore us again unto salvation. Through.
\subsubsec{Postcommunion}
\lett{A}{ssist} us, O God, our helper and defender: that our heart and flesh being renewed by the strength of continence and the newness of chastity, we who have offered unto thy majesty this sacrifice may be cleansed thereby from all temptations. Through.

\subby{To obtain the Grace of Humility}
\subsubsec{Collect}
\lett{O}{God,} who resistest the proud, and givest grace unto the humble: grant to us the virtue of true humility, after the pattern wherewith thine only-begotten Son manifested himself unto thy faithful people; that we may in no wise exalt ourselves to provoke thine anger, but may in lowliness of heart obtain thy grace. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, that this oblation may obtain for us the grace of true humility: and likewise take from our hearts the lust of the flesh, and the lust of the eyes, and the pride of life; that living soberly, righteously, and godly, we may attain unto the rewards of everlasting life. Through.
\subsubsec{Postcommunion}
\lett{M}{ay} this sacrament, O Lord, which we have received, cleanse us from every stain of sin: that we, shewing forth humility, may be brought unto thy heavenly kingdom. Through.

\subby{To obtain the Grace of Patience}
\subsubsec{Collect}
\lett{O}{God,} who by the long-suffering of thine only-begotten Son didst tread down the pride of our ancient enemy: grant us, we beseech thee, so rightly to call to mind the sufferings which of his goodness he bare for our sakes; that after his example we may patiently endure all our troubles and adversities. Through the same.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, graciously to receive our gifts and oblations: which we devoutly offer to thy majesty, beseeching thee that thou wouldest vouchsafe to bestow upon us the gift of patience. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that we who have received these sacred mysteries may be restored thereby unto the grace from which we have fallen: that by thy continual protection both here and ever we may be enabled to endure patiently in all adversities. Through.

\subby{To obtain the Grace of Charity}
\subsubsec{Collect}
\lett{O}{God,} who makest all things work together for good to them that love thee: pour into our hearts the steadfast affection of thy charity; that as thou dost put into our minds good desires, so we may not be turned from the same by any temptations. Through.
\subsubsec{Secret}
\lett{O}{God,} who by the sacraments and commandments dost renew us after thy likeness: stablish our goings in thy paths; that like as thou hast vouchsafed unto us the hope of obtaining thy gift of charity, so by these sacrifices which we offer thou wouldest make us effectually to be partakers of the same. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that the grace of the Holy Spirit may enlighten our hearts: that they may be abundantly renewed in the sweetness of perfect charity. Through . . . in the unity of the same Holy Spirit.

\subby{For Devoted Friends}
\subsubsec{Collect}
\lett{O}{God,} who by the grace of thy Holy Spirit hast endued the hearts of thy faithful people with the gifts of charity: grant to thy servants and handmaids, for whom we implore thy mercy, safety of mind and body; that they may love thee with all their strength, and with pure affection perform such things as are acceptable unto thee. Through . . . in the unity of the same.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, to have compassion upon thy servants and handmaids, for whom we offer unto thy majesty this sacrifice of praise: that through these holy things they may obtain the grace of thy heavenly benediction, and attain unto the glory of everlasting felicity. Through.
\subsubsec{Postcommunion}
\lett{O}{Lord,} who dost make us to taste of these divine mysteries, we humbly beseech thee: that like as we have offered these sacraments of salvation for the advancement in charity of thy servants; so they may be profitable unto them for their prosperity and peace. Through.

\subby{For Enemies}
\subsubsec{Collect}
\lett{O}{God,} who art the lover of peace, and preserver of charity: grant unto all our enemies true peace and charity: and vouchsafe unto them remission of all their sins, and by thy mighty power deliver us from their snares. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, graciously to accept the gifts which we offer: that by thy mercy we may be delivered from our enemies, and that they may receive of thee the pardon of their offences. Through.
\subsubsec{Postcommunion}
\lett{M}{ay} this communion, O Lord, deliver us from our offences: and defend us against the snares of our enemies. Through.

\subby{For a Prisoner or Captive}
\subsubsec{Collect}
\lett{O}{God,} who didst deliver blessed Peter the Apostle from his chains, and suffer him to depart unhurt: loose the chains of thy servant who now lieth in captivity; and by the merits of the same thine Apostle suffer him to depart unhurt. Through.
\subsubsec{Secret}
\lett{L}{et} thy plenteous benediction, O Lord, descend upon these sacrifices: that they may loose the chains of this thy servant in captivity, and make us speedily to rejoice in his deliverance. Through.
\subsubsec{Postcommunion}
\lett{O}{Lord,} we beseech thee, graciously hear our prayers: and by these sacraments which we have received deliver thy servant from the chains of his captivity. Through.

\subby{For Those at Sea}
\subsubsec{Collect}
\lett{O}{God,} who didst guide our fathers through the Red Sea, and didst bring them through the deep waters, singing praises to thy name: We humbly beseech thee; that thou wouldest defend thy servants now travelling by sea from all adversity, so that they, voyaging in tranquillity, may be brought by thee unto the haven where they would be. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, to receive the prayers of thy servants, with our sacrifices and oblations: that they for whom we celebrate these mysteries may be defended by thee against all adversities. Through.
\subsubsec{Postcommunion}
\lett{O}{Lord,} who hast sanctified us with this divine mystery, we humbly entreat and beseech thy majesty: that as thou dost suffer us to call to remembrance thy servants by means of these thy heavenly gifts; so through the wood of the holy Cross thou wouldest absolve them from their sins and deliver them from all perils and dangers. Through.

\subby{For the Safety of the Living}
\subsubsec{Collect}
\lett{S}{tretch} forth, O Lord, upon thy faithful people the right hand of thy heavenly succour: that they, seeking thee with their whole heart, may be found worthy to obtain those things which they ask according to thy will. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, to have compassion on our prayers, and graciously to accept these oblations of thy faithful people, which we present before thee for their safety: and that none may ask amiss or fail in his petition; grant that those things which we ask faithfully, we may obtain effectually. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, O Lord, unto thy faithful people steadfastness in faith and singleness of heart: that they, being stablished in divine charity, may not by any temptations be parted from the fulness of the same. Through.

\subby{For the Living \& Departed}
\subsubsec{Collect}
\lett{A}{lmighty} and everlasting God, who hast dominion both of the living and of the dead, and hast mercy upon all whom thou foreknowest will be thine in faith and works: we humbly beseech thee; that all those for whom we are minded to pour forth our prayers, whether in this present world they still be held in the flsh, or being delivered from the body have passed into that which is to come, may at the intercession of all thy Saints obtain of thy bountiful goodness the remission of all their sins. Through.
\subsubsec{Secret}
\lett{O}{God,} to whom alone is known the number of the elect, whom thou hast appointed unto heavenly felicity: grant, we beseech thee; that at the intercession of all thy Saints, the names of all those who have been commended to our prayers, and of all thy faithful people, may ever remain written in the book of those that are predestined to everlasting blessedness. Through.
\subsubsec{Postcommunion}
\lett{A}{lmighty} and merciful God, cleanse us we pray thee, by the sacraments which we have received: and, at the intercession of all thy Saints, grant; that this thy sacrament may not bring upon us guilt to our punishment but pardon to our salvation: that it may cleanse away sins, that it may confirm the feeble, that it may defend us against every danger in this world: that it may obtain for thy
faithful people both living and dead the remission of all their sins. Through.
\end{multicols}

\begin{rubric}
    The following propers are used for the faithful departed.
\end{rubric}

\begin{multicols}{2}
\subsection{For a Chief Bishop Departed}\label{RequiemPrayers}
\subsubsection{Collect}
\lett{O}{God,} who of thy unspeakable providence didst vouchsafe to number thy servant \textit{N.} in the company of thy high Priests: grant, we beseech thee; that like as he did fulfil on earth the office of thine Only-begotten Son, so he may be numbered in the everlasting fellowship of the same thy Saints. Through the same.

\subsubsec{Secret}
\lett{R}{eceive,} O Lord, we beseech thee, these our sacrifices for the soul of thy servant \textit{N.}, thy high Priest: that forasmuch as in this life thou didst give him the dignity of a Bishop, so thou wouldest bid him to be joined to the company of the Saints in thy heavenly kingdom. Through.
\subsubsec{Postcommunion}%Pope changed to Metropolitan
\lett{M}{ay} thy merciful loving-kindness, which we implore, benefit, we beseech thee, O Lord, the soul of thy servant the \textit{Metropolitan} \textit{N.}: that, by thy mercy he may attain the eternal fellowship of him in whom he hoped and believed. Through.

\subby{For a Departed Bishop}
\subsubsec{Collect}
\lett{O}{God,} who didst cause thy servant \textit{N.} to enjoy the dignity of a Bishop in the apostolic Priesthood: grant, we beseech thee; that he may evermore be joined unto the fellowship of the same. Through.
\subsubsec{Secret}
\lett{A}{ccept,} O Lord, we beseech thee, the sacrifice which we offer to thee for the soul of thy servant and Bishop \textit{N.}: that forasmuch as in this life thou didst bestow on him the dignity of a Bishop, so thou wouldest bid him to be joined unto the fellowship of thy Saints in the heavenly kingdom. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that thy merciful loving kindness, which we implore, may be profitable for the soul of thy servant and Bishop \textit{N.}: that by thy mercy he may attain unto the eternal fellowship of him in whom he hath hoped and believed. Through.
\begin{inhead}
or,
\end{inhead}
\vspace{-1\baselineskip}
\subsubsec{Collect}
\lett{G}{rant} unto us, O Lord, that the \textit{soul} of thy \textit{servant} and \textit{Bishop} \textit{N.} (and \textit{N.}), which thou hast taken out of the toil and conflict of this world, may be \textit{partaker} of the fellowship of thy Saints. Through.

%MANUAL ADJUSTMENT:
\vspace{-3ex}

\subsubsec{Secret}
\lett{G}{rant} unto us, we beseech thee, O Lord: that like as by the offering of this oblation thou didst grant unto the whole world remission of sins so the \textit{soul} of thy \textit{servant} and \textit{Bishop} \textit{N.} (and \textit{N.}) may be profited by the same. Through.

%MANUAL ADJUSTMENT:
\vspace{-3ex}

\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, almighty God, that the \textit{soul} of thy \textit{servant} and \textit{Bishop} \textit{N.} (and \textit{N.}), being purified by these sacrifices, may be found worthy to obtain thy pardon and everlasting refreshment. Through.

\subby{For a Departed Priest}
\subsubsec{Collect}
\lett{O}{God,} who didst cause thy \textit{servant} \textit{N.} (and \textit{N.}) to enjoy the dignity of a Priest in the apostolic Priesthood: grant, we beseech thee, that \textit{he} may evermore be joined unto the fellowship of the same. Through.
\subsubsec{Secret}
\lett{A}{ccept,} O Lord, we beseech thee, the sacrifice which we offer to thee for the \textit{soul} of thy \textit{servant} and \textit{Priest} \textit{N.} (and \textit{N.}): that forasmuch as in this life thou didst bestow on him the dignity of a Priest, so thou wouldest bid him to be joined unto the fellowship of thy Saints in the heavenly kingdom.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that thy merciful loving kindness, which we implore, may be profitable for the \textit{soul} of thy \textit{servant} and \textit{Priest} \textit{N.} (and \textit{N.}): that by thy mercy \textit{he} may attain unto the eternal fellowship of him in whom in whom \textit{he} hath hoped and believed. Through.
\begin{inhead}
    or,
\end{inhead}
\vspace{-1\baselineskip}
\subsubsec{Collect}
\lett{G}{rant,} we beseech thee, O Lord, that the \textit{soul} of thy \textit{servant} \textit{N.} (and \textit{N.}): whom while \textit{he} lived in this world thou didst adorn with the holy gifts of Priesthood; may ever rejoice in the glory of thy heavenly abode. Through.
\subsubsec{Secret}
\lett{R}{eceive,} O Lord, we beseech thee, these sacrifices, which we offer for the \textit{soul} of \textit{N.} (and \textit{N.}) thy \textit{servant} and \textit{Priest}: that as thou hast given \textit{him} the dignity of the Priesthood, thou wouldest also give him the reward of the same. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, almighty God: to the \textit{soul} of \textit{N.} (and \textit{N.}) thy \textit{servant} and \textit{Priest}, the fellowship of eternal blessedness in the congregation of the just. Through.

\subby{For a Man Departed}
\subsubsec{Collect}
\lett{I}{ncline} thine ear, O Lord, unto the prayers wherewith we humbly entreat thy mercy: that thou wouldest set the \textit{soul} of thy \textit{servant} \textit{N.} (and \textit{N.}), which thou hast bidden depart this life, in a place of peace and light; and make \textit{him} to be joined unto the fellowship of thy Saints. Through.
\subsubsec{Secret}
\lett{G}{rant} to us, we beseech thee, O Lord: that like as by the offering of this oblation thou didst bestow upon the whole world remission of sins: so the \textit{soul} of thy \textit{servant} \textit{N.} (and \textit{N.}) may be profited by the same. Through.
\subsubsec{Postcommunion}
\lett{A}{bsolve,} O Lord, we beseech thee, the \textit{soul} of thy \textit{servant} \textit{N.} (and \textit{N.}), from every bond of sin: that in the glory of the resurrection \textit{he} may be raised up amid thy Saints and elect unto newness of life. Through.


\subby{For a Woman Departed}
\subsubsec{Collect}
\lett{W}{e} beseech thee, O Lord, of thy loving-kindness to have mercy on the \textit{soul} of thy \textit{handmaid} \textit{N.} (and \textit{N.}): that being purged from the defilements of our mortal nature, \textit{she} may be restored to the portion of everlasting salvation. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, that by these sacrifices, whereby alone we have deliverance from our sins, the \textit{soul} of thy \textit{handmaid} \textit{N.} (and \textit{N.}) may be absolved from all \textit{her} offences: that through this our bounden duty of propitiation she may obtain thine everlasting mercy. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that forasmuch as the \textit{soul} of thine \textit{handmaid} \textit{N.} (and \textit{N.}) hath received this pledge of thy never-failing mercy: \textit{she} may be made partaker of thine everlasting brightness. Through.

\subby{For a Father and Mother}
\subsubsec{Collect}
\lett{O}{God,} who hast bidden us to honour our father and our mother: of thy loving-kindness have mercy on the souls of \textit{my father and mother} and forgive them their sins; and grant that \textit{I} may behold them in the joy of the eternal brightness. Through.
\subsubsec{Secret}
\lett{R}{eceive,} O Lord, the sacrifice which I offer unto thee for the souls of \textit{my father and mother}: and grant them in the land of the living everlasting blessedness; and join \textit{me} with them in the felicity of thy Saints. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that this partaking of thy heavenly sacrament may obtain for the souls of \textit{my father and mother} rest and light perpetual: and that with them \textit{I} may receive the crown of thy everlasting grace. Through.

\subby{For a Father Only}
\subsubsec{Collect}
\lett{O}{God,} who hast bidden us to honour our father and our mother: of thy loving-kindness have mercy on the soul of \textit{my} father and forgive him his sins; and grant that \textit{I} may behold him in the joy of the eternal brightness. Through.
\subsubsec{Secret}
\lett{R}{eceive,} O Lord, the sacrifice which I offer unto thee for the soul of \textit{my} father: and grant him in the land of the living everlasting blessedness; and join \textit{me} with him in the felicity of thy Saints. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that this partaking of thy heavenly sacrament may obtain for the soul of \textit{my} father rest and light perpetual; and that with him \textit{I} may receive the crown of thy everlasting grace. Through.

\subby{For a Mother Only}
\subsubsec{Collect}
\lett{O}{God,} who hast bidden us to honour our father and our mother: of thy loving-kindness have mercy on the soul of \textit{my} mother and forgive her her sins; and grant that \textit{I} may behold her in the joy of the eternal brightness. Through.
\subsubsec{Secret}
\lett{R}{eceive,} O Lord, the sacrifice which \textit{I} offer unto thee for the soul of \textit{my} mother: and grant her in the land of the living everlasting blessedness; and join \textit{me} with her in the felicity of thy Saints. Through.
\subsubsec{Postcommunion}
\lett{W}{e} beseech thee, O Lord, that this partaking of thy heavenly sacrament may obtain for the soul of \textit{my} mother rest and light perpetual: and that with her \textit{I} may receive the crown of thy everlasting grace. Through.

\subby{For Departed Brethren, Kinsfolk, \& Benefactors}
\subsubsec{Collect}
\lett{O}{God,} the giver of pardon and lover man's salvation: we beseech thee of thy mercy to grant ;
that the brethren, kinsfolk, and benefactors of our congregation who have passed out of this world, may at the intercession of blessed Mary, ever Virgin, and of all thy Saints attain unto the fellowship of eternal blessedness. Through.
\subsubsec{Secret}
\lett{O}{God,} whose mercies are without number, graciously receive our humble prayers: and through these sacraments of our salvation grant unto the souls of our brethren, kinsfolk, and benefactors, to whom thou hast given the confession of thy name, the remission of all their sins. Through.
\subsubsec{Postcommunion}
\lett{G}{rant,} we beseech thee, almighty and merciful God: that the souls of our brethren, kinsfolk, and benefactors, for whom we have offered unto thy majesty this sacrifice of praise; may by the power of this sacrament be cleansed from all their sins, and by thy mercy obtain the blessedness of everlasting light.
Through.

\subby{For Those who Rest in a Cemetery}
\subsubsec{Collect}
\lett{O}{God,} through whose mercy the souls of the faithful are at rest: mercifully grant to thy servants and handmaids, and to all that here and in all places do rest in Christ, the remission of their sins; that being delivered from all their iniquities, they may rejoice with thee for evermore. Through the same.
\subsubsec{Secret}
\lett{G}{raciously} receive, O Lord, this sacrifice which we offer for the souls of thy servants and handmaids, and of all Christian people who here and in all places are asleep in Christ: that by this wondrous sacrifice they may be delivered from the dreadful bonds of death, and be found worthy to enter into everlasting life. Through the same.
\subsubsec{Postcommunion}
\lett{O}{God,} the light of the souls of the faithful, assist our supplications: and grant unto thy servants and handmaids, whose bodies both here and in all places do rest in Christ, a place of refreshing, the blessedness of thy rest, and the glory of everlasting light. Through the same.

\subby{For Many Persons Departed}
\subsubsec{Collect}
\lett{O}{God,} whose property is ever to have mercy and to forgive: have compassion upon the souls of thy servants and handmaids, and forgive them all their sins; that being delivered from the bonds of this our mortal nature, they may be found worthy to enter into life. Through.
\subsubsec{Secret}
\lett{G}{rant} to us, we beseech thee, O Lord, that like as by the offering of this oblation thou didst bestow upon the whole world remission of sins: so the souls of thy servants and handmaids may be profited by the same. Through.
\subsubsec{Postcommunion}
\lett{O}{God,} to whom alone it belongeth to bestow healing after death: grant, we beseech thee; that the souls of thy servants and handmaids, being cleansed from the defilements of this earthly life, may be numbered in the portion of thy redeemed. Who livest and reignest.

\subby{Other Prayers for Many Persons Departed}
\subsubsec{Collect}
\lett{G}{rant,} O Lord, we pray thee, to the souls of thy servants and handmaids thy perpetual mercy: that as they have hoped and believed in thee, so it may be profitable unto them for evermore. Through.
\subsubsec{Secret}
\lett{W}{e} beseech thee, O Lord, mercifully to look upon these our gifts: that these things which we humbly offer to the glory of thy name, may be profitable unto them that are departed for the forgiveness of their sins. Through.
\subsubsec{Postcommunion}
\lett{W}{e} humbly offer unto thee, O Lord, our prayers for the souls of thy servants and handmaids: beseeching thee; that whatsoever they may have contracted in their earthly conversation being pardoned by thy goodness, they may be made partakers of the gladness of thy redeemed. Through.
\end{multicols}